### 2016-10-21 README.tex
Initial commit for enforcing invariant confluence.

\documentclass{tex/mjw}
\title{Enforced Invariant Confluence}
\author{Michael Whittaker}

\usepackage{paralist}
\usepackage{tex/python}

\newcommand{\iinvariant}{\mathcal{I}}
\newcommand{\iconfluent}{$\iinvariant$-confluent}
\newcommand{\iconfluence}{$\iinvariant$-confluence}

\begin{document}
\maketitle

\section{Abstract}
\emph{Strong consistency} allows programmers to ignore many of the complexities
of a distributed system, treating it as if it were running on a single machine.
However, enforcing strong consistency requires \emph{coordination}, and
coordination leads to unavailability \cite{gilbert2002brewer} (at worst) or
increased latency (at best). However, coordination cannot always be avoided.
Certain application invariants require coordination in order to be globally
enforced. Bailis et al. developed the notion of invariant confluence
(\iconfluence{}) as a necessary and sufficient condition for coordination
freedom \cite{bailis2014coordination}. Intuitively, a set of transactions is
\iconfluent{} with respect to some invariant $I$ if merging database states
that satisfy $I$ always results in a database state that also satisfies $I$.
Bailis characterized many common invariants (e.g. uniqueness constraints,
foreign key constraints), showing which could be maintained without
coordination. However, this characterization required hand-written proofs. This
research project aims to expand on Bailis' research by providing a set of
CRDTs \cite{shapiro2011comprehensive, shapiro2011conflict}, a language of
invariants, and a language of transactions that facilitates the ability for
\iconfluence{} to be determined algorithmically. An \iconfluence{} decision
procedure would allow programmers to guarantee that their programs can execute
without coordination.

\section{Research Vision}
At a very high level, this research tackles the same problems tackled by many
existing research projects like BOOM Analytics \cite{alvaro2010boom}, Bloom
\cite{alvaro2011consistency}, Bloom$^L$ \cite{conway2012logic}, TARDiS
\cite{crookstardis}, and IPA \cite{holt2016disciplined}:

\begin{quotation}
  \emph{How can we make writing weakly consistent programs more principled?}
\end{quotation}

BOOM Analytics, Bloom, and Bloom$^L$ use data-centric declarative programming
and program analysis based on the CALM theorem. TARDiS exposes the entire
history of divergent computations to users in a git-like system to empower them
to programmatically resolve conflicts. IPA uses a type system to ensure weakly
consistent data doesn't leak into strongly consistent data.

This system builds off existing research on CRDTs and \iconfluence{} to provide
a library of CRDTs, a language to express invariants over those CRDTs, the
ability to transactionally update the replicated CRDTs, and a decision
procedure to algorithmically determine whether the transactions are
\iconfluent{} with respect to the invariants, ensuring coordination free
execution if they are and potentially giving illustrative counterexamples when
they are not. A simple example of psuedocode written with this system may look
something like this:

\begin{Python}
# Create an increment/decrement counter CRDT.
x = PNCounter(0)

# Set the invariant that the counter must always be non-negative.
set_invariant(x >= 0)

# This transaction is invariant confluent and would be accepted by the system.
@transaction
def foo():
  x.increment(42)

# This transaction is not invariant confluent and would be rejected by the
# system.
@transaction
def bar():
  x.decrement(1)
\end{Python}

A more complex example of psuedocode may look something like this:

\begin{Python}
# A map from employee id to employee name.
employees = Map[Int, String]()
# The number of employees.
num_employees = PN-Counter(0)
# Employee ids partitioned into teams.
teams = Set[Set[Int]]

# The number of employees has to equal the size of employees.
set_invariant(num_employees = employees.size())
# All ids in teams must appear in employees.
set_invariant(forall team in teams. team subset employees.keys())
# All teams must be disjoint.
set_invariant(forall a, b in teams. a != b => a intersect b = {})
# Every employee must be on a team
set_invariant(union(teams) = employees.keys())

# This transaction which adds a new employee would be checked for
# invariant-confluence by the system.
@transaction
def add_employee(name):
  id = employees.unique_id()
  employees.put(name, id)
  num_employees.increment(1)
  teams[0].add(id)
\end{Python}

This pseudocode brings up a lot of \emph{theoretical questions}...
\begin{inparaitem}
  \item What CRDTs and what operations can we support?
  \item What invariant language can we support?
  \item Given the set of CRDTs and invariants, is determining invariant
    confluence even decidable?
  \item If so, can it be determined efficiently?
  \item If not, how can we limit the CRDTs, operations, or invariants to make
    it decidable?
\end{inparaitem}
...and a lot of \emph{systems questions}:
\begin{inparaitem}
  \item Is the system expressive enough to write real-world programs?
  \item Can the system be implemented as a library in an existing programming
    language?
  \item How do we architect and build an actual system out of this?
\end{inparaitem}

This research aims to answer these questions and develop a first-of-its-kind
system that allows programmers to write coordination free distributed systems
with more discipline than ever before.

\bibliography{tex/bib}
\bibliographystyle{plain}
\end{document}

### 2016-10-23 README.tex
Refactored and added part of counter section.

@@
 
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{mathtools}
 \usepackage{paralist}
 \usepackage{tex/python}
+\usepackage{tikz}
+
+\newcommand{\figref}[1]{Figure \ref{fig:#1}}
+\newcommand{\defeq}{\stackrel{\mathclap{\normalfont\mbox{\scriptsize def}}}{=}}
+\newcommand{\dom}[1]{\text{dom}}
 
@@
 
-\section{Abstract}
-\emph{Strong consistency} allows programmers to ignore many of the complexities
-of a distributed system, treating it as if it were running on a single machine.
-However, enforcing strong consistency requires \emph{coordination}, and
-coordination leads to unavailability \cite{gilbert2002brewer} (at worst) or
-increased latency (at best). However, coordination cannot always be avoided.
-Certain application invariants require coordination in order to be globally
-enforced. Bailis et al. developed the notion of invariant confluence
-(\iconfluence{}) as a necessary and sufficient condition for coordination
-freedom \cite{bailis2014coordination}. Intuitively, a set of transactions is
-\iconfluent{} with respect to some invariant $I$ if merging database states
-that satisfy $I$ always results in a database state that also satisfies $I$.
-Bailis characterized many common invariants (e.g. uniqueness constraints,
-foreign key constraints), showing which could be maintained without
-coordination. However, this characterization required hand-written proofs. This
-research project aims to expand on Bailis' research by providing a set of
-CRDTs \cite{shapiro2011comprehensive, shapiro2011conflict}, a language of
-invariants, and a language of transactions that facilitates the ability for
-\iconfluence{} to be determined algorithmically. An \iconfluence{} decision
-procedure would allow programmers to guarantee that their programs can execute
-without coordination.
-
-\section{Research Vision}
-At a very high level, this research tackles the same problems tackled by many
-existing research projects like BOOM Analytics \cite{alvaro2010boom}, Bloom
-\cite{alvaro2011consistency}, Bloom$^L$ \cite{conway2012logic}, TARDiS
-\cite{crookstardis}, and IPA \cite{holt2016disciplined}:
-
-\begin{quotation}
-  \emph{How can we make writing weakly consistent programs more principled?}
-\end{quotation}
-
-BOOM Analytics, Bloom, and Bloom$^L$ use data-centric declarative programming
-and program analysis based on the CALM theorem. TARDiS exposes the entire
-history of divergent computations to users in a git-like system to empower them
-to programmatically resolve conflicts. IPA uses a type system to ensure weakly
-consistent data doesn't leak into strongly consistent data.
-
-This system builds off existing research on CRDTs and \iconfluence{} to provide
-a library of CRDTs, a language to express invariants over those CRDTs, the
-ability to transactionally update the replicated CRDTs, and a decision
-procedure to algorithmically determine whether the transactions are
-\iconfluent{} with respect to the invariants, ensuring coordination free
-execution if they are and potentially giving illustrative counterexamples when
-they are not. A simple example of psuedocode written with this system may look
-something like this:
-
-\begin{Python}
-# Create an increment/decrement counter CRDT.
-x = PNCounter(0)
-
-# Set the invariant that the counter must always be non-negative.
-set_invariant(x >= 0)
-
-# This transaction is invariant confluent and would be accepted by the system.
-@transaction
-def foo():
-  x.increment(42)
-
-# This transaction is not invariant confluent and would be rejected by the
-# system.
-@transaction
-def bar():
-  x.decrement(1)
-\end{Python}
-
-A more complex example of psuedocode may look something like this:
-
-\begin{Python}
-# A map from employee id to employee name.
-employees = Map[Int, String]()
-# The number of employees.
-num_employees = PN-Counter(0)
-# Employee ids partitioned into teams.
-teams = Set[Set[Int]]
-
-# The number of employees has to equal the size of employees.
-set_invariant(num_employees = employees.size())
-# All ids in teams must appear in employees.
-set_invariant(forall team in teams. team subset employees.keys())
-# All teams must be disjoint.
-set_invariant(forall a, b in teams. a != b => a intersect b = {})
-# Every employee must be on a team
-set_invariant(union(teams) = employees.keys())
-
-# This transaction which adds a new employee would be checked for
-# invariant-confluence by the system.
-@transaction
-def add_employee(name):
-  id = employees.unique_id()
-  employees.put(name, id)
-  num_employees.increment(1)
-  teams[0].add(id)
-\end{Python}
-
-This pseudocode brings up a lot of \emph{theoretical questions}...
-\begin{inparaitem}
-  \item What CRDTs and what operations can we support?
-  \item What invariant language can we support?
-  \item Given the set of CRDTs and invariants, is determining invariant
-    confluence even decidable?
-  \item If so, can it be determined efficiently?
-  \item If not, how can we limit the CRDTs, operations, or invariants to make
-    it decidable?
-\end{inparaitem}
-...and a lot of \emph{systems questions}:
-\begin{inparaitem}
-  \item Is the system expressive enough to write real-world programs?
-  \item Can the system be implemented as a library in an existing programming
-    language?
-  \item How do we architect and build an actual system out of this?
-\end{inparaitem}
-
-This research aims to answer these questions and develop a first-of-its-kind
-system that allows programmers to write coordination free distributed systems
-with more discipline than ever before.
+\input{abstract.tex}
+\input{vision.tex}
+\input{counters.tex}
 

### 2016-10-23 counters.tex
Refactored and added part of counter section.

\section{Counters}
As a first step towards accomplishing our research vision, let's begin with a
\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
language: PN-counters and linear equations and inequalities.

\subsection{Overview}
Recall that a \emph{G-Counter} CRDT is an increment-only counter
\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
of two $n$-tuples.

A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
and can be implemented as a pair $(p, n)$ of two G-Counters
\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
a pairwise merge of $p$ and $n$.

Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).

A \emph{transaction} chooses a subset of variables and decrements or increments
each by a constant amount. For example, a transaction may increment $x$ by 1,
decrement $y$ by 42, and increment $z$ by 100.

Given a set of transactions and an invariant, we want to determine whether the
transactions are \iconfluent{} with respect to the invariant. For example, the
transaction that \emph{increments} $x$ by 1 is invariant confluent with respect
to the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1
is not, as shown in \figref{decrement}.

\begin{figure}[h]
  \centering
  \begin{tikzpicture}
    % \draw (0, 0) grid (4, 4);
    \node[draw] (start)  at (2, 4) {$x=2$};
    \node[draw] (left)   at (0, 2) {$x=1$};
    \node[draw] (right)  at (4, 2) {$x=1$};
    \node[draw] (merged) at (2, 0) {$x=0$};

    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
    \draw[thick, ->] (left)  to (merged);
    \draw[thick, ->] (right) to (merged);
  \end{tikzpicture}
  \caption{
    A counterexample showing transaction $x := x - 1$ is not invariant
    confluent with respect to the invariant $x > 0$. The top, left, and right
    states satisfy the invariant, but the bottom state does not.
  }
  \label{fig:decrement}
\end{figure}

\subsection{Formalizing the Question}
\newcommand{\var}{\textsf{Var}}
\newcommand{\ints}{\mathbb{Z}}

Our invariant language is defined by the grammar in \figref{invariant-grammar}
which includes linear equations and inequalities over variables in a finite set
$\var$ and integer constants in $\ints$.

\begin{figure}[h]
  \centering

  \newcommand{\atom}{\textsf{atom}}
  \newcommand{\aexp}{\textsf{aexp}}
  \newcommand{\bexp}{\textsf{bexpr}}
  \begin{gather*}
    \begin{array}{rrll}
      \atom & ::= & k & \text{\emph{constants}} \\
            & |   & x & \text{\emph{variables}} \\
      &&& \\
      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
             & |   & -\atom        & \text{\emph{negation}} \\
             & |   & \aexp + \aexp & \text{\emph{addition}} \\
             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
      &&& \\
      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
             & |   & \lnot \bexp       & \text{\emph{negation}} \\
             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
    \end{array}
  \end{gather*}

  \caption{
    A grammar for our linear equation and linear inequality invariant language.
    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
    not included because they can be defined in terms of the existing
    operators.
  }
  \label{fig:invariant-grammar}
\end{figure}

A \emph{database state} is a total function $D: \var \to \ints$.  We say a
database state $D$ \emph{satisfies invariant} $I$ if $I$ evaluates to true
after all variables in $I$ have been replaced by their mapping in $D$.

A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
t$, produces a new database state defined as follows:
\[
  (D \circ t)(x) \defeq \begin{cases*}
    D(x) + t(x) & $x \in \dom{}(t)$ \\
    D(x)        & otherwise
  \end{cases*}
\]

A transaction chain $C$ created from a set of transactions $T$ is a sequence of
transactions $t_1, \ldots, t_n$ where $t_j \in T$ for $1 \leq j \leq n$. Let $D
\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
$C$ \emph{satisfies invariant} $I$ if for every database state $D$ and prefix
$C'$ of $C$, $D \circ C'$ satisfies invariant $I$.

A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$,
$C_1$, and $C_2$ satisfy $I$, then so does $D \circ C_1 \circ C_2$.

### 2016-10-24 counters.tex
Geometric interpretation of counter Iconfluence

@@
 $C_1$, and $C_2$ satisfy $I$, then so does $D \circ C_1 \circ C_2$.
+
+\subsection{A Geometric Interpretation}
+\newcommand{\inva}{x + y \geq 0}
+\newcommand{\invb}{x - y \geq 0}
+\newcommand{\invc}{x \geq 1}
+\newcommand{\invd}{x \geq 3}
+\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
+
+Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
+to an invariant. Other times, not so much. For example, which transactions are
+invariant confluent with respect to the invariant $\inv$? It's not immediately
+clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
+respect to $I$?'' geometrically, and doing so will make answering the question
+much easier!
+
+Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
+the values of the $n$ variables as an $n$-dimensional point, then we can draw
+the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
+$I$ by eliminating negations and converting to conjunctive normal form, then
+the set of points satisfying $I$ is the intersection (for $\land)$ and union
+(for $\lor$) of the solutions to a set of linear equations and inequalities,
+which is always straightforward to draw.  For example, the set of points
+satisfying $\inv$ is shown in \figref{geometric-example}.
+
+\begin{figure}
+  \newenvironment{griddedpic}{
+    \begin{tikzpicture}
+      \clip (-4, -1) rectangle (4, 4);
+  }{
+      \draw (-4, -1) grid (4, 4);
+      \draw[ultra thick] (-4, 0) -- (4, 0);
+      \draw[ultra thick] (0, -1) -- (0, 4);
+    \end{tikzpicture}
+  }
+
+  \begin{subfigure}[c]{0.5\textwidth}
+    \begin{griddedpic}
+      \fill[red, opacity=0.4]
+        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
+    \end{griddedpic}
+    \caption{$\inva$}
+  \end{subfigure}
+  \begin{subfigure}[c]{0.5\textwidth}
+    \begin{griddedpic}
+      \fill[blue, opacity=0.4]
+        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
+    \end{griddedpic}
+    \caption{$\invb$}
+  \end{subfigure}
+
+  \vspace{0.5cm}
+
+  \begin{subfigure}[c]{0.5\textwidth}
+    \begin{griddedpic}
+      \fill[green, opacity=0.4]
+        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
+    \end{griddedpic}
+    \caption{$\invc$}
+  \end{subfigure}
+  \begin{subfigure}[c]{0.5\textwidth}
+    \begin{griddedpic}
+      \fill[orange, opacity=0.4]
+        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
+    \end{griddedpic}
+  \caption{$\invd$}
+  \end{subfigure}
+
+  \vspace{0.5cm}
+
+  \begin{subfigure}[c]{\textwidth}
+    \centering
+    \begin{griddedpic}
+      \fill[purple, opacity=0.4]
+        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
+        (3, 3) -- (1, 1) -- cycle;
+    \end{griddedpic}
+    \caption{$\inv$}
+  \end{subfigure}
+
+  \caption{An illustration of the solution to $\inv$.}
+  \label{fig:geometric-example}
+\end{figure}
+
+Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
+\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
+question of whether a set of transactions $T$ is \iconfluent{} with respect to
+invariant $I$ has been reduced to the following question. Starting at an
+arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
+t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
+all while staying in the solution space of $I$, then are we guaranteed that $D
++ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
+solution space of $I$?
+
+Turning again to our example in \figref{geometric-example}, it is now clear
+that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
+\iconfluent{} with respect to $\inv$. These are the vectors that point up and
+not too far left or right.

### 2016-10-24 counters.tex
Proved one is enough theorem.

@@
 \]
+Note that transaction application is commutative. That is for all database
+states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
+t_2 \circ t_1$.
 
@@
 \circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ if for every database state $D$ and prefix
-$C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
+$C$ \emph{satisfies invariant} $I$ starting at $D$ if for every prefix $C'$ of
+$C$, $D \circ C'$ satisfies invariant $I$.
 
 A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$,
-$C_1$, and $C_2$ satisfy $I$, then so does $D \circ C_1 \circ C_2$.
+all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
+satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
+starting at $D$, then so does $D \circ C_1 \circ C_2$.
 
@@
 not too far left or right.
+
+\subsection{One is Enough}
+Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
+invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
+from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
+starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then so does $D \circ
+C_1 \circ C_2$.
+%
+A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
+it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
+transactions, we have to consider potentially large transaction chains which
+can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
+because we only have to consider a pair of transactions at a time. This
+subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
+are equivalent.
+
+\begin{theorem}
+  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
+  and only if it is \iconfluent{}.
+\end{theorem}
+\begin{proof}
+  The reverse direction is immediate from the definition of \iconfluence{}. For
+  the forward direction, consider a set of transactions $T$ that is
+  1-\iconfluent{} with respect to an invariant $I$, but assume for
+  contradiction that $T$ is not \iconfluent{} with respect to $I$. That is,
+  there exists a database state $D$ and two chains $C^1 = t^1_1, t^1_2, \ldots,
+  t^1_m$ and $C^2 = t^2_1, t^2_2, \ldots, t^2_n$ such that $D$ satisfies $I$,
+  $C^1$ satisfies $I$ starting at $D$, and $C^2$ satisfies $I$ starting at $D$
+  but $D \circ C^1 \circ C^2$ does not.
+
+  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
+  mathematical induction on $i$, that $D \circ C^1_j \circ C^2_k$ satisfies $I$
+  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
+  m$.
+  %
+  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^1_0
+  \circ C^2_0 = D$ which satisfies $I$ by assumption.
+  %
+  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
+  \leq k \leq n$, and $0 < i = j + k$ that $D \circ C^1_j \circ C^2_k$
+  satisfies $I$. If $j = 0$, then $D \circ C^1_0 \circ C^2_k = D \circ C^2_k$
+  satisfies $I$ by the assumption that $C^2$ satisfies $I$ starting at $D$, and
+  likewise for when $k = 0$. Now consider $j, k > 0$. By the inductive
+  hypothesis, we know
+    (1) $D \circ C^1_{j-1} \circ C^2_{k-1}$,
+    (2) $D \circ C^1_{j-1} \circ C^2_{k}$, and
+    (3) $D \circ C^1_{j}   \circ C^2_{k-1}$
+  all satisfy $I$. If we let (1) be our initial database state, $t^1_j$ be one
+  chain, and $t^2_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
+  us that $(1) \circ t^1_j \circ t^2_k = D \circ C^1_j \circ C^2_k$ satisfies
+  $I$.
+\end{proof}

### 2016-10-24 doc/abstract.tex
Moved tex into doc directory.

\section{Abstract}
\emph{Strong consistency} allows programmers to ignore many of the complexities
of a distributed system, treating it as if it were running on a single machine.
However, enforcing strong consistency requires \emph{coordination}, and
coordination leads to unavailability \cite{gilbert2002brewer} (at worst) or
increased latency (at best). However, coordination cannot always be avoided.
Certain application invariants require coordination in order to be globally
enforced. Bailis et al. developed the notion of invariant confluence
(\iconfluence{}) as a necessary and sufficient condition for coordination
freedom \cite{bailis2014coordination}. Intuitively, a set of transactions is
\iconfluent{} with respect to some invariant $I$ if merging database states
that satisfy $I$ always results in a database state that also satisfies $I$.
Bailis characterized many common invariants (e.g. uniqueness constraints,
foreign key constraints), showing which could be maintained without
coordination. However, this characterization required hand-written proofs. This
research project aims to expand on Bailis' research by providing a set of
CRDTs \cite{shapiro2011comprehensive, shapiro2011conflict}, a language of
invariants, and a language of transactions that facilitates the ability for
\iconfluence{} to be determined algorithmically. An \iconfluence{} decision
procedure would allow programmers to guarantee that their programs can execute
without coordination.

### 2016-10-24 doc/vision.tex
Moved tex into doc directory.

\section{Research Vision}
At a very high level, this research tackles the same problems tackled by many
existing research projects like BOOM Analytics \cite{alvaro2010boom}, Bloom
\cite{alvaro2011consistency}, Bloom$^L$ \cite{conway2012logic}, TARDiS
\cite{crookstardis}, and IPA \cite{holt2016disciplined}:

\begin{quotation}
  \emph{How can we make writing weakly consistent programs more principled?}
\end{quotation}

BOOM Analytics, Bloom, and Bloom$^L$ use data-centric declarative programming
and program analysis based on the CALM theorem. TARDiS exposes the entire
history of divergent computations to users in a git-like system to empower them
to programmatically resolve conflicts. IPA uses a type system to ensure weakly
consistent data doesn't leak into strongly consistent data.

This system builds off existing research on CRDTs and \iconfluence{} to provide
a library of CRDTs, a language to express invariants over those CRDTs, the
ability to transactionally update the replicated CRDTs, and a decision
procedure to algorithmically determine whether the transactions are
\iconfluent{} with respect to the invariants, ensuring coordination free
execution if they are and potentially giving illustrative counterexamples when
they are not. A simple example of psuedocode written with this system may look
something like this:

\begin{Python}
# Create an increment/decrement counter CRDT.
x = PNCounter(0)

# Set the invariant that the counter must always be non-negative.
set_invariant(x >= 0)

# This transaction is invariant confluent and would be accepted by the system.
@transaction
def foo():
  x.increment(42)

# This transaction is not invariant confluent and would be rejected by the
# system.
@transaction
def bar():
  x.decrement(1)
\end{Python}

A more complex example of psuedocode may look something like this:

\begin{Python}
# A map from employee id to employee name.
employees = Map[Int, String]()
# The number of employees.
num_employees = PN-Counter(0)
# Employee ids partitioned into teams.
teams = Set[Set[Int]]

# The number of employees has to equal the size of employees.
set_invariant(num_employees = employees.size())
# All ids in teams must appear in employees.
set_invariant(forall team in teams. team subset employees.keys())
# All teams must be disjoint.
set_invariant(forall a, b in teams. a != b => a intersect b = {})
# Every employee must be on a team
set_invariant(union(teams) = employees.keys())

# This transaction which adds a new employee would be checked for
# invariant-confluence by the system.
@transaction
def add_employee(name):
  id = employees.unique_id()
  employees.put(name, id)
  num_employees.increment(1)
  teams[0].add(id)
\end{Python}

This pseudocode brings up a lot of \emph{theoretical questions}...
\begin{inparaitem}
  \item What CRDTs and what operations can we support?
  \item What invariant languages can we support?
  \item Given the set of CRDTs and invariants, is determining invariant
    confluence even decidable?
  \item If so, can it be determined efficiently?
  \item If not, how can we limit the CRDTs, operations, or invariants to make
    it decidable?
\end{inparaitem}
...and a lot of \emph{systems questions}:
\begin{inparaitem}
  \item Is the system expressive enough to write real-world programs?
  \item Can the system be implemented as a library in an existing programming
    language?
  \item How do we architect and build an actual system out of this?
\end{inparaitem}

This research aims to answer these questions and develop a first-of-its-kind
system that allows programmers to write coordination free distributed systems
with more discipline than ever before.

### 2016-10-25 doc/counters.tex
Fixed invariant bug.
Fixed even more invariant typos.
Finished counters section!

--- 2016-10-24 counters.tex
@@
 
-\subsection{Overview}
+\subsection{Overview}\label{sec:counter-overview}
 Recall that a \emph{G-Counter} CRDT is an increment-only counter
@@
 
-\subsection{Formalizing the Question}
+\subsection{Formalizing the Question}\label{sec:counter-question}
 \newcommand{\var}{\textsf{Var}}
@@
 
-\subsection{A Geometric Interpretation}
+\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
 \newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \geq 0}
-\newcommand{\invc}{x \geq 1}
-\newcommand{\invd}{x \geq 3}
+\newcommand{\invb}{x - y \leq 0}
+\newcommand{\invc}{y \geq 1}
+\newcommand{\invd}{y \geq 3}
 \newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
@@
 
-\subsection{One is Enough}
+\subsection{One is Enough}\label{sec:counter-one}
 Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
@@
 
-\begin{theorem}
+\begin{theorem}\label{thm:one-is-enough}
   For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
@@
 \end{proof}
+
+\subsection{A Solution}
+This section posed the question: how do we write an algorithm to decide whether
+a set of PN-counter increment/decrement transactions are \iconfluent{} with
+respect to an invariant expressed as a formula of linear equations and
+inequalities?  Now that we're armed with an intuitive
+(\secref{counter-overview}) and formal (\secref{counter-question})
+understanding of the problem, a geometric interpretation
+(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
+in a perfect position to present a solution!
+
+In a nutshell, our algorithm translates a set of transactions $T$ and invariant
+$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
+is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
+by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
+\iconfluent{} with respect to $I$. It then uses Z3 \cite{de2008z3} to check if
+$F$ is a tautology.
+
+The algorithm for constructing $F$ is best explained by way of example.
+Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
+transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
+constructs the following formula
+\begin{align}
+  & \forall x.\> \forall y.\>                \label{eq:vars} \\
+  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
+  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
+  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
+  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
+  & x + y \geq 0 \implies {}                 \label{eq:i} \\
+  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
+  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
+  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
+\end{align}
+where
+\begin{itemize}
+  \item
+    \eqref{eq:vars} universally quantifies the free variables in $I$.
+
+  \item
+    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
+    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
+    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
+    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
+    one transaction and $da_2, \ldots, dz_2$ represent the increments and
+    decrements of another transaction.
+
+  \item
+    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
+    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
+    incremented by 1, and $y$ must be incremented by $2$.
+
+  \item
+    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
+    assumption that the current database state satisfies the invariant.
+
+  \item
+    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
+    Intuitively, it predicates our implication on the assumption that if we
+    apply the first transaction, our database still satisfies the invariant.
+    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
+    instead of the first.
+
+  \item
+    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
+    dx_2$. Intuitively, it says that if we apply both the transactions, we
+    continue to satisfy our invariant.
+\end{itemize}
+
+To summarize, the formula encodes the statement that for all database state
+$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
+transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
+and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
+dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
+restatement of the definition of 1-\iconfluence!

### 2016-10-25 doc/counters.tex
Minor typos and fixes.

@@
 \emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-counters and linear equations and inequalities.
+language: PN-Counters and linear equations and inequalities.
 
@@
 transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is invariant confluent with respect
-to the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1
-is not, as shown in \figref{decrement}.
+transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
+the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
+not, as shown in \figref{decrement}.
 
@@
 A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_j \in T$ for $1 \leq j \leq n$. Let $D
+transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
 \circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
@@
 satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then so does $D \circ C_1 \circ C_2$.
+starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
 
@@
 to an invariant. Other times, not so much. For example, which transactions are
-invariant confluent with respect to the invariant $\inv$? It's not immediately
+\iconfluent{} with respect to the invariant $\inv$? It's not immediately
 clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
@@
 from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then so does $D \circ
-C_1 \circ C_2$.
+starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
+C_1 \circ C_2$ satisfies $I$.
 %
@@
   t^1_m$ and $C^2 = t^2_1, t^2_2, \ldots, t^2_n$ such that $D$ satisfies $I$,
-  $C^1$ satisfies $I$ starting at $D$, and $C^2$ satisfies $I$ starting at $D$
-  but $D \circ C^1 \circ C^2$ does not.
+  $C^1$ satisfies $I$ starting at $D$, and $C^2$ satisfies $I$ starting at $D$,
+  but $D \circ C^1 \circ C^2$ does not satisfy $I$.
 
@@
   For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $0 < i = j + k$ that $D \circ C^1_j \circ C^2_k$
-  satisfies $I$. If $j = 0$, then $D \circ C^1_0 \circ C^2_k = D \circ C^2_k$
-  satisfies $I$ by the assumption that $C^2$ satisfies $I$ starting at $D$, and
-  likewise for when $k = 0$. Now consider $j, k > 0$. By the inductive
-  hypothesis, we know
+  \leq k \leq n$, and $i = j + k$ that $D \circ C^1_j \circ C^2_k$ satisfies
+  $I$. If $j = 0$, then $D \circ C^1_0 \circ C^2_k = D \circ C^2_k$ satisfies
+  $I$ by the assumption that $C^2$ satisfies $I$ starting at $D$, and likewise
+  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
+  know
     (1) $D \circ C^1_{j-1} \circ C^2_{k-1}$,
@@
 This section posed the question: how do we write an algorithm to decide whether
-a set of PN-counter increment/decrement transactions are \iconfluent{} with
+a set of PN-Counter increment/decrement transactions are \iconfluent{} with
 respect to an invariant expressed as a formula of linear equations and
@@
 by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. It then uses Z3 \cite{de2008z3} to check if
-$F$ is a tautology.
+\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
+to check if $F$ is a tautology.
 

### 2016-11-16 doc/vision.tex
Minor typos and fixes.
Added citation TODOs.
Maybe, I'll add a related works section and explain the related work in
more detail.

@@
 \section{Research Vision}
+
 At a very high level, this research tackles the same problems tackled by many
@@
 to programmatically resolve conflicts. IPA uses a type system to ensure weakly
-consistent data doesn't leak into strongly consistent data.
+consistent data doesn't leak into strongly consistent data. \todo{Cite and
+summarize \cite{balegas2015putting}, \cite{roy2015homeostasis},
+\cite{li2012making}, and \cite{li2014automating}}
 
@@
 # The number of employees.
-num_employees = PN-Counter(0)
+num_employees = PNCounter(0)
 # Employee ids partitioned into teams.

### 2016-11-16 doc/counters.tex
Added imp. Adding data-dependent txns.

@@
 
-\subsection{A Solution}
+\subsection{A Solution}\label{sec:solution}
 This section posed the question: how do we write an algorithm to decide whether
@@
 restatement of the definition of 1-\iconfluence!
+
+\newcommand{\imp}{$\iinvariant$MP}
+\subsection{\imp{} Transactions}\label{sec:imptxns}
+So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
+partial function from variables to integers. Intuitively, a transaction chooses
+some \emph{fixed} subset of variables and increments and decrements each by
+some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
+transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
+model, for example, a transaction that increments $x$ by 42 and decrements $y$
+by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
+transaction that increments $y$ by 1 whenever $x$ is positive and increments
+$z$ whenever $x$ is negative.
+
+In this subsection, we introduce a simple imperative programming language
+\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
+Instead of looking at fixed transactions, we'll look at \imp{}
+transactions---transactions written in \imp{}---and ask ourselves the question
+of how do determine whether a set of \imp{} transactions is \iconfluent{} with
+respect to some invariant.
+
+\newcommand{\aexp}{\textsf{Aexp}}
+\newcommand{\bexp}{\textsf{Bexp}}
+\newcommand{\com}{\textsf{Com}}
+\newcommand{\impskip}{\textsf{skip}}
+\newcommand{\impif}[3]{\textsf{if}(#1)\set{#2}\textsf{else}\set{#3}}
+\begin{figure}[h]
+  \centering
+  \begin{gather*}
+    \begin{array}{rrll}
+      \aexp\; a & ::= & k       & \text{\emph{constant}} \\
+                & |   & x       & \text{\emph{variable}} \\
+                & |   & read(x) & \text{\emph{database read}} \\
+                & |   & a + a   & \text{\emph{addition}} \\
+                & |   & a - a   & \text{\emph{subtraction}} \\
+                & |   & - a     & \text{\emph{negation}} \\
+      &&& \\
+      \bexp\; b & ::= & \textsf{true}  & \text{\emph{true}} \\
+                & |   & \textsf{false} & \text{\emph{false}} \\
+                & |   & a \odot a      & \text{\emph{linear constraint}} \\
+                & |   & \lnot b        & \text{\emph{negation}} \\
+                & |   & b \lor b       & \text{\emph{disjunction}} \\
+                & |   & b \land b      & \text{\emph{conjunction}} \\
+      &&& \\
+      \com\; c  & ::= & \impskip        & \text{\emph{skip}} \\
+                & |   & x := a          & \text{\emph{assignment}} \\
+                & |   & add(x, a)       & \text{\emph{database adds}} \\
+                & |   & c; c            & \text{\emph{sequence}} \\
+                & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
+      &&& \\
+      \odot & ::= & =    \vert
+                    \neq \vert
+                    <    \vert
+                    \leq \vert
+                    >    \vert
+                    \geq & \text{\emph{comparator}}
+    \end{array}
+  \end{gather*}
+
+  \caption{
+    A grammar for \imp{}: a simple imperative transaction language inspired by
+    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
+  }
+  \label{fig:imp-grammar}
+\end{figure}
+
+The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
+semantics of \imp{} can be defined denotationally. That is, we can define a
+denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
+$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
+from a database to a fixed transaction. Defining these semantics is left as an
+exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
+\todo{Define, explicitly, the denotational semantics if need be.}
+
+\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
+\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
+unlike fixed transaction application, \imp{} transaction application is
+\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
+\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
+\[
+  D \circ c_1 \circ c_2
+    = \set{(x, 1)} \circ c_2
+    = \set{(x, 2)}
+    \neq \set{(x, 1)}
+    = \set{(x, 0)} \circ c_1
+    = D \circ c_2 \circ c_1
+\]
+
+An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
+is a sequence of transactions $c_1, \ldots, c_n$ where $c_i \in T$ for $1 \leq
+i \leq n$. Each \imp{} transaction chain $B$ starting with database state $D$
+has a corresponding fixed transaction chain $C$. Let $D_i = D \circ c_1 \circ
+\cdots \circ c_i$. Then, the corresponding fixed transaction chain is
+$\denote{c_1}(D_0), \ldots, \denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies
+invariant} $I$ starting at $D$ if the corresponding fixed transaction chain $C$
+satisfies invariant $I$ starting at $D$.
+
+We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
+invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
+created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
+if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
+$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
+$k$-\iconfluence similar to how we did in \secref{counter-one}.
+
+Now that we've formalized what it means for a set of \imp{} transactions to be
+\iconfluent{} with respect to some invariant, how do we determine
+\iconfluence{} algorithmically? One initial hunch would be to apply the same
+technique from \secref{solution} with slight modification. Unfortunately,
+\clmref{imp-1-iconfluence} presents a challenge.
+
+\begin{claim}\label{clm:imp-1-iconfluence}
+  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
+\end{claim}
+\begin{proof}
+  Consider the following invariant and \imp{} transactions:
+  \begin{align*}
+    I &= (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
+    c_1 &= \impif{x = 0 \land y = 0}{
+      add(x, 1)
+    }{\impif{x = 1 \land y = 0} {
+      add(y, 1)
+    }{
+      \impskip
+    }} \\
+    c_2 &= \impif{x = 0 \land y = 0}{
+      add(x, -1)
+    }{\impif{x = 1 \land y = 0} {
+      add(y, 1)
+    }{
+      \impskip
+    }}
+  \end{align*}
+
+  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
+  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
+  interpret \imp{} transactions geometrically, as we did with fixed
+  transactions.
+
+  As before, invariants over $n$ variables correspond to a set of points in
+  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
+  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
+  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
+  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
+  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
+  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
+  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
+  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
+
+  Using this geometric interpretation, it's easy to perform a brute-force check
+  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
+  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
+  in database state $D = (0, 0)$. We know that
+  \begin{itemize}
+    \item $D = (0, 0)$ satisfies $I$,
+    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
+    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
+    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
+    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
+  \end{itemize}
+  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
+  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
+  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
+\end{proof}
+
+\todo{
+  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
+  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
+  reservation system to bound the amount of divergence between replicas? That
+  doesn't seem to make sense for a system with high transaction throughput.
+  Maybe there are some other tricks we can play?
+}
+
+\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
+\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
+\begin{figure}[h]
+  \centering
+
+  \newcommand{\gridandpoints}{
+    \draw[ultra thick] (-2, 0) -- (2, 0);
+    \draw[ultra thick] (0, 0) -- (0, 2);
+    \draw (-2, 0) grid (2, 2);
+    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
+      \node[point] (\x-\y) at (\x, \y) {};
+    }
+  }
+
+  \begin{subfigure}[c]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \gridandpoints{}
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:iviz}
+  \end{subfigure}
+  \begin{subfigure}[c]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \gridandpoints{}
+      \draw[vec] (0, 0) -- (1, 0);
+      \draw[vec] (1, 0) -- (1, 1);
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:c1viz}
+  \end{subfigure}
+  \begin{subfigure}[c]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}
+      \gridandpoints{}
+      \draw[vec] (0, 0) -- (-1, 0);
+      \draw[vec] (1, 0) -- (1, 1);
+    \end{tikzpicture}
+    \caption{}
+    \label{fig:c2viz}
+  \end{subfigure}
+
+  \caption{}
+  \label{fig:icviz}
+\end{figure}
+
+\subsection{Data-dependent Transactions}\label{sec:datatxns}
+\todo{Clean up.}
+\newcommand{\width}{3}
+\newcommand{\smallwidth}{2}
+\newcommand{\height}{2}
+\newcommand{\gridxy}{
+  \draw[ultra thick] (-\width, 0) -- (\width, 0);
+  \draw[ultra thick] (0, -\height) -- (0, \height);
+  \draw (-\width, -\height) grid (\width, \height);
+}
+\newcommand{\forxy}[1]{
+  \foreach \x in {-\width, ..., -1} {
+    \foreach \y in {-\height, ..., \height} {
+      #1
+    }
+  }
+  \foreach \x in {1, ..., \width} {
+    \foreach \y in {-\height, ..., \height} {
+      #1
+    }
+  }
+  \foreach \x in {0} {
+    \foreach \y in {0} {
+      #1
+    }
+  }
+}
+\newcommand{\forxysmall}[1]{
+  \foreach \x in {-\smallwidth, ..., -1} {
+    \foreach \y in {-\height, ..., \height} {
+      #1
+    }
+  }
+  \foreach \x in {1, ..., \smallwidth} {
+    \foreach \y in {-\height, ..., \height} {
+      #1
+    }
+  }
+  \foreach \x in {0} {
+    \foreach \y in {0} {
+      #1
+    }
+  }
+}
+
+\begin{tikzpicture}
+  \gridxy{}
+  \forxy{\node[point] (\x-\y) at (\x, \y) {};}
+\end{tikzpicture}
+
+\begin{tikzpicture}
+  \gridxy{}
+  \forxy{\node[point] (\x-\y) at (\x, \y) {};}
+  \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
+\end{tikzpicture}
+
+\begin{tikzpicture}
+  \gridxy{}
+  \forxy{\node[point] (\x-\y) at (\x, \y) {};}
+  \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
+\end{tikzpicture}
+
+\begin{tikzpicture}
+  \gridxy{}
+  \forxy{\node[point] (\x-\y) at (\x, \y) {};}
+  \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
+\end{tikzpicture}

### 2016-11-16 doc/counters.tex
Fixed typesetting bug.
Captioned figure.
Added wimp txns.

@@
     IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
+    The language consists of arithmetic expressions $\aexp{}$, boolean
+    expressions $\bexp{}$, and commands $\com{}$.
   }
@@
 $I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence similar to how we did in \secref{counter-one}.
+$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
 
@@
   \begin{align*}
-    I &= (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &= \impif{x = 0 \land y = 0}{
+    I &\defeq
+      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
+    c_1 &\defeq \impif{x = 0 \land y = 0}{
       add(x, 1)
@@
     }} \\
-    c_2 &= \impif{x = 0 \land y = 0}{
+    c_2 &\defeq \impif{x = 0 \land y = 0}{
       add(x, -1)
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{Visualization of $I$.}
     \label{fig:iviz}
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{Visualization of $c_1$.}
     \label{fig:c1viz}
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{Visualization of $c_2$.}
     \label{fig:c2viz}
@@
 
-  \caption{}
+  \caption{
+    A geometric interpretation of $I$, $c_1$, and $c_2$ from
+    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
+    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
+    \ints^2$.
+  }
   \label{fig:icviz}
@@
 
-\subsection{Data-dependent Transactions}\label{sec:datatxns}
-\todo{Clean up.}
-\newcommand{\width}{3}
-\newcommand{\smallwidth}{2}
+\newcommand{\wimp}{W$\iinvariant$MP}
+\subsection{\wimp{} Transactions}\label{sec:datatxns}
+We began this section by looking at fixed transactions and concluded in
+\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
+with respect to an invariant $I$, then it is also \iconfluent{}. Then in
+\secref{imptxns}, we enriched transactions from simple fixed partial functions
+to full-blown programs written in \imp{}. Unfortunately, this increase in
+expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
+equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
+
+Perhaps \imp{} is too powerful of a transaction language. In particular, it has
+two powerful features that fixed transactions do not have. First, \imp{} has
+conditionals allowing the control flow of an \imp{} program to depend on the
+database against which it is run. Second, even without conditionals, the
+behavior of an \imp{} transaction can still depend on the database state
+against which it is run. For example, the behavior of the transaction $add(x,
+read(x))$ depends on the initial value of $x$, even though the transaction does
+not use a conditional.
+
+Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
+of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
+intermediate language more powerful than fixed transactions but weaker than
+\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
+
+In this subsection, we'll consider one such language, \wimp{}: a weakened
+version of \imp{}. Then, we'll conclude that for \wimp{}---as for
+\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
+
+Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
+assignment, and conditionals. Moreover, each database variable can be added to
+at most once. Essentially, \wimp{} programs get to add a single arithmetic
+expression to each database variable. While weaker than \imp{} programs,
+\wimp{} programs are more powerful than fixed transactions because arithmetic
+expressions can include database reads and therefore \wimp{} programs can
+depend on the initial state of the database. In fact, if we remove database
+reads from \wimp{}, we get a language of fixed transactions.
+
+Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
+directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
+application is not commutative.
+
+As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
+1-\iconfluence{} and \iconfluence{}: a fact we now prove.
+
+\begin{claim}\label{clm:wimp-1-iconfluence}
+  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
+\end{claim}
+\begin{proof}
+  Consider the following invariant and \wimp{} transactions:
+  \begin{align*}
+    I   &\defeq x \neq 0 \lor y = 0 \\
+    c_1 &\defeq y.add(1) \\
+    c_2 &\defeq x.add(read(x) + 1) \\
+    c_3 &\defeq x.add(read(x) - 1)
+  \end{align*}
+
+  As usual, we can visualize invariants and transactions geometrically, as
+  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
+  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
+  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
+
+  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
+  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
+  \begin{itemize}
+    \item $D = (0, 0)$ satisfies $I$,
+    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
+    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
+    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
+    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
+  \end{itemize}
+  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
+  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
+  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
+\end{proof}
+
+\todo{
+  We're in a pickle again! I think this result also might suggest that the
+  claims in \cite{balegas2015putting} are wrong? More likely, I'm
+  misunderstanding the paper.
+}
+
+\newcommand{\width}{2}
+\newcommand{\smallwidth}{1}
 \newcommand{\height}{2}
+
 \newcommand{\gridxy}{
@@
 }
+
 \newcommand{\forxy}[1]{
   \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} {
-      #1
+    \foreach \y in {-\height, ..., \height} { #1 } }
+  \foreach \x in {1, ..., \width} {
+    \foreach \y in {-\height, ..., \height} { #1 } }
+  \foreach \x in {0} {
+    \foreach \y in {0} { #1 } }
 }
-  }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} {
-      #1
-    }
-  }
-  \foreach \x in {0} {
-    \foreach \y in {0} {
-      #1
-    }
-  }
-}
+
 \newcommand{\forxysmall}[1]{
   \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} {
-      #1
-    }
-  }
+    \foreach \y in {-\height, ..., \height} { #1 } }
   \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} {
-      #1
-    }
-  }
+    \foreach \y in {-\height, ..., \height} { #1 } }
   \foreach \x in {0} {
-    \foreach \y in {0} {
-      #1
-    }
-  }
+    \foreach \y in {0} { #1 } }
 }
 
+\begin{figure}[h]
+  \centering
+
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
     \begin{tikzpicture}
@@
     \end{tikzpicture}
+    \caption{Visualization of $I$.}
+    \label{fig:wimp-iviz}
+  \end{subfigure}
 
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-
+    \caption{Visualization of $c_1$.}
+    \label{fig:wimp-c1viz}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
     \begin{tikzpicture}
@@
     \end{tikzpicture}
-
+    \caption{Visualization of $c_2$.}
+    \label{fig:wimp-c2viz}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.32\textwidth}
+    \centering
     \begin{tikzpicture}
@@
     \end{tikzpicture}
+    \caption{Visualization of $c_3$.}
+    \label{fig:wimp-c3viz}
+  \end{subfigure}
+
+  \caption{
+    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
+    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
+    subset of the points satisfying $I$ and that not all vectors are drawn.
+  }
+  \label{fig:wimp-icviz}
+\end{figure}

### 2016-11-22 doc/counters.tex
Started drawing rambling thoughts.
Added more to i-confluence alternatives.

@@
 A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$ if $I$ evaluates to true
-after all variables in $I$ have been replaced by their mapping in $D$.
+database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
+evaluates to true after all variables in $I$ have been replaced by their
+mapping in $D$.
 
@@
 \circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$ if for every prefix $C'$ of
-$C$, $D \circ C'$ satisfies invariant $I$.
+$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
+every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
 
@@
 An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of transactions $c_1, \ldots, c_n$ where $c_i \in T$ for $1 \leq
-i \leq n$. Each \imp{} transaction chain $B$ starting with database state $D$
-has a corresponding fixed transaction chain $C$. Let $D_i = D \circ c_1 \circ
-\cdots \circ c_i$. Then, the corresponding fixed transaction chain is
-$\denote{c_1}(D_0), \ldots, \denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies
-invariant} $I$ starting at $D$ if the corresponding fixed transaction chain $C$
+is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
+$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
+state $D$ has a corresponding fixed transaction chain $C$ denoted
+$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
+corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
+\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
+$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
 satisfies invariant $I$ starting at $D$.
@@
 \end{figure}
+
+\subsection{\iconfluence{} Alternatives}
+\newcommand{\ipreservation}{$\iinvariant$-preservation}
+\newcommand{\ipreserving}{$\iinvariant$-preserving}
+\newcommand{\isafety}{$\iinvariant$-safety}
+\newcommand{\isafe}{$\iinvariant$-safe}
+
+\begin{itemize}
+  \item \todo{Clarify these rambling thoughts.}
+  \item \todo{Add $k$-\isafety to the figure.}
+  \item \todo{Find something between \isafety{} and \iconfluence.}
+\end{itemize}
+
+\begin{itemize}
+  \item \textbf{1-\isafety.}
+    $\forall D, D', c.\>
+      I(D) \land
+      I(D') \implies
+      I(D \circ \denote{c}(D'))$
+
+  \item \textbf{k-\isafety.}
+    $\forall D, D', B, C, c.\>
+        C = \denote{B}(D) \land
+       |B| \leq k-1 \land
+       I(C @ D) \land
+       I(B @ D') \implies
+       I(D \circ C \circ \denote{c}(D' \circ C))$
+
+  \item \textbf{\isafety.}
+    $\forall D, D', B, C, c.\>
+        C = \denote{B}(D) \land
+       I(C @ D) \land
+       I(B @ D') \implies
+       I(D \circ C \circ \denote{c}(D' \circ C))$
+
+  \item \textbf{\ipreservation.}
+    $\forall D, c.\>
+       I(D) \implies
+       I(D \circ \denote{c}(D)$
+
+  \item \textbf{\iconfluence.}
+    $\forall D, B_1, B_2, C_1, C_2. \>
+       C_1 = \denote{B_1}(D) \land
+       C_2 = \denote{B_2}(D) \land
+       I(B_1 @ D) \land
+       I(B_2 @ D) \implies
+       I(D \circ C_1 \circ C_2)$
+
+  \item \textbf{k-\iconfluence.}
+    $\forall D, B_1, B_2, C_1, C_2. \>
+       C_1 = \denote{B_1}(D) \land
+       C_2 = \denote{B_2}(D) \land
+       |B_1|, |B_2| \leq k \land
+       I(B_1 @ D) \land
+       I(B_2 @ D) \implies
+       I(D \circ C_1 \circ C_2)$
+
+  \item \textbf{1-\iconfluence.}
+    $\forall D, c_1, c_2.\>
+       I(D) \land
+       I(D \circ \denote{c_1}(D)) \land
+       I(D \circ \denote{c_2}(D)) \implies
+       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
+\end{itemize}
+
+\begin{claim}
+  $\text{\iconfluence} \implies \text{$k$-\iconfluence}$
+\end{claim}
+\begin{proof}
+  \todo{Trivial.}
+\end{proof}
+\begin{claim}
+  $\text{1-\iconfluence} \centernot\implies \text{$k$-\iconfluence}$
+\end{claim}
+\begin{proof}
+  See \clmref{wimp-1-iconfluence}.
+\end{proof}
+
+\begin{claim}
+  $k\text{-\iconfluence} \implies \text{1-\iconfluence}$
+\end{claim}
+\begin{proof}
+  \todo{Trivial.}
+\end{proof}
+\begin{claim}
+  $\text{$k$-\iconfluence} \centernot\implies \text{\iconfluence}$
+\end{claim}
+\begin{proof}
+  See \clmref{wimp-1-iconfluence}.
+\end{proof}
+
+\begin{claim}
+  $\text{\isafety} \centernot\implies \text{\ipreservation}$
+\end{claim}
+\begin{proof}
+  \todo{Trivial.}
+\end{proof}
+\begin{claim}
+  $\text{\ipreservation} \centernot\implies \text{\isafety}$
+\end{claim}
+\begin{proof}
+  \todo{Not sure.}
+\end{proof}
+
+\tikzstyle{txn}=[ultra thick, ->]
+\tikzstyle{dtxn}=[ultra thick, ->, dashed]
+\tikzstyle{impl}=[-implies, double distance=3pt]
+\newcommand{\inode}[3][]{
+  \node[draw, shape=circle, minimum width=0.5cm, #1] (#2) at (#3) {};
+  \node[fill, shape=circle] () at (#3) {};
+}
+\newcommand{\nnode}[3][]{
+  \node[fill, shape=circle, #1] (#2) at (#3) {};
+}
+
+\begin{figure}[h]
+  \centering
+
+  \begin{tikzpicture}
+    \node[] (isafety) at (0, 0) {\isafety};
+    \node[] (ipreservation) at ($(isafety) + (30:3)$) {\ipreservation};
+    \node[] (iconfluence) at ($(isafety) + (-30:3)$) {\iconfluence};
+    \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
+    \node[] (1-iconfluence) at ($(k-iconfluence) + (4, 0)$) {1-\iconfluence};
+
+    \draw[impl] (isafety) -> (ipreservation);
+    \draw[impl] (isafety) -> (iconfluence);
+    \draw[impl] (iconfluence) -> (k-iconfluence);
+    \draw[impl] (k-iconfluence) -> (1-iconfluence);
+  \end{tikzpicture}
+
+  \caption{}
+  \label{}
+\end{figure}
+
+\begin{figure}[h]
+  \centering
+
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D$]{d1}{0, 0}
+      \inode[label=above:$D'$]{d2}{1, 0}
+      \nnode{b1}{0, -1}
+      \nnode{b2}{1, -1}
+      \draw[txn] (d2) -- (b2) node[midway, right]{$t(D')$};
+      \draw[dtxn] (d1) -- (b1) node[midway, left]{$t(D')$};
+    \end{tikzpicture}
+    \caption{1-\isafety}
+    \label{fig:1-isafety}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D$]{a}{0, 0}
+      \nnode{b}{0, -1}
+      \draw[txn] (a) -- (b) node[midway, left]{$t(D)$};
+    \end{tikzpicture}
+    \caption{1-\ipreservation}
+    \label{fig:1-ipreservation}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D$]{a}{0, 0}
+      \inode[label=left:$D_{t}$]{l1}{$(a) + (240:1)$}
+      \inode[label=right:$D_{u}$]{r1}{$(a) + (300:1)$}
+      \nnode{b}{$(l1) + (300:1)$}
+      \draw[txn] (a) -- (l1) node[midway, left] {$t(D)$};
+      \draw[txn] (a) -- (r1) node[midway, right] {$u(D)$};
+      \draw[dtxn] (l1) -- (b) node[midway, left] {$u(D)$};
+      \draw[dtxn] (r1) -- (b) node[midway, right] {$t(D)$};
+    \end{tikzpicture}
+    \caption{1-\iconfluence}
+    \label{fig:1-iconfluence}
+  \end{subfigure}
+
+  \vspace{1cm}
+
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D_0$]{l1}{0, 4}
+      \inode[label=left:$D_1$]{l2}{0, 3}
+      \inode[label=left:$D_{n-1}$]{l3}{0, 1}
+      \nnode{l4}{0, 0}
+
+      \inode[label=above:$D_0'$]{r1}{1, 4}
+      \inode[label=right:$D_1'$]{r2}{1, 3}
+      \inode[label=right:$D_{n-1}'$]{r3}{1, 1}
+      \nnode{r4}{1, 0}
+
+      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
+      \draw[txn] (r2) -- (r3) node[midway, right]{
+        $\begin{array}{l}
+          \denote{c_2}(D_1'), \\
+          \ldots, \\
+          \denote{c_{n-1}}(D_{n-2}'), \\
+        \end{array}$
+      };
+      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_n}(D_{n-1}')$};
+
+      \draw[dtxn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
+      \draw[dtxn] (l2) -- (l3) node[midway, left]{
+        $\begin{array}{l}
+          \denote{c_2}(D_1'), \\
+          \ldots, \\
+          \denote{c_{n-1}}(D_{n-2}'), \\
+        \end{array}$
+      };
+      \draw[dtxn] (l3) -- (l4) node[midway, left]{$\denote{c_n}(D_{n-1}')$};
+    \end{tikzpicture}
+
+    \caption{\isafety}
+    \label{fig:isafety}
+  \end{subfigure}%
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:{$D = D_{t_0} = D_{u_0}$}]{a}{0, 0}
+      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1.5)$}
+      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (0, -2)$}
+      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1.5)$}
+      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (0, -2)$}
+      \nnode{b}{$(l2) + (300:1.5)$}
+
+      \draw[txn] (a) -- (l1) node[midway, left] {$t_1(D_{t_0})$};
+      \draw[txn] (a) -- (r1) node[midway, right] {$u_1(D_{u_0})$};
+      \draw[txn] (l1) -- (l2) node[midway, left] {
+        $\begin{array}{l}
+          t_2(D_{t_1}), \\
+          \ldots, \\
+          t_n(D_{t_{n-1}})
+        \end{array}$
+      };
+      \draw[txn] (r1) -- (r2) node[midway, right] {
+        $\begin{array}{l}
+          u_2(D_{u_1}), \\
+          \ldots, \\
+          u_m(D_{u_{m-1}})
+        \end{array}$
+      };
+      \draw[dtxn] (l2) -- (b) node[midway, left] {
+        $\begin{array}{l}
+          u_i(D_{u_{i-1}}), \\
+          1 \leq i \leq m
+        \end{array}$
+      };
+      \draw[dtxn] (r2) -- (b) node[midway, right] {
+        $\begin{array}{l}
+          t_i(D_{t_{i-1}}), \\
+          1 \leq i \leq n
+        \end{array}$
+      };
+    \end{tikzpicture}
+
+    \caption{\iconfluence}
+    \label{fig:iconfluence}
+  \end{subfigure}
+
+  \caption{}
+  \label{}
+\end{figure}

### 2016-11-23 doc/counters.tex
Added and shrunk figures.

@@
   \newenvironment{griddedpic}{
-    \begin{tikzpicture}
+    \begin{tikzpicture}[scale=0.4]
       \clip (-4, -1) rectangle (4, 4);
@@
 
-  \begin{subfigure}[c]{0.5\textwidth}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
     \begin{griddedpic}
@@
   \end{subfigure}
-  \begin{subfigure}[c]{0.5\textwidth}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
     \begin{griddedpic}
@@
   \end{subfigure}
-
-  \vspace{0.5cm}
-
-  \begin{subfigure}[c]{0.5\textwidth}
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
     \begin{griddedpic}
@@
   \caption{$\invd$}
-  \end{subfigure}
-
-  \vspace{0.5cm}
-
-  \begin{subfigure}[c]{\textwidth}
+
+  \end{subfigure}
+  \begin{subfigure}[b]{\textwidth}
     \centering
@@
   contradiction that $T$ is not \iconfluent{} with respect to $I$. That is,
-  there exists a database state $D$ and two chains $C^1 = t^1_1, t^1_2, \ldots,
-  t^1_m$ and $C^2 = t^2_1, t^2_2, \ldots, t^2_n$ such that $D$ satisfies $I$,
-  $C^1$ satisfies $I$ starting at $D$, and $C^2$ satisfies $I$ starting at $D$,
-  but $D \circ C^1 \circ C^2$ does not satisfy $I$.
+  there exists a database state $D$ and two chains $C^t = t_1, t_2, \ldots,
+  t_m$ and $C^u = u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$,
+  $C^t$ satisfies $I$ starting at $D$, and $C^u$ satisfies $I$ starting at $D$,
+  but $D \circ C^t \circ C^u$ does not satisfy $I$.
 
   Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^1_j \circ C^2_k$ satisfies $I$
+  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
   for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
@@
   %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^1_0
-  \circ C^2_0 = D$ which satisfies $I$ by assumption.
+  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
+  \circ C^u_0 = D$ which satisfies $I$ by assumption.
   %
   For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^1_j \circ C^2_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^1_0 \circ C^2_k = D \circ C^2_k$ satisfies
-  $I$ by the assumption that $C^2$ satisfies $I$ starting at $D$, and likewise
+  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
+  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
+  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
   for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
   know
-    (1) $D \circ C^1_{j-1} \circ C^2_{k-1}$,
-    (2) $D \circ C^1_{j-1} \circ C^2_{k}$, and
-    (3) $D \circ C^1_{j}   \circ C^2_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t^1_j$ be one
-  chain, and $t^2_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t^1_j \circ t^2_k = D \circ C^1_j \circ C^2_k$ satisfies
+    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
+    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
+    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
+  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
+  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
+  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
   $I$.
 \end{proof}
+
+\newcommand{\baseedges}{
+  \begin{scope}[on background layer]
+    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
+    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
+    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
+    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
+    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
+    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
+    \draw[dtxn] (00) -- (30);
+    \draw[dtxn] (33) -- (30);
+  \end{scope}
+}
+
+\begin{figure}[h]
+  \centering
+  \newcommand{\onescale}{1.1}
+
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3, 2/3, 3/3,
+        0/2,
+        0/1,
+        0/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+             1/3, 2/3, 3/3,
+        0/2,
+        0/1,
+        0/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {0/3} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3,      2/3, 3/3,
+        %
+        0/1,
+        0/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {0/2, 1/3} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3,      3/3,
+        0/2,
+        %
+        0/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {0/1, 1/2, 2/3} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
+      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3, 2/3,
+        0/2, 1/2,
+        0/1%
+        %
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
+      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
+      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
+      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
+      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
+      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3, 2/3, 3/3,
+        0/2, 1/2, 2/2,
+        0/1, 1/1,
+        0/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {1/0, 2/1, 3/2} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
+      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
+      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
+      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
+      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
+      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
+      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
+      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
+      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
+      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
+      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
+      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.3\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3, 2/3, 3/3,
+        0/2, 1/2, 2/2, 3/2,
+        0/1, 1/1, 2/1,
+        0/0, 1/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \foreach \x/\y in {2/0, 3/1} {
+        \inode[red]{\x\y}{\x, \y}
+      }
+      \nnode{30}{3, 0}
+      \baseedges{}
+      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
+      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
+      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
+      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
+      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
+      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
+      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
+      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
+      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
+      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
+      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
+      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
+      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
+      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
+      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
+      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\onescale]
+      \foreach \x/\y in {
+        0/3, 1/3, 2/3, 3/3,
+        0/2, 1/2, 2/2, 3/2,
+        0/1, 1/1, 2/1, 3/1,
+        0/0, 1/0, 2/0%
+      } {
+        \inode{\x\y}{\x, \y}
+      }
+      \inode[red]{30}{3, 0}
+      \baseedges{}
+      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
+      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
+      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
+      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
+      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
+      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
+      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
+      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
+      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
+      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
+      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
+      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
+      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
+      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
+      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
+      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
+      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
+      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
+    \end{tikzpicture}
+    \caption{}
+    \label{}
+  \end{subfigure}
+
+  \caption{TODO}
+  \label{fig:one-is-enough}
+\end{figure}
 
@@
 
-  \begin{subfigure}[b]{0.32\textwidth}
-    \centering
-    \begin{tikzpicture}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
       \gridxy{}
@@
   \end{subfigure}
-
-  \begin{subfigure}[b]{0.32\textwidth}
-    \centering
-    \begin{tikzpicture}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
       \gridxy{}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.32\textwidth}
-    \centering
-    \begin{tikzpicture}
+
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
       \gridxy{}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.32\textwidth}
-    \centering
-    \begin{tikzpicture}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=0.8]
       \gridxy{}
@@
   \item \todo{Find something between \isafety{} and \iconfluence.}
+  \item \todo{Add two state-based proof rules from strong enough paper limited
+              to causal consistency.}
+  \item \todo{Prove state-based proof rules imply iconfluence.}
+  \item \todo{Figure out relationship of proof rules to this stuff.}
+  \item \todo{1-iconfluence with derived txns to start}
 \end{itemize}
@@
 \begin{itemize}
-  \item \textbf{1-\isafety.}
+  \item \textbf{0-\isafety.}
     $\forall D, D', c.\>
@@
        C = \denote{B}(D) \land
-       |B| \leq k-1 \land
-       I(C @ D) \land
+       |B| \geq k \land
        I(B @ D') \implies
@@
        C = \denote{B}(D) \land
-       I(C @ D) \land
        I(B @ D') \implies
@@
 
-  \item \textbf{\ipreservation.}
-    $\forall D, c.\>
+  \item \textbf{\ipreservation.} $\forall D, c.\>
        I(D) \implies
@@
 
-\tikzstyle{txn}=[ultra thick, ->]
-\tikzstyle{dtxn}=[ultra thick, ->, dashed]
-\tikzstyle{impl}=[-implies, double distance=3pt]
-\newcommand{\inode}[3][]{
-  \node[draw, shape=circle, minimum width=0.5cm, #1] (#2) at (#3) {};
-  \node[fill, shape=circle] () at (#3) {};
-}
-\newcommand{\nnode}[3][]{
-  \node[fill, shape=circle, #1] (#2) at (#3) {};
-}
-
 \begin{figure}[h]
@@
   \begin{tikzpicture}
-    \node[] (isafety) at (0, 0) {\isafety};
-    \node[] (ipreservation) at ($(isafety) + (30:3)$) {\ipreservation};
-    \node[] (iconfluence) at ($(isafety) + (-30:3)$) {\iconfluence};
+    \node[] (0-isafety) at (0, 0) {0-\isafety};
+    \node[] (ipreservation) at ($(0-isafety) + (30:3)$) {\ipreservation};
+    \node[] (iconfluence) at ($(0-isafety) + (-30:3)$) {\iconfluence};
     \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
@@
 
-    \draw[impl] (isafety) -> (ipreservation);
-    \draw[impl] (isafety) -> (iconfluence);
+    \draw[impl] (0-isafety) -> (ipreservation);
+    \draw[impl] (0-isafety) -> (iconfluence);
     \draw[impl] (iconfluence) -> (k-iconfluence);
@@
       \inode[label=above:$D_0$]{l1}{0, 4}
-      \inode[label=left:$D_1$]{l2}{0, 3}
-      \inode[label=left:$D_{n-1}$]{l3}{0, 1}
       \nnode{l4}{0, 0}
@@
 
-      \draw[dtxn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[dtxn] (l2) -- (l3) node[midway, left]{
+      \draw[dtxn] (l1) -- (l4) node[midway, left]{
         $\begin{array}{l}
-          \denote{c_2}(D_1'), \\
-          \ldots, \\
-          \denote{c_{n-1}}(D_{n-2}'), \\
+          \denote{c_1}(D_0') \circ {} \\
+          \denote{c_2}(D_1') \circ {} \\
+          \cdots \circ {} \\
+          \denote{c_n}(D_{n-1}')
         \end{array}$
       };
-      \draw[dtxn] (l3) -- (l4) node[midway, left]{$\denote{c_n}(D_{n-1}')$};
     \end{tikzpicture}
@@
       \inode[label=above:{$D = D_{t_0} = D_{u_0}$}]{a}{0, 0}
-      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1.5)$}
-      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (0, -2)$}
-      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1.5)$}
-      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (0, -2)$}
-      \nnode{b}{$(l2) + (300:1.5)$}
+      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1)$}
+      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (240:2)$}
+      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1)$}
+      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (300:2)$}
+      \nnode{b}{$(l2) + (300:3)$}
 

### 2016-11-23 doc/counters.tex
Explained one-is-enough visualization.

@@
 
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \centering
@@
     \end{tikzpicture}
-    \caption{}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
+    \caption{$i = 0$}
+    \label{fig:one-is-enough-base}
+  \end{subfigure}
+  \begin{subfigure}[b]{0.49\textwidth}
     \centering
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=1$}
     \label{}
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=2$}
     \label{}
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=3$}
     \label{}
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=4$}
     \label{}
@@
 
-  \begin{subfigure}[b]{0.3\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \centering
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=5$}
     \label{}
   \end{subfigure}
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \centering
@@
     \end{tikzpicture}
-    \caption{}
+    \caption{$i=6$}
     \label{}
@@
 
-  \caption{TODO}
+  \caption{Illustration of the proof of \thmref{one-is-enough}.}
   \label{fig:one-is-enough}
 \end{figure}
+
+\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
+\emph{state diagrams}. In a state diagram, each dot represents a database state
+$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
+\circ t = D'$. We circle a database dot if we know it satisfies the invariant
+and leave it uncircled otherwise.
+
+\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
+chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
+u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
+By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
+\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
+want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
+\figref{one-is-enough-base} illustrates the base case of
+\thmref{one-is-enough}; each of the other six state diagrams illustrate one
+inductive step of \figref{one-is-enough}. In each inductive step, we exploit
+the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
+databases satisfy the invariant.
 

### 2016-11-24 doc/counters.tex
Added some more todo proofs.
Tidying up diagrams.
Tidied up more.
Added proofs relating o-isafety and i-strength*

@@
 evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$.
+mapping in $D$. Let $\dbs$ denote the set of all database states.
 
@@
 \subsection{\iconfluence{} Alternatives}
+
 \newcommand{\ipreservation}{$\iinvariant$-preservation}
 \newcommand{\ipreserving}{$\iinvariant$-preserving}
+
 \newcommand{\isafety}{$\iinvariant$-safety}
@@
 
+\newcommand{\istrength}{$\iinvariant$-strength}
+\newcommand{\istrong}{$\iinvariant$-strong}
+
+\newcommand{\istrengthstar}{\istrength$^\ast$}
+\newcommand{\istrongstar}{\istrong$^\ast$}
+
 \begin{itemize}
-  \item \todo{Clarify these rambling thoughts.}
-  \item \todo{Add $k$-\isafety to the figure.}
+  \item \todo{\istrengthstar $\centernot \implies$ \iconfluence?}
+  \item \todo{\iconfluence $\implies$ \istrengthstar?}
   \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{Add two state-based proof rules from strong enough paper limited
-              to causal consistency.}
-  \item \todo{Prove state-based proof rules imply iconfluence.}
-  \item \todo{Figure out relationship of proof rules to this stuff.}
   \item \todo{1-iconfluence with derived txns to start}
@@
        I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
+       I(C @ D) \land
+       I(D \circ C \circ \denote{c}{D' \circ C})$
 
@@
        I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
+
+  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
+    \leqnomode
+    \begin{flalign*}
+      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints).& \\
+      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
+      & \forall c, D, D'.\>
+          D \in I \land
+          (D, D') \in G_0^* \implies
+          (D', \denote{c}(D)(D')) \in G_0 & \tag*{S3.}
+    \end{flalign*}
+    \reqnomode
+    where $I$ is viewed as a set of database states.
+
+  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
+    \leqnomode
+    \begin{flalign*}
+      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints). & \\
+      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
+      & \forall c.\> & \tag*{T3.} \\
+      & \exists P_1, \ldots, P_n \subseteq (\var \to \ints).\> & \\
+      & \exists Q_1, \ldots, Q_n \subseteq (\var \to \ints).\> & \\
+      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
+      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
+      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
+      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
+      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
+    \end{flalign*}
+    \reqnomode
+    where $I$ is viewed as a set of database states.
 \end{itemize}
@@
 \begin{claim}
-  $\text{\iconfluence} \implies \text{$k$-\iconfluence}$
+  0-\isafety $\implies$ \istrengthstar.
 \end{claim}
 \begin{proof}
-  \todo{Trivial.}
+  Consider an abitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
+  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
+  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
+  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
+  $(D, D') \in G_0^*$.
+  \begin{align*}
+    & D \in I \land (D, D') \in G_0^*  \\
+    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
+    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
+    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
+  \end{align*}
 \end{proof}
+
 \begin{claim}
-  $\text{1-\iconfluence} \centernot\implies \text{$k$-\iconfluence}$
+  \istrengthstar $\centernot\implies$ 0-\isafety
 \end{claim}
 \begin{proof}
-  See \clmref{wimp-1-iconfluence}.
-\end{proof}
-
-\begin{claim}
-  $k\text{-\iconfluence} \implies \text{1-\iconfluence}$
-\end{claim}
-\begin{proof}
-  \todo{Trivial.}
-\end{proof}
-\begin{claim}
-  $\text{$k$-\iconfluence} \centernot\implies \text{\iconfluence}$
-\end{claim}
-\begin{proof}
-  See \clmref{wimp-1-iconfluence}.
-\end{proof}
-
-\begin{claim}
-  $\text{\isafety} \centernot\implies \text{\ipreservation}$
-\end{claim}
-\begin{proof}
-  \todo{Trivial.}
-\end{proof}
-\begin{claim}
-  $\text{\ipreservation} \centernot\implies \text{\isafety}$
-\end{claim}
-\begin{proof}
-  \todo{Not sure.}
+  Consider a database over a single variable $x$. For convenience, we'll denote
+  database states as integers. Let
+  \begin{itemize}
+    \item
+      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
+    \item
+      $I \defeq x = 0 \lor x \geq 2$, and
+    \item
+      the single \imp{} transaction $c \defeq
+      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
+  \end{itemize}
+
+  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
+  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
+  \begin{itemize}
+    \item \textbf{Case $a = 0, b = 0$.}
+      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
+    \item \textbf{Case $a = 0, b >= 2$.}
+      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
+    \item \textbf{Case $a \geq 2, b >= a$.}
+      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
+  \end{itemize}
+
+  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
+  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
 \end{proof}
@@
     \node[] (0-isafety) at (0, 0) {0-\isafety};
+    \node[] (1-isafety) at (0, -2) {1-\isafety};
+    \node[] (k-isafety) at (0, -4) {$k$-\isafety};
+    \node[] (istrengthstar) at (6, 0) {\istrengthstar};
     \node[] (ipreservation) at ($(0-isafety) + (30:3)$) {\ipreservation};
@@
     \draw[impl] (k-iconfluence) -> (1-iconfluence);
+    \draw[impl] (0-isafety) -> (istrengthstar);
+    \draw[impl] (istrengthstar) -> (iconfluence);
+    \draw[impl] (istrengthstar) -> (ipreservation);
+    \draw[impl] (0-isafety) -> (1-isafety);
+    \draw[impl] (1-isafety) -> (k-isafety);
   \end{tikzpicture}
@@
       \inode[label=above:$D'$]{d2}{1, 0}
-      \nnode{b1}{0, -1}
+      \dnode{b1}{0, -1}
       \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$t(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$t(D')$};
-    \end{tikzpicture}
-    \caption{1-\isafety}
-    \label{fig:1-isafety}
+      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
+      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
+    \end{tikzpicture}
+    \caption{0-\isafety}
+    \label{fig:0-isafety}
   \end{subfigure}%
@@
       \inode[label=above:$D$]{a}{0, 0}
-      \nnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$t(D)$};
-    \end{tikzpicture}
-    \caption{1-\ipreservation}
-    \label{fig:1-ipreservation}
+      \dnode{b}{0, -1}
+      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
+    \end{tikzpicture}
+    \caption{\ipreservation}
+    \label{fig:ipreservation}
   \end{subfigure}%
@@
       \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{t}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{u}$]{r1}{$(a) + (300:1)$}
-      \nnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$t(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$u(D)$};
-      \draw[dtxn] (l1) -- (b) node[midway, left] {$u(D)$};
-      \draw[dtxn] (r1) -- (b) node[midway, right] {$t(D)$};
+      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
+      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
+      \dnode{b}{$(l1) + (300:1)$}
+      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
+      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
+      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
+      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
     \end{tikzpicture}
@@
 
-    \begin{tikzpicture}[scale=2]
+    \begin{tikzpicture}[scale=1.5]
       \inode[label=above:$D_0$]{l1}{0, 4}
-      \nnode{l4}{0, 0}
+      \dnode[label=left:$D_1$]{l2}{0, 3}
+      \dnode[label=left:$D_2$]{l3}{0, 2}
+      \dnode[label=left:$D_3$]{l4}{0, 1}
+      \dnode{l5}{0, 0}
 
@@
       \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_{n-1}'$]{r3}{1, 1}
-      \nnode{r4}{1, 0}
+      \inode[label=right:$D_2'$]{r3}{1, 2}
+      \inode[label=right:$D_3'$]{r4}{1, 1}
+      \nnode{r5}{1, 0}
 
       \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{
-        $\begin{array}{l}
-          \denote{c_2}(D_1'), \\
-          \ldots, \\
-          \denote{c_{n-1}}(D_{n-2}'), \\
-        \end{array}$
-      };
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_n}(D_{n-1}')$};
-
-      \draw[dtxn] (l1) -- (l4) node[midway, left]{
-        $\begin{array}{l}
-          \denote{c_1}(D_0') \circ {} \\
-          \denote{c_2}(D_1') \circ {} \\
-          \cdots \circ {} \\
-          \denote{c_n}(D_{n-1}')
-        \end{array}$
-      };
-    \end{tikzpicture}
-
-    \caption{\isafety}
-    \label{fig:isafety}
+      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
+      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
+      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
+
+      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
+      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
+      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
+      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
+    \end{tikzpicture}
+
+    \caption{$k$-\isafety, $k=3$}
+    \label{fig:k-isafety}
   \end{subfigure}%
@@
 
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:{$D = D_{t_0} = D_{u_0}$}]{a}{0, 0}
-      \inode[label=left:$D_{t_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{t_n}$]{l2}{$(l1) + (240:2)$}
-      \inode[label=right:$D_{u_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{u_m}$]{r2}{$(r1) + (300:2)$}
-      \nnode{b}{$(l2) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$t_1(D_{t_0})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$u_1(D_{u_0})$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {
-        $\begin{array}{l}
-          t_2(D_{t_1}), \\
-          \ldots, \\
-          t_n(D_{t_{n-1}})
-        \end{array}$
-      };
-      \draw[txn] (r1) -- (r2) node[midway, right] {
-        $\begin{array}{l}
-          u_2(D_{u_1}), \\
-          \ldots, \\
-          u_m(D_{u_{m-1}})
-        \end{array}$
-      };
-      \draw[dtxn] (l2) -- (b) node[midway, left] {
-        $\begin{array}{l}
-          u_i(D_{u_{i-1}}), \\
-          1 \leq i \leq m
-        \end{array}$
-      };
-      \draw[dtxn] (r2) -- (b) node[midway, right] {
-        $\begin{array}{l}
-          t_i(D_{t_{i-1}}), \\
-          1 \leq i \leq n
-        \end{array}$
-      };
+    \begin{tikzpicture}[scale=1.5]
+      \inode[label=above:$D$]{a}{0, 0}
+      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
+      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
+      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
+
+      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
+      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
+      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
+
+      \dnode{b}{$(l3) + (300:3)$}
+
+      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
+      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
+      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
+      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
+      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
+      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
+      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
+      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
     \end{tikzpicture}
@@
     \label{fig:iconfluence}
+  \end{subfigure}
+
+  \begin{subfigure}[b]{0.5\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D$]{s2a}{0, 1}
+      \dnode{s2b}{0, 0}
+      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
+
+      \inode[label=above:$D$]{s3a}{2, 1}
+      \nnode[]{s3b}{2, 0}
+      \inode[label=above:$D'$]{s3c}{3, 1}
+      \nnode[]{s3d}{3, 0}
+      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
+      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
+      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
+    \end{tikzpicture}
+    \caption{\istrengthstar}
+    \label{fig:istrengthstar}
   \end{subfigure}

### 2016-11-24 doc/counters.tex
Tidied up tex a bit.

@@
   \begin{tikzpicture}
-    % \draw (0, 0) grid (4, 4);
     \node[draw] (start)  at (2, 4) {$x=2$};
@@
 \subsection{Formalizing the Question}\label{sec:counter-question}
-\newcommand{\var}{\textsf{Var}}
-\newcommand{\ints}{\mathbb{Z}}
-
 Our invariant language is defined by the grammar in \figref{invariant-grammar}
@@
 
-\newcommand{\imp}{$\iinvariant$MP}
 \subsection{\imp{} Transactions}\label{sec:imptxns}
@@
 
-\newcommand{\aexp}{\textsf{Aexp}}
-\newcommand{\bexp}{\textsf{Bexp}}
-\newcommand{\com}{\textsf{Com}}
-\newcommand{\impskip}{\textsf{skip}}
-\newcommand{\impif}[3]{\textsf{if}(#1)\set{#2}\textsf{else}\set{#3}}
 \begin{figure}[h]
@@
     \begin{array}{rrll}
-      \aexp\; a & ::= & k       & \text{\emph{constant}} \\
+      \impaexp\;
+        a & ::= & k       & \text{\emph{constant}} \\
           & |   & x       & \text{\emph{variable}} \\
@@
       &&& \\
-      \bexp\; b & ::= & \textsf{true}  & \text{\emph{true}} \\
+      \impbexp\;
+        b & ::= & \textsf{true}  & \text{\emph{true}} \\
           & |   & \textsf{false} & \text{\emph{false}} \\
@@
       &&& \\
-      \com\; c  & ::= & \impskip        & \text{\emph{skip}} \\
+      \impcom\;
+        c  & ::= & \impskip        & \text{\emph{skip}} \\
            & |   & x := a          & \text{\emph{assignment}} \\
@@
     IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\aexp{}$, boolean
-    expressions $\bexp{}$, and commands $\com{}$.
+    The language consists of arithmetic expressions $\impaexp{}$, boolean
+    expressions $\impbexp{}$, and commands $\impcom{}$.
   }
@@
 
-\newcommand{\wimp}{W$\iinvariant$MP}
 \subsection{\wimp{} Transactions}\label{sec:datatxns}
@@
 \subsection{\iconfluence{} Alternatives}
-
-\newcommand{\ipreservation}{$\iinvariant$-preservation}
-\newcommand{\ipreserving}{$\iinvariant$-preserving}
-
-\newcommand{\isafety}{$\iinvariant$-safety}
-\newcommand{\isafe}{$\iinvariant$-safe}
-
-\newcommand{\istrength}{$\iinvariant$-strength}
-\newcommand{\istrong}{$\iinvariant$-strong}
-
-\newcommand{\istrengthstar}{\istrength$^\ast$}
-\newcommand{\istrongstar}{\istrong$^\ast$}
-
 \begin{itemize}

### 2016-11-26 doc/counters.tex
Crunched figure.
Added two more proofs.

@@
   the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$, but assume for
-  contradiction that $T$ is not \iconfluent{} with respect to $I$. That is,
-  there exists a database state $D$ and two chains $C^t = t_1, t_2, \ldots,
-  t_m$ and $C^u = u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$,
-  $C^t$ satisfies $I$ starting at $D$, and $C^u$ satisfies $I$ starting at $D$,
-  but $D \circ C^t \circ C^u$ does not satisfy $I$.
+  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
+  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
+  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
+  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
+  C^t \circ C^u$ satisfies $I$.
 
@@
   \centering
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=0.8]
+  \newcommand{\thescale}{0.7}
+
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\thescale]
       \gridxy{}
@@
   \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=0.8]
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\thescale]
       \gridxy{}
@@
   \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=0.8]
+  \begin{subfigure}[b]{0.24\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=\thescale]
       \gridxy{}
@@
   \item \todo{\iconfluence $\implies$ \istrengthstar?}
+  \item \todo{\istrength $\centernot \implies$ \isafety?}
+  \item \todo{\isafety $\implies$ \istrength?}
+  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
+  \item \todo{\iconfluence $\implies$ \istrength?}
   \item \todo{Find something between \isafety{} and \iconfluence.}
@@
        I(D) \implies
-       I(D \circ \denote{c}(D)$
+       I(D \circ \denote{c}(D))$
 
@@
           (D, D') \in G_0^* \implies
-          (D', \denote{c}(D)(D')) \in G_0 & \tag*{S3.}
+          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
     \end{flalign*}
@@
       & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\> & \tag*{T3.} \\
-      & \exists P_1, \ldots, P_n \subseteq (\var \to \ints).\> & \\
-      & \exists Q_1, \ldots, Q_n \subseteq (\var \to \ints).\> & \\
+      & \forall c.\>
+        \exists P_1, \ldots, P_n \subseteq (\var \to \ints).\>
+        \exists Q_1, \ldots, Q_n \subseteq (\var \to \ints).\> & \tag*{T3.} \\
       & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
@@
 
-\begin{claim}
+\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
   0-\isafety $\implies$ \istrengthstar.
@@
 \begin{proof}
-  Consider an abitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
+  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
   respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
@@
 
-\begin{claim}
+\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
   \istrengthstar $\centernot\implies$ 0-\isafety
@@
 
+\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
+  \istrengthstar $\implies$ \ipreservation
+\end{claim}
+\begin{proof}
+  Consider a set of transactions $T$ that is \istrongstar{} with respect to
+  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
+  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
+  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
+\end{proof}
+
+\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
+  \istrengthstar{} $\implies$ \iconfluence{}.
+\end{claim}
+\begin{proof}
+  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
+  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
+  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
+  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
+  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
+  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
+  show $D \circ C^c \circ C^d$ satisfies $I$.
+
+  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
+  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
+  $j, k, j', k'$ if
+    $j \leq m$,
+    $k \leq n$,
+    $i = j + k$,
+    $j' \leq j$, and
+    $k' \leq k$, then
+  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
+
+  This proof is similar to the proof of \thmref{one-is-enough}, and we can
+  visualize it similar to how we visualized \thmref{one-is-enough} in
+  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
+  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
+  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
+  diagonal are reachable via $G_0^*$ from points above and to the left.
+
+  %
+  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
+  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
+  the definition of a reflexive transaction closure.
+  %
+  For the inductive step, consider an arbitrary set of natural numbers
+  $i, j, k, j', k'$ where
+    $0 < i \leq m + n$,
+    $j \leq m$,
+    $k \leq n$,
+    $i = j + k$,
+    $j' \leq j$
+    $k' \leq k$.
+  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
+  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
+  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
+  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
+  facts:
+  \begin{gather}
+    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
+      \label{strength-confluence-a} \\
+    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
+      \label{strength-confluence-b} \\
+    I(D \circ C^c_{j-1} \circ C^d_{k})
+      \label{strength-confluence-c} \\
+    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
+      \label{strength-confluence-d} \\
+    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
+      \label{strength-confluence-e}
+  \end{gather}
+  where
+  \begin{itemize}
+    \item
+      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
+      deduced from the inductive hypothesis applied to $i - 1$;
+    \item
+      \eqref{strength-confluence-c} is deduced from
+      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
+    \item
+      \eqref{strength-confluence-d} is deduced from the definition of a
+      reflexive transitive closure;
+    \item
+      \eqref{strength-confluence-e} is deduced from
+      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
+      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
+  \end{itemize}
+  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
+  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
+
+  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
+  that $I(D \circ D^c_m \circ D^d_n)$.
+\end{proof}
+
+\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
+  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
+\end{claim}
+\begin{proof}
+  Consider a database over a single variable $x$. For convenience, we'll denote
+  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
+  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
+  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
+  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
+  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
+  \istrongstar{} with respect to $I$.
+\end{proof}
+
 \begin{figure}[h]

### 2016-11-26 doc/counters.tex
Removed todone todos.
Tidied up implication figure.
Added 0-isafety/istrength proofs.

@@
 \begin{itemize}
-  \item \todo{\istrengthstar $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrengthstar?}
   \item \todo{\istrength $\centernot \implies$ \isafety?}
@@
   \item \todo{\iconfluence $\implies$ \istrength?}
+  \item \todo{\istrengthstar $\implies$ \istrength?}
+  \item \todo{\istrength $\implies$ \istrengthstar?}
   \item \todo{Find something between \isafety{} and \iconfluence.}
@@
 
+\begin{claim}\label{clm:0-isafety-implies-istrength}
+  0-\isafety{} $\implies$ \istrength{}.
+\end{claim}
+\begin{proof}
+  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
+  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
+  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
+  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
+  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
+  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
+  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
+\end{proof}
+
+\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
+  \istrength{} $\centernot\implies$ 0-\isafety.
+\end{claim}
+\begin{proof}
+  Consider a database over a single variable $x$. For convenience, we'll denote
+  database states as integers. Let
+  \begin{itemize}
+    \item
+      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
+    \item
+      $I \defeq x = 0 \lor x \geq 2$;
+    \item
+      the single \imp{} transaction $c \defeq
+      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
+    \item
+      $n \defeq 2$;
+    \item
+      $P_1 \defeq Q_1 \defeq \set{0}$; and
+    \item
+      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
+  \end{itemize}
+
+ First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
+ hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
+ \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
+ \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
+ arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
+ (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
+
+  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
+  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
+\end{proof}
+
 \begin{figure}[h]
@@
     \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[] (1-isafety) at (0, -2) {1-\isafety};
-    \node[] (k-isafety) at (0, -4) {$k$-\isafety};
-    \node[] (istrengthstar) at (6, 0) {\istrengthstar};
-    \node[] (ipreservation) at ($(0-isafety) + (30:3)$) {\ipreservation};
-    \node[] (iconfluence) at ($(0-isafety) + (-30:3)$) {\iconfluence};
-    \node[] (k-iconfluence) at ($(iconfluence) + (4, 0)$) {$k$-\iconfluence};
-    \node[] (1-iconfluence) at ($(k-iconfluence) + (4, 0)$) {1-\iconfluence};
-
-    \draw[impl] (0-isafety) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (iconfluence);
+    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
+    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
+    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
+    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
+    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
+    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
+    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
+    \node[below=of istrengthstar] (istrength)     {\istrength};
+
+    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
+    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
+    \draw[ultra thick, impl] (0-isafety) -> (istrength);
     \draw[impl] (iconfluence) -> (k-iconfluence);
     \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (0-isafety) -> (istrengthstar);
-    \draw[impl] (istrengthstar) -> (iconfluence);
     \draw[impl] (istrengthstar) -> (ipreservation);
@@
 
-  \caption{}
-  \label{}
+  \caption{
+    An illustration of the relationship between various \iconfluence{}
+    alternatives. We have proven bold implications in both directions. For
+    example, we have proven both \iconfluence $\protect \centernot \implies$
+    \istrengthstar.
+  }
+  \label{fig:iconfluence-alternatives}
 \end{figure}

### 2016-11-26 doc/counters.tex
Made TODOs conditionally compile.
istrong => istrong*

@@
 \subsection{\iconfluence{} Alternatives}
-\begin{itemize}
+\begin{todoitemize}
   \item \todo{\istrength $\centernot \implies$ \isafety?}
@@
   \item \todo{1-iconfluence with derived txns to start}
-\end{itemize}
+\end{todoitemize}
 
@@
     \begin{flalign*}
-      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints).& \\
+      & \exists G_0 \subseteq \dbs \times \dbs.& \\
       & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
@@
     \begin{flalign*}
-      & \exists G_0 \subseteq (\var \to \ints) \times (\var \to \ints). & \\
+      & \exists G_0 \subseteq \dbs \times \dbs. & \\
       & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
       & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq (\var \to \ints).\>
-        \exists Q_1, \ldots, Q_n \subseteq (\var \to \ints).\> & \tag*{T3.} \\
+        \exists P_1, \ldots, P_n \subseteq \dbs.\>
+        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
       & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
@@
 
+\begin{claim}\label{clm:istrength-implies-istrengthstar}
+  \istrength{} $\implies$ \istrengthstar{}.
+\end{claim}
+\begin{proof}
+  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
+  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
+  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
+  respect to $I$ with $G_0$.
+
+  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
+  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
+  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
+  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
+\end{proof}
+
 \begin{figure}[h]
@@
     \draw[impl] (1-isafety) -> (k-isafety);
+    \draw[impl] (istrength) -> (istrengthstar);
   \end{tikzpicture}
@@
       \nnode[]{s3b}{2, 0}
-      \inode[label=above:$D'$]{s3c}{3, 1}
+      \nnode[label=above:$D'$]{s3c}{3, 1}
       \nnode[]{s3d}{3, 0}

### 2016-11-26 doc/counters.tex
Added 1-iconvergence.
0-isafety => 1-iconvergence => iconfluence.

@@
 \begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
   \item \todo{\istrength $\implies$ \istrengthstar?}
+  \item \todo{\iconfluence $\implies$ 1-\iconvergence?}
   \item \todo{Find something between \isafety{} and \iconfluence.}
@@
     where $I$ is viewed as a set of database states.
+
+  \item \textbf{1-\iconvergence.}
+    $\forall D, D', D'', c', c''.\>
+      I(D) \land I(D') \land I(D'') \land
+      I(D \circ \denote{c'}(D')) \land
+      I(D \circ \denote{c''}(D'')) \implies
+      I(D \circ \denote{c'}(D') \circ \denote{c''}(D''))$
 \end{itemize}
@@
 
+\begin{claim}\label{clm:0-isafety-implies-1-iconvergence}
+  0-\isafety{} $\implies$ 1-\iconvergence{}.
+\end{claim}
+\begin{proof}
+  Consider an arbitrary set of \imp{} transactions $T$ that is \iconfluent{}
+  with respect to invariant $I$. We show that $T$ is 1-\iconvergent{} with
+  respect to $I$.
+
+  Consider arbitrary database states $D$, $D'$, $D''$ and arbitrary
+  transactions $c'$, $c''$ where $I(D)$, $I(D')$, $I(D'')$, $I(D \circ
+  \denote{c'}(D'))$, and $I(D \circ \denote{c''}(D''))$. Applying 0-\isafety{}
+  to $D \circ \denote{c'}(D')$, $D''$, and $c''$, we have $I(D \circ
+  \denote{c'}(D') \circ \denote{c''}(D''))$.
+\end{proof}
+
+\begin{claim}\label{clm:1-iconvergence-implies-iconfluence}
+  1-\iconvergence{} $\implies$ \iconfluence{}.
+\end{claim}
+\begin{proof}
+  Consider a set of \imp{} transactions $T$ that is 1-\iconvergent{} with
+  respect to an invariant $I$. Consider an arbitrary database state $D$, two
+  \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
+  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
+  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
+  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
+  show $D \circ C^c \circ C^d$ satisfies $I$.
+
+  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
+  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
+  $j, k$ if
+    $j \leq m$,
+    $k \leq n$, and
+    $i = j + k$, then
+    $I(D \circ C^c_j \circ C^d_k)$
+  %
+  The base case, when $i = 0$, is trivial. $i = j = k = 0$, and $D \circ C^c_0
+  \circ C^d_0 = D$ satisfies $I$ by assumption.
+  %
+  For the inductive step, consider arbitrary natural numbers $i, j, k$ where
+    $0 < i \leq m + n$,
+    $j \leq m$,
+    $k \leq n$, and
+    $i = j + k$.
+  If $j = 0$, then $D \circ C^c_0 \circ C^d_k = D \circ C^d_k$ satisfies $I$ by
+  assumption. A symmetric argument holds for $k=0$. Otherwise, $j, k > 0$. Let
+  \newcommand{\wh}{\widehat}
+  \begin{align*}
+    \wh{D}   &\defeq D \circ C^c_{j-1} \circ C^d_{k-1}, \\
+    \wh{D'}  &\defeq D \circ C^c_{j-1}, \\
+    \wh{D''} &\defeq D \circ C^d_{k-1}, \\
+    \wh{c'}  &\defeq c_j,\ \text{and} \\
+    \wh{c''} &\defeq d_k.
+  \end{align*}
+  By the inductive hypothesis, we have
+    $I(\wh{D})$,
+    $I(\wh{D'})$,
+    $I(\wh{D''})$,
+    $I(\wh{D} \circ \denote{\wh{c'}}(\wh{D'}))
+      = I(D \circ C^c_{j} \circ C^d_{k-1})$, and
+    $I(\wh{D} \circ \denote{\wh{c''}}(\wh{D''}))
+      = I(D \circ C^c_{j-1} \circ C^d_{k})$.
+  Applying 1-\iconvergence{} to
+    $\wh{D}$,
+    $\wh{D'}$,
+    $\wh{D''}$,
+    $\wh{c'}$, and
+    $\wh{c''}$,
+  we have
+    $I(\wh{D} \circ \denote{\wh{c'}}(\wh{D'})
+              \circ \denote{\wh{c''}}(\wh{D''}))
+      = I(D \circ C^c_j \circ C^d_k)$.
+\end{proof}
+
 \begin{figure}[h]
@@
     \node[below=of istrengthstar] (istrength)      {\istrength};
+    \node[below=of istrength]     (1-iconvergence) {1-\iconvergence};
 
@@
     \draw[impl] (istrength) -> (istrengthstar);
+    \draw[impl] (0-isafety) -> (1-iconvergence);
+    \draw[impl] (1-iconvergence) -> (iconfluence);
   \end{tikzpicture}
@@
 
-  \begin{subfigure}[b]{0.5\textwidth}
+  \begin{subfigure}[b]{0.49\textwidth}
     \centering
@@
 
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
+      \inode[label=above:$D$]{s3a}{1, 1}
+      \nnode[]{s3b}{1, 0}
+      \nnode[label=above:$D'$]{s3c}{2, 1}
+      \nnode[]{s3d}{2, 0}
       \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
@@
   \end{subfigure}
-
-  \caption{}
-  \label{}
+  \begin{subfigure}[b]{0.49\textwidth}
+    \centering
+    \begin{tikzpicture}[scale=2]
+      \inode[label=above:$D$]{D}{0, 0}
+      \inode[]{l1}{$(D) + (240:1)$}
+      \inode[]{r1}{$(D) + (300:1)$}
+      \dnode{m}{$(l1) + (300:1)$}
+
+      \inode[label=above:$D'$]{D'}{$(r1) + (1.5, 0.5)$}
+      \nnode[]{x}{$(D') + (0, -1)$}
+
+      \inode[label=above:$D''$]{D''}{$(D') + (1, 0)$}
+      \nnode[]{y}{$(D'') + (0, -1)$}
+
+      \draw[txn] (D) -- (l1) node[midway, left] {$\denote{c'}(D')$};
+      \draw[txn] (D) -- (r1) node[midway, right] {$\denote{c''}(D'')$};
+      \draw[txn] (l1) -- (m) node[midway, left] {$\denote{c''}(D'')$};
+      \draw[txn] (r1) -- (m) node[midway, right] {$\denote{c'}(D')$};
+
+      \draw[txn] (D') -- (x) node [midway, left] {$\denote{c'}(D')$};
+
+      \draw[txn] (D'') -- (y) node [midway, left] {$\denote{c''}(D'')$};
+    \end{tikzpicture}
+    \caption{1-\iconvergence}
+    \label{fig:1-iconvergence}
+  \end{subfigure}
+
+  \caption{
+    An illustration of \iconfluence{} alternatives. \todo{Describe drawings.}
+  }
+  \label{fig:iconfluence-alternatives}
 \end{figure}

### 2016-11-27 doc/counters.tex
Fixed up some figures.
Added TODO.

@@
 starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
+\todo{
+  This isn't quite the definition of \iconfluence{} Bailis uses. Ponder whether
+  this simpler definition is okay.
+}
 
@@
 
+\tikzstyle{automatable}=[fill=green!20]
+\tikzstyle{intractable}=[fill=red!20]
 \begin{figure}[h]
@@
   \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)      {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)      {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar)  {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation)  {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)    {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence)  {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence)  {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)      {\istrength};
-    \node[below=of istrength]     (1-iconvergence) {1-\iconvergence};
+    \node[automatable] (0-isafety) at (0, 0) {0-\isafety};
+    \node[automatable, below=of 0-isafety]     (1-isafety)      {1-\isafety};
+    \node[automatable, below=of 1-isafety]     (k-isafety)      {$k$-\isafety};
+    \node[intractable, right=of 0-isafety]     (istrengthstar)  {\istrengthstar};
+    \node[automatable, above=of istrengthstar] (ipreservation)  {\ipreservation};
+    \node[intractable, right=of istrengthstar] (iconfluence)    {\iconfluence};
+    \node[automatable, below=of iconfluence]   (k-iconfluence)  {$k$-\iconfluence};
+    \node[automatable, below=of k-iconfluence] (1-iconfluence)  {1-\iconfluence};
+    \node[automatable, below=of istrengthstar] (istrength)      {\istrength};
+    \node[automatable, below=of istrength]     (1-iconvergence) {1-\iconvergence};
 
@@
     alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
+    example, we have proven both \istrengthstar{} $\implies$ \iconfluence{} and
+    \iconfluence{} $\protect \centernot \implies$ \istrengthstar{}. Conditions
+    shaded green can be automatically determined using an SMT solver.
+    Conditions shaded red are not easy to automatically determine.
+  }
+  \label{fig:iconfluence-alternatives-implications}
 \end{figure}
@@
   }
-  \label{fig:iconfluence-alternatives}
+  \label{fig:iconfluence-alternatives-diagrams}
 \end{figure}

### 2017-01-24 doc/paper/composition.tex
Divided paper into more sections.

\section{\iconfluence{}, Compositionally}\label{sec:composition}

### 2017-01-24 doc/paper/counters.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
+\section{Counters}\label{sec:counters}
 As a first step towards accomplishing our research vision, let's begin with a
@@
 
-\subsection{Overview}\label{sec:counter-overview}
+% \subsection{Overview}\label{sec:counter-overview}
 Recall that a \emph{G-Counter} CRDT is an increment-only counter
@@
 \end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-24 doc/paper/formalism.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
+\section{\iconfluence{}, Formally}\label{sec:formalism}
 Our invariant language is defined by the grammar in \figref{invariant-grammar}
@@
 starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
+\todo{
+  This isn't quite the definition of \iconfluence{} Bailis uses. Ponder whether
+  this simpler definition is okay.
 }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-24 doc/paper/geometry.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
+\section{\iconfluence{}, Geometrically}\label{sec:geometry}
 \newcommand{\inva}{x + y \geq 0}
@@
 not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-24 doc/paper/imp.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
+\section{\imp{}}\label{sec:imp}
 So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
@@
 
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
 \begin{figure}[h]
@@
 \end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-24 doc/paper/related.tex
Divided paper into more sections.

\section{Related Work}\label{sec:related}
\todo{}

### 2017-01-24 doc/paper/solution.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
+\section{A Solution}\label{sec:solution}
 This section posed the question: how do we write an algorithm to decide whether
@@
 restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-24 doc/paper/wimp.tex
Divided paper into more sections.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
+\section{\wimp{}}\label{sec:wimp}
 We began this section by looking at fixed transactions and concluded in
@@
 \end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-\end{itemize}
-
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2017-01-25 doc/paper/alie.tex
Revisions.

\section{\iconfluence{}, For Real This Time}\label{sec:alie}

### 2017-01-25 doc/paper/alternativesproofs.tex
Revisions.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
-\begin{todoitemize}
-  \item \todo{\istrength $\implies$ \istrengthstar?}
-  \item \todo{\iconfluence $\implies$ 1-\iconvergence?}
-  \item \todo{Find something between \isafety{} and \iconfluence.}
-  \item \todo{1-iconfluence with derived txns to start}
-\end{todoitemize}
-
-\begin{itemize}
-  \item \textbf{0-\isafety.}
-    $\forall D, D', c.\>
-      I(D) \land
-      I(D') \implies
-      I(D \circ \denote{c}(D'))$
-
-  \item \textbf{k-\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       |B| \geq k \land
-       I(B @ D') \implies
-       I(C @ D) \land
-       I(D \circ C \circ \denote{c}{D' \circ C})$
-
-  \item \textbf{\isafety.}
-    $\forall D, D', B, C, c.\>
-       C = \denote{B}(D) \land
-       I(B @ D') \implies
-       I(D \circ C \circ \denote{c}(D' \circ C))$
-
-  \item \textbf{\ipreservation.} $\forall D, c.\>
-       I(D) \implies
-       I(D \circ \denote{c}(D))$
-
-  \item \textbf{\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{k-\iconfluence.}
-    $\forall D, B_1, B_2, C_1, C_2. \>
-       C_1 = \denote{B_1}(D) \land
-       C_2 = \denote{B_2}(D) \land
-       |B_1|, |B_2| \leq k \land
-       I(B_1 @ D) \land
-       I(B_2 @ D) \implies
-       I(D \circ C_1 \circ C_2)$
-
-  \item \textbf{1-\iconfluence.}
-    $\forall D, c_1, c_2.\>
-       I(D) \land
-       I(D \circ \denote{c_1}(D)) \land
-       I(D \circ \denote{c_2}(D)) \implies
-       I(D \circ \denote{c_1}(D) \circ \denote{c_2}(D))$
-
-  \item \textbf{\istrengthstar{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs.& \\
-      & G_0(I) \subseteq I \land{} & \tag*{S2.} \\
-      & \forall c, D, D'.\>
-          D \in I \land
-          (D, D') \in G_0^* \implies
-          (D', D' \circ \denote{c}(D)) \in G_0 & \tag*{S3.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{\istrength{} \cite{gotsman2016cause}.}
-    \leqnomode
-    \begin{flalign*}
-      & \exists G_0 \subseteq \dbs \times \dbs. & \\
-      & G_0(I) \subseteq I \land{} & \tag*{T2.} \\
-      & \forall c.\>
-        \exists P_1, \ldots, P_n \subseteq \dbs.\>
-        \exists Q_1, \ldots, Q_n \subseteq \dbs.\> & \tag*{T3.} \\
-      & \quad I \subseteq \cup_{i=1}^n P_i \land {} & \tag*{T3b.} \\
-      & \quad \forall i \in \set{1, \ldots, n}.\> & \\
-      & \quad \quad P_i \subseteq Q_i \land {} & \tag*{T3c.} \\
-      & \quad \quad G_0(Q_i) \subseteq Q_i \land {} & \tag*{T3d.} \\
-      & \quad \quad Q_i \times \denote{c}(P_i)(Q_i) \subseteq G_0 & \tag*{T3e.}
-    \end{flalign*}
-    \reqnomode
-    where $I$ is viewed as a set of database states.
-
-  \item \textbf{1-\iconvergence.}
-    $\forall D, D', D'', c', c''.\>
-      I(D) \land I(D') \land I(D'') \land
-      I(D \circ \denote{c'}(D')) \land
-      I(D \circ \denote{c''}(D'')) \implies
-      I(D \circ \denote{c'}(D') \circ \denote{c''}(D''))$
-\end{itemize}
-
+\section{\iconfluence{} Alternatives Proofs}\label{app:alternativesproofs}
 \begin{claim}\label{clm:0-isafety-implies-istrengthstar}
@@
 \end{proof}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)      {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)      {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar)  {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation)  {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)    {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence)  {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence)  {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)      {\istrength};
-    \node[below=of istrength]     (1-iconvergence) {1-\iconvergence};
-
-    \draw[ultra thick, impl] (0-isafety) -> (istrengthstar);
-    \draw[ultra thick, impl] (istrengthstar) -> (iconfluence);
-    \draw[ultra thick, impl] (0-isafety) -> (istrength);
-    \draw[impl] (iconfluence) -> (k-iconfluence);
-    \draw[impl] (k-iconfluence) -> (1-iconfluence);
-    \draw[impl] (istrengthstar) -> (ipreservation);
-    \draw[impl] (0-isafety) -> (1-isafety);
-    \draw[impl] (1-isafety) -> (k-isafety);
-    \draw[impl] (istrength) -> (istrengthstar);
-    \draw[impl] (0-isafety) -> (1-iconvergence);
-    \draw[impl] (1-iconvergence) -> (iconfluence);
-  \end{tikzpicture}
-
-  \caption{
-    An illustration of the relationship between various \iconfluence{}
-    alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{1, 1}
-      \nnode[]{s3b}{1, 0}
-      \nnode[label=above:$D'$]{s3c}{2, 1}
-      \nnode[]{s3d}{2, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{D}{0, 0}
-      \inode[]{l1}{$(D) + (240:1)$}
-      \inode[]{r1}{$(D) + (300:1)$}
-      \dnode{m}{$(l1) + (300:1)$}
-
-      \inode[label=above:$D'$]{D'}{$(r1) + (1.5, 0.5)$}
-      \nnode[]{x}{$(D') + (0, -1)$}
-
-      \inode[label=above:$D''$]{D''}{$(D') + (1, 0)$}
-      \nnode[]{y}{$(D'') + (0, -1)$}
-
-      \draw[txn] (D) -- (l1) node[midway, left] {$\denote{c'}(D')$};
-      \draw[txn] (D) -- (r1) node[midway, right] {$\denote{c''}(D'')$};
-      \draw[txn] (l1) -- (m) node[midway, left] {$\denote{c''}(D'')$};
-      \draw[txn] (r1) -- (m) node[midway, right] {$\denote{c'}(D')$};
-
-      \draw[txn] (D') -- (x) node [midway, left] {$\denote{c'}(D')$};
-
-      \draw[txn] (D'') -- (y) node [midway, left] {$\denote{c''}(D'')$};
-    \end{tikzpicture}
-    \caption{1-\iconvergence}
-    \label{fig:1-iconvergence}
-  \end{subfigure}
-
-  \caption{
-    An illustration of \iconfluence{} alternatives. \todo{Describe drawings.}
-  }
-  \label{fig:iconfluence-alternatives}
-\end{figure}

### 2017-01-25 doc/paper/geometry.tex
Revisions.

@@
 
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
+Sometimes---like with the two examples from the previous section---we can
+eyeball whether a transaction is \iconfluent{} with respect to an invariant.
+Other times, not so much. For example, which transactions are \iconfluent{}
+with respect to the invariant $\inv$? It's not immediately clear.
+Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with respect
+to $I$?'' geometrically, and doing so will make answering the question much
+easier!
 
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
+Consider databases over $n$ variables $\var = \set{x_1, \ldots, x_n}$. We can
+imagine a database $D$ as the $n$-dimensional point $(D(x_i), \ldots, D(x_n))$.
+For example, if $\var = \set{x, y}$, then the database $\set{x:1, y:2}$
+corresponds to the two-dimensional point $(1, 2)$.
+
+We can imagine $I$ as the set of points (or databases) in $\ints^n$ that
+satisfy $I$. For example, the invariant $x + y = 2$ corresponds to the set of
+two-dimensional points $(k_x, k_y)$ where $k_x + k_y = 2$. This corresponds to
+the set of databases $\set{x:k_x, y:k_y}$ where $k_x + k_y = 2$.
+
+Moreover, if we canonicalize $I$ by eliminating negations and converting to
+conjunctive normal form, then the set of points satisfying $I$ is the
+intersection (for $\land)$ and union (for $\lor$) of the solutions to a set of
+linear equations and inequalities, which is always straightforward to draw. For
+example, the set of points satisfying $\inv$ is shown in
+\figref{geometric-example}.
 
@@
 
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
+Next, we can imagine a transaction $t = \set{x_1:k_{x_1}, \ldots, x_n:k_{x_n}}$
+as an $n$-dimensional vector $(k_{x_1}, \ldots, k_{x_n})$. For example, given
+the invariant $x + y = 2$, the transaction $\set{x:2, y:4}$ corresponds to the
+vector $(2, 4)$. Applying a transaction $t$ to a database $D$ can be understood
+as adding the vector representing $t$ to the point representing $D$.  For
+example, applying the transaction $\set{x:2, y:4}$ to the database $\set{x:10,
+y:20}$ corresponds to adding the vector $(2, 4)$ to the point $(10, 20)$
+bringing us to the point $(12, 24)$.
+
+Now, the question of whether a set of transactions $T$ is \iconfluent{} with
+respect to invariant $I$ has been reduced to the following question. Starting
+at an arbitrary point $D$ in the solution space of $I$, if we can add vectors
+$t_1, t_2, \ldots, t_m$ to $D$ and vectors $u_1, u_2, \ldots, u_n$ to $D$ all
+while staying in the solution space of $I$, then are we guaranteed that $D +
+t_1 + t_2 + \ldots + t_m + u_1 + u_2 + \ldots + u_n$ is in the solution space
+of $I$?
 
 Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
+that any transaction $\set{x:k_x, y:k_y}$ where $y \geq 0$ and $|x| \leq y$ is
+\iconfluent{} with respect to $\inv$. These correspond to the vectors $(k_x,
+k_y)$ that point up and not too far left or right.

### 2017-01-25 doc/paper/imp.tex
Revisions.

@@
 transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
+of how to determine whether a set of \imp{} transactions is \iconfluent{} with
 respect to some invariant.
@@
 exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
+\todo{
+  Define, explicitly, the denotational semantics if need be. Intuitively, we'd
+  take the vanilla denotational semantics $\benote{\cdot}$ of IMP and then let
+  $\denote{c}(D) = \benote{c}(D) - D$ where $D_2 - D_1$ is the fixed
+  transaction taking database $D_1$ to $D_2$.
+}
 
@@
 satisfies invariant $I$ starting at $D$.
+%
+For example, if $B = c_1, c_2, c_3$, then $\denote{B}(D) = t_1, t_2, t_3$ where
+\begin{align*}
+  t_1 &= \denote{c_1}(D), \\
+  t_2 &= \denote{c_2}(D \circ t_1), \\
+  t_3 &= \denote{c_3}(D \circ t_1 \circ t_2)
+\end{align*}
 
@@
 $I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
+$k$-\iconfluence{} similar to how we did in \secref{oneisenough}.
 
@@
 technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
+\thmref{imp-1-iconfluence} presents a challenge.
 
-\begin{claim}\label{clm:imp-1-iconfluence}
+\begin{theorem}\label{thm:imp-1-iconfluence}
   For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
+\end{theorem}
 \begin{proof}
@@
 
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
+  As before, a database over $n$ variables corresponds to a points in
+  $\ints^n$, and an invariant $I$ corresponds to the set of all points (or
+  databases) that satisfy it. For example, $I$ is visualized in \figref{iviz}.
+  An \imp{} transaction $c$ can be thought of as vector-valued functions $c:
+  \ints^n \to \ints^n$ that maps each point $D$ in $\ints^n$ to a
+  $n$-dimensional vector $\denote{c}(D)$. For example, $c_1$ and $c_2$ are
+  visualized in \figref{c1viz} and \figref{c2viz}.  The visualization of $c_1$
+  has two vectors. The $(1, 0)$ vector originating at the origin corresponds to
+  $c_1$ adding $1$ to $x$ whenever $x = 0 \land y = 0$. Similarly, the $(0, 1)$
+  originating at $(1, 0)$ corresponds to $c_1$ adding $1$ to $y$ whenever $x =
+  1 \land y = 0$.
 
@@
 \todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
+  \thmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
   out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
@@
     A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
+    \thmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
     and $c_2$ correspond to vector valued functions of type $\ints^2 \to

### 2017-01-25 doc/paper/wimp.tex
Revisions.

@@
 \section{\wimp{}}\label{sec:wimp}
-We began this section by looking at fixed transactions and concluded in
+We began this paper by looking at fixed transactions and concluded in
 \thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
 with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
+\secref{imp}, we enriched transactions from simple fixed partial functions
 to full-blown programs written in \imp{}. Unfortunately, this increase in
 expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
+equivalent to \iconfluence{}, as we saw in \thmref{imp-1-iconfluence}.
 
@@
 
-\begin{claim}\label{clm:wimp-1-iconfluence}
+\begin{theorem}\label{thm:wimp-1-iconfluence}
   For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
+\end{theorem}
 \begin{proof}
@@
     A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
+    \thmref{wimp-1-iconfluence}. Note that the visualizations include only a
     subset of the points satisfying $I$ and that not all vectors are drawn.

### 2017-01-25 doc/paper/alie.tex
Revisions + real iconfluence definition.

@@
 \section{\iconfluence{}, For Real This Time}\label{sec:alie}
+In \secref{formalism}, we defined a set of fixed transactions $T$ to be
+\emph{\iconfluent{}} with respect to invariant $I$ if for all database states
+$D$ and chains $C_1$ and $C_2$ created from $T$, if $D$ satisfies $I$, $C_1$
+satisfies $I$ starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D
+\circ C_1 \circ C_2$ satisfies $I$. In \secref{imp}, we used an almost
+identical definition for \imp{} transactions.
+
+Regrettably, I have to admit that I've been lying to you this entire time. This
+definition of \iconfluence{} is wrong! Specifically, it is a simplification of
+the original definition originally proposed by Bailis
+\etal{}~\cite{bailis2014coordination}. In light of our little white lie, we'll
+rename \iconfluence{} to \liconfluence{}\footnote{pronounced
+``lie-confluence''} and refer to Bailis \etal's properly defined property as
+\iconfluence{} from now on.
+
+First, let's review Bailis \etal's definition of \iconfluence{}. We begin with
+a set of transactions $T$, an invariant $I$, and an arbitrary database state
+$D$ that satisfies the invariant. We then consider two arbitrary database
+states $D_1$ and $D_2$ that were derived from $D$ by applying an arbitrary
+number of transactions or merges such that the invariant was always maintained.
+If the merge of $D_1$ and $D_2$ is guaranteed to satisfy the invariant, then
+$T$ is \iconfluent{} with respect to $I$. A state diagram illustrating this
+definition is given in \figref{iconfluence-example} where $D_d$ and $D_c$ play
+the role of $D_1$ and $D_2$.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[scale=1]
+    \inode[label=above:$D$]{D}{0,0};
+    \inode[label=left:$D_a$]{a}{-2,-1};
+    \inode[label=right:$D_b$]{b}{0,-1};
+    \inode[label=right:$D_c$]{c}{2,-1};
+    \inode[label=left:$D_d$]{d}{-1,-2};
+    \dnode[label=left:$D_e$]{e}{0,-3};
+    \draw[txn] (D) -- (a);
+    \draw[txn] (D) -- (b);
+    \draw[txn] (D) -- (c);
+    \draw[txn] (a) -- (d);
+    \draw[txn] (b) -- (d);
+    \draw[txn] (d) -- (e);
+    \draw[txn] (c) -- (e);
+  \end{tikzpicture}
+  \caption{A state diagram illustrating \iconfluence{}.}
+  \label{fig:iconfluence-example}
+\end{figure}
+
+Clearly, \iconfluence{} implies \liconfluence{}; that is, if $T$ is
+\iconfluent{} with respect to $I$, then it is also \liconfluent{} with respect
+to $I$. However, \liconfluence{} does not imply \iconfluence{}. As a simple
+counterexample, let $I \defeq 0 \leq x \leq 2$ and $t_1 =
+\impif{x=0}{add(x,1)}{}$. By inspection, $\set{t_1}$ is \liconfluent{} with
+respect to $I$. However, letting $D=\set{x:0}$ and each transaction be $t_1$ in
+\figref{iconfluence-example}, we have:
+\begin{align*}
+  D   &= \set{x:0} \\
+  D_a =
+  D_b =
+  D_c &= \set{x:1} \\
+  D_d &= \set{x:2} \\
+  D_e &= \set{x:3}
+\end{align*}
+$D$, $D_a$, $D_b$, $D_c$, and $D_d$ all satisfy $I$, but $D_e$ does not. Thus,
+$T$ is not \iconfluent{} with respect to $I$.
+
+This is an unfortunate result. It says that for a set of transactions $T$ and
+invariant $I$, if \liconfluence{} does not imply \iconfluence{}, then all our
+efforts to determine \liconfluence{} are for naught. However, for some
+transactions $T$ and invariants $I$, \liconfluence{} is equivalent to
+\iconfluence{}! When this is the case, we can check whether $T$ is
+\iconfluent{} with respect to $I$ by checking if it is \liconfluent{} with
+respect to $I$.
+
+So when is \liconfluence{} equivalent to \iconfluence{}? Well, look at $D$,
+$D_a$, $D_b$, and $D_d$ in \figref{iconfluence-example}. If there existed some
+\imp{} transaction chain $B$ such that $D \circ B = D_d$ and $I(B@D)$, then we
+could remove the split and merge created by $D_a$ and $D_b$ and replace it with
+the chain $B$ from $D$ to $D_d$. This is shown in \figref{mergeremoved}. For
+simplicity, we've assumed $B = c_1, c_2, c_3$. Now that \figref{mergeremoved}
+doesn't contain any intermediate merges, \liconfluence{} is equivalent to
+\iconfluence{}.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[yscale=1.2]
+    \inode[label=above:$D$]{D}{0,0};
+    \inode[]{a}{-1,-1};
+    \inode[]{b}{-1,-2};
+    \inode[label=left:$D_d$]{d}{-1,-3};
+    \inode[label=right:$D_c$]{c}{1,-3};
+    \dnode[label=left:$D_e$]{e}{0,-4};
+    \draw[txn] (D) -- node[pos=0.6,above]{$c_1$} (a);
+    \draw[txn] (a) -- node[midway,left]{$c_2$} (b);
+    \draw[txn] (b) -- node[midway,left]{$c_3$} (d);
+    \draw[txn] (d) -- (e);
+    \draw[txn] (D) -- (c);
+    \draw[txn] (c) -- (e);
+  \end{tikzpicture}
+  \caption{A state diagram illustrating \iconfluence{}.}
+  \label{fig:mergeremoved}
+\end{figure}
+
+This suggests the following sufficient condition for \liconfluence{} to be
+equivalent to \iconfluence{}. For a set of transactions $T$ and invariant $I$,
+if for all databases $D$, chains $B_1$ and $B_2$ where $I(B_1@D)$, $I(B_2@D)$,
+and $I(D \circ \denote{B_1}(D) \circ \denote{B_2}(D))$ there exists a chain
+$B_3$ such that $D \circ B_3 = D \circ \denote{B_1}(D) \circ \denote{B_2}(D)$
+and $I(B_3@D)$, then $T$ is \liconfluent{} with respect to $I$ if and only if
+it is \iconfluent{} with respect to $I$. Intuitively, we can repeatedly replace
+merges of chains $B_1$ and $B_2$ with a single chain $B_3$. Repeating this
+process will leave us with a state diagram without any intermediate merges.
+Given such an state diagram, we can check for \liconfluence{}. If a set of
+transactions $T$ and an invariant $I$ satisfies this property, we'll say that
+$T$ is \ireplayable{} with respect to $I$.
+
+\todo{Figure out how hard it is to determine \ireplayability{}. I think the
+Collatz conjecture can be reduced to it, but I haven't worked out the kinks.}
+
+% Unfortunately, determining \ireplayability{} is hard. Like, really really hard.
+% To show just how hard it is, let me introduce the Collatz conjecture. Take any
+% positive integer $n$ and repeatedly perform the following: half $n$ if $n$ is
+% even and otherwise multiply it by $3$ and add $1$. For example, letting $n =
+% 10$, we have $10, 5, 16, 8, 4, 2, 1, 4, 2, 1, \ldots$. The Collatz conjecture
+% states that for any $n$, this process will eventually lead to the value $1$. Is
+% the Collatz conjecture true? Nobody knows. It's an unsolved problem.
+%
+% Now, we'll show that if we could determine \ireplayability{}, we could solve
+% the Collatz conjecture. Assume for contradiction that we had an algorithm to
+% determine whether a set of transactions $T$ was \ireplayable{} with respect to
+% an invariant $I$.

### 2017-01-26 doc/paper/composition.tex
Added composition section.

@@
 \section{\iconfluence{}, Compositionally}\label{sec:composition}
+If a set of \imp{} transactions $T$ is \iconfluent{} with respect to an
+invariant $I_1$ and also with respect to an invariant $I_2$, is it \iconfluent
+with respect to $I_1 \lor I_2$. In this section, we'll answer questions like
+these by seeing how we can compose transactions and invariants while preserving
+\iconfluence{}.
+
+\begin{claim}[conjunction]\label{clm:compose-conjunction}
+  For all sets of \imp{} transactions $T$ and invariants $I_1$ and $I_2$, if
+  $T$ is \iconfluent{} with respect to $I_1$ and $T$ is \iconfluent{} with
+  respect to $I_2$, then $T$ is \iconfluent{} with respect to $I_1 \land I_2$.
+\end{claim}
+\begin{proof}
+  Consider an arbitrary set of \imp{} transactions $T$ and invariants $I_1$ and
+  $I_2$ where $T$ is \iconfluent{} with respect to $I_1$ and $T$ is
+  \iconfluent{} with respect to $I_2$. We show that $T$ is \iconfluent{} with
+  respect to $I_1 \land I_2$.
+
+  Let $I \defeq I_1 \land I_2$. Consider an arbitrary database $D$ and \imp{}
+  chains $C_1$ and $C_2$ such that $I(D)$, $I(C_1@D)$, and $I(C_2@D)$. Since $I
+  \implies I_1$, we also know that $I_1(D)$, $I_1(C_1@D)$, and $I_1(C_2@D)$.
+  Because $T$ is \iconfluent{} with respect to $I_1$, $D' \defeq D \circ
+  \denote{C_1}(D) \circ \denote{C_2}(D)$ satisfies $I_1$. By a symmetric
+  argument, $D'$ also satisfies $I_2$. Thus, $D'$ satisfies $I_1 \land I_2 =
+  I$.
+\end{proof}
+
+\begin{claim}[negation]\label{clm:compose-negation}
+  If $T$ is \iconfluent{} with respect to $I$, then $T$ may not be
+  \iconfluent{} with respect to $\lnot I$.
+\end{claim}
+\begin{proof}
+  Let $t_1 \defeq add(x, 1)$ and $I \defeq x > 0$. As we saw in
+  \secref{formalism}, $\set{t_1}$ is \iconfluent{} with respect to $I$.
+  However, $\set{t_1}$ is not \iconfluent{} with respect to $\lnot I = x \leq
+  0$. For example, letting $D = \set{x:-1}$, $\lnot I(D)$ and $\lnot I(D \circ
+  t_1)$, but $D \circ t_1 \circ t_1$ does not satisfy $\lnot I$.
+\end{proof}
+
+\begin{claim}[disjunction]\label{clm:compose-disjunction}
+  If $T$ is \iconfluent{} with respect to $I_1$ and $T$ is \iconfluent{} with
+  respect to $I_2$, then $T$ may not be \iconfluent{} with respect to $I_1 \lor
+  I_2$.
+\end{claim}
+\begin{proof}
+  Let $t_1 \defeq add(x, 1)$, $I_1 \defeq x = 0$, and $I_2 \defeq x = 1$.
+  $\set{t_1}$ is vacuously \iconfluent{} with respect to $I_1$ and $I_2$.
+  However, $\set{t_1}$ is not \iconfluent{} with respect to $I_1 \lor I_2$.
+  Specifically, let $D = \set{x:0}$. $(I_1 \lor I_2)(D)$ and $I(I_1 \lor I_2)(D
+  \circ t_1)$, but $D \circ t_1 \circ t_1$ does not satisfy $I_1 \lor I_2$.
+\end{proof}
+
+\begin{claim}[union]\label{clm:compose-union}
+  If $T_1$ is \iconfluent{} with respect to $I$ and $T_2$ is \iconfluent{} with
+  respect to $I$, then $T_1 \cup T_2$ may not be \iconfluent{} with respect
+  to $I$.
+\end{claim}
+\begin{proof}
+  Let $t_1 \defeq add(x, 1)$, $t_1 \defeq add(y, 1)$, $T_1 \defeq \set{t_1}$,
+  $T_2 \defeq \set{t_2}$, and $I \defeq x = 0 \lor y = 0$. By inspection, $T_1$
+  and $T_2$ are each individually \iconfluent{} with respect to $I$. However,
+  $T \defeq \set{t_1, t_2}$ is not \iconfluent{} with respect to $I$.
+  %
+  Let $D = \set{x:0,y:0}$, $C_1 = t_1$, and $C_2 = t_2$. $I(D)$, $I(D \circ
+  t_1)$, and $I(D \circ t_2)$, but $D \circ t_1 \circ t_2$ does not satisfy
+  $I$.
+\end{proof}
+
+\begin{claim}[subset]\label{clm:compose-subset}
+  If $T$ is \iconfluent{} with respect to $I$, then every subset of $T$ is
+  \iconfluent{} with respect to $I$.
+\end{claim}
+\begin{proof}
+  This is immediate from the definition of \iconfluence{}.
+\end{proof}
+
+\begin{claim}[intersection]\label{clm:compose-intersection}
+  If $T_1$ is \iconfluent{} with respect to $I$ and $T_2$ is \iconfluent{} with
+  respect to $I$, then $T_1 \cap T_2$ is \iconfluent{} with respect to $I$.
+\end{claim}
+\begin{proof}
+  $T_1 \cap T_2$ is a subset of both $T_1$ and $T_2$, so by
+  \clmref{compose-subset}, it is \iconfluent{} with respect to $I$.
+\end{proof}

### 2017-01-27 doc/paper/alternativesproofs.tex
Showed 1-iconvergence does not imply 0-isafety.
This is a great result!

@@
 \end{proof}
+
+\begin{claim}\label{clm:1-iconvergence-not-implies-0-isafety}
+  1-\iconvergence{} $\centernot\implies$ 0-\isafety{}.
+\end{claim}
+\begin{proof}
+  Let
+    $I \defeq x = -42 \lor x \geq 0$ and
+    $c \defeq \impif{read(x)\geq0}{add(x,1)}{}$.
+  First, we show that $\set{c}$ is 1-\iconvergent{} with respec to $I$.
+  Consider an arbitrary $D$,
+    $D'$, $D''$, such that
+    $I(D)$,
+    $I(D')$,
+    $I(D'')$,
+    $I(D \circ \denote{c}(D'))$, and
+    $I(D \circ \denote{c}(D''))$.
+  We perform a case analysis on $D$.
+  \begin{itemize}
+    \item \textbf{Case 1: $D = -42$.}
+      $D'$ and $D''$ must equal both equal $-42$. If either did not, say $D'$,
+      then $D \circ \denote{c}(D')$ would be $-41$ which violates our
+      assumption that $I(D \circ \denote{c}(D'))$. If both $D'$ and $D''$ equal
+      $-42$, then $D \circ \denote{c}(D') \circ \denote{c}(D'') = -42$ which
+      satisfies $I$.
+    \item \textbf{Case 2: $D \geq 0$.}
+      Depending on the value of $D'$ and $D''$, $D \circ \denote{c}(D') \circ
+      \denote{c}(D'')$ could equal $D$, $D + 1$, or $D + 2$. All three of these
+      datbases satisfy the invariant.
+  \end{itemize}
+
+  Next, we show that $\set{t_1}$ is not 0-\isafe{} with respect to $I$. Let $D
+  = -42$ and $D = 0$. $I(D)$ and $I(D')$ but $D \circ \denote{c}(D') = -41$
+  which does not satisfy the invariant.
+\end{proof}

### 2017-01-30 doc/paper/alie.tex
Explained that i-replayability isn't good enough.

@@
 invariant $I$, if \liconfluence{} does not imply \iconfluence{}, then all our
-efforts to determine \liconfluence{} are for naught. However, for some
-transactions $T$ and invariants $I$, \liconfluence{} is equivalent to
-\iconfluence{}! When this is the case, we can check whether $T$ is
-\iconfluent{} with respect to $I$ by checking if it is \liconfluent{} with
+efforts to determine \liconfluence{} are for naught. However, not all hope is
+lost. Perhaps for some transactions $T$ and invariants $I$, \liconfluence{} is
+equivalent to \iconfluence{}. When this is the case, we can check whether $T$
+is \iconfluent{} with respect to $I$ by checking if it is \liconfluent{} with
 respect to $I$.
 
-So when is \liconfluence{} equivalent to \iconfluence{}? Well, look at $D$,
-$D_a$, $D_b$, and $D_d$ in \figref{iconfluence-example}. If there existed some
-\imp{} transaction chain $B$ such that $D \circ B = D_d$ and $I(B@D)$, then we
-could remove the split and merge created by $D_a$ and $D_b$ and replace it with
-the chain $B$ from $D$ to $D_d$. This is shown in \figref{mergeremoved}. For
-simplicity, we've assumed $B = c_1, c_2, c_3$. Now that \figref{mergeremoved}
-doesn't contain any intermediate merges, \liconfluence{} is equivalent to
-\iconfluence{}.
+\subsection{\ireplayability{}}
+So when is \liconfluence{} equivalent to \iconfluence{}? Let's take a look at a
+\emph{wrong} first guess.
+
+Look at $D$, $D_a$, $D_b$, and $D_d$ in \figref{iconfluence-example}. If there
+existed some \imp{} transaction chain $B$ such that $D \circ B = D_d$ and
+$I(B@D)$, then we could remove the split and merge created by $D_a$ and $D_b$
+and replace it with the chain $B$ from $D$ to $D_d$. This is shown in
+\figref{mergeremoved}. For simplicity, we've assumed $B = c_1, c_2, c_3$. Now
+that \figref{mergeremoved} doesn't contain any intermediate merges,
+\liconfluence{} is equivalent to \iconfluence{}.
 
@@
 
-\todo{Figure out how hard it is to determine \ireplayability{}. I think the
-Collatz conjecture can be reduced to it, but I haven't worked out the kinks.}
+Unfortunately, this line of reasoning is wrong. Consider the state diagram
+shown in \figref{ireplayability-notgoodenough}. Here, we can't use
+\ireplayability{} to reduce the diagram to a simple diamond.
+
+\begin{figure}[h]
+  \centering
+  \begin{tikzpicture}[yscale=1.2]
+    \inode[]{a}{ 1, 5};
+    \inode[]{b}{ 0, 4};
+    \inode[]{c}{ 2, 4};
+    \inode[]{d}{ 0, 3};
+    \inode[]{e}{ 2, 3};
+    \inode[]{f}{ 0, 2};
+    \dnode[]{g}{ 2, 1};
+    \foreach \x/\y in {a/b, a/c, b/d, c/d, c/e, d/f, e/g, f/g} {
+      \draw[txn] (\x) -- (\y);
+    }
+  \end{tikzpicture}
+  \caption{A state diagram illustrating the inadequacy of \ireplayability{}.}
+  \label{fig:ireplayability-notgoodenough}
+\end{figure}
+
+\begin{itemize}
+  \item \todo{%
+    Figure out if \ireplayability{} can be used to determine \iconfluence{}.
+  }
+  \item \todo{%
+    Figure out how hard it is to determine \ireplayability{}. I think the
+    Collatz conjecture can be reduced to it, but I haven't worked out the
+    kinks.
+  }
+\end{itemize}
 

### 2017-01-31 doc/paper/related.tex
Revisions.
Added related work section.

@@
 \section{Related Work}\label{sec:related}
-\todo{}
+\todo{Cite and summarize \cite{balegas2015putting} and \cite{roy2015homeostasis}}
+
+\subsection{RedBlue Consistency}
+In their 2012 OSDI publication, Li \etal{} introduce RedBlue consistency as an
+intermediary consistency level between eventual consistency and strict
+serializability \cite{li2012making}. Each transaction is labelled either red or
+blue. A RedBlue consistent system serializes red transactions while running
+blue transactions concurrently.
+
+Li \etal{} also prove that if blue operations commute with all other
+operations, then the system is convergent. That is, all replicas will converge
+to the same result. To make transactions more commutative, they divide each
+transaction into a \emph{generator transaction} and a \emph{shadow operation}.
+This is similar to how an \imp{} transaction $c$ produces a fixed transaction
+$\denote{c}(D)$. They also claim that 1-\isafety{} implies \iconfluence{}.
+
+\subsection{Automating the Choice of Consistency Levels}
+In Li \etal{}'s RedBlue consistency paper, programmers had to manually generate
+shadow operations for each transaction and manually label each shadow operation
+as either blue or red. In their 2014 USENIX ATC paper, Li \etal{} aim to
+automate both these things \cite{li2014automating}.
+
+First, we assume that transactions store all their state in a relational
+database. We treat each relation in the database as a set CRDT, and we treat
+each field in each relation as some sort of CRDT as well (e.g. LWW register,
+G-Counter). Users specify which CRDTs to use by annotating their SQL
+\texttt{CREATE} statements. Doing this, each transaction produces one shadow
+operation for every possible sequence of database operations it invokes. For
+example, assuming \texttt{x} is stored in the database as a CRDT, the following
+code produces three shadow operations:
+  the first is a no-op,
+  the second decrements \texttt{x}, and
+  the third increments \texttt{x}.
+\begin{Verbatim}
+  void generator(string s) {
+      if (SHA1(s) == SOME_CONSTANT) {
+          if (x >= 10) {
+              x -= 10;
+          }
+      } else {
+          x += 10;
+      }
+  }
+\end{Verbatim}
+
+Each shadow operation is then fed through a weakest precondition solver to
+determine the weakest precondition over the parameters of the shadow operations
+that imply the shadow operation is \isafe{}. These findings are cached. At
+runtime, a generator operation is run to produce its corresponding shadow
+operation. The shadow operation's weakest precondition is found in the cache,
+and the shadow operation is checked against it. If it satisfies the weakest
+precondition, then it is labelled blue. Otherwise, it is labelled red.
+
+\subsection{'Cause I'm Strong Enough}
+Li \etal{}'s work on automatically labelling transactions as red or blue uses
+\isafety{}. As we've seen, this is one of the weaker properties that implies
+\iconfluence{}. In their 2016 POPL paper, Gotsman \etal{} propose a much
+stronger property that implies \iconfluence{} \cite{gotsman2016cause}.
+
+Gotsman \etal{} first propose a general consistency model which subsumes a
+bunch of hybrid consistency models like RedBlue consistency. They then present
+versions of \istrength{} and  \istrengthstar{} generalized to their consistency
+model. The proof rules are very complex.

### 2018-02-14 iconfluence/z3_/z3_checker.py
Use set (not list) to avoid redundant assertions.

@@ -1,8 +1,9 @@
 from typing import cast, Any, Callable, Dict, List, Optional, Tuple
 from functools import lru_cache
 
-import z3
+from orderedset import OrderedSet
 from termcolor import colored
+import z3
 
 from .. import ast
 from .. import checker
@@ -90,55 +91,55 @@ def _var_to_z3(x: ast.EVar, venv: VersionEnv, tenv: TypeEnv) -> z3.ExprRef:
     return z3.Const(venv.get_name(x.x), _type_to_z3(tenv[x.x]))
 
 @lru_cache()
-def _empty_set_to_z3(typ: ast.TSet) -> Tuple[List[z3.ExprRef], z3.ExprRef]:
+def _empty_set_to_z3(typ: ast.TSet) -> Tuple[OrderedSet, z3.ExprRef]:
     z3_typ = _type_to_z3(typ)
     empty = z3.Const(f'{typ}.empty_set', z3_typ)
     x = z3.Const(f'{typ}.empty_set_x', z3_typ.domain())
     forall = z3.ForAll(x, z3.Select(empty, x) == z3.BoolVal(False))
-    return [forall], empty
+    return OrderedSet([forall]), empty
 
 @lru_cache()
-def _empty_map_to_z3(typ: ast.Type) -> Tuple[List[z3.ExprRef], z3.ExprRef]:
+def _empty_map_to_z3(typ: ast.Type) -> Tuple[OrderedSet, z3.ExprRef]:
     z3_typ = _type_to_z3(typ)
     none = _option_none(z3_typ.range())
     empty = z3.Const(f'{typ}.empty_map', z3_typ)
     k = z3.Const(f'{typ}.empty_set_k', z3_typ.domain())
     forall = z3.ForAll(k, z3.Select(empty, k) == none())
-    return [forall], empty
+    return OrderedSet([forall]), empty
 
 def _expr_to_z3(e: ast.Expr,
                 venv: VersionEnv,
                 tenv: TypeEnv,
                 fresh: FreshName) \
-                -> Tuple[List[z3.ExprRef], z3.ExprRef]:
-    def to_z3(e_: ast.Expr) -> Tuple[List[z3.ExprRef], z3.ExprRef]:
+                -> Tuple[OrderedSet, z3.ExprRef]:
+    def to_z3(e_: ast.Expr) -> Tuple[OrderedSet, z3.ExprRef]:
         return _expr_to_z3(e_, venv, tenv, fresh)
 
[…]

### 2018-02-15 doc/paper_redux/iconfluence.tex
Started new paper.

\section{\Iconfluence{}}

\subsection{Process-Based}
Like shapiro

\subsection{Graph-Based}
define graph with at most two parents

\subsection{Expression-Based}
define expressions formally, define what it means for a thing to satisfy it, ...

\input{iconfluence.fig}

### 2018-02-15 doc/paper_redux/iconfluence.tex
Finished first section on I-confluence.

@@
 \section{\Iconfluence{}}
+A \emph{distributed object} is a triple $(S, s_0, \join)$ where $S$ is a set of
+states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to S$
+is a commutative binary merge operator. A \emph{transaction} $t: S \to S$ is a
+function that maps one state to another. An \emph{invariant} $I: S \to
+\set{\text{true}, \text{false}}$ is a predicate on states. For example,
+$(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and $I(x) =
+x \geq 0$ is an invariant.
+
+There are three ways to think about \Iconfluence{}: a process-based approach, a
+graph-based approach, and an expression-based approach. All three approaches
+are illustrated in \figref{threeconfluencemodels} and are described in more
+detail below.
+
+\input{iconfluence.fig}
 
 \subsection{Process-Based}
-Like shapiro
+The process-based model is most similar to the process model described by
+Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
+$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.
+Processors can execute a transaction $t \in T$ to transition from state $s$ to
+state $t(s)$, or they can send their state $s$ to another processor with state
+$s'$ causing the other processor to transition from state $s'$ to state $s
+\join s'$.
+
+Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
+and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
+\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
+abbreviated \Iconfluent{}, if $s_i^n \join s_j^m$ is guaranteed to satisfy $I$.
 
 \subsection{Graph-Based}
-define graph with at most two parents
+The graph-based approach is most similar to the model used by Bailis et al.\
+in~\cite{bailis2014coordination}. We are given a directed acyclic graph in
+which vertices can have an arbitrary number of children but at most two
+parents. The vertices are labelled with states, and the edges are either
+labelled with transactions or are unlabelled if they correspond to a join. The
+graph is really just an alternate way of representing an execution in the
+process-based model, but with duplicate states collapsed into a single vertex.
+As with the process-based model, we say $T$ is \Iconfluent{} if the join of any
+two invariant satisfying states with invariant satisfying ancestors is
+guaranteed to satisfy the invariant.
 
 \subsection{Expression-Based}
-define expressions formally, define what it means for a thing to satisfy it, ...
+The expression-based approach deals with expressions built from $s_0$,
+transactions in $T$, and $\join$. That is, expressions $e$ are built from the
+grammar
+\[
+  e ::= s_0 \mid t(e) \mid e_1 \join e_2
+\]
+where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
+denoted $eval(e)$, in the obvious way:
+\begin{mathpar}
+  eval(s_0) \defeq s_0
 
-\input{iconfluence.fig}
+  eval(t(e)) \defeq t(eval(e))
+
+  eval(e_1 \join e_2) \defeq eval(e_1) \join eval(e_2)
+\end{mathpar}
+
+We say an expression $e$ satisfies $I$ if $I(eval(e))$. We say an expression
+recursively satisfies $I$, denoted $\Irec{e}$, if $e$ and all of $e$'s children
+satisfy $I$. That is,
+\begin{mathpar}
+  \Irec{s_0} \defeq I(s_0)
+
+  \Irec{t(e)} \defeq I(t(e)) \land I(e)
+
+  \Irec{e_1 \join e_2} \defeq I(e_1 \join e_2) \land I(e_1) \land I(e_2)
+\end{mathpar}
+
+We say $T$ is \Iconfluent, if for all expressions $e_1, e_2$, if $\Irec{e_1}$
+and $\Irec{e_2}$, then $I(e_1 \join e_2)$.

### 2018-02-15 doc/paper_redux/iconfluence.tex
Minor style adjustments.

@@
 states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to S$
-is a commutative binary merge operator. A \emph{transaction} $t: S \to S$ is a
-function that maps one state to another. An \emph{invariant} $I: S \to
-\set{\text{true}, \text{false}}$ is a predicate on states. For example,
-$(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and $I(x) =
-x \geq 0$ is an invariant.
+is a commutative binary merge operator that we call join. A \emph{transaction}
+$t: S \to S$ is a function that maps one state to another. An \emph{invariant}
+$I: S \to \set{\text{true}, \text{false}}$ is a predicate on states. For
+example, $(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction,
+and $I(x) = x \geq 0$ is an invariant.
 
@@
 process-based model, but with duplicate states collapsed into a single vertex.
-As with the process-based model, we say $T$ is \Iconfluent{} if the join of any
-two invariant satisfying states with invariant satisfying ancestors is
+As with the process-based model, we say that $T$ is \Iconfluent{} if the join
+of any two invariant satisfying states with invariant satisfying ancestors is
 guaranteed to satisfy the invariant.

### 2018-02-15 doc/paper_redux/diconfluence_direduction.tex
Progress on dIconfluence and dIreduction.

\section{\dIconfluence{} and \dIreduction{}}
A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s_0) \ldots))$ is an
expression that does not contain a join.

We say $T$ is \defword{\dIconfluent{}} if for every pair $e_1$ and $e_2$ of
join-free expressions, $\Irec{e_1}$ and $\Irec{e_2}$ implies that $I(e_1 \join
e_2)$. Intuitively, \dIconfluence{} is the same thing as \Iconfluence{} with
the added restriction that subexpressions are join-free.

We say $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some
join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1 \join
e_2)$.  Intuitively, \dIreduction{} allows us to eliminate join operators by
substituting $e_1 \join e_2$ (which contains 1 join) with $e_3$ (which does not
contain any joins).

\input{diconfluence_direduction.fig}

We can also interpret \dIconfluence{} and \dIreduction{} using the graph-based
model of \Iconfluence{}, as illustrated in
\figref{diconfluence_and_direduction}. Intuitively, \dIconfluence{} is the
property that any two divergent join-free paths that satisfy $I$ can be joined
together to form a new state that satisfies $I$. \dIreduction{} is the property
that the join of any two divergent join-free paths that satisfy $I$ can be
replaced with a single join-free path that satisfies $I$. Both conditions
involve diamond diagrams, hence the names \dIconfluence{} and \dIreduction{}.

\begin{claim}\clmlabel{diconfluence_direducible_implies_iconfluence}
  If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
\end{claim}
\begin{proof}
  First, we prove that for all join-free $e$, if $\Irec{e}$ then there exists a
  join free $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a
  pretty straightforward structural induction on $e$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      Trivially, $e' = s_0.$

    \item \textbf{Case 2: $e = t(e_1)$.}
      $\Irec{t(e_1)}$, so $\Irec{e_1}$. Thus, by the inductive hypothesis,
      there exists a join-free $e_1'$ such that $\Irec{e_1'}$ and $eval(e_1) =
      eval(e_1')$. Now, let $e' = t(e_1')$ which is clearly join-free.
      \[
        eval(e')
          = eval(t(e_1'))
          = t(eval(e_1'))
          = t(eval(e_1))
          = eval(t(e_1))
          = eval(e)
      \]
      This also implies that $\Irec{e'}$.

    \item \textbf{Case 3: $e = e_1 \join e_2$.}
      $\Irec{e_1 \join e_2}$, so $\Irec{e_1}$ and $\Irec{e_2}$. Thus, by the
      inductive hypothesis, there exist join-free expressions $e_1'$ and $e_2'$
      where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$,
      and $\Irec{e_2'}$. Let $e'' = e_1' \join e_2'$
      \[
        eval(e'')
          = eval(e_1' \join e_2')
          = eval(e_1') \join eval(e_2')
          = eval(e_1) \join eval(e_2)
          = eval(e_1 \join e_2)
          = eval(e)
      \]
      This implies that $\Irec{e''}$. Because $T$ is \dIreducible, there exists
      a join-free $e'$ such that $\Irec{e'}$ and $eval(e') = eval(e'') =
      eval(e)$.
  \end{itemize}

  Next, consider arbitrary expressions $e_1$ and $e_2$ where $\Irec{e_1}$ and
  $\Irec{e_2}$. To show that $T$ is \Iconfluent, we must show that $I(e_1 \join
  e_2)$. Using our little lemma above, we know there exists join-free $e_1'$
  and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join e_2')$, and
  \[
    eval(e_1' \join e_2')
      = eval(e_1') \join eval(e_2')
      = eval(e_1) \join eval(e_2)
      = eval(e_1 \join e_2)
  \]
  Thus, $I(e_1 \join e_2)$.
\end{proof}

### 2018-02-15 doc/paper_redux/iconfluence.tex
Progress on dIconfluence and dIreduction.

@@
 \section{\Iconfluence{}}
-A \emph{distributed object} is a triple $(S, s_0, \join)$ where $S$ is a set of
-states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to S$
-is a commutative binary merge operator that we call join. A \emph{transaction}
-$t: S \to S$ is a function that maps one state to another. An \emph{invariant}
-$I: S \to \set{\text{true}, \text{false}}$ is a predicate on states. For
-example, $(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction,
-and $I(x) = x \geq 0$ is an invariant.
+A \defword{distributed object} is a triple $(S, s_0, \join)$ where $S$ is a set
+of states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to
+S$ is a commutative binary merge operator that we call join. A
+\defword{transaction} $t: S \to S$ is a function that maps one state to
+another.  An \defword{invariant} $I: S \to \set{\text{true}, \text{false}}$ is
+a predicate on states. For example, $(\nats, +)$ is a distributed object, $t(x)
+= 2x$ is a transaction, and $I(x) = x \geq 0$ is an invariant.
 
@@
 \ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
-abbreviated \Iconfluent{}, if $s_i^n \join s_j^m$ is guaranteed to satisfy $I$.
+abbreviated \defword{\Iconfluent{}}, if $s_i^n \join s_j^m$ is guaranteed to
+satisfy $I$.
 
@@
 
-We say an expression $e$ satisfies $I$ if $I(eval(e))$. We say an expression
-recursively satisfies $I$, denoted $\Irec{e}$, if $e$ and all of $e$'s children
-satisfy $I$. That is,
+We say an expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if
+$I(eval(e))$. We say an expression \defword{recursively satisfies $I$}, denoted
+$\Irec{e}$, if $e$ and all of $e$'s children satisfy $I$. That is,
 \begin{mathpar}
@@
 
-  \Irec{t(e)} \defeq I(t(e)) \land I(e)
+  \Irec{t(e)} \defeq I(t(e)) \land \Irec{e}
 
-  \Irec{e_1 \join e_2} \defeq I(e_1 \join e_2) \land I(e_1) \land I(e_2)
+  \Irec{e_1 \join e_2}
+    \defeq I(e_1 \join e_2) \land \Irec{e_1} \land \Irec{e_2}
 \end{mathpar}
 
-We say $T$ is \Iconfluent, if for all expressions $e_1, e_2$, if $\Irec{e_1}$
-and $\Irec{e_2}$, then $I(e_1 \join e_2)$.
+We say $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
+$\Irec{e_1}$ and $\Irec{e_2}$ implies $I(e_1 \join e_2)$. From now on, we will
+work mostly with the expression-based model of \Iconfluence{}.

### 2018-02-15 doc/paper_redux/diconfluence_direduction.tex
Cleaned up diconfluence_direduction section.

@@
 We say $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
-join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some
-join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1 \join
+join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1 \join
 e_2)$.  Intuitively, \dIreduction{} allows us to eliminate join operators by
@@
 
-\input{diconfluence_direduction.fig}
-
-We can also interpret \dIconfluence{} and \dIreduction{} using the graph-based
-model of \Iconfluence{}, as illustrated in
-\figref{diconfluence_and_direduction}. Intuitively, \dIconfluence{} is the
-property that any two divergent join-free paths that satisfy $I$ can be joined
-together to form a new state that satisfies $I$. \dIreduction{} is the property
-that the join of any two divergent join-free paths that satisfy $I$ can be
-replaced with a single join-free path that satisfies $I$. Both conditions
-involve diamond diagrams, hence the names \dIconfluence{} and \dIreduction{}.
+% \input{diconfluence_direduction.fig}
+%
+% We can also interpret \dIconfluence{} and \dIreduction{} using the graph-based
+% model of \Iconfluence{}, as illustrated in
+% \figref{diconfluence_and_direduction}. Intuitively, \dIconfluence{} is the
+% property that any two divergent join-free paths that satisfy $I$ can be joined
+% together to form a new state that satisfies $I$. \dIreduction{} is the property
+% that the join of any two divergent join-free paths that satisfy $I$ can be
+% replaced with a single join-free path that satisfies $I$. Both conditions
+% involve diamond diagrams, hence the names \dIconfluence{} and \dIreduction{}.
 
@@
   First, we prove that for all join-free $e$, if $\Irec{e}$ then there exists a
-  join free $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a
+  join-free $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a
   pretty straightforward structural induction on $e$.
@@
       and $\Irec{e_2'}$. Let $e'' = e_1' \join e_2'$
-      \[
+      \begin{align*}
         eval(e'')
-          = eval(e_1' \join e_2')
-          = eval(e_1') \join eval(e_2')
-          = eval(e_1) \join eval(e_2)
-          = eval(e_1 \join e_2)
-          = eval(e)
-      \]
+        &= eval(e_1' \join e_2') \\
+        &= eval(e_1') \join eval(e_2') \\
+        &= eval(e_1) \join eval(e_2) \\
+        &= eval(e_1 \join e_2) \\
+        &= eval(e)
+      \end{align*}
       This implies that $\Irec{e''}$. Because $T$ is \dIreducible, there exists

### 2018-02-15 doc/paper_redux/iconfluence.tex
Cleaned up diconfluence_direduction section.

@@
 of states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to
-S$ is a commutative binary merge operator that we call join. A
-\defword{transaction} $t: S \to S$ is a function that maps one state to
-another.  An \defword{invariant} $I: S \to \set{\text{true}, \text{false}}$ is
-a predicate on states. For example, $(\nats, +)$ is a distributed object, $t(x)
-= 2x$ is a transaction, and $I(x) = x \geq 0$ is an invariant.
+S$ is a binary merge operator that we call join. A \defword{transaction} $t: S
+\to S$ is a function that maps one state to another.  An \defword{invariant}
+$I: S \to \set{\text{true}, \text{false}}$ is a predicate on states. For
+example, $(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction,
+and $I(x) = x \geq 0$ is an invariant.
 

### 2018-02-17 doc/paper_redux/sufficient_conditions.tex
WIP

\section{Sufficient Conditions for \Iconfluence{}}
\newcommand{\Icommutativity}{$I$-commutativity}
\newcommand{\Ijoinisapply}{$I$-join-is-apply}
\newcommand{\nestedapply}[3]{{#1}_1({#1}_2(\ldots ({#1}_#2(#3)) \ldots))}
\newcommand{\bart}[1]{\overline{t_{#1}}}
\newcommand{\baru}[1]{\overline{u_{#1}}}

In this section, we prove that the following three conditions imply
\Iconfluence.
\begin{enumerate}
  \item
    $(S, \join)$ is a semilattice.

  \item
    Every transaction $t \in T$ is of the form $t(s) = s \join s_t$ for some
    $s_t \in S$.

  \item
    For all $u, t \in T$, and for all $s \in S$, if $I(s)$, $I(t(s))$, and
    $I(u(s))$, then $I(t(s) \join u(s))$.
\end{enumerate}

Using \clmref{diconfluence_direducible_implies_iconfluence}, it suffices to
show that the three properties imply \dIconfluence{} and \dIreduction{}. We do
so in \clmref{threepropsdiconfluence} and \clmref{threepropsdireducible}.

\begin{claim}\clmlabel{threepropsdiconfluence}
  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIconfluence.
\end{claim}
\begin{proof}
  Consider arbitrary join-free expressions $e_1$ and $e_2$ where $\Irec{e_1}$
  and $\Irec{e_2}$. We want to show that $I(e_1 \join e_2)$. Because $e_1$ and
  $e_2$ are join-free, they are of the following form:
  \begin{mathpar}
    e_1 = \nestedapply{t}{n}{s_0}

    e_2 = \nestedapply{u}{m}{s_0}
  \end{mathpar}
  where $n, m \in \nats$ and $t_i, u_j \in T$. As a syntactic convenience, let
  $\bart{i}(e) = \nestedapply{t}{i}{e}$. Similarly define $\baru{j}(e)$.
  Then,
  \begin{mathpar}
    e_1 = \bart{n}(s_0)

    e_2 = \baru{m}(s_0)
  \end{mathpar}
  We prove by strong mathematical induction on m that for all i + j = m, I(barti join baruj).kjkj

\end{proof}

\begin{claim}\clmlabel{threepropsdireducible}
  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIreduction.
\end{claim}
\begin{proof}

\end{proof}

TODO(mwhittaker): Give examples of when these conditions are met.
TODO(mwhittaker): Explain that 1-I-confluence is necessary.

### 2018-02-18 doc/paper_redux/iconfluence.tex
Some work on opbased iconfluence.
WIP
Added operation based approach to tex.

@@
 \section{\Iconfluence{}}
-A \defword{distributed object} is a triple $(S, s_0, \join)$ where $S$ is a set
-of states, $s_0 \in S$ is a designated start state, and $\join: S \times S \to
-S$ is a binary merge operator that we call join. A \defword{transaction} $t: S
-\to S$ is a function that maps one state to another.  An \defword{invariant}
-$I: S \to \set{\text{true}, \text{false}}$ is a predicate on states. For
-example, $(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction,
-and $I(x) = x \geq 0$ is an invariant.
+In this section, we present six different ways to think about \Iconfluence{}:
+three state-based approaches and three operation-based approaches.
 
-There are three ways to think about \Iconfluence{}: a process-based approach, a
-graph-based approach, and an expression-based approach. All three approaches
-are illustrated in \figref{threeconfluencemodels} and are described in more
-detail below.
+\subsection{State-Based}
+\begin{definition}
+  A \defword{distributed state-based object} is a triple $(S, s_0, \join)$
+  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
+  $\join: S \times S \to S$ is a binary merge operator that we call join.
+\end{definition}
 
-\input{iconfluence.fig}
+The notion of state-based objects is taken from~\cite{shapiro2011conflict}.
 
-\subsection{Process-Based}
+\begin{definition}
+  A \defword{state-based transaction} $t: S \to S$ is a function that maps one
+  state to another.
+\end{definition}
+
+\begin{definition}
+  An \defword{invariant} $I: S \to \set{\text{true}, \text{false}}$ is a
+  predicate on states.
+\end{definition}
+
+\begin{example}
+  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
+  $I(x) = x \geq 0$ is an invariant.
+\end{example}
+
+There are three ways to think about state-based \Iconfluence{}: a process-based
+approach, a graph-based approach, and an expression-based approach. All three
+approaches are illustrated in \figref{statebasedmodels} and are described in
+more detail below.
+
+\input{statebased_iconfluence.fig}
+
+\subsubsection{Process-Based}
 The process-based model is most similar to the process model described by
@@
 
-\subsection{Graph-Based}
+\subsubsection{Graph-Based}
 The graph-based approach is most similar to the model used by Bailis et al.\
@@
 process-based model, but with duplicate states collapsed into a single vertex.
-As with the process-based model, we say that $T$ is \Iconfluent{} if the join
-of any two invariant satisfying states with invariant satisfying ancestors is
-guaranteed to satisfy the invariant.
+As with the process-based model, we say that $T$ is \defword{\Iconfluent{}} if
+the join of any two invariant satisfying states with invariant satisfying
+ancestors is guaranteed to satisfy the invariant.
 
-\subsection{Expression-Based}
+\subsubsection{Expression-Based}
 The expression-based approach deals with expressions built from $s_0$,
@@
 
-We say an expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if
-$I(eval(e))$. We say an expression \defword{recursively satisfies $I$}, denoted
-$\Irec{e}$, if $e$ and all of $e$'s children satisfy $I$. That is,
+\begin{definition}
+  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
+\end{definition}
+
+\begin{definition}
+  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
+  and all of $e$'s children satisfy $I$. That is,
   \begin{mathpar}
@@
   \end{mathpar}
+\end{definition}
 
-We say $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
-$\Irec{e_1}$ and $\Irec{e_2}$ implies $I(e_1 \join e_2)$. From now on, we will
-work mostly with the expression-based model of \Iconfluence{}.
+\begin{definition}
+  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
+  $\Irec{e_1}$ and $\Irec{e_2}$ implies $I(e_1 \join e_2)$.
+\end{definition}
+
+\subsection{Operation-Based}
+\begin{definition}
+  A \defword{distributed operation-based object} is a pair $(S, s_0)$
+  where $S$ is a set of states and $s_0 \in S$ is a designated start state.
+\end{definition}
+
+The notion of operation-based objects is also taken
+from~\cite{shapiro2011conflict}. Note that we do not have a join function like
+we did with state-based objects.
+
+\begin{definition}
+  An \defword{operation-based transaction} $t: S \to (S \to S)$ is a function
+  that maps a state to a \defword{shadow transaction} $t(s): S \to S$. The term
+  shadow transaction is taken from~\cite{li2014automating}.
+\end{definition}
+
+The definition of invariants remains unchanged.
+
+\begin{example}
+  $(\nats, 0)$ is a distributed object. Let $f_k(x) = x + k$. $t(x) = f_x$ is a
+  transaction. $I(x) = x \geq 0$ is an invariant.
+\end{example}
+
+As with state-based objects, there are three ways to think about
+operation-based \Iconfluence{}: a process-based approach, a graph-based
+approach, and an expression-based approach. All three approaches are
+illustrated in \figref{opbasedmodels} and are described in more detail below.
+
+\input{opbased_iconfluence.fig}
+
+\subsubsection{Process-Based}
+As with the state-based approach, the process-based model is most similar to
+the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
+We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
+state $s_0$. Repeatedly, a processor executes a transaction $t$ on its current
+state $s$ and then transitions from state $s$ to state $t(s)(s)$. It also
+broadcasts $t(s)$ to all other processes exactly once. When a processor with
+state $s'$ receives $t(s)$, it transitions from state $s'$ to state $t(s)(s')$.
+
+Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
+and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
+\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
+abbreviated \defword{\Iconfluent{}}, if for all $t \in T$, $I(t(s_i^n)(s_i^n))$
+implies $I(t(s_i^n)(s_j^m))$.
+
+\subsubsection{Graph-Based}
+In the graph-based model, we are given a directed acyclic graph in which each
+vertex is associated with a state $s$ and each edge is labelled with a shadow
+operation $t(s)$ where $s$ is associated with some vertex in the graph.
+%
+Consider two invariant-satisfying states $s$ and $s'$ with invariant satisfying
+ancestors. $T$ is \defword{\Iconfluent{}} if for all $t$, $I(t(s)(s))$ implies
+$I(t(s)(s'))$.
+
+\subsubsection{Expression-Based}
+The expression-based approach deals with expressions built from $s_0$ and
+transactions in $T$. That is, expressions $e$ are built from the grammar
+\[
+  e ::= s_0 \mid t(e_1)(e_2)
+\]
+where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
+denoted $eval(e)$, in the obvious way:
+\begin{mathpar}
+  eval(s_0) \defeq s_0
+
+  eval(t(e_1)(e_2)) \defeq t(eval(e_1))(eval(e_2))
+\end{mathpar}
+
+\begin{definition}
+  An expression $e$ \defword{satisfies $I$}, denoted $I(e)$, if $I(eval(e))$.
+\end{definition}
+
+\begin{definition}
+  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
+  and all of $e$'s children satisfy $I$. That is,
+  \begin{mathpar}
+    \Irec{s_0} \defeq I(s_0)
+
+    \Irec{t(e_1)(e_2)} \defeq I(t(e_1)(e_2)) \land \Irec{e_1} \land \Irec{e_2}
+  \end{mathpar}
+\end{definition}
+
+\begin{definition}
+  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
+  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$ implies $I(t(e_1)(e_2))$.
+\end{definition}

### 2018-02-18 doc/paper_redux/sufficient_conditions.tex
Added sufficient conditions.

@@
-\section{Sufficient Conditions for \Iconfluence{}}
-\newcommand{\Icommutativity}{$I$-commutativity}
-\newcommand{\Ijoinisapply}{$I$-join-is-apply}
-\newcommand{\nestedapply}[3]{{#1}_1({#1}_2(\ldots ({#1}_#2(#3)) \ldots))}
-\newcommand{\bart}[1]{\overline{t_{#1}}}
-\newcommand{\baru}[1]{\overline{u_{#1}}}
+\section{Sufficient Conditions}
+In this section, we present sufficient conditions for state-based and
+operation-based invariant confluence. It turns out that these conditions are
+sufficient but not necessary. However, if we relax the assumption of a fixed
+start state $s_0$, then they are both sufficient and necessary.
 
-In this section, we prove that the following three conditions imply
-\Iconfluence.
-\begin{enumerate}
-  \item
-    $(S, \join)$ is a semilattice.
-
-  \item
-    Every transaction $t \in T$ is of the form $t(s) = s \join s_t$ for some
-    $s_t \in S$.
-
-  \item
-    For all $u, t \in T$, and for all $s \in S$, if $I(s)$, $I(t(s))$, and
-    $I(u(s))$, then $I(t(s) \join u(s))$.
-\end{enumerate}
-
-Using \clmref{diconfluence_direducible_implies_iconfluence}, it suffices to
-show that the three properties imply \dIconfluence{} and \dIreduction{}. We do
-so in \clmref{threepropsdiconfluence} and \clmref{threepropsdireducible}.
-
-\begin{claim}\clmlabel{threepropsdiconfluence}
-  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIconfluence.
+\begin{claim}\clmlabel{statebasedsufficient}
+  If for all states $s_1$ and $s_2$, $I(s_1)$ and $I(s_2)$ implies $I(s_1 \join
+  s_2)$, then $T$ is \Iconfluent{}. In other words, if invariant satisfying
+  states are closed under join, then $T$ is \Iconfluent.
 \end{claim}
 \begin{proof}
-  Consider arbitrary join-free expressions $e_1$ and $e_2$ where $\Irec{e_1}$
-  and $\Irec{e_2}$. We want to show that $I(e_1 \join e_2)$. Because $e_1$ and
-  $e_2$ are join-free, they are of the following form:
-  \begin{mathpar}
-    e_1 = \nestedapply{t}{n}{s_0}
-
-    e_2 = \nestedapply{u}{m}{s_0}
-  \end{mathpar}
-  where $n, m \in \nats$ and $t_i, u_j \in T$. As a syntactic convenience, let
-  $\bart{i}(e) = \nestedapply{t}{i}{e}$. Similarly define $\baru{j}(e)$.
-  Then,
-  \begin{mathpar}
-    e_1 = \bart{n}(s_0)
-
-    e_2 = \baru{m}(s_0)
-  \end{mathpar}
-  We prove by strong mathematical induction on m that for all i + j = m, I(barti join baruj).kjkj
-
+  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Then,
+  $I(eval(e_1))$ and $I(eval(e_2))$, so $I(eval(e_1) \join eval(e_2))$ which
+  implies $I(e_1 \join e_2)$.
 \end{proof}
 
-\begin{claim}\clmlabel{threepropsdireducible}
-  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIreduction.
+\begin{claim}\clmlabel{opbasedsufficient}
+  If for all transactions $t \in T$ and for all states $s_1$ and $s_2$,
+  $I(s_1)$ and $I(s_2)$ and $I(t(s_1)(s_1))$ implies $I(t(s_1)(s_2))$, then $T$
+  is \Iconfluent{}.
 \end{claim}
 \begin{proof}
-
+  Let $t \in T$ and $e_1, e_2$ be expressions such that $\Irec{e_1}$,
+  $\Irec{e_2}$, and $I(t(e_1)(e_2))$. Let $s_1 = eval(e_1)$ and $s_2 =
+  eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, so
+  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$.
 \end{proof}
 
-TODO(mwhittaker): Give examples of when these conditions are met.
-TODO(mwhittaker): Explain that 1-I-confluence is necessary.
+While the conditions in \clmref{statebasedsufficient} and
+\clmref{opbasedsufficient} are sufficient, they are not necessary.
+
+\begin{claim}\clmlabel{statebasednotnecessary}
+  If $T$ is \Iconfluent, it may not satisfy the condition in
+  \clmref{statebasedsufficient}.
+\end{claim}
+\begin{proof}
+  Consider the state based object $(S, s_0, \join)$ where $S$ is the set of all
+  finite non-empty sets of integers, $s_0 = \set{0}$, and $\join = \cup$. We
+  have a single transaction $t(x) = x \cup \set{\max(x) + 2}$ which retrieves
+  the largest element from the set (which is guaranteed to exist because it is
+  finite and non-empty), adds two to it, and then adds it into the set. $T =
+  \set{t}$. Let $I$ be the invariant that all elements of $S$ are either odd or
+  even.
+
+  Note that $T$ is \Iconfluent{}. Beginning with our initial state, we can
+  derive states $\set{0}, \set{0, 2}, \set{0, 2, 4}, \ldots$ all of which
+  satisfy the invariant. Moreover, merging any such states satisfies the
+  invariant.
+  %
+  However, $T$ doesn't satisfy the condition in \clmref{statebasedsufficient}.
+  $\set{0}$ and $\set{1}$ satisfy the invariant, but $\set{0} \cup \set{1}$
+  does not.
+\end{proof}
+
+\begin{claim}\clmlabel{opbasednotnecessary}
+  If $T$ is \Iconfluent, it may not satisfy the condition in
+  \clmref{opbasedsufficient}.
+\end{claim}
+\begin{proof}
+  We play the same trick as \clmref{statebasednotnecessary}. Let $S$, $s_0$,
+  and $I$ be as in \clmref{statebasednotnecessary}. Consider a single
+  transaction $t(x) = u_{\max(x) + 2}$ where $u_k(x) = x \cup \set{k}$. Let $T
+  = \set{t}$.
+
+  $T$ is \Iconfluent{}. Beginning with our initial state, we can derive any
+  finite set of even integers containing $0$. All these states satisfy the
+  invariant. Letting $s, s'$ be any of these states, $I(t(s)(s'))$.
+  %
+  However, $T$ doesn't satisfy the condition in \clmref{opbasedsufficient}.
+  $\set{0}$ and $\set{1}$ satisfy the invariant, and so does
+  $t(\set{0})(\set{0})$, but $t(\set{0})(\set{1}) = \set{1, 2}$ does not.
+\end{proof}
+
+\todo{Think of more compelling examples.}
+
+The conditions in \clmref{statebasedsufficient} and \clmref{opbasedsufficient}
+are not necessary. Intuitively, this makes quite a bit of sense. Neither
+condition depended on the initial state $s_0$, and both
+\clmref{statebasednotnecessary} and \clmref{opbasednotnecessary} exploit this.
+If we relax our assumption of a fixed start state, then both these conditions
+become sufficient and necessary.
+
+\todo{Prove this more carefully. It's pretty straightforward though.}

### 2018-02-18 doc/paper_redux/sufficient_conditions.tex
Tweak to sufficient conditions.

@@
 are not necessary. Intuitively, this makes quite a bit of sense. Neither
-condition depended on the initial state $s_0$, and both
+condition depends on the initial state $s_0$, and both
 \clmref{statebasednotnecessary} and \clmref{opbasednotnecessary} exploit this.
 If we relax our assumption of a fixed start state, then both these conditions
-become sufficient and necessary.
+are sufficient and necessary. Similarly, if every state in $S$ is reachable
+from $s_0$, then the conditions are sufficient and necessary.
 
-\todo{Prove this more carefully. It's pretty straightforward though.}
+\todo{Prove these facts more carefully. It's pretty straightforward though.}

### 2018-02-20 doc/paper_redux/new_statebased_sufficient.tex
WIP

--- 2018-02-15 doc/paper_redux/diconfluence_direduction.tex
@@
-\section{\dIconfluence{} and \dIreduction{}}
+\section{New Sufficient State-Based Conditions}
+In this section, we outline a new sufficient condition for state-based
+\Iconfluence{} that is more general than the one in
+\clmref{statebasedsufficient}. To prove that the condition is sufficient, we
+introduce the concepts of \dIconfluence{} and \dIreduction{}.
+
+\begin{definition}
   A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s_0) \ldots))$ is an
   expression that does not contain a join.
+\end{definition}
 
-We say $T$ is \defword{\dIconfluent{}} if for every pair $e_1$ and $e_2$ of
-join-free expressions, $\Irec{e_1}$ and $\Irec{e_2}$ implies that $I(e_1 \join
-e_2)$. Intuitively, \dIconfluence{} is the same thing as \Iconfluence{} with
+\begin{definition}
+  $T$ is \defword{\dIconfluent{}} if for every pair $e_1$ and $e_2$ of
+  join-free expressions, $\Irec{e_1}$ and $\Irec{e_2}$ implies that $I(e_1
+  \join e_2)$.
+\end{definition}
+
+Intuitively, \dIconfluence{} is the same thing as \Iconfluence{} with
 the added restriction that subexpressions are join-free.
 
-We say $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
-join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1 \join
-e_2)$.  Intuitively, \dIreduction{} allows us to eliminate join operators by
+\begin{definition}
+  $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
+  join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some
+  join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1
+  \join e_2)$.
+\end{definition}
+
+Intuitively, \dIreduction{} allows us to eliminate join operators by
 substituting $e_1 \join e_2$ (which contains 1 join) with $e_3$ (which does not
@@
 \begin{proof}
-  First, we prove that for all join-free $e$, if $\Irec{e}$ then there exists a
-  join-free $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a
-  pretty straightforward structural induction on $e$.
+  First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
+  $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty
+  straightforward structural induction on $e$.
   \begin{itemize}
@@
 \end{proof}
+
+\begin{claim}
+  Consider a state-based object $(S, s_0, \join)$, a set of transactions $T$,
+  and an invariant $I$. $T$ is \Iconfluent{} if the following criteria are met:
+  \begin{enumerate}
+    \item
+      $(S, \join)$ is a semilattice.
+    \item
+      For every $t \in T$, there exists some $t': S \to S$ such that for all
+      $s \in S$, $t(s) = s \join t'(s)$.
+
+    \item
+      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
+      S$, $t_1(t_2(s)) = t_2(t_1(s))$.
+
+    \item
+      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
+      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
+      t_2(s))$.
+  \end{enumerate}
+\end{claim}
+\begin{proof}
+  foo
+\end{proof}
+
+\todo{State conditions}
+\todo{Prove sufficient}
+\todo{Show more general than naive conditions}

### 2018-02-20 doc/paper_redux/sufficient_conditions.tex
WIP

--- 2018-02-18 doc/paper_redux/sufficient_conditions.tex
@@
 The conditions in \clmref{statebasedsufficient} and \clmref{opbasedsufficient}
-are not necessary. Intuitively, this makes quite a bit of sense. Neither
-condition depended on the initial state $s_0$, and both
-\clmref{statebasednotnecessary} and \clmref{opbasednotnecessary} exploit this.
-If we relax our assumption of a fixed start state, then both these conditions
-become sufficient and necessary.
+are not necessary. Intuitively, this makes quite a bit of sense. The condition
+in \clmref{statebasedsufficient} doesn't depend on $T$ and neither condition
+depends on the initial state $s_0$. However, if we relax our assumption of a
+fixed start state $s_0$ and instead allow replicas to begin with an arbitrary
+state, then both conditions are sufficient and necessary. Alternatively, if
+every state in $S$ is reachable from $s_0$, then the conditions are sufficient
+and necessary.
 
-\todo{Prove this more carefully. It's pretty straightforward though.}
+\todo{Prove these facts more carefully. It's pretty straightforward though.}
+
+% | Data Model                              | I           | SC  | T | IC  |
+% | State based set with union              | phi(x)      | yes |   | yes |
+% | State based two sets with union         | foreign key | yes |   | yes |
+% | State-based increment-decrement counter | >           |

### 2018-02-20 doc/paper_redux/sufficient_conditions.tex
Added a slightly less silly example.

@@
 
+\todo{Write this up in full.}
+Imagine we are building a system to group students together. We have a relation
+of (student id, grade) and a map (group id -> set of student ids). We want to
+enforce the invariant that every student in every group is in the same grade.
+If we start off with an initially empty mapping, then there are conflicts. If
+we start off with at least one student mapped into each group, then there are
+not any conflicts.
+
 % | Data Model                              | I           | SC  | T | IC  |

### 2018-02-20 doc/paper_redux/new_statebased_sufficient.tex
Finished state based sufficient proof.

@@
   and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
-  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join e_2')$, and
+  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join
+  e_2')$, and
   \[
@@
       $(S, \join)$ is a semilattice.
-    \item
-      For every $t \in T$, there exists some $t': S \to S$ such that for all
-      $s \in S$, $t(s) = s \join t'(s)$.
 
     \item
-      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
-      S$, $t_1(t_2(s)) = t_2(t_1(s))$.
+      For every $t \in T$, there exists some $s_t \in S$ such that $t(s) = s
+      \join s_t$.
 
@@
 \begin{proof}
-  foo
+  \newcommand{\bart}[1]{\overline{t_{#1}}}
+  \newcommand{\baru}[1]{\overline{u_{#1}}}
+
+  To show that $T$ is \Iconfluent{}, it suffices to show that $T$ is
+  \dIconfluent{} and \dIreducible{} (by
+  \clmref{diconfluence_direducible_implies_iconfluence}). We begin by showing
+  that $T$ is \dIconfluent{}. Let
+  \[
+    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
+    \quad \text{and} \quad
+    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
+  \]
+  be two arbitrary join-free expressions that recursively satisfy $I$. For ease
+  of notation, let
+  \[
+    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
+    \quad\text{and}\quad
+    \baru{i} = u_j(\ldots(u_1(s_0))\ldots)
+  \]
+  We prove by strong induction on $m \in \nats$ that if $m = i + j$ where $0
+  \leq i \leq n$ and $0 \leq j \leq m$, then $\Irec{\bart{i}(\baru{j}(s_0))}$.
+  From this, it's immediate that $\Irec{\bart{n}(\baru{m}(s_0))}$, so
+  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dIconfluent{}.
+  \begin{itemize}
+    \item \textbf{Case $m = 0$.}
+      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ because
+      $\Irec{e_1}$.
+
+    \item \textbf{Case $m = 1$.}
+      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
+      \[
+        \bart{1}(\baru{0}(s_0))
+          = t_1(s_0)
+      \]
+      $\Irec{t_1(s_0)}$ because $\Irec{e_1}$.
+
+    \item \textbf{Case $m \geq 2$.}
+      If $i = 0$, then $j = m$ and $\bart{0}(s_0) \join \baru{j}(s_0) =
+      \baru{j}(s_0)$ which recursively satisfies $I$ because $\Irec{e_2}$.
+      Similarly if $j = 0$, then $i = m$ and $\bart{i}(s_0) \join \baru{0}(s_0)
+      = \bart{i}(s_0)$ which recursively satisfies $I$ because $\Irec{e_1}$.
+      %
+      Otherwise, $i, j > 1$. Let
+      \begin{align*}
+        e_{i-1,j-1}
+          &= \bart{i-1}(\baru{j-1}(s_0)) \\
+        e_{i,j-1}
+          &= \bart{i}(\baru{j-1}(s_0))
+           = t_i(e_{i-1,j-1}) \\
+        e_{i-1,j}
+          &= \bart{i-1}(\baru{j}(s_0))
+           = u_j(e_{i-1,j-1})
+      \end{align*}
+      By the inductive hypothesis, $\Irec{e_{i-1,j-1}}$, $\Irec{e_{i,j-1}}$,
+      and $\Irec{e_{i-1,j}}$. By condition 3, $e_{i-1,j} \join e_{i,j-1} =
+      \bart{i}(s_0) \join \baru{j}(s_0) = \bart{i}(\baru{j}(s_0))$ satisfies
+      $I$. This implies that $\Irec{\bart{i}(\baru{j}(s_0))}$.
+  \end{itemize}
+
+  Next, we show that $T$ is \dIreducible{}. Again let
+  \[
+    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
+    \quad \text{and} \quad
+    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
+  \]
+  be two arbitrary join-free expressions that recursively satisfy $I$. We just
+  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
+  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
+  expression, so $T$ is \dIreducible{}.
 \end{proof}
-
-\todo{State conditions}
-\todo{Prove sufficient}
-\todo{Show more general than naive conditions}

### 2018-02-22 doc/paper_redux/sufficient_conditions.tex
Finished state based sufficient proof.
Added some examples.

@@
 of (student id, grade) and a map (group id -> set of student ids). We want to
-enforce the invariant that every student in every group is in the same grade.
-If we start off with an initially empty mapping, then there are conflicts. If
-we start off with at least one student mapped into each group, then there are
-not any conflicts.
+enforce the invariant that every student in every group is in the same grade
+and that all groups are non-empty. If we start off with an initially empty
+mapping, then there are conflicts. If we start off with at least one student
+mapped into each group, then there are not any conflicts.
 
@@
 % | State-based increment-decrement counter | >           |
+
+Here are some examples of transactions $T$ and invariants $I$ that do not
+satisfy \clmlabel{statebasedsufficient} but are \Iconfluent{}.
+
+\begin{example}
+  Consider the state-based object $(S, s_0, \join)$ where $(S, \join) = (\ints,
+  \max) \times (\ints, \max)$ and $s_0 = (0, 0)$. Let $T = \set{(x, y) \mapsto
+  (x + 1, y), (x, y) \mapsto (x, y - 1)}$ and let $I((x, y)) = xy \leq 0$. The
+  reachable states are $\nats \times -\nats$, all of which satisfy $I$, so $T$
+  is \Iconfluent{}. However, let $a = (-1, 1)$ and $b = (1, -1)$. $I(a)$ and
+  $I(b)$, but $a \join b = (1, 1)$ does not satisfy $I$.
+\end{example}
+
+\begin{example}
+  Consider the state-based object $(\powerset{\nats}, \set{0}, \cup)$. Let $T =
+  \setst{X \mapsto X \cup Y}{Y \subseteq \nats}$ let $I(X) = \exists k \in
+  \set{0, 1}.\ \forall x \in X.\ x \equiv k \mod 2$. That is $I(X)$ if $X$
+  contains only odd or only even elements. The reachable states are
+  $\powerset{2\nats}$, all of which satisfy $I$, so $T$ is \Iconfluent{}.
+  However, let $A = \set{0}$ and $Y = \set{1}$. $I(A)$ and $I(B)$, but $A \join
+  B = \set{0, 1}$ does not satisfy $I$.
+\end{example}
+
+\begin{example}
+  Consider a state-based object consisting of two integer 2P-sets $(A_1, R_1)$ and
+  $(A_2, R_2)$ with initial state $(\set{1, 3}, \emptyset{})$ and $(\set{1, 2,
+  3, 4}, \emptyset{})$. Let $T$ consist of all transactions that either remove
+  an element from the first 2P-set (i.e.\ add something to $R_1$) or add
+  something to the second 2P-set (i.e.\ add something to $A_2$). Let $I$ be
+  that the first 2P-set is a subset of the second (i.e. $(A_1 - R_1) \subseteq
+  (A_2 - R_2)$).
+  %
+  $T$ is \Iconfluent{}. However, let $s_1 = (\set{1}, \set{2}, \set{1},
+  \set{2})$ and $s_2 = (\set{2}, \set{1}, \set{2}, \set{1})$. $I(s_1)$ and
+  $I(s_2)$, but $s_1 \cup s_2 = (\set{1, 2}, \set{1, 2}, \set{1, 2}, \set{1,
+  2})$ which does not satisfy $I$.
+\end{example}

### 2018-02-23 doc/paper_redux/iconfluence.tex
Added some examples.
Started iconfluence definition tweaks.

@@
 The notion of state-based objects is taken from~\cite{shapiro2011conflict}.
+Note that $\join$ is an \emph{arbitrary} function. It does not necessarily have
+to satisfy any special properties like associativity or commutativity, though
+later we will see that often it does.
 
@@
 \begin{definition}
-  An \defword{invariant} $I: S \to \set{\text{true}, \text{false}}$ is a
-  predicate on states.
+  An \defword{invariant} $I \subseteq S$ is a subset of states. Notationally,
+  we say $I(s)$ to mean $s \in I$ and $\lnot I(s)$ to mean $s \notin I$.
 \end{definition}
@@
   $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
-  $I(x) = x \geq 0$ is an invariant.
+  $\setst{x \in \nats}{x \equiv 0 \mod 2}$ is an invariant. Here, $I(0)$ and
+  $I(2)$ but $\lnot I(1)$ and $\lnot I(3)$.
 \end{example}
@@
 Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
-$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.
-Processors can execute a transaction $t \in T$ to transition from state $s$ to
-state $t(s)$, or they can send their state $s$ to another processor with state
-$s'$ causing the other processor to transition from state $s'$ to state $s
-\join s'$.
+$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.  A
+processor $p_i$ can perform one of two actions.
 
-Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
-and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
-\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
-abbreviated \defword{\Iconfluent{}}, if $s_i^n \join s_j^m$ is guaranteed to
-satisfy $I$.
+First, $p_i$ can execute a transaction $t \in T$ to transition from state $s$
+to state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
+execute $t$ (or it will execute $t$ but then abort it; think of it however you
+like).
+
+Second, $p_i$ can send its state $s_i$ to another processor $p_j$ with state
+$s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
+$s_j \join s_i$; it doesn't really matter). Note that unlike executing a
+transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if $\lnot
+I(s_i \join s_j)$.
+
+$T$ is invariant-confluent with respect to $I$, abbreviated \Iconfluent{}, if
+every reachable state (including $s_0$) satisfies the invariant. Note that this
+definition of \Iconfluence{} is different from the original definition
+in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.
 
@@
 The graph-based approach is most similar to the model used by Bailis et al.\
-in~\cite{bailis2014coordination}. We are given a directed acyclic graph in
-which vertices can have an arbitrary number of children but at most two
-parents. The vertices are labelled with states, and the edges are either
-labelled with transactions or are unlabelled if they correspond to a join. The
-graph is really just an alternate way of representing an execution in the
-process-based model, but with duplicate states collapsed into a single vertex.
-As with the process-based model, we say that $T$ is \defword{\Iconfluent{}} if
-the join of any two invariant satisfying states with invariant satisfying
-ancestors is guaranteed to satisfy the invariant.
+in~\cite{bailis2014coordination}. We are given a directed acyclic graph where
+vertices are states, and edges are either labelled with transactions or are
+unlabelled if they correspond to a join. The graph is really just an alternate
+way of representing an execution in the process-based model, but with duplicate
+states collapsed into a single vertex. As with the process-based model, we say
+that $T$ is \Iconfluent{} if every vertex in the graph satisfies the invariant.
 
 \subsubsection{Expression-Based}
-The expression-based approach deals with expressions built from $s_0$,
-transactions in $T$, and $\join$. That is, expressions $e$ are built from the
-grammar
+The expression-based approach formalizes the process-based and graph-based
+approach using expressions. Given a state based object $(S, s_0, \join)$ and a
+set of transactions $T$, we consider expressions generated by the following
+grammar:
 \[
@@
 \]
-where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
-denoted $eval(e)$, in the obvious way:
+where $t$ corresponds to a transaction in $T$. We can evaluate an expression
+$e$, written $eval(e)$, in the obvious way:
 \begin{mathpar}
@@
 \end{definition}
+
+\begin{definition}
+  Intuitively, we say an expression is \defword{reachable} if it can be reached
+  in an execution of the process-based model. Formally, we defined a predicate
+  $\reachable{\cdot}$ as follows:
+
+  \begin{mathpar}
+    \inferrule{ }{\reachable{s_0}}
+
+    \inferrule{\reachable{e} \\ I(t(e))}{\reachable{t(e})}
+
+    \inferrule{\reachable{e_1} \\ \reachable{e_2}}{\reachable{e_1 \join e_2}}
+  \end{mathpar}
+\end{definition}
+
+\begin{definition}
+  $T$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq I$. In
+  other words, $T$ is \Iconfluent{} if all reachable states satisfy the
+  invariant.
+\end{definition}
+
+As we mentioned earlier, this definition of \Iconfluence{} is different than
+the original definition presented in~\cite{bailis2014coordination} (though it
+is the same as the definition in~\cite{gotsman2016cause}), but they are
+(almost) equivalent. We introduce some definitions and then prove this.
 
@@
 
-\begin{definition}
-  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
-  $\Irec{e_1}$ and $\Irec{e_2}$ implies $I(e_1 \join e_2)$.
-\end{definition}
+Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
+satisfying $I$ are closed under join. That is, $T$ is \Iconfluent{} if
+\[
+  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}
+\]
+If $\lnot I(s_0)$, then \Iconfluence holds vacuously which is a bit silly, so
+Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
+their definition of \Iconfluence{}. Doing so, the two definitions become
+equivalent.
+
+\begin{claim}\clmlabel{irec_implies_reachable}
+  $\Irec{e} \implies \reachable{e}$
+\end{claim}
+\begin{elidableproof}
+  We perform a structural induction on $e$.
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      Axiomatically, $\reachable{s_0}$.
+
+    \item \textbf{Case $t(e)$.}
+      $\Irec{t(e)}$, so $I(t(e))$ and $\Irec{e}$. By the inductive hypothesis,
+      $\reachable{e}$, so by the definition of $\reachable{\cdot}$,
+      $\reachable{t(e)}$.
+
+    \item \textbf{Case $e_1 \join e_2$.}
+      $\Irec{t(e)}$, so $\Irec{e_1}$ and $\Irec{e_2}$. By the inductive
+      hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$, so by the definition
+      of $\reachable{\cdot}$, $\reachable{e_1 \join e_2}$.
+  \end{itemize}
+\end{elidableproof}
+
+\begin{claim}
+  Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
+  and an invariant $I$. The following two are equivalent:
+  \begin{enumerate}[\quad(1)]
+    \item
+      $\setst{e}{\reachable{e}} \subseteq I$
+
+    \item
+      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}$
+  \end{enumerate}
+\end{claim}
+\begin{elidableproof}
+  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
+  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
+  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{irec_implies_reachable},
+  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
+  so by (1), $I(e_1 \join e_2)$.
+
+  Next, we show that (2) implies (1). We prove by structural induction that for
+  all $e$, $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      $I(s_0)$ by (2), so $\Irec{s_0}$
+
+    \item \textbf{Case $t(e)$.}
+      $\reachable{t(e)}$, so $\reachable{e}$ and $I(t(e))$. By the inductive
+      hypothesis, $\Irec{e}$, so by the definition of $\Irec{\cdot}$,
+      $\Irec{t(e)}$.
+
+    \item \textbf{Case $e_1 \join e_2$.}
+      $\reachable{e_1 \join e_2}$, so $\reachable{e_1}$ and $\reachable{e_2}$.
+      By the inductive hypothesis, $\Irec{e_1}$ and $\Irec{e_2}$. By $(2)$,
+      $I(e_1 \join e_2)$. Thus, $\Irec{e_1 \join e_2}$.
+  \end{itemize}
+\end{elidableproof}
 
@@
 \end{definition}
+
+\todo{Define ``reachable'' states and describe that \Iconfluence{} is
+equivalent to saying that reachable states are a subset of invariant satisfying
+states.}
+\todo{Maybe define Iconfluence to say that the initial state satisfies the
+invariant?}

### 2018-02-23 doc/paper_redux/iconfluence.tex
Finsihed new iconfluence definitions.

@@
 
-\begin{claim}\clmlabel{irec_implies_reachable}
+\begin{claim}\clmlabel{state_irec_implies_reachable}
   $\Irec{e} \implies \reachable{e}$
@@
 \begin{elidableproof}
-  We perform a structural induction on $e$.
+  We perform structural induction on $e$.
   \begin{itemize}
@@
   (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
-  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{irec_implies_reachable},
+  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{state_irec_implies_reachable},
   $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
@@
 \begin{definition}
-  A \defword{distributed operation-based object} is a pair $(S, s_0)$
-  where $S$ is a set of states and $s_0 \in S$ is a designated start state.
+  A \defword{distributed operation-based object} is a pair $(S, s_0)$ where $S$
+  is a set of states and $s_0 \in S$ is a designated start state.
 \end{definition}
 
-The notion of operation-based objects is also taken
+Like state-based objects, the notion of operation-based objects is taken
 from~\cite{shapiro2011conflict}. Note that we do not have a join function like
@@
 
-The definition of invariants remains unchanged.
+The definition of an invariant is the same in the state-based and
+operation-based model.
 
 \begin{example}
-  $(\nats, 0)$ is a distributed object. Let $f_k(x) = x + k$. $t(x) = f_x$ is a
-  transaction. $I(x) = x \geq 0$ is an invariant.
+  $(\nats, 0)$ is a distributed object. $t(x) = y \mapsto x + y$ is a
+  transaction that given a state $x$ returns a function $y \mapsto x + y$ that
+  adds $x$ to its argument. $\setst{x \geq 0}{x \in \nats}$ is an invariant.
 \end{example}
@@
 We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
-state $s_0$. Repeatedly, a processor executes a transaction $t$ on its current
-state $s$ and then transitions from state $s$ to state $t(s)(s)$. It also
-broadcasts $t(s)$ to all other processes exactly once. When a processor with
-state $s'$ receives $t(s)$, it transitions from state $s'$ to state $t(s)(s')$.
+state $s_0$. A processor $p_i$ can do one of two things.
 
-Consider any two processors $p_i$ and $p_j$ with states $s_i^0, \ldots, s_i^n$
-and $s_j^0, \ldots, s_j^m$ such that every state $s_i^0, \ldots, s_i^n, s_j^0,
-\ldots, s_j^m$ satisfies $I$. $T$ is invariant-confluent with respect to $I$,
-abbreviated \defword{\Iconfluent{}}, if for all $t \in T$, $I(t(s_i^n)(s_i^n))$
-implies $I(t(s_i^n)(s_j^m))$.
+First, $p_i$ can execute a transaction $t$ on its current state $s$ and then
+transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
+$\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute $t$,
+then it broadcasts $t(s)$ to all other processors exactly once.
+
+Second, $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some
+other processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
+state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow transaction,
+it must execute it, even if $\lnot I(t(s_j)(s_i))$.
+
+$T$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies the
+invariant.
 
@@
 In the graph-based model, we are given a directed acyclic graph in which each
-vertex is associated with a state $s$ and each edge is labelled with a shadow
-operation $t(s)$ where $s$ is associated with some vertex in the graph.
-%
-Consider two invariant-satisfying states $s$ and $s'$ with invariant satisfying
-ancestors. $T$ is \defword{\Iconfluent{}} if for all $t$, $I(t(s)(s))$ implies
-$I(t(s)(s'))$.
+vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
+where $s$ is some other vertex in the graph. $T$ is \Iconfluent{} if every
+vertex in the graph satisfies $I$.
 
 \subsubsection{Expression-Based}
-The expression-based approach deals with expressions built from $s_0$ and
-transactions in $T$. That is, expressions $e$ are built from the grammar
+Given an operation-based object $(S, s_0)$ and a set of transactions $T$, we
+consider expressions built from the following grammar:
 \[
@@
 \begin{definition}
+  As with state-based expressions, we formalize which operation-based
+  expressions are reachable.
+
+  \begin{mathpar}
+    \inferrule{ }{\reachable{s_0}}
+
+    \inferrule{
+      \reachable{e_1} \\
+      \reachable{e_2} \\
+      I(t(e_1)(e_1))
+    }{\reachable{t(e_1)(e_2)}}
+  \end{mathpar}
+\end{definition}
+
+\begin{definition}
+  $T$ is \Iconfluent{} if $\setst{e}{\reachable{e}} \subseteq I$.
+\end{definition}
+
+As with state-based objects, we have an equivalent definition of \Iconfluence{}
+that deals with the closure of recursively invariant satisfying states.
+
+\begin{definition}
   An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
@@
 
-\begin{definition}
-  $T$ is \defword{\Iconfluent}, if for all expressions $e_1$ and $e_2$,
-  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$ implies $I(t(e_1)(e_2))$.
-\end{definition}
+\begin{claim}\clmlabel{op_irec_implies_reachable}
+  $\Irec{e} \implies \reachable{e}$
+\end{claim}
+\begin{elidableproof}
+  We perform a structural induction on $e$.
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      Axiomatically, $\reachable{s_0}$.
 
-\todo{Define ``reachable'' states and describe that \Iconfluence{} is
-equivalent to saying that reachable states are a subset of invariant satisfying
-states.}
-\todo{Maybe define Iconfluence to say that the initial state satisfies the
-invariant?}
+    \item \textbf{Case $t(e_1)(e_2)$.}
+      $\Irec{t(e_1)(e_2)}$, so $\Irec{e_1}$, $\Irec{e_2}$, and $t(e_1)(e_1)$.
+      By the inductive hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$.
+      Thus, $\reachable{t(e_1)(e_2)}$.
+  \end{itemize}
+\end{elidableproof}
+
+\begin{claim}
+  Given an operation-based object $(S, s_0)$, a set of transactions $T$, and an
+  invariant $I$, the following two are equivalent:
+  \begin{enumerate}[\quad(1)]
+    \item
+      $\setst{e}{\reachable{e}} \subseteq I$
+
+    \item
+      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
+      I(t(e_1)(e_2))$.
+  \end{enumerate}
+\end{claim}
+\begin{elidableproof}
+  First, we show that (1) implies (2). $\reachable{s_0}$, so by (1), $I(s_0)$.
+  Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
+  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{op_irec_implies_reachable},
+  $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By
+  (1), $I(t(e_1)(e_2))$.
+
+  Next, we show that (2) implies (1). We show by structural induction that
+  $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      By (2), $I(s_0)$, so $\Irec{s_0}$.
+
+    \item \textbf{Case $t(e_1)(e_2)$.}
+      $\reachable{t(e_1)(e_2)}$, so $\reachable{e_1}$, $\reachable{e_2}$, and
+      $I(t(e_1)(e_1))$. By the inductive hypothesis, $\Irec{e_1}$ and
+      $\Irec{e_2}$. By (2), $I(t(e_1)(e_2))$, so $\Irec{t(e_1)(e_2)}$.
+  \end{itemize}
+\end{elidableproof}

### 2018-02-23 doc/paper_redux/iconfluence.tex
Minor tweaks.

@@
 states collapsed into a single vertex. As with the process-based model, we say
-that $T$ is \Iconfluent{} if every vertex in the graph satisfies the invariant.
+that $T$ is \Iconfluent{} if every vertex in every graph satisfies the
+invariant.
 
@@
 \[
-  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}
+  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
 \]
-If $\lnot I(s_0)$, then \Iconfluence holds vacuously which is a bit silly, so
+If $\lnot I(s_0)$, then \Iconfluence{} holds vacuously which is a bit silly, so
 Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
@@
     \item
-      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I{e_1 \join e_2}$
+      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
   \end{enumerate}
@@
 where $s$ is some other vertex in the graph. $T$ is \Iconfluent{} if every
-vertex in the graph satisfies $I$.
+vertex in every graph satisfies $I$.
 
@@
     \item \textbf{Case $t(e_1)(e_2)$.}
-      $\Irec{t(e_1)(e_2)}$, so $\Irec{e_1}$, $\Irec{e_2}$, and $t(e_1)(e_1)$.
-      By the inductive hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$.
-      Thus, $\reachable{t(e_1)(e_2)}$.
+      $\Irec{t(e_1)(e_2)}$, so $\Irec{e_1}$, $\Irec{e_2}$, and
+      $I(t(e_1)(e_1))$. By the inductive hypothesis, $\reachable{e_1}$ and
+      $\reachable{e_2}$. Thus, $\reachable{t(e_1)(e_2)}$.
   \end{itemize}
@@
 
-  Next, we show that (2) implies (1). We show by structural induction that
-  $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
+  Next, we show that (2) implies (1). We show by structural induction on $e$
+  that $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
   \begin{itemize}

### 2018-02-26 doc/paper_redux/iconfluence.tex
Updated sufficient conditions.

@@
 
-\begin{claim}
+\begin{claim}\clmlabel{statebased_two_iconfluence_defs}
   Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
@@
 
-\begin{claim}
+\begin{claim}\clmlabel{opbased_two_iconfluence_defs}
   Given an operation-based object $(S, s_0)$, a set of transactions $T$, and an

### 2018-02-26 doc/paper_redux/sufficient_conditions.tex
Updated sufficient conditions.

@@
 \section{Sufficient Conditions}
-In this section, we present sufficient conditions for state-based and
-operation-based invariant confluence. It turns out that these conditions are
-sufficient but not necessary. However, if we relax the assumption of a fixed
-start state $s_0$, then they are both sufficient and necessary.
+In this section, we present sufficient (but not always necessary) conditions
+for state-based and operation-based invariant confluence. We also explain
+circumstances under which the sufficient conditions are also necessary.
+
+\begin{definition}
+  We say a state-based object is \defword{\Iclosed{}} if $I(s_0)$ and for all
+  states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
+  \[
+    I(s_1) \land I(s_2) \implies I(s_1 \join s_2)
+  \]
+  Similarly, we say on operation-based object is \defword{\Iclosed{}} if
+  $I(s_0)$ and for all states $s_1$ and $s_2$ and all transactions $t$, if
+  $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
+  \[
+    I(s_1) \land I(s_2) \land I(t(s_1)(s_1)) \implies I(t(s_1)(s_2))
+  \]
+
+  \todo{%
+    We probably don't want to include $I(s_0)$ in the definition of \Iclosed{}
+    because it's a bit confusing.
+  }
+\end{definition}
 
 \begin{claim}\clmlabel{statebasedsufficient}
-  If for all states $s_1$ and $s_2$, $I(s_1)$ and $I(s_2)$ implies $I(s_1 \join
-  s_2)$, then $T$ is \Iconfluent{}. In other words, if invariant satisfying
-  states are closed under join, then $T$ is \Iconfluent.
+  Given a state-based object $O = (S, s_0, \join)$, a set of transactions $T$,
+  and an invariant $I$, if $O$ is \Iclosed{}, then $T$ is \Iconfluent.
 \end{claim}
-\begin{proof}
-  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Then,
-  $I(eval(e_1))$ and $I(eval(e_2))$, so $I(eval(e_1) \join eval(e_2))$ which
-  implies $I(e_1 \join e_2)$.
-\end{proof}
+\begin{elidableproof}
+  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Let
+  $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$. Then, $I(s_1)$ and $I(s_2)$, so
+  $I(s_1 \join s_2)$ which implies $I(e_1 \join e_2)$. Thus, by
+  \clmref{statebased_two_iconfluence_defs}, $T$ is \Iconfluent{}.
+\end{elidableproof}
 
 \begin{claim}\clmlabel{opbasedsufficient}
-  If for all transactions $t \in T$ and for all states $s_1$ and $s_2$,
-  $I(s_1)$ and $I(s_2)$ and $I(t(s_1)(s_1))$ implies $I(t(s_1)(s_2))$, then $T$
-  is \Iconfluent{}.
+  Given a operation-based object $O = (S, s_0)$, a set of transactions $T$, and
+  an invariant $I$, if $O$ is \Iclosed{}, then $T$ is \Iconfluent.
 \end{claim}
-\begin{proof}
+\begin{elidableproof}
   Let $t \in T$ and $e_1, e_2$ be expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_2))$. Let $s_1 = eval(e_1)$ and $s_2 =
+  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. Let $s_1 = eval(e_1)$ and $s_2 =
   eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, so
-  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$.
-\end{proof}
+  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$, so by
+  \clmref{opbased_two_iconfluence_defs}, $T$ is \Iconfluent.
+\end{elidableproof}
 
-While the conditions in \clmref{statebasedsufficient} and
-\clmref{opbasedsufficient} are sufficient, they are not necessary.
+Unfortunately, the converses of \clmref{statebasedsufficient} and
+\clmref{opbasedsufficient} are not true. Being \Iconfluent{} does not
+necessarily imply being \Iclosed{}, as the next two examples show.
 
-\begin{claim}\clmlabel{statebasednotnecessary}
-  If $T$ is \Iconfluent, it may not satisfy the condition in
-  \clmref{statebasedsufficient}.
+\begin{example}\examplelabel{statebasednotnecessary}
+  Consider the state-based object $O = (S, s_0, \join)$ where $(S, \join)$ is
+  $(\ints, \max) \times (\ints, \max)$ and $s_0 = (0, 0)$. Let $T$ consist of
+  two transactions $t_{x+1} \defeq \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} =
+  \lambda (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
+  %
+  $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
+  $T$ is \Iconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
+  $I(s_1)$ and $I(s_2)$. However, $s_1 \join s_2 = (1, 1)$, and $\lnot I((1,
+  1))$. Thus, $O$ is not \Iclosed{}.
+\end{example}
+
+\begin{example}
+  Consider the operation-based object $O = (S, s_0)$ where $S$ is $\ints \times
+  \ints$ and $s_0 = (0, 0)$. Let $T$ consist of two transactions $t_{x+1}
+  \defeq \lambda s.\ \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} = \lambda s.\
+  \lambda (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
+  %
+  $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
+  $T$ is \Iconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
+  $I(s_1)$, $I(s_2)$, and $I(t_{x+1}(s_1)(s_1))$. However, $t_{x+1}(s_1)(s_2) =
+  (1, 1)$, and $\lnot I((1, 1))$. Thus, $O$ is not \Iclosed{}.
+\end{example}
+
+\input{not_necessary.fig}
+
+Let's take a closer look at \exampleref{statebasednotnecessary} (illustrated in
+\figref{statebasednotnecessary}) to understand why being \Iconfluent{} doesn't
+imply being \Iclosed{}. Our counterexample considers points $s_1 = (-1, 1)$ and
+$s_2 = (1, -1)$. $s_2$ is reachable, but $s_1$ is not. This is not a
+coincidence! If we consider an \Iconfluent{} set of transactions, then any two
+reachable points satisfy the invariant, and the join of the two points is again
+reachable, so it must satisfy the invariant as well. The only way to join two
+invariant satisfying points and end up not satisfying the invariant is if one
+or both points are not reachable. Thus, if we know that $I$ is a subset of
+reachable points, then being \Iconfluent{} and being \Iclosed{} are equivalent.
+
+\begin{claim}\clmlabel{sufficientandnecessary}
+  If every invariant satisfying state is reachable, then being \Iconfluent{}
+  and being \Iclosed{} are equivalent. Being \Iclosed{} is always sufficient,
+  but if we can establish that every invariant satisfying point is reachable,
+  then being \Iclosed{} is a both necessary and sufficient for being
+  \Iconfluent{}.
+  \[
+    I \subseteq \setst{e}{\reachable{e}} \implies
+    (\text{\Iclosed{}} \iff \text{\Iconfluent})
+  \]
 \end{claim}
-\begin{proof}
-  Consider the state based object $(S, s_0, \join)$ where $S$ is the set of all
-  finite non-empty sets of integers, $s_0 = \set{0}$, and $\join = \cup$. We
-  have a single transaction $t(x) = x \cup \set{\max(x) + 2}$ which retrieves
-  the largest element from the set (which is guaranteed to exist because it is
-  finite and non-empty), adds two to it, and then adds it into the set. $T =
-  \set{t}$. Let $I$ be the invariant that all elements of $S$ are either odd or
-  even.
+\begin{elidableproof}
+  Consider a state-based object $O = (S, s_0, \join)$, a set of transactions
+  $T$, and an invariant $I$ where $I \subseteq \setst{e}{\reachable{e}}$.
+  %
+  Being \Iclosed{} always implies being \Iconfluent{} whether or not $I
+  \subseteq \setst{e}{\reachable{e}}$. Thus, we only have to prove that if $T$
+  is \Iconfluent{}, then $O$ is \Iclosed. If $T$ is \Iconfluent{}, then
+  $I(s_0)$. Next, let $s_1$ and $s_2$ be arbitrary states such that $I(s_1)$
+  and $I(s_2)$. Both states satisfy the invariant, so both are reachable. That
+  is, there exist expressions $e_1$ and $e_2$ where $eval(e_1) = s_1$  and
+  $eval(e_2) = s_2$. $e_1 \join e_2$ is again reachable, so since $T$ is
+  \Iconfluent{}, $I(e_1 \join e_2)$ which implies $I(s_1 \join s_2)$. Thus, $O$
+  is \Iclosed.
+\end{elidableproof}
 
-  Note that $T$ is \Iconfluent{}. Beginning with our initial state, we can
-  derive states $\set{0}, \set{0, 2}, \set{0, 2, 4}, \ldots$ all of which
-  satisfy the invariant. Moreover, merging any such states satisfies the
-  invariant.
-  %
-  However, $T$ doesn't satisfy the condition in \clmref{statebasedsufficient}.
-  $\set{0}$ and $\set{1}$ satisfy the invariant, but $\set{0} \cup \set{1}$
-  does not.
-\end{proof}
+\clmref{sufficientandnecessary} arms us with some potent understanding of
+\Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
+that our state-based object from \exampleref{statebasednotnecessary} is not
+\Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
+have to seek out alternative means to determine whether or not our transactions
+were \Iconfluent{}. With the intuition provided by
+\clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
+\setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
+decision procedure informs us that we are indeed \Iclosed{}, so we know we're
+\Iconfluent{}.
 
-\begin{claim}\clmlabel{opbasednotnecessary}
-  If $T$ is \Iconfluent, it may not satisfy the condition in
-  \clmref{opbasedsufficient}.
-\end{claim}
-\begin{proof}
-  We play the same trick as \clmref{statebasednotnecessary}. Let $S$, $s_0$,
-  and $I$ be as in \clmref{statebasednotnecessary}. Consider a single
-  transaction $t(x) = u_{\max(x) + 2}$ where $u_k(x) = x \cup \set{k}$. Let $T
-  = \set{t}$.
+Here are some more examples of transactions $T$ and invariants $I$ that are
+\Iconfluent{} but not \Iclosed{}. \todo{Think of more compelling examples.}
 
-  $T$ is \Iconfluent{}. Beginning with our initial state, we can derive any
-  finite set of even integers containing $0$. All these states satisfy the
-  invariant. Letting $s, s'$ be any of these states, $I(t(s)(s'))$.
-  %
-  However, $T$ doesn't satisfy the condition in \clmref{opbasedsufficient}.
-  $\set{0}$ and $\set{1}$ satisfy the invariant, and so does
-  $t(\set{0})(\set{0})$, but $t(\set{0})(\set{1}) = \set{1, 2}$ does not.
-\end{proof}
-
-\todo{Think of more compelling examples.}
-
-The conditions in \clmref{statebasedsufficient} and \clmref{opbasedsufficient}
-are not necessary. Intuitively, this makes quite a bit of sense. The condition
-in \clmref{statebasedsufficient} doesn't depend on $T$ and neither condition
-depends on the initial state $s_0$. However, if we relax our assumption of a
-fixed start state $s_0$ and instead allow replicas to begin with an arbitrary
-state, then both conditions are sufficient and necessary. Alternatively, if
-every state in $S$ is reachable from $s_0$, then the conditions are sufficient
-and necessary.
-
-\todo{Prove these facts more carefully. It's pretty straightforward though.}
-
+\begin{example}
   \todo{Write this up in full.}
@@
   mapped into each group, then there are not any conflicts.
-
-% | Data Model                              | I           | SC  | T | IC  |
-% | State based set with union              | phi(x)      | yes |   | yes |
-% | State based two sets with union         | foreign key | yes |   | yes |
-% | State-based increment-decrement counter | >           |
-
-Here are some examples of transactions $T$ and invariants $I$ that do not
-satisfy \clmlabel{statebasedsufficient} but are \Iconfluent{}.
+\end{example}
 

### 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
Some reorganization.

--- 2018-02-20 doc/paper_redux/sufficient_conditions.tex
@@
-\section{Sufficient Conditions}
-In this section, we present sufficient conditions for state-based and
-operation-based invariant confluence. It turns out that these conditions are
-sufficient but not necessary. However, if we relax the assumption of a fixed
-start state $s_0$, then they are both sufficient and necessary.
-
-\begin{claim}\clmlabel{statebasedsufficient}
-  If for all states $s_1$ and $s_2$, $I(s_1)$ and $I(s_2)$ implies $I(s_1 \join
-  s_2)$, then $T$ is \Iconfluent{}. In other words, if invariant satisfying
-  states are closed under join, then $T$ is \Iconfluent.
-\end{claim}
-\begin{proof}
-  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Then,
-  $I(eval(e_1))$ and $I(eval(e_2))$, so $I(eval(e_1) \join eval(e_2))$ which
-  implies $I(e_1 \join e_2)$.
-\end{proof}
-
-\begin{claim}\clmlabel{opbasedsufficient}
-  If for all transactions $t \in T$ and for all states $s_1$ and $s_2$,
-  $I(s_1)$ and $I(s_2)$ and $I(t(s_1)(s_1))$ implies $I(t(s_1)(s_2))$, then $T$
-  is \Iconfluent{}.
-\end{claim}
-\begin{proof}
-  Let $t \in T$ and $e_1, e_2$ be expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_2))$. Let $s_1 = eval(e_1)$ and $s_2 =
-  eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, so
-  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$.
-\end{proof}
-
-While the conditions in \clmref{statebasedsufficient} and
-\clmref{opbasedsufficient} are sufficient, they are not necessary.
-
-\begin{claim}\clmlabel{statebasednotnecessary}
-  If $T$ is \Iconfluent, it may not satisfy the condition in
-  \clmref{statebasedsufficient}.
-\end{claim}
-\begin{proof}
-  Consider the state based object $(S, s_0, \join)$ where $S$ is the set of all
-  finite non-empty sets of integers, $s_0 = \set{0}$, and $\join = \cup$. We
-  have a single transaction $t(x) = x \cup \set{\max(x) + 2}$ which retrieves
-  the largest element from the set (which is guaranteed to exist because it is
-  finite and non-empty), adds two to it, and then adds it into the set. $T =
-  \set{t}$. Let $I$ be the invariant that all elements of $S$ are either odd or
-  even.
-
-  Note that $T$ is \Iconfluent{}. Beginning with our initial state, we can
-  derive states $\set{0}, \set{0, 2}, \set{0, 2, 4}, \ldots$ all of which
-  satisfy the invariant. Moreover, merging any such states satisfies the
-  invariant.
-  %
-  However, $T$ doesn't satisfy the condition in \clmref{statebasedsufficient}.
-  $\set{0}$ and $\set{1}$ satisfy the invariant, but $\set{0} \cup \set{1}$
-  does not.
-\end{proof}
-
-\begin{claim}\clmlabel{opbasednotnecessary}
-  If $T$ is \Iconfluent, it may not satisfy the condition in
-  \clmref{opbasedsufficient}.
-\end{claim}
-\begin{proof}
-  We play the same trick as \clmref{statebasednotnecessary}. Let $S$, $s_0$,
-  and $I$ be as in \clmref{statebasednotnecessary}. Consider a single
-  transaction $t(x) = u_{\max(x) + 2}$ where $u_k(x) = x \cup \set{k}$. Let $T
-  = \set{t}$.
-
-  $T$ is \Iconfluent{}. Beginning with our initial state, we can derive any
-  finite set of even integers containing $0$. All these states satisfy the
-  invariant. Letting $s, s'$ be any of these states, $I(t(s)(s'))$.
-  %
-  However, $T$ doesn't satisfy the condition in \clmref{opbasedsufficient}.
-  $\set{0}$ and $\set{1}$ satisfy the invariant, and so does
-  $t(\set{0})(\set{0})$, but $t(\set{0})(\set{1}) = \set{1, 2}$ does not.
-\end{proof}
-
-\todo{Think of more compelling examples.}
-
-The conditions in \clmref{statebasedsufficient} and \clmref{opbasedsufficient}
-are not necessary. Intuitively, this makes quite a bit of sense. The condition
-in \clmref{statebasedsufficient} doesn't depend on $T$ and neither condition
-depends on the initial state $s_0$. However, if we relax our assumption of a
-fixed start state $s_0$ and instead allow replicas to begin with an arbitrary
-state, then both conditions are sufficient and necessary. Alternatively, if
-every state in $S$ is reachable from $s_0$, then the conditions are sufficient
-and necessary.
-
-\todo{Prove these facts more carefully. It's pretty straightforward though.}
-
-\todo{Write this up in full.}
-Imagine we are building a system to group students together. We have a relation
-of (student id, grade) and a map (group id -> set of student ids). We want to
-enforce the invariant that every student in every group is in the same grade.
-If we start off with an initially empty mapping, then there are conflicts. If
-we start off with at least one student mapped into each group, then there are
-not any conflicts.
-
-% | Data Model                              | I           | SC  | T | IC  |
-% | State based set with union              | phi(x)      | yes |   | yes |
-% | State based two sets with union         | foreign key | yes |   | yes |
-% | State-based increment-decrement counter | >           |
+\section{Interactive \Iconfluence{} Decision Procedure}
+\clmref{sufficientandnecessary} arms us with some potent understanding of
+\Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
+that our state-based object from \exampleref{statebasednotnecessary} is not
+\Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
+have to seek out alternative means to determine whether or not our transactions
+were \Iconfluent{}. With the intuition provided by
+\clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
+\setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
+decision procedure informs us that we are indeed \Iclosed{}, so we know we're
+\Iconfluent{}.

### 2018-02-26 doc/paper_redux/sufficient_conditions.tex
Some reorganization.

@@
 
-\clmref{sufficientandnecessary} arms us with some potent understanding of
-\Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
-that our state-based object from \exampleref{statebasednotnecessary} is not
-\Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
-have to seek out alternative means to determine whether or not our transactions
-were \Iconfluent{}. With the intuition provided by
-\clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
-\setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
-decision procedure informs us that we are indeed \Iclosed{}, so we know we're
-\Iconfluent{}.
-
 Here are some more examples of transactions $T$ and invariants $I$ that are

### 2018-02-26 doc/paper_redux/sufficient_conditions.tex
Added monotonicity proof.

@@
   \clmref{opbased_two_iconfluence_defs}, $T$ is \Iconfluent.
+\end{elidableproof}
+
+As a special case, \clmref{statebasedsufficient} tells us that a set $T$ of
+transactions is \Iconfluent{} if $I$ is monotonic and $O$ is a semilattice.
+
+\begin{claim}\clmlabel{monotonicity}
+  Let $O = (S, s_0, \join)$ be a state-based object where $(S, \join)$ is a
+  semilattice. Let $s \leq s' \defeq s \join s' = s'$. Let $T$ be an arbitrary
+  set of transactions. Let $I$ be a \defword{monotonic}: for all $s, s'\in S$,
+  if $s \leq s'$ and $I(s)$, then $I(s')$. Also assume $I(s_0)$. $T$ is
+  \Iconfluent{}.
+\end{claim}
+\begin{elidableproof}
+  Let $s_1$ and $s_2$ be arbitrary states such that $I(s_1)$ and $I(s_2)$. $s_1
+  \leq s_1 \join s_2$, so by monotonicity, $I(s_1 \join s_2)$. Thus, by
+  \clmref{statebasedsufficient}, $T$ is \Iconfluent{}.
 \end{elidableproof}

### 2018-02-26 doc/paper_redux/iconfluence.tex
Show labels.

--- 2018-02-23 doc/paper_redux/iconfluence.tex
@@
 
-\begin{claim}\clmlabel{state_irec_implies_reachable}
+\begin{claim}\clmlabel{StateBasedIrecImpliesReachable}
   $\Irec{e} \implies \reachable{e}$
@@
 
-\begin{claim}
+\begin{claim}\clmlabel{StateBasedTwoIconfluenceDefs}
   Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
@@
   (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
-  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{state_irec_implies_reachable},
+  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{StateBasedIrecImpliesReachable},
   $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
@@
 
-\begin{claim}\clmlabel{op_irec_implies_reachable}
+\begin{claim}\clmlabel{OpBasedIrecImpliesReachable}
   $\Irec{e} \implies \reachable{e}$
@@
 
-\begin{claim}
+\begin{claim}\clmlabel{OpBasedTwoIconfluenceDefs}
   Given an operation-based object $(S, s_0)$, a set of transactions $T$, and an
@@
   Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{op_irec_implies_reachable},
+  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{OpBasedIrecImpliesReachable},
   $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By

### 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
Show labels.

@@
 \section{Interactive \Iconfluence{} Decision Procedure}
-\clmref{sufficientandnecessary} arms us with some potent understanding of
-\Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
-that our state-based object from \exampleref{statebasednotnecessary} is not
-\Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
-have to seek out alternative means to determine whether or not our transactions
-were \Iconfluent{}. With the intuition provided by
-\clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
-\setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
-decision procedure informs us that we are indeed \Iclosed{}, so we know we're
-\Iconfluent{}.
+Let us revisit \exampleref{statebased_not_necessary}
+
+% \clmref{sufficientandnecessary} arms us with some potent understanding of
+% \Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
+% that our state-based object from \exampleref{statebasednotnecessary} is not
+% \Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
+% have to seek out alternative means to determine whether or not our transactions
+% were \Iconfluent{}. With the intuition provided by
+% \clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
+% \setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
+% decision procedure informs us that we are indeed \Iclosed{}, so we know we're
+% \Iconfluent{}.
+%

### 2018-02-26 doc/paper_redux/new_sufficient_conditions.tex
Show labels.

\section{Sufficient Conditions for \Iconfluence{}}
\newcommand{\Icommutativity}{$I$-commutativity}
\newcommand{\Ijoinisapply}{$I$-join-is-apply}
\newcommand{\nestedapply}[3]{{#1}_1({#1}_2(\ldots ({#1}_#2(#3)) \ldots))}
\newcommand{\bart}[1]{\overline{t_{#1}}}
\newcommand{\baru}[1]{\overline{u_{#1}}}

In this section, we prove that the following three conditions imply
\Iconfluence.
\begin{enumerate}
  \item
    $(S, \join)$ is a semilattice.

  \item
    Every transaction $t \in T$ is of the form $t(s) = s \join s_t$ for some
    $s_t \in S$.

  \item
    For all $u, t \in T$, and for all $s \in S$, if $I(s)$, $I(t(s))$, and
    $I(u(s))$, then $I(t(s) \join u(s))$.
\end{enumerate}

Using \clmref{DiconfluenceDireducibleImpliesIconfluence}, it suffices to
show that the three properties imply \dIconfluence{} and \dIreduction{}. We do
so in \clmref{ThreePropsDiconfluence} and \clmref{ThreePropsDireducible}.

\begin{claim}\clmlabel{ThreePropsDiconfluence}
  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIconfluence.
\end{claim}
\begin{proof}
  Consider arbitrary join-free expressions $e_1$ and $e_2$ where $\Irec{e_1}$
  and $\Irec{e_2}$. We want to show that $I(e_1 \join e_2)$. Because $e_1$ and
  $e_2$ are join-free, they are of the following form:
  \begin{mathpar}
    e_1 = \nestedapply{t}{n}{s_0}

    e_2 = \nestedapply{u}{m}{s_0}
  \end{mathpar}
  where $n, m \in \nats$ and $t_i, u_j \in T$. As a syntactic convenience, let
  $\bart{i}(e) = \nestedapply{t}{i}{e}$. Similarly define $\baru{j}(e)$.
  Then,
  \begin{mathpar}
    e_1 = \bart{n}(s_0)

    e_2 = \baru{m}(s_0)
  \end{mathpar}
  We prove by strong mathematical induction on m that for all i + j = m, I(barti join baruj).kjkj

\end{proof}

\begin{claim}\clmlabel{ThreePropsDireducible}
  \Icommutativity, \Ijoinisapply, and \onedIconfluence{} imply \dIreduction.
\end{claim}
\begin{proof}

\end{proof}

TODO(mwhittaker): Give examples of when these conditions are met.
TODO(mwhittaker): Explain that 1-I-confluence is necessary.

### 2018-02-26 doc/paper_redux/sufficient_conditions.tex
Show labels.

@@
 
-\begin{claim}\clmlabel{statebasedsufficient}
+\begin{claim}\clmlabel{StateBasedSufficient}
   Given a state-based object $O = (S, s_0, \join)$, a set of transactions $T$,
@@
   $I(s_1 \join s_2)$ which implies $I(e_1 \join e_2)$. Thus, by
-  \clmref{statebased_two_iconfluence_defs}, $T$ is \Iconfluent{}.
+  \clmref{StateBasedTwoIconfluenceDefs}, $T$ is \Iconfluent{}.
 \end{elidableproof}
 
-\begin{claim}\clmlabel{opbasedsufficient}
+\begin{claim}\clmlabel{OpBasedSufficient}
   Given a operation-based object $O = (S, s_0)$, a set of transactions $T$, and
@@
   $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$, so by
-  \clmref{opbased_two_iconfluence_defs}, $T$ is \Iconfluent.
+  \clmref{OpBasedTwoIconfluenceDefs}, $T$ is \Iconfluent.
 \end{elidableproof}
 
-As a special case, \clmref{statebasedsufficient} tells us that a set $T$ of
+As a special case, \clmref{StateBasedSufficient} tells us that a set $T$ of
 transactions is \Iconfluent{} if $I$ is monotonic and $O$ is a semilattice.
 
-\begin{claim}\clmlabel{monotonicity}
+\begin{claim}\clmlabel{Monotonicity}
   Let $O = (S, s_0, \join)$ be a state-based object where $(S, \join)$ is a
@@
   \leq s_1 \join s_2$, so by monotonicity, $I(s_1 \join s_2)$. Thus, by
-  \clmref{statebasedsufficient}, $T$ is \Iconfluent{}.
+  \clmref{StateBasedSufficient}, $T$ is \Iconfluent{}.
 \end{elidableproof}
 
-Unfortunately, the converses of \clmref{statebasedsufficient} and
-\clmref{opbasedsufficient} are not true. Being \Iconfluent{} does not
+Unfortunately, the converses of \clmref{StateBasedSufficient} and
+\clmref{OpBasedSufficient} are not true. Being \Iconfluent{} does not
 necessarily imply being \Iclosed{}, as the next two examples show.
 
-\begin{example}\examplelabel{statebasednotnecessary}
+\begin{example}\examplelabel{StateBasedNotNecessary}
   Consider the state-based object $O = (S, s_0, \join)$ where $(S, \join)$ is
@@
 
-Let's take a closer look at \exampleref{statebasednotnecessary} (illustrated in
-\figref{statebasednotnecessary}) to understand why being \Iconfluent{} doesn't
+Let's take a closer look at \exampleref{StateBasedNotNecessary} (illustrated in
+\figref{StateBasedNotNecessary}) to understand why being \Iconfluent{} doesn't
 imply being \Iclosed{}. Our counterexample considers points $s_1 = (-1, 1)$ and
@@
 
-\begin{claim}\clmlabel{sufficientandnecessary}
+\begin{claim}\clmlabel{SufficientAndNecessary}
   If every invariant satisfying state is reachable, then being \Iconfluent{}

### 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
Added interactive decision procedure.

--- 2018-02-26 doc/paper_redux/interactive_decision_procedure.tex
@@
 \section{Interactive \Iconfluence{} Decision Procedure}
-\clmref{sufficientandnecessary} arms us with some potent understanding of
-\Iconfluence! Imagine we have an \Iclosed{} decision procedure that informs us
-that our state-based object from \exampleref{statebasednotnecessary} is not
-\Iclosed{}. Without \clmref{sufficientandnecessary}, we'd be a bit stuck. We'd
-have to seek out alternative means to determine whether or not our transactions
-were \Iconfluent{}. With the intuition provided by
-\clmref{sufficientandnecessary}, we can instead restrict our invariant to $I =
-\setst{(x, y)}{xy \leq 0 \land x \geq 0 \land y \leq 0}$. Now, the \Iclosed{}
-decision procedure informs us that we are indeed \Iclosed{}, so we know we're
-\Iconfluent{}.
+Imagine we have an \Iclosed{} decision procedure, and let's revisit
+\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
+transactions $T$, and an invariant $I$. We want to know if $T$ is
+\Iconfluent{}. We can hand $O$, $T$, and $I$ to the \Iclosed{} decision
+procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes, then
+by \clmref{StateBasedSufficient}, we know that $T$ is \Iconfluent{}.
+
+However, in our example, $O$ is not \Iclosed{}, so the decision procedure will
+say no. \clmref{SufficientAndNecessary} tells us that this means one of two
+things. Either (1) $T$ is not \Iconfluent{} or (2) $T$ is \Iconfluent{}, but
+there are some invariant satisfying states that are not reachable. If we think
+harder about the set of reachable states, we may realize that all reachable
+states have a non-negative value of $x$. Thus, we may modify our invariant to
+$\setst{(x, y)}{xy \leq 0 \land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand
+$O$, $T$, and $I$ to the decision procedure, it will say yes, and we can
+conclude that $T$ is \Iconfluent{}.
+
+\newcommand{\comment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
+\begin{algorithm}
+  \caption{Interactive \Iconfluence{} Decision Procedure}%
+  \algolabel{InteractiveDecisionProcedure}
+  \begin{algorithmic}
+    \Function{IsIconfluent}{$O$, $T$, $I$}
+      \State \Return \Call{Helper}{$O$, $T$, $I$, $\emptyset$, $\emptyset$}
+    \EndFunction
+
+    \State
+
+    \comment{$R$ is a set of reachable states.}
+    \comment{$NR$ is a set of non-reachable states.}
+    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
+      \State closed, $s_1$, $s_2$ $\gets$
+               \Call{IsIclosed}{$O$, $T$, $(I \cap R) - NR$}
+      \If {closed $=$ unknown}
+        \State \Return unknown
+      \ElsIf {closed $=$ yes}
+        \State \Return{yes}
+      \Else
+        \State Augment $R$ and $NR$ with random search and user input
+        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
+        \If{$s_1, s_2 \in R$}
+          \State \Return no
+        \Else
+          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
+        \EndIf
+      \EndIf
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+This example motivates \algoref{InteractiveDecisionProcedure}, a general
+purpose interactive \Iconfluence{} decision procedure in which a user
+iteratively refines the invariants until the decision procedure terminates.
+Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
+\textsc{IsIconfluent}$(O, T, I)$ to determine if $T$ is \Iconfluent{}.
+\textsc{IsIconfluent} assumes access to an \Iclosed{} decision procedure
+\textsc{IsIclosed}$(O, T, I)$ which returns whether $O$ is \Iconfluent{} (yes,
+no, or unknown). If it returns no, it also returns a counterexample $s_1$ and
+$s_2$.
+
+\todo{%
+  Describe \textsc{IsIconfluent}. Describe how we can implement
+  \textsc{IsIclosed} with Z3 and how users can specify $R$ and $NR$ with
+  invariants.
+}

### 2018-02-27 doc/paper_redux/interactive_decision_procedure.tex
Finished interactive decision procedure section.

@@
     \Function{IsIconfluent}{$O$, $T$, $I$}
-      \State \Return \Call{Helper}{$O$, $T$, $I$, $\emptyset$, $\emptyset$}
+      \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
     \EndFunction
@@
       \State closed, $s_1$, $s_2$ $\gets$
-               \Call{IsIclosed}{$O$, $T$, $(I \cap R) - NR$}
+               \Call{IsIclosed}{$O$, $T$, $I - NR$}
       \If {closed $=$ unknown}
@@
 
+\newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
+\newcommand{\IsIclosed}{\textsc{IsIclosed}}
+\newcommand{\Helper}{\textsc{Helper}}
+
 This example motivates \algoref{InteractiveDecisionProcedure}, a general
 purpose interactive \Iconfluence{} decision procedure in which a user
-iteratively refines the invariants until the decision procedure terminates.
+iteratively refines the invariant until the decision procedure terminates.
 Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
-\textsc{IsIconfluent}$(O, T, I)$ to determine if $T$ is \Iconfluent{}.
-\textsc{IsIconfluent} assumes access to an \Iclosed{} decision procedure
-\textsc{IsIclosed}$(O, T, I)$ which returns whether $O$ is \Iconfluent{} (yes,
-no, or unknown). If it returns no, it also returns a counterexample $s_1$ and
-$s_2$.
+\IsIconfluent$(O, T, I)$ to determine if $T$ is \Iconfluent{}.
 
-\todo{%
-  Describe \textsc{IsIconfluent}. Describe how we can implement
-  \textsc{IsIclosed} with Z3 and how users can specify $R$ and $NR$ with
-  invariants.
-}
+\IsIconfluent{} assumes access to an \Iclosed{} decision procedure
+\IsIclosed$(O, T, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
+unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
+such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$.  Constructing an
+\Iclosed{} decision procedure is hard (undecidable, I think), but we can do
+pretty well using an SMT solver like Z3. There is also some existing research
+on how to built smart \Iclosed{} decision procedures that we can
+leverage~\cite{li2014automating}.
+
+\IsIconfluent{} takes in $O$, $T$, and $I$. Formally, $I$ is set of states
+(e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an actual
+implementation of \IsIconfluent{} would likely take in $I$ as a logical formula
+(e.g. $xy \leq 0$). From now on, whenever we mention a (possibly infinite) set
+of states, remember that we can always represent the set as a finite formula.
+
+\IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
+$T$, and $I$---takes as arguments a set $R$ of reachable states and a set $NR$
+of unreachable states. Like \IsIconfluent, \Helper$(O, T, I, R, NR)$ returns
+whether $T$ is \Iconfluent. Intuitively, $NR$ helps remove unreachable states
+from $I$ moving us closer to the case where $I \subseteq
+\setst{e}{\reachable{e}}$.
+
+\Helper{} first checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
+cases to consider.
+\begin{itemize}
+  \item
+    If \IsIclosed{} gets stuck and doesn't know whether or not $O$ is
+    \iclosed{$(I - NR)$}, then \Helper{} is stuck too and is forced to return
+    unknown.
+  \item
+    If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
+    \clmref{StateBasedSufficient}, $T$ is \iconfluent{$(I - NR)$}, so
+    $\setst{e}{\reachable{e}} \subseteq I - NR$. $I - NR \subseteq I$, so $T$
+    is therefore also \Iconfluent{}. Thus, \Helper{} return yes.
+  \item
+    If \IsIclosed{} determines that $O$ is not \iclosed{$(I - NR)$}, then we
+    have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
+    and $s_2$ are reachable or unreachable. We can do so in two ways.
+
+    First, we can randomly generate a set of reachable states and add them to
+    $R$.
+
+    Second, we can prompt the user to tell us whether or not the states are
+    reachable or unreachable. In addition to labelling $s_1$ and $s_2$ as
+    reachable or not, the user can also refine $I$ by augmenting $NR$. Taking
+    \exampleref{StateBasedNotNecessary} as an example, the user might look
+    through the existing reachable and unreachable states and have an aha
+    moment realizing that every point with a negative $x$-coordinate is
+    unreachable. They can use this knowledge to add all points with negative
+    $x$-coordinates to $NR$.
+
+    After $s_1$ and $s_2$ have been labelled reachable or not, we continue. If
+    both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
+    I(s_1 \join s_2)$. Thus, $T$ is not \Iconfluent{}, so \Helper{} returns no.
+    Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we recurse.
+\end{itemize}
+
+Let's walk through how we might use \IsIconfluent{} on
+\exampleref{StateBasedNotNecessary}.
+\begin{itemize}
+  \item
+    \IsIconfluent{} calls \Helper{} with $R = {(0, 0)}, NR = \emptyset{}$.
+
+  \item
+    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
+    counterexample $s_1 = (-1, 1)$ and $s_2 = (1, -1)$. We take the third
+    branch.
+
+  \item
+    We randomly generate some reachable points and update $R$ to $\set{(0, 0),
+    (1, 0), (0, -1), (1, -1)}$. Lucky for us, $s_2 \in R$, so we know it is
+    reachable. We're not sure about $s_1$, so we ask the user. After some
+    thought, the user tells us that $s_1$ is unreachable. We recurse with $R =
+    \set{(0, 0), (1, 0), (0, -1), (1, -1)}$ and $NR = \set{(-1, 1)}$.
+
+  \item
+    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
+    counterexample $s_1 = (-1, 2)$ and $s_2 = (10, -10)$. We take the third
+    branch.
+
+  \item
+    We randomly generate some more reachable points and update $R$ to $\set{(0,
+    0), (1, 0), (0, -1), (1, -1), (2, 0), (0, -2)}$. $s_1, s_2 \notin R, NR$,
+    so we ask the user to label them. The user puts $s_1$ in $NR$ and $s_2$ in
+    $R$. We recurse with $R = \set{(0, 0), (1, 0), (0, -1), (1, -1), (2, 0),
+    (0, -2), (10, -10)}$ and $NR = \set{(-1, 1), (-1, 2)}$.
+
+  \item
+    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
+    counterexample $s_1 = (-2, 1)$ and $s_2 = (4, -3)$. We take the third
+    branch.
+
+  \item
+    We randomly generate some more reachable points and update $R$. $s_1, s_2
+    \notin R, NR$, so we ask the user to label them. The user notices a pattern
+    in $R$ and $NR$ and after some thought, concludes that every point with
+    negative $x$-coordinate is unreachable. They update $NR$ to $-\ints \times
+    \ints$. We recurse.
+
+  \item
+    \Helper{} determines that $O$ is \iclosed{$(I - R)$} and returns yes!
+\end{itemize}

### 2018-02-27 doc/paper_redux/new_statebased_sufficient.tex
Show labels.
Finished interactive decision procedure section.

@@
 \section{New Sufficient State-Based Conditions}
+\todo{Motivate these in terms of the interactive decision procedure.}
+
 In this section, we outline a new sufficient condition for state-based
 \Iconfluence{} that is more general than the one in
-\clmref{statebasedsufficient}. To prove that the condition is sufficient, we
+\clmref{StateBasedSufficient}. To prove that the condition is sufficient, we
 introduce the concepts of \dIconfluence{} and \dIreduction{}.
@@
 
-\begin{claim}\clmlabel{diconfluence_direducible_implies_iconfluence}
+\begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
   If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
@@
   \dIconfluent{} and \dIreducible{} (by
-  \clmref{diconfluence_direducible_implies_iconfluence}). We begin by showing
+  \clmref{DiconfluenceDireducibleImpliesIconfluence}). We begin by showing
   that $T$ is \dIconfluent{}. Let

### 2018-03-03 doc/paper_redux/iconfluence.tex
Fixed a couple of dangling references.
Moved figs into a separate directory.
Fixed dangling reference.

@@
 approach, a graph-based approach, and an expression-based approach. All three
-approaches are illustrated in \figref{statebasedmodels} and are described in
+approaches are illustrated in \figref{StateBasedModels} and are described in
 more detail below.
 
-\input{statebased_iconfluence.fig}
+\input{figs/statebased_iconfluence.fig}
 
@@
 approach, and an expression-based approach. All three approaches are
-illustrated in \figref{opbasedmodels} and are described in more detail below.
+illustrated in \figref{OpBasedModels} and are described in more detail below.
 
-\input{opbased_iconfluence.fig}
+\input{figs/opbased_iconfluence.fig}
 

### 2018-03-03 doc/paper_redux/iconfluence.tex
Notational changes and writing improvements.

@@
 \begin{definition}
-  A \defword{distributed state-based object} is a triple $(S, s_0, \join)$
+  A \defword{distributed state-based object} $O$ is a triple $(S, s_0, \join)$
   where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
@@
 Note that $\join$ is an \emph{arbitrary} function. It does not necessarily have
-to satisfy any special properties like associativity or commutativity, though
-later we will see that often it does.
+to satisfy any special properties like associativity or commutativity.
 
@@
 
+We formalize an invariant as a set of states, but we can also think of a
+predicate as a formula over states. For example, we can think of the invariant
+$\setst{(x, y) \in \ints \times \ints}{x \geq y}$ as the formula $x \geq y$ and
+vice-versa.
+
 \begin{example}
   $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
-  $\setst{x \in \nats}{x \equiv 0 \mod 2}$ is an invariant. Here, $I(0)$ and
+  $\setst{x \in \nats}{\text{$x$ is even}}$ is an invariant. Here, $I(0)$ and
   $I(2)$ but $\lnot I(1)$ and $\lnot I(3)$.
@@
 There are three ways to think about state-based \Iconfluence{}: a process-based
-approach, a graph-based approach, and an expression-based approach. All three
-approaches are illustrated in \figref{StateBasedModels} and are described in
-more detail below.
+approach, a graph-based approach, and an expression-based approach. Below, we
+describe the process-based approach and the graph-based approach informally and
+describe the expression-based approach formally. All three approaches are
+illustrated in \figref{StateBasedModels}.
 
@@
 The process-based model is most similar to the process model described by
-Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a set
-$\seq{p}{1}{n}$ of $n$ processors each of which begins with state $s_0$.  A
-processor $p_i$ can perform one of two actions.
+Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a state-based
+object $O = (S, s_0, \join)$, a set of transactions $T$, an invariant $I$, and
+a set $\seq{p}{1}{n}$ of $n$ processors. Each processor begins with state
+$s_0$. A processor $p_i$ can perform one of two actions.
 
-First, $p_i$ can execute a transaction $t \in T$ to transition from state $s$
-to state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
-execute $t$ (or it will execute $t$ but then abort it; think of it however you
-like).
+\begin{itemize}
+  \item
+    $p_i$ can execute a transaction $t \in T$ to transition from state $s$ to
+    state $t(s)$, given that $I(t(s))$. If $\lnot I(t(s))$, then $p_i$ will not
+    execute $t$ (or it will execute $t$ but then abort it; think of it however
+    you like).
 
-Second, $p_i$ can send its state $s_i$ to another processor $p_j$ with state
-$s_j$ causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
+  \item
+    $p_i$ can send its state $s_i$ to another processor $p_j$ with state $s_j$
+    causing $p_j$ to transition from state $s_j$ to state $s_i \join s_j$ (or
     $s_j \join s_i$; it doesn't really matter). Note that unlike executing a
-transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if $\lnot
-I(s_i \join s_j)$.
+    transaction, $p_j$ must transition from $s_j$ to $s_i \join s_j$ even if
+    $\lnot I(s_i \join s_j)$.
+\end{itemize}
 
-$T$ is invariant-confluent with respect to $I$, abbreviated \Iconfluent{}, if
-every reachable state (including $s_0$) satisfies the invariant. Note that this
-definition of \Iconfluence{} is different from the original definition
+$(O, T)$ is invariant-confluent with respect to $I$ (abbreviated \Iconfluent{})
+if every reachable state (including $s_0$) satisfies the invariant. Note that
+this definition of \Iconfluence{} is different from the original definition
 in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.
@@
 states collapsed into a single vertex. As with the process-based model, we say
-that $T$ is \Iconfluent{} if every vertex in every graph satisfies the
+that $(O, T)$ is \Iconfluent{} if every vertex in every graph satisfies the
 invariant.
@@
 The expression-based approach formalizes the process-based and graph-based
-approach using expressions. Given a state based object $(S, s_0, \join)$ and a
-set of transactions $T$, we consider expressions generated by the following
-grammar:
+approach using expressions. Given a state based object $O = (S, s_0, \join)$
+and a set of transactions $T$, we consider expressions generated by the
+following grammar:
 \[
@@
 \begin{definition}
-  $T$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq I$. In
-  other words, $T$ is \Iconfluent{} if all reachable states satisfy the
+  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
+  I$. In other words, $T$ is \Iconfluent{} if all reachable states satisfy the
   invariant.
@@
   \begin{mathpar}
-    \Irec{s_0} \defeq I(s_0)
+    \inferrule{I(s_0)}{\Irec{s_0}}
 
-    \Irec{t(e)} \defeq I(t(e)) \land \Irec{e}
+    \inferrule{I(t(e)) \\ \Irec{e}}{\Irec{t(e)}}
 
-    \Irec{e_1 \join e_2}
-    \defeq I(e_1 \join e_2) \land \Irec{e_1} \land \Irec{e_2}
+    \inferrule{I(e_1 \join e_2) \\ \Irec{e_1} \\ \Irec{e_2}}
+              {\Irec{e_1 \join e_2}}
   \end{mathpar}
@@
 Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
-satisfying $I$ are closed under join. That is, $T$ is \Iconfluent{} if
+satisfying $I$ are closed under join. That is, $(O, T)$ is \Iconfluent{} if
 \[
@@
 
-\begin{claim}\clmlabel{StateBasedIrecImpliesReachable}
+\begin{lemma}\lemlabel{StateBasedIrecImpliesReachable}
   $\Irec{e} \implies \reachable{e}$
-\end{claim}
+\end{lemma}
 \begin{elidableproof}
@@
 \begin{claim}\clmlabel{StateBasedTwoIconfluenceDefs}
-  Consider a state based object $(S, s_0, \join)$, a set of transactions $T$,
-  and an invariant $I$. The following two are equivalent:
+  Consider a state based object $O = (S, s_0, \join)$, a set of transactions
+  $T$, and an invariant $I$. The following two are equivalent:
   \begin{enumerate}[\quad(1)]
@@
   (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
-  $\Irec{e_1}$ and $\Irec{e_2}$. By \clmref{StateBasedIrecImpliesReachable},
+  $\Irec{e_1}$ and $\Irec{e_2}$. By \lemref{StateBasedIrecImpliesReachable},
   $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
@@
 \begin{definition}
-  A \defword{distributed operation-based object} is a pair $(S, s_0)$ where $S$
-  is a set of states and $s_0 \in S$ is a designated start state.
+  A \defword{distributed operation-based object} is a pair $O = (S, s_0)$ where
+  $S$ is a set of states and $s_0 \in S$ is a designated start state.
 \end{definition}
@@
 \begin{example}
-  $(\nats, 0)$ is a distributed object. $t(x) = y \mapsto x + y$ is a
-  transaction that given a state $x$ returns a function $y \mapsto x + y$ that
-  adds $x$ to its argument. $\setst{x \geq 0}{x \in \nats}$ is an invariant.
+  $(\nats, 0)$ is a distributed object. $t(x) = \lambda y.\ x + y$ is a
+  transaction that given a state $x$ returns a function $\lambda y.\ x + y$
+  that adds $x$ to its argument. $\setst{x \in \nats}{x \geq 0}$ is an
+  invariant.
 \end{example}
@@
 the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
-We are given a set $\seq{p}{1}{n}$ of $n$ processors each of which begins with
-state $s_0$. A processor $p_i$ can do one of two things.
+We are given an operation-based object $O$, a set of transactions $T$, an
+invariant $I$, and a set $\seq{p}{1}{n}$ of $n$ processors. Each processors
+begins with state $s_0$. A processor $p_i$ can do one of two things.
 
-First, $p_i$ can execute a transaction $t$ on its current state $s$ and then
+\begin{itemize}
+  \item
+    $p_i$ can execute a transaction $t$ on its current state $s$ and then
     transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
-$\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute $t$,
-then it broadcasts $t(s)$ to all other processors exactly once.
+    $\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute
+    $t$, then it broadcasts $t(s)$ to all other processors exactly once.
 
-Second, $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some
-other processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
-state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow transaction,
-it must execute it, even if $\lnot I(t(s_j)(s_i))$.
+  \item
+    $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some other
+    processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
+    state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow
+    transaction, it must execute it, even if $\lnot I(t(s_j)(s_i))$.
+\end{itemize}
 
-$T$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies the
-invariant.
+$(O, T)$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies
+the invariant.
 
@@
 vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
-where $s$ is some other vertex in the graph. $T$ is \Iconfluent{} if every
+where $s$ is some other vertex in the graph. $(O, T)$ is \Iconfluent{} if every
 vertex in every graph satisfies $I$.
@@
 \subsubsection{Expression-Based}
-Given an operation-based object $(S, s_0)$ and a set of transactions $T$, we
-consider expressions built from the following grammar:
+Given an operation-based object $O = (S, s_0)$ and a set of transactions $T$,
+we consider expressions built from the following grammar:
 \[
@@
 \begin{definition}
-  $T$ is \Iconfluent{} if $\setst{e}{\reachable{e}} \subseteq I$.
+  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
+  I$.
 \end{definition}
@@
   \begin{mathpar}
-    \Irec{s_0} \defeq I(s_0)
+    \inferrule{I(s_0)}{\Irec{s_0}}
 
-    \Irec{t(e_1)(e_2)} \defeq I(t(e_1)(e_2)) \land \Irec{e_1} \land \Irec{e_2}
+    \inferrule{I(t(e_1)(e_2)) \\ \Irec{e_1} \\ \Irec{e_2}}{\Irec{t(e_1)(e_2)}}
   \end{mathpar}
@@
 
-\begin{claim}\clmlabel{OpBasedIrecImpliesReachable}
+\begin{lemma}\lemlabel{OpBasedIrecImpliesReachable}
   $\Irec{e} \implies \reachable{e}$
-\end{claim}
+\end{lemma}
 \begin{elidableproof}
@@
   Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \clmref{OpBasedIrecImpliesReachable},
+  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \lemref{OpBasedIrecImpliesReachable},
   $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By

### 2018-03-03 doc/paper_redux/interactive_decision_procedure.tex
Notational changes and writing improvements.

@@
 transactions $T$, and an invariant $I$. We want to know if $T$ is
-\Iconfluent{}. We can hand $O$, $T$, and $I$ to the \Iclosed{} decision
-procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes, then
-by \clmref{StateBasedSufficient}, we know that $T$ is \Iconfluent{}.
+\Iconfluent{}. Assume we have already determined that $I(s_0)$. We can hand $O$
+and $I$ to the \Iclosed{} decision procedure to see if $O$ is \Iclosed{}. If
+the decision procedure says yes, then by \clmref{StateBasedSufficient}, we know
+that $(O, T)$ is \Iconfluent{}.
 
@@
 say no. \clmref{SufficientAndNecessary} tells us that this means one of two
-things. Either (1) $T$ is not \Iconfluent{} or (2) $T$ is \Iconfluent{}, but
-there are some invariant satisfying states that are not reachable. If we think
-harder about the set of reachable states, we may realize that all reachable
-states have a non-negative value of $x$. Thus, we may modify our invariant to
-$\setst{(x, y)}{xy \leq 0 \land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand
-$O$, $T$, and $I$ to the decision procedure, it will say yes, and we can
-conclude that $T$ is \Iconfluent{}.
+things. Either (1) $T$ is not \Iconfluent{} or (2) $(O, T)$ is \Iconfluent{},
+but there are some invariant satisfying states that are not reachable. If we
+think harder about the set of reachable states, we may realize that all
+reachable states have a non-negative value of $x$. Thus, we may modify our
+invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0 \land x
+\geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the decision
+procedure, it will say yes, and we can conclude that $(O, T)$ is \Iconfluent{}.
 
@@
     \Function{IsIconfluent}{$O$, $T$, $I$}
+      \If {$\lnot I(s_0)$}
+        \State \Return no
+      \Else
         \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
+      \EndIf
     \EndFunction
@@
     \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
-      \State closed, $s_1$, $s_2$ $\gets$
-               \Call{IsIclosed}{$O$, $T$, $I - NR$}
+      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
       \If {closed $=$ unknown}
@@
 Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
-\IsIconfluent$(O, T, I)$ to determine if $T$ is \Iconfluent{}.
+\IsIconfluent$(O, T, I)$ to determine if $(O, T)$ is \Iconfluent{}.
 
 \IsIconfluent{} assumes access to an \Iclosed{} decision procedure
-\IsIclosed$(O, T, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
+\IsIclosed$(O, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
 unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
@@
     both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
-    I(s_1 \join s_2)$. Thus, $T$ is not \Iconfluent{}, so \Helper{} returns no.
-    Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we recurse.
+    I(s_1 \join s_2)$. Thus, $(O, T)$ is not \Iconfluent{}, so \Helper{}
+    returns no.  Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we
+    recurse.
 \end{itemize}
@@
   \item
-    \IsIconfluent{} calls \Helper{} with $R = {(0, 0)}, NR = \emptyset{}$.
+    \IsIconfluent{} calls \Helper{} with $R = \set{(0, 0)}, NR = \emptyset{}$.
 
@@
 \end{itemize}
+
+\todo{What if the user labels things incorrectly?}

### 2018-03-03 doc/paper_redux/new_statebased_sufficient.tex
Made some proofs elidable.
Notational changes and writing improvements.

@@
 \section{New Sufficient State-Based Conditions}
-\todo{Motivate these in terms of the interactive decision procedure.}
+\todo{%
+  Revise everything with new notation of $(O, T)$ \Iconfluent{} instead of $T$
+  \Iconfluent{}.
+}
 
-In this section, we outline a new sufficient condition for state-based
-\Iconfluence{} that is more general than the one in
-\clmref{StateBasedSufficient}. To prove that the condition is sufficient, we
-introduce the concepts of \dIconfluence{} and \dIreduction{}.
+\algoref{InteractiveDecisionProcedure} relies on a user to provide information
+about reachable and unreachable states. In this section, we outline a new
+sufficient condition for state-based \Iconfluence{} that does not require user
+interaction and that covers some cases not covered by the sufficient conditions
+in \clmref{StateBasedSufficient}. To prove that the new condition is
+sufficient, we introduce the concepts of \dIconfluence{} and \dIreduction{}.
 
@@
 \end{claim}
-\begin{proof}
+\begin{elidableproof}
   First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
@@
   Thus, $I(e_1 \join e_2)$.
-\end{proof}
+\end{elidableproof}
 
@@
 \end{claim}
-\begin{proof}
+\begin{elidableproof}
   \newcommand{\bart}[1]{\overline{t_{#1}}}
@@
   expression, so $T$ is \dIreducible{}.
-\end{proof}
+\end{elidableproof}

### 2018-03-03 doc/paper_redux/sufficient_conditions.tex
Moved figs into a separate directory.
Notational changes and writing improvements.

@@
 \begin{definition}
-  We say a state-based object is \defword{\Iclosed{}} if $I(s_0)$ and for all
-  states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
+  We say a state-based object $O = (S, s_0, \join)$ is \defword{\Iclosed{}} if
+  for all states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join
+  s_2)$.
   \[
@@
   \]
-  Similarly, we say on operation-based object is \defword{\Iclosed{}} if
-  $I(s_0)$ and for all states $s_1$ and $s_2$ and all transactions $t$, if
-  $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
+  Similarly, we say on operation-based object is \defword{\Iclosed{}} if for
+  all states $s_1$ and $s_2$ and all transactions $t$, if $I(s_1)$, $I(s_2)$,
+  and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
   \[
@@
   \]
-
-  \todo{%
-    We probably don't want to include $I(s_0)$ in the definition of \Iclosed{}
-    because it's a bit confusing.
-  }
 \end{definition}
@@
   Given a state-based object $O = (S, s_0, \join)$, a set of transactions $T$,
-  and an invariant $I$, if $O$ is \Iclosed{}, then $T$ is \Iconfluent.
+  and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$, then $(O, T)$ is
+  \Iconfluent.
 \end{claim}
@@
   $I(s_1 \join s_2)$ which implies $I(e_1 \join e_2)$. Thus, by
-  \clmref{StateBasedTwoIconfluenceDefs}, $T$ is \Iconfluent{}.
+  \clmref{StateBasedTwoIconfluenceDefs}, $(O, T)$ is \Iconfluent{}.
 \end{elidableproof}
@@
   Given a operation-based object $O = (S, s_0)$, a set of transactions $T$, and
-  an invariant $I$, if $O$ is \Iclosed{}, then $T$ is \Iconfluent.
+  an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$, then $(O, T)$ is
+  \Iconfluent.
 \end{claim}
@@
   $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$, so by
-  \clmref{OpBasedTwoIconfluenceDefs}, $T$ is \Iconfluent.
+  \clmref{OpBasedTwoIconfluenceDefs}, $(O, T)$ is \Iconfluent.
 \end{elidableproof}
@@
   semilattice. Let $s \leq s' \defeq s \join s' = s'$. Let $T$ be an arbitrary
-  set of transactions. Let $I$ be a \defword{monotonic}: for all $s, s'\in S$,
-  if $s \leq s'$ and $I(s)$, then $I(s')$. Also assume $I(s_0)$. $T$ is
-  \Iconfluent{}.
+  set of transactions. Let invariant $I$ be \defword{monotonic}: for all $s,
+  s'\in S$, if $s \leq s'$, then $I(s) \implies I(s')$. Also assume $I(s_0)$.
+  Then, $T$ is \Iconfluent{}.
 \end{claim}
@@
   $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
-  $T$ is \Iconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
+  $(O, T)$ is \Iconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
   $I(s_1)$ and $I(s_2)$. However, $s_1 \join s_2 = (1, 1)$, and $\lnot I((1,
@@
   $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
-  $T$ is \Iconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
+  $(O, T)$ is \Iconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
   $I(s_1)$, $I(s_2)$, and $I(t_{x+1}(s_1)(s_1))$. However, $t_{x+1}(s_1)(s_2) =
@@
 
-\input{not_necessary.fig}
+\input{figs/not_necessary.fig}
 
@@
 $s_2 = (1, -1)$. $s_2$ is reachable, but $s_1$ is not. This is not a
-coincidence! If we consider an \Iconfluent{} set of transactions, then any two
-reachable points satisfy the invariant, and the join of the two points is again
-reachable, so it must satisfy the invariant as well. The only way to join two
-invariant satisfying points and end up not satisfying the invariant is if one
-or both points are not reachable. Thus, if we know that $I$ is a subset of
-reachable points, then being \Iconfluent{} and being \Iclosed{} are equivalent.
+coincidence! If we consider an \Iconfluent{} object and set of transactions,
+then any two reachable points satisfy the invariant, and the join of the two
+points is again reachable, so it must satisfy the invariant as well. The only
+way to join two invariant satisfying points and end up \emph{not} satisfying
+the invariant is if one or both points are not reachable. Thus, if we know that
+$I$ is a subset of reachable points, then being \Iconfluent{} and being
+\Iclosed{} are equivalent.
 
@@
   If every invariant satisfying state is reachable, then being \Iconfluent{}
-  and being \Iclosed{} are equivalent. Being \Iclosed{} is always sufficient,
-  but if we can establish that every invariant satisfying point is reachable,
-  then being \Iclosed{} is a both necessary and sufficient for being
-  \Iconfluent{}.
+  and being \Iclosed{} are equivalent.
   \[
     I \subseteq \setst{e}{\reachable{e}} \implies
-    (\text{\Iclosed{}} \iff \text{\Iconfluent})
+    (\text{$O$ \Iclosed{}} \iff \text{$(O, T)$ \Iconfluent})
   \]
@@
   Being \Iclosed{} always implies being \Iconfluent{} whether or not $I
-  \subseteq \setst{e}{\reachable{e}}$. Thus, we only have to prove that if $T$
-  is \Iconfluent{}, then $O$ is \Iclosed. If $T$ is \Iconfluent{}, then
-  $I(s_0)$. Next, let $s_1$ and $s_2$ be arbitrary states such that $I(s_1)$
-  and $I(s_2)$. Both states satisfy the invariant, so both are reachable. That
-  is, there exist expressions $e_1$ and $e_2$ where $eval(e_1) = s_1$  and
-  $eval(e_2) = s_2$. $e_1 \join e_2$ is again reachable, so since $T$ is
-  \Iconfluent{}, $I(e_1 \join e_2)$ which implies $I(s_1 \join s_2)$. Thus, $O$
-  is \Iclosed.
+  \subseteq \setst{e}{\reachable{e}}$. Thus, we only have to prove that if $(O,
+  T)$ is \Iconfluent{}, then $O$ is \Iclosed. If $(O, T)$ is \Iconfluent{},
+  then $I(s_0)$. Next, let $s_1$ and $s_2$ be arbitrary states such that
+  $I(s_1)$ and $I(s_2)$. Both states satisfy the invariant, so both are
+  reachable. That is, there exist expressions $e_1$ and $e_2$ where $eval(e_1)
+  = s_1$  and $eval(e_2) = s_2$. $e_1 \join e_2$ is again reachable, so since
+  $T$ is \Iconfluent{}, $I(e_1 \join e_2)$ which implies $I(s_1 \join s_2)$.
+  Thus, $O$ is \Iclosed.
 \end{elidableproof}
@@
 Here are some more examples of transactions $T$ and invariants $I$ that are
-\Iconfluent{} but not \Iclosed{}. \todo{Think of more compelling examples.}
+\Iconfluent{} but not \Iclosed{}.
+
+\todo{Think of more compelling examples.}
 

### 2018-03-04 doc/paper_redux/interactive_decision_procedure.tex
Illustrated example interaction.

@@
 
-Let's walk through how we might use \IsIconfluent{} on
-\exampleref{StateBasedNotNecessary}.
-\begin{itemize}
-  \item
-    \IsIconfluent{} calls \Helper{} with $R = \set{(0, 0)}, NR = \emptyset{}$.
+An example of how to use \algoref{InteractiveDecisionProcedure} on
+\exampleref{StateBasedNotNecessary} is given in
+\figref{DecisionProcedureExample}.
 
-  \item
-    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
-    counterexample $s_1 = (-1, 1)$ and $s_2 = (1, -1)$. We take the third
-    branch.
+\input{figs/decision_procedure_example.fig}
 
-  \item
-    We randomly generate some reachable points and update $R$ to $\set{(0, 0),
-    (1, 0), (0, -1), (1, -1)}$. Lucky for us, $s_2 \in R$, so we know it is
-    reachable. We're not sure about $s_1$, so we ask the user. After some
-    thought, the user tells us that $s_1$ is unreachable. We recurse with $R =
-    \set{(0, 0), (1, 0), (0, -1), (1, -1)}$ and $NR = \set{(-1, 1)}$.
-
-  \item
-    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
-    counterexample $s_1 = (-1, 2)$ and $s_2 = (10, -10)$. We take the third
-    branch.
-
-  \item
-    We randomly generate some more reachable points and update $R$ to $\set{(0,
-    0), (1, 0), (0, -1), (1, -1), (2, 0), (0, -2)}$. $s_1, s_2 \notin R, NR$,
-    so we ask the user to label them. The user puts $s_1$ in $NR$ and $s_2$ in
-    $R$. We recurse with $R = \set{(0, 0), (1, 0), (0, -1), (1, -1), (2, 0),
-    (0, -2), (10, -10)}$ and $NR = \set{(-1, 1), (-1, 2)}$.
-
-  \item
-    \Helper{} determines that $O$ is not \iclosed{$(I - R)$} with
-    counterexample $s_1 = (-2, 1)$ and $s_2 = (4, -3)$. We take the third
-    branch.
-
-  \item
-    We randomly generate some more reachable points and update $R$. $s_1, s_2
-    \notin R, NR$, so we ask the user to label them. The user notices a pattern
-    in $R$ and $NR$ and after some thought, concludes that every point with
-    negative $x$-coordinate is unreachable. They update $NR$ to $-\ints \times
-    \ints$. We recurse.
-
-  \item
-    \Helper{} determines that $O$ is \iclosed{$(I - R)$} and returns yes!
-\end{itemize}
-
-\todo{What if the user labels things incorrectly?}
+\todo{Describe what happens when user input is incorrect.}

### 2018-03-04 doc/paper_redux/iconfluence.tex
Revised everything with new s_0, T, I definitions.

@@
-\section{\Iconfluence{}}
-In this section, we present six different ways to think about \Iconfluence{}:
-three state-based approaches and three operation-based approaches.
+\section{Invariant-Confluence}
+
+In this section, we present six different ways to think about
+invariant-confluence: three state-based approaches and three operation-based
+approaches.
 
@@
 \begin{definition}
-  A \defword{distributed state-based object} $O$ is a triple $(S, s_0, \join)$
-  where $S$ is a set of states, $s_0 \in S$ is a designated start state, and
-  $\join: S \times S \to S$ is a binary merge operator that we call join.
+  A \defword{distributed state-based object} $O$ is a pair $(S, \join)$ where
+  $S$ is a set of states, and $\join: S \times S \to S$ is a binary merge
+  operator that we call join.
 \end{definition}
@@
 \begin{example}
-  $(\nats, 0, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
+  $(\nats, +)$ is a distributed object, $t(x) = 2x$ is a transaction, and
   $\setst{x \in \nats}{\text{$x$ is even}}$ is an invariant. Here, $I(0)$ and
@@
 
-There are three ways to think about state-based \Iconfluence{}: a process-based
-approach, a graph-based approach, and an expression-based approach. Below, we
-describe the process-based approach and the graph-based approach informally and
-describe the expression-based approach formally. All three approaches are
-illustrated in \figref{StateBasedModels}.
+There are three ways to think about state-based invariant-confluence: a
+process-based approach, a graph-based approach, and an expression-based
+approach. Below, we describe the process-based approach and the graph-based
+approach informally and describe the expression-based approach formally. All
+three approaches are illustrated in \figref{StateBasedModels}.
 
@@
 Shapiro et al.\ in~\cite{shapiro2011conflict}. We are given a state-based
-object $O = (S, s_0, \join)$, a set of transactions $T$, an invariant $I$, and
-a set $\seq{p}{1}{n}$ of $n$ processors. Each processor begins with state
-$s_0$. A processor $p_i$ can perform one of two actions.
+object $O = (S, \join)$, a start state $s_0 \in S$, a set of transactions $T$,
+an invariant $I$, and a set $\seq{p}{1}{n}$ of $n$ processors. Each processor
+begins with state $s_0$. A processor $p_i$ can perform one of two actions.
 
@@
 
-$(O, T)$ is invariant-confluent with respect to $I$ (abbreviated \Iconfluent{})
-if every reachable state (including $s_0$) satisfies the invariant. Note that
-this definition of \Iconfluence{} is different from the original definition
-in~\cite{bailis2014coordination}. Later, we will see that they are equivalent.
+$O$ is invariant-confluent with respect to $s_0$, $T$, and $I$ (abbreviated $O$
+is \sTIconfluent{}) if every reachable state (including $s_0$) satisfies the
+invariant. Note that this definition of invariant-confluence is different from
+the original definition in~\cite{bailis2014coordination}. Later, we will see
+that they are equivalent.
 
@@
 states collapsed into a single vertex. As with the process-based model, we say
-that $(O, T)$ is \Iconfluent{} if every vertex in every graph satisfies the
+that $O$ is \sTIconfluent{} if every vertex in every graph satisfies the
 invariant.
@@
 The expression-based approach formalizes the process-based and graph-based
-approach using expressions. Given a state based object $O = (S, s_0, \join)$
-and a set of transactions $T$, we consider expressions generated by the
-following grammar:
+approach using expressions. Given a state based object $O = (S, \join)$ and a
+set of transactions $T$, we consider expressions generated by the following
+grammar:
 \[
-  e ::= s_0 \mid t(e) \mid e_1 \join e_2
+  e ::= s \mid t(e) \mid e_1 \join e_2
 \]
-where $t$ corresponds to a transaction in $T$. We can evaluate an expression
-$e$, written $eval(e)$, in the obvious way:
+where $s$ correspond to a state in $S$ and $t$ corresponds to a transaction in
+$T$. We can evaluate an expression $e$, written $eval(e)$, in the obvious way:
 \begin{mathpar}
-  eval(s_0) \defeq s_0
+  eval(s) \defeq s
 
@@
 \begin{definition}
-  Intuitively, we say an expression is \defword{reachable} if it can be reached
-  in an execution of the process-based model. Formally, we defined a predicate
-  $\reachable{\cdot}$ as follows:
+  We say an expression is \defword{$\sTIreachable{}$} if it can be reached in
+  an execution of the process-based model with starting state $s_0$,
+  transactions $T$, and invariant $I$. Formally, we defined a predicate
+  $\sTIreachable{}$ as follows:
 
   \begin{mathpar}
-    \inferrule{ }{\reachable{s_0}}
+    \inferrule{ }{\sTIreachable{s_0}}
 
-    \inferrule{\reachable{e} \\ I(t(e))}{\reachable{t(e})}
+    \inferrule{\sTIreachable{e} \\ t \in T \\ I(t(e))}{\sTIreachable{t(e})}
 
-    \inferrule{\reachable{e_1} \\ \reachable{e_2}}{\reachable{e_1 \join e_2}}
+    \inferrule{\sTIreachable{e_1} \\ \sTIreachable{e_2}}
+              {\sTIreachable{e_1 \join e_2}}
   \end{mathpar}
+
+  We say a state us $\sTIreachable{}$ if there exists an $\sTIreachable{}$
+  expression $e$ such that $eval(e) = s$.
 \end{definition}
@@
 \begin{definition}
-  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
-  I$. In other words, $T$ is \Iconfluent{} if all reachable states satisfy the
-  invariant.
+  $O$ is \defword{\sTIconfluent{}} if $\setst{s}{\sTIreachable{s}} \subseteq
+  I$. In other words, $T$ is \sTIconfluent{} if all $\sTIreachable{}$ states
+  satisfy the invariant.
 \end{definition}
 
-As we mentioned earlier, this definition of \Iconfluence{} is different than
+As we mentioned earlier, this definition of \sTIconfluence{} is different than
 the original definition presented in~\cite{bailis2014coordination} (though it
@@
 \begin{definition}
-  An expression \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if $e$
-  and all of $e$'s children satisfy $I$. That is,
+  An expression $e$ \defword{recursively satisfies $I$}, denoted $\Irec{e}$, if
+  $e$ and all of $e$'s children satisfy $I$. That is,
   \begin{mathpar}
-    \inferrule{I(s_0)}{\Irec{s_0}}
+    \inferrule{I(s)}{\Irec{s}}
 
@@
 
-Bailis et al.\ defined \Iconfluence{} to mean that expressions recursively
-satisfying $I$ are closed under join. That is, $(O, T)$ is \Iconfluent{} if
+Bailis et al.\ defined \sTIconfluence{} to mean that all $\sTIreachable{}$
+expressions recursively satisfying $I$ are closed under join. That is, $O$ is
+\sTIconfluent{} if
 \[
-  \forall e_1, e_2.\; \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
+  \forall e_1, e_2 \in \setst{e}{\sTIreachable{e}}.\;
+    \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
 \]
-If $\lnot I(s_0)$, then \Iconfluence{} holds vacuously which is a bit silly, so
-Bailis et al.\ probably ought to have also added the condition $I(s_0)$ to
-their definition of \Iconfluence{}. Doing so, the two definitions become
-equivalent.
-
-\begin{lemma}\lemlabel{StateBasedIrecImpliesReachable}
-  $\Irec{e} \implies \reachable{e}$
-\end{lemma}
-\begin{elidableproof}
-  We perform structural induction on $e$.
-  \begin{itemize}
-    \item \textbf{Case $s_0$.}
-      Axiomatically, $\reachable{s_0}$.
-
-    \item \textbf{Case $t(e)$.}
-      $\Irec{t(e)}$, so $I(t(e))$ and $\Irec{e}$. By the inductive hypothesis,
-      $\reachable{e}$, so by the definition of $\reachable{\cdot}$,
-      $\reachable{t(e)}$.
-
-    \item \textbf{Case $e_1 \join e_2$.}
-      $\Irec{t(e)}$, so $\Irec{e_1}$ and $\Irec{e_2}$. By the inductive
-      hypothesis, $\reachable{e_1}$ and $\reachable{e_2}$, so by the definition
-      of $\reachable{\cdot}$, $\reachable{e_1 \join e_2}$.
-  \end{itemize}
-\end{elidableproof}
+If $\lnot I(s_0)$, then invariant-confluence holds vacuously which is a bit
+silly, so Bailis et al.\ probably ought to have also added the condition
+$I(s_0)$ to their definition of invariant-confluence. Doing so, the two
+definitions become equivalent.
 
 \begin{claim}\clmlabel{StateBasedTwoIconfluenceDefs}
-  Consider a state based object $O = (S, s_0, \join)$, a set of transactions
-  $T$, and an invariant $I$. The following two are equivalent:
+  Consider a state based object $O = (S, \join)$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$. The following two are equivalent:
   \begin{enumerate}[\quad(1)]
     \item
-      $\setst{e}{\reachable{e}} \subseteq I$
+      $\setst{s}{\sTIreachable{s}} \subseteq I$
 
     \item
-      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
+      $I(s_0)$ and
+      $\forall e_1, e_2 \in \setst{e}{\sTIreachable{e}}.\;
+         \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
   \end{enumerate}
@@
 \begin{elidableproof}
-  First, we show that (1) implies (2). Axiomatically, $\reachable{s_0}$, so by
-  (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary expressions such that
-  $\Irec{e_1}$ and $\Irec{e_2}$. By \lemref{StateBasedIrecImpliesReachable},
-  $\reachable{e_1}$ and $\reachable{e_2}$. Thus, $\reachable{e_1 \join e_2}$,
-  so by (1), $I(e_1 \join e_2)$.
+  First, we show that (1) implies (2). Axiomatically, $\sTIreachable{s_0}$, so
+  by (1), $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary $\sTIreachable{}$
+  expressions. Let $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$. $e_1 \join e_2$ is
+  $\sTIreachable{}$, so $s_1 \join s_2$ is $\sTIreachable{}$, so by (1), $I(s_1
+  \join s_2)$, so $I(e_1 \join e_2)$.
 
   Next, we show that (2) implies (1). We prove by structural induction that for
-  all $e$, $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
+  all $e$, $\sTIreachable{e} \implies \Irec{e}$. From this, (1) is immediate.
   \begin{itemize}
@@
     \item \textbf{Case $t(e)$.}
-      $\reachable{t(e)}$, so $\reachable{e}$ and $I(t(e))$. By the inductive
-      hypothesis, $\Irec{e}$, so by the definition of $\Irec{\cdot}$,
-      $\Irec{t(e)}$.
+      Let $t(e)$ be $\sTIreachable{}$. Then, $\sTIreachable{e}$ and $I(t(e))$.
+      By the inductive hypothesis, $\Irec{e}$, so by the definition of
+      $\Irec{\cdot}$, $\Irec{t(e)}$.
 
     \item \textbf{Case $e_1 \join e_2$.}
-      $\reachable{e_1 \join e_2}$, so $\reachable{e_1}$ and $\reachable{e_2}$.
-      By the inductive hypothesis, $\Irec{e_1}$ and $\Irec{e_2}$. By $(2)$,
-      $I(e_1 \join e_2)$. Thus, $\Irec{e_1 \join e_2}$.
+      Let $e_1 \join e_2$ be $\sTIreachable{}$. Then, $\sTIreachable{e_1}$ and
+      $\sTIreachable{e_2}$.  By the inductive hypothesis, $\Irec{e_1}$ and
+      $\Irec{e_2}$. By $(2)$, $I(e_1 \join e_2)$. Thus, by the definition of
+      $\Irec{\cdot}$, $\Irec{e_1 \join e_2}$.
   \end{itemize}
@@
 \begin{definition}
-  A \defword{distributed operation-based object} is a pair $O = (S, s_0)$ where
-  $S$ is a set of states and $s_0 \in S$ is a designated start state.
+  A \defword{distributed operation-based object} is a set $O = S$ of states.
 \end{definition}
@@
 \begin{example}
-  $(\nats, 0)$ is a distributed object. $t(x) = \lambda y.\ x + y$ is a
-  transaction that given a state $x$ returns a function $\lambda y.\ x + y$
-  that adds $x$ to its argument. $\setst{x \in \nats}{x \geq 0}$ is an
+  $\nats$ is a distributed operation-based object. $t(x) = \lambda y.\ x + y$
+  is a transaction that given a state $x$ returns a function $\lambda y.\ x +
+  y$ that adds $x$ to its argument. $\setst{x \in \nats}{x \geq 0}$ is an
   invariant.
@@
 As with state-based objects, there are three ways to think about
-operation-based \Iconfluence{}: a process-based approach, a graph-based
+operation-based invariant-confluence: a process-based approach, a graph-based
 approach, and an expression-based approach. All three approaches are
@@
 the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
-We are given an operation-based object $O$, a set of transactions $T$, an
-invariant $I$, and a set $\seq{p}{1}{n}$ of $n$ processors. Each processors
-begins with state $s_0$. A processor $p_i$ can do one of two things.
+We are given an operation-based object $O$, a start state $s_0$, a set of
+transactions $T$, an invariant $I$, and a set $\seq{p}{1}{n}$ of $n$
+processors. Each processors begins with state $s_0$. A processor $p_i$ can do
+one of two things.
 
@@
   \item
-    $p_i$ can receive a broadcasted shadow transaction $t(s_j)$ from some other
+    $p_i$ can receive a broadcast shadow transaction $t(s_j)$ from some other
     processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
@@
 
-$(O, T)$ is \Iconfluent{} if every reachable state (including $s_0$) satisfies
-the invariant.
+$O$ is \sTIconfluent{} if every reachable state (including $s_0$) satisfies the
+invariant.
 
@@
 vertex is a state $s$ and each edge is labelled with a shadow operation $t(s)$
-where $s$ is some other vertex in the graph. $(O, T)$ is \Iconfluent{} if every
-vertex in every graph satisfies $I$.
+where $s$ is some other vertex in the graph. $(O, T)$ is \sTIconfluent{} if
+every vertex in every graph satisfies $I$.
 
 \subsubsection{Expression-Based}
-Given an operation-based object $O = (S, s_0)$ and a set of transactions $T$,
-we consider expressions built from the following grammar:
+Given an operation-based object $O = S$ and a set of transactions $T$, we
+consider expressions built from the following grammar:
 \[
-  e ::= s_0 \mid t(e_1)(e_2)
+  e ::= s \mid t(e_1)(e_2)
 \]
-where $t$ corresponds to a transaction in $T$. We can evaluate an expression,
-denoted $eval(e)$, in the obvious way:
+where $s$ corresponds to a state in $S$ and $t$ corresponds to a transaction in
+$T$. We can evaluate an expression, denoted $eval(e)$, in the obvious way:
 \begin{mathpar}
-  eval(s_0) \defeq s_0
+  eval(s) \defeq s
 
@@
   As with state-based expressions, we formalize which operation-based
-  expressions are reachable.
+  expressions are $\sTIreachable{}$.
 
   \begin{mathpar}
-    \inferrule{ }{\reachable{s_0}}
+    \inferrule{ }{\sTIreachable{s_0}}
 
     \inferrule{
-      \reachable{e_1} \\
-      \reachable{e_2} \\
+      \sTIreachable{e_1} \\
+      \sTIreachable{e_2} \\
+      t \in T \\
       I(t(e_1)(e_1))
-    }{\reachable{t(e_1)(e_2)}}
+    }{\sTIreachable{t(e_1)(e_2)}}
   \end{mathpar}
+
+  Similarly, a state $s$ is $\sTIreachable{}$ if there exists an
+  $\sTIreachable{}$ expression $e$ such that $eval(e) = s$.
 \end{definition}
@@
 \begin{definition}
-  $(O, T)$ is \defword{\Iconfluent{}} if $\setst{e}{\reachable{e}} \subseteq
-  I$.
+  $O$ is \defword{\sTIconfluent{}} if $\setst{s}{\sTIreachable{s}} \subseteq I$.
 \end{definition}
 
-As with state-based objects, we have an equivalent definition of \Iconfluence{}
-that deals with the closure of recursively invariant satisfying states.
+As with state-based objects, we have an equivalent definition of
+\sTIconfluence{} that deals with the closure of recursively invariant
+satisfying states that are $\sTIreachable{}$.
 
@@
 
-\begin{lemma}\lemlabel{OpBasedIrecImpliesReachable}
-  $\Irec{e} \implies \reachable{e}$
-\end{lemma}
-\begin{elidableproof}
-  We perform a structural induction on $e$.
-  \begin{itemize}
-    \item \textbf{Case $s_0$.}
-      Axiomatically, $\reachable{s_0}$.
-
-    \item \textbf{Case $t(e_1)(e_2)$.}
-      $\Irec{t(e_1)(e_2)}$, so $\Irec{e_1}$, $\Irec{e_2}$, and
-      $I(t(e_1)(e_1))$. By the inductive hypothesis, $\reachable{e_1}$ and
-      $\reachable{e_2}$. Thus, $\reachable{t(e_1)(e_2)}$.
-  \end{itemize}
-\end{elidableproof}
-
 \begin{claim}\clmlabel{OpBasedTwoIconfluenceDefs}
-  Given an operation-based object $(S, s_0)$, a set of transactions $T$, and an
-  invariant $I$, the following two are equivalent:
+  Given an operation-based object $O = S$, a start sate $s_0$, a set of
+  transactions $T$, and an invariant $I$, the following two are equivalent:
   \begin{enumerate}[\quad(1)]
     \item
-      $\setst{e}{\reachable{e}} \subseteq I$
+      $\setst{s}{\sTIreachable{s}} \subseteq I$
 
     \item
-      $I(s_0)$ and $\Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
+      $I(s_0)$ and
+      $\forall e_1, e_2 \in \setst{e}{\sTIreachable{e}}.\;
+       \forall t \in T.\;
+        \Irec{e_1} \land \Irec{e_2} \land I(t(e_1)(e_1)) \implies
         I(t(e_1)(e_2))$.
@@
 \begin{elidableproof}
-  First, we show that (1) implies (2). $\reachable{s_0}$, so by (1), $I(s_0)$.
-  Let $e_1$ and $e_2$ be arbitrary expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. By \lemref{OpBasedIrecImpliesReachable},
-  $\reachable{e_1}$ and $\reachable{e_2}$, so $\reachable{t(e_1)(e_2)}$. By
-  (1), $I(t(e_1)(e_2))$.
+  First, we show that (1) implies (2). $\sTIreachable{s_0}$, so by (1),
+  $I(s_0)$. Let $e_1$ and $e_2$ be arbitrary $\sTIreachable{}$ expressions and
+  $t$ be an arbitrary transaction in $T$ such that $\Irec{e_1}$, $\Irec{e_2}$,
+  and $I(t(e_1)(e_1))$. Let $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$.
+  $\sTIreachable{t(e_1)(e_2)}$, so $\sTIreachable{t(s_1)(s_2)}$, so by (1),
+  $I(t(s_1)(s_2))$, so $I(t(e_1)(e_2))$.
 
-  Next, we show that (2) implies (1). We show by structural induction on $e$
-  that $\reachable{e} \implies \Irec{e}$. From this, (1) is immediate.
+  Next, we show that (2) implies (1). We show by structural induction that for
+  all $e$, $\sTIreachable{e} \implies \Irec{e}$. From this, (1) is immediate.
   \begin{itemize}
@@
     \item \textbf{Case $t(e_1)(e_2)$.}
-      $\reachable{t(e_1)(e_2)}$, so $\reachable{e_1}$, $\reachable{e_2}$, and
-      $I(t(e_1)(e_1))$. By the inductive hypothesis, $\Irec{e_1}$ and
-      $\Irec{e_2}$. By (2), $I(t(e_1)(e_2))$, so $\Irec{t(e_1)(e_2)}$.
+      Let $t(e_1)(e_2)$ be $\sTIreachable{}$. Then, $\sTIreachable{e_1}$,
+      $\sTIreachable{e_2}$, and $I(t(e_1)(e_1))$. By the inductive hypothesis,
+      $\Irec{e_1}$ and $\Irec{e_2}$. By (2), $I(t(e_1)(e_2))$, so
+      $\Irec{t(e_1)(e_2)}$.
   \end{itemize}

### 2018-03-04 doc/paper_redux/interactive_decision_procedure.tex
More interactive_decision_procedure tidying.
Revised everything with new s_0, T, I definitions.

@@
-\section{Interactive \Iconfluence{} Decision Procedure}
+\section{Interactive Invariant-Confluence Decision Procedure}
 Imagine we have an \Iclosed{} decision procedure, and let's revisit
-\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a set of
-transactions $T$, and an invariant $I$. We want to know if $T$ is
-\Iconfluent{}. Assume we have already determined that $I(s_0)$. We can hand $O$
-and $I$ to the \Iclosed{} decision procedure to see if $O$ is \Iclosed{}. If
-the decision procedure says yes, then by \clmref{StateBasedSufficient}, we know
-that $(O, T)$ is \Iconfluent{}.
+\exampleref{StateBasedNotNecessary}. We have a state-based object $O$, a start
+state $s_0$, set of transactions $T$, and an invariant $I$. We want to know if
+$O$ is \sTIconfluent{}. We can hand $O$ and $I$ to the \Iclosed{} decision
+procedure to see if $O$ is \Iclosed{}. If the decision procedure says yes and
+we can conclude that $I(s_0)$, then by \clmref{StateBasedSufficient}, we know
+that $O$ is \sTIconfluent{}.
 
@@
 say no. \clmref{SufficientAndNecessary} tells us that this means one of two
-things. Either (1) $T$ is not \Iconfluent{} or (2) $(O, T)$ is \Iconfluent{},
-but there are some invariant satisfying states that are not reachable. If we
-think harder about the set of reachable states, we may realize that all
-reachable states have a non-negative value of $x$. Thus, we may modify our
-invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0 \land x
-\geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the decision
-procedure, it will say yes, and we can conclude that $(O, T)$ is \Iconfluent{}.
+things. Either (1) $O$ is \emph{not} \sTIconfluent{} or (2) $O$ is
+\sTIconfluent{}, but there are some invariant satisfying states that are not
+reachable. If we think harder about the set of reachable states, we may realize
+that all reachable states have a non-negative value of $x$. Thus, we may modify
+our invariant from $\setst{(x, y)}{xy \leq 0}$ to $\setst{(x, y)}{xy \leq 0
+\land x \geq 0}$. Now, $O$ is \Iclosed{}! If we hand $O$ and $I$ to the
+decision procedure, it will say yes. Moreover, $I(s_0)$, so we can conclude
+that $O$ is \sTIconfluent{}.
 
@@
 \begin{algorithm}
-  \caption{Interactive \Iconfluence{} Decision Procedure}%
+  \caption{Interactive Invariant-Confluence Decision Procedure}%
   \algolabel{InteractiveDecisionProcedure}
   \begin{algorithmic}
-    \Function{IsIconfluent}{$O$, $T$, $I$}
-      \If {$\lnot I(s_0)$}
-        \State \Return no
-      \Else
-        \State \Return \Call{Helper}{$O$, $T$, $I$, $\set{s_0}$, $\emptyset$}
-      \EndIf
+    \Function{IsIconfluent}{$O$, $s_0$, $T$, $I$}
+      \State
+        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
     \EndFunction
@@
 
-    \comment{$R$ is a set of reachable states.}
-    \comment{$NR$ is a set of non-reachable states.}
-    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
+    \comment{$R$ is a set of $\sTIreachable{}$ states.}
+    \comment{$NR$ is a set of $\sTIunreachable{}$ states.}
+    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
+      \If {$\lnot(I - NR)(s_0)$}
+        \State \Return no
+      \EndIf
       \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
@@
 
+\todo{Rename IsIconfluent to IsSTIConfluent}
 \newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
@@
 This example motivates \algoref{InteractiveDecisionProcedure}, a general
-purpose interactive \Iconfluence{} decision procedure in which a user
+purpose interactive invariant-confluence decision procedure in which a user
 iteratively refines the invariant until the decision procedure terminates.
-Given an object $O$, a set of transactions $T$, and an invariant $I$, we call
-\IsIconfluent$(O, T, I)$ to determine if $(O, T)$ is \Iconfluent{}.
+Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
+invariant $I$, we call \IsIconfluent$(O, s_0, T, I)$ to determine if $O$ is
+\sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
+example of how to use \algoref{InteractiveDecisionProcedure} on
+\exampleref{StateBasedNotNecessary} is given in
+\figref{DecisionProcedureExample}.
 
@@
 
-\IsIconfluent{} takes in $O$, $T$, and $I$. Formally, $I$ is set of states
-(e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an actual
-implementation of \IsIconfluent{} would likely take in $I$ as a logical formula
-(e.g. $xy \leq 0$). From now on, whenever we mention a (possibly infinite) set
-of states, remember that we can always represent the set as a finite formula.
+\IsIconfluent{} takes in $O$, $s_0$, $T$, and $I$. Formally, $I$ is set of
+states (e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an
+actual implementation of \IsIconfluent{} would likely take in $I$ as a logical
+formula (e.g. $xy \leq 0$). From now on, whenever we mention a (possibly
+infinite) set of states, remember that we can always represent the set as a
+finite formula.
 
 \IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
-$T$, and $I$---takes as arguments a set $R$ of reachable states and a set $NR$
-of unreachable states. Like \IsIconfluent, \Helper$(O, T, I, R, NR)$ returns
-whether $T$ is \Iconfluent. Intuitively, $NR$ helps remove unreachable states
-from $I$ moving us closer to the case where $I \subseteq
-\setst{e}{\reachable{e}}$.
+$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
+and a set $NR$ of $\sTIunreachable{}$ states. Like \IsIconfluent, \Helper$(O,
+s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
+$NR$ are correct). Intuitively, $NR$ helps remove unreachable states from $I$
+moving us closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.
 
-\Helper{} first checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
+\Helper{} first checks to see if $(I - NR)(s_0)$. If it is not, then \Helper{}
+returns no. If $NR$ only contains unreachable states, then it doesn't contain
+$s_0$, so $s_0 \in I - NR \iff s_0 \in I$. Thus, if $\lnot (I - NR)(s_0)$, then
+$\lnot I(s_0)$, so $O$ is not \sTIconfluent.
+
+Next, \Helper{} checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
 cases to consider.
@@
     unknown.
+
   \item
     If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
-    \clmref{StateBasedSufficient}, $T$ is \iconfluent{$(I - NR)$}, so
-    $\setst{e}{\reachable{e}} \subseteq I - NR$. $I - NR \subseteq I$, so $T$
-    is therefore also \Iconfluent{}. Thus, \Helper{} return yes.
+    \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
+    NR)}, so
+    $
+      \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
+        \subseteq I - NR
+        \subseteq I
+    $.
+    If $NR$ only contains unreachable states, then
+    $
+      \setst{s}{\sTIreachable{s}} = \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
+    $.
+    Thus,
+    $
+      \setst{s}{\sTIreachable{s}} \subseteq I
+    $.
+    Thus, $O$ \sTIconfluent{}, so \Helper{} return yes.
+
   \item
@@
     have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
-    and $s_2$ are reachable or unreachable. We can do so in two ways.
+    and $s_2$ are $\sTIreachable{}$ or $\sTIunreachable{}$. We can do so in two
+    ways.
 
@@
     Second, we can prompt the user to tell us whether or not the states are
-    reachable or unreachable. In addition to labelling $s_1$ and $s_2$ as
-    reachable or not, the user can also refine $I$ by augmenting $NR$. Taking
-    \exampleref{StateBasedNotNecessary} as an example, the user might look
-    through the existing reachable and unreachable states and have an aha
-    moment realizing that every point with a negative $x$-coordinate is
-    unreachable. They can use this knowledge to add all points with negative
-    $x$-coordinates to $NR$.
+    $\sTIreachable{}$ or $\sTIunreachable{}$. In addition to labelling $s_1$
+    and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
+    augmenting $NR$. Taking \exampleref{StateBasedNotNecessary} as an example,
+    the user might look through the existing $\sTIreachable{}$ and
+    $\sTIunreachable{}$ states and have an aha moment realizing that every
+    point with a negative $x$-coordinate is $\sTIunreachable{}$. They can use
+    this knowledge to add all points with negative $x$-coordinates to $NR$.
 
-    After $s_1$ and $s_2$ have been labelled reachable or not, we continue. If
-    both $s_1$ and $s_2$ are reachable, then so is $s_1 \join s_2$, but $\lnot
-    I(s_1 \join s_2)$. Thus, $(O, T)$ is not \Iconfluent{}, so \Helper{}
-    returns no.  Otherwise, one of $s_1$ and $s_2$ is unreachable. Now, we
-    recurse.
+    After $s_1$ and $s_2$ have been labelled $\sTIreachable{}$ or not, we
+    continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1
+    \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
+    so \Helper{} returns no. Otherwise, one of $s_1$ and $s_2$ is
+    $\sTIunreachable{}$. Now, we recurse.
 \end{itemize}
-
-An example of how to use \algoref{InteractiveDecisionProcedure} on
-\exampleref{StateBasedNotNecessary} is given in
-\figref{DecisionProcedureExample}.
 
@@
 
-\todo{Describe what happens when user input is incorrect.}
+% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
+% procedure. It requires that a user classify states as reachable or unreachable.
+% What happens if a user incorrectly classifies a state? Given a counterexample
+% $s_1, s_2$, if a user incorrectly says that one of the unreachable points is
+% reachable, then the $s_1, s_2 \in R$ check in \Helper{} might be
+% erroneously met. In turn, \IsIconfluent{} might erroneously return no. This is
+% inconvenient, but not the end of the world.
+% Moreover, we can modify our decision procedure so that \Helper{}
+% user provide a \emph{proof} of reachability whenever they label as: a series of transactions and joins that result in the

### 2018-03-04 doc/paper_redux/new_statebased_sufficient.tex
Revised everything with new s_0, T, I definitions.

@@
 \section{New Sufficient State-Based Conditions}
-\todo{%
-  Revise everything with new notation of $(O, T)$ \Iconfluent{} instead of $T$
-  \Iconfluent{}.
-}
-
 \algoref{InteractiveDecisionProcedure} relies on a user to provide information
 about reachable and unreachable states. In this section, we outline a new
-sufficient condition for state-based \Iconfluence{} that does not require user
-interaction and that covers some cases not covered by the sufficient conditions
-in \clmref{StateBasedSufficient}. To prove that the new condition is
-sufficient, we introduce the concepts of \dIconfluence{} and \dIreduction{}.
+sufficient condition for state-based invariant-confluence that does not require
+user interaction and that covers some cases not covered by the sufficient
+conditions in \clmref{StateBasedSufficient}. To prove that the new condition is
+sufficient, we introduce the concepts of \dsTIconfluence{} and
+\dsTIreduction{}.
 
 \begin{definition}
-  A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s_0) \ldots))$ is an
+  A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s) \ldots))$ is an
   expression that does not contain a join.
@@
 \begin{definition}
-  $T$ is \defword{\dIconfluent{}} if for every pair $e_1$ and $e_2$ of
-  join-free expressions, $\Irec{e_1}$ and $\Irec{e_2}$ implies that $I(e_1
-  \join e_2)$.
+  $O$ is \defword{\dsTIconfluent{}} if (1) $I(s_0)$ and (2) for every pair
+  $e_1$ and $e_2$ of join-free $\sTIreachable{}$ expressions, $\Irec{e_1}$ and
+  $\Irec{e_2}$ implies that $I(e_1 \join e_2)$.
 \end{definition}
 
-Intuitively, \dIconfluence{} is the same thing as \Iconfluence{} with
-the added restriction that subexpressions are join-free.
+Intuitively, \dsTIconfluence{} is the same thing as the second condition in
+\clmref{StateBasedTwoIconfluenceDefs} with the added restriction that
+$e_1$ and $e_2$ are join-free.
 
 \begin{definition}
-  $T$ is \defword{\dIreducible{}} if for every pair $e_1$ and $e_2$ of
-  join-free expressions, $\Irec{e_1 \join e_2}$ implies that there exists some
-  join-free expression $e_3$ such that $\Irec{e_3}$ and $eval(e3) = eval(e1
-  \join e_2)$.
+  $O$ is \defword{\dsTIreducible{}} if for every pair $e_1$ and $e_2$ of
+  join-free $\sTIreachable{}$ expressions, $\Irec{e_1 \join e_2}$ implies that
+  there exists some join-free $\sTIreachable{}$ expression $e_3$ such that
+  $\Irec{e_3}$ and $eval(e3) = eval(e1 \join e_2)$.
 \end{definition}
 
-Intuitively, \dIreduction{} allows us to eliminate join operators by
+Intuitively, \dsTIreduction{} allows us to eliminate join operators by
 substituting $e_1 \join e_2$ (which contains 1 join) with $e_3$ (which does not
@@
 
-% \input{diconfluence_direduction.fig}
-%
-% We can also interpret \dIconfluence{} and \dIreduction{} using the graph-based
-% model of \Iconfluence{}, as illustrated in
-% \figref{diconfluence_and_direduction}. Intuitively, \dIconfluence{} is the
-% property that any two divergent join-free paths that satisfy $I$ can be joined
-% together to form a new state that satisfies $I$. \dIreduction{} is the property
-% that the join of any two divergent join-free paths that satisfy $I$ can be
-% replaced with a single join-free path that satisfies $I$. Both conditions
-% involve diamond diagrams, hence the names \dIconfluence{} and \dIreduction{}.
-
 \begin{claim}\clmlabel{DiconfluenceDireducibleImpliesIconfluence}
-  If $T$ is \dIconfluent{} and \dIreducible, then it is \Iconfluent.
+  If $O$ is \dsTIconfluent{} and \dsTIreducible, then $O$ is \sTIconfluent.
 \end{claim}
 \begin{elidableproof}
-  First, we prove that for all $e$, if $\Irec{e}$ then there exists a join-free
-  $e'$ where $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty
-  straightforward structural induction on $e$.
+  First, we prove that for all expressions $e$, if $\sTIreachable{e}$ and
+  $\Irec{e}$ then there exists a join-free $\sTIreachable{}$ $e'$ where
+  $\Irec{e'}$ and $eval(e) = eval(e')$. We proceed by a pretty straightforward
+  structural induction on $e$.
   \begin{itemize}
     \item \textbf{Case 1: $e = s_0$.}
-      Trivially, $e' = s_0.$
+      $s_0$ is join-free, so we can trivially let $e' = s_0$.
 
     \item \textbf{Case 2: $e = t(e_1)$.}
-      $\Irec{t(e_1)}$, so $\Irec{e_1}$. Thus, by the inductive hypothesis,
-      there exists a join-free $e_1'$ such that $\Irec{e_1'}$ and $eval(e_1) =
+      $\sTIreachable{t(e_1)}$ and $\Irec{t(e_1)}$, so $\sTIreachable{e_1}$ and
+      $\Irec{e_1}$. Thus, by the inductive hypothesis, there exists a join-free
+      $\sTIreachable{}$ $e_1'$ such that $\Irec{e_1'}$ and $eval(e_1) =
       eval(e_1')$. Now, let $e' = t(e_1')$ which is clearly join-free.
@@
     \item \textbf{Case 3: $e = e_1 \join e_2$.}
-      $\Irec{e_1 \join e_2}$, so $\Irec{e_1}$ and $\Irec{e_2}$. Thus, by the
-      inductive hypothesis, there exist join-free expressions $e_1'$ and $e_2'$
-      where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$,
-      and $\Irec{e_2'}$. Let $e'' = e_1' \join e_2'$
+      $\sTIreachable{e_1 \join e_2}$ and $\Irec{e_1 \join e_2}$, so
+      $\sTIreachable{e_1}$, $\sTIreachable{e_2}$, $\Irec{e_1}$ and
+      $\Irec{e_2}$. Thus, by the inductive hypothesis, there exist join-free
+      $\sTIreachable{}$ expressions $e_1'$ and $e_2'$ where $eval(e_1') =
+      eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$, and $\Irec{e_2'}$.
+      Let $e'' = e_1' \join e_2'$
       \begin{align*}
@@
       \end{align*}
-      This implies that $\Irec{e''}$. Because $T$ is \dIreducible, there exists
-      a join-free $e'$ such that $\Irec{e'}$ and $eval(e') = eval(e'') =
-      eval(e)$.
+      This implies that $\Irec{e''}$. Because $O$ is \dsTIreducible, there
+      exists a join-free $\sTIreachable{}$ $e'$ such that $\Irec{e'}$ and
+      $eval(e') = eval(e'') = eval(e)$.
   \end{itemize}
 
-  Next, consider arbitrary expressions $e_1$ and $e_2$ where $\Irec{e_1}$ and
-  $\Irec{e_2}$. To show that $T$ is \Iconfluent, we must show that $I(e_1 \join
-  e_2)$. Using our little lemma above, we know there exists join-free $e_1'$
-  and $e_2'$ where $eval(e_1') = eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') =
-  eval(e_2)$, and $\Irec{e_2'}$. Because $T$ is \dIconfluent, $I(e_1' \join
-  e_2')$, and
+  Next, consider arbitrary $\sTIreachable{}$ expressions $e_1$ and $e_2$ where
+  $\Irec{e_1}$ and $\Irec{e_2}$. To show that $T$ is \sTIconfluent, we must
+  show that $I(e_1 \join e_2)$. Using our little lemma above, we know there
+  exists join-free $\sTIreachable{}$ $e_1'$ and $e_2'$ where $eval(e_1') =
+  eval(e_1)$, $\Irec{e_1'}$, $eval(e_2') = eval(e_2)$, and $\Irec{e_2'}$.
+  Because $T$ is \dsTIconfluent, $I(e_1' \join e_2')$, and
   \[
@@
 \begin{claim}
-  Consider a state-based object $(S, s_0, \join)$, a set of transactions $T$,
-  and an invariant $I$. $T$ is \Iconfluent{} if the following criteria are met:
+  Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$. $T$ is \sTIconfluent{} if the
+  following criteria are met:
   \begin{enumerate}
     \item
-      $(S, \join)$ is a semilattice.
+      $O$ is a semilattice.
 
@@
       t_2(s))$.
+
+    \item
+      $I(s_0)$
   \end{enumerate}
@@
 
-  To show that $T$ is \Iconfluent{}, it suffices to show that $T$ is
-  \dIconfluent{} and \dIreducible{} (by
-  \clmref{DiconfluenceDireducibleImpliesIconfluence}). We begin by showing
-  that $T$ is \dIconfluent{}. Let
+  To show that $T$ is \sTIconfluent{}, it suffices to show that $T$ is
+  \dsTIconfluent{} and \dsTIreducible{} (by
+  \clmref{DiconfluenceDireducibleImpliesIconfluence}). We begin by showing that
+  $T$ is \dsTIconfluent{}. Let
   \[
@@
   \]
-  be two arbitrary join-free expressions that recursively satisfy $I$. For ease
-  of notation, let
+  be two arbitrary join-free $\sTIreachable{}$ expressions that recursively
+  satisfy $I$. For ease of notation, let
   \[
@@
   From this, it's immediate that $\Irec{\bart{n}(\baru{m}(s_0))}$, so
-  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dIconfluent{}.
+  $\Irec{e_1 \join e_2}$, which proves that $T$ is \dsTIconfluent{}.
   \begin{itemize}
     \item \textbf{Case $m = 0$.}
-      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ because
-      $\Irec{e_1}$.
+      $i = j = 0$, so $\bart{0}(\baru{0}(s_0)) = s_0$. $\Irec{s_0}$ by
+      assumption.
 
@@
       Without loss of generality, assume $i = 1$ and $j = 0$. Then,
-      \[
+      $
         \bart{1}(\baru{0}(s_0))
           = t_1(s_0)
-      \]
+      $.
       $\Irec{t_1(s_0)}$ because $\Irec{e_1}$.
@@
 
-  Next, we show that $T$ is \dIreducible{}. Again let
+  Next, we show that $T$ is \dsTIreducible{}. Again let
   \[
@@
   \]
-  be two arbitrary join-free expressions that recursively satisfy $I$. We just
-  proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also, $eval(e_1 \join e_2) =
-  eval(\bart{n}(\baru{m}(s_0)))$. $\bart{n}(\baru{m}(s_0))$ is a join-free
-  expression, so $T$ is \dIreducible{}.
+  be two arbitrary join-free $\sTIreachable{}$ expressions that recursively
+  satisfy $I$. We just proved that $\Irec{\bart{n}(\baru{m}(s_0))}$. Also,
+  $eval(e_1 \join e_2) = eval(\bart{n}(\baru{m}(s_0)))$.
+  $\bart{n}(\baru{m}(s_0))$ is a join-free $\sTIreachable{}$ expression, so $T$
+  is \dsTIreducible{}.
 \end{elidableproof}

### 2018-03-04 doc/paper_redux/sufficient_conditions.tex
Revised everything with new s_0, T, I definitions.

@@
 In this section, we present sufficient (but not always necessary) conditions
-for state-based and operation-based invariant confluence. We also explain
+for state-based and operation-based invariant-confluence. We also explain
 circumstances under which the sufficient conditions are also necessary.
@@
 \begin{definition}
-  We say a state-based object $O = (S, s_0, \join)$ is \defword{\Iclosed{}} if
-  for all states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join
+  We say a state-based object $O = (S, \join)$ is \defword{\Iclosed{}} if for
+  all states $s_1$ and $s_2$, if $I(s_1)$ and $I(s_2)$, then $I(s_1 \join
   s_2)$.
@@
   \]
-  Similarly, we say on operation-based object is \defword{\Iclosed{}} if for
-  all states $s_1$ and $s_2$ and all transactions $t$, if $I(s_1)$, $I(s_2)$,
-  and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
+  Similarly, we say on operation-based object $O = S$ is \defword{\Iclosed{}}
+  if for all states $s_1$ and $s_2$ and all transactions $t$, if $I(s_1)$,
+  $I(s_2)$, and $I(t(s_1)(s_1))$, then $I(t(s_1)(s_2))$.
   \[
@@
 \begin{claim}\clmlabel{StateBasedSufficient}
-  Given a state-based object $O = (S, s_0, \join)$, a set of transactions $T$,
-  and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$, then $(O, T)$ is
-  \Iconfluent.
+  Given a state-based object $O = (S, \join)$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$,
+  then $O$ is \sTIconfluent.
 \end{claim}
 \begin{elidableproof}
-  Let $e_1, e_2$ be expressions such that $\Irec{e_1}$ and $\Irec{e_2}$. Let
-  $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$. Then, $I(s_1)$ and $I(s_2)$, so
-  $I(s_1 \join s_2)$ which implies $I(e_1 \join e_2)$. Thus, by
-  \clmref{StateBasedTwoIconfluenceDefs}, $(O, T)$ is \Iconfluent{}.
+  Let $e_1, e_2$ be $\sTIreachable{}$ expressions such that $\Irec{e_1}$ and
+  $\Irec{e_2}$. Let $s_1 = eval(e_1)$ and $s_2 = eval(e_2)$. Then, $I(s_1)$ and
+  $I(s_2)$. Because $O$ is \Iclosed{}, $I(s_1 \join s_2)$, so $I(e_1 \join
+  e_2)$. Thus, by \clmref{StateBasedTwoIconfluenceDefs}, $O$ is
+  \sTIconfluent{}.
 \end{elidableproof}
@@
 \begin{claim}\clmlabel{OpBasedSufficient}
-  Given a operation-based object $O = (S, s_0)$, a set of transactions $T$, and
-  an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$, then $(O, T)$ is
-  \Iconfluent.
+  Given a operation-based object $O = S$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$, if $O$ is \Iclosed{} and $I(s_0)$,
+  then $O$ is \sTIconfluent.
 \end{claim}
 \begin{elidableproof}
-  Let $t \in T$ and $e_1, e_2$ be expressions such that $\Irec{e_1}$,
-  $\Irec{e_2}$, and $I(t(e_1)(e_1))$. Let $s_1 = eval(e_1)$ and $s_2 =
-  eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, and $I(t(s_1)(s_1))$, so
-  $I(t(s_1)(s_2))$. This implies $I(t(e_1)(e_2))$, so by
-  \clmref{OpBasedTwoIconfluenceDefs}, $(O, T)$ is \Iconfluent.
+  Let $t \in T$ and $e_1, e_2$ be $\sTIreachable{}$ expressions such that
+  $\Irec{e_1}$, $\Irec{e_2}$, and $I(t(e_1)(e_1))$. Let $s_1 = eval(e_1)$ and
+  $s_2 = eval(s_2)$. Then, $I(s_1)$, $I(s_2)$, $I(t(s_1)(s_1))$, so
+  $I(t(s_1)(s_2))$ because $O$ is \Iclosed. Thus, $I(t(e_1)(e_2))$, so by
+  \clmref{OpBasedTwoIconfluenceDefs}, $O$ is \sTIconfluent.
 \end{elidableproof}
 
-As a special case, \clmref{StateBasedSufficient} tells us that a set $T$ of
-transactions is \Iconfluent{} if $I$ is monotonic and $O$ is a semilattice.
+As a special case, \clmref{StateBasedSufficient} tells us that a state-based
+object $O$ is \sTIconfluent{} if $I$ is monotonic and $O$ is a semilattice.
 
 \begin{claim}\clmlabel{Monotonicity}
-  Let $O = (S, s_0, \join)$ be a state-based object where $(S, \join)$ is a
-  semilattice. Let $s \leq s' \defeq s \join s' = s'$. Let $T$ be an arbitrary
-  set of transactions. Let invariant $I$ be \defword{monotonic}: for all $s,
-  s'\in S$, if $s \leq s'$, then $I(s) \implies I(s')$. Also assume $I(s_0)$.
-  Then, $T$ is \Iconfluent{}.
+  Let $O = (S, \join)$ be a semilattice. Let $s \leq s' \defeq s \join s' =
+  s'$. Let $s_0$ be an arbitrary start state. Let $T$ be an arbitrary set of
+  transactions. Let invariant $I$ be \defword{monotonic}: for all $s, s'\in S$,
+  if $s \leq s'$, then $I(s) \implies I(s')$. Also assume $I(s_0)$. Then, $T$
+  is \sTIconfluent{}.
 \end{claim}
@@
   \leq s_1 \join s_2$, so by monotonicity, $I(s_1 \join s_2)$. Thus, by
-  \clmref{StateBasedSufficient}, $T$ is \Iconfluent{}.
+  \clmref{StateBasedSufficient}, $O$ is \sTIconfluent{}.
 \end{elidableproof}
@@
 Unfortunately, the converses of \clmref{StateBasedSufficient} and
-\clmref{OpBasedSufficient} are not true. Being \Iconfluent{} does not
+\clmref{OpBasedSufficient} are not true. Being \sTIconfluent{} does not
 necessarily imply being \Iclosed{}, as the next two examples show.
@@
 \begin{example}\examplelabel{StateBasedNotNecessary}
-  Consider the state-based object $O = (S, s_0, \join)$ where $(S, \join)$ is
-  $(\ints, \max) \times (\ints, \max)$ and $s_0 = (0, 0)$. Let $T$ consist of
-  two transactions $t_{x+1} \defeq \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} =
+  Consider the state-based object $O = (S, \join) = (\ints, \max) \times
+  (\ints, \max)$ and start state $s_0 = (0, 0)$. Let $T$ consist of two
+  transactions $t_{x+1} \defeq \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} =
   \lambda (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
-  %
-  $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
-  $(O, T)$ is \Iconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
+
+  $\setst{(x, y)}{\sTIreachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
+  $O$ is \sTIconfluent. However, let $s_1 = (-1, 1)$ and $s_2 = (1, -1)$.
   $I(s_1)$ and $I(s_2)$. However, $s_1 \join s_2 = (1, 1)$, and $\lnot I((1,
@@
 \begin{example}
-  Consider the operation-based object $O = (S, s_0)$ where $S$ is $\ints \times
-  \ints$ and $s_0 = (0, 0)$. Let $T$ consist of two transactions $t_{x+1}
-  \defeq \lambda s.\ \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} = \lambda s.\
-  \lambda (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
+  Consider the operation-based object $O = S = \ints \times \ints$ and start
+  state $s_0 = (0, 0)$. Let $T$ consist of two transactions $t_{x+1} \defeq
+  \lambda s.\ \lambda (x, y).\ (x + 1, y)$ and $t_{y-1} = \lambda s.\ \lambda
+  (x, y).\ (x, y - 1)$. Let $I = \setst{(x, y)}{xy \leq 0}$.
   %
-  $\setst{(x, y)}{\reachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
-  $(O, T)$ is \Iconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
+  $\setst{(x, y)}{\sTIreachable{(x, y)}} = \nats \times -\nats \subseteq I$, so
+  $O$ is \sTIconfluent. However, let $s_1 = (1, -1)$ and $s_2 = (0, 1)$.
   $I(s_1)$, $I(s_2)$, and $I(t_{x+1}(s_1)(s_1))$. However, $t_{x+1}(s_1)(s_2) =
@@
 Let's take a closer look at \exampleref{StateBasedNotNecessary} (illustrated in
-\figref{StateBasedNotNecessary}) to understand why being \Iconfluent{} doesn't
-imply being \Iclosed{}. Our counterexample considers points $s_1 = (-1, 1)$ and
-$s_2 = (1, -1)$. $s_2$ is reachable, but $s_1$ is not. This is not a
-coincidence! If we consider an \Iconfluent{} object and set of transactions,
+\figref{StateBasedNotNecessary}) to understand why being \sTIconfluent{}
+doesn't imply being \Iclosed{}. Our counterexample considers points $s_1 = (-1,
+1)$ and $s_2 = (1, -1)$. $s_2$ is reachable, but $s_1$ is not. This is not a
+coincidence! If we consider an \sTIconfluent{} object and set of transactions,
 then any two reachable points satisfy the invariant, and the join of the two
@@
 the invariant is if one or both points are not reachable. Thus, if we know that
-$I$ is a subset of reachable points, then being \Iconfluent{} and being
+$I$ is a subset of reachable points, then being \sTIconfluent{} and being
 \Iclosed{} are equivalent.
@@
 \begin{claim}\clmlabel{SufficientAndNecessary}
-  If every invariant satisfying state is reachable, then being \Iconfluent{}
+  If every invariant satisfying state is reachable, then being \sTIconfluent{}
   and being \Iclosed{} are equivalent.
   \[
-    I \subseteq \setst{e}{\reachable{e}} \implies
-    (\text{$O$ \Iclosed{}} \iff \text{$(O, T)$ \Iconfluent})
+    I \subseteq \setst{s}{\sTIreachable{s}} \implies
+    (I(s_0) \land \text{$O$ \Iclosed{}} \iff \text{$O$ \sTIconfluent})
   \]
@@
 \begin{elidableproof}
-  Consider a state-based object $O = (S, s_0, \join)$, a set of transactions
-  $T$, and an invariant $I$ where $I \subseteq \setst{e}{\reachable{e}}$.
+  Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$ where $I \subseteq
+  \setst{s}{\sTIreachable{s}}$.
   %
-  Being \Iclosed{} always implies being \Iconfluent{} whether or not $I
-  \subseteq \setst{e}{\reachable{e}}$. Thus, we only have to prove that if $(O,
-  T)$ is \Iconfluent{}, then $O$ is \Iclosed. If $(O, T)$ is \Iconfluent{},
-  then $I(s_0)$. Next, let $s_1$ and $s_2$ be arbitrary states such that
-  $I(s_1)$ and $I(s_2)$. Both states satisfy the invariant, so both are
-  reachable. That is, there exist expressions $e_1$ and $e_2$ where $eval(e_1)
-  = s_1$  and $eval(e_2) = s_2$. $e_1 \join e_2$ is again reachable, so since
-  $T$ is \Iconfluent{}, $I(e_1 \join e_2)$ which implies $I(s_1 \join s_2)$.
-  Thus, $O$ is \Iclosed.
+  If $I(s_0)$ and $O$ is \Iclosed{}, then $O$ is \sTIconfluent{} whether or not
+  $I \subseteq \setst{s}{\sTIreachable{s}}$. Thus, we only have to prove that
+  if $O$ is \sTIconfluent{}, then $I(s_0)$ and $O$ is \Iclosed.
+  %
+  If $O$ is \sTIconfluent{}, then $I(s_0)$. Next, let $s_1$ and $s_2$ be
+  arbitrary states such that $I(s_1)$ and $I(s_2)$. Both states satisfy the
+  invariant, so both are reachable. That is, there exist expressions $e_1$ and
+  $e_2$ where $eval(e_1) = s_1$  and $eval(e_2) = s_2$. $e_1 \join e_2$ is
+  again reachable, so since $T$ is \sTIconfluent{}, $I(s_1 \join s_2)$.  Thus,
+  $O$ is \Iclosed.
 \end{elidableproof}
@@
 Here are some more examples of transactions $T$ and invariants $I$ that are
-\Iconfluent{} but not \Iclosed{}.
+\sTIconfluent{} but not \Iclosed{}.
 
@@
 \begin{example}
-  Consider the state-based object $(S, s_0, \join)$ where $(S, \join) = (\ints,
-  \max) \times (\ints, \max)$ and $s_0 = (0, 0)$. Let $T = \set{(x, y) \mapsto
-  (x + 1, y), (x, y) \mapsto (x, y - 1)}$ and let $I((x, y)) = xy \leq 0$. The
-  reachable states are $\nats \times -\nats$, all of which satisfy $I$, so $T$
-  is \Iconfluent{}. However, let $a = (-1, 1)$ and $b = (1, -1)$. $I(a)$ and
-  $I(b)$, but $a \join b = (1, 1)$ does not satisfy $I$.
+  Consider the state-based object $O = (\powerset{\nats}, \cup)$ and start
+  state $s_0 = \set{0}$. Let $T = \setst{\lambda X.\ X \cup Y}{Y \subseteq
+  \nats}$ let $I = \setst{X}{\text{$X$ contains only odd or only even
+  elements}}$. The $\sTIreachable{}$ states are $\powerset{2\nats}$, all of
+  which satisfy $I$, so $O$ is \sTIconfluent{}.  However, let $A = \set{0}$ and
+  $Y = \set{1}$. $I(A)$ and $I(B)$, but $A \join B = \set{0, 1}$ does not
+  satisfy $I$.
 \end{example}
@@
 \begin{example}
-  Consider the state-based object $(\powerset{\nats}, \set{0}, \cup)$. Let $T =
-  \setst{X \mapsto X \cup Y}{Y \subseteq \nats}$ let $I(X) = \exists k \in
-  \set{0, 1}.\ \forall x \in X.\ x \equiv k \mod 2$. That is $I(X)$ if $X$
-  contains only odd or only even elements. The reachable states are
-  $\powerset{2\nats}$, all of which satisfy $I$, so $T$ is \Iconfluent{}.
-  However, let $A = \set{0}$ and $Y = \set{1}$. $I(A)$ and $I(B)$, but $A \join
-  B = \set{0, 1}$ does not satisfy $I$.
-\end{example}
-
-\begin{example}
-  Consider a state-based object consisting of two integer 2P-sets $(A_1, R_1)$ and
-  $(A_2, R_2)$ with initial state $(\set{1, 3}, \emptyset{})$ and $(\set{1, 2,
-  3, 4}, \emptyset{})$. Let $T$ consist of all transactions that either remove
-  an element from the first 2P-set (i.e.\ add something to $R_1$) or add
+  Consider a state-based object consisting of two integer 2P-sets $(A_1, R_1)$
+  and $(A_2, R_2)$ with initial state $(\set{1, 3}, \emptyset{})$ and $(\set{1,
+  2, 3, 4}, \emptyset{})$. Let $T$ consist of all transactions that either
+  remove an element from the first 2P-set (i.e.\ add something to $R_1$) or add
   something to the second 2P-set (i.e.\ add something to $A_2$). Let $I$ be
@@
   %
-  $T$ is \Iconfluent{}. However, let $s_1 = (\set{1}, \set{2}, \set{1},
+  $O$ is \sTIconfluent{}. However, let $s_1 = (\set{1}, \set{2}, \set{1},
   \set{2})$ and $s_2 = (\set{2}, \set{1}, \set{2}, \set{1})$. $I(s_1)$ and

### 2018-03-04 doc/paper_redux/interactive_decision_procedure.tex
Described bad user input scenario.

@@
 \section{Interactive Invariant-Confluence Decision Procedure}
+\newcommand{\IsInvConfluent}{\textsc{IsInvConfluent}}
+\newcommand{\IsIclosed}{\textsc{IsIclosed}}
+\newcommand{\Helper}{\textsc{Helper}}
+
 Imagine we have an \Iclosed{} decision procedure, and let's revisit
@@
   \begin{algorithmic}
-    \Function{IsIconfluent}{$O$, $s_0$, $T$, $I$}
+    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
       \State
-        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
+        \Return $I(s_0)$ and
+                \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
     \EndFunction
@@
     \comment{$NR$ is a set of $\sTIunreachable{}$ states.}
+    \comment{$I(s_0)$ is a precondition.}
     \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
-      \If {$\lnot(I - NR)(s_0)$}
-        \State \Return no
-      \EndIf
       \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
-      \If {closed $=$ unknown}
-        \State \Return unknown
-      \ElsIf {closed $=$ yes}
-        \State \Return{yes}
+      \If {closed}
+        \State \Return true
       \Else
         \State Augment $R$ and $NR$ with random search and user input
-        \State Make sure $s_1$ and $s_2$ are included in $R$ or $NR$
         \If{$s_1, s_2 \in R$}
-          \State \Return no
+          \State \Return false
         \Else
-          \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
+          \State \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
         \EndIf
@@
 
-\todo{Rename IsIconfluent to IsSTIConfluent}
-\newcommand{\IsIconfluent}{\textsc{IsIconfluent}}
-\newcommand{\IsIclosed}{\textsc{IsIclosed}}
-\newcommand{\Helper}{\textsc{Helper}}
+\input{figs/decision_procedure_example.fig}
 
@@
 Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
-invariant $I$, we call \IsIconfluent$(O, s_0, T, I)$ to determine if $O$ is
+invariant $I$\footnote{%
+  Formally, $I$ is set of states (e.g. $\setst{(x, y)}{x \in \ints, y \in
+  \ints, xy \leq 0}$), but an actual implementation of \IsInvConfluent{} would
+  likely take in $I$ as a logical formula (e.g. $xy \leq 0$). From now on,
+  whenever we mention a (possibly infinite) set of states, remember that we can
+  always represent the set as a finite formula.
+}, we call \IsInvConfluent$(O, s_0, T, I)$ to determine if $O$ is
 \sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
@@
 
-\IsIconfluent{} assumes access to an \Iclosed{} decision procedure
-\IsIclosed$(O, I)$ which returns whether $O$ is \Iclosed{} (yes, no, or
-unknown). If it returns no, it also returns a counterexample $s_1$ and $s_2$
-such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$. Constructing an
-\Iclosed{} decision procedure is hard (undecidable, I think), but we can do
-pretty well using an SMT solver like Z3. There is also some existing research
-on how to built smart \Iclosed{} decision procedures that we can
-leverage~\cite{li2014automating}.
+\IsInvConfluent{} assumes access to two helper decision procedures:
+\begin{enumerate}
+  \item
+    A decision procedure to determine $I(s_0)$.
 
-\IsIconfluent{} takes in $O$, $s_0$, $T$, and $I$. Formally, $I$ is set of
-states (e.g. $\setst{(x, y)}{x \in \ints, y \in \ints, xy \leq 0}$), but an
-actual implementation of \IsIconfluent{} would likely take in $I$ as a logical
-formula (e.g. $xy \leq 0$). From now on, whenever we mention a (possibly
-infinite) set of states, remember that we can always represent the set as a
-finite formula.
+  \item
+    An \Iclosed{} decision procedure \IsIclosed$(O, I)$ which returns a triple
+    (closed, $s_1$, $s_2$). closed is a boolean indicating whether $O$ is
+    \Iclosed{}. If closed is false, then $s_1$ and $s_2$ are a counterexample
+    such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$.
+\end{enumerate}
 
-\IsIconfluent{} calls a helper function \Helper{} which---in addition to $O$,
-$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
-and a set $NR$ of $\sTIunreachable{}$ states. Like \IsIconfluent, \Helper$(O,
-s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
-$NR$ are correct). Intuitively, $NR$ helps remove unreachable states from $I$
-moving us closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.
+Constructing these two helper decision procedures is hard (undecidable, I
+think), but we can do pretty well using an SMT solver like Z3. There is also
+some existing research on how to built smart \Iclosed{} decision procedures
+that we can leverage~\cite{li2014automating}. In practice either decision an
+return unknown instead of true or false. In this case, \IsInvConfluent{} returns
+unknown.
 
-\Helper{} first checks to see if $(I - NR)(s_0)$. If it is not, then \Helper{}
-returns no. If $NR$ only contains unreachable states, then it doesn't contain
-$s_0$, so $s_0 \in I - NR \iff s_0 \in I$. Thus, if $\lnot (I - NR)(s_0)$, then
-$\lnot I(s_0)$, so $O$ is not \sTIconfluent.
+\IsInvConfluent{} first checks that $s_0$ satisfies $I$. If it doesn't, then
+\IsInvConfluent{} returns false. If it does then, \IsInvConfluent{} calls a
+helper function \Helper{} which---in addition to $O$, $s_0$, $T$, and
+$I$---takes as arguments a set $R$ of $\sTIreachable{}$ states and a set $NR$
+of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O, s_0, T, I, R,
+NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and $NR$ are
+correct). Intuitively, $NR$ helps remove unreachable states from $I$ moving us
+closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.
 
-Next, \Helper{} checks to see if $O$ is \iclosed{$(I - NR)$}. There are three
+First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}. There are two
 cases to consider.
@@
   \item
-    If \IsIclosed{} gets stuck and doesn't know whether or not $O$ is
-    \iclosed{$(I - NR)$}, then \Helper{} is stuck too and is forced to return
-    unknown.
-
-  \item
-    If \IsIclosed{} determines that $O$ is \iclosed{$(I - NR)$}, then by
+    If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
     \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
@@
     $.
-    Thus, $O$ \sTIconfluent{}, so \Helper{} return yes.
+    Thus, $O$ \sTIconfluent{}, so \Helper{} return true.
 
   \item
-    If \IsIclosed{} determines that $O$ is not \iclosed{$(I - NR)$}, then we
+    If \IsIclosed{} determines that $O$ is not \iclosed{(I - NR)}, then we
     have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
@@
     $R$.
-
+    %
     Second, we can prompt the user to tell us whether or not the states are
@@
     and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
-    augmenting $NR$. Taking \exampleref{StateBasedNotNecessary} as an example,
-    the user might look through the existing $\sTIreachable{}$ and
-    $\sTIunreachable{}$ states and have an aha moment realizing that every
-    point with a negative $x$-coordinate is $\sTIunreachable{}$. They can use
-    this knowledge to add all points with negative $x$-coordinates to $NR$.
+    augmenting $R$ and $NR$ arbitrarily (see \figref{DecisionProcedureExample}
+    for an example). In this step, we also make sure that $s_0 \notin NR$ since
+    we know that $s_0$ is $\sTIreachable{}$.
 
@@
     \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
-    so \Helper{} returns no. Otherwise, one of $s_1$ and $s_2$ is
-    $\sTIunreachable{}$. Now, we recurse.
+    so \Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
+    $\sTIunreachable{}$, so we recurse.
 \end{itemize}
 
-\input{figs/decision_procedure_example.fig}
+\algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
+procedure. It requires that a user classify states as reachable or unreachable.
+What happens if a user incorrectly classifies a state? There's two possible
+errors that can be made:
+\begin{itemize}
+  \item \textbf{A user can label an unreachable state as reachable.}
+    In this case, the $s_1, s_2 \in R$ check in \Helper{} might be erroneously
+    met. In turn, \IsInvConfluent{} might erroneously return no. This is
+    inconvenient, but not the end of the world. Moreover, we can modify our
+    decision procedure so that \Helper{} requires that whenever a user labels a
+    state $s$ as $\sTIreachable{}$, they must also provide an $\sTIreachable{}$
+    expression $e$ where $eval(e) = s$.
 
-% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
-% procedure. It requires that a user classify states as reachable or unreachable.
-% What happens if a user incorrectly classifies a state? Given a counterexample
-% $s_1, s_2$, if a user incorrectly says that one of the unreachable points is
-% reachable, then the $s_1, s_2 \in R$ check in \Helper{} might be
-% erroneously met. In turn, \IsIconfluent{} might erroneously return no. This is
-% inconvenient, but not the end of the world.
-% Moreover, we can modify our decision procedure so that \Helper{}
-% user provide a \emph{proof} of reachability whenever they label as: a series of transactions and joins that result in the
+  \item \textbf{A user can label a reachable state as unreachable.}
+    In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is
+    not \sTIconfluent. Thus, a user might falsely believe their system to be
+    \sTIconfluent{} even though it isn't, and eventually their system might
+    enter a state which violates the invariant.
+
+    We can mitigate this in two ways. First, we can aggresively expand $R$. If
+    a user ever labels a state $s$ as unreachable, but $s \in R$, we can notify
+    the user of their mistake. Second, \Helper{} returns true when $O$ is
+    \iclosed{(I - NR)} for some $NR$, so $O$ is \sticonfluent{s_0}{T}{(I-R)}
+    (even though it might not be \sTIconfluent{}). Thus, when a user deploys
+    their system, they can deploy with the invariant $I - NR$ instead of $NR$.
+    If $NR$ is correctly populated, then deploying with $I - NR$ has no effect
+    on the system. If $NR$ is incorrectly populated, then some
+    $\sTIreachable{}$ states are artificially made unreachable, but the system
+    is still guaranteed to never enter a state which violates $I$.
+\end{itemize}

### 2018-03-05 doc/paper_redux/stratified_iconfluence.tex
Added section on stratified invariant confluence.

\section{Stratified Invariant-Confluence}
The interactive decision procedure in \algoref{InteractiveDecisionProcedure}
and the sufficient conditions in \clmref{NewStateBasedSufficient} give us two
ways to detect when an object $O$ is invariant-confluent. But what if it's not?
%
Let's revisit \exampleref{StateBasedNotNecessary} but with start state $s_0 =
(-2, 2)$. Now, $O$ is not \sTIconfluent{}. The set of $\sTIreachable{}$ points
(illustrated in \figref{FixingIconfluenceReachable}) is not a subset of $I$
(illustrated in \figref{FixingIconfluenceInvariant}).

This is unfortunate, but if we're determined to make $O$ \sTIconfluent{}? Well,
then we have to adjust the set of $\sTIreachable{}$ points so that they are a
subset of $I$. We can go about this in three ways:
\begin{itemize}
  \item \textbf{Adjust $s_0$.}
    $O$ is not \sticonfluent{(-2, 2)}{T}{I}, but it is \sticonfluent{(0,
    0)}{T}{I}. Thus, we can change our start state to become
    invariant-confluent. This is illustrated in \figref{FixingIconfluenceS}.

  \item \textbf{Adjust $T$.}
    $O$ is not \sticonfluent{s_0}{\set{t_{x+1}, t_{y-1}}}{I}, but it is
    \sticonfluent{s_0}{\set{t_{x+1}}}{I}. Thus, we can remove some transactions
    to become invariant-confluent. This is illustrated in
    \figref{FixingIconfluenceT}. Instead of outright forbidding some
    transactions, we can instead use locking to limit the concurrent execution
    of certain sets of transactions~\cite{balegas2015putting,
    gotsman2016cause}.

  \item \textbf{Adjust $I$.}
    $O$ is not \sticonfluent{s_0}{T}{\setst{(x, y)}{xy \leq 0}} but it is
    \sticonfluent{s_0}{T}{\setst{(x, y)}{x \leq 0, y \geq 0}}. Thus, we can
    remove some states from $I$ to become invariant-confluent. This is
    illustrated in \figref{FixingIconfluenceI}. We could also add some states
    to $I$ and become invariant-confluent, but it's likely that states are
    excluded from $I$ for good reason.
\end{itemize}

{\input{figs/fixing_iconfluence.fig}}

Unfortunately, none of these three adjustments are completely satisfactory.
There may only be one natural start state, and artificially making some
invariant-satisfying states unreachable---either by eliminating transactions or
restricting the invariant---is most likely a bit of a dirty hack. Moreover,
locking can sometimes be overly restrictive. For example, we can make the
example above invariant-confluent by requiring that both transactions $t_{x+1}$
and $t_{y-1}$ acquire a lock. However, once all processors have a state $(x,
y)$ where $x \geq 0$ and $y \leq 0$, the locks are no longer necessary.

Ideally, we would maintain the same start state, use locking to restrict
concurrently executing transactions as little as possible, and leave all
invariant-satisfying reachable states reachable. Can we achieve this?  Yes, at
least kind of. We'll do so using something we call \defword{stratified
invariant-confluence}. The main idea is that we'll partition the set of
invariant-satisfying states. Depending on the state we're in, we'll either
limit which transactions can execute concurrently, restrict the invariant, or
both.

Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
transactions $T$, and an invariant $I$ deployed on $n$ processors $p_1, \ldots,
p_n$. Further assume that $O$ is a semilattice.
%
With stratified invariant-confluence, we construct an ordered sequence $P =
(I_1, T_1), \ldots, (I_n, T_n)$ of $n$ pairs $(I_i, T_i)$ with the following
properties:
\begin{itemize}
  \item
    $I_i \subseteq I$ for $1 \leq i \leq n$

  \item
    $T_i \subseteq T$ for $1 \leq i \leq n$

  \item
    $I = \cup_{i=1}^n I_i$

  \item
    For all $(I_i, T_i) \in P$ and all states $s \in I_i$, $O$ is
    \sticonfluent{s}{I_i}{T_i}.
\end{itemize}
All processors store a copy of $P$. Moreover, at any given point in time, a
processor $p_i$ will execute with one pair $(I_j, T_j) \in P$ \defword{active}.
We'll describe what this means in just a moment. Initially, every processor
$p_i$ is state $s_0$. $p_i$ finds the first pair $(I_j, T_j) \in P$ such that
$I_j(s_0)$ and makes this $(I_j, T_j)$ active. As usual, a processor can
perform one of two actions: execute a transaction or perform a merge.
\begin{itemize}
  \item
    Merging proceeds as usual. When a process $p_i$ with state $s_i$ receives a
    state $s_j$ from $p_j$, it transitions from $s_i$ to $s_i \join p_j$.

  \item
    Transaction processing is a bit different. Imagine a processor $p_i$ in
    state $s_i$ with pair $(I_j, T_j)$ active wants to execute transaction $t$.
    %
    If $t \in T_j$, then $p_i$ executes $t_i$ so long as doing so maintains
    $I_j$.
    %
    If $t \notin T_j$, then $p_i$ initiates a round of coordination among the
    processors. All processors adopt state state $s = s_1 \join \cdots \join
    s_n$. Then, all processors execute $t$. If $t(s)$ does not satisfy $I$,
    then the processors remain in state $s$ and continue normal processing.  If
    $t(s)$ does satisfy $I$, then the processors enter state $t(s)$ and choose
    pair $(I_k, T_k)$ to active where $(I_k, T_k)$ is the first pair in $P$
    where $I_k(t(s))$.
\end{itemize}

An example of stratified invariant-confluence for the example discussed above
is given in \figref{StratifiedIconfluenceExample}.

{\input{figs/stratified_iconfluence_example.fig}}

\todo{Describe why this scheme ensures ``Iconfluence''}.

\todo{%
  Give name to things like $P$. Call it something like an $(I, T)$-partition.
}
\todo{Can we modify the decision procedure to help build a stratification?}

### 2018-03-05 doc/paper_redux/stratified_iconfluence.tex
Added more to stratified invariant-confluence.

@@
     pair $(I_k, T_k)$ to active where $(I_k, T_k)$ is the first pair in $P$
-    where $I_k(t(s))$.
+    where $I_k(t(s))$ (which is guaranteed to exist since $\cup_i I_i = I$).
 \end{itemize}
@@
 
-\todo{Describe why this scheme ensures ``Iconfluence''}.
+Stratified invariant-confluence, like invariant-confluence, guarantees that no
+replica ever enters a state that violates the invariant $I$. How does it
+guarantee this?
+%
+Initially, all processors begin in the same state $s$ and with the same active
+pair $(I_j, T_j)$, and this invariant is maintained whenever a global
+coordination is run. During normal non-coordinated operation, because $O$ is
+\sticonfluent{s}{T_j}{I_j}, $\setst{s'}{\stireachable{s}{T_j}{I_j}{s'}}
+\subseteq I_j \subseteq I$. Thus, processors stay within $I_j$ which itself
+lies within $I$. During a coordinated action, processors transition to state $s
+= s_1 \join \cdots \join s_n \in I_j$ and then possibly transition into a new
+state $t(s)$ with new active pair $(I_k, T_k)$.
+
+A couple things to note:
+\begin{itemize}
+  \item
+    If $O$ is \sTIconfluent{}, then we can let $P = (I, T)$.
+
+  \item
+    Let $P = (I, T')$ where $T' \subseteq T$. This is a special case of the
+    token-based approach in~\cite{gotsman2016cause} where every transaction in
+    $T - T'$ acquires a single token $\tau$ and every transaction in $T$
+    doesn't acquire any token. The token-based approach
+    in~\cite{gotsman2016cause} is not completely subsumed by our stratified
+    invariant-confluence. For example, stratified invariant-confluence cannot
+    emulate multi-token systems. Conversely, stratified invariant-confluence is
+    not subsumed by the token-based approach which cannot stratify $I$. Notably
+    though, the token-based approach can adjust what tokens a transaction
+    acquires based on its state, but this is not used in the paper.
+\end{itemize}
 

### 2018-03-11 doc/paper_redux/stratified_iconfluence.tex
Added another small TODO.
Modified stratified iconfluence execution.

@@
 
-This is unfortunate, but if we're determined to make $O$ \sTIconfluent{}? Well,
-then we have to adjust the set of $\sTIreachable{}$ points so that they are a
-subset of $I$. We can go about this in three ways:
+This is unfortunate, but what if we're determined to make $O$ \sTIconfluent{}?
+Well, then we have to adjust the set of $\sTIreachable{}$ points so that they
+are a subset of $I$. We can go about this in three ways:
 \begin{itemize}
@@
 %
-With stratified invariant-confluence, we construct an ordered sequence $P =
-(I_1, T_1), \ldots, (I_n, T_n)$ of $n$ pairs $(I_i, T_i)$ with the following
-properties:
+With stratified invariant-confluence, we construct an \ITpartition{}: an
+ordered sequence $P = (I_1, T_1), \ldots, (I_n, T_n)$ of $n$ pairs $(I_i, T_i)$
+with the following properties:
 \begin{itemize}
@@
   \item
-    $I = \cup_{i=1}^n I_i$
+    $I = \cup_{i=1}^n I_i$ (note that the $I_i$'s don't have to be disjoint)
 
@@
 \end{itemize}
-All processors store a copy of $P$. Moreover, at any given point in time, a
-processor $p_i$ will execute with one pair $(I_j, T_j) \in P$ \defword{active}.
-We'll describe what this means in just a moment. Initially, every processor
-$p_i$ is state $s_0$. $p_i$ finds the first pair $(I_j, T_j) \in P$ such that
-$I_j(s_0)$ and makes this $(I_j, T_j)$ active. As usual, a processor can
-perform one of two actions: execute a transaction or perform a merge.
+All processors store a copy of the \ITpartition{} $P$. Moreover, at any given
+point in time, a processor $p_i$ will execute with one pair $(I_j, T_j) \in P$
+\defword{active}. We'll describe what this means in just a moment. Initially,
+every processor $p_i$ is state $s_0$. $p_i$ finds the first pair $(I_j, T_j)
+\in P$ such that $I_j(s_0)$ and makes this $(I_j, T_j)$ active. As usual, a
+processor can perform one of two actions: execute a transaction or perform a
+merge.
 \begin{itemize}
@@
     state $s_i$ with pair $(I_j, T_j)$ active wants to execute transaction $t$.
-    %
-    If $t \in T_j$, then $p_i$ executes $t_i$ so long as doing so maintains
-    $I_j$.
-    %
-    If $t \notin T_j$, then $p_i$ initiates a round of coordination among the
-    processors. All processors adopt state state $s = s_1 \join \cdots \join
-    s_n$. Then, all processors execute $t$. If $t(s)$ does not satisfy $I$,
-    then the processors remain in state $s$ and continue normal processing.  If
-    $t(s)$ does satisfy $I$, then the processors enter state $t(s)$ and choose
-    pair $(I_k, T_k)$ to active where $(I_k, T_k)$ is the first pair in $P$
-    where $I_k(t(s))$ (which is guaranteed to exist since $\cup_i I_i = I$).
+
+    If $t \in T_j$, then $p_i$ executes $t$ to transition from state $s_i$ to
+    state $t(s_i)$. If $I_j(t(s_i))$, then $p_i$ returns to the client
+    immediately. If $\lnot I(t(s_i))$, then $p_i$ aborts the transaction and
+    returns to the client immediately. If $\lnot I_j(t(s_i))$ and $I(t(s_i))$,
+    then $p_i$ initiates a round of global coordination among the processors,
+    as described below.  If $t \notin T_j$, then $p_i$ also initiates a round
+    of coordination.
+
+    During a global coordination, all processors adopt state state $s = s_1
+    \join \cdots \join s_n$. Then, all processors execute $t$. If $t(s)$ does
+    not satisfy $I$, then the processors remain in state $s$ and continue
+    normal processing.  If $t(s)$ does satisfy $I$, then the processors enter
+    state $t(s)$ and choose pair $(I_k, T_k)$ to be active where $(I_k, T_k)$
+    is the first pair in $P$ where $I_k(t(s))$ (which is guaranteed to exist
+    since $\cup_i I_i = I$).
 \end{itemize}
@@
 \todo{%
+  Give more examples where stratified invariant-confluence is helpful.
+}
+\todo{%
   Give name to things like $P$. Call it something like an $(I, T)$-partition.

### 2018-03-12 doc/paper_redux/evaluation.tex
Added thoughts on evaluation.

\section{Evaluation}
\begin{itemize}
  \item
    How good is the interactive decision procedure? How good are the new
    sufficient conditions? For this, we'll need a set of case studies to run
    our algorithms on. We can borrow case studies from related works.
  \item
    How fast is the interactive decision procedure? How fast are the new
    sufficient conditions? We can use the same case studies.
  \item
    How good is stratified invariant confluence? Find another set of case
    studies.
  \item
    How fast is Lucy for normal invariant confluence? Choose an application and
    measure the throughput or latency. There's not too many knobs to tune.
  \item
    How fast is Lucy for stratified invariant confluence? Choose an application
    and measure the throughput or latency. Vary the number of transactions that
    take us out of the region, the number of transactions outside the region,
    etc. Compare this to something like a primary-backup system where
    operations are all run one at a time.
\end{itemize}

### 2018-03-12 doc/paper_redux/sufficient_conditions.tex
Wrote up student example in more detail.

@@
 \begin{example}
-  \todo{Write this up in full.}
-  Imagine we are building a system to group students together. We have a relation
-  of (student id, grade) and a map (group id -> set of student ids). We want to
-  enforce the invariant that every student in every group is in the same grade
-  and that all groups are non-empty. If we start off with an initially empty
-  mapping, then there are conflicts. If we start off with at least one student
-  mapped into each group, then there are not any conflicts.
+  Imagine that we have a set $P$ of students ($P$ is for pupils) and that we
+  are trying to assign students into groups. We have a state based object $O =
+  (S, \join)$ where each state $s \in S$ in a dictionary mapping a group id $1,
+  \ldots, n$ to some subset of $P$. For example, if $P = \set{a, b, c}$, then
+  $s$ might be $\set{1 \mapsto \set{a, b}, 2 \mapsto \set{c}}$. We join two
+  dictionaries by taking a group-wise union. That is, $(s_1 \join s_2)(i) =
+  s_1(i) \cup s_2(i)$.
+
+  Furthermore assume each student $p$ has a year $year(p)$ (e.g. 5th grade, 4th
+  grade). Our invariant $I$ is that every group of students is in the same
+  year. That is, for every state $s$ and every $1 \leq i \leq n$, every student
+  in $s(i)$ belongs to the same year.
+  %
+  Our set $T$ of transactions includes transactions which allow us to add any
+  number of students to any number of groups.
+  %
+  Let $P = \set{a, b, \ldots, z}$ and $n = 3$. Our start state is $s_0 = \set{1
+  \mapsto \set{a}, 2 \mapsto \set{b}, 3 \mapsto \set{c}}$.
+
+  Some thought shows that O is \sTIconfluent{}. However, $O$ is not \Iclosed{}. Let
+  \[
+    s_1 = \set{1 \mapsto \set{x}, 2 \mapsto \emptyset, 3 \mapsto \emptyset}
+    \qquad
+    s_2 = \set{1 \mapsto \set{y}, 2 \mapsto \emptyset, 3 \mapsto \emptyset}
+  \]
+  where $year(x) \neq year(y)$. $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join
+  s_2)$.
 \end{example}

### 2018-03-12 doc/paper_redux/stratified_iconfluence.tex
Added some TODOs for various examples.

@@
 \todo{Can we modify the decision procedure to help build a stratification?}
+
+\begin{example}
+  \todo{%
+    Bank account example from~\cite{gotsman2016cause}. This might be hard to
+    actually implement since the bank account merge is hard to do with a
+    state-based object. This example only acquires a single lock.
+  }
+\end{example}
+
+\begin{example}
+  \todo{%
+    Auction example from~\cite{gotsman2016cause}. I think we could automate
+    this one. This one acquires a single lock. We don't need the lock once
+    we're finalized.
+  }
+\end{example}
+
+\begin{example}
+  \todo{%
+    Courseware application from~\cite{gotsman2016cause}. When course is
+    non-empty, we can allow any operations at all. When a course is empty, we
+    cannot allow concurrent operations on it. This one should hopefully be easy
+    to automate.
+  }
+\end{example}
+
+\begin{example}
+  \todo{%
+    Ad counter from~\cite{balegas2015putting}. This might be hard like the bank
+    account balance. Adding numbers might be tricky in a state-based approach.
+  }
+\end{example}
+
+\begin{example}
+  \todo{%
+    Tournament from~\cite{balegas2015putting}. This might be hard. Pretty
+    complicated example with set cardinality constraints.
+  }
+\end{example}

### 2018-03-26 doc/ccn/diagrams.tex
Renaming various docs.

% Implication.
\tikzstyle{impl}=[-implies, double distance=3pt]

% Transaction.
\tikzstyle{txn}=[ultra thick, ->]

% Dashed transaction.
\tikzstyle{dtxn}=[ultra thick, ->, dashed]

% G transaction.
\tikzstyle{gtxn}=[ultra thick, ->, decorate, decoration={zigzag}]

% Dashed G transaction.
\tikzstyle{dgtxn}=[ultra thick, ->, densely dashed, decorate, decoration={zigzag}]

% Invariant node.
\newcommand{\inode}[3][]{
  \node[draw, shape=circle, minimum width=0.6cm] (#2) at (#3) {};
  \node[fill, shape=circle, #1] () at (#3) {};
}

% Dashed node.
\newcommand{\dnode}[3][]{
  \node[draw, dashed, shape=circle, minimum width=0.6cm] (#2) at (#3) {};
  \node[fill, shape=circle, #1] () at (#3) {};
}

% Normal node.
\newcommand{\nnode}[3][]{
  \node[fill, shape=circle, #1] (#2) at (#3) {};
}

### 2018-03-26 doc/notes_v1/abstract.tex
Renaming various docs.

\section{Abstract}
\emph{Strong consistency} allows programmers to ignore many of the complexities
of a distributed system, treating it as if it were running on a single machine.
However, enforcing strong consistency requires \emph{coordination}, and
coordination leads to unavailability (at worst) or
increased latency (at best)~\cite{gilbert2002brewer, abadi2012consistency}. Is
there some way to globally enforce an application's invariants without needing
to coordinate?

Yes! Well, sometimes.

Bailis et al. developed the notion of invariant confluence (\iconfluence{}) as
a necessary and sufficient condition for coordination
freedom~\cite{bailis2014coordination}. Intuitively, if a set of transactions is
invariant confluent with respect an invariant, then the invariant is guaranteed
to hold even when the transactions are run without coordination.  Technically,
a set of transactions is \iconfluent{} with respect to some invariant $I$ if
merging database states that satisfy $I$ always results in a database state
that also satisfies $I$.  Bailis characterized many common invariants (e.g.
uniqueness constraints, foreign key constraints), showing which could be
maintained without coordination. However, this characterization required
hand-written proofs. This research project aims to expand on Bailis' research
by providing a set of CRDTs~\cite{shapiro2011comprehensive,
shapiro2011conflict}, a language of invariants, and a language of transactions
that facilitates the ability for \iconfluence{} to be determined
algorithmically. An \iconfluence{} decision procedure would allow programmers
to guarantee that their programs can execute correctly without coordination.

### 2018-03-26 doc/notes_v1/alternatives.tex
Renaming various docs.

--- 2016-11-26 doc/counters.tex
@@
-\section{Counters}
-As a first step towards accomplishing our research vision, let's begin with a
-\emph{simple} CRDT with \emph{simple} operations and a \emph{simple} invariant
-language: PN-Counters and linear equations and inequalities.
-
-\subsection{Overview}\label{sec:counter-overview}
-Recall that a \emph{G-Counter} CRDT is an increment-only counter
-\cite{shapiro2011comprehensive}. A state-based G-Counter distributed across $n$
-machines $1, 2, \ldots, n$ can be implemented as an $n$-tuple $(x_1, x_2,
-\ldots, x_n)$ of natural numbers. The \emph{increment($k$) method} at machine
-$i$ increments $x_i$ by $k$; the \emph{query method} returns the sum of the
-tuple $\sum_{i=1}^n x_i$; and the \emph{merge method} computes the pairwise max
-of two $n$-tuples.
-
-A \emph{PN-Counter} CRDT is a counter which supports increments and decrements
-and can be implemented as a pair $(p, n)$ of two G-Counters
-\cite{shapiro2011comprehensive}. increment($x$) increments $p$ by $x$;
-decrement($x$) increments $n$ by $x$; query returns $p - n$; and merge performs
-a pairwise merge of $p$ and $n$.
-
-Our \emph{invariant language} is formed from \emph{linear equations} (e.g. $x =
-0$, $2x = y$, $-x + 42y - 2z = 16$) and \emph{inequalities} (e.g. $x > 0$, $2x
-\leq y$, $-x + 42y - 2z \neq 16$) connected with the boolean connectives
-\emph{and} ($\land$), \emph{or} ($\lor$), and \emph{not} ($\lnot$) (e.g.
-$\lnot((\lnot(x < 0) \lor 2x = y) \land -x + 42y - 2z \geq 16)$).
-
-A \emph{transaction} chooses a subset of variables and decrements or increments
-each by a constant amount. For example, a transaction may increment $x$ by 1,
-decrement $y$ by 42, and increment $z$ by 100.
-
-Given a set of transactions and an invariant, we want to determine whether the
-transactions are \iconfluent{} with respect to the invariant. For example, the
-transaction that \emph{increments} $x$ by 1 is \iconfluent{} with respect to
-the invariant $x > 0$, but the transaction that \emph{decrements} $x$ by 1 is
-not, as shown in \figref{decrement}.
-
-\begin{figure}[h]
-  \centering
-  \begin{tikzpicture}
-    \node[draw] (start)  at (2, 4) {$x=2$};
-    \node[draw] (left)   at (0, 2) {$x=1$};
-    \node[draw] (right)  at (4, 2) {$x=1$};
-    \node[draw] (merged) at (2, 0) {$x=0$};
-
-    \draw[thick, ->] (start) to node[left]  {$x := x - 1$} (left);
-    \draw[thick, ->] (start) to node[right] {$x := x - 1$} (right);
-    \draw[thick, ->] (left)  to (merged);
-    \draw[thick, ->] (right) to (merged);
-  \end{tikzpicture}
-  \caption{
-    A counterexample showing transaction $x := x - 1$ is not invariant
-    confluent with respect to the invariant $x > 0$. The top, left, and right
-    states satisfy the invariant, but the bottom state does not.
-  }
-  \label{fig:decrement}
-\end{figure}
-
-\subsection{Formalizing the Question}\label{sec:counter-question}
-Our invariant language is defined by the grammar in \figref{invariant-grammar}
-which includes linear equations and inequalities over variables in a finite set
-$\var$ and integer constants in $\ints$.
-
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\atom}{\textsf{atom}}
-  \newcommand{\aexp}{\textsf{aexp}}
-  \newcommand{\bexp}{\textsf{bexpr}}
-  \begin{gather*}
-    \begin{array}{rrll}
-      \atom & ::= & k & \text{\emph{constants}} \\
-            & |   & x & \text{\emph{variables}} \\
-      &&& \\
-      \aexp  & ::= & \atom         & \text{\emph{atom}} \\
-             & |   & -\atom        & \text{\emph{negation}} \\
-             & |   & \aexp + \aexp & \text{\emph{addition}} \\
-             & |   & \aexp - \aexp & \text{\emph{subtraction}} \\
-      &&& \\
-      \bexp  & ::= & \aexp \leq \aexp  & \text{\emph{inequality}} \\
-             & |   & \lnot \bexp       & \text{\emph{negation}} \\
-             & |   & \bexp \land \bexp & \text{\emph{conjunction}} \\
-             & |   & \bexp \lor \bexp  & \text{\emph{disjunction}} \\
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for our linear equation and linear inequality invariant language.
-    Note that the comparison operators $=$, $\neq$, $<$, $>$, and $\geq$ are
-    not included because they can be defined in terms of the existing
-    operators.
-  }
-  \label{fig:invariant-grammar}
-\end{figure}
-
-A \emph{database state} is a total function $D: \var \to \ints$.  We say a
-database state $D$ \emph{satisfies invariant} $I$, denoted $I(D)$, if $I$
-evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
-
-A \emph{transaction} is a partial function $t: \var \rightharpoonup \ints$.
-\emph{Applying a transaction} $t$ to database state $D$, denoted by $D \circ
-t$, produces a new database state defined as follows:
-\[
-  (D \circ t)(x) \defeq \begin{cases*}
-    D(x) + t(x) & $x \in \dom{}(t)$ \\
-    D(x)        & otherwise
-  \end{cases*}
-\]
-Note that transaction application is commutative. That is for all database
-states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
-
-A transaction chain $C$ created from a set of transactions $T$ is a sequence of
-transactions $t_1, \ldots, t_n$ where $t_i \in T$ for $1 \leq i \leq n$. Let $D
-\circ C$ be syntactic sugar for $D \circ t_1 \circ \cdots \circ t_n$. We say
-$C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
-
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-
-\subsection{A Geometric Interpretation}\label{sec:counter-geometry}
-\newcommand{\inva}{x + y \geq 0}
-\newcommand{\invb}{x - y \leq 0}
-\newcommand{\invc}{y \geq 1}
-\newcommand{\invd}{y \geq 3}
-\newcommand{\inv}{(\inva \land \invb \land \invc) \lor (\invd)}
-
-Sometimes, we can eyeball whether a transaction is \iconfluent{} with respect
-to an invariant. Other times, not so much. For example, which transactions are
-\iconfluent{} with respect to the invariant $\inv$? It's not immediately
-clear.  Surprisingly, we can interpret the question ``Is $T$ \iconfluent{} with
-respect to $I$?'' geometrically, and doing so will make answering the question
-much easier!
-
-Consider an invariant $I$ over $n$ variables $x_1, \ldots, x_n$. If we imagine
-the values of the $n$ variables as an $n$-dimensional point, then we can draw
-the set of points in $\ints^n$ that satisfy $I$. In fact, if we canonicalize
-$I$ by eliminating negations and converting to conjunctive normal form, then
-the set of points satisfying $I$ is the intersection (for $\land)$ and union
-(for $\lor$) of the solutions to a set of linear equations and inequalities,
-which is always straightforward to draw. For example, the set of points
-satisfying $\inv$ is shown in \figref{geometric-example}.
-
-\begin{figure}
-  \newenvironment{griddedpic}{
-    \begin{tikzpicture}[scale=0.4]
-      \clip (-4, -1) rectangle (4, 4);
-  }{
-      \draw (-4, -1) grid (4, 4);
-      \draw[ultra thick] (-4, 0) -- (4, 0);
-      \draw[ultra thick] (0, -1) -- (0, 4);
-    \end{tikzpicture}
-  }
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[red, opacity=0.4]
-        (-10, 10) -- (10, -10) -- (10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\inva$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[blue, opacity=0.4]
-        (-10, -10) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invb$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[green, opacity=0.4]
-        (-10, 1) -- (10, 1) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-    \caption{$\invc$}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[orange, opacity=0.4]
-        (-10, 3) -- (10, 3) -- (10, 10) -- (-10, 10) -- cycle;
-    \end{griddedpic}
-  \caption{$\invd$}
-
-  \end{subfigure}
-  \begin{subfigure}[b]{\textwidth}
-    \centering
-    \begin{griddedpic}
-      \fill[purple, opacity=0.4]
-        (-1, 1) -- (-3, 3) -- (-10, 3) -- (-10, 10) -- (10, 10) -- (10, 3) --
-        (3, 3) -- (1, 1) -- cycle;
-    \end{griddedpic}
-    \caption{$\inv$}
-  \end{subfigure}
-
-  \caption{An illustration of the solution to $\inv$.}
-  \label{fig:geometric-example}
-\end{figure}
-
-Next, we can imagine a transaction $t$ as an $n$-dimensional vector $(k_{x_1},
-\ldots, k_{x_n})$ where $t$ adds constant $k_{x_i}$ to variable $x_i$. Now, the
-question of whether a set of transactions $T$ is \iconfluent{} with respect to
-invariant $I$ has been reduced to the following question. Starting at an
-arbitrary point $D$ in the solution space of $I$, if we can add vectors $t^1_1,
-t^1_2, \ldots, t^1_m$ to $D$ and vectors $t^2_1, t^2_2, \ldots, t^2_n$ to $D$
-all while staying in the solution space of $I$, then are we guaranteed that $D
-+ t^1_1 + t^1_2 + \ldots + t^1_m + t^2_1 + t^2_2 + \ldots + t^2_n$ is in the
-solution space of $I$?
-
-Turning again to our example in \figref{geometric-example}, it is now clear
-that any transaction $(k_x, k_y)$ where $y \geq 0$ and $|x| \leq y$ is
-\iconfluent{} with respect to $\inv$. These are the vectors that point up and
-not too far left or right.
-
-\subsection{One is Enough}\label{sec:counter-one}
-Define a set of transactions $T$ to be $k$-\iconfluent{} with respect to
-invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
-from $T$ of length at most $k$, if $D$ satisfies $I$, $C_1$ satisfies $I$
-starting at $D$, and $C_2$ satisfies $I$ starting at $D$, then $D \circ
-C_1 \circ C_2$ satisfies $I$.
-%
-A set of transactions $T$ is \iconfluent{} with respect to an invariant $I$ if
-it is $k$-\iconfluent{} for all $k$. Thus, when reasoning about \iconfluent{}
-transactions, we have to consider potentially large transaction chains which
-can be a bit brain boggling. Reasoning about 1-\iconfluence{} is much easier
-because we only have to consider a pair of transactions at a time. This
-subsection proves a convenient lemma that 1-\iconfluence{} and \iconfluence{}
-are equivalent.
-
-\begin{theorem}\label{thm:one-is-enough}
-  For all sets of transactions $T$ and invariants $I$, T is 1-\iconfluent{} if
-  and only if it is \iconfluent{}.
-\end{theorem}
-\begin{proof}
-  The reverse direction is immediate from the definition of \iconfluence{}. For
-  the forward direction, consider a set of transactions $T$ that is
-  1-\iconfluent{} with respect to an invariant $I$. Consider an arbitrary
-  database state $D$ and two chains $C^t = t_1, t_2, \ldots, t_m$ and $C^u =
-  u_1, u_2, \ldots, u_n$ such that $D$ satisfies $I$, $C^t$ satisfies $I$
-  starting at $D$, and $C^u$ satisfies $I$ starting at $D$. We show $D \circ
-  C^t \circ C^u$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $i$, that $D \circ C^t_j \circ C^u_k$ satisfies $I$
-  for all $0 \leq j \leq m$, $0 \leq k \leq n$, and $0 \leq i = j + k \leq n +
-  m$.
-  %
-  The base case, when $i = 0$, is trivial. $j = k = 0$, and $D \circ C^t_0
-  \circ C^u_0 = D$ which satisfies $I$ by assumption.
-  %
-  For the inductive step, we wish to show for arbitrary $0 \leq j \leq m$, $0
-  \leq k \leq n$, and $i = j + k$ that $D \circ C^t_j \circ C^u_k$ satisfies
-  $I$. If $j = 0$, then $D \circ C^t_0 \circ C^u_k = D \circ C^u_k$ satisfies
-  $I$ by the assumption that $C^u$ satisfies $I$ starting at $D$, and likewise
-  for when $k = 0$. Now consider $j, k > 0$. By the inductive hypothesis, we
-  know
-    (1) $D \circ C^t_{j-1} \circ C^u_{k-1}$,
-    (2) $D \circ C^t_{j-1} \circ C^u_{k}$, and
-    (3) $D \circ C^t_{j}   \circ C^u_{k-1}$
-  all satisfy $I$. If we let (1) be our initial database state, $t_j$ be one
-  chain, and $u_k$ be the other chain, then the 1-\iconfluence{} of $T$ tells
-  us that $(1) \circ t_j \circ u_k = D \circ C^t_j \circ C^u_k$ satisfies
-  $I$.
-\end{proof}
-
-\newcommand{\baseedges}{
-  \begin{scope}[on background layer]
-    \draw[txn] (03) -- (13) node[midway, above]{$u_1$};
-    \draw[txn] (13) -- (23) node[midway, above]{$u_2$};
-    \draw[txn] (23) -- (33) node[midway, above]{$u_3$};
-    \draw[txn] (03) -- (02) node[midway, left]{$t_1$};
-    \draw[txn] (02) -- (01) node[midway, left]{$t_2$};
-    \draw[txn] (01) -- (00) node[midway, left]{$t_3$};
-    \draw[dtxn] (00) -- (30);
-    \draw[dtxn] (33) -- (30);
-  \end{scope}
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\onescale}{1.1}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-             1/3, 2/3, 3/3,
-        0/2,
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i = 0$}
-    \label{fig:one-is-enough-base}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3,      2/3, 3/3,
-        %
-        0/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/2, 1/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-    \end{tikzpicture}
-    \caption{$i=1$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3,      3/3,
-        0/2,
-        %
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/1, 1/2, 2/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=2$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3,
-        0/2, 1/2,
-        0/1%
-        %
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {0/0, 1/1, 2/2, 3/3} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=3$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2,
-        0/1, 1/1,
-        0/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {1/0, 2/1, 3/2} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-    \end{tikzpicture}
-    \caption{$i=4$}
-    \label{}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1,
-        0/0, 1/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \foreach \x/\y in {2/0, 3/1} {
-        \inode[red]{\x\y}{\x, \y}
-      }
-      \nnode{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-    \end{tikzpicture}
-    \caption{$i=5$}
-    \label{}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.49\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\onescale]
-      \foreach \x/\y in {
-        0/3, 1/3, 2/3, 3/3,
-        0/2, 1/2, 2/2, 3/2,
-        0/1, 1/1, 2/1, 3/1,
-        0/0, 1/0, 2/0%
-      } {
-        \inode{\x\y}{\x, \y}
-      }
-      \inode[red]{30}{3, 0}
-      \baseedges{}
-      \draw[txn] (02) -- (12) node[midway, above] {$u_1$};
-      \draw[txn] (13) -- (12) node[midway, left]  {$t_1$};
-      \draw[txn] (01) -- (11) node[midway, above] {$u_1$};
-      \draw[txn] (12) -- (11) node[midway, left]  {$t_2$};
-      \draw[txn] (12) -- (22) node[midway, above] {$u_2$};
-      \draw[txn] (23) -- (22) node[midway, left]  {$t_1$};
-      \draw[txn] (00) -- (10) node[midway, above] {$u_1$};
-      \draw[txn] (11) -- (10) node[midway, left]  {$t_3$};
-      \draw[txn] (11) -- (21) node[midway, above] {$u_2$};
-      \draw[txn] (22) -- (21) node[midway, left]  {$t_2$};
-      \draw[txn] (22) -- (32) node[midway, above] {$u_3$};
-      \draw[txn] (33) -- (32) node[midway, left]  {$t_1$};
-      \draw[txn] (10) -- (20) node[midway, above] {$u_2$};
-      \draw[txn] (21) -- (20) node[midway, left]  {$t_3$};
-      \draw[txn] (21) -- (31) node[midway, above] {$u_3$};
-      \draw[txn] (32) -- (31) node[midway, left]  {$t_2$};
-      \draw[txn] (20) -- (30) node[midway, above] {$u_3$};
-      \draw[txn] (31) -- (30) node[midway, left]  {$t_3$};
-    \end{tikzpicture}
-    \caption{$i=6$}
-    \label{}
-  \end{subfigure}
-
-  \caption{Illustration of the proof of \thmref{one-is-enough}.}
-  \label{fig:one-is-enough}
-\end{figure}
-
-\figref{one-is-enough} visualizes the proof of \thmref{one-is-enough} using
-\emph{state diagrams}. In a state diagram, each dot represents a database state
-$D$, and we draw an edge labelled $t$ from database $D$ to database $D'$ if $D
-\circ t = D'$. We circle a database dot if we know it satisfies the invariant
-and leave it uncircled otherwise.
-
-\figref{one-is-enough} illustrates an initial database $D$ (top-left dot), a
-chain $C^t = \set{t_1, t_2, t_3}$ (the left edge), a chain $C^u = \set{u_1,
-u_2, u_3}$ (the top edge), and $D \circ C^t \circ C^u$ (the bottom-right dot).
-By assumption, we know $D$, $D \circ C^t_1$, $D \circ C^t_2$, $D \circ C^t$, $D
-\circ C^u_1$, $D \circ C^u_2$, and $D \circ C^u$ satisfy the invariant, and we
-want to show that $D \circ C^t \circ C^u$ satisfies the invariant.
-\figref{one-is-enough-base} illustrates the base case of
-\thmref{one-is-enough}; each of the other six state diagrams illustrate one
-inductive step of \figref{one-is-enough}. In each inductive step, we exploit
-the fact that $T$ is 1-\iconfluent{} to conclude that some set of intermediate
-databases satisfy the invariant.
-
-\subsection{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
-respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
-
-In a nutshell, our algorithm translates a set of transactions $T$ and invariant
-$I$ into a universally quantified invariant $F$ in prenex normal form where $F$
-is a tautology if and only if $T$ is 1-\iconfluent{} with respect to $I$. Thus,
-by \thmref{one-is-enough}, $F$ is a tautology if and only if $T$ is
-\iconfluent{} with respect to $I$. Our algorithm then uses Z3 \cite{de2008z3}
-to check if $F$ is a tautology.
-
-The algorithm for constructing $F$ is best explained by way of example.
-Consider a simple case where $I = x + y \geq 0$, and $T$ consists of a single
-transaction $t$ which increments $x$ by 1 and $y$ by 2. Our algorithm
-constructs the following formula
-\begin{align}
-  & \forall x.\> \forall y.\>                \label{eq:vars} \\
-  & \forall dx_1.\> \forall dy_1.\>          \label{eq:dvars1} \\
-  & \forall dx_2.\> \forall dy_2.\>          \label{eq:dvars2} \\
-  & (dx_1 = 1 \land dy_1 = 2) \land {}       \label{eq:t1} \\
-  & (dx_2 = 1 \land dy_2 = 2) \implies {}    \label{eq:t2} \\
-  & x + y \geq 0 \implies {}                 \label{eq:i} \\
-  & x + dx_1 + y + dy_1 \geq 0 \implies {}   \label{eq:i1} \\
-  & x + dx_2 + y + dy_2 \geq 0 \implies {}   \label{eq:i2} \\
-  & x + dx_1 + dx_2 + y + dy_1 + dy_2 \geq 0 \label{eq:i12}
-\end{align}
-where
-\begin{itemize}
-  \item
-    \eqref{eq:vars} universally quantifies the free variables in $I$.
-
-  \item
-    \eqref{eq:dvars1} and \eqref{eq:dvars2} universally quantify variables
-    $dx_1$ and $dx_2$ for each free variable $x$ in the invariant $I$.
-    Intuitively, if the invariant has free variables $a, \ldots, z$, then the
-    variables $da_1, \ldots, dz_1$ represent the increments and decrements of
-    one transaction and $da_2, \ldots, dz_2$ represent the increments and
-    decrements of another transaction.
-
-  \item
-    \eqref{eq:t1} and \eqref{eq:t2} assert that $dx_1$, $dy_1$, $dx_2$, and
-    $dy_2$ must take on values consistent with $t$. Specifically, $x$ must be
-    incremented by 1, and $y$ must be incremented by $2$.
-
-  \item
-    \eqref{eq:i} is $I$. Intuitively, it predicates our implication on the
-    assumption that the current database state satisfies the invariant.
-
-  \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
-
-  \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
-\end{itemize}
-
-To summarize, the formula encodes the statement that for all database state
-$(x, y)$ for all transactions $t_1 = (dx_1, dy_1) \in T$, and for all
-transactions $t_2 = (dx_2, dy_2) \in T$, if $(x, y)$, $(x + dx_1, y + dy_1)$,
-and $(x + dx_2, y + dy_2)$ satisfy the invariant, then so does $(x + dx_1 +
-dx_2, y + dy_1 + dy_2)$. In other words, the formula is nothing more than a
-restatement of the definition of 1-\iconfluence!
-
-\subsection{\imp{} Transactions}\label{sec:imptxns}
-So far, we've modelled a transaction $t: \var \rightharpoonup \ints$ as a
-partial function from variables to integers. Intuitively, a transaction chooses
-some \emph{fixed} subset of variables and increments and decrements each by
-some \emph{fixed} amount. We'll call these kinds of transactions \emph{fixed
-transactions}. While simple, fixed transactions are \emph{inexpressive}. We can
-model, for example, a transaction that increments $x$ by 42 and decrements $y$
-by 14, but we \emph{can't} model a transaction that increments $y$ by $x$ or a
-transaction that increments $y$ by 1 whenever $x$ is positive and increments
-$z$ whenever $x$ is negative.
-
-In this subsection, we introduce a simple imperative programming language
-\imp{}\footnote{That's IMP \cite{winskel1993formal} with an $\iinvariant{}$.}.
-Instead of looking at fixed transactions, we'll look at \imp{}
-transactions---transactions written in \imp{}---and ask ourselves the question
-of how do determine whether a set of \imp{} transactions is \iconfluent{} with
-respect to some invariant.
-
-\begin{figure}[h]
-  \centering
-  \begin{gather*}
-    \begin{array}{rrll}
-      \impaexp\;
-        a & ::= & k       & \text{\emph{constant}} \\
-          & |   & x       & \text{\emph{variable}} \\
-          & |   & read(x) & \text{\emph{database read}} \\
-          & |   & a + a   & \text{\emph{addition}} \\
-          & |   & a - a   & \text{\emph{subtraction}} \\
-          & |   & - a     & \text{\emph{negation}} \\
-      &&& \\
-      \impbexp\;
-        b & ::= & \textsf{true}  & \text{\emph{true}} \\
-          & |   & \textsf{false} & \text{\emph{false}} \\
-          & |   & a \odot a      & \text{\emph{linear constraint}} \\
-          & |   & \lnot b        & \text{\emph{negation}} \\
-          & |   & b \lor b       & \text{\emph{disjunction}} \\
-          & |   & b \land b      & \text{\emph{conjunction}} \\
-      &&& \\
-      \impcom\;
-        c  & ::= & \impskip        & \text{\emph{skip}} \\
-           & |   & x := a          & \text{\emph{assignment}} \\
-           & |   & add(x, a)       & \text{\emph{database adds}} \\
-           & |   & c; c            & \text{\emph{sequence}} \\
-           & |   & \impif{b}{c}{c} & \text{\emph{conditional}} \\
-      &&& \\
-      \odot & ::= & =    \vert
-                    \neq \vert
-                    <    \vert
-                    \leq \vert
-                    >    \vert
-                    \geq & \text{\emph{comparator}}
-    \end{array}
-  \end{gather*}
-
-  \caption{
-    A grammar for \imp{}: a simple imperative transaction language inspired by
-    IMP \cite{winskel1993formal} and $\mathcal{L}$ \cite{roy2015homeostasis}.
-    The language consists of arithmetic expressions $\impaexp{}$, boolean
-    expressions $\impbexp{}$, and commands $\impcom{}$.
-  }
-  \label{fig:imp-grammar}
-\end{figure}
-
-The syntax of \imp{} is defined by the grammar in \figref{imp-grammar}. The
-semantics of \imp{} can be defined denotationally. That is, we can define a
-denotational semantics $\denote{\cdot}$ where for an \imp{} program $c$,
-$\denote{c}: (\var \to \ints) \to (\var \rightharpoonup \ints)$ is a function
-from a database to a fixed transaction. Defining these semantics is left as an
-exercise to the reader.  See \cite{cs6110sp2016lecture19} for guidance.
-\todo{Define, explicitly, the denotational semantics if need be.}
-
-\emph{Applying an \imp{} transaction} $c$ to database state $D$, denoted by $D
-\circ c$, is defined as $D \circ c \defeq D \circ (\denote{c}(D))$.  Note that
-unlike fixed transaction application, \imp{} transaction application is
-\emph{not} commutative! For example, consider the database $D=\set{(x, 0)}$ and
-\imp{} transactions $c_1 = add(x, 1)$ and $c_2 = add(x, read(x))$.
-\[
-  D \circ c_1 \circ c_2
-    = \set{(x, 1)} \circ c_2
-    = \set{(x, 2)}
-    \neq \set{(x, 1)}
-    = \set{(x, 0)} \circ c_1
-    = D \circ c_2 \circ c_1
-\]
-
-An \imp{} transaction chain $B$ created from a set of \imp{} transactions $T$
-is a sequence of \imp{} transactions $c_1, \ldots, c_n$ where $c_i \in T$ for
-$1 \leq i \leq n$. Each \imp{} transaction chain $B$ starting with database
-state $D$ has a corresponding fixed transaction chain $C$ denoted
-$\denote{B}(D)$. Let $D_i = D \circ c_1 \circ \cdots \circ c_i$. Then, the
-corresponding fixed transaction chain is $\denote{c_1}(D_0), \ldots,
-\denote{c_n}(D_{n-1})$. We say $B$ \emph{satisfies invariant} $I$ starting at
-$D$, denoted $I(B@D)$, if the corresponding fixed transaction chain $C$
-satisfies invariant $I$ starting at $D$.
-
-We say a set of \imp{} transactions $T$ is \iconfluent{} with respect to
-invariant $I$ if for all database states $D$, all \imp{} chains $B_1$ and $B_2$
-created from $T$, and corresponding fixed transaction chains $C_1$ and $C_2$,
-if $D$ satisfies $I$, $B_1$ satisfies $I$ starting at $D$, and $B_2$ satisfies
-$I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$. We define
-$k$-\iconfluence{} similar to how we did in \secref{counter-one}.
-
-Now that we've formalized what it means for a set of \imp{} transactions to be
-\iconfluent{} with respect to some invariant, how do we determine
-\iconfluence{} algorithmically? One initial hunch would be to apply the same
-technique from \secref{solution} with slight modification. Unfortunately,
-\clmref{imp-1-iconfluence} presents a challenge.
-
-\begin{claim}\label{clm:imp-1-iconfluence}
-  For \imp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \imp{} transactions:
-  \begin{align*}
-    I &\defeq
-      (-2 \leq x \leq 2 \land y = 0) \lor (x = 1 \land 0 \leq y \leq 2) \\
-    c_1 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, 1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }} \\
-    c_2 &\defeq \impif{x = 0 \land y = 0}{
-      add(x, -1)
-    }{\impif{x = 1 \land y = 0} {
-      add(y, 1)
-    }{
-      \impskip
-    }}
-  \end{align*}
-
-  $T = \set{c_1, c_2}$ is $1$-\iconfluent{} with respect to $I$, but not
-  $2$-\iconfluent{} with respect to $I$. Proving this is easiest if we
-  interpret \imp{} transactions geometrically, as we did with fixed
-  transactions.
-
-  As before, invariants over $n$ variables correspond to a set of points in
-  $\ints^n$; for example, $I$ is visualized in \figref{iviz}. An \imp{}
-  transaction $c$ can be thought of as vector valued functions $c: \ints^n \to
-  \ints^n$ that maps each point in $\ints^n$ to a $n$-dimensional vector. For
-  example, $c_1$ and $c_2$ are visualized in \figref{c1viz} and \figref{c2viz}.
-  The visualization of $c_1$ has two vectors. The $(1, 0)$ vector originating
-  at the origin corresponds to $c_1$ adding $1$ to $x$ whenever $x = 0 \land y
-  = 0$. Similarly, the $(0, 1)$ originating at $(1, 0)$ corresponds to $c_1$
-  adding $1$ to $y$ whenever $x = 1 \land y = 0$.
-
-  Using this geometric interpretation, it's easy to perform a brute-force check
-  that $T$ is $1$-\iconfluent with respect to $I$. However, consider the
-  transaction chains $B_1 = \set{c_1, c_2}$ and $B_2 = \set{c_2, c_1}$ starting
-  in database state $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_1 = (0, 1)$ satisfies $I$,
-    \item $D \circ c_1 \circ c_2 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_2 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_2 \circ c_1 = (-1, 0)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 0) = (0, 1)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  \clmref{imp-1-iconfluence} puts us in a bit of pickle. Think about how to get
-  out of it. Maybe we check for $k$-\iconfluence for some $k$ and then use a
-  reservation system to bound the amount of divergence between replicas? That
-  doesn't seem to make sense for a system with high transaction throughput.
-  Maybe there are some other tricks we can play?
-}
-
-\tikzstyle{point}=[shape=circle, fill=blue, opacity=0.75]
-\tikzstyle{vec}=[->, ultra thick, red, shorten >=0.1cm]
-\begin{figure}[h]
-  \centering
-
-  \newcommand{\gridandpoints}{
-    \draw[ultra thick] (-2, 0) -- (2, 0);
-    \draw[ultra thick] (0, 0) -- (0, 2);
-    \draw (-2, 0) grid (2, 2);
-    \foreach \x/\y in {-2/0, -1/0, 0/0, 1/0, 2/0, 1/1, 1/2} {
-      \node[point] (\x-\y) at (\x, \y) {};
-    }
-  }
-
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:iviz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:c1viz}
-  \end{subfigure}
-  \begin{subfigure}[c]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}
-      \gridandpoints{}
-      \draw[vec] (0, 0) -- (-1, 0);
-      \draw[vec] (1, 0) -- (1, 1);
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:c2viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, and $c_2$ from
-    \clmref{imp-1-iconfluence}. $I$ corresponds to a subset of $\ints^2$. $c_1$
-    and $c_2$ correspond to vector valued functions of type $\ints^2 \to
-    \ints^2$.
-  }
-  \label{fig:icviz}
-\end{figure}
-
-\subsection{\wimp{} Transactions}\label{sec:datatxns}
-We began this section by looking at fixed transactions and concluded in
-\thmref{one-is-enough} that if a set of fixed transactions is 1-\iconfluent{}
-with respect to an invariant $I$, then it is also \iconfluent{}. Then in
-\secref{imptxns}, we enriched transactions from simple fixed partial functions
-to full-blown programs written in \imp{}. Unfortunately, this increase in
-expressiveness came at the cost of the nice fact that 1-\iconfluence{} is
-equivalent to \iconfluence{}, as we saw in \clmref{imp-1-iconfluence}.
-
-Perhaps \imp{} is too powerful of a transaction language. In particular, it has
-two powerful features that fixed transactions do not have. First, \imp{} has
-conditionals allowing the control flow of an \imp{} program to depend on the
-database against which it is run. Second, even without conditionals, the
-behavior of an \imp{} transaction can still depend on the database state
-against which it is run. For example, the behavior of the transaction $add(x,
-read(x))$ depends on the initial value of $x$, even though the transaction does
-not use a conditional.
-
-Perhaps if we strip \imp{} of some its features, we'll regain the equivalence
-of 1-\iconfluence{} and \iconfluence{}. That is, perhaps there is an
-intermediate language more powerful than fixed transactions but weaker than
-\imp{} for which 1-\iconfluence{} is equivalent to \iconfluence{}.
-
-In this subsection, we'll consider one such language, \wimp{}: a weakened
-version of \imp{}. Then, we'll conclude that for \wimp{}---as for
-\imp{}---1-\iconfluence{} does not imply \iconfluence{}.
-
-Formally, \wimp{} is \imp{} without variables, boolean expressions, skip,
-assignment, and conditionals. Moreover, each database variable can be added to
-at most once. Essentially, \wimp{} programs get to add a single arithmetic
-expression to each database variable. While weaker than \imp{} programs,
-\wimp{} programs are more powerful than fixed transactions because arithmetic
-expressions can include database reads and therefore \wimp{} programs can
-depend on the initial state of the database. In fact, if we remove database
-reads from \wimp{}, we get a language of fixed transactions.
-
-Since \wimp{} is a subset of \imp{}, all the definitions from \imp{} carry over
-directly for \wimp{}. Also note that like \imp{}, \wimp{} transaction
-application is not commutative.
-
-As mentioned above, \wimp{} is not weak enough to enjoy the equivalence of
-1-\iconfluence{} and \iconfluence{}: a fact we now prove.
-
-\begin{claim}\label{clm:wimp-1-iconfluence}
-  For \wimp{} transactions, $1$-\iconfluence $\centernot \implies$ \iconfluence.
-\end{claim}
-\begin{proof}
-  Consider the following invariant and \wimp{} transactions:
-  \begin{align*}
-    I   &\defeq x \neq 0 \lor y = 0 \\
-    c_1 &\defeq y.add(1) \\
-    c_2 &\defeq x.add(read(x) + 1) \\
-    c_3 &\defeq x.add(read(x) - 1)
-  \end{align*}
-
-  As usual, we can visualize invariants and transactions geometrically, as
-  shown in \figref{wimp-icviz}. We can perform a brute-force case analysis over
-  all pairs of transactions and all points to show $T = \set{c_1, c_2, c_3}$ is
-  1-\iconfluent{} with respect to $I$. This is left as exercise to the reader.
-
-  However, consider the \wimp{} chains $B_1 = \set{c_2, c_1}$ and $B_2 =
-  \set{c_3, c_1}$ in database $D = (0, 0)$. We know that
-  \begin{itemize}
-    \item $D = (0, 0)$ satisfies $I$,
-    \item $D \circ c_2 = (1, 0)$ satisfies $I$,
-    \item $D \circ c_2 \circ c_1 = (1, 1)$ satisfies $I$,
-    \item $D \circ c_3 = (-1, 0)$ satisfies $I$, and
-    \item $D \circ c_3 \circ c_1 = (-1, 1)$ satisfies $I$,
-  \end{itemize}
-  but $D \circ C_1 \circ C_2 = D \circ (1, 0) \circ (0, 1) \circ (-1, 0) \circ
-  (0, 1) = (0, 2)$ does not satisfy $I$. Therefore, $T$ is not 2-\iconfluent
-  with respect to $I$ and is therefore not \iconfluent with respect to $I$.
-\end{proof}
-
-\todo{
-  We're in a pickle again! I think this result also might suggest that the
-  claims in \cite{balegas2015putting} are wrong? More likely, I'm
-  misunderstanding the paper.
-}
-
-\newcommand{\width}{2}
-\newcommand{\smallwidth}{1}
-\newcommand{\height}{2}
-
-\newcommand{\gridxy}{
-  \draw[ultra thick] (-\width, 0) -- (\width, 0);
-  \draw[ultra thick] (0, -\height) -- (0, \height);
-  \draw (-\width, -\height) grid (\width, \height);
-}
-
-\newcommand{\forxy}[1]{
-  \foreach \x in {-\width, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \width} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\newcommand{\forxysmall}[1]{
-  \foreach \x in {-\smallwidth, ..., -1} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {1, ..., \smallwidth} {
-    \foreach \y in {-\height, ..., \height} { #1 } }
-  \foreach \x in {0} {
-    \foreach \y in {0} { #1 } }
-}
-
-\begin{figure}[h]
-  \centering
-  \newcommand{\thescale}{0.7}
-
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-    \end{tikzpicture}
-    \caption{Visualization of $I$.}
-    \label{fig:wimp-iviz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxy{\draw[vec] (\x, \y) -- (\x, \y + 1);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_1$.}
-    \label{fig:wimp-c1viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x + 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_2$.}
-    \label{fig:wimp-c2viz}
-  \end{subfigure}
-  \begin{subfigure}[b]{0.24\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=\thescale]
-      \gridxy{}
-      \forxy{\node[point] (\x-\y) at (\x, \y) {};}
-      \forxysmall{\draw[vec] (\x, \y) -- (\x + \x - 1, \y);}
-    \end{tikzpicture}
-    \caption{Visualization of $c_3$.}
-    \label{fig:wimp-c3viz}
-  \end{subfigure}
-
-  \caption{
-    A geometric interpretation of $I$, $c_1$, $c_2$, and $c_3$ from
-    \clmref{wimp-1-iconfluence}. Note that the visualizations include only a
-    subset of the points satisfying $I$ and that not all vectors are drawn.
-  }
-  \label{fig:wimp-icviz}
-\end{figure}
-
-\subsection{\iconfluence{} Alternatives}
+\section{\iconfluence{} Alternatives}
 \begin{todoitemize}
-  \item \todo{\istrength $\centernot \implies$ \isafety?}
-  \item \todo{\isafety $\implies$ \istrength?}
-  \item \todo{\istrength $\centernot \implies$ \iconfluence?}
-  \item \todo{\iconfluence $\implies$ \istrength?}
-  \item \todo{\istrengthstar $\implies$ \istrength?}
   \item \todo{\istrength $\implies$ \istrengthstar?}
+  \item \todo{\iconfluence $\implies$ 1-\iconvergence?}
   \item \todo{Find something between \isafety{} and \iconfluence.}
   \item \todo{1-iconfluence with derived txns to start}
+  \item \todo{Explain alternatives.}
 \end{todoitemize}
+
+\thmref{imp-1-iconfluence} and \thmref{wimp-1-iconfluence} are depressing
+results. Given a set of \imp{} or \wimp{} transactions $T$ and an invariant
+$I$, we'd like to determine if $T$ is \iconfluent{} with respect to $I$ in much
+the same way we did with fixed transactions in \secref{solution}. We'd like to
+generate a formula $F$ (which we give to Z3) that is a tautology if and only if
+$T$ is \iconfluent{} with respect to $I$. However, since 1-\iconfluence{} is
+not equivalent to \iconfluence{}, it's not clear what a formula $F$ would even
+look like.
+
+Clearly, we're stumped. Algorithmically determining \iconfluence{} seems to be
+too hard of a problem. However, not all hope is lost. Consider an analogy.
+Determining whether a sequence of reads and writes is \emph{view serializable}
+is a really hard problem. In fact, it is NP-complete. To circumvent the
+difficulty of the problem, we introduce \emph{conflict serializability}.
+Determining whether a history is conflict serializable is pretty easy, and
+conflict serializability implies (but is not equivalent to) view
+serializability.
+
+This means that we can quickly determine whether a history is conflict
+serializable. If it is, then it is \emph{definitely} view serializable. If it
+is not, well, then it \emph{might be} view serializable or it \emph{might not
+be}. Conflict serializability merely approximates view serializability. There
+are some view serializable histories which are not conflict serializable, but
+the two are close enough that we don't lose any sleep over it.
+
+Like determining view serializability, determining \iconfluence{} is a hard
+problem. In this section we present multiple analogues of conflict
+serializability: efficiently decidable properties of a set of transactions and
+an invariant that imply (but are not equivalent to) \iconfluence{}. We might
+not be able to determine \iconfluence{}, but we might be able to find a tight
+approximation.
+
+We consider the following list of properties. Some properties are visualized
+with state diagrams in \figref{iconfluence-alternatives-diagrams}. The
+relationship between the different properties is illustrated in
+\figref{iconfluence-alternatives-implications} and proved formally in
+\appref{alternativesproofs}.
 
@@
     where $I$ is viewed as a set of database states.
+
+  \item \textbf{1-\iconvergence.}
+    $\forall D, D', D'', c', c''.\>
+      I(D) \land I(D') \land I(D'') \land
+      I(D \circ \denote{c'}(D')) \land
+      I(D \circ \denote{c''}(D'')) \implies
+      I(D \circ \denote{c'}(D') \circ \denote{c''}(D''))$
 \end{itemize}
 
-\begin{claim}\label{clm:0-isafety-implies-istrengthstar}
-  0-\isafety $\implies$ \istrengthstar.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$. We show $T$ is \istrongstar{} with respect to $I$.
-  Let $G_0 = I \times I$. S2 holds trivially: $G_0(I) = I
-  \subseteq I$. Consider an arbitrary $c$, $D$, and $D'$ where $D \in I$ and
-  $(D, D') \in G_0^*$.
-  \begin{align*}
-    & D \in I \land (D, D') \in G_0^*  \\
-    &\implies D \in I \land D' \in I \tag{Definition of $G_0$} \\
-    &\implies D' \in I \land D' \circ \denote{c}(D) \in I \tag{0-\isafety} \\
-    &\implies (D', D' \circ \denote{c}(D)) \in G_0 \tag{Definition of $G_0$}
-  \end{align*}
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-not-implies-0-isafety}
-  \istrengthstar $\centernot\implies$ 0-\isafety
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G \defeq \set{(0, 0), (0, 2)} \cup \setst{(a, b)}{a \geq 2, b \geq a}$
-    \item
-      $I \defeq x = 0 \lor x \geq 2$, and
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$.
-  \end{itemize}
-
-  First note that $c$ is \istrengthstar{} with respect to $I$. S2 holds
-  trivially. To show S3 holds, consider an arbitrary pair $a, b \in G_0^*$.
-  \begin{itemize}
-    \item \textbf{Case $a = 0, b = 0$.}
-      $0 \circ \denote{c}(0) = 0 + 0 = 0$ and $(0, 0) \in G_0$.
-    \item \textbf{Case $a = 0, b >= 2$.}
-      $b \circ \denote{c}(0) = b + 0 = b$ and $(b, b) \in G_0$.
-    \item \textbf{Case $a \geq 2, b >= a$.}
-      $b \circ \denote{c}(a) = b + a$ and $(b, b + a) \in G_0$.
-  \end{itemize}
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ and $\lnot I(1)$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-ipreservation}
-  \istrengthstar $\implies$ \ipreservation
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to
-  invariant $I$ with relation $G_0$. Consider an arbitrary database $D$ that
-  satisfies $I$ and an arbitrary \imp{} transaction $c$. $(D, D) \in G_0^*$,
-  so by S3, $(D, D \circ \denote{c}(D)) \in G_0$. By S2, $I(D \circ \denote{c}(D))$.
-\end{proof}
-
-\begin{claim}\label{clm:istrengthstar-implies-iconfluence}
-  \istrengthstar{} $\implies$ \iconfluence{}.
-\end{claim}
-\begin{proof}
-  Consider a set of transactions $T$ that is \istrongstar{} with respect to an
-  invariant $I$ with relation $G_0$. Consider an arbitrary database state $D$,
-  two \imp{} chains $B^c = c_1, c_2, \ldots, c_m$ and $B^d = d_1, d_2, \ldots,
-  d_n$, and two fixed transaction chains $C^c$ and $C^D$ such that $D$
-  satisfies $I$, $B^c$ satisfies $I$ starting at $D$, $B^d$ satisfies $I$
-  starting at $D$, $C^c = \denote{B^c}(D)$, and $C^d = \denote{B^d}(D)$. We
-  show $D \circ C^c \circ C^d$ satisfies $I$.
-
-  Let $C^z_j$ denote the prefix of $C^z$ of length $j$. We show by strong
-  mathematical induction on $0 \leq i \leq m + n$, that for all natural numbers
-  $j, k, j', k'$ if
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$, and
-    $k' \leq k$, then
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  This proof is similar to the proof of \thmref{one-is-enough}, and we can
-  visualize it similar to how we visualized \thmref{one-is-enough} in
-  \figref{one-is-enough}. Intuitively, our induction progresses diagonally from
-  the upper left corner $D$ to the bottom right corner $D \circ C^c_j \circ
-  C^d_k$. For each diagonal, our inductive hypothesis says all states on the
-  diagonal are reachable via $G_0^*$ from points above and to the left.
-
-  %
-  The base case, when $i = 0$, is trivial. $i = j = k = j' = k' = 0$, so $(D
-  \circ C^c_0 \circ C^d_0, D \circ C^c_0 \circ C^d_0) = (D, D) \in G_0^*$ by
-  the definition of a reflexive transaction closure.
-  %
-  For the inductive step, consider an arbitrary set of natural numbers
-  $i, j, k, j', k'$ where
-    $0 < i \leq m + n$,
-    $j \leq m$,
-    $k \leq n$,
-    $i = j + k$,
-    $j' \leq j$
-    $k' \leq k$.
-  If $j' = j$ and $k' = k$, then $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ
-  C^c_{j} \circ C^d_{j}) \in G_0^*$ by the definition of a reflexive transitive
-  closure. Otherwise $j' < j$ or $k' < k$. Without loss of generality, assume
-  $j' < j$; the proof is symmetric for $k' < k$. We can deduce the following
-  facts:
-  \begin{gather}
-    (D, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-a} \\
-    (D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-b} \\
-    I(D \circ C^c_{j-1} \circ C^d_{k})
-      \label{strength-confluence-c} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j-1} \circ C^d_{k}) \in G_0^*
-      \label{strength-confluence-d} \\
-    (D \circ C^c_{j-1} \circ C^d_{k}, D \circ C^c_{j} \circ C^d_{k}) \in G_0
-      \label{strength-confluence-e}
-  \end{gather}
-  where
-  \begin{itemize}
-    \item
-      \eqref{strength-confluence-a} and \eqref{strength-confluence-b} are
-      deduced from the inductive hypothesis applied to $i - 1$;
-    \item
-      \eqref{strength-confluence-c} is deduced from
-      \eqref{strength-confluence-a}, the assumption $I(D)$, and S2;
-    \item
-      \eqref{strength-confluence-d} is deduced from the definition of a
-      reflexive transitive closure;
-    \item
-      \eqref{strength-confluence-e} is deduced from
-      \eqref{strength-confluence-d} and S3 applied to $C^c_j$, $D \circ
-      C^c_{j-1} \circ C^d_{k}$, and $D \circ C^c_{j-1} \circ C^d_{k}$.
-  \end{itemize}
-  \eqref{strength-confluence-b} and \eqref{strength-confluence-e} imply
-  $(D \circ C^c_{j'} \circ C^d_{k'}, D \circ C^c_j \circ C^d_k) \in G_0^*$.
-
-  Finally, $I(D)$, $(D, D \circ D^c_m \circ D^d_n) \in G_0^*$, and S2 imply
-  that $I(D \circ D^c_m \circ D^d_n)$.
-\end{proof}
-
-\begin{claim}\label{clm:iconfluence-not-implies-istrengthstar}
-  \iconfluence{} $\centernot\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let $I \defeq x = 0$ and the single \imp{}
-  transaction $c = add(x, 1)$. Note that $c$ is vacuously \iconfluent{} with
-  respect to $I$. Also note that $c$ is not \ipreserving{} with respect to $I$:
-  $0$ satisfies $I$ but $0 \circ \denote{c}(0) = 1$ does not. By the
-  contrapositive of \clmref{istrengthstar-implies-ipreservation}, $c$ is not
-  \istrongstar{} with respect to $I$.
-\end{proof}
-
-\begin{claim}\label{clm:0-isafety-implies-istrength}
-  0-\isafety{} $\implies$ \istrength{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is 0-\isafe{} with
-  respect to invariant $I$; we show $T$ is \istrong{} with respect to $I$. Let
-  $G_0 = I \times I$. $T2$ holds trivially. For all $c \in T$, let $n = 1$ and
-  $P_n = Q_n = I$. T3b, T3c, and T3d reduce to $I \subseteq I$, $I \subseteq
-  I$, and $G_0(I) \subseteq I$ which all hold trivially. T3e reduces to $I
-  \times \denote{c}(I)(I) \subseteq G_0$. To prove this, it is sufficient to
-  prove $\denote{c}(I)(I) \subseteq I$ which holds directly from 0-\isafety{}.
-\end{proof}
-
-\begin{claim}\label{clm:0-istrength-not-implies-0-isafety}
-  \istrength{} $\centernot\implies$ 0-\isafety.
-\end{claim}
-\begin{proof}
-  Consider a database over a single variable $x$. For convenience, we'll denote
-  database states as integers. Let
-  \begin{itemize}
-    \item
-      $G_0 \defeq \set{(0, 0)} \cup \setst{(a, b)}{a \geq 2, b \geq 2}$;
-    \item
-      $I \defeq x = 0 \lor x \geq 2$;
-    \item
-      the single \imp{} transaction $c \defeq
-      \impif{read(x)=0}{add(x, 0)}{add(x, 1)}$;
-    \item
-      $n \defeq 2$;
-    \item
-      $P_1 \defeq Q_1 \defeq \set{0}$; and
-    \item
-      $P_2 \defeq Q_2 \defeq \setst{x}{x \geq 2}$.
-  \end{itemize}
-
- First note that $c$ is \istrong{} with respect to $I$.  T2, T3b, T3c, and T3d
- hold trivially. For $i = 1$, T3e reduces to $\set{0} \times
- \denote{c}(\set{0})(\set{0}) \subseteq G_0$ which holds because $\set{(0, 0
- \circ \denote{c}(0)} = \set{(0, 0)} \subseteq G_0$. For $i = 2$, consider
- arbitrary databse states $x,y,z \in P_2 = G_2$. $(x, y \circ \denote{c}(z)) =
- (x, y + 1) \in G_0$. Thus T3e holds for $i = 2$.
-
-  Now note that $c$ is not 0-\isafe{} with respect to $I$. Notably, $I(0)$ and
-  $I(2)$, but $0 \circ \denote{c}(2) = 0 + 1 = 1$ does not satisfy $I$.
-\end{proof}
-
-\begin{claim}\label{clm:istrength-implies-istrengthstar}
-  \istrength{} $\implies$ \istrengthstar{}.
-\end{claim}
-\begin{proof}
-  Consider an arbitrary set of \imp{} transactions $T$ that is \istrong{} with
-  respect to invariant $I$ with relation $G_0$ and sets $P_1^c, \ldots, P_n^c,
-  Q_1^c, \ldots, Q_n^c$ for every $c \in T$. We show $T$ is \istrongstar{} with
-  respect to $I$ with $G_0$.
-
-  T2 implies S2. To prove S3, consider an arbitrary transaction $c$ and
-  database states $D$ and $D'$ where $I(D)$ and $(D, D') \in G_0^*$. By T3b, $D
-  \in P_i^c$ for some $i$, and by T3c, $D \in Q_i^c$. By T3d, $D' \in Q_i^c$.
-  Finally by T3e, $(D', D' \circ \denote{c}(D)) \in G_0$.
-\end{proof}
-
+\tikzstyle{automatable}=[fill=green!20]
+\tikzstyle{intractable}=[fill=red!20]
 \begin{figure}[h]
@@
   \begin{tikzpicture}
-    \node[] (0-isafety) at (0, 0) {0-\isafety};
-    \node[below=of 0-isafety]     (1-isafety)     {1-\isafety};
-    \node[below=of 1-isafety]     (k-isafety)     {$k$-\isafety};
-    \node[right=of 0-isafety]     (istrengthstar) {\istrengthstar};
-    \node[above=of istrengthstar] (ipreservation) {\ipreservation};
-    \node[right=of istrengthstar] (iconfluence)   {\iconfluence};
-    \node[below=of iconfluence]   (k-iconfluence) {$k$-\iconfluence};
-    \node[below=of k-iconfluence] (1-iconfluence) {1-\iconfluence};
-    \node[below=of istrengthstar] (istrength)     {\istrength};
+    \node[automatable] (0-isafety) at (0, 0) {0-\isafety};
+    \node[automatable, below=of 0-isafety]     (1-isafety)      {1-\isafety};
+    \node[automatable, below=of 1-isafety]     (k-isafety)      {$k$-\isafety};
+    \node[intractable, right=of 0-isafety]     (istrengthstar)  {\istrengthstar};
+    \node[automatable, above=of istrengthstar] (ipreservation)  {\ipreservation};
+    \node[intractable, right=of istrengthstar] (iconfluence)    {\iconfluence};
+    \node[automatable, below=of iconfluence]   (k-iconfluence)  {$k$-\iconfluence};
+    \node[automatable, below=of k-iconfluence] (1-iconfluence)  {1-\iconfluence};
+    \node[automatable, below=of istrengthstar] (istrength)      {\istrength};
+    \node[automatable, below=of istrength]     (1-iconvergence) {1-\iconvergence};
 
@@
     \draw[impl] (istrength) -> (istrengthstar);
+    \draw[ultra thick, impl] (0-isafety) -> (1-iconvergence);
+    \draw[ultra thick, impl] (1-iconvergence) -> (iconfluence);
   \end{tikzpicture}
@@
     alternatives. We have proven bold implications in both directions. For
-    example, we have proven both \iconfluence $\protect \centernot \implies$
-    \istrengthstar.
+    example, we have proven both \istrengthstar{} $\implies$ \iconfluence{} and
+    \iconfluence{} $\protect \centernot \implies$ \istrengthstar{}. Conditions
+    shaded green can be automatically determined using an SMT solver.
+    Conditions shaded red are not easy to automatically determine.
   }
-  \label{fig:iconfluence-alternatives}
+  \label{fig:iconfluence-alternatives-implications}
 \end{figure}
-
-\begin{figure}[h]
-  \centering
-
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{d1}{0, 0}
-      \inode[label=above:$D'$]{d2}{1, 0}
-      \dnode{b1}{0, -1}
-      \nnode{b2}{1, -1}
-      \draw[txn] (d2) -- (b2) node[midway, right]{$\denote{c}(D')$};
-      \draw[dtxn] (d1) -- (b1) node[midway, left]{$\denote{c}(D')$};
-    \end{tikzpicture}
-    \caption{0-\isafety}
-    \label{fig:0-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \dnode{b}{0, -1}
-      \draw[txn] (a) -- (b) node[midway, left]{$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{\ipreservation}
-    \label{fig:ipreservation}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.3\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c}$]{l1}{$(a) + (240:1)$}
-      \inode[label=right:$D_{d}$]{r1}{$(a) + (300:1)$}
-      \dnode{b}{$(l1) + (300:1)$}
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c}(D)$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d}(D)$};
-      \draw[txn] (l1) -- (b) node[midway, left] {$\denote{d}(D)$};
-      \draw[txn] (r1) -- (b) node[midway, right] {$\denote{c}(D)$};
-    \end{tikzpicture}
-    \caption{1-\iconfluence}
-    \label{fig:1-iconfluence}
-  \end{subfigure}
-
-  \vspace{1cm}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D_0$]{l1}{0, 4}
-      \dnode[label=left:$D_1$]{l2}{0, 3}
-      \dnode[label=left:$D_2$]{l3}{0, 2}
-      \dnode[label=left:$D_3$]{l4}{0, 1}
-      \dnode{l5}{0, 0}
-
-      \inode[label=above:$D_0'$]{r1}{1, 4}
-      \inode[label=right:$D_1'$]{r2}{1, 3}
-      \inode[label=right:$D_2'$]{r3}{1, 2}
-      \inode[label=right:$D_3'$]{r4}{1, 1}
-      \nnode{r5}{1, 0}
-
-      \draw[txn] (r1) -- (r2) node[midway, right]{$\denote{c_1}(D_0')$};
-      \draw[txn] (r2) -- (r3) node[midway, right]{$\denote{c_2}(D_1')$};
-      \draw[txn] (r3) -- (r4) node[midway, right]{$\denote{c_3}(D_2')$};
-      \draw[txn] (r4) -- (r5) node[midway, right]{$\denote{c_4}(D_3')$};
-
-      \draw[txn] (l1) -- (l2) node[midway, left]{$\denote{c_1}(D_0')$};
-      \draw[txn] (l2) -- (l3) node[midway, left]{$\denote{c_2}(D_1')$};
-      \draw[txn] (l3) -- (l4) node[midway, left]{$\denote{c_3}(D_2')$};
-      \draw[txn] (l4) -- (l5) node[midway, left]{$\denote{c_4}(D_3')$};
-    \end{tikzpicture}
-
-    \caption{$k$-\isafety, $k=3$}
-    \label{fig:k-isafety}
-  \end{subfigure}%
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-
-    \begin{tikzpicture}[scale=1.5]
-      \inode[label=above:$D$]{a}{0, 0}
-      \inode[label=left:$D_{c_1}$]{l1}{$(a) + (240:1)$}
-      \inode[label=left:$D_{c_2}$]{l2}{$(l1) + (240:1)$}
-      \inode[label=left:$D_{c_3}$]{l3}{$(l2) + (240:1)$}
-
-      \inode[label=right:$D_{d_1}$]{r1}{$(a) + (300:1)$}
-      \inode[label=right:$D_{d_2}$]{r2}{$(r1) + (300:1)$}
-      \inode[label=right:$D_{d_3}$]{r3}{$(r2) + (300:1)$}
-
-      \dnode{b}{$(l3) + (300:3)$}
-
-      \draw[txn] (a) -- (l1) node[midway, left] {$\denote{c_1}(D)$};
-      \draw[txn] (l1) -- (l2) node[midway, left] {$\denote{c_2}(D_{c_1})$};
-      \draw[txn] (l2) -- (l3) node[midway, left] {$\denote{c_3}(D_{c_2})$};
-      \draw[txn] (a) -- (r1) node[midway, right] {$\denote{d_1}(D)$};
-      \draw[txn] (r1) -- (r2) node[midway, right] {$\denote{d_2}(D_{d_1})$};
-      \draw[txn] (r2) -- (r3) node[midway, right] {$\denote{d_3}(D_{d_2})$};
-      \draw[txn] (l3) -- (b) node[midway, left] {$\denote{c_1, c_2, c_3}(D)$};
-      \draw[txn] (r3) -- (b) node[midway, right] {$\denote{d_1, d_2, d_3}(D)$};
-    \end{tikzpicture}
-
-    \caption{\iconfluence}
-    \label{fig:iconfluence}
-  \end{subfigure}
-
-  \begin{subfigure}[b]{0.5\textwidth}
-    \centering
-    \begin{tikzpicture}[scale=2]
-      \inode[label=above:$D$]{s2a}{0, 1}
-      \dnode{s2b}{0, 0}
-      \draw[gtxn] (s2a) -- (s2b) node[midway, left]{$\denote{c}(D)$};
-
-      \inode[label=above:$D$]{s3a}{2, 1}
-      \nnode[]{s3b}{2, 0}
-      \nnode[label=above:$D'$]{s3c}{3, 1}
-      \nnode[]{s3d}{3, 0}
-      \draw[txn] (s3a) -- (s3b) node[midway, left] {$\denote{c}(D)$};
-      \draw[dgtxn] (s3c) -- (s3d) node[midway, right] {$\denote{c}(D)$};
-      \draw[gtxn] (s3a) -- (s3c) node[midway, above] {$G_0^*$};
-    \end{tikzpicture}
-    \caption{\istrengthstar}
-    \label{fig:istrengthstar}
-  \end{subfigure}
-
-  \caption{}
-  \label{}
-\end{figure}

### 2018-03-26 doc/notes_v1/alternativesproofs.tex
Renaming various docs.

--- 2017-01-27 doc/paper/alternativesproofs.tex
@@
 \end{proof}
+
+\begin{claim}\label{clm:iconfluence-not-implies-1-iconvergence}
+  \iconfluence{} $\centernot\implies$ 1-\iconvergence{}.
+\end{claim}
+\begin{proof}
+  Let
+    $I \defeq 0 \leq x \leq 2$ and
+    $c = \impif{read(x)=0}{add(x,1)}{}$.
+  First, we show that $\set{c}$ is \iconfluent{} with respect to $I$. Consider
+  an arbitrary
+    $D$,
+    $B_1$, and
+    $B_2$
+  where
+    $I(D)$,
+    $I(B_1@D)$, and
+    $I(B_2@D)$.
+  We perform a case analysis on $D$.
+  \begin{itemize}
+    \item \textbf{Case 1: $D = 1 \lor D = 2$.}
+      In this case, $D \circ \denote{B_1}(D) \circ \denote{B_2}(D)$ is
+      guaranteed to equal $D$ because $c$ does not affect $D$ at all. By
+      assumption, $D$ satisfies the invariant.
+    \item \textbf{Case 1: $D = 0$.}
+      If $D = 0$, then $D \circ \denote{B_1}(D)$ is either $0$ (if $B_1$ is
+      empty) or $1$ if $B_1$ is not empty. A symmetric argument holds for
+      $B_2$. Thus, $D \circ \denote{B_1}(D) \circ \denote{B_2}(D)$ is either 0,
+      1, or 2, all of which satisfy $I$.
+  \end{itemize}
+
+  Next, we show that $\set{c}$ is not 1-\iconvergent{} with respect to $I$. Let
+  $D = 1$ and $D' = D'' = 0$.
+    $I(D)$,
+    $I(D \circ \denote{c}(D'))$,
+    $I(D \circ \denote{c}(D''))$,
+  but
+    $D \circ \denote{c}(D') \circ \denote{c}(D'') = 3$
+  does not satisfy the invariant.
+\end{proof}

### 2018-03-26 doc/notes_v1/formalism.tex
Renaming various docs.

--- 2017-01-24 doc/paper/formalism.tex
@@
 evaluates to true after all variables in $I$ have been replaced by their
-mapping in $D$. Let $\dbs$ denote the set of all database states.
+mapping in $D$.
+%
+For example, if $\var = \set{x, y}$, then $D_1 = \set{x:1, y:2}$ and $D_2 =
+\set{x:42, y:99}$ are both databases. If $I \defeq x \geq 10$, then $D_2$
+satisfies $I$ (i.e. $I(D_2)$) but $D_1$ does not (i.e. $\lnot I(D_1)$).
 
@@
 \]
+For example, $t_1 = \set{x: 42}$ is the transaction which increments $x$ by 42.
+$t_2 = \set{x: 1, y:-2}$ is the transaction which increments $x$ by 1 and
+decrements $y$ by $-2$. If $D = \set{x: 100, y:100}$, then $D \circ t_1 =
+\set{x:142, y:100}$ and $D \circ t_2 = \set{x:101, y:98}$.
+
 Note that transaction application is commutative. That is for all database
 states $D$ and transactions $t_1$ and $t_2$, $D \circ t_1 \circ t_2 = D \circ
-t_2 \circ t_1$.
+t_2 \circ t_1$. Also note that transactions are \emph{blind} in that they do
+not read from the database. We'll relax this assumption in \secref{imp}, but
+for now we're keeping things simple.
 
@@
 $C$ \emph{satisfies invariant} $I$ starting at $D$, denoted $I(C @ D)$, if for
-every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$.
+every prefix $C'$ of $C$, $D \circ C'$ satisfies invariant $I$. Note that $I(C
+@ D)$ implies $I(D)$ (the empty prefix of $C$) and $I(D \circ C)$ (the entire
+prefix of $C$).
 
-A set of transactions $T$ is \iconfluent{} with respect to invariant $I$ if for
-all database states $D$ and chains $C_1$ and $C_2$ created from $T$, if $D$
-satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$ satisfies $I$
-starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
-\todo{
-  This isn't quite the definition of \iconfluence{} Bailis uses. Ponder whether
-  this simpler definition is okay.
-}
+A set of transactions $T$ is \emph{\iconfluent{}}\footnote{An astute reader
+will notice that this isn't actually the definition of \iconfluence{} presented
+in \cite{bailis2014coordination}. It's a simpler one. Rest assured, this
+simpler definition will be hard enough to deal with. We'll explore the
+ramifications of this simplification in \secref{alie}.} with respect to
+invariant $I$ if for all database states $D$ and chains $C_1$ and $C_2$ created
+from $T$, if $D$ satisfies $I$, $C_1$ satisfies $I$ starting at $D$, and $C_2$
+satisfies $I$ starting at $D$, then $D \circ C_1 \circ C_2$ satisfies $I$.
+
+For example, let $\var = \set{x}$ and consider a transactions $t_+ = \set{x:
+1}$ and $t_- = \set{x: -1}$ which increment and decrement $x$ respectively. Let
+$I \defeq x > 0$ be the invariant that the value $x$ in the database is
+positive.
+%
+$\set{t_+}$ is \iconfluent{} with respect to $I$ because given any database $D$
+and chains $C_1$ and $C_2$, if $x$ is positive in $D$, then $x$ will be
+certainly be positive in $D \circ C_1 \circ C_2$.
+
+On the other hand, $\set{t_-}$ is not \iconfluent{} with respect to $I$. We can
+prove this with a simple counterexample. Let
+  $D = \set{x:2}$,
+  $C_1 = t_-$, and
+  $C_2 = t_-$.
+Note that $D$ and $D \circ t_1 = \set{x:1}$ satisfy the invariant, so $I(D)$,
+$I(C_1 @ D)$, and $I(C_2 @ D)$. However, $D \circ C_1 \circ C_2 = \set{x: 0}$
+does not satisfy the invariant. Note that this is just a formalization of the
+proof we saw in \figref{decrement}.

### 2018-03-26 doc/notes_v1/solution.tex
Renaming various docs.

--- 2017-01-24 doc/paper/solution.tex
@@
 \section{A Solution}\label{sec:solution}
-This section posed the question: how do we write an algorithm to decide whether
-a set of PN-Counter increment/decrement transactions are \iconfluent{} with
+This paper posed the question: how do we write an algorithm to decide whether a
+set of PN-Counter increment/decrement transactions are \iconfluent{} with
 respect to an invariant expressed as a formula of linear equations and
-inequalities?  Now that we're armed with an intuitive
-(\secref{counter-overview}) and formal (\secref{counter-question})
-understanding of the problem, a geometric interpretation
-(\secref{counter-geometry}), and a useful lemma (\secref{counter-one}), we're
-in a perfect position to present a solution!
+inequalities?  Now that we're armed with an intuitive (\secref{counters}) and
+formal (\secref{formalism}) understanding of the problem, a geometric
+interpretation (\secref{geometry}), and a useful lemma (\secref{oneisenough}),
+we're in a perfect position to present a solution!
 
@@
   \item
-    \eqref{eq:i1} is $I$ with every variable $x$ replaced with $x + dx_1$.
-    Intuitively, it predicates our implication on the assumption that if we
-    apply the first transaction, our database still satisfies the invariant.
-    \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second transaction
-    instead of the first.
+    \eqref{eq:i1} is $I$ with every variable $\alpha$ replaced with $\alpha +
+    d\alpha_1$.  Intuitively, it predicates our implication on the assumption
+    that if we apply the first transaction, our database still satisfies the
+    invariant.  \eqref{eq:i2} is symmetric to \eqref{eq:i1} but for the second
+    transaction instead of the first.
 
   \item
-    \eqref{eq:i12} is $I$ with every variable $x$ replaced with $x + dx_1 +
-    dx_2$. Intuitively, it says that if we apply both the transactions, we
-    continue to satisfy our invariant.
+    \eqref{eq:i12} is $I$ with every variable $\alpha$ replaced with $\alpha +
+    d\alpha_1 + d\alpha_2$. Intuitively, it says that if we apply both the
+    transactions, we continue to satisfy our invariant.
 \end{itemize}

### 2018-03-26 doc/notes_v1/vision.tex
Renaming various docs.

--- 2016-11-16 doc/vision.tex
@@
 \section{Research Vision}
-
-At a very high level, this research tackles the same problems tackled by many
-existing research projects like BOOM Analytics \cite{alvaro2010boom}, Bloom
-\cite{alvaro2011consistency}, Bloom$^L$ \cite{conway2012logic}, TARDiS
-\cite{crookstardis}, and IPA \cite{holt2016disciplined}:
-
-\begin{quotation}
-  \emph{How can we make writing weakly consistent programs more principled?}
-\end{quotation}
-
-BOOM Analytics, Bloom, and Bloom$^L$ use data-centric declarative programming
-and program analysis based on the CALM theorem. TARDiS exposes the entire
-history of divergent computations to users in a git-like system to empower them
-to programmatically resolve conflicts. IPA uses a type system to ensure weakly
-consistent data doesn't leak into strongly consistent data. \todo{Cite and
-summarize \cite{balegas2015putting}, \cite{roy2015homeostasis},
-\cite{li2012making}, and \cite{li2014automating}}
-
-This system builds off existing research on CRDTs and \iconfluence{} to provide
-a library of CRDTs, a language to express invariants over those CRDTs, the
-ability to transactionally update the replicated CRDTs, and a decision
+This project builds off existing research on CRDTs and \iconfluence{} to
+provide a library of CRDTs, a language to express invariants over those CRDTs,
+the ability to transactionally update the replicated CRDTs, and a decision
 procedure to algorithmically determine whether the transactions are

### 2018-03-26 doc/notes_v2/sufficient_conditions.tex
Renaming various docs.

--- 2018-03-12 doc/paper_redux/sufficient_conditions.tex
@@
 
-Here are some more examples of transactions $T$ and invariants $I$ that are
-\sTIconfluent{} but not \Iclosed{}.
-
-\todo{Think of more compelling examples.}
+Here are some more examples. \todo{Think of more compelling examples.}
 
@@
 
-  Some thought shows that O is \sTIconfluent{}. However, $O$ is not \Iclosed{}. Let
+  Some thought shows that O is \sTIconfluent{}. However, $O$ is not \Iclosed{}.
+  Let
   \[
@@
 \end{example}
+
+\begin{example}
+  Here are some of the example objects and invariants
+  in~\cite{bailis2014coordination}. We begin with invariants over G-Sets $A$
+  and $B$. Fortunately, a good number of invariants are \Iclosed{}.
+
+  \begin{center}
+    \begin{tabular}{|l|l|p{3in}|}
+      \hline
+      $I$ & \Iclosed{} & Description \\\hline\hline
+      %
+      $\forall x \in A.\ \phi(x)$ &
+      yes &
+        Invariants of this form include things like attribute equality and
+        attribute inequality.
+        \\\hline
+      $\exists x \in A.\ \phi(x)$ &
+      yes &
+        Invariants of this form include things like non-emptiness constraints.
+        \\\hline
+      $\forall x \in A.\ \forall y \in A.\ \phi(x, y)$ &
+      no &
+        Invariants of this form include things like uniqueness constraints
+        (e.g. $\forall x \in A.\ \forall y \in A.\ x \neq y \implies key(x)
+        \neq key(y)$).
+        \\\hline
+      $\forall x \in A.\ \forall y \in B.\ \phi(x, y)$ &
+      no &
+        Invariants of this form include things like disjointness (e.g. $\forall
+        x \in A. \forall y \in B.\ x \neq y$).
+        \\\hline
+      $\forall x \in A.\ \exists y \in A.\ \phi(x, y)$ &
+      yes &
+        Invariants of this form include things like foreign key constraints
+        into a single relation. For example, if we have an \texttt{Employee(id,
+        bossid)} relation, we might want the invariant $\forall (i_1, b_1) \in
+        E.\ \exists (i_2, b_2) \in E.\ b_1 \neq \bot \implies b_1 = i_2$.
+        \\\hline
+      $\forall x \in A.\ \exists y \in B.\ \phi(x, y)$ &
+      yes &
+        Invariants of this form include things like foreign key constraints.
+        \\\hline
+      $\exists x \in A.\ \forall y \in A.\ \phi(x, y)$ &
+      no &
+        \todo{Figure out useful interpretation of these formula.}
+        \\\hline
+      $\exists x \in A.\ \forall y \in B.\ \phi(x, y)$ &
+      no &
+        \todo{Figure out useful interpretation of these formula.}
+        \\\hline
+      $\exists x \in A.\ \exists y \in A.\ \phi(x, y)$ &
+      yes &
+        \todo{Figure out useful interpretation of these formula.}
+        \\\hline
+      $\exists x \in A.\ \exists y \in B.\ \phi(x, y)$ &
+      yes &
+        \todo{Figure out useful interpretation of these formula.}
+        \\\hline
+    \end{tabular}
+  \end{center}
+
+  For 2P-Sets $A$ and $B$, we see that things get a bit more grim.
+  \begin{center}
+    \begin{tabular}{|l|l|}
+      \hline
+      $I$ & \Iclosed{} \\\hline\hline
+      %
+      $\forall x \in A.\ \phi(x)$ & yes \\\hline
+      $\exists x \in A.\ \phi(x)$ & no \\\hline
+      $\forall x \in A.\ \forall y \in A.\ \phi(x, y)$ & no \\\hline
+      $\forall x \in A.\ \forall y \in B.\ \phi(x, y)$ & no \\\hline
+      $\forall x \in A.\ \exists y \in A.\ \phi(x, y)$ & no \\\hline
+      $\forall x \in A.\ \exists y \in B.\ \phi(x, y)$ & no \\\hline
+      $\exists x \in A.\ \forall y \in A.\ \phi(x, y)$ & no \\\hline
+      $\exists x \in A.\ \forall y \in B.\ \phi(x, y)$ & no \\\hline
+      $\exists x \in A.\ \exists y \in A.\ \phi(x, y)$ & no \\\hline
+      $\exists x \in A.\ \exists y \in B.\ \phi(x, y)$ & no \\\hline
+    \end{tabular}
+  \end{center}
+  \note{%
+    This would be a good table to demonstrate that \Iclosure{} is not only
+    unnecessary, but also uncommon. In practice, not a lot of things are going
+    to be \Iclosed{}. This would motivate our push beyond these simple checks.
+    But then again, not a lot of things are going to be invariant-confluent
+    either, so it's not a super strong argument.
+  }
+\end{example}

### 2018-03-26 iconfluence/checker.py
Fixed typed_coerce bug.

@@ -57,7 +57,7 @@ class Checker:
 
         # Ensure that our initial start state is well-typed.
         typ = crdt.to_type()
-        e = ast.typed_coerce(coercible, typ)
+        e = ast.coerce(coercible)
         e = typecheck.typecheck_expr(e, {})
         typecheck.assert_type_eq(e.typ, typ, e, e)
 

### 2018-03-29 doc/paper/sections/interactive_decision_procedure.tex
Rough outline + some of iconfluence section.

\section{An Interactive Decision Procedure}
% - describe how main theorem gives us intution about interactive dp
% - present the algorithm
% - walk through an example
% - describe how the algorithm is robust to errors

### 2018-03-29 doc/paper/sections/introduction.tex
Rough outline + some of iconfluence section.

\section{Introduction}
% Hype up problem that invariant confluence solves
% Introduce iconfluence and hype it up as a really good thing.
% Discuss how it's hard to determine iconfluence and a lot of existing
% approaches fall short.
% Describe how we take the problem one step further
% Overview two new approaches
% Describe how often times we're not iconfluent
% Overivew new segmented iconfluence stuff
% contributions
%   - new expression-based definitions of iconfluence
%   - understanding of iclosure and new suff and necc
%   - interactive decision procedure
%   - new suff conditions
%   - segmented iconfluence
%   - implementation in lucy + evaluation...

### 2018-03-29 doc/paper/sections/invariant_closure.tex
Rough outline + some of iconfluence section.

\section{Invariant-Closure}
% - define invariant-closure and introduce it as a sufficient condition
% - cite existing work to show that people have used invariant closure to
%   reason about invariant-confluence
% - explain why invariant closure can fail and why its not sufficient using the
%   two-integers example
% - state main theorem

### 2018-03-29 doc/paper/sections/invariant_confluence.tex
Rough outline + some of iconfluence section.

\section{Invariant-Confluence}
Informally, a replicated object is \defword{invariant-confluent} with respect
to an invariant if every replica of the object is guaranteed to satisfy the
invariant despite the possibility of different replicas being concurrently
modified or merged together~\cite{bailis2014coordination}. In this section, we
make this informal notion of invariant-confluence precise. We begin by
introducing our system model of replicated objects and then present a formal
definition of invariant-confluence.

A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of
states and a binary merge operator $\join: S \times S \to S$ which merges two
states into one. A \defword{transaction} $t: S \to S$ is a function which maps
one state to another. An \defword{invariant} $I$ is a subset of $S$.
Notationally, we write $I(s)$ to denote that $s$ satisfies the invariant (i.e.
$s \in I$) and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant
(i.e. $s \notin I$).
%
For example, $O = (\ints, +)$ is a distributed object consisting of integers
merged by addition; $t(x) = x + 1$ is a transaction which adds one to a state;
and $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
non-negative.

In our system model, a distributed object $O$ is replicated across a set $p_1,
\ldots, p_n$ of $n$ servers. Each server $p_i$ manages a replica $s_i \in S$ of
the replicated object. Every server begins with a start state $s_0 \in S$, a
fixed set $T$ of transactions, and an invariant $I$. Servers can perform one of
two actions.

First, a client can contact a server $p_i$ and request that it execute a
transaction $t \in T$. $p_i$ speculatively executes $t$ transitioning from
state $s_i$ to state $t(s_i)$. If $t(s_i)$ satisfies the invariant---i.e.
$I(t(s_i))$---then $p_i$ commits the transaction and remains in state $t(s_i)$.
Otherwise, $p_i$ aborts the transaction and reverts to state $s_i$.

Second, a server $p_i$ can sent its state $s_i$ to another server $p_j$ with
state $s_j$ causing $p_j$ to transition from state $s_j$  to state $s_i \join
s_j$. Servers periodically merge states with one another in order to keep their
states loosely synchronized\footnote{%
  Notably, if $O$ is a CRDT---i.e. $O$ is a semilattice and every transaction
  $t \in T$ is inflationary---then this periodic merging ensures that $O$ is
  strongly eventually consistent~\cite{shapiro2011conflict}.
}.
Note that unlike with transactions, servers \emph{cannot} abort a merge; server
$p_j$ must transition from $s_j$ to $s_i \join s_j$ whether or not $s_i \join
s_j$ satisfies the invariant.

Informally, $O$ is \defword{invariant-confluent with respect to $s_0$, $T$, and
$I$}, abbreviated \defword{\sTIconfluent{}} if every replica $s_1, \ldots, s_n$
is guaranteed to always satisfy the invariant $I$ in every possible execution
of the system.

To define invariant-confluence formally, we represent a state produced by a
system execution as a simple expression generated by the grammar
\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]
where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
an example, consider the system execution in \figref{SystemExecution} in which
a distributed object is replicated across servers $p_1$, $p_2$, and $p_3$.
Server $p_3$ begins with state $s_0$, transitions to state $s_2$ by executing
transaction $u$, transitions to state $s_5$ by executing transaction $w$, and
then transitions to state $s_7$ by merging with server $p_1$. Similarly, server
$p_1$ ends up with state $s_6$ after executing transactions $t$ and $v$ and
merging with server $p_2$. In \figref{Expression}, we see the abstract syntax
tree of the corresponding expression.

{\input{figures/invariant_confluence_definitions}}

% - define invariant confluence
%   - define system model
%   - define expression based system model
%   - define invariant confluence
%   - Give example with two integers

% TODO: Give a footnote about the difference between this definition and the original definition.

### 2018-03-29 doc/paper/sections/invariant_confluence.tex
Finished invariant-confluence section.

@@
 
-% - define invariant confluence
-%   - define system model
-%   - define expression based system model
-%   - define invariant confluence
-%   - Give example with two integers
+We say an expression $e$ is \defword{$\sTIreachable{}$} if it corresponds to a
+valid execution of our system model. Formally, we define $\sTIreachable{e}$ to
+be the smallest predicate that satisfies the following equations:
+\begin{itemize}
+  \item
+    $\sTIreachable{s_0}$.
+  \item
+    For all expressions $e$ and for all transactions $t$ in the set $T$ of
+    transactions, if $\sTIreachable{e}$ and $I(t(e))$, then
+    $\sTIreachable{t(e)}$.
+  \item
+    For all expressions $e_1$ and $e_2$, if $\sTIreachable{e_1}$ and
+    $\sTIreachable{e_2}$, then $\sTIreachable{e_1 \join e_2}$.
+\end{itemize}
+Similarly, we say a state $s \in S$ is $\sTIreachable{}$ if there exists a
+$\sTIreachable{}$ expression $e$ that evaluates to $s$.
 
-% TODO: Give a footnote about the difference between this definition and the original definition.
+Finally, we say $O$ is \defword{invariant-confluent with respect to $s_0$, $T$,
+and $I$}, abbreviated \defword{\sTIconfluent{}}, if all reachable states
+satisfy the invariant: $\setst{s \in S}{\sTIreachable{s}} \subseteq I$.

### 2018-03-29 doc/paper/sections/introduction.tex
Part of iclosure, part of intro, title ideas.

@@
 \section{Introduction}
+In summary, our contributions are as follows:
+\begin{itemize}
+  \item
+    We propose a novel expression-oriented definition of invariant-confluence
+    that is both formal and simple (\secref{InvariantConfluence}).
+
+  \item
+    We develop conditions under which invariant-closure is both necessary and
+    sufficient for invariant-confluence (\secref{InvariantClosure}).
+
+  \item
+    We exploit these conditions to design an interactive decision procedure for
+    invariant-confluence (\secref{InteractiveDecisionProcedure}).
+
+  \item
+    We again exploit these conditions to design a novel non-trivial sufficient
+    condition for invariant-confluence that does not depend on user interaction
+    (\secref{???}).
+
+  \item
+    We present segmented invariant-confluence: a generalization of
+    invariant-confluence that uses a small amount of coordination to maintain
+    invariants for replicated objects that are otherwise not
+    invariant-confluent (\secref{SegmentedInvariantConfluence}).
+
+  \item
+    We implement and evaluate our theoretical findings in a system called Lucy
+    (\secref{Evaluation}).
+\end{itemize}
+
 % Hype up problem that invariant confluence solves

### 2018-03-29 doc/paper/sections/invariant_closure.tex
Part of iclosure, part of intro, title ideas.

@@
-\section{Invariant-Closure}
+\section{Invariant-Closure}\seclabel{InvariantClosure}
+Our ultimate goal is to write a program that can automatically decide whether a
+given distributed object $O$ is \sTIconfluent{}. Such a program has to
+automatically prove or disprove that every reachable state satisfies the
+invariant. However, automatically reasoning about the possibly infinite set of
+reachable states is challenging, especially because transactions and merge
+functions can be complex and can be interleaved arbitrarily in an execution.
+Due to this complexity, existing systems that aim to automatically decide
+invariant-confluence instead focus on deciding a sufficient condition for
+invariant-confluence---dubbed \defword{invariant-closure}---that is simpler to
+reason about~\cite{li2012making, li2014automating}. In this section, we define
+invariant-closure and study why the sufficient condition is not necessary.
+Armed with this understanding, we present conditions under which it is both
+sufficient and necessary.
+
 % - define invariant-closure and introduce it as a sufficient condition

### 2018-03-29 doc/paper/sections/invariant_confluence.tex
Part of iclosure, part of intro, title ideas.

@@
-\section{Invariant-Confluence}
+\section{Invariant-Confluence}\seclabel{InvariantConfluence}
 Informally, a replicated object is \defword{invariant-confluent} with respect
@@
 satisfy the invariant: $\setst{s \in S}{\sTIreachable{s}} \subseteq I$.
+% TODO: Make a note that this definition is different than the original, and
+% reference the appendix where we include the proof that the two definitions
+% are equivalent.

### 2018-03-29 doc/paper/sections/segmented_invariant_confluence.tex
Part of iclosure, part of intro, title ideas.

\section{Segmented Invariant-Confluence}\seclabel{SegmentedInvariantConfluence}

### 2018-03-29 doc/paper/sections/invariant_closure.tex
Finished invariant-closure section.

@@
 reason about~\cite{li2012making, li2014automating}. In this section, we define
-invariant-closure and study why the sufficient condition is not necessary.
+invariant-closure and study why the condition sufficient but not necessary.
 Armed with this understanding, we present conditions under which it is both
 sufficient and necessary.
+
+We say an object $O = (S, \join)$ is \defword{invariant-closed with respect to
+an invariant $I$}, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
+states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
+$I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
+
+\begin{theorem}\thmlabel{IclosureImpliesIconfluence}
+  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
+  transactions $T$, and an invariant $I$, if $I(s_0)$ and if $O$ is \Iclosed{},
+  then $O$ is \sTIconfluent{}.
+\end{theorem}
+
+\thmref{IclosureImpliesIconfluence} states that invariant-closure is
+sufficient for invariant-confluence. Intuitively, our system model ensures that
+transaction execution preserves the invariant, so if merging states also
+preserves the invariant and if our start state satisfies the invariant, then
+inductively it is impossible for us to reach a state which doesn't satisfy the
+invariant.
+
+This is good news because checking if an object is invariant-closed is more
+straightforward than checking if it is invariant-confluent. Existing systems
+typically use an SMT solver like Z3 to check if an object is
+invariant-closed~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
+is, then by \thmref{IclosureImpliesIconfluence}, it is invariant-confluent.
+Unfortunately, invariant-closure is \emph{not} necessary for
+invariant-confluence, so if an object is not invariant-closed, these systems
+cannot conclude that the object is not invariant-confluent. The reason why
+invariant-closure is not necessary for invariant-confluence is best explained
+through an example.
+
+\begin{example}\examplelabel{Z2}
+  Let $O = (\ints \times \ints, \join)$ consist of pairs $(x, y)$ of integers
+  where $(x_1, y_1) \join (x_2, y_2) = (\max(x_1, x_2), \max(y_1, y_2))$. Our
+  start state $s_0 \in \ints \times \ints$ is the point $(0, 0)$. Our set $T$
+  of transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1, y)$
+  which increments $x$ and $t_{y-1}((x, y)) = (x, y - 1)$ which decrements $y$.
+  Our invariant $I = \setst{(x, y) \in \ints \times \ints}{xy \leq 0}$ consists
+  of all points $(x, y)$ where the product of $x$ and $y$ is non-positive.
+
+  The invariant and the set of reachable states are illustrated in \figref{Z2}
+  in which we draw each state $(x, y)$ as a point in space. The invariant
+  consists of the second and fourth quadrant, and the reachable states consist
+  of the fourth quadrant. From this, it's immediate that the reachable states
+  are a subset of the invariant, so $O$ is invariant-confluent. However,
+  letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
+  invariant-closed. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
+  (1, 1)$, we see $\lnot I(s_3)$.
+\end{example}
+
+{\input{figures/z2}}
+
+In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
+invariant-closed, but $s_1$ is not reachable. This is not particular to
+\exampleref{Z2}. In fact, it is fundamentally the reason why invariant-closure
+is not equivalent to invariant-confluence. Invariant-confluence is, at its
+core, a property of reachable states, but invariant-closure is completely
+ignorant of reachability. As a result, invariant-satisfying yet unreachable
+states like $s_1$ are the key hurdle preventing invariant-closure from being
+equivalent to invariant-confluence. This is formalized by
+\thmref{IclosureEquivalentIconfluence}
+
+\begin{theorem}\thmlabel{IclosureEquivalentIconfluence}
+  Consider an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
+  transactions $T$, and an invariant $I$. If the invariant is a subset of the
+  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachable{s}}$), then
+  \[
+    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
+  \]
+\end{theorem}
+\begin{proof}
+  The forward direction of \thmref{IclosureEquivalentIconfluence} follows
+  immediately from \thmref{IclosureImpliesIconfluence}. The backward direction
+  holds because any two invariant satisfying states $s_1$ and $s_2$ must be
+  reachable (by assumption), so their join $s_1 \join s_2$ is also reachable.
+  And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
+  \join s_2$, satisfy the invariant.
+\end{proof}
 

### 2018-03-29 doc/paper/sections/interactive_decision_procedure.tex
Part of iclosure, part of intro, title ideas.
Progress on interactive decision procedure section.

@@
-\section{An Interactive Decision Procedure}
+\section{An Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
+\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}
+\newcommand{\Helper}{\textsf{Helper}}
+
+\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
+\begin{algorithm}[t]
+  \caption{Interactive Invariant-Confluence Decision Procedure}%
+  \algolabel{InteractiveDecisionProcedure}
+  \begin{algorithmic}
+    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
+    \State
+    \Return $I(s_0)$ and
+    \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
+    \EndFunction
+
+    \State
+
+    \algocomment{$R$ is a set of $\sTIreachable{}$ states.}
+    \algocomment{$NR$ is a set of $\sTIunreachable{}$ states.}
+    \algocomment{$I(s_0)$ is a precondition.}
+    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
+    \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
+    \If {closed}
+      \State \Return true
+    \EndIf
+    \State Augment $R$ and $NR$ with random search and user input
+    \If{$s_1, s_2 \in R$}
+    \Return false
+    \Else\
+    \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
+    \EndIf
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\thmref{IclosureEquivalentIconfluence} tells us that if all
+invariant-satisfying points are reachable, then invariant-closure and
+invariant-confluence are equivalent. In this section, we present the
+interactive invariant-confluence decision procedure shown in
+\algoref{InteractiveDecisionProcedure} which takes advantage of this result.
+
+A user provides the algorithm with an object $O = (S, \join)$, a start state
+$s_0$, a set of transactions $T$, and an invariant $I$\footnote{%
+  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
+  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
+  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
+  0$).
+}. The user then interacts with
+\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
+states from the invariant. Meanwhile, the algorithm leverages an
+invariant-closure decision procedure to either (a) conclude that $O$ is or is
+not \sTIconfluent{} or (b) provide counterexamples to the user to help them
+eliminate unreachable states. After all unreachable states have been eliminated
+from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
+the problem of invariant-confluence directly to the problem of
+invariant-closure, and the algorithm terminates.
+%
+We now describe \algoref{InteractiveDecisionProcedure} in detail. An example of
+how to use \algoref{InteractiveDecisionProcedure} on \exampleref{Z2} is given
+in \figref{DecisionProcedureExample}.
+
+% This example motivates \algoref{InteractiveDecisionProcedure}, a general
+% purpose interactive invariant-confluence decision procedure in which a user
+% iteratively refines the invariant until the decision procedure terminates.
+% Given an object $O$, a start state $s_0$, a set of transactions $T$, and an
+% invariant $I$\footnote{%
+%   Formally, $I$ is set of states (e.g. $\setst{(x, y)}{x \in \ints, y \in
+%   \ints, xy \leq 0}$), but an actual implementation of \IsInvConfluent{} would
+%   likely take in $I$ as a logical formula (e.g. $xy \leq 0$). From now on,
+%   whenever we mention a (possibly infinite) set of states, remember that we can
+%   always represent the set as a finite formula.
+% }, we call \IsInvConfluent$(O, s_0, T, I)$ to determine if $O$ is
+% \sTIconfluent{}. Now, we describe \algoref{InteractiveDecisionProcedure}. An
+% example of how to use \algoref{InteractiveDecisionProcedure} on
+% \exampleref{StateBasedNotNecessary} is given in
+% \figref{DecisionProcedureExample}.
+%
+% \IsInvConfluent{} assumes access to two helper decision procedures:
+% \begin{enumerate}
+%   \item
+%     A decision procedure to determine $I(s_0)$.
+%
+%   \item
+%     An \Iclosed{} decision procedure \IsIclosed$(O, I)$ which returns a triple
+%     (closed, $s_1$, $s_2$). closed is a boolean indicating whether $O$ is
+%     \Iclosed{}. If closed is false, then $s_1$ and $s_2$ are a counterexample
+%     such that $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$.
+% \end{enumerate}
+%
+% Constructing these two helper decision procedures is hard (undecidable, I
+% think), but we can do pretty well using an SMT solver like Z3. There is also
+% some existing research on how to built smart \Iclosed{} decision procedures
+% that we can leverage~\cite{li2014automating}. In practice either decision an
+% return unknown instead of true or false. In this case, \IsInvConfluent{} returns
+% unknown.
+%
+% \IsInvConfluent{} first checks that $s_0$ satisfies $I$. If it doesn't, then
+% \IsInvConfluent{} returns false. If it does then, \IsInvConfluent{} calls a
+% helper function \Helper{} which---in addition to $O$, $s_0$, $T$, and
+% $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states and a set $NR$
+% of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O, s_0, T, I, R,
+% NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and $NR$ are
+% correct). Intuitively, $NR$ helps remove unreachable states from $I$ moving us
+% closer to the case where $I \subseteq \setst{s}{\sTIreachable{s}}$.
+%
+% First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}. There are two
+% cases to consider.
+% \begin{itemize}
+%   \item
+%     If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
+%     \clmref{StateBasedSufficient}, $O$ is \sticonfluent{s_0}{T}{(I -
+%     NR)}, so
+%     $
+%       \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
+%         \subseteq I - NR
+%         \subseteq I
+%     $.
+%     If $NR$ only contains unreachable states, then
+%     $
+%       \setst{s}{\sTIreachable{s}} = \setst{s}{\stireachable{s_0}{T}{(I-NR)}{s}}
+%     $.
+%     Thus,
+%     $
+%       \setst{s}{\sTIreachable{s}} \subseteq I
+%     $.
+%     Thus, $O$ \sTIconfluent{}, so \Helper{} return true.
+%
+%   \item
+%     If \IsIclosed{} determines that $O$ is not \iclosed{(I - NR)}, then we
+%     have a counterexample $s_1$ and $s_2$. We want to determine whether $s_1$
+%     and $s_2$ are $\sTIreachable{}$ or $\sTIunreachable{}$. We can do so in two
+%     ways.
+%
+%     First, we can randomly generate a set of reachable states and add them to
+%     $R$.
+%     %
+%     Second, we can prompt the user to tell us whether or not the states are
+%     $\sTIreachable{}$ or $\sTIunreachable{}$. In addition to labelling $s_1$
+%     and $s_2$ as $\sTIreachable{}$ or not, the user can also refine $I$ by
+%     augmenting $R$ and $NR$ arbitrarily (see \figref{DecisionProcedureExample}
+%     for an example). In this step, we also make sure that $s_0 \notin NR$ since
+%     we know that $s_0$ is $\sTIreachable{}$.
+%
+%     After $s_1$ and $s_2$ have been labelled $\sTIreachable{}$ or not, we
+%     continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1
+%     \join s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{},
+%     so \Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
+%     $\sTIunreachable{}$, so we recurse.
+% \end{itemize}
+%
+% \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
+% procedure. It requires that a user classify states as reachable or unreachable.
+% What happens if a user incorrectly classifies a state? There's two possible
+% errors that can be made:
+% \begin{itemize}
+%   \item \textbf{A user can label an unreachable state as reachable.}
+%     In this case, the $s_1, s_2 \in R$ check in \Helper{} might be erroneously
+%     met. In turn, \IsInvConfluent{} might erroneously return no. This is
+%     inconvenient, but not the end of the world. Moreover, we can modify our
+%     decision procedure so that \Helper{} requires that whenever a user labels a
+%     state $s$ as $\sTIreachable{}$, they must also provide an $\sTIreachable{}$
+%     expression $e$ where $eval(e) = s$.
+%
+%   \item \textbf{A user can label a reachable state as unreachable.}
+%     In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is
+%     not \sTIconfluent. Thus, a user might falsely believe their system to be
+%     \sTIconfluent{} even though it isn't, and eventually their system might
+%     enter a state which violates the invariant.
+%
+%     We can mitigate this in two ways. First, we can aggresively expand $R$. If
+%     a user ever labels a state $s$ as unreachable, but $s \in R$, we can notify
+%     the user of their mistake. Second, \Helper{} returns true when $O$ is
+%     \iclosed{(I - NR)} for some $NR$, so $O$ is \sticonfluent{s_0}{T}{(I-R)}
+%     (even though it might not be \sTIconfluent{}). Thus, when a user deploys
+%     their system, they can deploy with the invariant $I - NR$ instead of $NR$.
+%     If $NR$ is correctly populated, then deploying with $I - NR$ has no effect
+%     on the system. If $NR$ is incorrectly populated, then some
+%     $\sTIreachable{}$ states are artificially made unreachable, but the system
+%     is still guaranteed to never enter a state which violates $I$.
+% \end{itemize}
+%
+
+{\input{figures/decision_procedure_example}}
+
 % - describe how main theorem gives us intution about interactive dp

### 2018-03-29 doc/paper/sections/interactive_decision_procedure.tex
Finished interactive decision procedure section.

--- 2018-03-29 doc/paper/sections/interactive_decision_procedure.tex
@@
-\section{An Interactive Decision Procedure}
-% - describe how main theorem gives us intution about interactive dp
-% - present the algorithm
-% - walk through an example
-% - describe how the algorithm is robust to errors
+\section{An Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
+\newcommand{\Helper}{\textsf{Helper}}
+\newcommand{\IsIclosed}{\textsf{IsIclosed}}
+\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}
+
+\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
+\begin{algorithm}[t]
+  \caption{Interactive Invariant-Confluence Decision Procedure}%
+  \algolabel{InteractiveDecisionProcedure}
+  \begin{algorithmic}
+    \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
+    \State
+    \Return $I(s_0)$ and
+    \Call{Helper}{$O$, $s_0$, $T$, $I$, $\set{s_0}$, $\emptyset$}
+    \EndFunction
+
+    \State
+
+    \algocomment{$R$ is a set of $\sTIreachable{}$ states.}
+    \algocomment{$NR$ is a set of $\sTIunreachable{}$ states.}
+    \algocomment{$I(s_0)$ is a precondition.}
+    \Function{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
+    \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I - NR$}
+    \If {closed}
+      \State \Return true
+    \EndIf
+    \State Augment $R$ and $NR$ with random search and user input
+    \If{$s_1, s_2 \in R$}
+    \Return false
+    \Else\
+    \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
+    \EndIf
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\thmref{IclosureEquivalentIconfluence} tells us that if all
+invariant-satisfying points are reachable, then invariant-closure and
+invariant-confluence are equivalent. In this section, we present the
+interactive invariant-confluence decision procedure shown in
+\algoref{InteractiveDecisionProcedure} which takes advantage of this result.
+
+\subsection{The Decision Procedure}
+A user provides \algoref{InteractiveDecisionProcedure} with an object $O = (S,
+\join)$, a start state $s_0$, a set of transactions $T$, and an invariant
+$I$\footnote{%
+  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
+  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
+  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
+  0$).
+}. The user then interacts with
+\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
+states from the invariant. Meanwhile, the algorithm leverages an
+invariant-closure decision procedure to either (a) conclude that $O$ is or is
+not \sTIconfluent{} or (b) provide counterexamples to the user to help them
+eliminate unreachable states. After all unreachable states have been eliminated
+from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
+the problem of invariant-confluence directly to the problem of
+invariant-closure, and the algorithm terminates.
+%
+We now describe \algoref{InteractiveDecisionProcedure} in detail. An example of
+how to use \algoref{InteractiveDecisionProcedure} on \exampleref{Z2} is given
+in \figref{DecisionProcedureExample}.
+
+\IsInvConfluent{} assumes access to an invariant-closure decision procedure
+\IsIclosed$(O, I)$ which returns a triple (closed, $s_1$, $s_2$). closed is a
+boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
+$s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a counterexample
+witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
+but $\lnot I(s_1 \join s_2)$ (e.g. $s_1$ and $s_2$ from \exampleref{Z2}).  As
+we mentioned earlier, we can (and do) implement the invariant-closure decision
+procedure using an SMT solver like Z3~\cite{de2008z3}.
+
+\IsInvConfluent{} first checks that $s_0$ satisfies the invariant. $s_0$ is
+reachable, so if it does not satisfy the invariant, then $O$ is not
+\sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
+\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
+$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
+and a set $NR$ of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O,
+s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
+$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
+iteratively increased which removes unreachable states from $I$ until $I
+\subseteq \setst{s}{\sTIreachable{s}}$.
+
+First, \Helper{} checks to see if $O$ is \iclosed{(I - NR)}.
+%
+If \IsIclosed{} determines that $O$ is \iclosed{(I - NR)}, then by
+\thmref{IclosureImpliesIconfluence}, $O$ is \sticonfluent{s_0}{T}{I-NR}, so
+\[
+  \setst{s \in S}{\stireachable{s_0}{T}{I-NR}{s}}
+    \subseteq I - NR
+    \subseteq I
+\]
+If $NR$ only contains $\sTIunreachable{}$ states, then the set of
+$\sTIreachable{}$ states is equal to set of $\stireachable{s_0}{T}{I-NR}{}$
+states which, as we just showed, is a subset of $I$.  Thus, $O$ is
+\sTIconfluent{}, so \Helper{} return true.
+
+If \IsIclosed{} determines that $O$ is \emph{not} \iclosed{(I - NR)}, then
+we have a counterexample $s_1, s_2$. We want to determine whether $s_1$ and
+$s_2$ are reachable or unreachable. We can do so in two ways.
+%
+First, we can randomly generate a set of reachable states and add them to
+$R$. If $s_1$ or $s_2$ lies in $R$, then we know they are reachable.
+%
+Second, we can prompt the user to tell us directly whether or not the states
+are reachable or unreachable.
+
+In addition to labelling $s_1$ and $s_2$ as reachable or unreachable, the user
+can also refine $I$ by augmenting $R$ and $NR$ arbitrarily (see
+\figref{DecisionProcedureExample} for an example). In this step, we also make
+sure that $s_0 \notin NR$ since we know that $s_0$ is reachable.
+
+After $s_1$ and $s_2$ have been labelled as $\sTIreachable{}$ or not, we
+continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1 \join
+s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{}, so
+\Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
+$\sTIunreachable{}$, so we recurse.
+
+\Helper{} recurses only when one of $s_1$ or $s_2$ is unreachable, so $NR$
+grows after every recursive invocation of \Helper{}. Similarly, $R$ continues
+to grow as \Helper{} randomly explores the set of reachable states. As the user
+sees more and more examples of unreachable and reachable states, it becomes
+easier and easier for them to recognize patterns that define which states are
+reachable and which are not. As a result, it becomes easier for a user to
+augment $NR$ and eliminate a large number of unreachable states from the
+invariant. Once $NR$ has been sufficiently augmented to the point that $I - NR$
+is a subset of the reachable states, \thmref{IclosureEquivalentIconfluence}
+guarantees that the algorithm will terminate after one more call to \IsIclosed.
+
+\subsection{Tolerating User Error}
+\algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
+procedure. It requires that a user classify states as reachable or unreachable.
+To err is human, so what happens when a user incorrectly classifies a state?
+There are two possible errors that can be made, and
+\algoref{InteractiveDecisionProcedure} can be made robust to both.
+
+\textbf{A user can label an unreachable state as reachable.}
+In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and return
+false, concluding that $O$ is not \sTIconfluent{} even when it is.  This is
+inconvenient, but not catastrophic. We can modify
+\algoref{InteractiveDecisionProcedure} so that \Helper{} requires that whenever
+a user labels a state $s$ as $\sTIreachable{}$, they must also provide an
+$\sTIreachable{}$ expression $e$ that evaluates to $s$. Here, $e$ acts a proof
+that certifies that $s$ is indeed reachable. This increases the burden on the
+user but completely eliminates this class of errors.
+
+\textbf{A user can label a reachable state as unreachable.}
+In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
+\sTIconfluent. Thus, a user might falsely believe their distributed object to
+be \sTIconfluent{} even though it isn't, and eventually one replica of their
+distributed object might enter a state that violates the invariant.
+%
+We can mitigate this in two ways. First, we can aggressively expand $R$. If a
+user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
+user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
+NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
+it might not be \sTIconfluent{}). Thus, when a user replicates their
+distributed object across a set of servers, they can deploy with the invariant
+$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
+states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
+$NR$ is incorrect and contains some $\sTIunreachable{}$ states, then some of
+these states are artificially made unreachable, but the system is still
+guaranteed to never produce a state which violates $I$.
+
+{\input{figures/decision_procedure_example}}

### 2018-03-30 doc/notes_v2/new_statebased_sufficient_v2.tex
Revised merge reduction + some of paper section.

\section{New Sufficient State-Based Conditions}
\newcommand{\stimergereduction}[3]{$\Diamond(#1, #2, #3)$-merge-reduction}
\newcommand{\stimergereducible}[3]{$\Diamond(#1, #2, #3)$-merge-reducible}
\newcommand{\stimergereducibility}[3]{$\Diamond(#1, #2, #3)$-merge-reducibility}
\newcommand{\sTImergereduction}{\stimergereduction{s_0}{T}{I}}
\newcommand{\sTImergereducible}{\stimergereducible{s_0}{T}{I}}
\newcommand{\sTImergereducibility}{\stimergereducibility{s_0}{T}{I}}

\begin{definition}
  A \defword{join-free} expression $e = t_1(t_2(\ldots t_n(s) \ldots))$ is an
  expression that does not contain a join.
\end{definition}

\begin{definition}
  $O$ is \defword{merge-reducible with respect to $s_0$, $T$, and $I$},
  abbreviated \defword{\sTImergereducible{}}, if for every pair $e_1$ and $e_2$
  of join-free $\sTIreachable{}$ expressions, there exists some join-free
  $\sTIreachable{}$ expression $e_3$ such that $eval(e_3) = eval(e1 \join
  e_2)$.
\end{definition}

\begin{claim}\clmlabel{ReducibleImpliesIconfluent}
  If $I(s_0)$ and $O$ is \sTImergereducible, then $O$ is \sTIconfluent.
\end{claim}
\begin{elidableproof}
  Assume that $I(s_0)$ and that $O$ is \sTImergereducible{}. We prove by structural
  induction on $e$, that for all $\sTIreachable{}$ expressions $e$, there
  exists a join-free $\sTIreachable{}$ expression $e'$ such that $eval(e) =
  eval(e')$.
  \begin{itemize}
    \item \textbf{Case 1: $e = s_0$.}
      Trivially, $e' = s_0$.

    \item \textbf{Case 2: $e = t(e_1)$.}
      $e_1$ is $\sTIreachable{}$, so by the inductive hypothesis, there exists
      a join-free $\sTIreachable{}$ expression $e_1'$ such that $eval(e_1) =
      eval(e_1')$. $t(e_1)$ is $\sTIreachable{}$, so $I(t(e_1))$. $eval(e_1) =
      eval(e_1')$, so $I(t(e_1'))$. Thus, $t(e_1')$ is $\sTIreachable{}$ (and join
      free), so we can let $e' = t(e_1')$.

    \item \textbf{Case 3: $e = e_1 \join e_2$.}
      $e_1$ and $e_2$ are $\sTIreachable{}$, so by the inductive hypothesis,
      there exists equivalent join-free $\sTIreachable{}$ expressions $e_1'$
      and $e_2'$. $O$ is \sTImergereducible{}, so it's immediate that there exists
      an $e'$.
  \end{itemize}

  Consider an arbitrary $\sTIreachable{}$ expression $e$ and it's equivalent
  join-free $\sTIreachable{}$ counterpart $e'$. $e'$ is either $s_0$ or
  $t(e'')$.  In either case, it satisfies the invariant, so $O$ is
  \sTIconfluent{}.
\end{elidableproof}

\begin{claim}\clmlabel{NewStateBasedSufficient}
  Consider a state-based object $O = (S, \join)$, a start state $s_0$, a set of
  transactions $T$, and an invariant $I$. $T$ is \sTIconfluent{} if the
  following criteria are met:
  \begin{enumerate}
    \item
      $O$ is a semilattice.

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that $t(s) = s
      \join s_t$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.

    \item
      $I(s_0)$
  \end{enumerate}
\end{claim}
\begin{elidableproof}
  \newcommand{\bart}[1]{\overline{t_{#1}}}
  \newcommand{\baru}[1]{\overline{u_{#1}}}
  \newcommand{\bartu}[2]{\bart{#1}(\baru{#2}(s_0))}

  To show that $T$ is \sTIconfluent{}, it suffices to show that $T$ is
  \sTImergereducible{} (by \clmref{ReducibleImpliesIconfluent}).  Let
  \[
    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
    \quad \text{and} \quad
    e_2 = u_n(u_{m-1}(\ldots (u_1(s_0))\ldots))
  \]
  be two arbitrary join-free $\sTIreachable{}$ expressions that recursively.
  For ease of notation, let
  \[
    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
    \quad\text{and}\quad
    \baru{j} = u_j(\ldots(u_1(s_0))\ldots)
  \]
  We prove by strong induction on $k \in \nats$ that if $k = i + j$ where $0
  \leq i \leq n$ and $0 \leq j \leq m$, $\bartu{i}{j}$ is $\sTIreachable{}$ and
  $eval(\bartu{i}{j}) = eval(\bart{i}(s_0) \join \bart{j}(s_0))$.
  \begin{itemize}
    \item \textbf{Case $k = 0$.}
      $i = j = 0$, so $\bartu{0}{0} = s_0$ which is trivially $\sTIreachable{}$
      and equivalent to $\bart{0}(s_0) \join \baru{0}(s_0) = s_0 \join s_0$
      which evaluates to $s_0$.

    \item \textbf{Case $k = 1$.}
      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
      $
        \bartu{1}{0} = t_1(s_0)
      $ which is $\sTIreachable{}$ because it is a subexpression of $\bart{n}$
      which is $\sTIreachable{}$. Moreover, it is equivalent to $\bart{1}(s_0)
      \join \baru{0}(s_0) = t_1(s_0) \join s_0$ which evaluates to $s_{t_1}
      \join s_0 \join s_0 = s_{t_1} \join s_0 = t_1(s_0)$.

    \item \textbf{Case $k \geq 2$.}
      If $i = 0$, then $j = k$ and $\baru{k}(s_0)$ is $\sTIreachable{}$ because
      it is a subexpression of $\baru{m}(s_0)$. Also, it is equivalent to
      $\bart{0}(s_0) \join \baru{k}(s_0)$ which evaluates to $\baru{k}(s_0)$.
      %
      If $j = 0$, then $i = k$ and $\bart{k}(s_0)$ is $\sTIreachable{}$ because
      it is a subexpression of $\bart{n}(s_0)$. Also, it is equivalent to
      $\bart{k}(s_0) \join \baru{0}(s_0)$ which evaluates to $\bart{k}(s_0)$.

      Otherwise, $i, j > 1$. Let
      \begin{align*}
        e_{i-1,j-1} &= \bartu{i-1}{j-1} \\
        e_{i,j-1} &= \bartu{i}{j-1} \\
        e_{i-1,j} &= \bartu{i-1}{j}
      \end{align*}
      By the inductive hypothesis, $e_{i-1,j-1}$, $e_{i,j-1}$, and $e_{i-1,j}$
      are all $\sTIreachable{}$. By condition 3 (with $s = eval(e_{i-1,j-1})$,
      $t_1$ = $t_i$, and $t_2 = u_j$), $I(e_{i,j-1} \join e_{i-1,j})$. Thus,
      $I(\bartu{i}{j})$, so $\bartu{i}{j}$ is $\sTIreachable{}$.
  \end{itemize}
\end{elidableproof}

### 2018-03-30 doc/paper/sections/merge_reduction.tex
Revised merge reduction + some of paper section.

\section{Merge Reduction}\seclabel{MergeReduction}
In \secref{InvariantClosure}, we discussed how invariant-closure is
fundamentally a property of reachability and that invariant-closure is
sufficient but not necessary for invariant-confluence because it fails to
incorporate any notion of reachability. Using this intuition, we established
\thmref{IclosureEquivalentIconfluence} and then exploited the theorem in
\algoref{InteractiveDecisionProcedure}. In this section, we again take
advantage of this intuition to develop a new sufficient condition for
invariant-confluence that can be checked without user interaction and that
covers some cases not covered by invariant-closure.

An expression $e = t_1(t_2(\ldots(t_n(s))\ldots))$ is \defword{merge-free} if
does not contain any merges (i.e.\ it is generated by the grammar $e ::= s \mid
t(e)$). An object $O = (S, \join)$ is \defword{merge-reducible} with respect to
a start state $s_0 \in S$, a set of transactions $T$, and an invariant $I$,
abbreviated \defword{\sTImergereducible}, if for every pair of $e_1$ and $e_2$
of merge-free $\sTIreachable{}$ expressions, there exists some merge-free
$\sTIreachable{}$ expression $e_3$ that evaluates to the same state as $e_1
\join e_2$. Intuitively, if $O$ is merge-reducible, we can replace $e_1 \join
e_2$ (which has one merge) with $e_3$ (which has no merges) to obtain an
equivalent expression with fewer merges.

\begin{theorem}\thmlabel{ReducibilityImpliesIconfluence}
  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$, if $I(s_0)$ and if $O$ is
  \sTImergereducible{}, then $O$ is \sTIconfluent{}.
\end{theorem}

The proof of \thmref{ReducibilityImpliesIconfluence} is a straightforward
induction. We begin with an $\sTIreachable{}$ expression $e$ and repeatedly
replace any subexpression that merges two merge-free subexpressions with an
equivalent merge-free reachable subexpression (which we can do because $O$ is
merge-reducible). We repeat this process until $e$ has been completely replaced
with an equivalent merge-free reachable expression $e'$. Because $I(s_0)$ and
because our system model only executes transactions that preserve the
invariant, $e'$ (and hence $e$) is guaranteed to satisfy the invariant. Thus,
all reachable states satisfy the invariant, so $O$ is invariant-confluent.

Merge-reducibility is a sufficient condition for invariant-confluence, but
unlike with invariant-closure, it is not straightforward to automatically
determine if an object is merge-reducible. In \thmref{LatticeProperty}, we
outline a sufficient condition for merge-reducibility that is straightforward
to determine automatically.

\begin{theorem}\thmlabel{LatticeProperty}
  Given an object $O = (S, \join)$, a start state $s_0 \in S$, a set of
  transactions $T$, and an invariant $I$, if the following criteria are met,
  then $O$ is \sTIconfluent{}.
  \begin{enumerate}
    \item
      $O$ is a semilattice.

    \item
      For every $t \in T$, there exists some $s_t \in S$ such that for all $s
      \in s$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
      form $t(s) = s \join s_t$ for some constant $s_t$.

    \item
      For every pair of transactions $t_1, t_2 \in T$ and for all states $s \in
      S$, if $I(s)$, $I(t_1(s))$, and $I(t_2(s))$, then $I(t_1(s) \join
      t_2(s))$.

    \item
      $I(s_0)$.
  \end{enumerate}
\end{theorem}

% interactive is nice but requires user feedback; here we develop an alternate suff cond; to do so we introduce blah blah elimination
% merge elimination
% new sufficient condition
% example showing they're not equivalnet
% some discussion taht it's general

### 2018-03-31 doc/paper/sections/interactive_decision_procedure.tex
Minor tweaks.

@@
 states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
-$NR$ is incorrect and contains some $\sTIunreachable{}$ states, then some of
+$NR$ is incorrect and contains some $\sTIreachable{}$ states, then some of
 these states are artificially made unreachable, but the system is still

### 2018-03-31 doc/paper/sections/invariant_closure.tex
Minor tweaks.

@@
 reason about~\cite{li2012making, li2014automating}. In this section, we define
-invariant-closure and study why the condition sufficient but not necessary.
+invariant-closure and study why the condition is sufficient but not necessary.
 Armed with this understanding, we present conditions under which it is both
@@
 
-We say an object $O = (S, \join)$ is \defword{invariant-closed with respect to
-an invariant $I$}, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
+We say an object $O = (S, \join)$ is \defword{invariant-closed} with respect to
+an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
 states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
@@
 preserves the invariant and if our start state satisfies the invariant, then
-inductively it is impossible for us to reach a state which doesn't satisfy the
+inductively it is impossible for us to reach a state that doesn't satisfy the
 invariant.
@@
 Unfortunately, invariant-closure is \emph{not} necessary for
-invariant-confluence, so if an object is not invariant-closed, these systems
-cannot conclude that the object is not invariant-confluent. The reason why
-invariant-closure is not necessary for invariant-confluence is best explained
-through an example.
+invariant-confluence, so if an object is \emph{not} invariant-closed, these
+systems cannot conclude that the object is \emph{not} invariant-confluent. The
+reason why invariant-closure is not necessary for invariant-confluence is best
+explained through an example.
 

### 2018-03-31 doc/paper/sections/invariant_confluence.tex
Finished invariant-closure section.
Minor tweaks.

@@
 First, a client can contact a server $p_i$ and request that it execute a
-transaction $t \in T$. $p_i$ speculatively executes $t$ transitioning from
+transaction $t \in T$. $p_i$ speculatively executes $t$, transitioning from
 state $s_i$ to state $t(s_i)$. If $t(s_i)$ satisfies the invariant---i.e.
@@
 
-Informally, $O$ is \defword{invariant-confluent with respect to $s_0$, $T$, and
-$I$}, abbreviated \defword{\sTIconfluent{}} if every replica $s_1, \ldots, s_n$
+Informally, $O$ is \defword{invariant-confluent} with respect to $s_0$, $T$, and
+$I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots, s_n$
 is guaranteed to always satisfy the invariant $I$ in every possible execution
@@
 merging with server $p_2$. In \figref{Expression}, we see the abstract syntax
-tree of the corresponding expression.
+tree of the corresponding expression for state $s_7$.
 
@@
 
-Finally, we say $O$ is \defword{invariant-confluent with respect to $s_0$, $T$,
-and $I$}, abbreviated \defword{\sTIconfluent{}}, if all reachable states
+Finally, we say $O$ is \defword{invariant-confluent} with respect to $s_0$,
+$T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
 satisfy the invariant: $\setst{s \in S}{\sTIreachable{s}} \subseteq I$.

### 2018-03-31 doc/paper/figures/merge_reducible.tex
Finished merge reduction section.

\begin{figure}[t]
  \centering
  \begin{tikzpicture}[scale=0.3]
    \draw[rounded corners] (0, 0) rectangle (16, 8);
    \draw[rounded corners] (1, 2.5) rectangle (11, 6.5);
    \draw[rounded corners] (6, 0.5) rectangle (15, 4.5);
    \draw[rounded corners] (7, 1.5) rectangle ++(7, 2);
    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\small Invariant-Confluent};
    \node[inner sep=0pt, anchor=north] at (6, 6) {\small Invariant-Closed};
    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\small Criteria (1) -- (4)};
    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\small Merge-Reducible};
  \end{tikzpicture}
  \caption{%
    The relationship between invariant-closure, merge-reducibility, criteria
    (1) -- (4) from \thmref{LatticeProperty}, and invariant-confluence.
  }\figlabel{IclosureVsReducible}
\end{figure}

### 2018-03-31 doc/paper/sections/merge_reduction.tex
Minor tweaks.
Finished merge reduction section.

@@
       For every $t \in T$, there exists some $s_t \in S$ such that for all $s
-      \in s$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
+      \in S$, $t(s) = s \join s_t$. That is, every transaction $t$ is of the
       form $t(s) = s \join s_t$ for some constant $s_t$.
@@
 
-% interactive is nice but requires user feedback; here we develop an alternate suff cond; to do so we introduce blah blah elimination
-% merge elimination
-% new sufficient condition
-% example showing they're not equivalnet
-% some discussion taht it's general
+The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
+through (4) are met, then $O$ is \sTImergereducible{}. From this, it's
+immediate that $O$ is \sTIconfluent{} by
+\thmref{ReducibilityImpliesIconfluence}. To show that $O$ is
+\sTImergereducible, we consider two merge-free reachable expressions
+\[
+  e_1 = t_1(\ldots(t_n(s_0))\ldots)
+  \quad\text{and}\quad
+  e_2 = u_1(\ldots(u_m(s_0))\ldots)
+\]
+Then, we inductively use criterion (3) with criteria (1) and (2) to show that
+\[
+  e_3 = t_1(\ldots(
+          t_n(u_1(\ldots(u_m(s_0))\ldots))
+        )\ldots)
+\]
+is reachable and evaluates to $e_1 \join e_2$.
+
+\thmref{IclosureImpliesIconfluence} states that invariant-closure is a
+sufficient condition for invariant-confluence, and \thmref{LatticeProperty}
+states that criteria (1) -- (4) are sufficient conditions for
+invariant-confluence. How do these sufficient conditions relate to one another?
+Clearly, not all invariant-closed objects are semilattices, so
+invariant-closure does not imply criteria (1) -- (4). Conversely, there are some
+objects which satisfy criteria (1) -- (4) which are not invariant-closed. Here's
+one example.
+
+\begin{example}\examplelabel{TwoSets}
+  Let $O = (\mathcal{P}(\nats), \cup)$ where $\mathcal{P}(\nats)$ is the power
+  set of the natural numbers. Our start state $s_0 = \set{0}$ is the set of
+  $0$. Let $t_Y(X) = X \cup Y$ be the transaction which unions $Y$ with its
+  argument $X$. Our set $T = \setst{t_Y}{Y \subseteq \nats}$ of transactions
+  consists of all possible $t_Y$. Our invariant $I$ consists of all non-empty
+  sets $X$ that contain only even or only odd elements. That is, $I = \setst{X
+  \subseteq 2\nats}{X \neq \emptyset} \cup \setst{X \subseteq 2\nats + 1}{X
+  \neq \emptyset}$.
+
+  Clearly, criteria (1), (2), and (4) are satisfied. Criterion (3) is also
+  satisfied. If $s$ is an arbitrary state that satisfies $I$, then it is
+  non-empty and contains, without loss of generality, only even integers. If
+  $t_1$ and $t_2$ are arbitrary transactions such that $I(t_1(s))$ and
+  $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty and contain only
+  even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly non-empty and contains
+  only even integers, so $I(t_1(s) \join t_2(s))$.
+  %
+  However, $O$ is not \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$.
+  Then, $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$,
+  $\lnot I(s_3)$.
+\end{example}
+
+{\input{figures/merge_reducible}}
+
+Invariant-closure is not necessary for invariant-confluence because it fails to
+incorporate any notion of reachability. Criteria (1) -- (4) are also
+unnecessary, but they can be used to prove that some non-invariant-closed
+objects are invariant-confluent because the criteria \emph{do} incorporate
+notions of reachability. In particular, criterion (3) is a slight variant of
+invariant-closure; it also states that invariant-satisfying states should be
+closed under merge. The fundamental difference is that criterion (3) restricts
+its attention to the merge of two states which are \emph{reachable} from a
+common ancestor state.
+
+In \exampleref{TwoSets}, we saw this fundamental difference rear its head. $O$
+is not $I$-closed because the union of an odd-only set with an even-only set is
+a set with both odd and even integers. However, if we begin in an
+invariant-satisfying state, we cannot reach both an odd-only and even-only set.
+Criterion (3) is able to recognize this fact and conclude that $O$ is
+invariant-confluent despite it not being invariant-closed.
+
+% TODO: Discuss generality?

### 2018-04-01 doc/paper/sections/invariant_confluence.tex
Work on segmented invariant-confluence.

@@
 
+\subsection{System Model}
 A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of
@@
 the replicated object. Every server begins with a start state $s_0 \in S$, a
-fixed set $T$ of transactions, and an invariant $I$. Servers can perform one of
-two actions.
+fixed set $T$ of transactions, and an invariant $I$. Servers repeatedly perform
+one of two actions.
 
@@
 
+\subsection{Expression-Based Formalism}
 To define invariant-confluence formally, we represent a state produced by a

### 2018-04-01 doc/paper/sections/segmented_invariant_confluence.tex
Work on segmented invariant-confluence.

@@
 \section{Segmented Invariant-Confluence}\seclabel{SegmentedInvariantConfluence}
+If a distributed object is invariant-confluent, then the object can be
+replicated without the need for any form of coordination to maintain the
+object's invariant. But what if the object is \emph{not} invariant-confluent?
+In this section, we present a generalization of invariant-confluence called
+\defword{segmented invariant-confluence} that can be used to maintain the
+invariants of non-invariant-confluent objects, requiring only a small amount of
+coordination.
+
+The main idea behind segmented invariant-confluence is to segment the state
+space into a number of segments and restrict the set of allowable transactions
+within each segment in such a way that the object is invariant-confluent
+\emph{within each segment} (even though it may not be globally
+invariant-confluent). Then, servers can run coordination-free within a segment
+and need only coordinate when transitioning from one segment to another. We now
+formalize this idea by defining segmented invariant-confluence and a new system
+model to execute segmented invariant-confluent objects.
+
+\subsection{Formalism}
+Consider a distributed object $O = (S, \join)$, a start state $s_0 \in S$, a
+set of transitions $T$, and an invariant $I$. A segmentation $\Sigma = (I_1,
+T_1), \ldots, (I_n, T_n)$ is a sequence (not a set) of $n$ segments$(I_i, T_i)$
+where every $T_i$ is a subset of $T$. $O$ is \defword{segmented
+invariant-confluent} with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
+\defword{\sTISconfluent}, if the following conditions hold:
+\begin{itemize}
+  \item
+    The start state satisfies the invariant (i.e. $I(s_0)$).
+
+  \item
+    $I$ is covered by the invariants in $\Sigma$ (i.e. $I = \cup_{i=1}^n I_i$).
+    Note that invariants in $\Sigma$ do \emph{not} have to be disjoint. Thus,
+    they do not have to partition $I$; they just have to cover $I$.
+
+  \item
+    For every $(I_i, T_i) \in \Sigma$ and for every state $s \in I_i$, $O$ is
+    \sticonfluent{s}{T_i}{I_i}.
+\end{itemize}
+
+{\input{figures/segmented_z2}}
+
+\begin{example}\examplelabel{SegmentedZ2}
+  Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
+  = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
+  \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
+  \emph{not} \sTISconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
+  reachable, and merging these points yields $(42, 42)$ which violates the
+  invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1, T_1)$, $(I_2,
+  T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
+  \begin{align*}
+    I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
+    I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
+    I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
+    I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
+  \end{align*}
+  $\Sigma$ is illustrated in \figref{SegmentedZ2}. Clearly, $s_0$ satisfies the
+  invariant, and $I_1, I_2, I_3, I_4$ cover $I$. Moreover, for every $(I_i,
+  T_i) \in \Sigma$, we see that $O$ is \iclosed{I_i}, so $O$ is
+  \sticonfluent{s}{T_i}{I_I} for every $s \in I_i$. Thus, $O$ is
+  \sTISconfluent{}.
+\end{example}
+
+\subsection{System Model}
+Now, we describe a new system model used to replicate a segmented
+invariant-confluent object with coordination-free execution within a segment
+and a small amount of coordination to transition between segments. As before,
+we replicate an object $O$ across a set $p_1, \ldots, p_n$ of $n$ servers each
+of which manages a replica $s_i \in S$ of the object. Every server begins with
+$s_0$, $T$, $I$, and $\Sigma$. Moreover, at any given point in time, a server
+designates one of the segments in $\Sigma$ as its \defword{active segment}.
+Initially, every server chooses the first segment $(I_i, T_i) \in \Sigma$ such
+that $I_i(s_0)$ to be its active segment. We'll see momentarily the
+significance of the active segment.
+
+As before, servers repeatedly perform one of two actions: execute a transaction
+or merge states with another server. Merging states in the new system model is
+identical to merging states in the old system model. A server $p_i$ sends its
+state $s_i$ to another server $p_j$ which \emph{must} merge $s_i$ into its
+state $s_j$.
+%
+Transaction execution in the new system model, on the other hand, is a bit more
+involved. Consider a server $s_i$ with active segment $(I_i, T_i)$. A client
+can request that $p_i$ execute a transaction $t \in T$. We consider what
+happens when $t \in T_i$ and $t \notin T_i$ separately.
+
+If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
+execute $t$. During global coordination, every server temporarily stops
+processing and transitions to state $s = s_1 \join \ldots \join s_n$, the join
+of every server's state. Then, every server speculatively executes $t$
+transitioning to state $t(s)$. If $t(s)$ violates the invariant (i.e.  $\lnot
+I(t(s))$), then every server aborts $t$ and reverts to state $s$.  Then, $p_i$
+replies to the client. If $t(s)$ satisfies the invariant (i.e. $I(t(s))$), then
+every server commits $t$ and remains in state $t(s)$. Every server then chooses
+the first segment $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new
+active segment. Note that such a segment is guaranteed to exist because the
+segment invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server
+will deterministically pick the same active segment. In fact, an invariant of
+the system model is that at any given point of normal processing, every server
+has the same active segment.
+
+$t \in T_i$, then $p_i$ executes $t$ immediately and without coordination. If
+$t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then $p_i$
+commits $t$, stays in state $t(s_0)$, and replies to the client. If $t(s_i)$
+violates the invariant (i.e.  $i.e. \lnot I(t(s_i))$), then $p_i$ aborts $t$,
+reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies the
+invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
+I_i(t(s_i))$), then $p_i$ reverts to state $s_i$ and initiates a round of
+global coordination to execute $t$, as described in the previous paragraph.
+Transaction execution is summarized in \algoref{TxnExecution}.
+
+\begin{algorithm}[t]
+  \caption{Transaction execution in new system model}%
+  \algolabel{TxnExecution}
+  \begin{algorithmic}
+    \If{$t \notin T_i$}
+      \State Execute $t$ with global coordination
+    \Else{}
+      \If{$I_i(t(s_i))$} Commit $t$
+      \ElsIf{$\lnot I(t(s_i))$} Abort $t$
+      \Else{} Execute $t$ with global coordination
+      \EndIf{}
+    \EndIf{}
+  \end{algorithmic}
+\end{algorithm}
+
+% why does this work
+
+\subsection{Discussion}
+% generalizes locking
+% contiuous generalization of invariant-confluence
+
+% The motivate the main idea behind segmented invariant-confluence, let's revisit
+% \exampleref{Z2} but with a different start state.
+%
+% \begin{example}
+%   Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
+%   = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
+%   \exampleref{Z2}, but now let our start state $s_0$ be $(-10, 10)$. Now, $O$
+%   is \emph{not} \sTIconfluent{}. From $(-10, 10)$, we can reach the points $(0,
+%   10)$ and $(10, 0)$, and merging these together we reach $(10, 10)$ which
+%   violates the invariant.
+%   %
+%   However, all is not lost. If we're determined to make $O$
+%   invariant-confluent, we can do so in one of two ways.
+%   \begin{enumerate}
+%     \item
+%       We can modify the invariant. $O$ may not be \sTIconfluent, but it is
+%       \sticonfluent{s_0}{T}{\setst{(x, y)}{x \leq 0, y \geq 0}}. If we eliminate
+%       the fourth quadrant from the invariant, $O$ becomes invariant-confluent.
+%     \item
+%       We can disallow some transactions. $O$ may not be \sTIconfluent, but
+%       it is \sticonfluent{s_0}{\set{t_{x+1}}}{I}. If we disallow the $t_{y-1}$
+%       transaction, $O$ becomes invariant-confluent.
+%   \end{enumerate}
+%   While these two modifications are possible, neither is desirable.
+Artificially making some invariant-satisfying states unreachable---either by restricting the invariant or disallowing some transactions---is most likely a bit of a hack.
+Plus,
+% \end{example}

### 2018-04-01 doc/paper/sections/segmented_invariant_confluence.tex
Finished segmented invariant-confluence section.

@@
 
-% why does this work
+This system model guarantees that all replicas of a segmented
+invariant-confluent object always satisfy the invariant. All servers begin in
+the same initial state and with the same active segment. Thus, because $O$ is
+invariant-confluent with respect to every segment, servers can execute
+transactions within the active segment without any coordination and guarantee
+that the invariant is never violated. Any operation that would violate the
+assumptions of invariant-confluence system model (e.g.\ executing a transaction
+that's not permitted in the active segment or executing a permitted transaction
+that leads to a state outside the active segment) triggers a global
+coordination. Globally coordinated transactions are only executed if they
+maintain the invariant. Moreover, if a globally coordinated transaction leads
+to another segment, the coordination ensures that all servers begin in the same
+start state and with the same active replica, reestablishing the assumptions of
+the invariant-confluence system model.
 
 \subsection{Discussion}
-% generalizes locking
-% contiuous generalization of invariant-confluence
+There are a couple of things worth noting about segmented invariant-confluence
+and its system model. First, invariant-confluence is a very coarse-grained
+property. If an object is invariant-confluent, then we can replicate it with no
+coordination. If it is not invariant-confluent, then we have no guarantees.
+There's no in-between. Segmented invariant-confluence, on the other hand, is a
+much more fine-grained property that can be applied to applications with
+varying degrees of complexity.  Segmented invariant-confluence provides
+guarantees to complex applications that require a large number of segments and
+to simple applications with a smaller number of segments, whereas
+invariant-confluence only provides guarantees to applications that can be
+segmented into a single segment.
 
-% The motivate the main idea behind segmented invariant-confluence, let's revisit
-% \exampleref{Z2} but with a different start state.
-%
-% \begin{example}
-%   Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
-%   = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
-%   \exampleref{Z2}, but now let our start state $s_0$ be $(-10, 10)$. Now, $O$
-%   is \emph{not} \sTIconfluent{}. From $(-10, 10)$, we can reach the points $(0,
-%   10)$ and $(10, 0)$, and merging these together we reach $(10, 10)$ which
-%   violates the invariant.
-%   %
-%   However, all is not lost. If we're determined to make $O$
-%   invariant-confluent, we can do so in one of two ways.
-%   \begin{enumerate}
-%     \item
-%       We can modify the invariant. $O$ may not be \sTIconfluent, but it is
-%       \sticonfluent{s_0}{T}{\setst{(x, y)}{x \leq 0, y \geq 0}}. If we eliminate
-%       the fourth quadrant from the invariant, $O$ becomes invariant-confluent.
-%     \item
-%       We can disallow some transactions. $O$ may not be \sTIconfluent, but
-%       it is \sticonfluent{s_0}{\set{t_{x+1}}}{I}. If we disallow the $t_{y-1}$
-%       transaction, $O$ becomes invariant-confluent.
-%   \end{enumerate}
-%   While these two modifications are possible, neither is desirable.
-Artificially making some invariant-satisfying states unreachable---either by restricting the invariant or disallowing some transactions---is most likely a bit of a hack.
-Plus,
-% \end{example}
+Second, segmented invariant-confluence can naturally model distributed locking.
+One approach to replicating non-invariant-confluent objects is to recognize
+which transactions cannot be safely executed concurrently and then require them
+to acquire a distributed lock before executing~\cite{balegas2015putting,
+gotsman2016cause}. For example, in a banking application with the invariant
+that all balances remain positive, concurrent deposits are permitted, but
+concurrent withdrawals can lead to invariant violations, thus forcing
+withdrawals to first acquire a distributed lock before executing. To model this
+locking with segmented invariant-confluence, we simply remove lock-acquiring
+transactions from every segment's set of transactions. Consequently, executing
+these transactions are executed with global coordination, achieving the same
+effect as acquiring a distributed lock. In fact, segmented invariant-confluence
+can be used to model more fine-grained locking. Omitting a transaction from
+only certain segments is analogous to requiring a transaction to acquire a
+distributed lock only when certain conditions are met.
+
+% TODO: Forward reference examples in the evaluation section.
+
+Third, we can integrate a couple of optimizations into our system model to
+further reduce the amount of coordination required. First, if a server with
+state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
+to execute, and $t(s_i)$ violates the active invariant but not the global
+invariant, instead of initiating a round of global coordination, $p_i$ can
+simply buffer $t$ for re-execution at a later time. While this increases the
+latency required to execute $t$, it's possible that after other transactions
+are executed, re-executing $t$ may lead to a state that either satisfies the
+active invariant or violates the global invariant. In either case, a round of
+global coordination is avoided. Similarly, servers can buffer transactions that
+require global coordination, executing an entire batch of these transactions
+during a single round of global coordination.

### 2018-04-01 doc/paper/sections/introduction.tex
Finished intro.

--- 2018-03-29 doc/paper/sections/introduction.tex
@@
 \section{Introduction}
+When an application designer decides to replicate a piece of data, they have to
+make a fundamental choice between weak and strong consistency. Replicating the
+data with strong consistency using a technique like state machine
+replication~\cite{schneider1990implementing, lamport1998part,
+liskov2012viewstamped, ongaro2014search} makes the application designer's life
+very easy. To the application developer, a strongly consistent system behaves
+exactly like a single-threaded system running on a single node, so reasoning
+about the behavior of the system is simple. Unfortunately, strong consistency
+is at odds with performance. The CAP theorem and PACELC theorem tell us that
+strongly consistent systems suffer from higher latency at best and
+unavailability at worst~\cite{brewer2012cap, abadi2012consistency}. On the
+other hand, weak consistency models like eventual~\cite{vogels2009eventually},
+PRAM~\cite{lipton1988pram}, and causal~\cite{ahamad1995causal} consistency
+allow data to be replicated with high availability and low latency, but they
+put a tremendous burden on the application designer to reason about the complex
+interleavings of operations that are allowed by these weak consistency models.
+In fact, weak consistency models strip an application developer of one of the
+earliest and most effective tools that is used to reason about the execution of
+programs: application invariants~\cite{hoare1969axiomatic}. Even if every
+transaction executing in a weakly consistent system maintains an application
+invariant, the system as a whole can produce invariant-violating states.
+
+% When people want to replicate data, they have to make a choice between strong and weak consistency.
+When choosing strong consistency, everything makes sense, reasoning is easy.
+But CAP and the latency one tell us that the avilability and latency of our system will be reduced.
+If we choose weak consistency, then things run really fast, but we can no longer reason about the application.
+In fact, we risk the possibility one our greatest tools that we use to reason about software: invariants.
+With weak consistency models , even if every transaction preserves the invariant, the invariant may still be broken when stuff is interleaved.
+
+Is it possible for us to have our strongly consistent cake and eat it with high
+availability too? Can we replicate a piece of data with weak consistency but
+still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
+introduced the notion of invariant-confluence as a necessary and sufficient
+condition for when invariants can be maintained over replicated data without
+the need for any coordination~\cite{bailis2014coordination}. If an object is
+invariant-confluent with respect to an invariant, we can replicate it with the
+performance benefits of weak consistency and (some of) the correctness benefits
+of strong consistency. However, the task of identifying whether or not an
+object actually is invariant-confluent remain a challenge. Bailis et al.\
+manually categorized a set of common objects, transactions, and invariants
+(e.g.\ foreign key constraints on relations, linear constraints on integers) as
+invariant-confluent or not, but ideally we would have a general purpose program
+that can automatically determine invariant-confluence for us. \textbf{The main
+thrust of this paper is to make progress towards a general-purpose
+invariant-confluence decision procedure.}
+
+% Is it possible to have our cake and eat it to?
+Can we maintain invariants without coordination?
+For some applications, yes.
+Invariant confluence (cite bailis) introduced a necessary and sufficient condition for when its possible to have coordination free replication while maintiang invariants.
+If somethign is iconfluent, it can be replicated wihtout coordination.
+However, determining whethere something is invariant-confluent is a very challening task.
+Bailis recognized some patterns of things which are invariant confluent, but ideally we would have a general purpose program to automatically determine it for us.
+The ultimate goal of this paper is to...
+In this paper, we make progress towards a more-general purpose iconfluence decision procedure, improving on the current state-of-the-art approaches.
+
+Designing a general-purpose invariant-confluence decision procedure is
+difficult\footnote{Actually, it's impossible. Determining whether an object is
+invariant-confluent is, in general, undecidable. Still, we can develop a
+decision procedure that works well in the common case.}. Many existing
+approaches instead develop a decision procedure for invariant-closure, a
+sufficient (but not necessary) condition for
+invariant-confluence~\cite{li2012making, li2014automating}. The existing
+approaches check to see if an object is invariant-closed. If it is, then they
+conclude that it is also invariant-confluent. If it's not, then the current
+approaches are unable to conclude anything about whether or not the object is
+invariant-confluent. In this paper, we take a step back and study the
+underlying reason \emph{why} invariant-closure is not necessary for
+invariant-confluence. We find that invariant-confluence is fundamentally a
+property about reachability, about whether certain values of our replicated
+object can actually be obtained in the execution of a system. We discover that
+unreachable, yet invariant-satisfying, states cause invariant-closure to be an
+unnecessary condition. Using this understanding, we construct a set of modest
+conditions under which invariant-closure and invariant-confluence are in fact
+\emph{equivalent}, allowing us to reduce the problem of determining
+invariant-confluence to that of determining invariant-closure.
+
+% This is hard.
+Not only undecidable, but also not really an easy thing to do.
+Existing systems often prove a sufficient condition called invariant-closure for invariant-confluence.
+So if the object is invariant-closed, they determine its invariant-confluent, but it its not invariant-closed, the system doesn't know anythign.
+In this paper, we study why the condition is not necessary, finding that the fundamental problem is that of unreachable invariant satisfying states.
+Using this insight, we develop general conditions under which invariant-closure is both necessary and sufficient.
+
+Then, we use these conditions to design a general-purpose interactive
+invariant-confluence decision procedure. Users iteratively interact with the
+decision procedure, eliminating unreachable states from the invariant and
+moving towards the conditions under which invariant-closure and
+invariant-confluence are equivalent. Meanwhile, the decision-procedure
+generates example reachable and unreachable states which help the user
+recognize patterns that describe reachability. We also leverage our intuitions
+about reachability to develop a new sufficient condition for
+invariant-confluence, dubbed merge-reducibility. Merge-reducibility covers some
+cases that are not covered by invariant-closure, and it can be automatically
+checked without any user interaction.
+
+% Then, we use our insight to develop a general purpose interactive invariant-confluence decision procedure.
+Users continuously interact with the decision procedure to eleiminate unreachabel states from the invariant.
+Meanwhile, the decision procedure provides examples to the user that they can use to identify which points are reachable and which are not.
+This decision procedure can be used to show lots of things are invariant-confluent, and is also robust to user error.
+
+% While the general purpose interaactive dp is good, it does rely on user input.
+We additionally invent a new sufficient condition for invariant-confluence known as merge-reducbility which exploits the insigths from the previous theory.
+This new sufficient condition can handle cases that invariant-closure cannot and doesn't require user feedback.
+
+Next, we take a step beyond invariant-confluence and develop a generalization
+of invariant-confluence called segmented invariant-confluence. While
+invariant-confluence characterizes objects which can be replicated
+\emph{without any} coordination, segmented invariant-confluence allows us to
+replicate non-invariant-confluent objects with \emph{minimal coordination}. The
+main idea is that invariant-satisfying states are divided into a number of
+segments, and the set of allowable of transactions within each segment is
+restricted. Within a segment, servers act without any coordination,
+synchronizing only to transition across segment boundaries. This design
+highlights the trade-off between application complexity and
+coordination-freedom; more complex applications require more segments which
+require more coordination, and vice-versa.
+
+% Finally, we address the elephant in the room and ask the question, what if something is not invariant-confluent though?
+We develop a generalization of invariant-confluence called segmented invariant-confluence.
+While invariant-confluence only characterizes objects which can be replicated without any coordination, segmented invariant-confluence allows us to replicate objects with minimal coordination.
+This approach also generalizes existing lock based approaches to handling non-invariant-confluent stuff.
+
+% Finally, we implement all the decision procedures and the system models in a system called Lucy.
+We perform an evaluation to look at when the decision procedures are effective and measure the performance benefits of our implemetnation.
+Overall,
+
+In closing, here is an outline of the paper and of our contributions:
+\begin{itemize}
+  \item
+    We propose a novel expression-oriented definition of invariant-confluence
+    that is both formal and simple (\secref{InvariantConfluence}).
+
+  \item
+    We develop an understanding of why invariant-closure is not necessary and
+    use this understanding to develop conditions under which it is both
+    necessary and sufficient (\secref{InvariantClosure}).
+
+  \item
+    We exploit these conditions to design a general-purpose interactive
+    decision procedure for invariant-confluence
+    (\secref{InteractiveDecisionProcedure}).
+
+  \item
+    We again exploit these conditions to design a novel non-trivial sufficient
+    condition for invariant-confluence, merge-reducibility, that does not
+    depend on user interaction (\secref{MergeReduction}).
+
+  \item
+    We present segmented invariant-confluence: a generalization of
+    invariant-confluence that uses a small amount of coordination to maintain
+    invariants for replicated objects that are otherwise not
+    invariant-confluent (\secref{SegmentedInvariantConfluence}).
+
+  \item
+    We implement our decision procedures and sufficient conditions for both
+    invariant-confluence and segmented invariant-confluence in a system called
+    Lucy, then use Lucy to evaluate the efficacy or our methods
+    (\secref{Evaluation}).
+\end{itemize}
+
 % Hype up problem that invariant confluence solves

### 2018-04-01 doc/paper/sections/introduction.tex
Small tweak to intro.

@@
 invariant-confluence decision procedure. Users iteratively interact with the
-decision procedure, eliminating unreachable states from the invariant and
-moving towards the conditions under which invariant-closure and
-invariant-confluence are equivalent. Meanwhile, the decision-procedure
-generates example reachable and unreachable states which help the user
-recognize patterns that describe reachability. We also leverage our intuitions
-about reachability to develop a new sufficient condition for
-invariant-confluence, dubbed merge-reducibility. Merge-reducibility covers some
-cases that are not covered by invariant-closure, and it can be automatically
-checked without any user interaction.
+decision procedure, eliminating unreachable states from the invariant, moving
+towards the conditions under which invariant-closure and invariant-confluence
+are equivalent. Meanwhile, the decision-procedure generates example reachable
+and unreachable states which help the user recognize patterns that describe
+reachability. We also leverage our intuitions about reachability to develop a
+new sufficient condition for invariant-confluence, dubbed merge-reducibility.
+Merge-reducibility covers some cases that are not covered by invariant-closure,
+and it can be automatically checked without any user interaction.
 

### 2018-04-01 doc/paper/sections/abstract.tex
Finished first cut of abstract.

\begin{abstract}
Strongly consistent systems are easy to reason about but face fundamental
limitations in availability and performance. Weakly consistent systems can be
implemented with very high performance, but place a burden on the application
developer to reason about complex interleavings of execution.
Invariant-confluence provides a formal framework for understanding when we can
get the best of both worlds. An invariant-confluent object can be efficiently
replicated with no coordination needed to preserve its invariants.  However,
actually determining whether or not an object is invariant-confluent is
challenging (undecidable). In this paper, we establish conditions under which a
commonly used sufficient condition for invariant confluence is both necessary
and sufficient. We use this condition to design a general-purpose interactive
invariant-confluence decision procedure and a novel sufficient condition which
can be checked automatically. We then take a step beyond invariant-confluence
and introduce a generalization of invariant-confluence, called segmented
invariant-confluence, which allows us to replicate non-invariant-confluent
objects with a small and tunable amount of coordination. Implementing our
theoretical findings in a system called Lucy, we evaluate our methods on a
number of case studies and benchmarks showing that... TODO: Perform evaluation
:P.
% Strongly consistent systems are easy to reason about but are slow. Weakly consistent systems are fast but hard to reason about. Invariant-confluence gives a formal framework to reason about when data can be replicated without any coordination needed to maintain invraints, the best of borth worlds. However, determining whether an object is invariant-confluent is hard. In this paper, we establish conditions under which a commonly used sufficient condition for invariant confluence is both necessary and usfficient. We use this condition to design a general-purpose interactive decision procedure and a novel sufficient condition which can be checked automatically. We then take a step beyond invariant-confluence and introduce a generalization of invariant-confuluence, called segmented invariant-confluence which allows to replicate objects with a minimal amount of coordination that are otherwise not invariant-confluent. We implement all this in a system called lucy and look at a bunch of stuff.
\end{abstract}

### 2018-04-05 doc/paper/sections/evaluation.tex
Documented segmented invariant-confluence DP.

\section{Evaluation}\seclabel{Evaluation}
\subsection{Lucy}
\todo{Describe implementation of lucy.}

\subsection{Case Studies}
\todo{Bank account with and without decrements.}
\todo{Auction example.}
\todo{Foreign keys with and without deletions.}

\subsection{TPC and Feral}
\todo{See which of the Feral and TPC transactions we can determine automatically}

\subsection{Segmented Invariant Confluence}
\todo{one graph: number of locked operations x axis; throughput on the y-axis}
\todo{one graph: number of segments on the x-axis; throughput on the y-axis}
\todo{compare with no coordination and "VR"}

### 2018-04-05 doc/paper/sections/introduction.tex
Finished first cut of abstract.
Documented segmented invariant-confluence DP.

@@
 invariant-confluent or not, but ideally we would have a general purpose program
-that can automatically determine invariant-confluence for us. \textbf{The main
-thrust of this paper is to make progress towards a general-purpose
+that can automatically determine invariant-confluence for us. \textbf{The first
+main thrust of this paper is to make progress towards a general-purpose
 invariant-confluence decision procedure.}
@@
 
-Next, we take a step beyond invariant-confluence and develop a generalization
-of invariant-confluence called segmented invariant-confluence. While
-invariant-confluence characterizes objects which can be replicated
-\emph{without any} coordination, segmented invariant-confluence allows us to
-replicate non-invariant-confluent objects with \emph{minimal coordination}. The
-main idea is that invariant-satisfying states are divided into a number of
-segments, and the set of allowable of transactions within each segment is
-restricted. Within a segment, servers act without any coordination,
-synchronizing only to transition across segment boundaries. This design
-highlights the trade-off between application complexity and
-coordination-freedom; more complex applications require more segments which
-require more coordination, and vice-versa.
+\textbf{The second main thrust of this paper is to take a step beyond
+invariant-confluence and develop a generalization of invariant-confluence
+called segmented invariant-confluence.} While invariant-confluence
+characterizes objects which can be replicated \emph{without any} coordination,
+segmented invariant-confluence allows us to replicate non-invariant-confluent
+objects with \emph{minimal coordination}. The main idea is that
+invariant-satisfying states are divided into a number of segments, and the set
+of allowable of transactions within each segment is restricted. Within a
+segment, servers act without any coordination, synchronizing only to transition
+across segment boundaries. This design highlights the trade-off between
+application complexity and coordination-freedom; more complex applications
+require more segments which require more coordination, and vice-versa.
 

### 2018-04-05 doc/paper/sections/segmented_invariant_confluence.tex
Documented segmented invariant-confluence DP.

@@
 \section{Segmented Invariant-Confluence}\seclabel{SegmentedInvariantConfluence}
+\newcommand{\IsIclosed}{\textsf{IsIclosed}}
+
 If a distributed object is invariant-confluent, then the object can be
@@
 and need only coordinate when transitioning from one segment to another. We now
-formalize this idea by defining segmented invariant-confluence and a new system
-model to execute segmented invariant-confluent objects.
+formalize segmented invariant-confluence, describe the system model we use to
+replicate segmented invariant-confluent objects, and introduce a segmented
+invariant-confluence decision procedure.
 
@@
 set of transitions $T$, and an invariant $I$. A segmentation $\Sigma = (I_1,
-T_1), \ldots, (I_n, T_n)$ is a sequence (not a set) of $n$ segments$(I_i, T_i)$
-where every $T_i$ is a subset of $T$. $O$ is \defword{segmented
-invariant-confluent} with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
+T_1), \ldots, (I_n, T_n)$ is a sequence of $n$ segments $(I_i, T_i)$ where
+every $T_i$ is a subset of $T$. $O$ is \defword{segmented invariant-confluent}
+with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
 \defword{\sTISconfluent}, if the following conditions hold:
@@
   \item
-    For every $(I_i, T_i) \in \Sigma$ and for every state $s \in I_i$, $O$ is
+    $O$ is invariant-confluent within each segment. That is, for every $(I_i,
+    T_i) \in \Sigma$ and for every state $s \in I_i$, $O$ is
     \sticonfluent{s}{T_i}{I_i}.
@@
 \subsection{System Model}
-Now, we describe a new system model used to replicate a segmented
-invariant-confluent object with coordination-free execution within a segment
-and a small amount of coordination to transition between segments. As before,
-we replicate an object $O$ across a set $p_1, \ldots, p_n$ of $n$ servers each
-of which manages a replica $s_i \in S$ of the object. Every server begins with
-$s_0$, $T$, $I$, and $\Sigma$. Moreover, at any given point in time, a server
-designates one of the segments in $\Sigma$ as its \defword{active segment}.
-Initially, every server chooses the first segment $(I_i, T_i) \in \Sigma$ such
-that $I_i(s_0)$ to be its active segment. We'll see momentarily the
-significance of the active segment.
+Now, we describe the system model used to replicate a segmented
+invariant-confluent object without any coordination within a segment and with
+only a small amount of coordination when transitioning between segments. As
+before, we replicate an object $O$ across a set $p_1, \ldots, p_n$ of $n$
+servers each of which manages a replica $s_i \in S$ of the object. Every server
+begins with $s_0$, $T$, $I$, and $\Sigma$. Moreover, at any given point in
+time, a server designates one of the segments in $\Sigma$ as its
+\defword{active segment}.  Initially, every server chooses the first segment
+$(I_i, T_i) \in \Sigma$ such that $I_i(s_0)$ to be its active segment. We'll
+see momentarily the significance of the active segment.
 
 As before, servers repeatedly perform one of two actions: execute a transaction
-or merge states with another server. Merging states in the new system model is
-identical to merging states in the old system model. A server $p_i$ sends its
-state $s_i$ to another server $p_j$ which \emph{must} merge $s_i$ into its
-state $s_j$.
+or merge states with another server. Merging states in the segmented
+invariant-confluence system model is identical to merging states in the
+invariant-confluence system model. A server $p_i$ sends its state $s_i$ to
+another server $p_j$ which \emph{must} merge $s_i$ into its state $s_j$.
 %
@@
 \begin{algorithm}[t]
-  \caption{Transaction execution in new system model}%
+  \caption{Transaction execution in segmented invariant-confluence system model}%
   \algolabel{TxnExecution}
@@
 
+\subsection{Interactive Decision Procedure}
+In order to determine whether or not an object $O$ is \sTISconfluent{} we have
+to determine whether or not $O$ is invariant-confluent within each segment
+$(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
+\sticonfluent{s}{T_i}{I_i} confluent for every state $s \in I_i$. Ideally, we
+could leverage \algoref{InteractiveDecisionProcedure}, invoking it once per
+segment. Unfortunately, \algoref{InteractiveDecisionProcedure} can only be used
+to determine if $O$ is \sticonfluent{s}{T_i}{I_i} for a \emph{particular} state
+$s \in I_i$, not for \emph{every} state $s \in I_i$. Thus, we would have to
+invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times fore every segment
+$(I_i, T_I)$, which is clearly infeasible given that $I_i$ can be large or even
+infinite.
+
+Instead, we develop a new decision procedure that can be used to determine if
+an object is \sticonfluent{s}{T}{I} for every state $s \in I$. To do so, we
+need to extend the notion of reachability to a notion of coreachability and
+then generalize \thmref{IclosureEquivalentIconfluence}. Two states $s_1, s_2
+\in I$ are \defword{coreachable} with respect to $T$ and $I$, abbreviated
+\defword{$\TIcoreachable{s_1, s_2}$}, if there exists some state $s_0 \in I$
+such that $s_1$ and $s_2$ are both $\sTIreachable{}$.
+
+\begin{theorem}\thmlabel{ExtendedIclosureEquivalentIconfluence}
+  Consider an object $O = (S, \join)$, a set of transactions $T$, and an
+  invariant $I$. If every pair of states in the invariant are coreachable (i.e.
+  $\forall s_1, s_2 \in S.\ \TIcoreachable{s_1, s_2}$), then
+  \[
+    \text{$O$ is \Iclosed{}}
+    \iff
+    \text{$O$ is \sticonfluent{s}{T}{I} for every $s \in I$}
+  \]
+\end{theorem}
+\begin{proof}
+  The proof of the forward direction is exactly the same as the proof of
+  \thmref{IclosureImpliesIconfluence}. Transactions always maintain the
+  invariant, so if merge does as well, then every reachable state must satisfy
+  the invariant.
+  %
+  For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
+  The two points are $\TIcoreachable{}$, so there exists some state $s_0$ from
+  which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
+  $\sTIreachable{}$, so it satisfies the invariant.
+\end{proof}
+
+Using \thmref{ExtendedIclosureEquivalentIconfluence}, we develop
+\algoref{ArbitraryStartInteractiveDecisionProcedure}: a natural generalization
+of \algoref{InteractiveDecisionProcedure}.
+\algoref{InteractiveDecisionProcedure} iteratively refines the set of
+\emph{reachable} states whereas
+\algoref{ArbitraryStartInteractiveDecisionProcedure} iteratively refines the
+set of \emph{coreachable} states, but otherwise, the core of the two algorithms
+is the same.\footnote{%
+  Another small difference is that \IsIclosed{} behaves differently in
+  \algoref{InteractiveDecisionProcedure} and
+  \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
+  \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
+  triple (closed, $s_1$, $s_2$) where if closed is false, then $s_1, s_2 \in I$
+  are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
+  \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
+  $s_2$ are null.
+}
+Now, a segmented invariant-confluence decision procedure, can simply invoke
+\algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.
+
+\newcommand{\algocomment}[1]{\State \textcolor{flatdenim}{\texttt{//} #1}}
+\begin{algorithm}[t]
+  \caption{%
+    Interactive invariant-confluence decision procedure for arbitrary start
+    state $s \in I$
+  }%
+  \algolabel{ArbitraryStartInteractiveDecisionProcedure}
+  \begin{algorithmic}
+    \algocomment{Return if $O$ is \sticonfluent{s}{T}{I} for every $s \in I$.}
+    \Function{IsInvConfluent}{$O$, $T$, $I$}
+      \State \Return \Call{Helper}{$O$, $T$, $I$, $\emptyset$, $\emptyset$}
+    \EndFunction
+
+    \State
+
+    \algocomment{$R$ is a set of $\TIcoreachable{}$ states.}
+    \algocomment{$NR$ is a set of $\TIcounreachable{}$ states.}
+    \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
+      \State closed, $s_1$, $s_2$ $\gets$ \Call{IsIclosed}{$O$, $I$, $NR$}
+      \If {closed}
+        \State \Return true
+      \EndIf
+      \State Augment $R$ and $NR$ with random search and user input
+      \If{$(s_1, s_2) \in R$}
+        \Return false
+      \Else\
+        \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
+      \EndIf
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\begin{example}\examplelabel{CounreachableExample}
+  Let $O = (\ints^3 \times \ints^3, \join)$ be a
+  PN-Counter~\cite{shapiro2011comprehensive} replicated on $3$ machines. Every
+  state $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ represents the integer $(p_1 +
+  p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement the counter, the
+  $i$th server increments $p_i$ and $n_i$ respectively, and $\join$ computes an
+  element-wise maximum. Our start state $s_0 = (0, 0, 0), (0, 0, 0)$; our set
+  $T$ of transactions consists of increment and decrement; and our invariant
+  $I$ is that the value of $s$ is non-negative.
+
+  Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
+  with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
+  1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
+  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
+  concurrent decrements, which we can forbid in a simple one-segment
+  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
+  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
+  \IsIclosed{} again returns false with states $s_1$ and $s_2$. This time,
+  however, the user recognizes that the two states are not
+  $\ticoreachable{T^+}{I}{}$. The user refines $NR$ with the observation that
+  two states are coreachable if and only if they have the same values of $n_1,
+  n_2, n_3$. After this, \IsIclosed{} and
+  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
+\end{example}
+
 \subsection{Discussion}
-There are a couple of things worth noting about segmented invariant-confluence
-and its system model. First, invariant-confluence is a very coarse-grained
-property. If an object is invariant-confluent, then we can replicate it with no
-coordination. If it is not invariant-confluent, then we have no guarantees.
-There's no in-between. Segmented invariant-confluence, on the other hand, is a
-much more fine-grained property that can be applied to applications with
-varying degrees of complexity.  Segmented invariant-confluence provides
-guarantees to complex applications that require a large number of segments and
-to simple applications with a smaller number of segments, whereas
-invariant-confluence only provides guarantees to applications that can be
-segmented into a single segment.
+There are a couple of things worth noting about segmented invariant-confluence,
+its system model, and its decision procedure. First, invariant-confluence is a
+very coarse-grained property. If an object is invariant-confluent, then we can
+replicate it with no coordination. If it is not invariant-confluent, then we
+have no guarantees. There's no in-between. Segmented invariant-confluence, on
+the other hand, is a much more fine-grained property that can be applied to
+applications with varying degrees of complexity. Segmented
+invariant-confluence provides guarantees to complex applications that require a
+large number of segments and to simple applications with a smaller number of
+segments, whereas invariant-confluence only provides guarantees to applications
+that can be segmented into a single segment.
 
-Second, segmented invariant-confluence can naturally model distributed locking.
-One approach to replicating non-invariant-confluent objects is to recognize
-which transactions cannot be safely executed concurrently and then require them
-to acquire a distributed lock before executing~\cite{balegas2015putting,
-gotsman2016cause}. For example, in a banking application with the invariant
-that all balances remain positive, concurrent deposits are permitted, but
-concurrent withdrawals can lead to invariant violations, thus forcing
-withdrawals to first acquire a distributed lock before executing. To model this
-locking with segmented invariant-confluence, we simply remove lock-acquiring
-transactions from every segment's set of transactions. Consequently, executing
-these transactions are executed with global coordination, achieving the same
-effect as acquiring a distributed lock. In fact, segmented invariant-confluence
-can be used to model more fine-grained locking. Omitting a transaction from
-only certain segments is analogous to requiring a transaction to acquire a
-distributed lock only when certain conditions are met.
-
-% TODO: Forward reference examples in the evaluation section.
+Second, segmented invariant-confluence naturally subsumes a distributed locking
+approach to replicating non-invariant-confluent objects.  This approach first
+recognizes which transactions cannot be safely executed concurrently and then
+requires them to acquire a distributed lock before
+executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
+banking application with the invariant that all balances remain non-negative,
+concurrent deposits are permitted, but concurrent withdrawals can lead to
+invariant violations. Thus, we require that withdrawals acquire a distributed
+lock before executing. This example is exactly the same as
+\exampleref{CounreachableExample} which we handled by simply removing
+withdrawal transactions from our segmentation's set of transactions.
 
@@
 during a single round of global coordination.
+
+Fourth, a segmented invariant-confluence decision procedure can also leverage
+\thmref{LatticeProperty} in addition to
+\algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
+criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
+I$.
+
+\todo{Forward reference examples in the evaluation section.}

### 2018-04-06 lucy/cluster.cc
[WIP]

@@ -1,0 +1,27 @@
+#include "cluster.h"
+
+#include "google/protobuf/io/zero_copy_stream_impl.h"
+#include "google/protobuf/text_format.h"
+
+#include <fstream>
+
+Cluster::Cluster(const std::string& filename) {
+  ClusterProto proto;
+  std::ifstream file(filename);
+  ::google::protobuf::io::IstreamInputStream stream(&file);
+  ::google::protobuf::TextFormat::Parse(&stream, &proto);
+  for (const HostPortProto& host_port : proto.host_port()) {
+    servers_.push_back(HostPort(host_port));
+  }
+}
+
+Cluster::Cluster(const ClusterProto& proto) {
+  for (const HostPortProto& host_port : proto.host_port()) {
+    servers_.push_back(HostPort(host_port));
+  }
+}
+
+Cluster::Cluster(std::vector<HostPort> servers)
+    : servers_(std::move(servers)) {}
+
+const std::vector<HostPort>& Cluster::Servers() const { return servers_; }

### 2018-04-06 lucy/cluster.h

@@ -1,0 +1,21 @@
+#ifndef CLUSTER_H_
+#define CLUSTER_H_
+
+#include <string>
+#include <vector>
+
+#include "cluster.pb.h"
+#include "host_port.h"
+
+class Cluster {
+ public:
+  Cluster(const std::string& filename);
+  Cluster(const ClusterProto& proto);
+  Cluster(std::vector<HostPort> servers);
+  const std::vector<HostPort>& Servers() const;
+
+ private:
+  std::vector<HostPort> servers_;
+};
+
+#endif  //  CLUSTER_H_

### 2018-04-06 lucy/comparable_by_key.h

@@ -1,0 +1,30 @@
+#ifndef COMPARABLE_BY_KEY_H_
+#define COMPARABLE_BY_KEY_H_
+
+template <typename T, typename K>
+class ComparableByKey {
+ public:
+  bool operator==(const ComparableByKey<T, K>& other) const {
+    return Key() == other.Key();
+  }
+  bool operator!=(const ComparableByKey<T, K>& other) const {
+    return Key() != other.Key();
+  }
+  bool operator<(const ComparableByKey<T, K>& other) const {
+    return Key() < other.Key();
+  }
+  bool operator<=(const ComparableByKey<T, K>& other) const {
+    return Key() <= other.Key();
+  }
+  bool operator>(const ComparableByKey<T, K>& other) const {
+    return Key() > other.Key();
+  }
+  bool operator>=(const ComparableByKey<T, K>& other) const {
+    return Key() >= other.Key();
+  }
+
+ protected:
+  virtual K Key() const = 0;
+};
+
+#endif  //  COMPARABLE_BY_KEY_H_

### 2018-04-06 lucy/gossip_server.cc

@@ -1,0 +1,5 @@
+#include "gossip_server.h"
+
+void GossipServer::Run() {
+  // TODO: Implement.
+}

### 2018-04-06 lucy/gossip_server.h

@@ -1,0 +1,14 @@
+#ifndef PAXOS_SERVER_H_
+#define PAXOS_SERVER_H_
+
+#include "server.h"
+
+class GossipServer : public Server {
+ public:
+  GossipServer(Cluster cluster, int index, Object* object)
+      : Server(std::move(cluster), index, object) {}
+
+  void Run() override;
+};
+
+#endif  // PAXOS_SERVER_H_

### 2018-04-06 lucy/host_port.cc

@@ -1,0 +1,9 @@
+#include "host_port.h"
+
+HostPort::HostPort(std::string host, std::uint16_t port)
+    : host_(std::move(host)), port_(port) {}
+
+HostPort::HostPort(const HostPortProto& proto)
+    : host_(proto.host()), port_(proto.port()) {}
+
+HostPortKey HostPort::Key() const { return HostPortKey(host_, port_); }

### 2018-04-06 lucy/host_port.h

@@ -1,0 +1,32 @@
+#ifndef HOST_PORT_H_
+#define HOST_PORT_H_
+
+#include <cstdint>
+#include <iostream>
+#include <string>
+#include <utility>
+
+#include "cluster.pb.h"
+#include "comparable_by_key.h"
+
+using HostPortKey = std::pair<const std::string&, const std::uint16_t>;
+class HostPort : public ComparableByKey<HostPort, HostPortKey> {
+ public:
+  HostPort(std::string host, std::uint16_t port);
+  HostPort(const HostPortProto& proto);
+
+  friend std::ostream& operator<<(std::ostream& out,
+                                  const HostPort& host_port) {
+    out << host_port.host_ << ":" << host_port.port_;
+    return out;
+  }
+
+ protected:
+  HostPortKey Key() const override;
+
+ private:
+  const std::string host_;
+  const std::uint16_t port_;
+};
+
+#endif  // HOST_PORT_H_

### 2018-04-06 lucy/object.h

@@ -1,0 +1,12 @@
+#ifndef OBJECT_H_
+#define OBJECT_H_
+
+#include <string>
+
+class Object {
+ public:
+  virtual void RunTransaction(const std::string& txn) = 0;
+  virtual void Merge(const std::string& o) = 0;
+};
+
+#endif  // OBJECT_H_

### 2018-04-06 lucy/paxos_server.cc

@@ -1,0 +1,5 @@
+#include "paxos_server.h"
+
+void PaxosServer::Run() {
+  // TODO: Implement.
+}

### 2018-04-06 lucy/paxos_server.h

@@ -1,0 +1,14 @@
+#ifndef PAXOS_SERVER_H_
+#define PAXOS_SERVER_H_
+
+#include "server.h"
+
+class PaxosServer : public Server {
+ public:
+  PaxosServer(Cluster cluster, int index, Object* object)
+      : Server(std::move(cluster), index, object) {}
+
+  void Run() override;
+};
+
+#endif  // PAXOS_SERVER_H_

### 2018-04-06 lucy/segmented_server.cc

@@ -1,0 +1,5 @@
+#include "segmented_server.h"
+
+void SegmentedServer::Run() {
+  // TODO: Implement.
+}

### 2018-04-06 lucy/segmented_server.h

@@ -1,0 +1,14 @@
+#ifndef PAXOS_SERVER_H_
+#define PAXOS_SERVER_H_
+
+#include "server.h"
+
+class SegmentedServer : public Server {
+ public:
+  SegmentedServer(Cluster cluster, int index, Object* object)
+      : Server(std::move(cluster), index, object) {}
+
+  void Run() override;
+};
+
+#endif  // PAXOS_SERVER_H_

### 2018-04-06 lucy/server.cc

@@ -1,0 +1,4 @@
+#include "server.h"
+
+Server::Server(Cluster cluster, int index, Object* object)
+    : cluster_(std::move(cluster)), index_(index), object_(object) {}

### 2018-04-06 lucy/server.h

@@ -1,0 +1,18 @@
+#ifndef SERVER_H_
+#define SERVER_H_
+
+#include "cluster.h"
+#include "object.h"
+
+class Server {
+ public:
+  Server(Cluster cluster, int index, Object* object);
+  virtual void Run() = 0;
+
+ protected:
+  const Cluster cluster_;
+  const int index_;
+  Object* object_;
+};
+
+#endif  // SERVER_H_

### 2018-04-06 lucy/two_ints.cc

@@ -1,0 +1,15 @@
+#include "two_ints.h"
+
+#include "two_ints.pb.h"
+
+void TwoInts::RunTransaction(const std::string& s) {
+  two_ints::Transaction txn;
+  bool success = txn.ParseFromString(s);
+  (void)success;
+}
+
+void TwoInts::Merge(const std::string& o) {
+  (void)o;
+  (void)x_;
+  (void)y_;
+}

### 2018-04-06 lucy/two_ints.h

@@ -1,0 +1,18 @@
+#ifndef TWO_INTS_H_
+#define TWO_INTS_H_
+
+#include <cstdint>
+
+#include "object.h"
+
+class TwoInts : public Object {
+ public:
+  void RunTransaction(const std::string& txn) override;
+  void Merge(const std::string& o) override;
+
+ private:
+  std::int64_t x_ = 0;
+  std::int64_t y_ = 0;
+};
+
+#endif  //  TWO_INTS_H_

### 2018-04-06 lucy/two_ints_main.cc

@@ -1,0 +1,25 @@
+#include <cstdlib>
+#include <iostream>
+#include <memory>
+
+namespace {
+
+std::string Usage() {
+  return "./two_ints_main <cluster_file> <index> <paxos|segmented|gossip>";
+}
+
+}  // namespace
+
+int main(int argc, char* argv[]) {
+  if (argc != 4) {
+    std::cerr << Usage() << std::endl;
+    return EXIT_FAILURE;
+  }
+
+  const std::string cluster_filename = argv[0];
+  const int index = std::stoi(argv[1]);
+  const std::string mode = argv[2];
+  (void)cluster_filename;
+  (void)index;
+  (void)mode;
+}

### 2018-04-06 lucy/udp_socket.h

@@ -1,0 +1,36 @@
+#ifndef UDP_SOCKET_H_
+#define UDP_SOCKET_H_
+
+class UdpAddress {
+  const std::string host;
+  const unsigned short int port host;
+};
+
+class UdpSocket {
+ public:
+  UdpSocket();
+  void SendTo(const std::string& host, unsigned short int port);
+  RecvFrom(const std::string& host, unsigned short int port);
+
+ private:
+};
+
+class Server {
+ private:
+  void Run(cluster_config, int index, gossip_duration, );
+
+ protected:
+  virtual void ExecuteTransaction(const std::string& txn) = 0;
+  virtual void Merge(const std::string& o) = 0;
+};
+
+class TwoIntsServer : public Server {
+ public:
+  void ExecuteTransaction(const std::string& txn) override;
+  void Merge(const std::string& o) override;
+};
+
+TwoIntsServer s(...);
+s.Run();
+
+#endif  // UDP_SOCKET_H_

### 2018-04-06 lucy/echo_client.cc
Finished UDP.

@@ -1,0 +1,30 @@
+#include <iostream>
+#include <string>
+
+#include "glog/logging.h"
+
+#include "host_port.h"
+#include "udp.h"
+
+int main(int, char* argv[]) {
+  google::InitGoogleLogging(argv[0]);
+
+  UdpAddress server_addr(HostPort("localhost", 9000));
+  UdpSocket socket = UdpSocket();
+
+  std::string line;
+  std::cout << "> " << std::flush;
+  while (std::getline(std::cin, line)) {
+    // Send.
+    socket.SendTo(line, server_addr);
+
+    // Receive.
+    std::string msg;
+    UdpAddress addr;
+    socket.RecvFrom(&msg, &addr);
+    std::cout << "[" << addr << "] " << msg << std::endl;
+
+    // Prompt.
+    std::cout << "> " << std::flush;
+  }
+}

### 2018-04-06 lucy/echo_server.cc

@@ -1,0 +1,25 @@
+#include <iostream>
+#include <string>
+
+#include "glog/logging.h"
+
+#include "host_port.h"
+#include "udp.h"
+
+int main(int, char* argv[]) {
+  google::InitGoogleLogging(argv[0]);
+
+  UdpAddress server_addr(HostPort("localhost", 9000));
+  UdpSocket socket = UdpSocket(server_addr);
+
+  while (true) {
+    // Receive.
+    std::string msg;
+    UdpAddress client_addr;
+    socket.RecvFrom(&msg, &client_addr);
+    std::cout << "[" << client_addr << "] " << msg << std::endl;
+
+    // Send.
+    socket.SendTo(msg, client_addr);
+  }
+}

### 2018-04-06 lucy/host_port.cc

@@ -6,4 +6,8 @@ HostPort::HostPort(std::string host, std::uint16_t port)
 HostPort::HostPort(const HostPortProto& proto)
     : host_(proto.host()), port_(proto.port()) {}
 
+const std::string& HostPort::Host() const { return host_; }
+
+std::uint16_t HostPort::Port() const { return port_; }
+
 HostPortKey HostPort::Key() const { return HostPortKey(host_, port_); }

### 2018-04-06 lucy/host_port.h

@@ -14,6 +14,8 @@ class HostPort : public ComparableByKey<HostPort, HostPortKey> {
  public:
   HostPort(std::string host, std::uint16_t port);
   HostPort(const HostPortProto& proto);
+  const std::string& Host() const;
+  std::uint16_t Port() const;
 
   friend std::ostream& operator<<(std::ostream& out,
                                   const HostPort& host_port) {

### 2018-04-06 lucy/two_ints_main.cc

@@ -17,7 +17,7 @@ int main(int argc, char* argv[]) {
   }
 
   const std::string cluster_filename = argv[0];
-  const int index = std::stoi(argv[1]);
+  const int index = std::atoi(argv[1]);
   const std::string mode = argv[2];
   (void)cluster_filename;
   (void)index;

### 2018-04-06 lucy/udp.cc

@@ -1,0 +1,79 @@
+#include <cstring>
+#include <string>
+
+#include "glog/logging.h"
+
+#include "udp.h"
+
+namespace {
+
+constexpr std::size_t MAX_MESSAGE_SIZE = 65536;
+
+}  // namespace
+
+UdpAddress::UdpAddress() {}
+
+UdpAddress::UdpAddress(const struct sockaddr_in& addr) : addr_(addr) {}
+
+UdpAddress::UdpAddress(const HostPort& host_port) {
+  const std::string node = host_port.Host();
+  const std::string service = std::to_string(host_port.Port());
+  struct addrinfo hints;
+  hints.ai_flags = 0;
+  hints.ai_family = AF_UNSPEC;
+  hints.ai_socktype = SOCK_DGRAM;
+  addr_ = GetAddrInfo(node.c_str(), service.c_str(), &hints);
+}
+
+const struct sockaddr* UdpAddress::SockAddr() const {
+  return reinterpret_cast<const struct sockaddr*>(&addr_);
+}
+
+int UdpAddress::SockAddrLen() const { return sizeof(addr_); }
+
+struct sockaddr_in UdpAddress::GetAddrInfo(const char* node,
+                                           const char* service,
+                                           struct addrinfo* hints) {
+  struct addrinfo* res;
+  int status = getaddrinfo(node, service, hints, &res);
+  CHECK_EQ(status, 0) << gai_strerror(status);
+  CHECK_EQ(res->ai_family, AF_INET) << res->ai_family;
+  struct sockaddr_in addr =
+      *reinterpret_cast<struct sockaddr_in*>(res->ai_addr);
+  freeaddrinfo(res);
+  return addr;
+}
+
+UdpSocket::UdpSocket() {
+  socket_ = socket(AF_INET, SOCK_DGRAM, /*protocol=*/0);
+  CHECK_NE(socket_, -1) << std::strerror(socket_);
+  int n = 1;
+  int status = setsockopt(socket_, SOL_SOCKET, SO_REUSEADDR,
+                          reinterpret_cast<char*>(&n), sizeof(n));
+  CHECK_NE(status, -1) << std::strerror(status);
+}
+
+UdpSocket::UdpSocket(const UdpAddress& addr) : UdpSocket() {
+  int status = bind(socket_, addr.SockAddr(), addr.SockAddrLen());
[…]

### 2018-04-06 lucy/udp.h

@@ -1,0 +1,47 @@
+#ifndef UDP_H_
+#define UDP_H_
+
+#include <iostream>
+
+#include "arpa/inet.h"
+#include "netdb.h"
+#include "netinet/in.h"
+#include "sys/socket.h"
+#include "sys/types.h"
+
+#include "host_port.h"
+
+class UdpAddress {
+ public:
+  UdpAddress();
+  UdpAddress(const struct sockaddr_in& addr);
+  UdpAddress(const HostPort& host_port);
+  const struct sockaddr* SockAddr() const;
+  int SockAddrLen() const;
+
+  friend std::ostream& operator<<(std::ostream& out, const UdpAddress& addr) {
+    char s[INET_ADDRSTRLEN];
+    inet_ntop(AF_INET, &(addr.addr_.sin_addr), s, INET_ADDRSTRLEN);
+    out << s << ":" << addr.addr_.sin_port;
+    return out;
+  }
+
+ private:
+  static struct sockaddr_in GetAddrInfo(const char* node, const char* service,
+                                        struct addrinfo* hints);
+
+  struct sockaddr_in addr_;
+};
+
+class UdpSocket {
+ public:
+  UdpSocket();
+  UdpSocket(const UdpAddress& addr);
+  void SendTo(const std::string& msg, const UdpAddress& addr);
+  void RecvFrom(std::string* msg, UdpAddress* addr);
+
+ private:
+  int socket_;
+};
+
+#endif  //  UDP_H_

### 2018-04-07 lucy/bank_account_client.h
Work on gossip server + client.

@@ -1,0 +1,17 @@
+#ifndef BANK_ACCOUNT_CLIENT_H_
+#define BANK_ACCOUNT_CLIENT_H_
+
+#include <cstdint>
+
+#include "client.h"
+
+class BankAccountClient : public Client {
+ public:
+  enum Result { COMMITTED, ABORTED };
+
+  std::int64_t Get(const UdpAddress& dst_addr);
+  Result Deposit(std::int64_t amount, const UdpAddress& dst_addr);
+  Result Withdraw(std::int64_t amount, const UdpAddress& dst_addr);
+};
+
+#endif  //  BANK_ACCOUNT_CLIENT_H_

### 2018-04-07 lucy/bank_account_client_main.cc

@@ -1,0 +1,56 @@
+#include <cstdlib>
+#include <iostream>
+#include <string>
+#include <vector>
+
+#include "glog/logging.h"
+
+#include "bank_account_client.h"
+#include "cluster.h"
+#include "string_util.h"
+
+namespace {
+
+void Usage() {
+  std::cerr << "./bank_account_client_main <cluster_file>" << std::endl;
+  ;
+}
+
+void ReplUsage() {
+  std::cout << "get | deposit <n> | withdraw <n>" << std::endl;
+}
+
+}  // namespace
+
+int main(int argc, char* argv[]) {
+  google::InitGoogleLogging(argv[0]);
+
+  if (argc != 2) {
+    Usage();
+    return EXIT_FAILURE;
+  }
+
+  const std::string cluster_filename = argv[1];
+  Cluster cluster(cluster_filename);
+  BankAccountClient client;
+
+  std::string line;
+  std::cout << "> " << std::flush;
+  std::size_t replica_ = 0;
+  while (std::getline(std::cin, line)) {
+    std::vector<std::string> words = Split(line);
+    const UdpAddress& dst_addr = cluster.UdpAddrs()[replica_];
+    if (words.size() == 1 && words[0] == "get") {
+      std::cout << client.Get(dst_addr) << std::endl;
+    } else if (words.size() == 2 && words[0] == "deposit") {
+      client.Deposit(std::stoi(words[1]), dst_addr);
+    } else if (words.size() == 2 && words[0] == "withdraw") {
+      client.Withdraw(std::stoi(words[1]), dst_addr);
+    } else {
+      ReplUsage();
+    }
+
+    replica_ = (replica_ + 1) % cluster.Size();
+    std::cout << "> " << std::flush;
+  }
+}

### 2018-04-07 lucy/client.cc

@@ -1,0 +1,24 @@
+#include "client.h"
+
+#include "glog/logging.h"
+
+#include "server.pb.h"
+
+std::string Client::Run(const std::string& txn, const UdpAddress& dst_addr) {
+  // Send request.
+  ServerMessage request;
+  request.set_type(ServerMessage::TXN_REQUEST);
+  request.mutable_txn_request()->set_txn(txn);
+  std::string request_str;
+  request.SerializeToString(&request_str);
+  socket_.SendTo(request_str, dst_addr);
+
+  // Get reply.
+  std::string reply_str;
+  socket_.RecvFrom(&reply_str, nullptr);
+  ServerMessage reply;
+  reply.ParseFromString(reply_str);
+  CHECK_EQ(reply.type(), ServerMessage::TXN_REPLY);
+  CHECK(reply.has_txn_reply());
+  return reply.txn_reply().reply();
+}

### 2018-04-07 lucy/gossip_server.cc

@@ -1,5 +1,60 @@
 #include "gossip_server.h"
 
+#include "glog/logging.h"
+
+#include "rand_util.h"
+
 void GossipServer::Run() {
-  // TODO: Implement.
+  while (true) {
+    std::string msg;
+    UdpAddress src_addr;
+    socket_.RecvFrom(&msg, &src_addr);
+
+    ServerMessage proto;
+    proto.ParseFromString(msg);
+    switch (proto.type()) {
+      case ServerMessage::MERGE_REQUEST: {
+        CHECK(proto.has_merge_request());
+        HandleMergeRequest(proto.merge_request(), src_addr);
+      }
+      case ServerMessage::TXN_REQUEST: {
+        CHECK(proto.has_txn_request());
+        HandleTxnRequest(proto.txn_request(), src_addr);
+      }
+      default: { LOG(FATAL) << "Unexpected server message type"; }
+    }
+
+    num_requests_serviced_++;
+    if (num_requests_serviced_ % num_requests_per_gossip_ == 0) {
+      // We could be sending this merge request to ourselves. While this is
+      // inefficient, it's not incorrect.
+      int dst_replica = RandomInt(0, cluster_.Size());
+      const UdpAddress& dst_addr = cluster_.UdpAddrs()[dst_replica];
+
+      ServerMessage msg;
+      msg.set_type(ServerMessage::MERGE_REQUEST);
+      msg.mutable_merge_request()->set_object(object_->Get());
+      std::string s;
+      msg.SerializeToString(&s);
+      socket_.SendTo(s, dst_addr);
+    }
+  }
+}
+
+void GossipServer::HandleMergeRequest(const MergeRequest& merge_request,
+                                      const UdpAddress& src_addr) {
+  DLOG(INFO) << "Received MergeRequest from " << src_addr << ".";
+  (void)src_addr;
+  object_->Merge(merge_request.object());
+}
+
+void GossipServer::HandleTxnRequest(const TxnRequest& txn_request,
+                                    const UdpAddress& src_addr) {
+  DLOG(INFO) << "Received TxnRequest from " << src_addr << ".";
+  ServerMessage msg;
+  msg.set_type(ServerMessage::TXN_REPLY);
+  msg.mutable_txn_reply()->set_reply(object_->Run(txn_request.txn()));
+  std::string s;
+  msg.SerializeToString(&s);
+  socket_.SendTo(s, src_addr);
 }

### 2018-04-07 lucy/rand_util.h

@@ -1,0 +1,6 @@
+#ifndef RAND_UTIL_H_
+#define RAND_UTIL_H_
+
+int RandomInt(int low_inclusive, int high_exclusive);
+
+#endif  // RAND_UTIL_H_

### 2018-04-07 lucy/string_util.h

@@ -1,0 +1,9 @@
+#ifndef STRING_UTIL_H_
+#define STRING_UTIL_H_
+
+#include <string>
+#include <vector>
+
+std::vector<std::string> Split(const std::string& s);
+
+#endif  //  STRING_UTIL_H_

### 2018-04-16 lucy/benchmark_master.cc
Updated benchmarks. Ready to roll.

@@ -7,7 +7,7 @@
 #include "host_port.h"
 #include "proto_util.h"
 
-static constexpr std::size_t NUM_CLIENTS_PER_SERVER = 2;
+static constexpr std::size_t NUM_CLIENTS_PER_SERVER = 3;
 
 void BenchmarkMaster::ServersStart(const BenchmarkServerStartRequest& start) {
   // Send requests.

### 2018-04-16 lucy/benchmark_master_main.cc

@@ -37,8 +37,9 @@ void VaryWithdraws(const std::size_t num_servers, BenchmarkMaster *master) {
   std::ofstream vary_withdraws_file("vary_withdraws.csv");
 
   for (ServerType server_type : {GOSSIP, SEGMENTED_MASTER, PAXOS}) {
-    std::vector<double> fraction_withdraws = {0,   0.01, 0.025, 0.05,
-                                              0.1, 0.2,  0.3,   0.4};
+    std::vector<double> fraction_withdraws = {0,   0.01, 0.02, 0.03, 0.04, 0.05,
+                                              0.1, 0.2,  0.3,  0.4,  0.5,  0.6,
+                                              0.7, 0.8,  0.9,  1.0};
     for (double fraction_withdraw : fraction_withdraws) {
       LOG(INFO) << "=====================================================";
       LOG(INFO) << "server_type       = " << ServerTypeToString(server_type);
@@ -85,7 +86,8 @@ void VarySegments(std::size_t num_servers, BenchmarkMaster *master) {
   std::ofstream vary_segments_file("vary_segments.csv");
 
   for (ServerType server_type : {GOSSIP, SEGMENTED_MASTER, PAXOS}) {
-    std::vector<std::uint64_t> segment_lengths = {2, 3, 4, 5, 6, 7, 8, 9, 10};
+    std::vector<std::uint64_t> segment_lengths = {2, 3, 4,  5,  6,  7,
+                                                  8, 9, 10, 20, 30, 40};
     for (double segment_length : segment_lengths) {
       LOG(INFO) << "=====================================================";
       LOG(INFO) << "server_type    = " << ServerTypeToString(server_type);
@@ -131,7 +133,8 @@ void VaryNodes(const std::size_t total_num_servers, BenchmarkMaster *master) {
   std::ofstream vary_nodes_file("vary_nodes.csv");
 
   for (ServerType server_type : {GOSSIP, SEGMENTED_MASTER, PAXOS}) {
-    std::vector<std::uint64_t> num_servers = {2, 3, 4, 8, 16, 24, 32};
+    std::vector<std::uint64_t> num_servers = {2, 3,  4,  5,  6,  7,
+                                              8, 16, 20, 24, 28, 32};
     for (double num_server : num_servers) {
       if (num_server > total_num_servers) {
         continue;
@@ -153,7 +156,7 @@ void VaryNodes(const std::size_t total_num_servers, BenchmarkMaster *master) {
[…]

### 2018-04-16 lucy/client.cc
Increased client retry.

@@ -14,7 +14,7 @@ Client::Client(ServerType server_type, const Cluster& server_cluster,
       server_type_(server_type),
       server_cluster_(server_cluster) {
   resend_pending_txn_timer_ = loop->RegisterTimer(
-      std::chrono::milliseconds(400), [this]() { ResendPendingTxn(); });
+      std::chrono::milliseconds(1000), [this]() { ResendPendingTxn(); });
 }
 
 UdpAddress Client::GetServerAddress() const {

### 2018-04-16 lucy/paxos_server.cc
Added another log to paxos server.

@@ -112,6 +112,9 @@ void PaxosServer::HandlePrepareOk(const PrepareOk& prepare_ok,
     return;
   }
 
+  VLOG(1) << "PaxosServer leader received a PrepareOk for transaction "
+          << txn_index << "from all other replicas!";
+
   // If we have received a PrepareOk from every other server, then we're ready
   // to try and commit this transaction. We add it to waiting_for_commit_ and
   // also clean up the metadata in prepare_ok_replies_ and

### 2018-04-16 lucy/paxos_server.cc
Fixed paxos bug.

@@ -73,6 +73,7 @@ void PaxosServer::HandleTxnRequest(const TxnRequest& txn_request,
   // the other servers and once we've committed all previous transactions, we
   // will commit this transaction.
   waiting_for_prepare_oks_.insert({txn_index, {txn_request, src_addr}});
+  prepare_ok_replies_[txn_index];
 
   // Send Prepare messages to all the other servers.
   ServerMessage msg;
@@ -113,7 +114,7 @@ void PaxosServer::HandlePrepareOk(const PrepareOk& prepare_ok,
   }
 
   VLOG(1) << "PaxosServer leader received a PrepareOk for transaction "
-          << txn_index << "from all other replicas!";
+          << txn_index << " from all other replicas!";
 
   // If we have received a PrepareOk from every other server, then we're ready
   // to try and commit this transaction. We add it to waiting_for_commit_ and

### 2018-04-17 benchmark/clone_install_build_launch.sh
Updated benchmark scripts.

@@ -15,6 +15,7 @@ main() {
     fi
 
     cd enforced_invariant_confluence/lucy
+    git pull
     make DEBUG=0
 
     if tmux has-session -t iconfluence &> /dev/null; then
@@ -27,22 +28,33 @@ main() {
     tmux new-window -n client_2
     tmux new-window -n client_3
 
-    public_ip="$(hostname -I)"
-    server_index="$(grep -n $public_ip ~/benchmark_server_cluster.txt | cut -d : -f 1)"
-    server_index="$((server_index - 1))"
-    client_1_index="$((server_index * 3))"
-    client_2_index="$((client_1_index + 1))"
-    client_3_index="$((client_1_index + 2))"
+    private_ip="$(hostname -I)"
+    echo "private_ip = $private_ip"
+    # Launch the servers, but only if we're not the master.
+    if grep $private_ip $HOME/benchmark_client_cluster.txt &> /dev/null; then
+        echo "STARTING SERVERS"
+        server_index="$(grep -n $private_ip ~/benchmark_server_cluster.txt | cut -d : -f 1)"
+        server_index="$((server_index - 1))"
+        client_1_index="$((server_index * 3))"
+        client_2_index="$((client_1_index + 1))"
+        client_3_index="$((client_1_index + 2))"
 
-    echo "server_index = $server_index"
-    echo "client_1_index = $client_1_index"
-    echo "client_2_index = $client_2_index"
-    echo "client_3_index = $client_3_index"
+        echo "server_index = $server_index"
+        echo "client_1_index = $client_1_index"
+        echo "client_2_index = $client_2_index"
+        echo "client_3_index = $client_3_index"
 
-    tmux send-keys -t iconfluence:server.0 "GLOG_logtostderr=1 ./benchmark_server_main ~/benchmark_server_cluster.txt ~/server_cluster.txt $server_index" C-m
-    tmux send-keys -t iconfluence:client_1.0 "GLOG_logtostderr=1 ./benchmark_client_main ~/benchmark_client_cluster.txt ~/server_cluster.txt $client_1_index" C-m
[…]

### 2018-04-17 benchmark/create_cluster_files.py

@@ -41,9 +41,9 @@ def main():
 
     with open("benchmark_client_cluster.txt", "w") as f:
         for private_ip in private_ips:
-            f.write('host_port {{ host: "{}" port: 8000 }}\n'.format(private_ip))
-            f.write('host_port {{ host: "{}" port: 8001 }}\n'.format(private_ip))
-            f.write('host_port {{ host: "{}" port: 8002 }}\n'.format(private_ip))
+            f.write('host_port {{ host: "{}" port: 9000 }}\n'.format(private_ip))
+            f.write('host_port {{ host: "{}" port: 9001 }}\n'.format(private_ip))
+            f.write('host_port {{ host: "{}" port: 9002 }}\n'.format(private_ip))
 
 if __name__ == '__main__':
     main()

### 2018-04-17 benchmark/launch_ec2_instances.sh

@@ -33,7 +33,7 @@ main() {
         --instance-type "$instance_type" \
         --key-name "$key_name" \
         --security-group-ids "$security_group_ids" \
-        --count 9
+        --count 32
         # --dry-run
 }
 

### 2018-04-17 benchmark/plot.py
Added plotting script.

@@ -1,0 +1,86 @@
+from collections import defaultdict
+import csv
+import os
+import sys
+
+import matplotlib.pyplot as plt
+
+VARY_WITHDRAWS = 'vary_withdraws.csv'
+VARY_SEGMENTS = 'vary_segments.csv'
+VARY_NODES = 'vary_nodes.csv'
+
+def load_csv(filename):
+    with open(filename, 'r') as f:
+        reader = csv.reader(f, delimiter=',')
+        return [(row[0], float(row[1]), float(row[2])) for row in reader]
+
+def split_data(data):
+    splits = defaultdict(lambda: ([], []))
+    for (server_type, x, y) in data:
+        splits[server_type][0].append(x)
+        splits[server_type][1].append(y)
+    return splits
+
+def plot_vary_withdraws(data):
+    splits = split_data(data)
+
+    gossip_x, gossip_y = splits['gossip']
+    segmented_x, segmented_y = splits['segmented_master']
+    paxos_x, paxos_y = splits['paxos']
+
+    plt.figure()
+    plt.plot(gossip_x, gossip_y, label='eventual', marker='o')
+    plt.plot(segmented_x, segmented_y, label='invariant confluent', marker='s')
+    plt.plot(paxos_x, paxos_y, label='linearizable', marker='^')
+    plt.legend()
+    plt.xlabel('Fraction Withdrawals')
+    plt.ylabel('Throughput (txns/s)')
+    plt.savefig('vary_withdraws.pdf')
+    plt.close()
+
+def plot_vary_segments(data):
+    splits = split_data(data)
+
+    gossip_x, gossip_y = splits['gossip']
+    segmented_x, segmented_y = splits['segmented_master']
+    paxos_x, paxos_y = splits['paxos']
+
+    plt.figure()
+    plt.plot(gossip_x, gossip_y, label='eventual', marker='o')
+    plt.plot(segmented_x, segmented_y, label='invariant confluent', marker='s')
+    plt.plot(paxos_x, paxos_y, label='linearizable', marker='^')
+    plt.legend()
+    plt.xlabel('Segment Side Length')
+    plt.ylabel('Throughput (txns/s)')
+    plt.savefig('vary_segments.pdf')
+    plt.close()
+
+def plot_vary_nodes(data):
+    splits = split_data(data)
+
+    gossip_x, gossip_y = splits['gossip']
[…]

### 2018-04-17 lucy/paxos_server.cc
Tweaking paxos hyperparams.

@@ -12,7 +12,7 @@ PaxosServer::PaxosServer(const Cluster& cluster, replica_index_t replica_index,
   if (AmLeader()) {
     LOG(INFO) << "PaxosServer leader listening on "
               << cluster_.UdpAddrs()[replica_index_] << ".";
-    const std::chrono::milliseconds delay(100);
+    const std::chrono::milliseconds delay(500);
     resend_prepares_timer_ =
         loop->RegisterTimer(delay, [this]() { LeaderResendPrepares(); });
     resend_prepares_timer_.Start();

### 2018-04-17 lucy/paxos_server.cc
Another paxos tweak.

@@ -12,7 +12,7 @@ PaxosServer::PaxosServer(const Cluster& cluster, replica_index_t replica_index,
   if (AmLeader()) {
     LOG(INFO) << "PaxosServer leader listening on "
               << cluster_.UdpAddrs()[replica_index_] << ".";
-    const std::chrono::milliseconds delay(500);
+    const std::chrono::milliseconds delay(75);
     resend_prepares_timer_ =
         loop->RegisterTimer(delay, [this]() { LeaderResendPrepares(); });
     resend_prepares_timer_.Start();

### 2018-04-17 lucy/paxos_server.cc
More paxos tweaks.

@@ -12,7 +12,7 @@ PaxosServer::PaxosServer(const Cluster& cluster, replica_index_t replica_index,
   if (AmLeader()) {
     LOG(INFO) << "PaxosServer leader listening on "
               << cluster_.UdpAddrs()[replica_index_] << ".";
-    const std::chrono::milliseconds delay(75);
+    const std::chrono::milliseconds delay(20);
     resend_prepares_timer_ =
         loop->RegisterTimer(delay, [this]() { LeaderResendPrepares(); });
     resend_prepares_timer_.Start();

### 2018-04-17 lucy/paxos_server.cc
Even more paxos tweaks.

@@ -12,7 +12,7 @@ PaxosServer::PaxosServer(const Cluster& cluster, replica_index_t replica_index,
   if (AmLeader()) {
     LOG(INFO) << "PaxosServer leader listening on "
               << cluster_.UdpAddrs()[replica_index_] << ".";
-    const std::chrono::milliseconds delay(20);
+    const std::chrono::milliseconds delay(100);
     resend_prepares_timer_ =
         loop->RegisterTimer(delay, [this]() { LeaderResendPrepares(); });
     resend_prepares_timer_.Start();
@@ -64,6 +64,12 @@ void PaxosServer::OnRecvFollower(const std::string& msg,
 void PaxosServer::HandleTxnRequest(const TxnRequest& txn_request,
                                    const UdpAddress& src_addr) {
   VLOG(1) << "PaxosServer received TxnRequest from " << src_addr << ".";
+
+  // Drop this request if we're already waiting for too many requests.
+  if (waiting_for_prepare_oks_.size() > 100) {
+    VLOG(1) << "PaxosServer request buffer is full and is dropping a request.";
+    return;
+  }
 
   // Assign this transaction a new transaction index.
   const txn_index_t txn_index = txn_index_;

### 2018-04-17 lucy/bank_account.cc
Start bank account off with higher balance.

@@ -8,7 +8,7 @@
 BankAccount::BankAccount(std::size_t num_replicas, replica_index_t replica)
     : num_replicas_(num_replicas),
       replica_(replica),
-      p_(/*count=*/num_replicas, /*value=*/0),
+      p_(/*count=*/num_replicas, /*value=*/50 * 1000 * 1000),
       n_(/*count=*/num_replicas, /*value=*/0) {
   CHECK_GT(num_replicas, 0);
   CHECK_GE(replica, 0);

### 2018-04-17 lucy/paxos_server.cc
Trying a variant of paxos.

@@ -122,6 +122,15 @@ void PaxosServer::HandlePrepareOk(const PrepareOk& prepare_ok,
   VLOG(1) << "PaxosServer leader received a PrepareOk for transaction "
           << txn_index << " from all other replicas!";
 
+  // Reply to the client. We should be waiting to commit the transaction first,
+  // but this hack simulates a much more optimized variant of Paxos.
+  ServerMessage msg;
+  msg.mutable_txn_reply()->set_request_id(
+      waiting_for_prepare_oks_[txn_index].txn_request.request_id());
+  msg.mutable_txn_reply()->set_reply(
+      object_->ExecTxn(waiting_for_prepare_oks_[txn_index].txn_request.txn()));
+  SendTo(msg, waiting_for_prepare_oks_[txn_index].src_addr);
+
   // If we have received a PrepareOk from every other server, then we're ready
   // to try and commit this transaction. We add it to waiting_for_commit_ and
   // also clean up the metadata in prepare_ok_replies_ and
@@ -158,14 +167,15 @@ void PaxosServer::LeaderCommitReadyTransactions() {
 
   auto it = waiting_for_commit_.begin();
   while (it->first == num_committed_) {
-    // Execute the transaction, and send a reply to the client.
-    const PendingTxn& pending_txn = it->second;
-    ServerMessage msg;
-    msg.mutable_txn_reply()->set_request_id(
-        pending_txn.txn_request.request_id());
-    msg.mutable_txn_reply()->set_reply(
-        object_->ExecTxn(pending_txn.txn_request.txn()));
-    SendTo(msg, pending_txn.src_addr);
+    // TODO: Clean up.
+    // // Execute the transaction, and send a reply to the client.
+    // const PendingTxn& pending_txn = it->second;
+    // ServerMessage msg;
+    // msg.mutable_txn_reply()->set_request_id(
+    //     pending_txn.txn_request.request_id());
+    // msg.mutable_txn_reply()->set_reply(
+    //     object_->ExecTxn(pending_txn.txn_request.txn()));
+    // SendTo(msg, pending_txn.src_addr);
 
     num_committed_++;
     it = waiting_for_commit_.erase(it);

### 2018-04-17 benchmark/plot.py
Updated plot.py.

@@ -5,10 +5,6 @@ import sys
 
 import matplotlib.pyplot as plt
 
-VARY_WITHDRAWS = 'vary_withdraws.csv'
-VARY_SEGMENTS = 'vary_segments.csv'
-VARY_NODES = 'vary_nodes.csv'
-
 def load_csv(filename):
     with open(filename, 'r') as f:
         reader = csv.reader(f, delimiter=',')
@@ -29,9 +25,9 @@ def plot_vary_withdraws(data):
     paxos_x, paxos_y = splits['paxos']
 
     plt.figure()
-    plt.plot(gossip_x, gossip_y, label='eventual', marker='o')
-    plt.plot(segmented_x, segmented_y, label='invariant confluent', marker='s')
-    plt.plot(paxos_x, paxos_y, label='linearizable', marker='^')
+    plt.semilogy(gossip_x, gossip_y, label='eventual', marker='o')
+    plt.semilogy(segmented_x, segmented_y, label='invariant confluent', marker='s')
+    plt.semilogy(paxos_x, paxos_y, label='linearizable', marker='^')
     plt.legend()
     plt.xlabel('Fraction Withdrawals')
     plt.ylabel('Throughput (txns/s)')
@@ -46,9 +42,9 @@ def plot_vary_segments(data):
     paxos_x, paxos_y = splits['paxos']
 
     plt.figure()
-    plt.plot(gossip_x, gossip_y, label='eventual', marker='o')
-    plt.plot(segmented_x, segmented_y, label='invariant confluent', marker='s')
-    plt.plot(paxos_x, paxos_y, label='linearizable', marker='^')
+    plt.semilogy(gossip_x, gossip_y, label='eventual', marker='o')
+    plt.semilogy(segmented_x, segmented_y, label='invariant confluent', marker='s')
+    plt.semilogy(paxos_x, paxos_y, label='linearizable', marker='^')
     plt.legend()
     plt.xlabel('Segment Side Length')
     plt.ylabel('Throughput (txns/s)')
@@ -63,9 +59,9 @@ def plot_vary_nodes(data):
     paxos_x, paxos_y = splits['paxos']
 
     plt.figure()
-    plt.plot(gossip_x, gossip_y, label='eventual', marker='o')
-    plt.plot(segmented_x, segmented_y, label='invariant confluent', marker='s')
-    plt.plot(paxos_x, paxos_y, label='linearizable', marker='^')
[…]

### 2018-04-17 benchmark/plot.py
Added figures.

@@ -3,7 +3,24 @@ import csv
 import os
 import sys
 
+import matplotlib
 import matplotlib.pyplot as plt
+
+EVENTUAL_FMT = {
+    'label': 'eventual',
+    'marker': 'o',
+    'linewidth': 2,
+}
+IC_FMT = {
+    'label': 'invariant confluent',
+    'marker': 'o',
+    'linewidth': 2,
+}
+LIN_FMT = {
+    'label': 'linearizable',
+    'marker': '^',
+    'linewidth': 2,
+}
 
 def load_csv(filename):
     with open(filename, 'r') as f:
@@ -25,10 +42,10 @@ def plot_vary_withdraws(data):
     paxos_x, paxos_y = splits['paxos']
 
     plt.figure()
-    plt.semilogy(gossip_x, gossip_y, label='eventual', marker='o')
-    plt.semilogy(segmented_x, segmented_y, label='invariant confluent', marker='s')
-    plt.semilogy(paxos_x, paxos_y, label='linearizable', marker='^')
-    plt.legend()
+    plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
+    plt.semilogy(segmented_x, segmented_y, **IC_FMT)
+    plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
+    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.12), prop={'size': 12})
     plt.xlabel('Fraction Withdrawals')
     plt.ylabel('Throughput (txns/s)')
     plt.savefig('vary_withdraws.pdf')
@@ -42,12 +59,12 @@ def plot_vary_segments(data):
     paxos_x, paxos_y = splits['paxos']
 
     plt.figure()
-    plt.semilogy(gossip_x, gossip_y, label='eventual', marker='o')
-    plt.semilogy(segmented_x, segmented_y, label='invariant confluent', marker='s')
-    plt.semilogy(paxos_x, paxos_y, label='linearizable', marker='^')
-    plt.legend()
+    plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
+    plt.semilogy(segmented_x, segmented_y, **IC_FMT)
+    plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
     plt.xlabel('Segment Side Length')
     plt.ylabel('Throughput (txns/s)')
+    plt.tight_layout()
     plt.savefig('vary_segments.pdf')
[…]

### 2018-04-17 benchmark/plot.py
Plot tweaks.

@@ -41,14 +41,14 @@ def plot_vary_withdraws(data):
     segmented_x, segmented_y = splits['segmented_master']
     paxos_x, paxos_y = splits['paxos']
 
-    plt.figure()
+    plt.figure(figsize=(8, 3))
     plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
     plt.semilogy(segmented_x, segmented_y, **IC_FMT)
     plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
-    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.12), prop={'size': 12})
+    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.2), prop={'size': 12})
     plt.xlabel('Fraction Withdrawals')
     plt.ylabel('Throughput (txns/s)')
-    plt.savefig('vary_withdraws.pdf')
+    plt.savefig('vary_withdraws.pdf', bbox_inches='tight')
     plt.close()
 
 def plot_vary_segments(data):
@@ -58,14 +58,13 @@ def plot_vary_segments(data):
     segmented_x, segmented_y = splits['segmented_master']
     paxos_x, paxos_y = splits['paxos']
 
-    plt.figure()
+    plt.figure(figsize=(8, 3))
     plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
     plt.semilogy(segmented_x, segmented_y, **IC_FMT)
     plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
     plt.xlabel('Segment Side Length')
     plt.ylabel('Throughput (txns/s)')
-    plt.tight_layout()
-    plt.savefig('vary_segments.pdf')
+    plt.savefig('vary_segments.pdf', bbox_inches='tight')
     plt.close()
 
 def plot_vary_nodes(data):
@@ -75,14 +74,14 @@ def plot_vary_nodes(data):
     segmented_x, segmented_y = splits['segmented_master']
     paxos_x, paxos_y = splits['paxos']
 
-    plt.figure()
+    plt.figure(figsize=(8, 4))
     plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
     plt.semilogy(segmented_x, segmented_y, **IC_FMT)
     plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
[…]

### 2018-04-17 benchmark/plot.py
Updated figure labels.

@@ -12,7 +12,7 @@ EVENTUAL_FMT = {
     'linewidth': 2,
 }
 IC_FMT = {
-    'label': 'invariant confluent',
+    'label': 'segmented',
     'marker': 'o',
     'linewidth': 2,
 }

### 2018-04-17 doc/paper/sections/abstract.tex
Minor edits + starting on eval.

@@
 Strongly consistent systems are easy to reason about but face fundamental
-limitations in availability and performance. Weakly consistent systems can be
-implemented with very high performance, but place a burden on the application
-developer to reason about complex interleavings of execution.
+limitations in availability and performance, and weakly consistent systems can
+be implemented with very high performance but place a burden on the
+application developer to reason about complex interleavings of execution.
 Invariant-confluence provides a formal framework for understanding when we can
@@
 commonly used sufficient condition for invariant confluence is both necessary
-and sufficient. We use this condition to design a general-purpose interactive
-invariant-confluence decision procedure and a novel sufficient condition which
-can be checked automatically. We then take a step beyond invariant-confluence
-and introduce a generalization of invariant-confluence, called segmented
-invariant-confluence, which allows us to replicate non-invariant-confluent
-objects with a small and tunable amount of coordination. Implementing our
-theoretical findings in a system called Lucy, we evaluate our methods on a
-number of case studies and benchmarks showing that... TODO: Perform evaluation
-:P.
-% Strongly consistent systems are easy to reason about but are slow.
-Weakly consistent systems are fast but hard to reason about.
-Invariant-confluence gives a formal framework to reason about when data can be replicated without any coordination needed to maintain invraints, the best of borth worlds.
-However, determining whether an object is invariant-confluent is hard.
-In this paper, we establish conditions under which a commonly used sufficient condition for invariant confluence is both necessary and usfficient.
-We use this condition to design a general-purpose interactive decision procedure and a novel sufficient condition which can be checked automatically.
-We then take a step beyond invariant-confluence and introduce a generalization of invariant-confuluence, called segmented invariant-confluence which allows to replicate objects with a minimal amount of coordination that are otherwise not invariant-confluent.
-We implement all this in a system called lucy and look at a bunch of stuff.
+and sufficient and use this condition to design (a) a general-purpose
+interactive invariant-confluence decision procedure and (b) a novel sufficient
+condition which can be checked automatically. We then take a step beyond
+invariant-confluence and introduce a generalization of invariant-confluence
+called segmented invariant-confluence, which allows us to replicate
+non-invariant-confluent objects with a small and tunable amount of
+coordination. Implementing our theoretical findings in a system called Lucy, we
+evaluate our methods on a number of case studies and benchmarks showing that...
+TODO: Perform evaluation :P.
+% Strongly consistent systems are easy to reason about but are slow. Weakly
+% consistent systems are fast but hard to reason about. Invariant-confluence
+% gives a formal framework to reason about when data can be replicated without
+% any coordination needed to maintain invraints, the best of borth worlds.
+% However, determining whether an object is invariant-confluent is hard. In
+% this paper, we establish conditions under which a commonly used sufficient
+% condition for invariant confluence is both necessary and usfficient. We use
+% this condition to design a general-purpose interactive decision procedure and
+% a novel sufficient condition which can be checked automatically. We then take
+% a step beyond invariant-confluence and introduce a generalization of
+% invariant-confuluence, called segmented invariant-confluence which allows to
+% replicate objects with a minimal amount of coordination that are otherwise
+% not invariant-confluent. We implement all this in a system called lucy and
+% look at a bunch of stuff.
 \end{abstract}

### 2018-04-17 doc/paper/sections/evaluation.tex
Minor edits + starting on eval.

@@
 \section{Evaluation}\seclabel{Evaluation}
+In this section, we describe Lucy: an implementation of our decision procedures
+and system models. We then evaluate Lucy by answering the following questions:
+\begin{itemize}
+  \item
+    How practical is the interactive invariant-confluence decision procedure?
+    Can we use it to classify real-world transactions and invariants?
+  \item
+    How practical is segmented invariant-confluence? Are real-world workloads
+    amenable to segmentation?
+  \item
+    How efficient is the interactive invariant-confluence decision procedure?
+  \item
+    How efficiently can we replicate a segmented invariant-confluent object as
+    compared to alternative approaches like replicating with weak or strong
+    consistency?
+  \item
+    How does the performance of replicating a segmented invariant-confluence
+    object vary as we vary the workload, segmentation, and replication factor?
+\end{itemize}
+
 \subsection{Lucy}
-\todo{Describe implementation of lucy.}
+Lucy includes an implementation of the interactive decision procedure described
+in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
+procedure which checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
+implementation of the decision procedure described in
+\algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
+are implemented in roughly 2,500 lines of Python. Users specify objects,
+transactions, and invariants in a small Python DSL and interact with the
+interactive decision procedures using an interactive Python console. We use
+Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
+compiling objects and invariants into a formula which is satisfiable if and
+only if the object is \emph{not} invariant-confluent. If the object is
+invariant-closed, then Z3 concludes that the formula is unsatisfiable.
+Otherwise, if the object is not invariant-closed, then Z3 produces a
+counterexample witnessing the satisfiability of the formula.
 
-\subsection{Case Studies}
-\todo{Bank account with and without decrements.}
-\todo{Auction example.}
-\todo{Foreign keys with and without deletions.}
+Lucy also includes an implementation of the invariant-confluence and
+segmented-invariant confluence system models in roughly 3,500 lines of C++.
+Users specify objects, transactions, invariants, and segmentations in C++. Lucy
+then replicates the objects using invariant-confluence, segmented
+invariant-confluence, eventual consistency, or linearizability. In
+\secref{SegmentedInvariantConfluenceEval}, we compare the performance of these
+various consistency models.
+%
+Because fault-tolerance is largely an orthogonal concern to
+invariant-confluence, Lucy is implemented without fault-tolerance. It would be
+straightforward to add fault-tolerance to Lucy, but it would not affect our
+discussions or evaluation, so we leave it for future work. Moreover, users
+currently have to specify their workloads in Python (for the decision
+procedures) and C++ (for the runtime). In the future, we plan on removing this
+redundancy.
 
-\subsection{TPC and Feral}
-\todo{See which of the Feral and TPC transactions we can determine automatically}
+\subsection{Decision Procedures}
+% \todo{Bank account with and without decrements.}
+% \todo{Auction example.}
+% \todo{Foreign keys with and without deletions.}
+% \subsection{TPC and Feral}
+% \todo{See which of the Feral and TPC transactions we can determine automatically}
 
-\subsection{Segmented Invariant Confluence}
-\todo{one graph: number of locked operations x axis; throughput on the y-axis}
-\todo{one graph: number of segments on the x-axis; throughput on the y-axis}
-\todo{compare with no coordination and "VR"}
+\subsection{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluenceEval}
+% \todo{one graph: number of locked operations x axis; throughput on the y-axis}
+% \todo{one graph: number of segments on the x-axis; throughput on the y-axis}
+% \todo{compare with no coordination and "VR"}
+
+\begin{figure}[ht]
+  \centering
+
+  \begin{subfigure}[c]{\columnwidth}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/vary_withdraws.pdf}
+  \end{subfigure}
+  \begin{subfigure}[c]{\columnwidth}
+    \includegraphics[width=\columnwidth]{figures/vary_segments.pdf}
+  \end{subfigure}
+
+  \caption{}\figlabel{}
+\end{figure}
+
+\begin{figure}[ht]
+  \centering
+  \includegraphics[width=\columnwidth]{figures/vary_nodes.pdf}
+  \caption{}\figlabel{}
+\end{figure}
+
+\uselengthunit{in}\printlength{\columnwidth}

### 2018-04-17 doc/paper/sections/introduction.tex
Minor edits + starting on eval.

@@
 programs: application invariants~\cite{hoare1969axiomatic}. Even if every
-transaction executing in a weakly consistent system maintains an application
-invariant, the system as a whole can produce invariant-violating states.
+transaction executing in a weakly consistent system individually maintains an
+application invariant, the system as a whole can produce invariant-violating
+states.
 
@@
 of strong consistency. However, the task of identifying whether or not an
-object actually is invariant-confluent remain a challenge. Bailis et al.\
+object actually is invariant-confluent remains a challenge. Bailis et al.\
 manually categorized a set of common objects, transactions, and invariants

### 2018-04-17 doc/paper/sections/evaluation.tex
Progress on eval.

@@
 
-\subsection{Lucy}
+\subsection{Implementation}
 Lucy includes an implementation of the interactive decision procedure described
@@
 compiling objects and invariants into a formula which is satisfiable if and
-only if the object is \emph{not} invariant-confluent. If the object is
+only if the object is \emph{not} invariant-closed. If the object is
 invariant-closed, then Z3 concludes that the formula is unsatisfiable.
@@
 Users specify objects, transactions, invariants, and segmentations in C++. Lucy
-then replicates the objects using invariant-confluence, segmented
-invariant-confluence, eventual consistency, or linearizability. In
-\secref{SegmentedInvariantConfluenceEval}, we compare the performance of these
-various consistency models.
-%
+then replicates the objects using segmented invariant-confluence (or
+invariant-confluence if the segmentation contains a single segment without any
+disallowed transactions). Clients send every transaction request to a randomly
+selected server. When a server receives a transaction request, it executes
+\algoref{TxnExecution} to attempt to execute the transaction locally. If the
+transaction requires global coordination, then the server forwards the
+transaction request to a predetermined leader. When the leader receives a
+transaction request, it sends a coordination request to all other servers. When
+a server receives a coordination request from the leader, it stops processing
+transactions and sends the leader its state in a coordination reply. When the
+leader receives coordination replies from all other servers, it executes the
+transaction, and then sends its state to the other servers. When a server
+receives a new state, it adopts the state, computes its new active segment, and
+resumes normal processing. After every 100 transactions processed, a server
+sends a merge request to a randomly selected server. Merge requests are tagged
+with a monotonically increasing epoch number that is incremented by the master
+after every round of global coordination. This allows servers to discard merge
+requests from previous epochs.
+
+Lucy can also replicate an object with eventual consistency and with
+linearizability. With eventual consistency, clients send every transaction
+request to a randomly selected server. The server executes the transaction
+locally and returns immediately to the client, sending merge requests after
+every 100 transactions. With linearizability, clients send every transaction
+request to a predetermined leader. The leader relays the transaction request to
+all other servers, and when it receives replies from them, it executes the
+transaction and replies to the client. This communication pattern mimics the
+``normal operation'' of state machine replication protocols like
+Paxos~\cite{lamport1998part} and Viewstamped
+Replicated~\cite{liskov2012viewstamped}. In
+\secref{SegmentedInvariantConfluenceEval}, we compare the performance of
+replicating with segmented invariant-confluence against the performance of
+replicating with eventual consistency and linearizability.
+
 Because fault-tolerance is largely an orthogonal concern to
@@
 
-\subsection{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluenceEval}
-% \todo{one graph: number of locked operations x axis; throughput on the y-axis}
-% \todo{one graph: number of segments on the x-axis; throughput on the y-axis}
-% \todo{compare with no coordination and "VR"}
+\subsection{Segmented Invariant Confluence}%
+\seclabel{SegmentedInvariantConfluenceEval}
+In this section, we evaluate the performance of replicating an object with
+segmented invariant-confluence as compared to replicating it with eventual
+consistency or linearizability. We begin with two benchmarks that demonstrate
+the same concept: the performance of segmented invariant-confluent replication
+varies with the amount of global coordination induced by either (a) performing
+a transaction that is disallowed within a segment or (b) transitioning between
+segments.
 
@@
 
-  \caption{}\figlabel{}
+  \caption{%
+    Segmented invariant-confluent replication throughput versus coordination,
+    induced by executing disallowed transactions (top) and by transitioning
+    across segments (bottom).
+  }\figlabel{ThroughputVsGlobalSyncs}
 \end{figure}
+
+\begin{benchmark}\benchlabel{VaryWithdraws}
+Consider again the PN-Counter from \exampleref{CounreachableExample} and the
+corresponding transactions, invariants, and single-segment segmentation that
+forbids concurrent decrements. We replicate this object on 32 servers
+deployed on 32 m5.xlarge EC2 instances within the same availability zone.  Each
+server has three colocated clients that issue deposit and withdrawal
+transactions. We replicate the object with eventual consistency, segmented
+invariant-confluence, and linearizability and measure the system's total
+throughput as we vary the fraction of client requests that are withdrawals. The
+results are shown in the top of \figref{ThroughputVsGlobalSyncs}.
+
+Both eventually consistent replication and linearizable replication are
+unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
+per second respectively. Expectedly, eventually consistent replication
+significantly outperforms linearizable replication because (a) transactions can
+be sent to any server (not just the leader) and (b) servers do not coordinate
+with each other at all.
+%
+Segmented invariant-confluent replication performs well for low-withdrawal
+workloads and performs increasingly poorly as we increase the fraction of
+withdrawal transactions, eventually becoming slower that linearizable
+replication. For example, with 5\% withdrawal transactions, segmented
+invariant-confluent replication performs an order of magnitude better than
+linearizable replication; with 50\% withdrawals, it performs as well; and with
+100\% withdrawals, it performs two times worse.
+
+These results are expected. Deposit transactions can execute without any
+coordination while withdrawal transactions require global coordination. As we
+increase the fraction of withdrawals, we increase the amount of coordination
+that the system has to perform which in turn drastically decreases the
+throughput. These results also offer two insights:
+%
+First, for low-withdrawal workloads, segmented invariant-confluent replication
+achieves a compromise between strong and weak consistency. It guarantees that
+invariants are maintained (which is impossible with eventual consistency if the
+object is not invariant-confluent) with performance many times better than
+strongly consistent replication.
+%
+Second, segmented invariant-confluent replication is poorly suited to workloads
+that require a large amount of coordination. For workloads without much inherit
+concurrency (e.g.\ withdraw-mostly workloads), maintaining invariants is best
+done with strong consistency. It provides stronger guarantees with better
+performance.
+\end{benchmark}
+
+\begin{benchmark}\benchlabel{VarySegmentLength}
+  Consider again the object, transactions, and invariants from \exampleref{Z2}
+  and \exampleref{SegmentedZ2}. As with \benchref{VaryWithdraws}, we replicate
+  the object across 32 servers. Clients issue 50\% increment $x$ transactions,
+  and 50\% decrement $y$ transactions. We consider a ``checkerboard''
+  segmentation $\Sigma_n = \setst{(I_{i, j}, T)}{i, j \in \ints}$ where segment
+  invariant $I_{i, j}$ consists of the square of points $\setst{(x, y)}{ni \leq
+  x < n(i + 1), nj \leq y < n(j + 1)}$ with side length $n$. For example,
+  $\Sigma_1$ places each point in its own segment, $\Sigma_2$ tessellates
+  $\ints^2$ with 2x2 squares, $\Sigma_3$ tessellates $\ints^2$ with 3x3
+  squares, and so on. We measure the throughput of the object replicated with
+  eventually consistent, segmented invariant-confluent, and linearizable
+  replication as we vary the segment side length $n$. The results are shown in
+  the bottom \figref{ThroughputVsGlobalSyncs}.
+
+  This benchmark tells the same tale as \benchref{VaryWithdraws}. Eventual
+  consistency and linearizability are unaffected by workload, and eventual
+  consistency outperforms linearizability by roughly two orders of magnitude.
+  In this example, the segmented invariant-confluent replication only requires
+  coordination when transitioning between segment boundaries, so as we increase
+  the segment side length, the throughput of the system increases
+  significantly.
+\end{benchmark}
 
@@
 \end{figure}
-
-\uselengthunit{in}\printlength{\columnwidth}

### 2018-04-18 doc/paper/sections/evaluation.tex
Finished benchmark part of eval.

@@
   \includegraphics[width=\columnwidth]{figures/vary_nodes.pdf}
-  \caption{}\figlabel{}
+  \caption{%
+    Throughput of eventually consistent, segmented invariant-confluent, and
+    linearizable replication measured against the number of
+    nodes.
+  }\figlabel{VaryNodes}
 \end{figure}
+
+\begin{benchmark}
+  In this benchmark, we measure the scale-out of segmented invariant-confluent
+  replication. We repeat \benchref{VaryWithdraws} with a 10\% withdrawal rate,
+  but this time we vary the number of servers we use to replicate our object.
+  When we replicate with $n$ servers, we use $3n$ clients (the $3$ colocated
+  clients on each server) as part of the workload. The results are shown in
+  \figref{VaryNodes}.
+
+  Eventually consistent replication scales perfectly with the number of nodes,
+  confirming the results in~\cite{bailis2014coordination}. With eventually
+  consistent replication, servers do not coordinate at all, so they are
+  completely unaffected by the number of servers. Linearizable replication, on
+  the other hand, scales up to about 3-5 servers before performance begins to
+  decrease. These numbers are consistent with typical deployments of
+  state-machine replication protocols like Paxos~\cite{chandra2007paxos}.
+  Because all messages are sent to the leader, the leader becomes the
+  bottleneck as the number of servers and clients increases. Moreover, the
+  leader must wait for responses from more servers, increasing the latency of
+  the slowest response which in turn decreases throughput.
+  Segmented invariant-confluent replication scales up to about 6-8 servers
+  before succumbing to the same scalability bottlenecks as linearizable
+  replication.
+
+  These results highlight the importance of coordination avoidance in
+  distributed databases. While segmented invariant-confluent replication scales
+  out slightly better than linearizable replication, both scale significantly
+  worse than eventually consistent replication even for a very low (i.e.\ 10\%)
+  withdrawal workload. This demonstrates that even a small amount of
+  coordination can significantly reduce the scalability of a system.
+\end{benchmark}

### 2018-04-18 doc/paper/sections/introduction.tex
Work on related work.

@@
 replication~\cite{schneider1990implementing, lamport1998part,
-liskov2012viewstamped, ongaro2014search} makes the application designer's life
-very easy. To the application developer, a strongly consistent system behaves
-exactly like a single-threaded system running on a single node, so reasoning
-about the behavior of the system is simple. Unfortunately, strong consistency
-is at odds with performance. The CAP theorem and PACELC theorem tell us that
+liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
+oki1988viewstamped, lamport2005generalized, lamport2006fast} makes the
+application designer's life very easy. To the application developer, a strongly
+consistent system behaves exactly like a single-threaded system running on a
+single node, so reasoning about the behavior of the system is
+simple~\cite{herlihy1990linearizability}.  Unfortunately, strong consistency is
+at odds with performance. The CAP theorem and PACELC theorem tell us that
 strongly consistent systems suffer from higher latency at best and
-unavailability at worst~\cite{brewer2012cap, abadi2012consistency}. On the
-other hand, weak consistency models like eventual~\cite{vogels2009eventually},
-PRAM~\cite{lipton1988pram}, and causal~\cite{ahamad1995causal} consistency
-allow data to be replicated with high availability and low latency, but they
-put a tremendous burden on the application designer to reason about the complex
-interleavings of operations that are allowed by these weak consistency models.
-In fact, weak consistency models strip an application developer of one of the
-earliest and most effective tools that is used to reason about the execution of
-programs: application invariants~\cite{hoare1969axiomatic}. Even if every
+unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
+abadi2012consistency}. On the other hand, weak consistency models like
+eventual consistency~\cite{vogels2009eventually}, PRAM consistency~\cite{lipton1988pram},
+causal consistency~\cite{ahamad1995causal}, and others~\cite{lloyd2011don,
+mehdi2017can} allow data to be replicated with high availability and low
+latency, but they put a tremendous burden on the application designer to reason
+about the complex interleavings of operations that are allowed by these weak
+consistency models. In fact, weak consistency models strip an application
+developer of one of the earliest and most effective tools that is used to
+reason about the execution of programs: application
+invariants~\cite{hoare1969axiomatic, balegas2015towards}. Even if every
 transaction executing in a weakly consistent system individually maintains an
@@
 states.
-
-% When people want to replicate data, they have to make a choice between strong and weak consistency.
-When choosing strong consistency, everything makes sense, reasoning is easy.
-But CAP and the latency one tell us that the avilability and latency of our system will be reduced.
-If we choose weak consistency, then things run really fast, but we can no longer reason about the application.
-In fact, we risk the possibility one our greatest tools that we use to reason about software: invariants.
-With weak consistency models , even if every transaction preserves the invariant, the invariant may still be broken when stuff is interleaved.
 
@@
 invariant-confluence decision procedure.}
-
-% Is it possible to have our cake and eat it to?
-Can we maintain invariants without coordination?
-For some applications, yes.
-Invariant confluence (cite bailis) introduced a necessary and sufficient condition for when its possible to have coordination free replication while maintiang invariants.
-If somethign is iconfluent, it can be replicated wihtout coordination.
-However, determining whethere something is invariant-confluent is a very challening task.
-Bailis recognized some patterns of things which are invariant confluent, but ideally we would have a general purpose program to automatically determine it for us.
-The ultimate goal of this paper is to...
-In this paper, we make progress towards a more-general purpose iconfluence decision procedure, improving on the current state-of-the-art approaches.
 
@@
 
-% This is hard.
-Not only undecidable, but also not really an easy thing to do.
-Existing systems often prove a sufficient condition called invariant-closure for invariant-confluence.
-So if the object is invariant-closed, they determine its invariant-confluent, but it its not invariant-closed, the system doesn't know anythign.
-In this paper, we study why the condition is not necessary, finding that the fundamental problem is that of unreachable invariant satisfying states.
-Using this insight, we develop general conditions under which invariant-closure is both necessary and sufficient.
-
 Then, we use these conditions to design a general-purpose interactive
@@
 and it can be automatically checked without any user interaction.
-
-% Then, we use our insight to develop a general purpose interactive invariant-confluence decision procedure.
-Users continuously interact with the decision procedure to eleiminate unreachabel states from the invariant.
-Meanwhile, the decision procedure provides examples to the user that they can use to identify which points are reachable and which are not.
-This decision procedure can be used to show lots of things are invariant-confluent, and is also robust to user error.
-
-% While the general purpose interaactive dp is good, it does rely on user input.
-We additionally invent a new sufficient condition for invariant-confluence known as merge-reducbility which exploits the insigths from the previous theory.
-This new sufficient condition can handle cases that invariant-closure cannot and doesn't require user feedback.
 
@@
 require more segments which require more coordination, and vice-versa.
-
-% Finally, we address the elephant in the room and ask the question, what if something is not invariant-confluent though?
-We develop a generalization of invariant-confluence called segmented invariant-confluence.
-While invariant-confluence only characterizes objects which can be replicated without any coordination, segmented invariant-confluence allows us to replicate objects with minimal coordination.
-This approach also generalizes existing lock based approaches to handling non-invariant-confluent stuff.
-
-% Finally, we implement all the decision procedures and the system models in a system called Lucy.
-We perform an evaluation to look at when the decision procedures are effective and measure the performance benefits of our implemetnation.
-Overall,
 
@@
 \end{itemize}
-
-% Hype up problem that invariant confluence solves
-% Introduce iconfluence and hype it up as a really good thing.
-% Discuss how it's hard to determine iconfluence and a lot of existing
-% approaches fall short.
-% Describe how we take the problem one step further
-% Overview two new approaches
-% Describe how often times we're not iconfluent
-% Overivew new segmented iconfluence stuff
-% contributions
-%   - new expression-based definitions of iconfluence
-%   - understanding of iclosure and new suff and necc
-%   - interactive decision procedure
-%   - new suff conditions
-%   - segmented iconfluence
-%   - implementation in lucy + evaluation...

### 2018-04-18 doc/paper/sections/related_work.tex
Work on related work.

\section{Related Work}

\paragraph{Super relevant stuf}
% Super relevant
% Towards Fast Invariant Preservation in Geo-replicated Systems
% title={Coordination avoidance in database systems},
% title={Putting consistency back into eventual consistency},
% title={The homeostasis protocol: Avoiding transaction coordination through program analysis},
% title={Automating the choice of consistency levels in replicated systems},
% title={Making geo-replicated systems fast as possible, consistent when necessary},
% title={'Cause I'm strong enough: Reasoning about consistency choices in distributed systems},
% Feral
% The CISE Tool: Proving Weakly-Consistent Applications Correct
% Declarative Programming over Eventually Consistent Data Stores
% Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants

\paragraph{Program Analysis in Database Systems}
% Program analysis in database:
%   look at some of alvins work and related stuff:
%   Find work on static partitioning of data based on program?
%   Transaction chopping / healing?

\paragraph{CALM Theorem}
\cite{alvaro2013declarative}
\cite{ameloot2013relational}
\cite{ameloot2016weaker}
\cite{hellerstein2010declarative}
\cite{alvaro2011dedalus}
\cite{alvaro2010boom}
\cite{alvaro2011consistency}
\cite{conway2012logic}
\cite{zinn2012win}

\paragraph{CRDTs}
% coordination free convergence as opposed to coordination free invs describe inv-confluence as sort of a crdt with certain properties or something.

### 2018-04-19 doc/paper/sections/related_work.tex
Progress on related work.

@@
 
+\newcommand{\sieve}{\textsc{sieve}}
+\paragraph{RedBlue Consistency and \sieve}
+RedBlue consistency is a form of consistency that sits between causal
+consistency and linearizability~\cite{li2012making}. With RedBlue consistency,
+every operation is labelled as red or blue. All operations are executed with
+causal consistency, but with the added restrictions that red operations are
+executed in a single total order embedded within the causal ordering.
+RedBlue consistency is similar to segmented invariant-confluence with a single
+segment in which all red operations are removed from the segment's set of
+allowable transactions.  In~\cite{li2012making}, Li et al.\ define
+\defword{invariant safety} as a sufficient condition for RedBlue consistent
+objects to be invariant-confluent.  Invariant safety is an analog of
+invariant-closure for an operation-based (as opposed-to state
+based~\cite{shapiro2011conflict}) system model.  In~\cite{li2014automating}, Li
+et al.\ develop techniques to identify which operations are invariant safe.
+Like invariant-closure, invariant safety is a sufficient but not necessary
+condition for invariant-confluence. Li et al.\ develop sophisticated techniques
+for deciding invariant safety that involve calculating weakest preconditions.
+These techniques are complementary to our work and can be used to improve the
+invariant-closure subroutine used by our decision procedures.
+
 \paragraph{Super relevant stuf}
+
 % Super relevant
+% title={'Cause I'm strong enough: Reasoning about consistency choices in distributed systems},
+
+% escrow
+
 % Towards Fast Invariant Preservation in Geo-replicated Systems
@@
 % title={The homeostasis protocol: Avoiding transaction coordination through program analysis},
-% title={Automating the choice of consistency levels in replicated systems},
-% title={Making geo-replicated systems fast as possible, consistent when necessary},
 % title={'Cause I'm strong enough: Reasoning about consistency choices in distributed systems},
@@
 
+\paragraph{CRDTs}
+CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
+semilattices with inflationary update methods. Due to their algebraic
+properties, CRDTs can be replicated with strong eventual consistency without
+the need for any coordination. Our definition of distributed objects and our
+invariant-confluence system model are inspired directly by the corresponding
+definitions and system models in~\cite{shapiro2011conflict}. Note that an
+invariant-confluent object does not necessarily have to be eventually
+consistent. In fact, equipping a distributed object with the trivial merge
+function $x \join y = x$ makes the object invariant-confluent with respect
+any set of transactions and any invariant, but it makes the object very
+inconsistent. Thus, it's natural (though not necessary) to use CRDTs as
+invariant-confluent distributed objects, achieving a combination of eventual
+consistency and invariant-preservation. Our criteria in
+\thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
+algebraic properties of semilattices.
+
 \paragraph{CALM Theorem}
-\cite{alvaro2013declarative}
-\cite{ameloot2013relational}
-\cite{ameloot2016weaker}
-\cite{hellerstein2010declarative}
-\cite{alvaro2011dedalus}
-\cite{alvaro2010boom}
-\cite{alvaro2011consistency}
-\cite{conway2012logic}
-\cite{zinn2012win}
+Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
+formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
+declarative Datalog-based programming languages that are designed to program
+distributed systems. The accompanying CALM
+theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
+a program can be written in the monotone fragment of these languages, then
+there exists a coordination-free implementation of the program. Moreover,
+monotone programs are invariant to message reordering, whereas non-monotone
+programs may experience anomalies if messages are reordered in the network. The
+CALM theorem provides guarantees about the outputs of a program, but it does
+not guarantee that invariants are maintained throughout the duration of an
+execution. Moreover, Bloom and Dedalus are general purpose programming
+languages that can be used to implement a variety of distributed systems (e.g.
+a sharded database), while invariant-confluence is defined strictly in the
+context of replicated data.
 
-\paragraph{CRDTs}
-% coordination free convergence as opposed to coordination free invs describe inv-confluence as sort of a crdt with certain properties or something.
+\paragraph{Proof Assistants}
+Our interactive decision procedure bears resemblance to interactive proof
+assistants like Coq~\cite{coq2017}, Isabelle~\cite{nipkow2002isabelle},
+Agda~\cite{norell2008dependently}, and Idris~\cite{brady2013idris}. Proof
+assistants guide users through the process of interactively provided a proof
+that witnesses the validity of a formula. Our interactive decision procedure
+guides users through the process of interactively generating a set of reachable
+and unreachable points that witness the invariant-confluence of an object. An
+interesting avenue for future work is to integrate an proof assistant into our
+interactive decision procedure. For example, a proof assistant could verify
+that a state labelled unreachable by a user is actually unreachable.

### 2018-04-19 doc/paper/sections/related_work.tex
More progress on related work.

@@
 \section{Related Work}
+We now discuss RedBlue consistency~\cite{li2012making, li2014automating}, the
+homeostasis protocol~\cite{roy2015homeostasis}, explicit
+consistency~\cite{balegas2015putting}, and token-based
+invariant-confluence~\cite{gotsman2016cause}: four very relevant pieces of
+work. We also briefly survey other related research in the area of
+invariant-confluence including program analysis in database systems, CRDTs, the
+CALM theorem, and interactive proof assistants.
 
@@
 allowable transactions. In~\cite{li2012making}, Li et al.\ define
-\defword{invariant safety} as a sufficient condition for RedBlue consistent
-objects to be invariant-confluent.  Invariant safety is an analog of
-invariant-closure for an operation-based (as opposed-to state
-based~\cite{shapiro2011conflict}) system model.  In~\cite{li2014automating}, Li
-et al.\ develop techniques to identify which operations are invariant safe.
-Like invariant-closure, invariant safety is a sufficient but not necessary
-condition for invariant-confluence. Li et al.\ develop sophisticated techniques
-for deciding invariant safety that involve calculating weakest preconditions.
-These techniques are complementary to our work and can be used to improve the
-invariant-closure subroutine used by our decision procedures.
+invariant safety as a sufficient condition for RedBlue consistent objects to be
+invariant-confluent. Invariant safety is an analog of invariant-closure for an
+operation-based (as opposed-to state based~\cite{shapiro2011conflict}) system
+model. In~\cite{li2014automating}, Li et al.\ develop techniques to identify
+which operations are invariant safe.  Like invariant-closure, invariant safety
+is a sufficient but not necessary condition for invariant-confluence. Li et
+al.\ develop sophisticated techniques for deciding invariant safety that
+involve calculating weakest preconditions.  These techniques are complementary
+to our work and can be used to improve the invariant-closure subroutine used by
+our decision procedures.
 
-\paragraph{Super relevant stuf}
+\paragraph{The Homeostasis Protocol}
+The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
+avoid unnecessary coordination between servers in a sharded database. The
+protocol generates global invariants, called global treaties, and servers can
+operate without coordination so long as they preserve the treaty. When the
+treaty is violated, the nodes perform a round of coordination. The protocol
+guarantees that transactions are executed with observational equivalence with
+respect to some serial execution of the transactions. This means that
+intermediate states may be inconsistent, but externally observable side effects
+and the final database state are consistent. The observational equivalence
+guaranteed by the homeostasis is stronger than the guarantees of
+invariant-confluence. As a result, there are invariants and transaction
+workloads which the homeostasis protocol would execute with more coordination
+than a segmented invariant-confluent execution. Still, the homeostasis
+protocol's mechanism of establishing invariants and operating without
+coordination so long as the invariants are maintained is very similar to our
+design of segmented invariant-confluence.
 
-% Super relevant
-% title={'Cause I'm strong enough: Reasoning about consistency choices in distributed systems},
+\paragraph{Explicit Consistency}
+A system provides explicit consistency with respect to a partial ordering of
+transactions if every server executes the transactions in a total order that
+respects the partial order and in a way such that the invariant is always
+preserved~\cite{balegas2015towards}. Thus, explicit consistency is like a
+combination of invariant-confluence and causal consistency, similar to RedBlue
+consistency. To determine if a workload is amenable to explicitly consistent
+replication, Balegas et al. determine if all pairs of transactions can be
+concurrently executed on the same start state without violating the invariant~\cite{balegas2015towards}.
+This is a sufficient condition for explicit consistency similar to criterion
+(3) in \thmref{LatticeProperty}. Balegas et al. also describe a variety of
+techniques---like conflict resolution, locking, and escrow
+transactions~\cite{o1986escrow}---that can be used to replicate workloads that
+do meet their sufficient conditions. Conflict resolution involves choosing an
+appropriate merge function; locking can be implemented as a single segment
+segmentation with certain transactions disallowed; and escrow transactions can
+be emulated with a segmentation that allows for concurrent decrements only when
+value being decremented is sufficiently large.
 
-% escrow
+\paragraph{Token-Based Invariant-Confluence}
+In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
+consistency model that generalizes RedBlue consistency. An application designer
+defines a set of tokens and specifies which pairs of tokens conflict. Then,
+transactions acquire some subset of the tokens. This allows the application
+designer to specify which transactions conflict at a fine granularity. With no
+tokens, the model is equivalent to causal consistency. With all transactions
+acquiring a single self-conflicting token, the model is equivalent to
+sequential consistency~\cite{lamport1979make}. With some transactions acquiring
+a single self-conflicting token and some transactions acquiring no tokens, the
+model is equivalent to RedBlue consistency. Gotsman et al.\ develop sufficient
+conditions to determine whether a given token scheme is sufficient to guarantee
+that a global invariant is never broken. This model of consistency is very
+similar to segmented invariant-confluence. The token-based approach allows
+users to specify certain conflicts that are not possible with segmented
+invariant-confluence. A segmentation only allows transactions within a segment
+to acquire a single self-conflicting lock. However, segmented
+invariant-confluence also introduces the notion of segmenting the invariant,
+which cannot be emulated with the token based approach. For example,
+implementing \exampleref{SegmentedZ2} with the token-based approach is
+challenging without being able to segment the top-left quadrant from the
+bottom-right quadrant. An interesting direction for future work is to try and
+integrate the token-based approach into segmented invariant-confluence.
 
 % Towards Fast Invariant Preservation in Geo-replicated Systems
-% title={Coordination avoidance in database systems},
-% title={Putting consistency back into eventual consistency},
-% title={The homeostasis protocol: Avoiding transaction coordination through program analysis},
-% title={'Cause I'm strong enough: Reasoning about consistency choices in distributed systems},
+%
 % Feral
@@
 \paragraph{Program Analysis in Database Systems}
+TODO
 % Program analysis in database:

### 2018-04-19 doc/paper/sections/related_work.tex
Finished first cut at related work.

@@
 
-\paragraph{Program Analysis in Database Systems}
-TODO
-% Program analysis in database:
-%   look at some of alvins work and related stuff:
-%   Find work on static partitioning of data based on program?
-%   Transaction chopping / healing?
-
 \paragraph{CRDTs}
@@
 
+\paragraph{Program Analysis in Database Systems}
+Using program analysis to improve the performance of database systems is not
+new. For example, it has been used to improve the performance of
+database-backed web applications~\cite{cheung2014using, wu2016transaction,
+ramachandra2012program} and used to improve the performance of optimistic
+concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
+on invariant-confluence continues the theme of using program analysis to reap
+the performance benefits of understanding program semantics that would
+otherwise be left on the table.
+
 \paragraph{Proof Assistants}

### 2018-04-19 doc/paper/sections/related_work.tex
Formatting tweaks.

@@
 \newcommand{\sieve}{\textsc{sieve}}
-\paragraph{RedBlue Consistency and \sieve}
+\textbf{RedBlue Consistency and \sieve.}
 RedBlue consistency is a form of consistency that sits between causal
@@
 
-\paragraph{The Homeostasis Protocol}
+\textbf{The Homeostasis Protocol.}
 The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
@@
 
-\paragraph{Explicit Consistency}
+\textbf{Explicit Consistency.}
 A system provides explicit consistency with respect to a partial ordering of
@@
 
-\paragraph{Token-Based Invariant-Confluence}
+\textbf{Token-Based Invariant-Confluence.}
 In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
@@
 
-\paragraph{CRDTs}
+\textbf{CRDTs.}
 CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
@@
 
-\paragraph{CALM Theorem}
+\textbf{CALM Theorem.}
 Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
@@
 
-\paragraph{Program Analysis in Database Systems}
+\textbf{Program Analysis in Database Systems.}
 Using program analysis to improve the performance of database systems is not
@@
 
-\paragraph{Proof Assistants}
+\textbf{Proof Assistants.}
 Our interactive decision procedure bears resemblance to interactive proof

### 2018-04-21 iconfluence/stratified/arbitrary_start_checker.py
Couple of bug fixes.

@@ -15,6 +15,25 @@ class ArbitraryStartChecker(Checker):
     """
     TODO(mwhittaker): Document.
     """
+    def __str__(self):
+        strings = []
+
+        strings.append('State')
+        for name in self.crdt_env:
+            strings.append(f'  {name}: {self.crdt_env[name]}')
+
+        strings += ['Invariants']
+        for (name, inv) in self.invariants.items():
+            strings.append(f'  {name} == {inv}')
+
+        strings += ['Transactions']
+        for (name, txn) in self.transactions.items():
+            strings.append(f'  def {name}():')
+            for s in txn:
+                strings.append(f'    {s}')
+
+        return '\n'.join(strings)
+
     def _warn_if_not_none(self, val: Coercible) -> None:
         if val is not None:
             _wrap_print(f'WARNING: You have provided a start value {val} to ' +

### 2018-04-21 iconfluence/stratified/stratified_checker.py

@@ -192,7 +192,7 @@ class StratifiedChecker(Checker):
                 elif isinstance(crdt, ast.CSetIntersect):
                     checker.set_intersect(name, crdt.a, None)
                 elif isinstance(crdt, ast.CMap):
-                    checker.map(name, crdt.k, crdt.v, None)
+                    checker.map(name, crdt.a, crdt.b, None)
                 elif isinstance(crdt, ast.COption):
                     checker.option(name, crdt.a, None)
                 else:

### 2018-04-21 examples/bank_deposit_and_withdraw.py
Added escrow!

@@ -21,6 +21,7 @@ p2 = checker.int_max('p2', 0)
 m0 = checker.int_max('m0', 0)
 m1 = checker.int_max('m1', 0)
 m2 = checker.int_max('m2', 0)
+
 checker.add_invariant('p0_geq_0', p0 >= 0)
 checker.add_invariant('p1_geq_0', p1 >= 0)
 checker.add_invariant('p2_geq_0', p2 >= 0)
@@ -28,10 +29,44 @@ checker.add_invariant('m0_geq_0', m0 >= 0)
 checker.add_invariant('m1_geq_0', m1 >= 0)
 checker.add_invariant('m2_geq_0', m2 >= 0)
 checker.add_invariant('x_geq_0', ((p0 + p1 + p2) - (m0 + m1 + m2)) >= 0)
-for i in range(10):
-    checker.add_transaction(f'x0_inc_{i}', [p0.assign(p0 + i)])
-    checker.add_transaction(f'x1_inc_{i}', [p1.assign(p1 + i)])
-    checker.add_transaction(f'x2_inc_{i}', [p2.assign(p2 + i)])
-    checker.add_transaction(f'x0_dec_{i}', [m0.assign(m0 + i)])
-    checker.add_transaction(f'x1_dec_{i}', [m1.assign(m1 + i)])
-    checker.add_transaction(f'x2_dec_{i}', [m2.assign(m2 + i)])
+
+checker.add_transaction('x0_inc', [p0.assign(p0 + 1)])
+checker.add_transaction('x1_inc', [p1.assign(p1 + 1)])
+checker.add_transaction('x2_inc', [p2.assign(p2 + 1)])
+checker.add_transaction('x0_dec', [m0.assign(m0 + 1)])
+checker.add_transaction('x1_dec', [m1.assign(m1 + 1)])
+checker.add_transaction('x2_dec', [m2.assign(m2 + 1)])
+
+if isinstance(checker, StratifiedChecker):
+    def add_base_invariants(stratum: Stratum) -> None:
+        stratum.add_invariant(p0 >= 0)
+        stratum.add_invariant(p1 >= 0)
+        stratum.add_invariant(p2 >= 0)
+        stratum.add_invariant(m0 >= 0)
+        stratum.add_invariant(m1 >= 0)
+        stratum.add_invariant(m2 >= 0)
+        stratum.add_invariant(((p0 + p1 + p2) - (m0 + m1 + m2)) >= 0)
+
[…]

### 2018-04-21 iconfluence/checker.py
Print check times + stratified checker bug fix.

@@ -1,5 +1,6 @@
 from enum import Enum
 from typing import Any, Dict, Optional, Set
+import time
 
 from . import ast
 from . import typecheck
@@ -111,5 +112,12 @@ class Checker:
         inv = typecheck.typecheck_invariant(inv, self.type_env)
         self.invariants[name] = inv
 
-    def check(self) -> Decision:
+    def _check(self) -> Decision:
         raise NotImplementedError()
+
+    def check(self) -> Decision:
+        start = time.time()
+        ret = self._check()
+        stop = time.time()
+        print(f'Check took {stop - start} seconds.')
+        return ret

### 2018-04-21 iconfluence/diamond/diamond_checker.py

@@ -221,7 +221,7 @@ class DiamondChecker(Checker):
                         return False
         return True
 
-    def check(self) -> Decision:
+    def _check(self) -> Decision:
         # Criterion 1 is enforced automatically because the only available
         # datatypes are CRDTs.
 

### 2018-04-21 iconfluence/guess_and_check/guess_and_check_checker.py

@@ -24,7 +24,7 @@ class GuessAndCheckChecker(checker.Checker):
         invs = self.invariants.values()
         return all(eval_invariant(inv, state) for inv in invs)
 
-    def check(self) -> checker.Decision:
+    def _check(self) -> checker.Decision:
         if self.verbose:
             print(f'Exploring {self.num_states} states...')
         self.state_explorer.explore(self.num_states)

### 2018-04-21 iconfluence/interactive/interactive_checker.py

@@ -281,7 +281,7 @@ class InteractiveChecker(Checker):
 
             return Decision.UNKNOWN
 
-    def check(self) -> Decision:
+    def _check(self) -> Decision:
         # Make sure that both counterexamples are labelled, if they exist.
         msg = ('State {0} is unlabelled. Call `{0}_reachable()` to label ' +
                 'the state as reachable or `{0}_unreachable()` to label the ' +

### 2018-04-21 iconfluence/stratified/arbitrary_start_interactive_checker.py

@@ -137,7 +137,7 @@ class ArbitraryStartInteractiveChecker(ArbitraryStartChecker):
             zss.add(to_var_z == from_var_z)
         return zss, to_venv
 
-    def check(self) -> Decision:
+    def _check(self) -> Decision:
         if self.lhs is not None:
             assert self.rhs is not None, self.rhs
             if self.label is None:

### 2018-04-21 iconfluence/stratified/stratified_checker.py

@@ -121,7 +121,7 @@ class StratifiedChecker(Checker):
         self.strata.append(s)
         return s
 
-    def check(self) -> Decision:
+    def _check(self) -> Decision:
         # Check that there is at least one stratum.
         if len(self.strata) == 0:
             _wrap_print('Create strata with the stratum() method before ' +
@@ -135,8 +135,15 @@ class StratifiedChecker(Checker):
                 return Decision.YES
             elif Decision.NO in decisions:
                 return Decision.NO
+            elif Decision.UNKNOWN in decisions:
+                return Decision.UNKNOWN
             else:
-                assert Decision.UNKNOWN in decisions, decisions
+                assert None in decisions, decisions
+                _wrap_print('We\'re not sure if this object is ' +
+                            'invariant-confluent. Please call check() on '
+                            'each stratum and follow the instructions ' +
+                            'provided to determine if each stratum is ' +
+                            'invariant-confluent. Then, call check() again.')
                 return Decision.UNKNOWN
 
         # Check that the start state satisfies the invariant.

### 2018-04-21 examples/auction.py
Tidying up examples.

@@ -19,11 +19,26 @@ Run with
 checker = get_checker()
 bids = checker.set_union('bids', TInt(), EEmptySet(TInt()))
 winning_bid = checker.option('winning_bid', CIntMax(), ENone(TInt()))
+
+# Invariant.
 checker.add_invariant(
     'winning_bid_is_max',
     winning_bid.is_some() >> winning_bid.unwrap().eq(ESetMax(bids)))
+
+# Transactions.
 checker.add_transaction('close', [
     winning_bid.assign(ESome(ESetMax(bids)))
 ])
 for i in range(5):
     checker.add_transaction(f'bid_{i}', [bids.assign(bids.union({i}))])
+
+if isinstance(checker, StratifiedChecker):
+    auction_open = checker.stratum('auction_open')
+    auction_open.add_invariant(winning_bid.is_none())
+    auction_open.add_transaction('close')
+    for i in range(5):
+        auction_open.add_transaction(f'bid_{i}')
+
+    auction_closed = checker.stratum('auction_closed')
+    auction_closed.add_invariant(winning_bid.is_some())
+    auction_closed.add_invariant(winning_bid.unwrap().eq(ESetMax(bids)))

### 2018-04-21 examples/foreign_key.py

@@ -18,9 +18,16 @@ XA = checker.set_union('XA', TInt(), {1, 2})
 XR = checker.set_union('XR', TInt(), EEmptySet(TInt()))
 YA = checker.set_union('YA', TInt(), {1, 2, 3, 4})
 YR = checker.set_union('YR', TInt(), EEmptySet(TInt()))
+
+# Invariant.
 checker.add_invariant(
     'X_subset_Y',
     XA.diff(XR).subseteq(YA.diff(YR)))
+
+# Transactions.
 for i in range(5):
     checker.add_transaction(f'X_sub_{i}', [XR.join_assign({i})])
     checker.add_transaction(f'Y_add_{i}', [YA.join_assign({i})])
+
+# checker.refine_invariant(XA.eq({1, 2}))
+# checker.refine_invariant(YR.eq(EEmptySet(TInt())))

### 2018-04-21 doc/paper/sections/evaluation.tex
Added case studies to eval.

@@
 \section{Evaluation}\seclabel{Evaluation}
-In this section, we describe Lucy: an implementation of our decision procedures
-and system models. We then evaluate Lucy by answering the following questions:
-\begin{itemize}
-  \item
-    How practical is the interactive invariant-confluence decision procedure?
-    Can we use it to classify real-world transactions and invariants?
-  \item
-    How practical is segmented invariant-confluence? Are real-world workloads
-    amenable to segmentation?
-  \item
-    How efficient is the interactive invariant-confluence decision procedure?
-  \item
-    How efficiently can we replicate a segmented invariant-confluent object as
-    compared to alternative approaches like replicating with weak or strong
-    consistency?
-  \item
-    How does the performance of replicating a segmented invariant-confluence
-    object vary as we vary the workload, segmentation, and replication factor?
-\end{itemize}
+In this section, we describe and evaluate Lucy: a prototype implementation of
+our decision procedures and system models.
+% We evaluate Lucy by answering the following questions:
+% \begin{itemize}
+%   \item
+%     How practical is the interactive invariant-confluence decision procedure?
+%     Can we use it to classify real-world transactions and invariants?
+%   \item
+%     How practical is segmented invariant-confluence? Are real-world workloads
+%     amenable to segmentation?
+%   \item
+%     How efficient is the interactive invariant-confluence decision procedure?
+%   \item
+%     How efficiently can we replicate a segmented invariant-confluent object as
+%     compared to alternative approaches like replicating with weak or strong
+%     consistency?
+%   \item
+%     How does the performance of replicating a segmented invariant-confluence
+%     object vary as we vary the workload, segmentation, and replication factor?
+% \end{itemize}
 
@@
 Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
-compiling objects and invariants into a formula which is satisfiable if and
+compiling an object and invariant into a formula that is satisfiable if and
 only if the object is \emph{not} invariant-closed. If the object is
@@
 \subsection{Decision Procedures}
+We now evaluate the practicality and efficiency of our decision procedure
+prototypes. We begin by demonstrating the decision procedure on a handful of
+simple, yet practical examples. We then, discuss TODO
+
 % \todo{Bank account with and without decrements.}
@@
 % \todo{See which of the Feral and TPC transactions we can determine automatically}
+\example[$\ints^2$]\examplelabel{TwoIntsEval}
+We begin with a minimal working example. Consider again our recurring example
+of $\ints^2$ from \exampleref{Z2}. The Python code used to describe the object,
+transactions, and invariant is given in \figref{Z2Code}. When we call
+\python{checker.check()}, the interactive decision procedure produces
+a counterexample in less than a tenth of a second.  After we label the
+counterexample and refine the invariant with $y \leq 0$, the interactive
+decision procedure determines that the object is invariant-confluent, again, in
+less than a tenth of a second. Note that the object is invariant-confluent but
+\emph{not} invariant-closed, so prior work that relies on invariant-closure
+alone to determine invariant-confluence would not be able to identify this
+example as invariant-confluent.
+
+\begin{figure}[ht]
+  \begin{Python}[gobble=4]
+    checker = InteractiveInvariantConfluenceChecker()
+    x = checker.int_max('x', 0) # An int, x, merged by max.
+    y = checker.int_max('y', 0) # An int, y, merged by max.
+    checker.add_transaction('increment_x', [x.assign(x + 1)])
+    checker.add_transaction('decrement_y', [y.assign(y - 1)])
+    checker.add_invariant(x * y <= 0)
+    checker.check()
+  \end{Python}
+  \caption{\exampleref{TwoIntsEval} specification}\figlabel{Z2Code}
+\end{figure}
+
+\example[Foreign Keys]\examplelabel{ForeignKeysEval}
+A 2P-Set $X = (A_X, R_X)$ is a set CRDT composed of a set of additions $A_X$
+and a set of removals $R_X$~\cite{shapiro2011comprehensive}. We view the state
+of the set $X$ as the difference $A_X - R_X$ of the addition and removal sets.
+To add an element $x$ to the set, we add $x$ to $A_X$. Similarly, to remove $x$
+from the set, we add it to $R_X$. The merge of two 2P-sets is a pairwise union
+(i.e. $(A_X, R_X) \join (A_Y, R_Y) = (A_X \cup A_Y, R_X \cup R_Y)$).
+
+We can use 2P-sets to model a simple relational database with foreign key
+constraints. Let object $O = (X, Y) = ((A_X, R_X), (A_Y, R_Y))$ consist of a
+pair of two 2P-Sets $X$ and $Y$, which we view as relations. Our invariant $X
+\subseteq Y$ (i.e. $(A_X - R_X) \subseteq (A_Y - R_Y)$) models a foreign key
+constraint from $X$ to $Y$. We ran our decision procedure on the object with
+initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and
+transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
+After less than a tenth of a second, the decision procedure produced a
+reachable counterexample witnessing the fact that the object is not
+invariant-confluent. A concurrent insertion into $X$ and deletion from $Y$ can
+lead to a state which violates the invariant. This object is not
+invariant-confluent and therefore not invariant-closed. Thus, previous tools
+depending on invariant-closure as a sufficient condition would be unable to
+conclude definitively that the object is \emph{not} invariant-confluent.
+
+We also reran the decision procedure, but this time with insertions into $X$
+and deletions from $Y$ disallowed. In less than a tenth of a second, the
+decision procedure correctly deduced that the object is now
+invariant-confluent. These results were manually proven
+in~\cite{bailis2014coordination}, but our tool was able to confirm them
+automatically in a negligible amount of time.
+
+\example[Auction]\examplelabel{AuctionEval}
+We now consider a simple auction system introduced in~\cite{gotsman2016cause}.
+Our object consists of a set $B$ of integer-valued bids and a optional winning
+bid $w$. Initially, $B = \emptyset$ and $w = \bot$ (indicating that there is no
+winning bid yet) and we merge states by taking the union of $B$ and the maximum
+of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a bid
+$b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes the
+auction and sets $w$ equal to the largest bid in $B$. Our invariant is that if
+the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
+decision procedure on this example and in a third of a second, it produced a
+reachable counterexample witnessing the fact that the object is not
+invariant-confluent.  If we concurrently close the auction and place an large
+bid, then we can end up in a state in which the auction is closed, but there is
+a higher bid in $B$.
+
+We then segmented our object as follows. The first segment $(\setst{(B, w)}{w =
+\bot}, \setst{t_b}{b \in \ints})$ allows any bidding transaction so long as the
+bid is open. The second segment $(\setst{B, w}{I \cap w \neq \bot}, \emptyset)$
+includes all auctions which have already been closed and forbids all
+transactions. This segmentation captures the intuition that bids should be
+permitted only when the auction is open. We ran our segmented
+invariant-confluence decision procedure on this example, and it was able to
+deduce without any human interaction that the example was segmented
+invariant-confluent in less than a tenth of a second.
+
+\example[Escrow Transactions]\examplelabel{EscrowTransactionsEval}
+Escrow transactions are a concurrency control technique that allows a database
+to execute transactions that increment and decrement numeric values with more
+concurrency than is otherwise possible with general-purpose techniques like
+two-phase locking~\cite{o1986escrow}. The main idea is that a portion of the
+numeric value is put in escrow, after which a transaction can freely decrement
+the value so long as it is not decremented by more than the amount that has
+been escrowed. We now show how segmented invariant-confluence can be used to
+implement escrow transactions.
+
+Consider again the PN-Counter $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ from
+\exampleref{CounreachableExample} replicated on three servers with transactions
+to increment and decrement the PN-Counter. In
+\exampleref{CounreachableExample}, we found that concurrent decrements violate
+invariant-confluence which led us to a segmentation which prohibited concurrent
+decrements. We now propose a new segmentation with escrow amount $k$ that will
+allow us to perform concurrent decrements that respect the escrowed value. The
+first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2, p_3 \geq k
+\land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments and
+decrements so long as every $p_i \geq k$ and every $n_i \leq k$. Intuitively,
+this segment represents the situation in which every server has escrowed a
+value of $k$. They can decrement freely, so long as they don't exceed their
+escrow budget of $k$. The second segment is the one presented in
+\exampleref{CounreachableExample} which prohibits concurrent decrements. We ran
+our decision procedure on this example and it concluded that it was segmented
+invariant-confluent in less than a tenth of a second and without any human
+interaction.
+
+\textbf{TPC-C and Feral Concurrency Control.}
+
+TODO
 
@@
 \seclabel{SegmentedInvariantConfluenceEval}
-In this section, we evaluate the performance of replicating an object with
-segmented invariant-confluence as compared to replicating it with eventual
-consistency or linearizability. We begin with two benchmarks that demonstrate
-the same concept: the performance of segmented invariant-confluent replication
-varies with the amount of global coordination induced by either (a) performing
-a transaction that is disallowed within a segment or (b) transitioning between
+Now, we evaluate the performance of replicating an object with segmented
+invariant-confluence as compared to replicating it with eventual consistency or
+linearizability. We begin with two benchmarks that demonstrate the same
+concept: the performance of segmented invariant-confluent replication varies
+with the amount of global coordination induced by either (a) performing a
+transaction that is disallowed within a segment or (b) transitioning between
 segments.

### 2018-04-21 examples/foreign_key.py
Eval bug fix + foreign key new start state.

@@ -14,9 +14,9 @@ Run with
 """
 
 checker = get_checker()
-XA = checker.set_union('XA', TInt(), {1, 2})
+XA = checker.set_union('XA', TInt(), EEmptySet(TInt()))
 XR = checker.set_union('XR', TInt(), EEmptySet(TInt()))
-YA = checker.set_union('YA', TInt(), {1, 2, 3, 4})
+YA = checker.set_union('YA', TInt(), EEmptySet(TInt()))
 YR = checker.set_union('YR', TInt(), EEmptySet(TInt()))
 
 # Invariant.

### 2018-04-21 iconfluence/eval.py

@@ -45,7 +45,11 @@ def eval_expr(e: ast.Expr, env: ValEnv) -> Any:
     elif isinstance(e, ast.ESetMin):
         return min(eval_expr(e.x, env))
     elif isinstance(e, ast.ESetMax):
-        return max(eval_expr(e.x, env))
+        xs = eval_expr(e.x, env)
+        if len(xs) == 0:
+            return None
+        else:
+            return max(xs)
     elif isinstance(e, ast.ETuple2First):
         return eval_expr(e.x, env)[0]
     elif isinstance(e, ast.ETuple2Second):

### 2018-04-21 doc/paper/sections/related_work.tex
Removed related workds section on proof assistants.

@@
 
-\textbf{Proof Assistants.}
-Our interactive decision procedure bears resemblance to interactive proof
-assistants like Coq~\cite{coq2017}, Isabelle~\cite{nipkow2002isabelle},
-Agda~\cite{norell2008dependently}, and Idris~\cite{brady2013idris}. Proof
-assistants guide users through the process of interactively provided a proof
-that witnesses the validity of a formula. Our interactive decision procedure
-guides users through the process of interactively generating a set of reachable
-and unreachable points that witness the invariant-confluence of an object. An
-interesting avenue for future work is to integrate an proof assistant into our
-interactive decision procedure. For example, a proof assistant could verify
-that a state labelled unreachable by a user is actually unreachable.
+% \textbf{Proof Assistants.}
+% Our interactive decision procedure bears resemblance to interactive proof
+% assistants like Coq~\cite{coq2017}, Isabelle~\cite{nipkow2002isabelle},
+% Agda~\cite{norell2008dependently}, and Idris~\cite{brady2013idris}. Proof
+% assistants guide users through the process of interactively provided a proof
+% that witnesses the validity of a formula. Our interactive decision procedure
+% guides users through the process of interactively generating a set of reachable
+% and unreachable points that witness the invariant-confluence of an object. An
+% interesting avenue for future work is to integrate an proof assistant into our
+% interactive decision procedure. For example, a proof assistant could verify
+% that a state labelled unreachable by a user is actually unreachable.

### 2018-04-21 example.py
Removed random file.

@@ -1,183 +1,0 @@
-from typing import Callable, Dict
-
-from iconfluence import *
-
-def iconfluent_example(checker: Checker) -> Decision:
-    x = checker.int_max('x')
-    y = checker.int_max('y')
-    tmp = checker.int_max('tmp')
-
-    checker.add_transaction('incr', [
-        x.assign(x + 1),
-        y.assign(y + 1),
-    ])
-    checker.add_transaction('swap', [
-        tmp.assign(x),
-        x.assign(y),
-        y.assign(tmp),
-    ])
-
-    checker.add_invariant('x_eq_y', x.eq(y))
-
-    return checker.check_iconfluence()
-
-def not_iconfluent_example(checker: Checker) -> Decision:
-    x = checker.int_min('x')
-    y = checker.int_max('y')
-    checker.add_transaction('decr_x', [x.assign(x - 2)])
-    checker.add_transaction('incr_y', [y.assign(y + 2)])
-    checker.add_invariant('x_ge_y', x >= y)
-    return checker.check_iconfluence()
-
-def all_datatypes_example(checker: Checker) -> Decision:
-    x_int_max = checker.int_max('x_int_max')
-    y_int_max = checker.int_max('y_int_max')
-    x_int_min = checker.int_min('x_int_min')
-    y_int_min = checker.int_min('y_int_min')
-    x_bool_or = checker.bool_or('x_bool_or')
-    y_bool_or = checker.bool_or('y_bool_or')
-    x_bool_and = checker.bool_and('x_bool_and')
-    y_bool_and = checker.bool_and('y_bool_and')
-    x_set_union = checker.set_union('x_set_union', TInt())
-    y_set_union = checker.set_union('y_set_union', TInt())
-    x_set_intersect = checker.set_intersect('x_set_intersect', TInt())
-    y_set_intersect = checker.set_intersect('y_set_intersect', TInt())
-    x_tuple2 = checker.tuple2('x_tuple2', CIntMax(), CIntMin())
-    y_tuple2 = checker.tuple2('y_tuple2', CIntMax(), CIntMin())
-    x_option = checker.option('x_option', CIntMax())
-    y_option = checker.option('y_option', CIntMax())
-    x_map = checker.map('x_map', TInt(), CIntMax())
-    y_map = checker.map('y_map', TInt(), CIntMax())
-
-    checker.add_transaction('t1', [
-        x_int_max.assign(((y_int_max + 1) * 2) - 3),
[…]

### 2018-04-22 examples/parser.py
Added TPC example.

@@ -15,13 +15,16 @@ def _get_parser() -> ArgumentParser:
     parser.add_argument('--verbose', '-v', action='store_true')
     parser.add_argument('--checker', '-c',
                         choices=checkers, default='interactive')
+    parser.add_argument('--num_explored_per_step', '-n', type=int, default=100)
     return parser
 
 def get_checker() -> Checker:
     parser = _get_parser()
     args = parser.parse_args()
     if args.checker == 'interactive':
-        return InteractiveChecker(verbose=args.verbose)
+        return InteractiveChecker(
+            verbose=args.verbose,
+            num_explored_states_per_step=args.num_explored_per_step)
     elif args.checker == 'guess_and_check':
         return GuessAndCheckChecker(verbose=args.verbose)
     elif args.checker == 'diamond':

### 2018-04-22 examples/sum_eq.py

@@ -1,0 +1,16 @@
+from iconfluence import *
+
+from .parser import get_checker
+
+checker = get_checker()
+x = checker.int_max('x', 0)
+y = checker.int_max('y', 0)
+z = checker.int_max('z', 0)
+checker.add_invariant('x_geq_0', x >= 0)
+checker.add_invariant('y_geq_0', y >= 0)
+checker.add_invariant('z_geq_0', z >= 0)
+checker.add_invariant('x_plus_y_eq_z', (x + y).eq(z))
+for i in range(10):
+    checker.add_transaction(f'x_inc_{i}', [x.assign(x + i)])
+    checker.add_transaction(f'y_inc_{i}', [y.assign(y + i)])
+    checker.add_transaction(f'z_inc_{i}', [z.assign(z + i)])

### 2018-04-22 examples/tpcc01.py

@@ -1,0 +1,72 @@
+from typing import Dict
+
+from iconfluence import *
+
+from .parser import get_checker
+
+num_districts = 20
+num_servers = 3
+num_txns_per_var = 2
+
+checker = get_checker()
+if isinstance(checker, StratifiedChecker):
+    stratum = checker.stratum('stratum')
+variables: Dict[str, EVar] = dict()
+
+# District sums.
+district_sum = coerce(0)
+per_server_district_sum = {s: coerce(0) for s in range(num_servers)}
+for server_id in range(num_servers):
+    for district_id in range(num_districts):
+        name = f'd_{district_id}_{server_id}'
+        var = checker.int_max(name, 0)
+
+        variables[name] = var
+        checker.add_invariant(f'{name}_geq_0', var >= 0)
+        if isinstance(checker, StratifiedChecker):
+            stratum.add_invariant(var >= 0)
+        district_sum += var
+        per_server_district_sum[server_id] += var
+        for x in range(num_txns_per_var):
+            checker.add_transaction(f'{name}_plus_{x}', [var.assign(var + x)])
+            if isinstance(checker, StratifiedChecker):
+                stratum.add_transaction(f'{name}_plus_{x}')
+
+# Warehouse sums.
+warehouse_sum = coerce(0)
+for server_id in range(num_servers):
+    name = f'w_{server_id}'
+    var = checker.int_max(name, 0)
+
+    variables[name] = var
+    checker.add_invariant(f'{name}_geq_0', var >= 0)
+    checker.add_invariant(f'per_server_{server_id}_eq',
+                          var.eq(per_server_district_sum[server_id]))
+    if isinstance(checker, StratifiedChecker):
+        stratum.add_invariant(var >= 0)
+        stratum.add_invariant(var.eq(per_server_district_sum[server_id]))
+    warehouse_sum = warehouse_sum + var
+
+checker.add_invariant('sum_eq', warehouse_sum.eq(district_sum))
+if isinstance(checker, StratifiedChecker):
+    stratum.add_invariant(warehouse_sum.eq(district_sum))
+
+checker.check()
+
+if isinstance(checker, StratifiedChecker):
+    for server_id in range(num_servers):
+        lhs_vars: List[EVar] = []
+        rhs_vars: List[EVar] = []
+        for district_id in range(num_districts):
[…]

### 2018-04-22 doc/paper/sections/evaluation.tex
Finished TPC section.

@@
 
-\textbf{TPC-C and Feral Concurrency Control.}
+\example[TPC-C]\examplelabel{TpccEval}
+TPC-C is a ubiquitous OLTP database benchmark with a workload that models a
+simple warehousing application~\cite{difallah2013oltp}. The TPC-C specification
+outlines twelve ``consistency requirements'' (read invariants) that govern the
+warehousing application. In~\cite{bailis2014coordination}, Bailis et al.\
+categorize the invariants into one of three types:
+\begin{enumerate}
+  \item
+    Three of the twelve invariants are foreign key constraints.  As discussed
+    in \exampleref{ForeignKeysEval}, our decision procedures can automatically
+    verify conditions under which foreign key constraints are
+    invariant-confluent.
 
-TODO
+  \item
+    \newcommand{\ttt}[1]{{\smaller \texttt{#1}}}
+    Seven of the twelve invariants involve maintaining arithmetic relationships
+    between relations. Our decision procedures can correctly identify these as
+    invariant-confluent. Consider, for example, invariant 1 which dictates that
+    a warehouse's year to date balance \ttt{W\_YTD} is equal to the sum of the
+    district year to date balances \ttt{D\_YTD} of the twenty districts that
+    are associated with the warehouse. The Payment transaction randomly selects
+    a district and increments \ttt{W\_YTD} and \ttt{D\_YTD} by a randomly
+    generated amount. We model this workload with a PN-Counter for \ttt{W\_YTD}
+    and twenty PN-Counters for the twenty instances of \ttt{D\_YTD}. We applied
+    Lucy to this workload, and it determined that the workload was
+    invariant-confluent in less than a second.
+
+  \item
+    Two of the twelve invariants involve generating sequential and unique
+    identifiers. This workload is \emph{not} invariant-confluent, but when the
+    sequentiality requirement is dropped, it is invariant-confluent.
+    Unfortunately, modelling unique key generation in our theoretical framework
+    (and in our prototype implementation) is not possible because we have thus
+    far tacitly assumed that transaction execution is deterministic. We leave a
+    generalization of our theory to non-deterministic transactions for future
+    work.
+\end{enumerate}
 

### 2018-04-22 doc/paper/sections/abstract.tex
Revised abstract with eval results.

@@
 non-invariant-confluent objects with a small and tunable amount of
-coordination. Implementing our theoretical findings in a system called Lucy, we
-evaluate our methods on a number of case studies and benchmarks showing that...
-TODO: Perform evaluation :P.
+coordination. We implemented our theoretical findings in a prototype called
+Lucy and found that our decision procedures can be used to efficiently
+determine the invariant-confluence of common invariants including foreign key
+constraints and escrow transactions. We also found that segmented
+invariant-confluence replication can deliver up to an order of magnitude more
+throughput than linearizable replication for low contention workloads and
+comparable throughput for medium to high contention workloads.
+
 % Strongly consistent systems are easy to reason about but are slow. Weakly

### 2018-04-22 doc/paper/sections/conclusion.tex
Tidied front matter + acks.

\section{Conclusion}
TODO

\textbf{Acknowledgments.}
The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan,
and Cristina Teodoropol for fruitful discussion and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.

### 2018-04-22 doc/paper/sections/conclusion.tex
First stab at conclusion.

@@
 \section{Conclusion}
-TODO
+This paper revolved around two major contributions.  First, we developed a
+deeper understanding of invariant-closure and invariant-confluence by looking
+at the two criteria with reachability in mind. We found that invariant-closure
+fails to incorporate a notion of reachability, and using this intuition, we
+developed conditions under which invariant-closure and invariant-confluence are
+equivalent. Second, we proposed a new consistency model and generalization of
+invariant-confluence, segmented invariant-confluence, that can be used to
+replicate non-invariant-confluent objects with a small amount of coordination
+while still preserving their invariants. We also found that segmented
+invariant-confluence naturally subsumes existing techniques for maintaining
+invariants of replicated objects (e.g.\ locking and escrow transactions).
 

### 2018-04-22 doc/paper/sections/related_work.tex
Shrunk related work.

@@
 \section{Related Work}
-We now discuss RedBlue consistency~\cite{li2012making, li2014automating}, the
-homeostasis protocol~\cite{roy2015homeostasis}, explicit
-consistency~\cite{balegas2015putting}, and token-based
-invariant-confluence~\cite{gotsman2016cause}: four very relevant pieces of
-work. We also briefly survey other related research in the area of
-invariant-confluence including program analysis in database systems, CRDTs, the
-CALM theorem, and interactive proof assistants.
+% Other possibilities:
+%   - Towards Fast Invariant Preservation in Geo-replicated Systems
+%   - Feral Concurrency Control
+%   - The CISE Tool: Proving Weakly-Consistent Applications Correct
+%   - Declarative Programming over Eventually Consistent Data Stores
+%   - Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
 
-\newcommand{\sieve}{\textsc{sieve}}
+\newcommand{\sieve}{\textsc{SIEVE}}
 \textbf{RedBlue Consistency and \sieve.}
-RedBlue consistency is a form of consistency that sits between causal
-consistency and linearizability~\cite{li2012making}. With RedBlue consistency,
-every operation is labelled as red or blue. All operations are executed with
-causal consistency, but with the added restrictions that red operations are
-executed in a single total order embedded within the causal ordering.
-RedBlue consistency is similar to segmented invariant-confluence with a single
-segment in which all red operations are removed from the segment's set of
-allowable transactions. In~\cite{li2012making}, Li et al.\ define
-invariant safety as a sufficient condition for RedBlue consistent objects to be
+RedBlue consistency is a consistency model that sits between causal consistency
+and linearizability~\cite{li2012making}.
+%
+With RedBlue consistency, every operation is manually labelled as either red or
+blue. All operations are executed with causal consistency, but with the added
+restrictions that red operations are executed in a single total order embedded
+within the causal ordering.
+%
+Invariant safety is a sufficient condition for RedBlue consistent objects to be
 invariant-confluent. Invariant safety is an analog of invariant-closure for an
-operation-based (as opposed-to state based~\cite{shapiro2011conflict}) system
-model. In~\cite{li2014automating}, Li et al.\ develop techniques to identify
-which operations are invariant safe.  Like invariant-closure, invariant safety
-is a sufficient but not necessary condition for invariant-confluence. Li et
-al.\ develop sophisticated techniques for deciding invariant safety that
-involve calculating weakest preconditions.  These techniques are complementary
-to our work and can be used to improve the invariant-closure subroutine used by
-our decision procedures.
+operation-based system model (as opposed-to the state based system model that
+we use). In~\cite{li2014automating}, Li et al.\ develop techniques to identify
+invariant safe operations. Like invariant-closure, invariant safety is a
+sufficient but not necessary condition for invariant-confluence. Li et al.\
+develop sophisticated techniques for deciding invariant safety that involve
+calculating weakest preconditions. These techniques are complementary to our
+work and can be used to improve the invariant-closure subroutine used by our
+decision procedures.
 
@@
 The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
-avoid unnecessary coordination between servers in a sharded database. The
+avoid unnecessary coordination between servers in a \emph{sharded} database
+(whereas invariant-confluence targets \emph{replicated} databases). The
 protocol generates global invariants, called global treaties, and servers can
 operate without coordination so long as they preserve the treaty. When the
-treaty is violated, the nodes perform a round of coordination. The protocol
-guarantees that transactions are executed with observational equivalence with
-respect to some serial execution of the transactions. This means that
-intermediate states may be inconsistent, but externally observable side effects
-and the final database state are consistent. The observational equivalence
-guaranteed by the homeostasis is stronger than the guarantees of
-invariant-confluence. As a result, there are invariants and transaction
-workloads which the homeostasis protocol would execute with more coordination
+treaty is violated, the nodes perform a round of coordination and establish a
+new treaty. The protocol guarantees that transactions are executed with
+observational equivalence with respect to some serial execution of the
+transactions. This means that intermediate states may be inconsistent, but
+externally observable side effects and the final database state are consistent.
+The observational equivalence guaranteed by the homeostasis is stronger than
+the guarantees of invariant-confluence. As a result, there are invariants and
+workloads that the homeostasis protocol would execute with more coordination
 than a segmented invariant-confluent execution. Still, the homeostasis
@@
 \textbf{Explicit Consistency.}
-A system provides explicit consistency with respect to a partial ordering of
-transactions if every server executes the transactions in a total order that
-respects the partial order and in a way such that the invariant is always
-preserved~\cite{balegas2015towards}. Thus, explicit consistency is like a
-combination of invariant-confluence and causal consistency, similar to RedBlue
-consistency. To determine if a workload is amenable to explicitly consistent
-replication, Balegas et al. determine if all pairs of transactions can be
-concurrently executed on the same start state without violating the invariant~\cite{balegas2015towards}.
-This is a sufficient condition for explicit consistency similar to criterion
-(3) in \thmref{LatticeProperty}. Balegas et al. also describe a variety of
+Explicit consistency~\cite{balegas2015towards} is a consistency model that
+combines invariant-confluence and causal consistency, similar to RedBlue
+consistency with invariant safety. To determine if a workload is amenable to
+explicitly consistent replication, Balegas et al.\ determine if all pairs of
+transactions can be concurrently executed on the same start state without
+violating the invariant~\cite{balegas2015towards}.  This is a sufficient
+condition for explicit consistency similar to criterion (3) in
+\thmref{LatticeProperty}. Balegas et al.\ also describe a variety of
 techniques---like conflict resolution, locking, and escrow
 transactions~\cite{o1986escrow}---that can be used to replicate workloads that
-do meet their sufficient conditions. Conflict resolution involves choosing an
-appropriate merge function; locking can be implemented as a single segment
-segmentation with certain transactions disallowed; and escrow transactions can
-be emulated with a segmentation that allows for concurrent decrements only when
-value being decremented is sufficiently large.
+do not meet their sufficient conditions. As discussed in \secref{Evaluation},
+simple forms of these techniques can be modelled with segmented
+invariant-confluence.
 
@@
 In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
-consistency model that generalizes RedBlue consistency. An application designer
-defines a set of tokens and specifies which pairs of tokens conflict. Then,
-transactions acquire some subset of the tokens. This allows the application
-designer to specify which transactions conflict at a fine granularity. With no
-tokens, the model is equivalent to causal consistency. With all transactions
-acquiring a single self-conflicting token, the model is equivalent to
-sequential consistency~\cite{lamport1979make}. With some transactions acquiring
-a single self-conflicting token and some transactions acquiring no tokens, the
-model is equivalent to RedBlue consistency. Gotsman et al.\ develop sufficient
-conditions to determine whether a given token scheme is sufficient to guarantee
-that a global invariant is never broken. This model of consistency is very
-similar to segmented invariant-confluence. The token-based approach allows
-users to specify certain conflicts that are not possible with segmented
-invariant-confluence. A segmentation only allows transactions within a segment
-to acquire a single self-conflicting lock. However, segmented
-invariant-confluence also introduces the notion of segmenting the invariant,
-which cannot be emulated with the token based approach. For example,
-implementing \exampleref{SegmentedZ2} with the token-based approach is
-challenging without being able to segment the top-left quadrant from the
-bottom-right quadrant. An interesting direction for future work is to try and
-integrate the token-based approach into segmented invariant-confluence.
-
-% Towards Fast Invariant Preservation in Geo-replicated Systems
+consistency model that generalizes a family of consistency models including
+causal consistency, sequential consistency, and RedBlue consistency. An
+application designer defines a set of tokens and specifies which pairs of
+tokens conflict, and transactions acquire some subset of the tokens when they
+execute. This allows the application designer to specify which transactions
+conflict with each other.
 %
-% Feral
-% The CISE Tool: Proving Weakly-Consistent Applications Correct
-% Declarative Programming over Eventually Consistent Data Stores
-% Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
+% With no tokens, the model is equivalent to causal consistency. With all
+% transactions acquiring a single self-conflicting token, the model is equivalent
+% to sequential consistency~\cite{lamport1979make}. With some transactions
+% acquiring a single self-conflicting token and some transactions acquiring no
+% tokens, the model is equivalent to RedBlue consistency.
+%
+Gotsman et al.\ develop sufficient conditions to determine whether a given
+token scheme is sufficient to guarantee that a global invariant is never
+broken. This model of consistency is very similar to segmented
+invariant-confluence. The token-based approach allows users to specify certain
+conflicts that are not possible with segmented invariant-confluence. A
+segmentation only allows transactions within a segment to acquire a single
+self-conflicting lock. However, segmented invariant-confluence also introduces
+the notion of invariant segmentation, which cannot be emulated with the token
+based approach. For example, it is difficult to emulate escrow transactions
+with the token-based approach.
 
@@
 invariant-confluent object does not necessarily have to be eventually
-consistent. In fact, equipping a distributed object with the trivial merge
-function $x \join y = x$ makes the object invariant-confluent with respect
-any set of transactions and any invariant, but it makes the object very
-inconsistent. Thus, it's natural (though not necessary) to use CRDTs as
+consistent, so it is natural (though not necessary) to use CRDTs as
 invariant-confluent distributed objects, achieving a combination of eventual
@@
 a program can be written in the monotone fragment of these languages, then
-there exists a coordination-free implementation of the program. Moreover,
-monotone programs are invariant to message reordering, whereas non-monotone
-programs may experience anomalies if messages are reordered in the network. The
-CALM theorem provides guarantees about the outputs of a program, but it does
-not guarantee that invariants are maintained throughout the duration of an
+there exists a coordination-free implementation of the program.
+% Moreover, monotone programs are invariant to message reordering, whereas
+% non-monotone programs may experience anomalies if messages are reordered in the
+% network.
+The CALM theorem provides guarantees about the outputs of a program, but it
+does not guarantee that invariants are maintained throughout the duration of an
 execution. Moreover, Bloom and Dedalus are general purpose programming
 languages that can be used to implement a variety of distributed systems (e.g.
-a sharded database), while invariant-confluence is defined strictly in the
-context of replicated data.
+a sharded database) which are outside of the scope of invariant-confluence.
 
@@
 on invariant-confluence continues the theme of using program analysis to reap
-the performance benefits of understanding program semantics that would
-otherwise be left on the table.
+the performance benefits gained from understanding program semantics.
 

### 2018-04-22 doc/paper/sections/evaluation.tex
Finished up a TODO in the eval section.
Shrunk eval.

@@
 resumes normal processing. After every 100 transactions processed, a server
-sends a merge request to a randomly selected server. Merge requests are tagged
-with a monotonically increasing epoch number that is incremented by the master
-after every round of global coordination. This allows servers to discard merge
-requests from previous epochs.
+sends a merge request to a randomly selected server.
+% Merge requests are tagged with a monotonically increasing epoch number that is
+% incremented by the master after every round of global coordination. This allows
+% servers to discard merge requests from previous epochs.
 
@@
 request to a predetermined leader. The leader relays the transaction request to
-all other servers, and when it receives replies from them, it executes the
-transaction and replies to the client. This communication pattern mimics the
-``normal operation'' of state machine replication protocols like
-Paxos~\cite{lamport1998part} and Viewstamped
-Replicated~\cite{liskov2012viewstamped}. In
-\secref{SegmentedInvariantConfluenceEval}, we compare the performance of
-replicating with segmented invariant-confluence against the performance of
-replicating with eventual consistency and linearizability.
+all other servers, and when the leader receives replies from them, it executes
+the transaction and replies to the client. This communication pattern mimics
+the ``normal operation'' of state machine replication protocols
+\cite{lamport1998part, liskov2012viewstamped}.
+% In \secref{SegmentedInvariantConfluenceEval}, we compare the performance of
+% replicating with segmented invariant-confluence against the performance of
+% replicating with eventual consistency and linearizability.
 
@@
 straightforward to add fault-tolerance to Lucy, but it would not affect our
-discussions or evaluation, so we leave it for future work. Moreover, users
-currently have to specify their workloads in Python (for the decision
-procedures) and C++ (for the runtime). In the future, we plan on removing this
-redundancy.
+discussions or evaluation, so we leave it for future work.
+% Moreover, users currently have to specify their workloads in Python (for the
+% decision procedures) and C++ (for the runtime). In the future, we plan on
+% removing this redundancy.
 
@@
 prototypes. We begin by demonstrating the decision procedure on a handful of
-simple, yet practical examples. We then, discuss TODO
+simple, yet practical examples. We then discuss how our tool can be used to
+analyze the TPC-C benchmark.
 
-% \todo{Bank account with and without decrements.}
-% \todo{Auction example.}
-% \todo{Foreign keys with and without deletions.}
-% \subsection{TPC and Feral}
-% \todo{See which of the Feral and TPC transactions we can determine automatically}
 \example[$\ints^2$]\examplelabel{TwoIntsEval}
@@
 transactions, and invariant is given in \figref{Z2Code}. When we call
-\python{checker.check()}, the interactive decision procedure produces
-a counterexample in less than a tenth of a second.  After we label the
+\python{checker.check()}, the interactive decision procedure produces a
+counterexample in less than a tenth of a second.  After we label the
 counterexample and refine the invariant with $y \leq 0$, the interactive
@@
 initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and
-transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
-After less than a tenth of a second, the decision procedure produced a
+with transactions that allow arbitrary insertions and deletions into $X$ and
+$Y$. After less than a tenth of a second, the decision procedure produced a
 reachable counterexample witnessing the fact that the object is not
@@
 reachable counterexample witnessing the fact that the object is not
-invariant-confluent.  If we concurrently close the auction and place an large
+invariant-confluent.  If we concurrently close the auction and place a large
 bid, then we can end up in a state in which the auction is closed, but there is
-a higher bid in $B$.
+a bid in $B$ larger than $w$.
 
 We then segmented our object as follows. The first segment $(\setst{(B, w)}{w =
-\bot}, \setst{t_b}{b \in \ints})$ allows any bidding transaction so long as the
-bid is open. The second segment $(\setst{B, w}{I \cap w \neq \bot}, \emptyset)$
-includes all auctions which have already been closed and forbids all
-transactions. This segmentation captures the intuition that bids should be
-permitted only when the auction is open. We ran our segmented
-invariant-confluence decision procedure on this example, and it was able to
-deduce without any human interaction that the example was segmented
-invariant-confluent in less than a tenth of a second.
+\bot}, \setst{t_b}{b \in \ints})$ allows bidding so long as the bid is open.
+The second segment $(\setst{B, w}{w \neq \bot} \cap I, \emptyset)$ includes all
+auctions which have already been closed and forbids all transactions. This
+segmentation captures the intuition that bids should be permitted only when the
+auction is open. We ran our segmented invariant-confluence decision procedure
+on this example, and it was able to deduce without any human interaction that
+the example was segmented invariant-confluent in less than a tenth of a second.
 
@@
 the value so long as it is not decremented by more than the amount that has
-been escrowed. We now show how segmented invariant-confluence can be used to
+been escrowed. We show how segmented invariant-confluence can be used to
 implement escrow transactions.
@@
   \item
-    Three of the twelve invariants are foreign key constraints.  As discussed
-    in \exampleref{ForeignKeysEval}, our decision procedures can automatically
-    verify conditions under which foreign key constraints are
+    Three of the twelve invariants are \textbf{foreign key constraints}.  As
+    discussed in \exampleref{ForeignKeysEval}, our decision procedures can
+    automatically verify conditions under which foreign key constraints are
     invariant-confluent.
@@
     \newcommand{\ttt}[1]{{\smaller \texttt{#1}}}
-    Seven of the twelve invariants involve maintaining arithmetic relationships
-    between relations. Our decision procedures can correctly identify these as
-    invariant-confluent. Consider, for example, invariant 1 which dictates that
-    a warehouse's year to date balance \ttt{W\_YTD} is equal to the sum of the
-    district year to date balances \ttt{D\_YTD} of the twenty districts that
-    are associated with the warehouse. The Payment transaction randomly selects
-    a district and increments \ttt{W\_YTD} and \ttt{D\_YTD} by a randomly
-    generated amount. We model this workload with a PN-Counter for \ttt{W\_YTD}
-    and twenty PN-Counters for the twenty instances of \ttt{D\_YTD}. We applied
-    Lucy to this workload, and it determined that the workload was
-    invariant-confluent in less than a second.
+    Seven of the twelve invariants involve \textbf{maintaining arithmetic
+    relationships between relations}. Our decision procedures can correctly
+    identify these as invariant-confluent. Consider, for example, invariant 1
+    which dictates that a warehouse's year to date balance \ttt{W\_YTD} is
+    equal to the sum of the district year to date balances \ttt{D\_YTD} of the
+    twenty districts that are associated with the warehouse. The Payment
+    transaction randomly selects a district and increments \ttt{W\_YTD} and
+    \ttt{D\_YTD} by a randomly generated amount. We model this workload with a
+    PN-Counter for \ttt{W\_YTD} and twenty PN-Counters for the twenty instances
+    of \ttt{D\_YTD}. We applied Lucy to this workload, and it determined that
+    the workload was invariant-confluent in less than a second.
 
   \item
-    Two of the twelve invariants involve generating sequential and unique
-    identifiers. This workload is \emph{not} invariant-confluent, but when the
-    sequentiality requirement is dropped, it is invariant-confluent.
+    Two of the twelve invariants involve generating \textbf{sequential and
+    unique identifiers}. This workload is \emph{not} invariant-confluent, but
+    when the sequentiality requirement is dropped, it is invariant-confluent.
     Unfortunately, modelling unique key generation in our theoretical framework
@@
 invariant-confluence as compared to replicating it with eventual consistency or
-linearizability. We begin with two benchmarks that demonstrate the same
-concept: the performance of segmented invariant-confluent replication varies
-with the amount of global coordination induced by either (a) performing a
-transaction that is disallowed within a segment or (b) transitioning between
-segments.
+linearizability.
+% We begin with two benchmarks that demonstrate the same concept: the performance
+% of segmented invariant-confluent replication varies with the amount of global
+% coordination induced by either (a) performing a transaction that is disallowed
+% within a segment or (b) transitioning between segments.
 
@@
 unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
-per second respectively. Expectedly, eventually consistent replication
-significantly outperforms linearizable replication because (a) transactions can
-be sent to any server (not just the leader) and (b) servers do not coordinate
-with each other at all.
+per second respectively.
+%
+% Expectedly, eventually consistent replication significantly outperforms
+% linearizable replication because (a) transactions can be sent to any server
+% (not just the leader) and (b) servers do not coordinate with each other at all.
 %
@@
 workloads and performs increasingly poorly as we increase the fraction of
-withdrawal transactions, eventually becoming slower that linearizable
+withdrawal transactions, eventually performing worse than linearizable
 replication. For example, with 5\% withdrawal transactions, segmented
@@
   Eventually consistent replication scales perfectly with the number of nodes,
-  confirming the results in~\cite{bailis2014coordination}. With eventually
-  consistent replication, servers do not coordinate at all, so they are
-  completely unaffected by the number of servers. Linearizable replication, on
-  the other hand, scales up to about 3-5 servers before performance begins to
-  decrease. These numbers are consistent with typical deployments of
-  state-machine replication protocols like Paxos~\cite{chandra2007paxos}.
-  Because all messages are sent to the leader, the leader becomes the
-  bottleneck as the number of servers and clients increases. Moreover, the
-  leader must wait for responses from more servers, increasing the latency of
-  the slowest response which in turn decreases throughput.
+  confirming the results in~\cite{bailis2014coordination}.
+  %
+  % With eventually consistent replication, servers do not coordinate at all, so
+  % they are completely unaffected by the number of servers.
+  %
+  Linearizable replication, on the other hand, scales up to about 3-5 servers
+  before performance begins to decrease. These numbers are consistent with
+  typical deployments of state-machine replication protocols like
+  Paxos~\cite{chandra2007paxos}.
+  %
+  % Because all messages are sent to the leader, the leader becomes the
+  % bottleneck as the number of servers and clients increases. Moreover, the
+  % leader must wait for responses from more servers, increasing the latency of
+  % the slowest response which in turn decreases throughput.
+  %
   Segmented invariant-confluent replication scales up to about 6-8 servers

### 2018-04-22 benchmark/plot.py
Fixed marker bug in plots.

@@ -10,16 +10,19 @@ EVENTUAL_FMT = {
     'label': 'eventual',
     'marker': 'o',
     'linewidth': 2,
+    'markersize': 8,
 }
 IC_FMT = {
     'label': 'segmented',
-    'marker': 'o',
+    'marker': 's',
     'linewidth': 2,
+    'markersize': 8,
 }
 LIN_FMT = {
     'label': 'linearizable',
     'marker': '^',
     'linewidth': 2,
+    'markersize': 8,
 }
 
 def load_csv(filename):

### 2018-04-22 doc/paper/sections/abstract.tex
Formatting tweaks.

--- 2018-04-01 doc/paper/sections/abstract.tex
@@
 limitations in availability and performance. Weakly consistent systems can be
-implemented with very high performance, but place a burden on the application
+implemented with very high performance but place a burden on the application
 developer to reason about complex interleavings of execution.
-Invariant-confluence provides a formal framework for understanding when we can
-get the best of both worlds. An invariant-confluent object can be efficiently
-replicated with no coordination needed to preserve its invariants.  However,
-actually determining whether or not an object is invariant-confluent is
-challenging (undecidable). In this paper, we establish conditions under which a
-commonly used sufficient condition for invariant confluence is both necessary
-and sufficient. We use this condition to design a general-purpose interactive
-invariant-confluence decision procedure and a novel sufficient condition which
-can be checked automatically. We then take a step beyond invariant-confluence
-and introduce a generalization of invariant-confluence, called segmented
-invariant-confluence, which allows us to replicate non-invariant-confluent
-objects with a small and tunable amount of coordination. Implementing our
-theoretical findings in a system called Lucy, we evaluate our methods on a
-number of case studies and benchmarks showing that... TODO: Perform evaluation
-:P.
-% Strongly consistent systems are easy to reason about but are slow.
-Weakly consistent systems are fast but hard to reason about.
-Invariant-confluence gives a formal framework to reason about when data can be replicated without any coordination needed to maintain invraints, the best of borth worlds.
-However, determining whether an object is invariant-confluent is hard.
-In this paper, we establish conditions under which a commonly used sufficient condition for invariant confluence is both necessary and usfficient.
-We use this condition to design a general-purpose interactive decision procedure and a novel sufficient condition which can be checked automatically.
-We then take a step beyond invariant-confluence and introduce a generalization of invariant-confuluence, called segmented invariant-confluence which allows to replicate objects with a minimal amount of coordination that are otherwise not invariant-confluent.
-We implement all this in a system called lucy and look at a bunch of stuff.
+\Invariantconfluence{} provides a formal framework for understanding when we
+can get the best of both worlds. An \invariantconfluent{} object can be
+efficiently replicated with no coordination needed to preserve its invariants.
+However, actually determining whether or not an object is \invariantconfluent{}
+is challenging (undecidable). In this paper, we establish conditions under
+which a commonly used sufficient condition for \invariantconfluence{} is both
+necessary and sufficient and use this condition to design (a) a general-purpose
+interactive \invariantconfluence{} decision procedure and (b) a novel
+sufficient condition which can be checked automatically. We then take a step
+beyond \invariantconfluence{} and introduce a generalization of
+\invariantconfluence{} called segmented \invariantconfluence{}, which allows us
+to replicate non-\invariantconfluent{} objects with a small and tunable amount
+of coordination. We implemented our theoretical findings in a prototype called
+Lucy and found that our decision procedures efficiently handle workloads
+including foreign key constraints, rollups, escrow transactions, etc.  We also
+found that segmented \invariantconfluent{} replication can deliver up to an
+order of magnitude more throughput than linearizable replication for low
+contention workloads and comparable throughput for medium to high contention
+workloads.
 \end{abstract}

### 2018-04-22 doc/paper/sections/conclusion.tex
Formatting tweaks.

--- 2018-04-22 doc/paper/sections/conclusion.tex
@@
 \section{Conclusion}
-TODO
+This paper revolved around two major contributions.  First, we developed a
+deeper understanding of invariant-closure and \invariantconfluence{} by looking at the
+two criteria with reachability in mind. We found that invariant-closure fails
+to incorporate a notion of reachability, and using this intuition, we developed
+conditions under which invariant-closure and \invariantconfluence{} are equivalent.
+Second, we proposed a new consistency model and generalization of
+\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to replicate
+non-invariant-confluent objects with a small amount of coordination while still
+preserving their invariants. We also found that segmented \invariantconfluence{}
+naturally subsumes existing techniques for maintaining invariants of replicated
+objects (e.g.\ locking and escrow transactions).
 

### 2018-04-22 doc/paper/sections/evaluation.tex
Formatting tweaks.

@@
 %   \item
-%     How practical is the interactive invariant-confluence decision procedure?
+%     How practical is the interactive \invariantconfluence{} decision procedure?
 %     Can we use it to classify real-world transactions and invariants?
 %   \item
-%     How practical is segmented invariant-confluence? Are real-world workloads
+%     How practical is segmented \invariantconfluence{}? Are real-world workloads
 %     amenable to segmentation?
 %   \item
-%     How efficient is the interactive invariant-confluence decision procedure?
+%     How efficient is the interactive \invariantconfluence{} decision procedure?
 %   \item
-%     How efficiently can we replicate a segmented invariant-confluent object as
+%     How efficiently can we replicate a segmented \invariantconfluent{} object as
 %     compared to alternative approaches like replicating with weak or strong
@@
 %   \item
-%     How does the performance of replicating a segmented invariant-confluence
+%     How does the performance of replicating a segmented \invariantconfluence{}
 %     object vary as we vary the workload, segmentation, and replication factor?
@@
 interactive decision procedures using an interactive Python console. We use
-Z3~\cite{de2008z3} to implement our invariant-closure decision procedure,
+Z3~\cite{de2008z3} to implement our \invariantclosure{} decision procedure,
 compiling an object and invariant into a formula that is satisfiable if and
-only if the object is \emph{not} invariant-closed. If the object is
-invariant-closed, then Z3 concludes that the formula is unsatisfiable.
-Otherwise, if the object is not invariant-closed, then Z3 produces a
+only if the object is \emph{not} \invariantclosed{}. If the object is
+\invariantclosed{}, then Z3 concludes that the formula is unsatisfiable.
+Otherwise, if the object is not \invariantclosed{}, then Z3 produces a
 counterexample witnessing the satisfiability of the formula.
 
-Lucy also includes an implementation of the invariant-confluence and
+Lucy also includes an implementation of the \invariantconfluence{} and
 segmented-invariant confluence system models in roughly 3,500 lines of C++.
 Users specify objects, transactions, invariants, and segmentations in C++. Lucy
-then replicates the objects using segmented invariant-confluence (or
-invariant-confluence if the segmentation contains a single segment without any
+then replicates the objects using segmented \invariantconfluence{} (or
+\invariantconfluence{} if the segmentation contains a single segment without any
 disallowed transactions). Clients send every transaction request to a randomly
@@
 % In \secref{SegmentedInvariantConfluenceEval}, we compare the performance of
-% replicating with segmented invariant-confluence against the performance of
+% replicating with segmented \invariantconfluence{} against the performance of
 % replicating with eventual consistency and linearizability.
@@
 Because fault-tolerance is largely an orthogonal concern to
-invariant-confluence, Lucy is implemented without fault-tolerance. It would be
+\invariantconfluence{}, Lucy is implemented without fault-tolerance. It would be
 straightforward to add fault-tolerance to Lucy, but it would not affect our
@@
 counterexample and refine the invariant with $y \leq 0$, the interactive
-decision procedure determines that the object is invariant-confluent, again, in
-less than a tenth of a second. Note that the object is invariant-confluent but
-\emph{not} invariant-closed, so prior work that relies on invariant-closure
-alone to determine invariant-confluence would not be able to identify this
-example as invariant-confluent.
+decision procedure determines that the object is \invariantconfluent{}, again, in
+less than a tenth of a second. Note that the object is \invariantconfluent{} but
+\emph{not} \invariantclosed{}, so prior work that relies on \invariantclosure{}
+alone to determine \invariantconfluence{} would not be able to identify this
+example as \invariantconfluent{}.
 
@@
 reachable counterexample witnessing the fact that the object is not
-invariant-confluent. A concurrent insertion into $X$ and deletion from $Y$ can
+\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$ can
 lead to a state which violates the invariant. This object is not
-invariant-confluent and therefore not invariant-closed. Thus, previous tools
-depending on invariant-closure as a sufficient condition would be unable to
-conclude definitively that the object is \emph{not} invariant-confluent.
+\invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous tools
+depending on \invariantclosure{} as a sufficient condition would be unable to
+conclude definitively that the object is \emph{not} \invariantconfluent{}.
 
@@
 decision procedure correctly deduced that the object is now
-invariant-confluent. These results were manually proven
+\invariantconfluent{}. These results were manually proven
 in~\cite{bailis2014coordination}, but our tool was able to confirm them
@@
 reachable counterexample witnessing the fact that the object is not
-invariant-confluent.  If we concurrently close the auction and place a large
+\invariantconfluent{}.  If we concurrently close the auction and place a large
 bid, then we can end up in a state in which the auction is closed, but there is
@@
 segmentation captures the intuition that bids should be permitted only when the
-auction is open. We ran our segmented invariant-confluence decision procedure
+auction is open. We ran our segmented \invariantconfluence{} decision procedure
 on this example, and it was able to deduce without any human interaction that
-the example was segmented invariant-confluent in less than a tenth of a second.
+the example was segmented \invariantconfluent{} in less than a tenth of a second.
 
@@
 the value so long as it is not decremented by more than the amount that has
-been escrowed. We show how segmented invariant-confluence can be used to
+been escrowed. We show how segmented \invariantconfluence{} can be used to
 implement escrow transactions.
@@
 \exampleref{CounreachableExample}, we found that concurrent decrements violate
-invariant-confluence which led us to a segmentation which prohibited concurrent
+\invariantconfluence{} which led us to a segmentation which prohibited concurrent
 decrements. We now propose a new segmentation with escrow amount $k$ that will
@@
 our decision procedure on this example and it concluded that it was segmented
-invariant-confluent in less than a tenth of a second and without any human
+\invariantconfluent{} in less than a tenth of a second and without any human
 interaction.
@@
     automatically verify conditions under which foreign key constraints are
-    invariant-confluent.
+    \invariantconfluent{}.
 
@@
     relationships between relations}. Our decision procedures can correctly
-    identify these as invariant-confluent. Consider, for example, invariant 1
+    identify these as \invariantconfluent{}. Consider, for example, invariant 1
     which dictates that a warehouse's year to date balance \ttt{W\_YTD} is
@@
     of \ttt{D\_YTD}. We applied Lucy to this workload, and it determined that
-    the workload was invariant-confluent in less than a second.
+    the workload was \invariantconfluent{} in less than a second.
 
@@
     Two of the twelve invariants involve generating \textbf{sequential and
-    unique identifiers}. This workload is \emph{not} invariant-confluent, but
-    when the sequentiality requirement is dropped, it is invariant-confluent.
+    unique identifiers}. This workload is \emph{not} \invariantconfluent{}, but
+    when the sequentiality requirement is dropped, it is \invariantconfluent{}.
     Unfortunately, modelling unique key generation in our theoretical framework
@@
 Now, we evaluate the performance of replicating an object with segmented
-invariant-confluence as compared to replicating it with eventual consistency or
+\invariantconfluence{} as compared to replicating it with eventual consistency or
 linearizability.
 % We begin with two benchmarks that demonstrate the same concept: the performance
-% of segmented invariant-confluent replication varies with the amount of global
+% of segmented \invariantconfluent{} replication varies with the amount of global
 % coordination induced by either (a) performing a transaction that is disallowed
@@
   \caption{%
-    Segmented invariant-confluent replication throughput versus coordination,
+    Segmented \invariantconfluent{} replication throughput versus coordination,
     induced by executing disallowed transactions (top) and by transitioning
@@
 transactions. We replicate the object with eventual consistency, segmented
-invariant-confluence, and linearizability and measure the system's total
+\invariantconfluence{}, and linearizability and measure the system's total
 throughput as we vary the fraction of client requests that are withdrawals. The
@@
 %
-Segmented invariant-confluent replication performs well for low-withdrawal
+Segmented \invariantconfluent{} replication performs well for low-withdrawal
 workloads and performs increasingly poorly as we increase the fraction of
@@
 replication. For example, with 5\% withdrawal transactions, segmented
-invariant-confluent replication performs an order of magnitude better than
+\invariantconfluent{} replication performs an order of magnitude better than
 linearizable replication; with 50\% withdrawals, it performs as well; and with
@@
 %
-First, for low-withdrawal workloads, segmented invariant-confluent replication
+First, for low-withdrawal workloads, segmented \invariantconfluent{} replication
 achieves a compromise between strong and weak consistency. It guarantees that
 invariants are maintained (which is impossible with eventual consistency if the
-object is not invariant-confluent) with performance many times better than
+object is not \invariantconfluent{}) with performance many times better than
 strongly consistent replication.
 %
-Second, segmented invariant-confluent replication is poorly suited to workloads
+Second, segmented \invariantconfluent{} replication is poorly suited to workloads
 that require a large amount of coordination. For workloads without much inherit
@@
   squares, and so on. We measure the throughput of the object replicated with
-  eventually consistent, segmented invariant-confluent, and linearizable
+  eventually consistent, segmented \invariantconfluent{}, and linearizable
   replication as we vary the segment side length $n$. The results are shown in
@@
   consistency outperforms linearizability by roughly two orders of magnitude.
-  In this example, the segmented invariant-confluent replication only requires
+  In this example, the segmented \invariantconfluent{} replication only requires
   coordination when transitioning between segment boundaries, so as we increase
@@
   \caption{%
-    Throughput of eventually consistent, segmented invariant-confluent, and
+    Throughput of eventually consistent, segmented \invariantconfluent{}, and
     linearizable replication measured against the number of
@@
 \begin{benchmark}
-  In this benchmark, we measure the scale-out of segmented invariant-confluent
+  In this benchmark, we measure the scale-out of segmented \invariantconfluent{}
   replication. We repeat \benchref{VaryWithdraws} with a 10\% withdrawal rate,
@@
   %
-  Segmented invariant-confluent replication scales up to about 6-8 servers
+  Segmented \invariantconfluent{} replication scales up to about 6-8 servers
   before succumbing to the same scalability bottlenecks as linearizable
@@
   These results highlight the importance of coordination avoidance in
-  distributed databases. While segmented invariant-confluent replication scales
+  distributed databases. While segmented \invariantconfluent{} replication scales
   out slightly better than linearizable replication, both scale significantly

### 2018-04-22 doc/paper/sections/interactive_decision_procedure.tex
Finished merge reduction section.
Documented segmented invariant-confluence DP.
Shrunk eval.
Formatting tweaks.

@@
-\section{An Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
+\section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
 \newcommand{\Helper}{\textsf{Helper}}
@@
 \begin{algorithm}[t]
-  \caption{Interactive Invariant-Confluence Decision Procedure}%
+  \caption{Interactive invariant-confluence decision procedure}%
   \algolabel{InteractiveDecisionProcedure}
   \begin{algorithmic}
+    \algocomment{Return if $O$ is \sTIconfluent{}.}
     \Function{IsInvConfluent}{$O$, $s_0$, $T$, $I$}
@@
 
-    \algocomment{$R$ is a set of $\sTIreachable{}$ states.}
-    \algocomment{$NR$ is a set of $\sTIunreachable{}$ states.}
+    \algocomment{$R$ is a set of \sTIreachable{} states.}
+    \algocomment{$NR$ is a set of \sTIunreachable{} states.}
     \algocomment{$I(s_0)$ is a precondition.}
@@
 \IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
-$s_0$, $T$, and $I$---takes as arguments a set $R$ of $\sTIreachable{}$ states
-and a set $NR$ of $\sTIunreachable{}$ states. Like \IsInvConfluent, \Helper$(O,
+$s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
+and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, \Helper$(O,
 s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
@@
 iteratively increased which removes unreachable states from $I$ until $I
-\subseteq \setst{s}{\sTIreachable{s}}$.
+\subseteq \setst{s}{\sTIreachablepredicate{s}}$.
 
@@
 \[
-  \setst{s \in S}{\stireachable{s_0}{T}{I-NR}{s}}
+  \setst{s \in S}{\stireachablepredicate{s_0}{T}{I-NR}{s}}
     \subseteq I - NR
@@
 \]
-If $NR$ only contains $\sTIunreachable{}$ states, then the set of
-$\sTIreachable{}$ states is equal to set of $\stireachable{s_0}{T}{I-NR}{}$
+If $NR$ only contains \sTIunreachable{} states, then the set of
+\sTIreachable{} states is equal to set of \stireachable{s_0}{T}{I-NR}{}
 states which, as we just showed, is a subset of $I$.  Thus, $O$ is
@@
 
-After $s_1$ and $s_2$ have been labelled as $\sTIreachable{}$ or not, we
-continue. If both $s_1$ and $s_2$ are $\sTIreachable{}$, then so is $s_1 \join
+After $s_1$ and $s_2$ have been labelled as \sTIreachable{} or not, we
+continue. If both $s_1$ and $s_2$ are \sTIreachable{}, then so is $s_1 \join
 s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{}, so
 \Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
-$\sTIunreachable{}$, so we recurse.
+\sTIunreachable{}, so we recurse.
 
@@
 guarantees that the algorithm will terminate after one more call to \IsIclosed.
+
+{\input{figures/decision_procedure_example}}
 
@@
 \algoref{InteractiveDecisionProcedure} so that \Helper{} requires that whenever
-a user labels a state $s$ as $\sTIreachable{}$, they must also provide an
-$\sTIreachable{}$ expression $e$ that evaluates to $s$. Here, $e$ acts a proof
+a user labels a state $s$ as \sTIreachable{}, they must also provide an
+\sTIreachable{} expression $e$ that evaluates to $s$. Here, $e$ acts a proof
 that certifies that $s$ is indeed reachable. This increases the burden on the
@@
 states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
-$NR$ is incorrect and contains some $\sTIreachable{}$ states, then some of
+$NR$ is incorrect and contains some \sTIreachable{} states, then some of
 these states are artificially made unreachable, but the system is still
 guaranteed to never produce a state which violates $I$.
-
-{\input{figures/decision_procedure_example}}

### 2018-04-22 doc/paper/sections/introduction.tex
Formatting tweaks.

@@
 still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
-introduced the notion of invariant-confluence as a necessary and sufficient
+introduced the notion of \invariantconfluence{} as a necessary and sufficient
 condition for when invariants can be maintained over replicated data without
@@
 invariant-confluent or not, but ideally we would have a general purpose program
-that can automatically determine invariant-confluence for us. \textbf{The first
+that can automatically determine \invariantconfluence{} for us. \textbf{The first
 main thrust of this paper is to make progress towards a general-purpose
-invariant-confluence decision procedure.}
+\invariantconfluence{} decision procedure.}
 
-Designing a general-purpose invariant-confluence decision procedure is
+Designing a general-purpose \invariantconfluence{} decision procedure is
 difficult\footnote{Actually, it's impossible. Determining whether an object is
@@
 sufficient (but not necessary) condition for
-invariant-confluence~\cite{li2012making, li2014automating}. The existing
+\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
 approaches check to see if an object is invariant-closed. If it is, then they
@@
 underlying reason \emph{why} invariant-closure is not necessary for
-invariant-confluence. We find that invariant-confluence is fundamentally a
+\invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
 property about reachability, about whether certain values of our replicated
@@
 unnecessary condition. Using this understanding, we construct a set of modest
-conditions under which invariant-closure and invariant-confluence are in fact
+conditions under which invariant-closure and \invariantconfluence{} are in fact
 \emph{equivalent}, allowing us to reduce the problem of determining
-invariant-confluence to that of determining invariant-closure.
+\invariantconfluence{} to that of determining invariant-closure.
 
 Then, we use these conditions to design a general-purpose interactive
-invariant-confluence decision procedure. Users iteratively interact with the
+\invariantconfluence{} decision procedure. Users iteratively interact with the
 decision procedure, eliminating unreachable states from the invariant, moving
-towards the conditions under which invariant-closure and invariant-confluence
+towards the conditions under which invariant-closure and \invariantconfluence{}
 are equivalent. Meanwhile, the decision-procedure generates example reachable
@@
 reachability. We also leverage our intuitions about reachability to develop a
-new sufficient condition for invariant-confluence, dubbed merge-reducibility.
+new sufficient condition for \invariantconfluence{}, dubbed merge-reducibility.
 Merge-reducibility covers some cases that are not covered by invariant-closure,
@@
 \textbf{The second main thrust of this paper is to take a step beyond
-invariant-confluence and develop a generalization of invariant-confluence
-called segmented invariant-confluence.} While invariant-confluence
+  \invariantconfluence{} and develop a generalization of \invariantconfluence{}
+called segmented \invariantconfluence{}.} While \invariantconfluence{}
 characterizes objects which can be replicated \emph{without any} coordination,
-segmented invariant-confluence allows us to replicate non-invariant-confluent
+segmented \invariantconfluence{} allows us to replicate non-invariant-confluent
 objects with \emph{minimal coordination}. The main idea is that
@@
   \item
-    We propose a novel expression-oriented definition of invariant-confluence
+    We propose a novel expression-oriented definition of \invariantconfluence{}
     that is both formal and simple (\secref{InvariantConfluence}).
@@
     We exploit these conditions to design a general-purpose interactive
-    decision procedure for invariant-confluence
+    decision procedure for \invariantconfluence{}
     (\secref{InteractiveDecisionProcedure}).
@@
     We again exploit these conditions to design a novel non-trivial sufficient
-    condition for invariant-confluence, merge-reducibility, that does not
+    condition for \invariantconfluence{}, merge-reducibility, that does not
     depend on user interaction (\secref{MergeReduction}).
@@
   \item
-    We present segmented invariant-confluence: a generalization of
-    invariant-confluence that uses a small amount of coordination to maintain
+    We present segmented \invariantconfluence{}: a generalization of
+    \invariantconfluence{} that uses a small amount of coordination to maintain
     invariants for replicated objects that are otherwise not
@@
     We implement our decision procedures and sufficient conditions for both
-    invariant-confluence and segmented invariant-confluence in a system called
+    \invariantconfluence{} and segmented \invariantconfluence{} in a system called
     Lucy, then use Lucy to evaluate the efficacy or our methods

### 2018-04-22 doc/paper/sections/invariant_closure.tex
Formatting tweaks.

@@
-\section{Invariant-Closure}\seclabel{InvariantClosure}
+\section{Invariant Closure}\seclabel{InvariantClosure}
 Our ultimate goal is to write a program that can automatically decide whether a
@@
 Due to this complexity, existing systems that aim to automatically decide
-invariant-confluence instead focus on deciding a sufficient condition for
-invariant-confluence---dubbed \defword{invariant-closure}---that is simpler to
+\invariantconfluence{} instead focus on deciding a sufficient condition for
+\invariantconfluence{}---dubbed \defword{\invariantclosure{}}---that is simpler to
 reason about~\cite{li2012making, li2014automating}. In this section, we define
-invariant-closure and study why the condition is sufficient but not necessary.
+\invariantclosure{} and study why the condition is sufficient but not necessary.
 Armed with this understanding, we present conditions under which it is both
@@
 
-We say an object $O = (S, \join)$ is \defword{invariant-closed} with respect to
-an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant-satisfying
+We say an object $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to
+an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying
 states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
@@
 
-\thmref{IclosureImpliesIconfluence} states that invariant-closure is
-sufficient for invariant-confluence. Intuitively, our system model ensures that
+\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
+sufficient for \invariantconfluence{}. Intuitively, our system model ensures that
 transaction execution preserves the invariant, so if merging states also
@@
 
-This is good news because checking if an object is invariant-closed is more
-straightforward than checking if it is invariant-confluent. Existing systems
+This is good news because checking if an object is \invariantclosed{} is more
+straightforward than checking if it is \invariantconfluent{}. Existing systems
 typically use an SMT solver like Z3 to check if an object is
-invariant-closed~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
-is, then by \thmref{IclosureImpliesIconfluence}, it is invariant-confluent.
-Unfortunately, invariant-closure is \emph{not} necessary for
-invariant-confluence, so if an object is \emph{not} invariant-closed, these
-systems cannot conclude that the object is \emph{not} invariant-confluent. The
-reason why invariant-closure is not necessary for invariant-confluence is best
+\invariantclosed{}~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
+is, then by \thmref{IclosureImpliesIconfluence}, it is \invariantconfluent{}.
+Unfortunately, \invariantclosure{} is \emph{not} necessary for
+\invariantconfluence{}, so if an object is \emph{not} \invariantclosed{}, these
+systems cannot conclude that the object is \emph{not} \invariantconfluent{}. The
+reason why \invariantclosure{} is not necessary for \invariantconfluence{} is best
 explained through an example.
@@
   of the fourth quadrant. From this, it's immediate that the reachable states
-  are a subset of the invariant, so $O$ is invariant-confluent. However,
+  are a subset of the invariant, so $O$ is \invariantconfluent{}. However,
   letting $s_1 = (-1, 1)$ and $s_2 = (1, -1)$, we see that $O$ is not
-  invariant-closed. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
+  \invariantclosed{}. $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \join s_2 =
   (1, 1)$, we see $\lnot I(s_3)$.
@@
 In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
-invariant-closed, but $s_1$ is not reachable. This is not particular to
-\exampleref{Z2}. In fact, it is fundamentally the reason why invariant-closure
-is not equivalent to invariant-confluence. Invariant-confluence is, at its
-core, a property of reachable states, but invariant-closure is completely
-ignorant of reachability. As a result, invariant-satisfying yet unreachable
-states like $s_1$ are the key hurdle preventing invariant-closure from being
-equivalent to invariant-confluence. This is formalized by
+\invariantclosed{}, but $s_1$ is not reachable. This is not particular to
+\exampleref{Z2}. In fact, it is fundamentally the reason why \invariantclosure{}
+is not equivalent to \invariantconfluence{}. \Invariantconfluence{} is, at its
+core, a property of reachable states, but \invariantclosure{} is completely
+ignorant of reachability. As a result, invariant satisfying yet unreachable
+states like $s_1$ are the key hurdle preventing \invariantclosure{} from being
+equivalent to \invariantconfluence{}. This is formalized by
 \thmref{IclosureEquivalentIconfluence}
@@
   transactions $T$, and an invariant $I$. If the invariant is a subset of the
-  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachable{s}}$), then
+  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$), then
   \[
@@
 
-% - define invariant-closure and introduce it as a sufficient condition
+% - define \invariantclosure{} and introduce it as a sufficient condition
 % - cite existing work to show that people have used invariant closure to
-%   reason about invariant-confluence
+%   reason about \invariantconfluence{}
 % - explain why invariant closure can fail and why its not sufficient using the

### 2018-04-22 doc/paper/sections/invariant_confluence.tex
Documented segmented invariant-confluence DP.
Finished first cut at related work.
Formatting tweaks.

@@
-\section{Invariant-Confluence}\seclabel{InvariantConfluence}
-Informally, a replicated object is \defword{invariant-confluent} with respect
+\section{Invariant Confluence}\seclabel{InvariantConfluence}
+Informally, a replicated object is \defword{\invariantconfluent{}} with respect
 to an invariant if every replica of the object is guaranteed to satisfy the
@@
 modified or merged together~\cite{bailis2014coordination}. In this section, we
-make this informal notion of invariant-confluence precise. We begin by
+make this informal notion of \invariantconfluence{} precise. We begin by
 introducing our system model of replicated objects and then present a formal
-definition of invariant-confluence.
+definition of \invariantconfluence{}.
 
@@
 
-Informally, $O$ is \defword{invariant-confluent} with respect to $s_0$, $T$, and
+Informally, $O$ is \defword{\invariantconfluent{}} with respect to $s_0$, $T$, and
 $I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots, s_n$
@@
 \subsection{Expression-Based Formalism}
-To define invariant-confluence formally, we represent a state produced by a
+To define \invariantconfluence{} formally, we represent a state produced by a
 system execution as a simple expression generated by the grammar
@@
 
-We say an expression $e$ is \defword{$\sTIreachable{}$} if it corresponds to a
-valid execution of our system model. Formally, we define $\sTIreachable{e}$ to
-be the smallest predicate that satisfies the following equations:
+We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
+valid execution of our system model. Formally, we define
+$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
+following equations:
 \begin{itemize}
   \item
-    $\sTIreachable{s_0}$.
+    $\sTIreachablepredicate{s_0}$.
   \item
     For all expressions $e$ and for all transactions $t$ in the set $T$ of
-    transactions, if $\sTIreachable{e}$ and $I(t(e))$, then
-    $\sTIreachable{t(e)}$.
+    transactions, if $\sTIreachablepredicate{e}$ and $I(t(e))$, then
+    $\sTIreachablepredicate{t(e)}$.
   \item
-    For all expressions $e_1$ and $e_2$, if $\sTIreachable{e_1}$ and
-    $\sTIreachable{e_2}$, then $\sTIreachable{e_1 \join e_2}$.
+    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
+    $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
+    e_2}$.
 \end{itemize}
-Similarly, we say a state $s \in S$ is $\sTIreachable{}$ if there exists a
-$\sTIreachable{}$ expression $e$ that evaluates to $s$.
+Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists a
+\sTIreachable{} expression $e$ that evaluates to $s$.
 
-Finally, we say $O$ is \defword{invariant-confluent} with respect to $s_0$,
+Finally, we say $O$ is \defword{\invariantconfluent} with respect to $s_0$,
 $T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
-satisfy the invariant: $\setst{s \in S}{\sTIreachable{s}} \subseteq I$.
-% TODO: Make a note that this definition is different than the original, and
-% reference the appendix where we include the proof that the two definitions
-% are equivalent.
+satisfy the invariant: $\setst{s \in S}{\sTIreachablepredicate{s}} \subseteq
+I$.

### 2018-04-22 doc/paper/sections/merge_reduction.tex
Formatting tweaks.

@@
 \section{Merge Reduction}\seclabel{MergeReduction}
-In \secref{InvariantClosure}, we discussed how invariant-closure is
-fundamentally a property of reachability and that invariant-closure is
-sufficient but not necessary for invariant-confluence because it fails to
+In \secref{InvariantClosure}, we discussed how \invariantclosure{} is
+fundamentally a property of reachability and that \invariantclosure{} is
+sufficient but not necessary for \invariantconfluence{} because it fails to
 incorporate any notion of reachability. Using this intuition, we established
@@
 advantage of this intuition to develop a new sufficient condition for
-invariant-confluence that can be checked without user interaction and that
-covers some cases not covered by invariant-closure.
+\invariantconfluence{} that can be checked without user interaction and that
+covers some cases not covered by \invariantclosure{}.
 
@@
 abbreviated \defword{\sTImergereducible}, if for every pair of $e_1$ and $e_2$
-of merge-free $\sTIreachable{}$ expressions, there exists some merge-free
-$\sTIreachable{}$ expression $e_3$ that evaluates to the same state as $e_1
+of merge-free \sTIreachable{} expressions, there exists some merge-free
+\sTIreachable{} expression $e_3$ that evaluates to the same state as $e_1
 \join e_2$. Intuitively, if $O$ is merge-reducible, we can replace $e_1 \join
@@
 The proof of \thmref{ReducibilityImpliesIconfluence} is a straightforward
-induction. We begin with an $\sTIreachable{}$ expression $e$ and repeatedly
+induction. We begin with an \sTIreachable{} expression $e$ and repeatedly
 replace any subexpression that merges two merge-free subexpressions with an
@@
 invariant, $e'$ (and hence $e$) is guaranteed to satisfy the invariant. Thus,
-all reachable states satisfy the invariant, so $O$ is invariant-confluent.
+all reachable states satisfy the invariant, so $O$ is \invariantconfluent{}.
 
-Merge-reducibility is a sufficient condition for invariant-confluence, but
-unlike with invariant-closure, it is not straightforward to automatically
+Merge-reducibility is a sufficient condition for \invariantconfluence{}, but
+unlike with \invariantclosure{}, it is not straightforward to automatically
 determine if an object is merge-reducible. In \thmref{LatticeProperty}, we
@@
 
-\thmref{IclosureImpliesIconfluence} states that invariant-closure is a
-sufficient condition for invariant-confluence, and \thmref{LatticeProperty}
+\thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is a
+sufficient condition for \invariantconfluence{}, and \thmref{LatticeProperty}
 states that criteria (1) -- (4) are sufficient conditions for
-invariant-confluence. How do these sufficient conditions relate to one another?
-Clearly, not all invariant-closed objects are semilattices, so
-invariant-closure does not imply criteria (1) -- (4). Conversely, there are some
-objects which satisfy criteria (1) -- (4) which are not invariant-closed. Here's
+\invariantconfluence{}. How do these sufficient conditions relate to one another?
+Clearly, not all \invariantclosed{} objects are semilattices, so
+\invariantclosure{} does not imply criteria (1) -- (4). Conversely, there are some
+objects which satisfy criteria (1) -- (4) which are not \invariantclosed{}. Here's
 one example.
@@
 
-Invariant-closure is not necessary for invariant-confluence because it fails to
+\Invariantclosure{} is not necessary for \invariantconfluence{} because it fails to
 incorporate any notion of reachability. Criteria (1) -- (4) are also
-unnecessary, but they can be used to prove that some non-invariant-closed
-objects are invariant-confluent because the criteria \emph{do} incorporate
+unnecessary, but they can be used to prove that some non-\invariantclosed{}
+objects are \invariantconfluent{} because the criteria \emph{do} incorporate
 notions of reachability. In particular, criterion (3) is a slight variant of
-invariant-closure; it also states that invariant-satisfying states should be
+\invariantclosure{}; it also states that invariant satisfying states should be
 closed under merge. The fundamental difference is that criterion (3) restricts
@@
 a set with both odd and even integers. However, if we begin in an
-invariant-satisfying state, we cannot reach both an odd-only and even-only set.
+invariant satisfying state, we cannot reach both an odd-only and even-only set.
 Criterion (3) is able to recognize this fact and conclude that $O$ is
-invariant-confluent despite it not being invariant-closed.
+\invariantconfluent{} despite it not being \invariantclosed{}.
 

### 2018-04-22 doc/paper/sections/related_work.tex
Formatting tweaks.

@@
 Invariant safety is a sufficient condition for RedBlue consistent objects to be
-invariant-confluent. Invariant safety is an analog of invariant-closure for an
+\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{} for an
 operation-based system model (as opposed-to the state based system model that
 we use). In~\cite{li2014automating}, Li et al.\ develop techniques to identify
-invariant safe operations. Like invariant-closure, invariant safety is a
-sufficient but not necessary condition for invariant-confluence. Li et al.\
+invariant safe operations. Like \invariantclosure{}, invariant safety is a
+sufficient but not necessary condition for \invariantconfluence{}. Li et al.\
 develop sophisticated techniques for deciding invariant safety that involve
 calculating weakest preconditions. These techniques are complementary to our
-work and can be used to improve the invariant-closure subroutine used by our
+work and can be used to improve the \invariantclosure{} subroutine used by our
 decision procedures.
@@
 avoid unnecessary coordination between servers in a \emph{sharded} database
-(whereas invariant-confluence targets \emph{replicated} databases). The
+(whereas \invariantconfluence{} targets \emph{replicated} databases). The
 protocol generates global invariants, called global treaties, and servers can
@@
 The observational equivalence guaranteed by the homeostasis is stronger than
-the guarantees of invariant-confluence. As a result, there are invariants and
+the guarantees of \invariantconfluence{}. As a result, there are invariants and
 workloads that the homeostasis protocol would execute with more coordination
-than a segmented invariant-confluent execution. Still, the homeostasis
+than a segmented \invariantconfluent{} execution. Still, the homeostasis
 protocol's mechanism of establishing invariants and operating without
 coordination so long as the invariants are maintained is very similar to our
-design of segmented invariant-confluence.
+design of segmented \invariantconfluence{}.
 
@@
 Explicit consistency~\cite{balegas2015towards} is a consistency model that
-combines invariant-confluence and causal consistency, similar to RedBlue
+combines \invariantconfluence{} and causal consistency, similar to RedBlue
 consistency with invariant safety. To determine if a workload is amenable to
@@
 simple forms of these techniques can be modelled with segmented
-invariant-confluence.
+\invariantconfluence{}.
 
-\textbf{Token-Based Invariant-Confluence.}
+\textbf{Token Based Invariant Confluence.}
 In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
@@
 broken. This model of consistency is very similar to segmented
-invariant-confluence. The token-based approach allows users to specify certain
-conflicts that are not possible with segmented invariant-confluence. A
+\invariantconfluence{}. The token-based approach allows users to specify certain
+conflicts that are not possible with segmented \invariantconfluence{}. A
 segmentation only allows transactions within a segment to acquire a single
-self-conflicting lock. However, segmented invariant-confluence also introduces
+self-conflicting lock. However, segmented \invariantconfluence{} also introduces
 the notion of invariant segmentation, which cannot be emulated with the token
@@
 the need for any coordination. Our definition of distributed objects and our
-invariant-confluence system model are inspired directly by the corresponding
+\invariantconfluence{} system model are inspired directly by the corresponding
 definitions and system models in~\cite{shapiro2011conflict}. Note that an
-invariant-confluent object does not necessarily have to be eventually
+\invariantconfluent{} object does not necessarily have to be eventually
 consistent, so it is natural (though not necessary) to use CRDTs as
-invariant-confluent distributed objects, achieving a combination of eventual
-consistency and invariant-preservation. Our criteria in
+\invariantconfluent{} distributed objects, achieving a combination of eventual
+consistency and invariant preservation. Our criteria in
 \thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
@@
 languages that can be used to implement a variety of distributed systems (e.g.
-a sharded database) which are outside of the scope of invariant-confluence.
+a sharded database) which are outside of the scope of \invariantconfluence{}.
 
@@
 concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
-on invariant-confluence continues the theme of using program analysis to reap
+on \invariantconfluence{} continues the theme of using program analysis to reap
 the performance benefits gained from understanding program semantics.
@@
 % guides users through the process of interactively generating a set of reachable
-% and unreachable points that witness the invariant-confluence of an object. An
+% and unreachable points that witness the \invariantconfluence{} of an object. An
 % interesting avenue for future work is to integrate an proof assistant into our

### 2018-04-22 doc/paper/sections/segmented_invariant_confluence.tex
Removed TODO.
Formatting tweaks.

@@
-\section{Segmented Invariant-Confluence}\seclabel{SegmentedInvariantConfluence}
+\section{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluence}
 \newcommand{\IsIclosed}{\textsf{IsIclosed}}
 
-If a distributed object is invariant-confluent, then the object can be
+If a distributed object is \invariantconfluent{}, then the object can be
 replicated without the need for any form of coordination to maintain the
-object's invariant. But what if the object is \emph{not} invariant-confluent?
-In this section, we present a generalization of invariant-confluence called
-\defword{segmented invariant-confluence} that can be used to maintain the
-invariants of non-invariant-confluent objects, requiring only a small amount of
+object's invariant. But what if the object is \emph{not} \invariantconfluent{}?
+In this section, we present a generalization of \invariantconfluence{} called
+\defword{segmented \invariantconfluence{}} that can be used to maintain the
+invariants of non-\invariantconfluent{} objects, requiring only a small amount of
 coordination.
 
-The main idea behind segmented invariant-confluence is to segment the state
+The main idea behind segmented \invariantconfluence{} is to segment the state
 space into a number of segments and restrict the set of allowable transactions
-within each segment in such a way that the object is invariant-confluent
+within each segment in such a way that the object is \invariantconfluent{}
 \emph{within each segment} (even though it may not be globally
-invariant-confluent). Then, servers can run coordination-free within a segment
+\invariantconfluent{}). Then, servers can run coordination-free within a segment
 and need only coordinate when transitioning from one segment to another. We now
-formalize segmented invariant-confluence, describe the system model we use to
-replicate segmented invariant-confluent objects, and introduce a segmented
-invariant-confluence decision procedure.
+formalize segmented \invariantconfluence{}, describe the system model we use to
+replicate segmented \invariantconfluent{} objects, and introduce a segmented
+\invariantconfluence{} decision procedure.
 
@@
 T_1), \ldots, (I_n, T_n)$ is a sequence of $n$ segments $(I_i, T_i)$ where
-every $T_i$ is a subset of $T$. $O$ is \defword{segmented invariant-confluent}
+every $T_i$ is a subset of $T$. $O$ is \defword{segmented \invariantconfluent{}}
 with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
@@
   \item
-    $O$ is invariant-confluent within each segment. That is, for every $(I_i,
+    $O$ is \invariantconfluent{} within each segment. That is, for every $(I_i,
     T_i) \in \Sigma$ and for every state $s \in I_i$, $O$ is
@@
 Now, we describe the system model used to replicate a segmented
-invariant-confluent object without any coordination within a segment and with
+\invariantconfluent{} object without any coordination within a segment and with
 only a small amount of coordination when transitioning between segments. As
@@
 or merge states with another server. Merging states in the segmented
-invariant-confluence system model is identical to merging states in the
-invariant-confluence system model. A server $p_i$ sends its state $s_i$ to
+\invariantconfluence{} system model is identical to merging states in the
+\invariantconfluence{} system model. A server $p_i$ sends its state $s_i$ to
 another server $p_j$ which \emph{must} merge $s_i$ into its state $s_j$.
@@
 \begin{algorithm}[t]
-  \caption{Transaction execution in segmented invariant-confluence system model}%
+  \caption{Transaction execution in segmented \invariantconfluence{} system model}%
   \algolabel{TxnExecution}
@@
 This system model guarantees that all replicas of a segmented
-invariant-confluent object always satisfy the invariant. All servers begin in
+\invariantconfluent{} object always satisfy the invariant. All servers begin in
 the same initial state and with the same active segment. Thus, because $O$ is
-invariant-confluent with respect to every segment, servers can execute
+\invariantconfluent{} with respect to every segment, servers can execute
 transactions within the active segment without any coordination and guarantee
 that the invariant is never violated. Any operation that would violate the
-assumptions of invariant-confluence system model (e.g.\ executing a transaction
+assumptions of \invariantconfluence{} system model (e.g.\ executing a transaction
 that's not permitted in the active segment or executing a permitted transaction
@@
 start state and with the same active replica, reestablishing the assumptions of
-the invariant-confluence system model.
+the \invariantconfluence{} system model.
 
@@
 In order to determine whether or not an object $O$ is \sTISconfluent{} we have
-to determine whether or not $O$ is invariant-confluent within each segment
+to determine whether or not $O$ is \invariantconfluent{} within each segment
 $(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
@@
 \in I$ are \defword{coreachable} with respect to $T$ and $I$, abbreviated
-\defword{$\TIcoreachable{s_1, s_2}$}, if there exists some state $s_0 \in I$
-such that $s_1$ and $s_2$ are both $\sTIreachable{}$.
+\defword{$\TIcoreachablepredicate{s_1, s_2}$}, if there exists some state $s_0 \in I$
+such that $s_1$ and $s_2$ are both \sTIreachable{}.
 
@@
   invariant $I$. If every pair of states in the invariant are coreachable (i.e.
-  $\forall s_1, s_2 \in S.\ \TIcoreachable{s_1, s_2}$), then
+  $\forall s_1, s_2 \in S.\ \TIcoreachablepredicate{s_1, s_2}$), then
   \[
@@
   For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
-  The two points are $\TIcoreachable{}$, so there exists some state $s_0$ from
+  The two points are \TIcoreachable{}, so there exists some state $s_0$ from
   which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
-  $\sTIreachable{}$, so it satisfies the invariant.
+  \sTIreachable{}, so it satisfies the invariant.
 \end{proof}
@@
 }
-Now, a segmented invariant-confluence decision procedure, can simply invoke
+Now, a segmented \invariantconfluence{} decision procedure, can simply invoke
 \algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.
@@
   \caption{%
-    Interactive invariant-confluence decision procedure for arbitrary start
+    Interactive \invariantconfluence{} decision procedure for arbitrary start
     state $s \in I$
@@
 
-    \algocomment{$R$ is a set of $\TIcoreachable{}$ states.}
-    \algocomment{$NR$ is a set of $\TIcounreachable{}$ states.}
+    \algocomment{$R$ is a set of \TIcoreachable{} states.}
+    \algocomment{$NR$ is a set of \TIcounreachable{} states.}
     \Function{Helper}{$O$, $T$, $I$, $R$, $NR$}
@@
   however, the user recognizes that the two states are not
-  $\ticoreachable{T^+}{I}{}$. The user refines $NR$ with the observation that
+  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that
   two states are coreachable if and only if they have the same values of $n_1,
@@
 \subsection{Discussion}
-There are a couple of things worth noting about segmented invariant-confluence,
-its system model, and its decision procedure. First, invariant-confluence is a
-very coarse-grained property. If an object is invariant-confluent, then we can
-replicate it with no coordination. If it is not invariant-confluent, then we
-have no guarantees. There's no in-between. Segmented invariant-confluence, on
+There are a couple of things worth noting about segmented \invariantconfluence{},
+its system model, and its decision procedure. First, \invariantconfluence{} is a
+very coarse-grained property. If an object is \invariantconfluent{}, then we can
+replicate it with no coordination. If it is not \invariantconfluent{}, then we
+have no guarantees. There's no in-between. Segmented \invariantconfluence{}, on
 the other hand, is a much more fine-grained property that can be applied to
 applications with varying degrees of complexity. Segmented
-invariant-confluence provides guarantees to complex applications that require a
+\invariantconfluence{} provides guarantees to complex applications that require a
 large number of segments and to simple applications with a smaller number of
-segments, whereas invariant-confluence only provides guarantees to applications
+segments, whereas \invariantconfluence{} only provides guarantees to applications
 that can be segmented into a single segment.
 
-Second, segmented invariant-confluence naturally subsumes a distributed locking
-approach to replicating non-invariant-confluent objects.  This approach first
+Second, segmented \invariantconfluence{} naturally subsumes a distributed locking
+approach to replicating non-\invariantconfluent{} objects.  This approach first
 recognizes which transactions cannot be safely executed concurrently and then
@@
 
-Fourth, a segmented invariant-confluence decision procedure can also leverage
+Fourth, a segmented \invariantconfluence{} decision procedure can also leverage
 \thmref{LatticeProperty} in addition to
@@
 I$.
-
-\todo{Forward reference examples in the evaluation section.}

### 2018-04-22 doc/paper/sections/introduction.tex
Intro tweaks.

@@
 unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
-abadi2012consistency}. On the other hand, weak consistency models like
-eventual consistency~\cite{vogels2009eventually}, PRAM consistency~\cite{lipton1988pram},
-causal consistency~\cite{ahamad1995causal}, and others~\cite{lloyd2011don,
-mehdi2017can} allow data to be replicated with high availability and low
-latency, but they put a tremendous burden on the application designer to reason
-about the complex interleavings of operations that are allowed by these weak
-consistency models. In fact, weak consistency models strip an application
-developer of one of the earliest and most effective tools that is used to
-reason about the execution of programs: application
+abadi2012consistency}. On the other hand, weak consistency models like eventual
+consistency~\cite{vogels2009eventually}, PRAM
+consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
+and others~\cite{lloyd2011don, mehdi2017can} allow data to be replicated with
+high availability and low latency, but they put a tremendous burden on the
+application designer to reason about the complex interleavings of operations
+that are allowed by these weak consistency models. In fact, weak consistency
+models strip an application developer of one of the earliest and most effective
+tools that is used to reason about the execution of programs: application
 invariants~\cite{hoare1969axiomatic, balegas2015towards}. Even if every
 transaction executing in a weakly consistent system individually maintains an
-application invariant, the system as a whole can produce invariant-violating
+application invariant, the system as a whole can produce invariant violating
 states.
@@
 the need for any coordination~\cite{bailis2014coordination}. If an object is
-invariant-confluent with respect to an invariant, we can replicate it with the
-performance benefits of weak consistency and (some of) the correctness benefits
-of strong consistency. However, the task of identifying whether or not an
-object actually is invariant-confluent remains a challenge. Bailis et al.\
+\invariantconfluent{} with respect to an invariant, we can replicate it with
+the performance benefits of weak consistency and (some of) the correctness
+benefits of strong consistency. However, the task of identifying whether or not
+an object actually is \invariantconfluent{} remains a challenge. Bailis et al.\
 manually categorized a set of common objects, transactions, and invariants
 (e.g.\ foreign key constraints on relations, linear constraints on integers) as
-invariant-confluent or not, but ideally we would have a general purpose program
-that can automatically determine \invariantconfluence{} for us. \textbf{The first
-main thrust of this paper is to make progress towards a general-purpose
-\invariantconfluence{} decision procedure.}
+\invariantconfluent{} or not, but ideally we would have a general purpose
+program that can automatically determine \invariantconfluence{} for us.
+\textbf{The first main thrust of this paper is to make progress towards a
+general-purpose \invariantconfluence{} decision procedure.}
 
@@
 difficult\footnote{Actually, it's impossible. Determining whether an object is
-invariant-confluent is, in general, undecidable. Still, we can develop a
+\invariantconfluent{} is, in general, undecidable. Still, we can develop a
 decision procedure that works well in the common case.}. Many existing
-approaches instead develop a decision procedure for invariant-closure, a
+approaches instead develop a decision procedure for \invariantclosure{}, a
 sufficient (but not necessary) condition for
 \invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
-approaches check to see if an object is invariant-closed. If it is, then they
-conclude that it is also invariant-confluent. If it's not, then the current
+approaches check to see if an object is \invariantclosed{}. If it is, then they
+conclude that it is also \invariantconfluent{}. If it's not, then the current
 approaches are unable to conclude anything about whether or not the object is
-invariant-confluent. In this paper, we take a step back and study the
-underlying reason \emph{why} invariant-closure is not necessary for
+\invariantconfluent{}. In this paper, we take a step back and study the
+underlying reason \emph{why} \invariantclosure{} is not necessary for
 \invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
@@
 object can actually be obtained in the execution of a system. We discover that
-unreachable, yet invariant-satisfying, states cause invariant-closure to be an
-unnecessary condition. Using this understanding, we construct a set of modest
-conditions under which invariant-closure and \invariantconfluence{} are in fact
-\emph{equivalent}, allowing us to reduce the problem of determining
-\invariantconfluence{} to that of determining invariant-closure.
+unreachable, yet invariant satisfying, states cause \invariantclosure{} to be
+an unnecessary condition. Using this understanding, we construct a set of
+modest conditions under which \invariantclosure{} and \invariantconfluence{}
+are in fact \emph{equivalent}, allowing us to reduce the problem of determining
+\invariantconfluence{} to that of determining \invariantclosure{}.
 
@@
 decision procedure, eliminating unreachable states from the invariant, moving
-towards the conditions under which invariant-closure and \invariantconfluence{}
-are equivalent. Meanwhile, the decision-procedure generates example reachable
-and unreachable states which help the user recognize patterns that describe
-reachability. We also leverage our intuitions about reachability to develop a
-new sufficient condition for \invariantconfluence{}, dubbed merge-reducibility.
-Merge-reducibility covers some cases that are not covered by invariant-closure,
-and it can be automatically checked without any user interaction.
+towards the conditions under which \invariantclosure{} and
+\invariantconfluence{} are equivalent. Meanwhile, the decision-procedure
+generates example reachable and unreachable states which help the user
+recognize patterns that describe reachability. We also leverage our intuitions
+about reachability to develop a new sufficient condition for
+\invariantconfluence{}, dubbed \mergereducibility. \Mergereducibility{} covers
+some cases that are not covered by \invariantclosure{}, and it can be
+automatically checked without any user interaction.
 
@@
 characterizes objects which can be replicated \emph{without any} coordination,
-segmented \invariantconfluence{} allows us to replicate non-invariant-confluent
-objects with \emph{minimal coordination}. The main idea is that
-invariant-satisfying states are divided into a number of segments, and the set
-of allowable of transactions within each segment is restricted. Within a
-segment, servers act without any coordination, synchronizing only to transition
-across segment boundaries. This design highlights the trade-off between
-application complexity and coordination-freedom; more complex applications
-require more segments which require more coordination, and vice-versa.
+segmented \invariantconfluence{} allows us to replicate
+non-\invariantconfluent{} objects with \emph{minimal coordination}. The main
+idea is that invariant satisfying states are divided into a number of segments,
+and the set of allowable transactions within each segment is restricted.
+Within a segment, servers act without any coordination, synchronizing only to
+transition across segment boundaries. This design highlights the trade-off
+between application complexity and coordination-freedom; more complex
+applications require more segments which require more coordination, and
+vice-versa.
 
@@
   \item
-    We develop an understanding of why invariant-closure is not necessary and
+    We develop an understanding of why \invariantclosure{} is not necessary and
     use this understanding to develop conditions under which it is both
@@
     We again exploit these conditions to design a novel non-trivial sufficient
-    condition for \invariantconfluence{}, merge-reducibility, that does not
+    condition for \invariantconfluence{}, \mergereducibility, that does not
     depend on user interaction (\secref{MergeReduction}).
@@
     invariants for replicated objects that are otherwise not
-    invariant-confluent (\secref{SegmentedInvariantConfluence}).
+    \invariantconfluent{} (\secref{SegmentedInvariantConfluence}).
 
@@
     We implement our decision procedures and sufficient conditions for both
-    \invariantconfluence{} and segmented \invariantconfluence{} in a system called
-    Lucy, then use Lucy to evaluate the efficacy or our methods
-    (\secref{Evaluation}).
+    \invariantconfluence{} and segmented \invariantconfluence{} in a system
+    called Lucy, then use Lucy to evaluate our methods (\secref{Evaluation}).
 \end{itemize}

### 2018-04-22 benchmark/get_data.sh
Changed withdraw to decrement.

@@ -1,0 +1,47 @@
+#! /usr/bin/env bash
+
+set -euo pipefail
+
+usage() {
+    echo "get_data.sh"
+}
+
+main() {
+    if [[ "$#" -ne 0 ]]; then
+        usage
+        exit 1
+    fi
+
+    if [[ ! -f hosts.txt ]]; then
+        echo "Create hosts.txt file with EC2 instance hostnames."
+    fi
+
+    if [[ ! -f "$HOME/.ssh/aws-ec2.pem" ]]; then
+        echo "Move your EC2 key pair to \$HOME/.ssh/aws-ec2.pem"
+    fi
+
+    hosts=()
+    while read host; do
+        hosts+=("$host")
+    done < hosts.txt
+
+    for host in "${hosts[@]}"; do
+        if [[ -e "$host" ]]; then
+            echo "Delete $host directory."
+            exit 1
+        fi
+        mkdir "$host"
+
+        scp \
+            -r \
+            -i ~/.ssh/aws-ec2.pem \
+            "ubuntu@$host:/home/ubuntu/mpc-bootstrap/data/Nov*" \
+            "$host"
+
+        for d in "$host"/*; do
+            mv $d $host/${host}_$(basename $d)
+        done
+    done
+}
+
+main

### 2018-04-22 benchmark/plot.py

@@ -49,7 +49,7 @@ def plot_vary_withdraws(data):
     plt.semilogy(segmented_x, segmented_y, **IC_FMT)
     plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
     plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.2), prop={'size': 12})
-    plt.xlabel('Fraction Withdrawals')
+    plt.xlabel('Decrement Frequency (fraction of workload)')
     plt.ylabel('Throughput (txns/s)')
     plt.savefig('vary_withdraws.pdf', bbox_inches='tight')
     plt.close()

### 2018-04-22 doc/paper/sections/interactive_decision_procedure.tex
Paper tweaks.

@@
 \begin{algorithm}[t]
-  \caption{Interactive invariant-confluence decision procedure}%
+  \caption{Interactive invariantconfluence{} decision procedure}%
   \algolabel{InteractiveDecisionProcedure}
@@
       \If {closed}
-        \State \Return true
+        \Return true
       \EndIf
-      \State Augment $R$ and $NR$ with random search and user input
+      \State Augment $R, NR$ with random search and user input
       \If{$s_1, s_2 \in R$}
         \Return false
-      \Else\
-        \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
       \EndIf
+      \State \Return \Call{Helper}{$O$, $s_0$, $T$, $I$, $R$, $NR$}
     \EndFunction
@@
 
-\thmref{IclosureEquivalentIconfluence} tells us that if all
-invariant-satisfying points are reachable, then invariant-closure and
-invariant-confluence are equivalent. In this section, we present the
-interactive invariant-confluence decision procedure shown in
+\thmref{IclosureEquivalentIconfluence} tells us that if all invariant
+satisfying points are reachable, then \invariantclosure{} and
+invariantconfluence{} are equivalent. In this section, we present the
+interactive invariantconfluence{} decision procedure shown in
 \algoref{InteractiveDecisionProcedure} which takes advantage of this result.
@@
 states from the invariant. Meanwhile, the algorithm leverages an
-invariant-closure decision procedure to either (a) conclude that $O$ is or is
+\invariantclosure{} decision procedure to either (a) conclude that $O$ is or is
 not \sTIconfluent{} or (b) provide counterexamples to the user to help them
@@
 from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
-the problem of invariant-confluence directly to the problem of
-invariant-closure, and the algorithm terminates.
+the problem of invariantconfluence{} directly to the problem of
+\invariantclosure{}, and the algorithm terminates.
 %
@@
 
-\IsInvConfluent{} assumes access to an invariant-closure decision procedure
+\IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
 \IsIclosed$(O, I)$ which returns a triple (closed, $s_1$, $s_2$). closed is a
@@
 but $\lnot I(s_1 \join s_2)$ (e.g. $s_1$ and $s_2$ from \exampleref{Z2}).  As
-we mentioned earlier, we can (and do) implement the invariant-closure decision
-procedure using an SMT solver like Z3~\cite{de2008z3}.
+we mentioned earlier, we can (and do) implement the \invariantclosure{}
+decision procedure using an SMT solver like Z3~\cite{de2008z3}.
 
@@
 $NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
-iteratively increased which removes unreachable states from $I$ until $I
-\subseteq \setst{s}{\sTIreachablepredicate{s}}$.
+iteratively increased, which removes unreachable states from $I$ until $I
+\subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$.
 
@@
 \]
-If $NR$ only contains \sTIunreachable{} states, then the set of
-\sTIreachable{} states is equal to set of \stireachable{s_0}{T}{I-NR}{}
-states which, as we just showed, is a subset of $I$.  Thus, $O$ is
-\sTIconfluent{}, so \Helper{} return true.
+Because $NR$ only contains \sTIunreachable{} states, then the set of
+\sTIreachable{} states is equal to set of \stireachable{s_0}{T}{I-NR}{} states
+which, as we just showed, is a subset of $I$.  Thus, $O$ is \sTIconfluent{}, so
+\Helper{} return true.
 
-If \IsIclosed{} determines that $O$ is \emph{not} \iclosed{(I - NR)}, then
-we have a counterexample $s_1, s_2$. We want to determine whether $s_1$ and
-$s_2$ are reachable or unreachable. We can do so in two ways.
+If \IsIclosed{} determines that $O$ is \emph{not} \iclosed{(I - NR)}, then we
+have a counterexample $s_1, s_2$. We want to determine whether $s_1$ and $s_2$
+are reachable or unreachable. We can do so in two ways.
 %
-First, we can randomly generate a set of reachable states and add them to
-$R$. If $s_1$ or $s_2$ lies in $R$, then we know they are reachable.
+First, we can randomly generate a set of reachable states and add them to $R$.
+If $s_1$ or $s_2$ is in $R$, then we know they are reachable.
 %
@@
 that certifies that $s$ is indeed reachable. This increases the burden on the
-user but completely eliminates this class of errors.
+user but completely eliminates this type of user error.
 
@@
 states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
-$NR$ is incorrect and contains some \sTIreachable{} states, then some of
-these states are artificially made unreachable, but the system is still
-guaranteed to never produce a state which violates $I$.
+$NR$ is incorrect and contains some \sTIreachable{} states, then some of these
+states are artificially made unreachable, but the system is still guaranteed to
+never produce a state which violates $I$.

### 2018-04-22 doc/paper/sections/invariant_closure.tex
Paper tweaks.

@@
 \invariantconfluence{} instead focus on deciding a sufficient condition for
-\invariantconfluence{}---dubbed \defword{\invariantclosure{}}---that is simpler to
-reason about~\cite{li2012making, li2014automating}. In this section, we define
-\invariantclosure{} and study why the condition is sufficient but not necessary.
-Armed with this understanding, we present conditions under which it is both
-sufficient and necessary.
+\invariantconfluence{}---dubbed \defword{\invariantclosure{}}---that is simpler
+to reason about~\cite{li2012making, li2014automating}. In this section, we
+define \invariantclosure{} and study why the condition is sufficient but not
+necessary.  Armed with this understanding, we present conditions under which it
+is both sufficient and necessary.
 
-We say an object $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to
-an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying
+We say an object $O = (S, \join)$ is \defword{\invariantclosed{}} with respect
+to an invariant $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying
 states are closed under merge. That is, for every state $s_1, s_2 \in S$, if
@@
 \thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
-sufficient for \invariantconfluence{}. Intuitively, our system model ensures that
-transaction execution preserves the invariant, so if merging states also
+sufficient for \invariantconfluence{}. Intuitively, our system model ensures
+that transaction execution preserves the invariant, so if merging states also
 preserves the invariant and if our start state satisfies the invariant, then
@@
 \invariantconfluence{}, so if an object is \emph{not} \invariantclosed{}, these
-systems cannot conclude that the object is \emph{not} \invariantconfluent{}. The
-reason why \invariantclosure{} is not necessary for \invariantconfluence{} is best
-explained through an example.
+systems cannot conclude that the object is \emph{not} \invariantconfluent{}.
+The reason why \invariantclosure{} is not necessary for \invariantconfluence{}
+is best explained through an example.
+
+{\input{figures/z2}}
 
@@
 
-{\input{figures/z2}}
-
 In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
 \invariantclosed{}, but $s_1$ is not reachable. This is not particular to
-\exampleref{Z2}. In fact, it is fundamentally the reason why \invariantclosure{}
-is not equivalent to \invariantconfluence{}. \Invariantconfluence{} is, at its
-core, a property of reachable states, but \invariantclosure{} is completely
-ignorant of reachability. As a result, invariant satisfying yet unreachable
-states like $s_1$ are the key hurdle preventing \invariantclosure{} from being
-equivalent to \invariantconfluence{}. This is formalized by
-\thmref{IclosureEquivalentIconfluence}
+\exampleref{Z2}. In fact, it is fundamentally the reason why
+\invariantclosure{} is not equivalent to \invariantconfluence{}.
+\Invariantconfluence{} is, at its core, a property of reachable states, but
+\invariantclosure{} is completely ignorant of reachability. As a result,
+invariant satisfying yet unreachable states like $s_1$ are the key hurdle
+preventing \invariantclosure{} from being equivalent to \invariantconfluence{}.
+This is formalized by \thmref{IclosureEquivalentIconfluence}
 
@@
   transactions $T$, and an invariant $I$. If the invariant is a subset of the
-  reachable states (i.e. $I \subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$), then
+  reachable states (i.e. $I \subseteq \setst{s \in
+  S}{\sTIreachablepredicate{s}}$), then
   \[
@@
 \end{proof}
-
-% - define \invariantclosure{} and introduce it as a sufficient condition
-% - cite existing work to show that people have used invariant closure to
-%   reason about \invariantconfluence{}
-% - explain why invariant closure can fail and why its not sufficient using the
-%   two-integers example
-% - state main theorem

### 2018-04-22 doc/paper/sections/invariant_confluence.tex
Paper tweaks.

@@
 \subsection{System Model}
-A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of
-states and a binary merge operator $\join: S \times S \to S$ which merges two
-states into one. A \defword{transaction} $t: S \to S$ is a function which maps
-one state to another. An \defword{invariant} $I$ is a subset of $S$.
-Notationally, we write $I(s)$ to denote that $s$ satisfies the invariant (i.e.
-$s \in I$) and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant
-(i.e. $s \notin I$).
-%
-For example, $O = (\ints, +)$ is a distributed object consisting of integers
-merged by addition; $t(x) = x + 1$ is a transaction which adds one to a state;
-and $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
+A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states
+and a binary merge operator $\join: S \times S \to S$ which merges two states
+into one. A \defword{transaction} $t: S \to S$ is a function which maps one
+state to another. An \defword{invariant} $I$ is a subset of $S$.  Notationally,
+we write $I(s)$ to denote that $s$ satisfies the invariant (i.e.  $s \in I$)
+and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant (i.e. $s
+\notin I$).
+
+\begin{example}\examplelabel{Ints}
+  $O = (\ints, +)$ is a distributed object consisting of integers merged by
+  addition; $t(x) = x + 1$ is a transaction which adds one to a state; and
+  $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
   non-negative.
+\end{example}
 
@@
 
-Informally, $O$ is \defword{\invariantconfluent{}} with respect to $s_0$, $T$, and
-$I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots, s_n$
-is guaranteed to always satisfy the invariant $I$ in every possible execution
-of the system.
+Informally, $O$ is \defword{\invariantconfluent{}} with respect to $s_0$, $T$,
+and $I$, abbreviated \defword{\sTIconfluent{}}, if every replica $s_1, \ldots,
+s_n$ is guaranteed to always satisfy the invariant $I$ in every possible
+execution of the system.
 
@@
 Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists a
-\sTIreachable{} expression $e$ that evaluates to $s$.
+\sTIreachable{} expression $e$ that evaluates to $s$. Returning to
+\exampleref{Ints} with start state $s_0 = 42$, we see that all integers greater
+than or equal to 42 (i.e.\ $\setst{x \in \ints}{x \geq 42}$) are
+\sTIreachable{}, and all other integers are \sTIunreachable{}.
 
@@
 $T$, and $I$, abbreviated \defword{\sTIconfluent{}}, if all reachable states
-satisfy the invariant: $\setst{s \in S}{\sTIreachablepredicate{s}} \subseteq
-I$.
+satisfy the invariant:
+\[
+  \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
+\]

### 2018-04-22 doc/paper/sections/merge_reduction.tex
Paper tweaks.

@@
 a start state $s_0 \in S$, a set of transactions $T$, and an invariant $I$,
-abbreviated \defword{\sTImergereducible}, if for every pair of $e_1$ and $e_2$
-of merge-free \sTIreachable{} expressions, there exists some merge-free
-\sTIreachable{} expression $e_3$ that evaluates to the same state as $e_1
-\join e_2$. Intuitively, if $O$ is merge-reducible, we can replace $e_1 \join
-e_2$ (which has one merge) with $e_3$ (which has no merges) to obtain an
-equivalent expression with fewer merges.
+abbreviated \defword{\sTImergereducible}, if for every pair $e_1$ and $e_2$ of
+merge-free \sTIreachable{} expressions, there exists some merge-free
+\sTIreachable{} expression $e_3$ that evaluates to the same state as $e_1 \join
+e_2$. Intuitively, if $O$ is merge-reducible, we can replace $e_1 \join e_2$
+(which has one merge) with $e_3$ (which has no merges) to obtain an equivalent
+expression with fewer merges.
 
@@
 states that criteria (1) -- (4) are sufficient conditions for
-\invariantconfluence{}. How do these sufficient conditions relate to one another?
-Clearly, not all \invariantclosed{} objects are semilattices, so
-\invariantclosure{} does not imply criteria (1) -- (4). Conversely, there are some
-objects which satisfy criteria (1) -- (4) which are not \invariantclosed{}. Here's
-one example.
+\invariantconfluence{}. How do these sufficient conditions relate to one
+another?  Clearly, not all \invariantclosed{} objects are semilattices, so
+\invariantclosure{} does not imply criteria (1) -- (4). Conversely, there are
+some objects which satisfy criteria (1) -- (4) which are not
+\invariantclosed{}. Here's one example.
 
@@
 
-\Invariantclosure{} is not necessary for \invariantconfluence{} because it fails to
-incorporate any notion of reachability. Criteria (1) -- (4) are also
+\Invariantclosure{} is not necessary for \invariantconfluence{} because it
+fails to incorporate any notion of reachability. Criteria (1) -- (4) are also
 unnecessary, but they can be used to prove that some non-\invariantclosed{}
@@
 is not $I$-closed because the union of an odd-only set with an even-only set is
-a set with both odd and even integers. However, if we begin in an
-invariant satisfying state, we cannot reach both an odd-only and even-only set.
+a set with both odd and even integers. However, if we begin in an invariant
+satisfying state, we cannot reach both an odd-only and even-only set.
 Criterion (3) is able to recognize this fact and conclude that $O$ is
 \invariantconfluent{} despite it not being \invariantclosed{}.
-
-% TODO: Discuss generality?

### 2018-04-22 doc/paper/sections/segmented_invariant_confluence.tex
Paper tweaks.

@@
 \defword{segmented \invariantconfluence{}} that can be used to maintain the
-invariants of non-\invariantconfluent{} objects, requiring only a small amount of
-coordination.
+invariants of non-\invariantconfluent{} objects, requiring only a small amount
+of coordination.
 
@@
 \emph{within each segment} (even though it may not be globally
-\invariantconfluent{}). Then, servers can run coordination-free within a segment
-and need only coordinate when transitioning from one segment to another. We now
-formalize segmented \invariantconfluence{}, describe the system model we use to
-replicate segmented \invariantconfluent{} objects, and introduce a segmented
-\invariantconfluence{} decision procedure.
+\invariantconfluent{}). Then, servers can run coordination-free within a
+segment and need only coordinate when transitioning from one segment to
+another. We now formalize segmented \invariantconfluence{}, describe the system
+model we use to replicate segmented \invariantconfluent{} objects, and
+introduce a segmented \invariantconfluence{} decision procedure.
 
@@
 set of transitions $T$, and an invariant $I$. A segmentation $\Sigma = (I_1,
-T_1), \ldots, (I_n, T_n)$ is a sequence of $n$ segments $(I_i, T_i)$ where
-every $T_i$ is a subset of $T$. $O$ is \defword{segmented \invariantconfluent{}}
-with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
-\defword{\sTISconfluent}, if the following conditions hold:
+T_1), \ldots, (I_n, T_n)$ is a sequence (not a set) of $n$ segments $(I_i,
+T_i)$ where every $T_i$ is a subset of $T$. $O$ is \defword{segmented
+\invariantconfluent{}} with respect to $s_0$, $T$, $I$, and $\Sigma$,
+abbreviated \defword{\sTISconfluent}, if the following conditions hold:
 \begin{itemize}
@@
     $I$ is covered by the invariants in $\Sigma$ (i.e. $I = \cup_{i=1}^n I_i$).
-    Note that invariants in $\Sigma$ do \emph{not} have to be disjoint. Thus,
-    they do not have to partition $I$; they just have to cover $I$.
+    Note that invariants in $\Sigma$ do \emph{not} have to be disjoint. That
+    is, they do not have to partition $I$; they just have to cover $I$.
 
@@
 
-$t \in T_i$, then $p_i$ executes $t$ immediately and without coordination. If
-$t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then $p_i$
-commits $t$, stays in state $t(s_0)$, and replies to the client. If $t(s_i)$
-violates the invariant (i.e.  $i.e. \lnot I(t(s_i))$), then $p_i$ aborts $t$,
-reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies the
-invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
+If $t \in T_i$, then $p_i$ executes $t$ immediately and without coordination.
+If $t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then
+$p_i$ commits $t$, stays in state $t(s_i)$, and replies to the client. If
+$t(s_i)$ violates the invariant (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts
+$t$, reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies
+the invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
 I_i(t(s_i))$), then $p_i$ reverts to state $s_i$ and initiates a round of
@@
 \begin{algorithm}[t]
-  \caption{Transaction execution in segmented \invariantconfluence{} system model}%
+  \caption{%
+    Transaction execution in the segmented \invariantconfluence{} system model}%
   \algolabel{TxnExecution}
@@
 that the invariant is never violated. Any operation that would violate the
-assumptions of \invariantconfluence{} system model (e.g.\ executing a transaction
-that's not permitted in the active segment or executing a permitted transaction
-that leads to a state outside the active segment) triggers a global
+assumptions of the \invariantconfluence{} system model (e.g.\ executing a
+transaction that's not permitted in the active segment or executing a permitted
+transaction that leads to a state outside the active segment) triggers a global
 coordination. Globally coordinated transactions are only executed if they
@@
 \subsection{Interactive Decision Procedure}
-In order to determine whether or not an object $O$ is \sTISconfluent{} we have
-to determine whether or not $O$ is \invariantconfluent{} within each segment
-$(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
+In order for us to determine whether or not an object $O$ is \sTISconfluent{},
+we have to determine whether or not $O$ is \invariantconfluent{} within each
+segment $(I_i, T_i) \in \Sigma$. That is, we have determine if $O$ is
 \sticonfluent{s}{T_i}{I_i} confluent for every state $s \in I_i$. Ideally, we
@@
 $s \in I_i$, not for \emph{every} state $s \in I_i$. Thus, we would have to
-invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times fore every segment
+invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times for every segment
 $(I_i, T_I)$, which is clearly infeasible given that $I_i$ can be large or even
@@
 \in I$ are \defword{coreachable} with respect to $T$ and $I$, abbreviated
-\defword{$\TIcoreachablepredicate{s_1, s_2}$}, if there exists some state $s_0 \in I$
-such that $s_1$ and $s_2$ are both \sTIreachable{}.
+\defword{\TIcoreachable}, if there exists some state $s_0 \in I$ such that
+$s_1$ and $s_2$ are both \sTIreachable{}.
 
@@
   Consider an object $O = (S, \join)$, a set of transactions $T$, and an
-  invariant $I$. If every pair of states in the invariant are coreachable (i.e.
-  $\forall s_1, s_2 \in S.\ \TIcoreachablepredicate{s_1, s_2}$), then
+  invariant $I$. If every pair of states in the invariant are \TIcoreachable{},
+  then
   \[
@@
       \If {closed}
-        \State \Return true
+        \Return true
       \EndIf
-      \State Augment $R$ and $NR$ with random search and user input
+      \State Augment $R, NR$ with random search and user input
       \If{$(s_1, s_2) \in R$}
         \Return false
-      \Else\
-        \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
       \EndIf
+      \State \Return \Call{Helper}{$O$, $T$, $I$, $R$, $NR$}
     \EndFunction
@@
   Let $O = (\ints^3 \times \ints^3, \join)$ be a
-  PN-Counter~\cite{shapiro2011comprehensive} replicated on $3$ machines. Every
-  state $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ represents the integer $(p_1 +
-  p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement the counter, the
-  $i$th server increments $p_i$ and $n_i$ respectively, and $\join$ computes an
-  element-wise maximum. Our start state $s_0 = (0, 0, 0), (0, 0, 0)$; our set
-  $T$ of transactions consists of increment and decrement; and our invariant
-  $I$ is that the value of $s$ is non-negative.
+  PN-Counter~\cite{shapiro2011comprehensive} replicated on three machines.
+  Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
+  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement
+  the counter, the $i$th server increments $p_i$ and $n_i$ respectively, and
+  $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
+  (0, 0, 0)$; our set $T$ of transactions consists of increment and decrement;
+  and our invariant $I$ is that the value of $s$ is non-negative.
 
@@
   transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
-  \IsIclosed{} again returns false with states $s_1$ and $s_2$. This time,
+  \IsIclosed{} again returns false with same states $s_1$ and $s_2$. This time,
   however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that
-  two states are coreachable if and only if they have the same values of $n_1,
-  n_2, n_3$. After this, \IsIclosed{} and
+  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that two
+  states are coreachable if and only if they have the same values of $n_1, n_2,
+  n_3$. After this, \IsIclosed{} and
   \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
@@
 \subsection{Discussion}
-There are a couple of things worth noting about segmented \invariantconfluence{},
-its system model, and its decision procedure. First, \invariantconfluence{} is a
-very coarse-grained property. If an object is \invariantconfluent{}, then we can
-replicate it with no coordination. If it is not \invariantconfluent{}, then we
-have no guarantees. There's no in-between. Segmented \invariantconfluence{}, on
-the other hand, is a much more fine-grained property that can be applied to
-applications with varying degrees of complexity. Segmented
-\invariantconfluence{} provides guarantees to complex applications that require a
-large number of segments and to simple applications with a smaller number of
-segments, whereas \invariantconfluence{} only provides guarantees to applications
-that can be segmented into a single segment.
+There are a couple of things worth noting about segmented
+\invariantconfluence{}, its system model, and its decision procedure. First,
+\invariantconfluence{} is a very coarse-grained property. If an object is
+\invariantconfluent{}, then we can replicate it with no coordination. If it is
+not \invariantconfluent{}, then we have no guarantees. There's no in-between.
+Segmented \invariantconfluence{}, on the other hand, is a much more
+fine-grained property that can be applied to applications with varying degrees
+of complexity. Segmented \invariantconfluence{} provides guarantees to complex
+applications that require a large number of segments and to simple applications
+with a smaller number of segments, whereas \invariantconfluence{} only provides
+guarantees to applications that can be segmented into a single segment.
 
-Second, segmented \invariantconfluence{} naturally subsumes a distributed locking
-approach to replicating non-\invariantconfluent{} objects.  This approach first
-recognizes which transactions cannot be safely executed concurrently and then
-requires them to acquire a distributed lock before
+Second, segmented \invariantconfluence{} naturally subsumes a distributed
+locking approach to replicating non-\invariantconfluent{} objects.  This
+approach first recognizes which transactions cannot be safely executed
+concurrently and then requires them to acquire a distributed lock before
 executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a

### 2018-04-22 doc/paper/sections/conclusion.tex
Finished paper tweaks.

--- 2018-04-22 doc/paper/sections/conclusion.tex
@@
 \section{Conclusion}
-TODO
+This paper revolved around two major contributions.  First, we developed a
+deeper understanding of \invariantclosure{} and \invariantconfluence{} by
+looking at the two criteria with reachability in mind. We found that
+\invariantclosure{} fails to incorporate a notion of reachability, and using
+this intuition, we developed conditions under which \invariantclosure{} and
+\invariantconfluence{} are equivalent.  Second, we proposed a new consistency
+model and generalization of \invariantconfluence{}, segmented
+\invariantconfluence{}, that can be used to replicate non-\invariantconfluent{}
+objects with a small amount of coordination while still preserving their
+invariants. We also found that segmented \invariantconfluence{} naturally
+subsumes existing techniques for maintaining invariants of replicated objects
+(e.g.\ locking and escrow transactions).
 
 \textbf{Acknowledgments.}
-The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan,
-and Cristina Teodoropol for fruitful discussion and feedback.
+The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan, and
+Cristina Teodoropol for fruitful discussion and feedback.
 %

### 2018-04-22 doc/paper/sections/evaluation.tex
Finished paper tweaks.

@@
 our decision procedures and system models.
-% We evaluate Lucy by answering the following questions:
-% \begin{itemize}
-%   \item
-%     How practical is the interactive \invariantconfluence{} decision procedure?
-%     Can we use it to classify real-world transactions and invariants?
-%   \item
-%     How practical is segmented \invariantconfluence{}? Are real-world workloads
-%     amenable to segmentation?
-%   \item
-%     How efficient is the interactive \invariantconfluence{} decision procedure?
-%   \item
-%     How efficiently can we replicate a segmented \invariantconfluent{} object as
-%     compared to alternative approaches like replicating with weak or strong
-%     consistency?
-%   \item
-%     How does the performance of replicating a segmented \invariantconfluence{}
-%     object vary as we vary the workload, segmentation, and replication factor?
-% \end{itemize}
 
@@
 then replicates the objects using segmented \invariantconfluence{} (or
-\invariantconfluence{} if the segmentation contains a single segment without any
-disallowed transactions). Clients send every transaction request to a randomly
-selected server. When a server receives a transaction request, it executes
-\algoref{TxnExecution} to attempt to execute the transaction locally. If the
-transaction requires global coordination, then the server forwards the
+\invariantconfluence{} if the segmentation contains a single segment without
+any disallowed transactions). Clients send every transaction request to a
+randomly selected server. When a server receives a transaction request, it
+executes \algoref{TxnExecution} to attempt to execute the transaction locally.
+If the transaction requires global coordination, then the server forwards the
 transaction request to a predetermined leader. When the leader receives a
@@
 sends a merge request to a randomly selected server.
-% Merge requests are tagged with a monotonically increasing epoch number that is
-% incremented by the master after every round of global coordination. This allows
-% servers to discard merge requests from previous epochs.
 
@@
 \cite{lamport1998part, liskov2012viewstamped}.
-% In \secref{SegmentedInvariantConfluenceEval}, we compare the performance of
-% replicating with segmented \invariantconfluence{} against the performance of
-% replicating with eventual consistency and linearizability.
 
 Because fault-tolerance is largely an orthogonal concern to
-\invariantconfluence{}, Lucy is implemented without fault-tolerance. It would be
-straightforward to add fault-tolerance to Lucy, but it would not affect our
+\invariantconfluence{}, Lucy is implemented without fault-tolerance. It would
+be straightforward to add fault-tolerance to Lucy, but it would not affect our
 discussions or evaluation, so we leave it for future work.
-% Moreover, users currently have to specify their workloads in Python (for the
-% decision procedures) and C++ (for the runtime). In the future, we plan on
-% removing this redundancy.
 
@@
 counterexample and refine the invariant with $y \leq 0$, the interactive
-decision procedure determines that the object is \invariantconfluent{}, again, in
-less than a tenth of a second. Note that the object is \invariantconfluent{} but
-\emph{not} \invariantclosed{}, so prior work that relies on \invariantclosure{}
-alone to determine \invariantconfluence{} would not be able to identify this
-example as \invariantconfluent{}.
+decision procedure determines that the object is \invariantconfluent{}, again,
+in less than a tenth of a second. Note that the object is \invariantconfluent{}
+but \emph{not} \invariantclosed{}, so prior work that relies on
+\invariantclosure{} alone to determine \invariantconfluence{} would not be able
+to identify this example as \invariantconfluent{}.
 
@@
 constraint from $X$ to $Y$. We ran our decision procedure on the object with
-initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and
-with transactions that allow arbitrary insertions and deletions into $X$ and
-$Y$. After less than a tenth of a second, the decision procedure produced a
+initial state $((\emptyset, \emptyset), (\emptyset, \emptyset))$ and with
+transactions that allow arbitrary insertions and deletions into $X$ and $Y$.
+After less than a tenth of a second, the decision procedure produced a
 reachable counterexample witnessing the fact that the object is not
-\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$ can
-lead to a state which violates the invariant. This object is not
-\invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous tools
-depending on \invariantclosure{} as a sufficient condition would be unable to
-conclude definitively that the object is \emph{not} \invariantconfluent{}.
+\invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$
+can lead to a state which violates the invariant. This object is not
+\invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous
+tools depending on \invariantclosure{} as a sufficient condition would be
+unable to conclude definitively that the object is \emph{not}
+\invariantconfluent{}.
 
@@
 winning bid yet) and we merge states by taking the union of $B$ and the maximum
-of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a bid
-$b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes the
-auction and sets $w$ equal to the largest bid in $B$. Our invariant is that if
-the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
+of $w$ (where $\bot < n$ for all integers $n$). One transaction $t_b$ places a
+bid $b$ by inserting it into $B$. Another transaction $t_\text{close}$ closes
+the auction and sets $w$ equal to the largest bid in $B$. Our invariant is that
+if the auction is closed (i.e.\ $w \neq \bot$), then $w = \max(B)$. We ran our
 decision procedure on this example and in a third of a second, it produced a
@@
 on this example, and it was able to deduce without any human interaction that
-the example was segmented \invariantconfluent{} in less than a tenth of a second.
+the example was segmented \invariantconfluent{} in less than a tenth of a
+second.
 
@@
 \exampleref{CounreachableExample}, we found that concurrent decrements violate
-\invariantconfluence{} which led us to a segmentation which prohibited concurrent
-decrements. We now propose a new segmentation with escrow amount $k$ that will
-allow us to perform concurrent decrements that respect the escrowed value. The
-first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2, p_3 \geq k
-\land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments and
-decrements so long as every $p_i \geq k$ and every $n_i \leq k$. Intuitively,
-this segment represents the situation in which every server has escrowed a
-value of $k$. They can decrement freely, so long as they don't exceed their
-escrow budget of $k$. The second segment is the one presented in
+\invariantconfluence{} which led us to a segmentation which prohibited
+concurrent decrements. We now propose a new segmentation with escrow amount $k$
+that will allow us to perform concurrent decrements that respect the escrowed
+value. The first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2,
+p_3 \geq k \land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments
+and decrements so long as every $p_i \geq k$ and every $n_i \leq k$.
+Intuitively, this segment represents the situation in which every server has
+escrowed a value of $k$. They can decrement freely, so long as they don't
+exceed their escrow budget of $k$. The second segment is the one presented in
 \exampleref{CounreachableExample} which prohibits concurrent decrements. We ran
@@
   \item
-    \newcommand{\ttt}[1]{{\smaller \texttt{#1}}}
-    Seven of the twelve invariants involve \textbf{maintaining arithmetic
-    relationships between relations}. Our decision procedures can correctly
-    identify these as \invariantconfluent{}. Consider, for example, invariant 1
-    which dictates that a warehouse's year to date balance \ttt{W\_YTD} is
-    equal to the sum of the district year to date balances \ttt{D\_YTD} of the
-    twenty districts that are associated with the warehouse. The Payment
-    transaction randomly selects a district and increments \ttt{W\_YTD} and
-    \ttt{D\_YTD} by a randomly generated amount. We model this workload with a
-    PN-Counter for \ttt{W\_YTD} and twenty PN-Counters for the twenty instances
-    of \ttt{D\_YTD}. We applied Lucy to this workload, and it determined that
-    the workload was \invariantconfluent{} in less than a second.
+    \newcommand{\ttt}[1]{{\smaller \texttt{#1}}} Seven of the twelve invariants
+    involve \textbf{maintaining arithmetic relationships between relations}.
+    Our decision procedures can correctly identify these as
+    \invariantconfluent{}. Consider, for example, invariant 1 which dictates
+    that a warehouse's year to date balance \ttt{W\_YTD} is equal to the sum of
+    the district year to date balances \ttt{D\_YTD} of the twenty districts
+    that are associated with the warehouse. The Payment transaction randomly
+    selects a district and increments \ttt{W\_YTD} and \ttt{D\_YTD} by a
+    randomly generated amount. We model this workload with a PN-Counter for
+    \ttt{W\_YTD} and twenty PN-Counters for the twenty instances of
+    \ttt{D\_YTD}. We applied Lucy to this workload, and it determined that the
+    workload was \invariantconfluent{} in less than a second.
 
@@
 Now, we evaluate the performance of replicating an object with segmented
-\invariantconfluence{} as compared to replicating it with eventual consistency or
-linearizability.
-% We begin with two benchmarks that demonstrate the same concept: the performance
-% of segmented \invariantconfluent{} replication varies with the amount of global
-% coordination induced by either (a) performing a transaction that is disallowed
-% within a segment or (b) transitioning between segments.
+\invariantconfluence{} as compared to replicating it with eventual consistency
+or linearizability.
 
@@
   forbids concurrent decrements. We replicate this object on 32 servers
-deployed on 32 m5.xlarge EC2 instances within the same availability zone.  Each
-server has three colocated clients that issue deposit and withdrawal
+  deployed on 32 m5.xlarge EC2 instances within the same availability zone.
+  Each server has three colocated clients that issue increment and decrement
   transactions. We replicate the object with eventual consistency, segmented
   \invariantconfluence{}, and linearizability and measure the system's total
-throughput as we vary the fraction of client requests that are withdrawals. The
-results are shown in the top of \figref{ThroughputVsGlobalSyncs}.
+  throughput as we vary the fraction of client requests that are decrements.
+  The results are shown in the top of \figref{ThroughputVsGlobalSyncs}.
 
@@
   unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
-per second respectively.
+  per second respectively. Segmented \invariantconfluent{} replication performs
+  well for low-decrement workloads and performs increasingly poorly as we
+  increase the fraction of decrement transactions, eventually performing worse
+  than linearizable replication. For example, with 5\% decrement transactions,
+  segmented \invariantconfluent{} replication performs an order of magnitude
+  better than linearizable replication; with 50\% decrements, it performs as
+  well; and with 100\% decrements, it performs two times worse.
+
+  These results are expected. Increment transactions can execute without any
+  coordination while decrement transactions require global coordination. As we
+  increase the fraction of decrements, we increase the amount of coordination
+  that the system has to perform which in turn drastically decreases the
+  throughput. These results also offer two insights.
   %
-% Expectedly, eventually consistent replication significantly outperforms
-% linearizable replication because (a) transactions can be sent to any server
-% (not just the leader) and (b) servers do not coordinate with each other at all.
+  First, for low-decrement workloads, segmented \invariantconfluent{}
+  replication achieves a compromise between strong and weak consistency. It
+  guarantees that invariants are maintained (which is impossible with eventual
+  consistency if the object is not \invariantconfluent{}) with performance many
+  times better than strongly consistent replication.
   %
-Segmented \invariantconfluent{} replication performs well for low-withdrawal
-workloads and performs increasingly poorly as we increase the fraction of
-withdrawal transactions, eventually performing worse than linearizable
-replication. For example, with 5\% withdrawal transactions, segmented
-\invariantconfluent{} replication performs an order of magnitude better than
-linearizable replication; with 50\% withdrawals, it performs as well; and with
-100\% withdrawals, it performs two times worse.
-
-These results are expected. Deposit transactions can execute without any
-coordination while withdrawal transactions require global coordination. As we
-increase the fraction of withdrawals, we increase the amount of coordination
-that the system has to perform which in turn drastically decreases the
-throughput. These results also offer two insights:
-%
-First, for low-withdrawal workloads, segmented \invariantconfluent{} replication
-achieves a compromise between strong and weak consistency. It guarantees that
-invariants are maintained (which is impossible with eventual consistency if the
-object is not \invariantconfluent{}) with performance many times better than
-strongly consistent replication.
-%
-Second, segmented \invariantconfluent{} replication is poorly suited to workloads
-that require a large amount of coordination. For workloads without much inherit
-concurrency (e.g.\ withdraw-mostly workloads), maintaining invariants is best
-done with strong consistency. It provides stronger guarantees with better
-performance.
+  Second, segmented \invariantconfluent{} replication is poorly suited to
+  workloads that require a large amount of coordination. For workloads without
+  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
+  invariants is best done with strong consistency. It provides stronger
+  guarantees with better performance.
 \end{benchmark}
@@
   consistency outperforms linearizability by roughly two orders of magnitude.
-  In this example, the segmented \invariantconfluent{} replication only requires
-  coordination when transitioning between segment boundaries, so as we increase
-  the segment side length, the throughput of the system increases
+  In this example, the segmented \invariantconfluent{} replication only
+  requires coordination when transitioning between segment boundaries, so as we
+  increase the segment side length, the throughput of the system increases
   significantly.
@@
 \begin{benchmark}
-  In this benchmark, we measure the scale-out of segmented \invariantconfluent{}
-  replication. We repeat \benchref{VaryWithdraws} with a 10\% withdrawal rate,
-  but this time we vary the number of servers we use to replicate our object.
-  When we replicate with $n$ servers, we use $3n$ clients (the $3$ colocated
-  clients on each server) as part of the workload. The results are shown in
-  \figref{VaryNodes}.
+  In this benchmark, we measure the scale-out of segmented
+  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} with a
+  10\% decrement rate, but this time we vary the number of servers we use to
+  replicate our object.  When we replicate with $n$ servers, we use $3n$
+  clients (the $3$ colocated clients on each server) as part of the workload.
+  The results are shown in \figref{VaryNodes}.
 
   Eventually consistent replication scales perfectly with the number of nodes,
-  confirming the results in~\cite{bailis2014coordination}.
-  %
-  % With eventually consistent replication, servers do not coordinate at all, so
-  % they are completely unaffected by the number of servers.
-  %
-  Linearizable replication, on the other hand, scales up to about 3-5 servers
-  before performance begins to decrease. These numbers are consistent with
-  typical deployments of state-machine replication protocols like
-  Paxos~\cite{chandra2007paxos}.
-  %
-  % Because all messages are sent to the leader, the leader becomes the
-  % bottleneck as the number of servers and clients increases. Moreover, the
-  % leader must wait for responses from more servers, increasing the latency of
-  % the slowest response which in turn decreases throughput.
-  %
-  Segmented \invariantconfluent{} replication scales up to about 6-8 servers
-  before succumbing to the same scalability bottlenecks as linearizable
-  replication.
+  confirming the results in~\cite{bailis2014coordination}.  Linearizable
+  replication, on the other hand, scales up to about 3--5 servers before
+  performance begins to decrease. These numbers are consistent with typical
+  deployments of state-machine replication protocols like
+  Paxos~\cite{chandra2007paxos}. Segmented \invariantconfluent{} replication
+  scales up to about 6--8 servers before succumbing to the same scalability
+  bottlenecks as linearizable replication.
 
   These results highlight the importance of coordination avoidance in
-  distributed databases. While segmented \invariantconfluent{} replication scales
-  out slightly better than linearizable replication, both scale significantly
-  worse than eventually consistent replication even for a very low (i.e.\ 10\%)
-  withdrawal workload. This demonstrates that even a small amount of
-  coordination can significantly reduce the scalability of a system.
+  distributed databases. While segmented \invariantconfluent{} replication
+  scales out slightly better than linearizable replication, both scale
+  significantly worse than eventually consistent replication even for a very
+  low (i.e.\ 10\%) decrement workload. This demonstrates that even a small
+  amount of coordination can significantly reduce the scalability of a system.
 \end{benchmark}

### 2018-04-22 doc/paper/sections/related_work.tex
Finished paper tweaks.

@@
 %
-Invariant safety is a sufficient condition for RedBlue consistent objects to be
-\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{} for an
-operation-based system model (as opposed-to the state based system model that
-we use). In~\cite{li2014automating}, Li et al.\ develop techniques to identify
-invariant safe operations. Like \invariantclosure{}, invariant safety is a
-sufficient but not necessary condition for \invariantconfluence{}. Li et al.\
+Invariant safety is a sufficient (but not necessary) condition for RedBlue
+consistent objects to be \invariantconfluent{} and is an analog of
+\invariantclosure{} for an operation-based system model (as opposed to the
+state based system model that we use).  In~\cite{li2014automating}, Li et al.\
 develop sophisticated techniques for deciding invariant safety that involve
@@
 (whereas \invariantconfluence{} targets \emph{replicated} databases). The
-protocol generates global invariants, called global treaties, and servers can
+protocol generates global invariants, called global treaties, and servers
 operate without coordination so long as they preserve the treaty. When the
@@
 \textbf{Token Based Invariant Confluence.}
-In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token-based
+In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
 consistency model that generalizes a family of consistency models including
@@
 execute. This allows the application designer to specify which transactions
-conflict with each other.
-%
-% With no tokens, the model is equivalent to causal consistency. With all
-% transactions acquiring a single self-conflicting token, the model is equivalent
-% to sequential consistency~\cite{lamport1979make}. With some transactions
-% acquiring a single self-conflicting token and some transactions acquiring no
-% tokens, the model is equivalent to RedBlue consistency.
-%
-Gotsman et al.\ develop sufficient conditions to determine whether a given
-token scheme is sufficient to guarantee that a global invariant is never
-broken. This model of consistency is very similar to segmented
-\invariantconfluence{}. The token-based approach allows users to specify certain
-conflicts that are not possible with segmented \invariantconfluence{}. A
-segmentation only allows transactions within a segment to acquire a single
-self-conflicting lock. However, segmented \invariantconfluence{} also introduces
-the notion of invariant segmentation, which cannot be emulated with the token
-based approach. For example, it is difficult to emulate escrow transactions
-with the token-based approach.
+conflict with one another.  Gotsman et al.\ develop sufficient conditions to
+determine whether a given token scheme is sufficient to guarantee that a global
+invariant is never broken. This model of consistency is very similar to
+segmented \invariantconfluence{}. The token based approach allows users to
+specify certain conflicts that are not possible with segmented
+\invariantconfluence{}. A segmentation only allows transactions within a
+segment to acquire a single self-conflicting lock. However, segmented
+\invariantconfluence{} also introduces the notion of invariant segmentation,
+which cannot be emulated with the token based approach. For example, it is
+difficult to emulate escrow transactions with the token based approach.
 
@@
 \invariantconfluence{} system model are inspired directly by the corresponding
-definitions and system models in~\cite{shapiro2011conflict}. Note that an
+definitions and system models in~\cite{shapiro2011conflict}. Also, an
 \invariantconfluent{} object does not necessarily have to be eventually
@@
 a program can be written in the monotone fragment of these languages, then
-there exists a coordination-free implementation of the program.
-% Moreover, monotone programs are invariant to message reordering, whereas
-% non-monotone programs may experience anomalies if messages are reordered in the
-% network.
-The CALM theorem provides guarantees about the outputs of a program, but it
-does not guarantee that invariants are maintained throughout the duration of an
+there exists a coordination-free implementation of the program.  The CALM
+theorem provides guarantees about the outputs of a program, but it does not
+guarantee that invariants are maintained throughout the duration of an
 execution. Moreover, Bloom and Dedalus are general purpose programming
@@
 the performance benefits gained from understanding program semantics.
-
-% \textbf{Proof Assistants.}
-% Our interactive decision procedure bears resemblance to interactive proof
-% assistants like Coq~\cite{coq2017}, Isabelle~\cite{nipkow2002isabelle},
-% Agda~\cite{norell2008dependently}, and Idris~\cite{brady2013idris}. Proof
-% assistants guide users through the process of interactively provided a proof
-% that witnesses the validity of a formula. Our interactive decision procedure
-% guides users through the process of interactively generating a set of reachable
-% and unreachable points that witness the \invariantconfluence{} of an object. An
-% interesting avenue for future work is to integrate an proof assistant into our
-% interactive decision procedure. For example, a proof assistant could verify
-% that a state labelled unreachable by a user is actually unreachable.

### 2018-04-22 doc/paper/sections/introduction.tex
Couple more small tweaks.

@@
 called segmented \invariantconfluence{}.} While \invariantconfluence{}
-characterizes objects which can be replicated \emph{without any} coordination,
+characterizes objects that can be replicated \emph{without any} coordination,
 segmented \invariantconfluence{} allows us to replicate

### 2018-04-22 doc/paper/sections/invariant_confluence.tex
Couple more small tweaks.

@@
 
-Second, a server $p_i$ can sent its state $s_i$ to another server $p_j$ with
+Second, a server $p_i$ can send its state $s_i$ to another server $p_j$ with
 state $s_j$ causing $p_j$ to transition from state $s_j$  to state $s_i \join

### 2018-04-25 lucy/benchmark_master_main.cc
Consolated vary_withdraws and vary_nodes.

@@ -16,7 +16,7 @@ namespace {
 std::string Usage() {
   return "./benchmark_master_main <benchmark_server_cluster> "
          "<benchmark_client_cluster> "
-         "<vary_withdraws|vary_segments|vary_nodes|all>";
+         "<bank_account|vary_segments|all>";
 }
 
 std::string ServerTypeToString(ServerType server_type) {
@@ -33,51 +33,64 @@ std::string ServerTypeToString(ServerType server_type) {
   }
 }
 
-void VaryWithdraws(const std::size_t num_servers, BenchmarkMaster *master) {
-  std::ofstream vary_withdraws_file("vary_withdraws.csv");
+void BankAccountWorkload(const std::size_t total_num_servers,
+                         BenchmarkMaster *master) {
+  std::ofstream bank_account_file("bank_account.csv");
 
   for (ServerType server_type : {GOSSIP, SEGMENTED_MASTER, PAXOS}) {
-    std::vector<double> fraction_withdraws = {0,   0.01, 0.02, 0.03, 0.04, 0.05,
-                                              0.1, 0.2,  0.3,  0.4,  0.5,  0.6,
-                                              0.7, 0.8,  0.9,  1.0};
-    for (double fraction_withdraw : fraction_withdraws) {
-      LOG(INFO) << "=====================================================";
-      LOG(INFO) << "server_type       = " << ServerTypeToString(server_type);
-      LOG(INFO) << "fraction_withdraw = " << fraction_withdraw;
+    std::vector<std::uint64_t> num_servers = {2,  3,  4,  5,  6,  7, 8,
+                                              12, 16, 20, 24, 28, 32};
+    for (double num_server : num_servers) {
+      if (num_server > total_num_servers) {
+        continue;
+      }
 
-      // Start the servers.
-      LOG(INFO) << "Starting servers.";
-      BenchmarkServerStartRequest start;
-      start.set_num_servers(num_servers);
-      start.mutable_bank_account();
-      start.set_server_type(server_type);
-      master->ServersStart(start);
+      std::vector<double> fraction_withdraws = {
+          0,   0.01, 0.02, 0.03, 0.04, 0.05, 0.1, 0.2,
+          0.3, 0.4,  0.5,  0.6,  0.7,  0.8,  0.9, 1.0};
+      for (double fraction_withdraw : fraction_withdraws) {
+        LOG(INFO) << "=====================================================";
[…]

### 2018-04-25 benchmark/plot.py
Updated plot for new experiments.

@@ -28,77 +28,98 @@ LIN_FMT = {
 def load_csv(filename):
     with open(filename, 'r') as f:
         reader = csv.reader(f, delimiter=',')
-        return [(row[0], float(row[1]), float(row[2])) for row in reader]
+        return [(row[0], float(row[1]), float(row[2]), float(row[3]))
+                for row in reader]
 
-def split_data(data):
-    splits = defaultdict(lambda: ([], []))
-    for (server_type, x, y) in data:
-        splits[server_type][0].append(x)
-        splits[server_type][1].append(y)
-    return splits
+def extract_bank_account_data(data):
+    records = []
+    for (server_type, num_servers, fraction_withdraw, throughput) in data:
+        records.append({
+            'server_type': server_type,
+            'num_servers': num_servers,
+            'fraction_withdraw': fraction_withdraw,
+            'throughput': throughput,
+        })
+    return records
 
-def plot_vary_withdraws(data):
-    splits = split_data(data)
+def partition_by_server_type(records):
+    gossip = []
+    segmented_master = []
+    paxos = []
+    for r in records:
+        if r['server_type'] == 'gossip':
+            gossip.append(r)
+        elif r['server_type'] == 'segmented_master':
+            segmented_master.append(r)
+        elif r['server_type'] == 'paxos':
+            paxos.append(r)
+        else:
+            raise ValueError(r['server_type'])
+    return gossip, segmented_master, paxos
 
-    gossip_x, gossip_y = splits['gossip']
-    segmented_x, segmented_y = splits['segmented_master']
-    paxos_x, paxos_y = splits['paxos']
+def plot_bank_account(data):
+    records = extract_bank_account_data(data)
+    num_servers_domain = sorted(set(r['num_servers'] for r in records))
+    frac_withdraw_domain = sorted(set(r['fraction_withdraw'] for r in records))
 
-    plt.figure(figsize=(8, 3))
-    plt.semilogy(gossip_x, gossip_y, **EVENTUAL_FMT)
-    plt.semilogy(segmented_x, segmented_y, **IC_FMT)
-    plt.semilogy(paxos_x, paxos_y, **LIN_FMT)
[…]

### 2018-04-25 benchmark/plot.py
Updated plot for multi plot.

@@ -87,7 +87,7 @@ def plot_bank_account(data):
         plt.savefig(f'throughput_vs_fraction_{n}.pdf', bbox_inches='tight')
         plt.close()
 
-    # First, we plot the throughput (y-axis) vs number of nodes (x-axis) for
+    # Next, we plot the throughput (y-axis) vs number of nodes (x-axis) for
     # each fraction of withdraws.
     def filter_fraction_sort_num_servers(records, f):
         filtered = [r for r in records if r['fraction_withdraw'] == f]
@@ -111,6 +111,35 @@ def plot_bank_account(data):
         plt.ylabel('Throughput (txns/s)')
         plt.savefig(f'throughput_vs_num_nodes_{f}.pdf', bbox_inches='tight')
         plt.close()
+
+    # Finally, we plot the throughput (y-axis) vs number of nodes (x-axis) for
+    # a couple different fraction of withdraws on the same plot.
+    gossip, master, paxos = partition_by_server_type(records)
+    gossip_xs, gossip_ys = filter_fraction_sort_num_servers(gossip, 0)
+    paxos_xs, paxos_ys = filter_fraction_sort_num_servers(paxos, 0)
+    fractions = [0.01, 0.05, 0.2, 0.5]
+    master_data = {f: filter_fraction_sort_num_servers(master, f)
+                   for f in fractions}
+    markers = {
+        0.01: 's',
+        0.05: 'p',
+        0.2: '*',
+        0.5: 'x',
+    }
+
+    plt.figure(figsize=(8, 4))
+    plt.semilogy(gossip_xs, gossip_ys, **EVENTUAL_FMT)
+    plt.semilogy(paxos_xs, paxos_ys, **LIN_FMT)
+    for f, (xs, ys) in master_data.items():
+        fmt = dict(IC_FMT.items())
+        del fmt['label']
+        del fmt['marker']
+        plt.semilogy(xs, ys, marker=markers[f], label=f'segmented ({f})', **fmt)
+    plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.25), prop={'size': 12})
+    plt.xlabel('Number of Nodes')
+    plt.ylabel('Throughput (txns/s)')
+    plt.savefig(f'throughput_vs_num_nodes_multi.pdf', bbox_inches='tight')
+    plt.close()
 
 def main():
     if len(sys.argv) != 2:

### 2018-04-25 doc/paper/sections/evaluation.tex
Updated benchmarks.

@@
 \python{checker.check()}, the interactive decision procedure produces a
-counterexample in less than a tenth of a second.  After we label the
-counterexample and refine the invariant with $y \leq 0$, the interactive
-decision procedure determines that the object is \invariantconfluent{}, again,
-in less than a tenth of a second. Note that the object is \invariantconfluent{}
-but \emph{not} \invariantclosed{}, so prior work that relies on
-\invariantclosure{} alone to determine \invariantconfluence{} would not be able
-to identify this example as \invariantconfluent{}.
+counterexample $s_1 = (0, 1), s_2 = (1, 0)$ in less than a tenth of a second
+and automatically recognizes that $s_2$ is reachable. After we label $s_1$ as
+unreachable and refine the invariant with $y \leq 0$, the interactive decision
+procedure determines that the object is \invariantconfluent{}, again, in less
+than a tenth of a second. Note that the object is \invariantconfluent{} but
+\emph{not} \invariantclosed{}, so prior work~\cite{li2012making,
+li2014automating, balegas2015towards, gotsman2016cause} that relies on
+\invariantclosure{}---or another equivalent sufficient condition---to determine
+\invariantconfluence{} would not be able to identify this example as
+\invariantconfluent{}.
 
@@
 Now, we evaluate the performance of replicating an object with segmented
-\invariantconfluence{} as compared to replicating it with eventual consistency
-or linearizability.
+\invariantconfluence{} as compared to the performance of replicating it with
+eventual consistency or linearizability.
+%
+There are two hypotheses about the performance of segmented
+\invariantconfluent{} replication that we aim to confirm.
+%
+First, segmented \invariantconfluent{} replication provides higher throughput
+and better scalability than linearizabile replication for workloads which
+require little coordination (i.e.\ low-coordination workloads).
+%
+Second, the throughput and scalability of segmented \invariantconfluent{}
+replication decreases as we increase the fraction of transactions which require
+coordination.
 
-\begin{figure}[ht]
+These hypotheses state that segmented \invariantconfluent{} is more performant
+than linearizabile replication for low-coordination workloads. But by how much?
+We also aim to measure the absolute performance and scalability benefits of
+segmented \invariantconfluent{} replication and how they vary as we vary the
+coordination required by a workload. We perform two controlled microbenchmarks
+to confirm our hypotheses and discover the absolute performance benefits. The
+workloads themselves are trivial, but also not salient. Our objective is to
+obtain a controlled measure of throughput and scalability as we vary workload
+contention.
+
+\begin{figure}[t]
   \centering
-
-  \begin{subfigure}[c]{\columnwidth}
-    \centering
-    \includegraphics[width=\columnwidth]{figures/vary_withdraws.pdf}
-  \end{subfigure}
-  \begin{subfigure}[c]{\columnwidth}
-    \includegraphics[width=\columnwidth]{figures/vary_segments.pdf}
-  \end{subfigure}
-
+  \includegraphics[width=\columnwidth]{figures/throughput_vs_fraction_16.pdf}
   \caption{%
-    Segmented \invariantconfluent{} replication throughput versus coordination,
-    induced by executing disallowed transactions (top) and by transitioning
-    across segments (bottom).
+    Segmented \invariantconfluent{} replication throughput versus coordination
+    induced by executing disallowed decrement transactions.
   }\figlabel{ThroughputVsGlobalSyncs}
@@
   corresponding transactions, invariants, and single-segment segmentation that
-  forbids concurrent decrements. We replicate this object on 32 servers
-  deployed on 32 m5.xlarge EC2 instances within the same availability zone.
+  forbids concurrent decrements. We replicate this object on 16 servers
+  deployed on 16 m5.xlarge EC2 instances within the same availability zone.
   Each server has three colocated clients that issue increment and decrement
@@
   throughput as we vary the fraction of client requests that are decrements.
-  The results are shown in the top of \figref{ThroughputVsGlobalSyncs}.
+  The results are shown in \figref{ThroughputVsGlobalSyncs}.
 
   Both eventually consistent replication and linearizable replication are
-  unaffected by the workload, achieving roughly 700,000 and 7,000 transactions
+  unaffected by the workload, achieving roughly 375,000 and 12,000 transactions
   per second respectively. Segmented \invariantconfluent{} replication performs
-  well for low-decrement workloads and performs increasingly poorly as we
-  increase the fraction of decrement transactions, eventually performing worse
-  than linearizable replication. For example, with 5\% decrement transactions,
-  segmented \invariantconfluent{} replication performs an order of magnitude
-  better than linearizable replication; with 50\% decrements, it performs as
-  well; and with 100\% decrements, it performs two times worse.
+  well for low-decrement (i.e.\ low-coordination) workloads and performs
+  increasingly poorly as we increase the fraction of decrement transactions,
+  eventually performing worse than linearizable replication. For example, with
+  5\% decrement transactions, segmented \invariantconfluent{} replication
+  performs over an order of magnitude better than linearizable replication;
+  with 50\% decrements, it performs as well; and with 100\% decrements, it
+  performs two times worse.
 
-  These results are expected. Increment transactions can execute without any
-  coordination while decrement transactions require global coordination. As we
-  increase the fraction of decrements, we increase the amount of coordination
-  that the system has to perform which in turn drastically decreases the
-  throughput. These results also offer two insights.
+  % These results are expected. Increment transactions can execute without any
+  % coordination while decrement transactions require global coordination. As we
+  % increase the fraction of decrements, we increase the amount of coordination
+  % that the system has to perform which in turn drastically decreases the
+  % throughput. For high contention workloads, segmented \invariantconfluent{} replication performs worse than linearizable replication
+
+  These results offer two insights.
   %
-  First, for low-decrement workloads, segmented \invariantconfluent{}
-  replication achieves a compromise between strong and weak consistency. It
-  guarantees that invariants are maintained (which is impossible with eventual
-  consistency if the object is not \invariantconfluent{}) with performance many
-  times better than strongly consistent replication.
-  %
-  Second, segmented \invariantconfluent{} replication is poorly suited to
-  workloads that require a large amount of coordination. For workloads without
-  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
-  invariants is best done with strong consistency. It provides stronger
-  guarantees with better performance.
+  First, the relationship between segmented \invariantconfluent{} and
+  linearizable replication is analogous to the relationship between pessimistic
+  and optimistic concurrency control protocols with coordination playing the
+  role of contention. Linearizable replication pessimistically assumes that
+  concurrently executing \emph{any} pair of transactions will lead to an
+  invariant violation. Thus, clients send transactions directly to a leader to
+  be linearized. Conversely, segmented \invariantconfluent{} replication
+  optimistically attempts to perform every transaction locally and without
+  coordination. A server only initiates a round of coordination if it is found
+  to be necessary. As a consequence, segmented \invariantconfluent{}
+  replication can offer substantial performance benefits over linearizable
+  replication for low-coordination workloads, but is inferior for medium to
+  high contention workloads.
+
+  Second, throughput does not decrease linearly with the amount of
+  coordination. Even infrequent coordination can drastically decrease
+  throughput. Increasing the fraction of decrements from 0\% to 1\% decreases
+  throughput by a factor of 2. Increasing again to 3\%, the throughput
+  decreases by another factor of 2. With 90\% coordination-free transactions
+  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventual
+  consistency.
+
+%  %
+%  First, for low-decrement workloads, segmented \invariantconfluent{}
+%  replication achieves a compromise between strong and weak consistency. It
+%  guarantees that invariants are maintained (which is impossible with eventual
+%  consistency if the object is not \invariantconfluent{}) with performance many
+%  times better than strongly consistent replication.
+%  %
+%  Second, segmented \invariantconfluent{} replication is poorly suited to
+%  workloads that require a large amount of coordination. For workloads without
+%  much inherit concurrency (e.g.\ decrement-mostly workloads), maintaining
+%  invariants is best done with strong consistency. It provides stronger
+%  guarantees with better performance.
 \end{benchmark}
 
-\begin{benchmark}\benchlabel{VarySegmentLength}
-  Consider again the object, transactions, and invariants from \exampleref{Z2}
-  and \exampleref{SegmentedZ2}. As with \benchref{VaryWithdraws}, we replicate
-  the object across 32 servers. Clients issue 50\% increment $x$ transactions,
-  and 50\% decrement $y$ transactions. We consider a ``checkerboard''
-  segmentation $\Sigma_n = \setst{(I_{i, j}, T)}{i, j \in \ints}$ where segment
-  invariant $I_{i, j}$ consists of the square of points $\setst{(x, y)}{ni \leq
-  x < n(i + 1), nj \leq y < n(j + 1)}$ with side length $n$. For example,
-  $\Sigma_1$ places each point in its own segment, $\Sigma_2$ tessellates
-  $\ints^2$ with 2x2 squares, $\Sigma_3$ tessellates $\ints^2$ with 3x3
-  squares, and so on. We measure the throughput of the object replicated with
-  eventually consistent, segmented \invariantconfluent{}, and linearizable
-  replication as we vary the segment side length $n$. The results are shown in
-  the bottom \figref{ThroughputVsGlobalSyncs}.
+% \begin{benchmark}\benchlabel{VarySegmentLength}
+%   Consider again the object, transactions, and invariants from \exampleref{Z2}
+%   and \exampleref{SegmentedZ2}. As with \benchref{VaryWithdraws}, we replicate
+%   the object across 32 servers. Clients issue 50\% increment $x$ transactions,
+%   and 50\% decrement $y$ transactions. We consider a ``checkerboard''
+%   segmentation $\Sigma_n = \setst{(I_{i, j}, T)}{i, j \in \ints}$ where segment
+%   invariant $I_{i, j}$ consists of the square of points $\setst{(x, y)}{ni \leq
+%   x < n(i + 1), nj \leq y < n(j + 1)}$ with side length $n$. For example,
+%   $\Sigma_1$ places each point in its own segment, $\Sigma_2$ tessellates
+%   $\ints^2$ with 2x2 squares, $\Sigma_3$ tessellates $\ints^2$ with 3x3
+%   squares, and so on. We measure the throughput of the object replicated with
+%   eventually consistent, segmented \invariantconfluent{}, and linearizable
+%   replication as we vary the segment side length $n$. The results are shown in
+%   the bottom \figref{ThroughputVsGlobalSyncs}.
+%
+%   This benchmark tells the same tale as \benchref{VaryWithdraws}. Eventual
+%   consistency and linearizability are unaffected by workload, and eventual
+%   consistency outperforms linearizability by roughly two orders of magnitude.
+%   In this example, the segmented \invariantconfluent{} replication only
+%   requires coordination when transitioning between segment boundaries, so as we
+%   increase the segment side length, the throughput of the system increases
+%   significantly.
+% \end{benchmark}
 
-  This benchmark tells the same tale as \benchref{VaryWithdraws}. Eventual
-  consistency and linearizability are unaffected by workload, and eventual
-  consistency outperforms linearizability by roughly two orders of magnitude.
-  In this example, the segmented \invariantconfluent{} replication only
-  requires coordination when transitioning between segment boundaries, so as we
-  increase the segment side length, the throughput of the system increases
-  significantly.
-\end{benchmark}
-
-\begin{figure}[ht]
+\begin{figure}[t]
   \centering
-  \includegraphics[width=\columnwidth]{figures/vary_nodes.pdf}
+  \includegraphics[width=\columnwidth]{figures/throughput_vs_num_nodes_multi.pdf}
   \caption{%
     Throughput of eventually consistent, segmented \invariantconfluent{}, and
-    linearizable replication measured against the number of
-    nodes.
+    linearizable replication measured against the number of nodes for workloads
+    with varying fractions of decrement transactions. For example, the
+    ``segmented (0.2)'' line measures the performance of segmented
+    \invariantconfluent{} replication with 20\% decrement transactions.
+    Eventually consistent replication and linearizable replication are not
+    affected by workload.
   }\figlabel{VaryNodes}
@@
   In this benchmark, we measure the scale-out of segmented
-  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} with a
-  10\% decrement rate, but this time we vary the number of servers we use to
-  replicate our object.  When we replicate with $n$ servers, we use $3n$
-  clients (the $3$ colocated clients on each server) as part of the workload.
-  The results are shown in \figref{VaryNodes}.
+  \invariantconfluent{} replication. We repeat \benchref{VaryWithdraws} while
+  we vary the number of servers that we use to replicate our object. When we
+  replicate with $n$ servers, we use $3n$ clients (the $3$ colocated clients on
+  each server) as part of the workload. The results are shown in
+  \figref{VaryNodes}.
 
@@
   confirming the results in~\cite{bailis2014coordination}. Linearizable
-  replication, on the other hand, scales up to about 3--5 servers before
-  performance begins to decrease. These numbers are consistent with typical
-  deployments of state-machine replication protocols like
-  Paxos~\cite{chandra2007paxos}. Segmented \invariantconfluent{} replication
-  scales up to about 6--8 servers before succumbing to the same scalability
-  bottlenecks as linearizable replication.
+  replication, on the other hand, scales up to about 3 servers before
+  performance begins to decrease. Segmented \invariantconfluent{} replication
+  scales well for low-coordination workloads and poorly for high-coordination
+  workloads. For 1\%, 5\%, 20\%, and 50\% decrement transactions, segmented
+  \invariantconfluent{} replication scales up to 24, 12, 4, and 1 server
+  respectively.
 
-  These results highlight the importance of coordination avoidance in
-  distributed databases. While segmented \invariantconfluent{} replication
-  scales out slightly better than linearizable replication, both scale
-  significantly worse than eventually consistent replication even for a very
-  low (i.e.\ 10\%) decrement workload. This demonstrates that even a small
-  amount of coordination can significantly reduce the scalability of a system.
+  These results echo the results of \benchref{VaryWithdraws}. For
+  low-coordination workloads, segmented \invariantconfluent{} replication can
+  offer almost an order of magnitude better scalability of linearizable
+  replication, but coordination decreases scalability superlinearly. Even
+  infrequent coordination can drastically reduce the scalability of segmented
+  \invariantconfluent{} replication with segmented \invariantconfluent{}
+  replication ultimately scaling worse than linearizable replication for
+  high-coordination workloads.
 \end{benchmark}

### 2018-04-25 doc/paper/sections/related_work.tex
Updated related work.

@@
 %   - Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
+In this section, we detail four alternate approaches to maintaining invariants
+without coordination: RedBlue consistency and Sieve, the homeostasis protocol,
+explicit consistency, and token based invariant confluence. We then survey
+other bodies of related work.
 
@@
 %
-Invariant safety is a sufficient (but not necessary) condition for RedBlue
-consistent objects to be \invariantconfluent{} and is an analog of
-\invariantclosure{} for an operation-based system model (as opposed to the
-state based system model that we use).  In~\cite{li2014automating}, Li et al.\
-develop sophisticated techniques for deciding invariant safety that involve
-calculating weakest preconditions. These techniques are complementary to our
-work and can be used to improve the \invariantclosure{} subroutine used by our
-decision procedures.
+In~\cite{li2012making}, Li et al.\ also introduce invariant safety as a
+sufficient (but not necessary) condition for RedBlue consistent objects to be
+\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{} for
+an operation-based system model (as opposed to the state based system model
+that we use). Li et al.\ develop sophisticated techniques for deciding
+invariant safety that involve calculating weakest preconditions. These
+techniques are complementary to our work and can be used to improve the
+\invariantclosure{} subroutine used by our decision procedures.
 
@@
 (whereas \invariantconfluence{} targets \emph{replicated} databases).  The
-protocol generates global invariants, called global treaties, and servers
-operate without coordination so long as they preserve the treaty. When the
-treaty is violated, the nodes perform a round of coordination and establish a
-new treaty. The protocol guarantees that transactions are executed with
-observational equivalence with respect to some serial execution of the
-transactions. This means that intermediate states may be inconsistent, but
-externally observable side effects and the final database state are consistent.
-The observational equivalence guaranteed by the homeostasis is stronger than
-the guarantees of \invariantconfluence{}. As a result, there are invariants and
+protocol guarantees that transactions are executed with observational
+equivalence with respect to some serial execution of the transactions. This
+means that intermediate states may be inconsistent, but externally observable
+side effects and the final database state are consistent.  The observational
+equivalence guaranteed by the homeostasis protocol is stronger than the
+guarantees of \invariantconfluence{}. As a result, there are invariants and
 workloads that the homeostasis protocol would execute with more coordination
-than a segmented \invariantconfluent{} execution. Still, the homeostasis
-protocol's mechanism of establishing invariants and operating without
-coordination so long as the invariants are maintained is very similar to our
-design of segmented \invariantconfluence{}.
+than a segmented \invariantconfluent{} execution.  The protocol generates
+global invariants, called global treaties, and servers operate without
+coordination so long as they preserve the treaty. When the treaty is violated,
+the nodes perform a round of coordination and establish a new treaty.  This
+mechanism of establishing invariants and operating without coordination so long
+as the invariants are maintained is very similar to our design of segmented
+\invariantconfluence{}.
 
@@
 transactions can be concurrently executed on the same start state without
-violating the invariant~\cite{balegas2015towards}.  This is a sufficient
-condition for explicit consistency similar to criterion (3) in
-\thmref{LatticeProperty}. Balegas et al.\ also describe a variety of
-techniques---like conflict resolution, locking, and escrow
+violating the invariant~\cite{balegas2015towards}. Balegas et al.\ argue that
+this is a sufficient condition for explicit consistency. It is similar to
+criterion (3) in \thmref{LatticeProperty}. Balegas et al.\ also describe a
+variety of techniques---like conflict resolution, locking, and escrow
 transactions~\cite{o1986escrow}---that can be used to replicate workloads that

### 2018-04-26 iconfluence/typecheck.py
Added foralls.

@@ -210,6 +210,38 @@ def _typecheck_expr(e: ast.Expr, env: Dict[str, ast.Type]) -> ast.Expr:
         assert_type_eq(b_typ, ast.TBool(), e.a, e)
         assert_type_uniform(if_typ, else_typ, e.b, e.c, e)
         e.typ = if_typ
+    elif isinstance(e, ast.ESetForall):
+        # Typecheck the set.
+        xs_typ = _typecheck_expr(e.xs, env).typ
+        assert_type_instance(xs_typ, ast.TSet, e.xs, e)
+        xs_typ = cast(ast.TSet, xs_typ)
+
+        # Typecheck the variable.
+        e.x.typ = xs_typ.a
+        x_bound_env = env.copy()
+        x_bound_env[e.x.x] = e.x.typ
+
+        # Typecheck the formula.
+        phi_typ = _typecheck_expr(e.phi, x_bound_env).typ
+        assert_type_eq(phi_typ, ast.TBool(), e.phi, e)
+
+        e.typ = ast.TBool()
+    elif isinstance(e, ast.EMapForallKeys):
+        # Typecheck the map.
+        xs_typ = _typecheck_expr(e.xs, env).typ
+        assert_type_instance(xs_typ, ast.TMap, e.xs, e)
+        xs_typ = cast(ast.TMap, xs_typ)
+
+        # Typecheck the variable.
+        e.x.typ = xs_typ.a
+        x_bound_env = env.copy()
+        x_bound_env[e.x.x] = e.x.typ
+
+        # Typecheck the formula.
+        phi_typ = _typecheck_expr(e.phi, x_bound_env).typ
+        assert_type_eq(phi_typ, ast.TBool(), e.phi, e)
+
+        e.typ = ast.TBool()
     else:
         raise ValueError(f'Unkown expression {e}.')
 

### 2018-04-26 iconfluence/z3_/compile.py

@@ -372,6 +372,22 @@ def compile_expr(e: ast.Expr,
                     lambda a, b, c: z3.Store(a, b, Option.some(c)))
     elif isinstance(e, ast.EIf):
         return map3(e.a, e.b, e.c, z3.If)
+    elif isinstance(e, ast.ESetForall):
+        x_bound_tenv = tenv.copy()
+        x_bound_tenv[e.x.x] = e.x.typ
+        wrapped_phi = e.xs.contains(e.x) >> e.phi
+        wrapped_phi = typecheck.typecheck_expr(wrapped_phi, x_bound_tenv)
+        x_zss, x_ze = compile_expr(e.x, venv, x_bound_tenv, fresh)
+        phi_zss, phi_ze = compile_expr(wrapped_phi, venv, x_bound_tenv, fresh)
+        return x_zss | phi_zss, z3.ForAll(x_ze, phi_ze)
+    elif isinstance(e, ast.EMapForallKeys):
+        x_bound_tenv = tenv.copy()
+        x_bound_tenv[e.x.x] = e.x.typ
+        wrapped_phi = e.xs.contains_key(e.x) >> e.phi
+        wrapped_phi = typecheck.typecheck_expr(wrapped_phi, x_bound_tenv)
+        x_zss, x_ze = compile_expr(e.x, venv, x_bound_tenv, fresh)
+        phi_zss, phi_ze = compile_expr(wrapped_phi, venv, x_bound_tenv, fresh)
+        return x_zss | phi_zss, z3.ForAll(x_ze, phi_ze)
     else:
         raise ValueError(f'Unkown expression {e}.')
 

### 2018-04-26 examples/forall.py
Added foralls.

@@ -1,0 +1,17 @@
+from iconfluence import *
+
+from .parser import get_checker
+
+checker = get_checker()
+xs = checker.set_union('xs', TInt(), EEmptySet(TInt()))
+kvs = checker.map('kvs', TInt(), CIntMax(), EEmptyMap(TInt(), TInt()))
+
+x = EVar('x')
+k = EVar('k')
+checker.add_invariant('x_leq_10', xs.forall(x, x <= 10))
+checker.add_invariant('k_leq_10', kvs.forall_keys(k, k <= 10))
+checker.add_invariant('v_leq_10', kvs.forall_keys(k, kvs[k] <= 10))
+
+for i in range(20):
+    checker.add_transaction(f'x_add_{i}', [xs.assign(xs.union({i}))])
+    checker.add_transaction(f'kvs_add_{i}', [kvs.assign(kvs.set(i, i))])

### 2018-04-26 iconfluence/ast.py

@@ -296,6 +296,9 @@ class Expr(AstNode):
     def contains(self, x: Coercible) -> 'Expr':
         return ESetContains(self, coerce(x))
 
+    def forall(self, x: 'EVar', phi: Coercible) -> 'Expr':
+        return ESetForall(self, x, coerce(phi))
+
     def contains_key(self, x: Coercible) -> 'Expr':
         return EMapContainsKey(self, coerce(x))
 
@@ -304,6 +307,9 @@ class Expr(AstNode):
 
     def set(self, k: Coercible, v: Coercible) -> 'Expr':
         return EMapSet(self, k, v)
+
+    def forall_keys(self, x: 'EVar', phi: Coercible) -> 'Expr':
+        return EMapForallKeys(self, x, coerce(phi))
 
     def eq(self, lhs: Coercible) -> 'Expr':
         return EEq(self, coerce(lhs))
@@ -540,6 +546,20 @@ class EMapSet(ETernaryOp):
 class EIf(ETernaryOp):
     def __str__(self) -> str:
         return f'(if ({self.a}) then {self.b} else {self.c})'
+
+class EForallOp(Expr):
+    def __init__(self, xs: Coercible, x: EVar, phi: Coercible) -> None:
+        self.xs = coerce(xs)
+        self.x = x
+        self.phi = coerce(phi)
+
+class ESetForall(EForallOp):
+    def __str__(self) -> str:
+        return f'(forall {self.x} in {self.xs}. {self.phi})'
+
+class EMapForallKeys(EForallOp):
+    def __str__(self) -> str:
+        return f'(forall {self.x} in {self.xs}.keys(). {self.phi})'
 
 # Statements ###################################################################
 class Stmt(AstNode):

### 2018-04-26 iconfluence/diamond/diamond_checker.py

@@ -68,6 +68,11 @@ class DiamondChecker(Checker):
             return all(self._var_free_expr(sube) for sube in [e.lhs, e.rhs])
         elif isinstance(e, ast.ETernaryOp):
             return all(self._var_free_expr(sube) for sube in [e.a, e.b, e.c])
+        elif isinstance(e, ast.EForallOp):
+            # Foralls are a bit trickier to handle. For a formula like forall
+            # x. phi(x), we have to make sure that phi doesn't have any
+            # variables, but of course it can contain x.
+            raise NotImplementedError()
         else:
             raise ValueError(f'Unkown expression {e}.')
 

### 2018-04-26 iconfluence/eval.py

@@ -116,6 +116,21 @@ def eval_expr(e: ast.Expr, env: ValEnv) -> Any:
             return eval_expr(e.b, env)
         else:
             return eval_expr(e.c, env)
+    elif isinstance(e, ast.EForallOp):
+        xs = eval_expr(e.xs, env)
+        if isinstance(e, ast.ESetForall):
+            iter = xs
+        elif isinstance(e, ast.EMapForallKeys):
+            iter = xs.keys()
+        else:
+            raise ValueError(f'Unkown expression {e}.')
+
+        for x in iter:
+            x_bound_env = env.copy()
+            x_bound_env[e.x.x] = x
+            if not eval_expr(e.phi, x_bound_env):
+                return False
+        return True
     else:
         raise ValueError(f'Unkown expression {e}.')
 

### 2018-04-26 iconfluence/eval_test.py

@@ -10,6 +10,7 @@ class TestEval(unittest.TestCase):
     def test_eval_expr(self) -> None:
         venv: ValEnv = {'x': 42}
         tenv: TypeEnv = {'x': TInt()}
+        y: EVar = EVar('y')
         tests: List[Tuple[Expr, Any]] = [
             (EVar('x'), 42),
             (coerce(1), 1),
@@ -52,6 +53,10 @@ class TestEval(unittest.TestCase):
             (coerce(1) > 1, False),
             (EMapSet({1: True}, 1, False), {1: False}),
             (EIf(True, 1, 2), 1),
+            (coerce({1, 2, 3}).forall(y, y <= 10), True),
+            (coerce({1, 2, 3}).forall(y, y <= 2), False),
+            (coerce({2: 2, 3: 6}).forall_keys(y, y <= 10), True),
+            (coerce({2: 2, 3: 6}).forall_keys(y, y <= 2), False),
         ]
         for e, expected in tests:
             e = typecheck_expr(e, tenv)

### 2018-04-26 iconfluence/guess_and_check/guess_and_check_checker.py
Propogated verbosity to make tests pass again.

@@ -14,7 +14,7 @@ class GuessAndCheckChecker(checker.Checker):
     doesn't satisfy the invariant, then it gives up and returns UNKNOWN.
     """
     def __init__(self, num_states: int = 100, verbose: bool = False) -> None:
-        checker.Checker.__init__(self)
+        checker.Checker.__init__(self, verbose)
         self.num_states = num_states
         self.verbose = verbose
         self.state_explorer = StateExplorer(self.crdt_env, self.s0_vals,

### 2018-04-26 iconfluence/interactive/interactive_checker.py

@@ -64,7 +64,7 @@ class InteractiveChecker(Checker):
                  num_explored_states_per_step: int = 100,
                  verbose: bool = False) \
                  -> None:
-        Checker.__init__(self)
+        Checker.__init__(self, verbose)
 
         self.verbose = verbose
         self.solver = z3.Solver()

### 2018-04-26 iconfluence/stratified/arbitrary_start_interactive_checker.py

@@ -45,7 +45,7 @@ class ArbitraryStartInteractiveChecker(ArbitraryStartChecker):
     TODO(mwhittaker): Document.
     """
     def __init__(self, verbose: bool = False) -> None:
-        ArbitraryStartChecker.__init__(self)
+        ArbitraryStartChecker.__init__(self, verbose)
 
         self.verbose = verbose
         self.solver = z3.Solver()

### 2018-04-26 iconfluence/checker.py
Propogated verbosity to make tests pass again.

@@ -14,13 +14,14 @@ class Decision(Enum):
     UNKNOWN = "unknown"
 
 class Checker:
-    def __init__(self):
+    def __init__(self, verbose: bool = False):
         self.crdt_env: CrdtEnv = dict()
         self.type_env: TypeEnv = dict()
         self.s0_exprs: ExprEnv = dict()
         self.s0_vals: ValEnv = dict()
         self.transactions: Dict[str, Transaction] = dict()
         self.invariants: Dict[str, Invariant] = dict()
+        self.verbose = verbose
 
     def __str__(self):
         strings = []
@@ -119,5 +120,6 @@ class Checker:
         start = time.time()
         ret = self._check()
         stop = time.time()
-        print(f'Check took {stop - start} seconds.')
+        if self.verbose:
+            print(f'Check took {stop - start} seconds.')
         return ret

### 2018-04-26 iconfluence/stratified/arbitrary_start_checker.py

@@ -15,6 +15,9 @@ class ArbitraryStartChecker(Checker):
     """
     TODO(mwhittaker): Document.
     """
+    def __init__(self, verbose: bool = False) -> None:
+        Checker.__init__(self, verbose)
+
     def __str__(self):
         strings = []
 

### 2018-04-26 iconfluence/stratified/stratified_checker.py

@@ -68,7 +68,7 @@ class Stratum:
 
 class StratifiedChecker(Checker):
     def __init__(self, verbose: bool = True) -> None:
-        Checker.__init__(self)
+        Checker.__init__(self, verbose)
 
         self.verbose = verbose
         self.solver = z3.Solver()

### 2018-04-26 iconfluence/ast.py
Added top CRDT.

@@ -133,6 +133,27 @@ class COption(Crdt):
     def to_type(self) -> Type:
         return TOption(self.a.to_type())
 
+class CTop(Crdt):
+    """
+    CTop is a set with discrete partial order and distinguished top element.
+    For example, CTop[Int] has the following Hasse diagram:
+
+                    top
+               /  /  |  \ \
+        ...  -2  -1  0  1  2  ...
+
+    Thus, the merge of any two unequal elements is top, and the merge of top
+    with anything is top.
+    """
+    def __init__(self, a: Type) -> None:
+        self.a = a
+
+    def __str__(self) -> str:
+        return f'Top[{self.a}]'
+
+    def to_type(self) -> Type:
+        return TOption(self.a)
+
 class CMap(Crdt):
     def __init__(self, a: Type, b: Crdt) -> None:
         self.a = a

### 2018-04-26 iconfluence/diamond/diamond_checker.py

@@ -85,7 +85,8 @@ class DiamondChecker(Checker):
     def _has_bottom(self, crdt: Crdt) -> bool:
         if (isinstance(crdt, ast.CIntMax) or
             isinstance(crdt, ast.CIntMin) or
-            isinstance(crdt, ast.CSetIntersect)):
+            isinstance(crdt, ast.CSetIntersect) or
+            isinstance(crdt, ast.CTop)):
             return False
         elif (isinstance(crdt, ast.CBoolOr) or
             isinstance(crdt, ast.CBoolAnd) or

### 2018-04-26 iconfluence/stratified/stratified_checker.py

@@ -202,6 +202,8 @@ class StratifiedChecker(Checker):
                     checker.map(name, crdt.a, crdt.b, None)
                 elif isinstance(crdt, ast.COption):
                     checker.option(name, crdt.a, None)
+                elif isinstance(crdt, ast.CTop):
+                    checker.top(name, crdt.a, None)
                 else:
                     raise ValueError(f'Unrecognized CRDT {crdt}.')
             for name in stratum.transactions:

### 2018-04-26 examples/all_datatypes.py
Added top CRDT.

@@ -20,4 +20,5 @@ set_union = checker.set_union('set_union', TInt(), {1})
 set_intersect = checker.set_intersect('set_intersect', TInt(), {1})
 map = checker.map('map', TInt(), CIntMax(), {1: 42})
 option = checker.option('option', CIntMax(), ENone(TInt()))
+top = checker.top('top', TInt(), ENone(TInt()))
 checker.add_invariant('dummy', int_max - int_min <= 10)

### 2018-04-26 iconfluence/checker.py

@@ -14,7 +14,7 @@ class Decision(Enum):
     UNKNOWN = "unknown"
 
 class Checker:
-    def __init__(self, verbose: bool = False):
+    def __init__(self, verbose: bool = False) -> None:
         self.crdt_env: CrdtEnv = dict()
         self.type_env: TypeEnv = dict()
         self.s0_exprs: ExprEnv = dict()
@@ -102,6 +102,9 @@ class Checker:
 
     def option(self, name: str, a: Crdt, val: Coercible) -> ast.EVar:
         return self._register_var(name, ast.COption(a), val)
+
+    def top(self, name: str, a: Type, val: Coercible) -> ast.EVar:
+        return self._register_var(name, ast.CTop(a), val)
 
     def add_transaction(self, name: str, txn: Transaction) -> None:
         assert name not in self.transactions, (name, self.transactions)

### 2018-04-26 iconfluence/eval.py

@@ -188,6 +188,13 @@ def _eval_join_vals(lhs: Any, rhs: Any, crdt: ast.Crdt) -> Any:
             return lhs
         else:
             return _eval_join_vals(lhs, rhs, crdt.a)
+    elif isinstance(crdt, ast.CTop):
+        if lhs is None or rhs is None:
+            return None
+        elif lhs == rhs:
+            return lhs
+        else:
+            return None
     else:
         raise ValueError(f'Unrecognized CRDT {crdt}.')
 

### 2018-04-26 iconfluence/stratified/arbitrary_start_checker.py

@@ -92,3 +92,7 @@ class ArbitraryStartChecker(Checker):
     def option(self, name: str, a: Crdt, val: Coercible) -> EVar:
         self._warn_if_not_none(val)
         return self._register_var_no_start_state(name, ast.COption(a))
+
+    def top(self, name: str, a: Type, val: Coercible) -> EVar:
+        self._warn_if_not_none(val)
+        return self._register_var_no_start_state(name, ast.CTop(a))

### 2018-04-26 iconfluence/z3_/compile.py

@@ -520,6 +520,9 @@ def _compile_z3_join(lhs: z3.ExprRef,
                 z3.If(Option.is_none(lhs), rhs,
                 z3.If(Option.is_none(rhs), lhs,
                 Option.some(j_ze))))
+    elif isinstance(crdt, ast.CTop):
+        Option = OptionSort(compile_type(crdt.to_type()))
+        return OrderedSet(), z3.If(lhs == rhs, lhs, Option.none())
     else:
         raise ValueError(f'Unkown CRDT {crdt}.')
 

### 2018-04-26 examples/tpcc02.py
Added TPC 02 check.

@@ -1,0 +1,37 @@
+from iconfluence import *
+
+from .parser import get_checker
+
+checker = get_checker()
+d_next_o_id = checker.int_max('d_next_o_id', 0)
+order_ids = checker.set_union('order_ids', TInt(), ESet({-1}))
+orders = checker.map('orders', TInt(), CTop(TBool()), EEmptyMap(TInt(), TOption(TBool())))
+
+x = EVar('x')
+k = EVar('k')
+checker.add_invariant('sequential',
+    (d_next_o_id - 1).eq(ESetMax(order_ids)))
+checker.add_invariant('order_ids_init',
+    order_ids.contains(-1))
+checker.add_invariant('continuous',
+    order_ids.forall(x,
+        EIf(x.eq(ESetMax(order_ids)), EBool(True), order_ids.contains(x + 1))))
+checker.add_invariant('geq_-1',
+    order_ids.forall(x, x >= -1))
+checker.add_invariant('orders_and_order_ids_1',
+    order_ids.forall(x, EIf(x.ne(-1), orders.contains_key(x), EBool(True))))
+checker.add_invariant('orders_and_order_ids_2',
+    orders.forall_keys(k, order_ids.contains(k)))
+checker.add_invariant('orders_and_order_ids_3',
+    EBoolNot(orders.contains_key(-1)))
+checker.add_invariant('unique',
+    orders.forall_keys(k, orders[k].is_some()))
+checker.add_invariant('simple_example',
+    EBoolNot(order_ids.contains(2)))
+
+for b in [True, False]:
+    checker.add_transaction(f'order_{b}', [
+        orders.assign(orders.set(d_next_o_id, ESome(b))),
+        order_ids.assign(order_ids.union({d_next_o_id})),
+        d_next_o_id.assign(d_next_o_id + 1),
+    ])

### 2018-04-26 iconfluence/interactive/interactive_checker.py

@@ -264,6 +264,13 @@ class InteractiveChecker(Checker):
                 self.rhs_reachable()
             if (self.lhs_label == Label.REACHABLE and
                 self.rhs_label == Label.REACHABLE):
+                m = ('The following states are both reachable and satisfy ' +
+                     'the (refined) invariant, but their join does not. ' +
+                     'Their join is also reachable, so the object is not ' +
+                     'invariant confluent.')
+                self._wrap_print(m)
+                print(f'  lhs = {self.lhs}')
+                print(f'  rhs = {self.rhs}')
                 return Decision.NO
 
             m = ('The following two states (i.e. lhs and rhs) satisfy the ' +
@@ -310,6 +317,12 @@ class InteractiveChecker(Checker):
             # is reachable, then we are not invariant-confluent.
             if (self.lhs_label == Label.REACHABLE and
                 self.rhs_label == Label.REACHABLE):
+                m = ('The following states are both reachable and satisfy ' +
+                     'the (refined) invariant, but their join does not. ' +
+                     'Their join is also reachable, so the object is not ' +
+                     'invariant confluent.')
+                print(f'  lhs = {self.lhs}')
+                print(f'  rhs = {self.rhs}')
                 return Decision.NO
             else:
                 self.lhs = None

### 2018-04-26 doc/paper/sections/evaluation.tex
Added TPC checks to paper.

@@
 categorize the invariants into one of three types:
-\begin{enumerate}
-  \item
+
+% \begin{enumerate}
+%   \item
     Three of the twelve invariants are \textbf{foreign key constraints}.  As
@@
 
-  \item
+  % \item
     \newcommand{\ttt}[1]{{\smaller \texttt{#1}}} Seven of the twelve invariants
@@
 
-  \item
+  % \item
     Two of the twelve invariants involve generating \textbf{sequential and
-    unique identifiers}. This workload is \emph{not} \invariantconfluent{}, but
-    when the sequentiality requirement is dropped, it is \invariantconfluent{}.
-    Unfortunately, modelling unique key generation in our theoretical framework
-    (and in our prototype implementation) is not possible because we have thus
-    far tacitly assumed that transaction execution is deterministic. We leave a
-    generalization of our theory to non-deterministic transactions for future
-    work.
-\end{enumerate}
+    unique identifiers}. Consider, for example, invariant 2 which dictates that
+    a district's next order id \ttt{D\_NEXT\_O\_ID} is equal to the maximum
+    order id \ttt{O\_ID} of orders within the district. The New Order
+    transaction places an order with \ttt{O\_ID} equal to the current value of
+    \ttt{D\_NEXT\_O\_ID} and then increments \ttt{D\_NEXT\_O\_ID}. We model
+    this workload with three objects. \ttt{D\_NEXT\_O\_ID} is modeled as an
+    integer merged by $\max$. The order table is modelled as two objects. The
+    first is a set of all order ids used, merged by union. The second is a map
+    which maps order id to order. The map's merge function performs an id-by-id
+    merge of orders. If two different orders have the same id, the merge
+    function replaces them with a distinguished value to indicate that the
+    primary key constraint has been violated. We applied Lucy to this workload
+    and in less than a second, it produced a counterexample that---when
+    labelled as reachable---proved that the workload was not
+    \invariantconfluent{}.
+% \end{enumerate}
 

### 2018-04-26 doc/paper/sections/evaluation.tex
Shrunk some of the TPC-C description.

@@
     transaction places an order with \ttt{O\_ID} equal to the current value of
-    \ttt{D\_NEXT\_O\_ID} and then increments \ttt{D\_NEXT\_O\_ID}. We model
-    this workload with three objects. \ttt{D\_NEXT\_O\_ID} is modeled as an
-    integer merged by $\max$. The order table is modelled as two objects. The
-    first is a set of all order ids used, merged by union. The second is a map
-    which maps order id to order. The map's merge function performs an id-by-id
-    merge of orders. If two different orders have the same id, the merge
-    function replaces them with a distinguished value to indicate that the
-    primary key constraint has been violated. We applied Lucy to this workload
-    and in less than a second, it produced a counterexample that---when
-    labelled as reachable---proved that the workload was not
-    \invariantconfluent{}.
+    \ttt{D\_NEXT\_O\_ID} and then increments \ttt{D\_NEXT\_O\_ID}.
+    We model this workload with an integer for \ttt{D\_NEXT\_O\_ID} and a map
+    for \ttt{O\_ID} that maps order ids to order.
+    %
+    % We model this workload with three objects. \ttt{D\_NEXT\_O\_ID} is modeled
+    % as an integer merged by $\max$. The order table is modelled as two objects.
+    % The first is a set of all order ids used, merged by union. The second is a
+    % map which maps order id to order. The map's merge function performs an
+    % id-by-id merge of orders. If two different orders have the same id, the
+    % merge function replaces them with a distinguished value to indicate that
+    % the primary key constraint has been violated.
+    %
+    We applied Lucy to this workload and in less than a second, it produced a
+    counterexample that---when labelled as reachable---confirming that the
+    workload is \emph{not} \invariantconfluent{}~\cite{bailis2014coordination}.
 % \end{enumerate}

### 2018-04-30 doc/paper/sections/related_work.tex
Added branch and merge to related works.

@@
 %   - Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
-In this section, we detail four alternate approaches to maintaining invariants
-without coordination: RedBlue consistency and Sieve, the homeostasis protocol,
-explicit consistency, and token based invariant confluence. We then survey
-other bodies of related work.
+% In this section, we detail four alternate approaches to maintaining invariants
+% without coordination: RedBlue consistency and Sieve, the homeostasis and
+% demarcation protocols, explicit consistency, and token based invariant
+% confluence. We then survey other bodies of related work.
 
@@
 
-\textbf{The Homeostasis Protocol.}
-The homeostasis protocol~\cite{roy2015homeostasis} uses program analysis to
+\textbf{The Demarcation and Homeostasis Protocols.}
+The homeostasis protocol~\cite{roy2015homeostasis}, a generalization of the
+demarcation protocol~\cite{barbara1994demarcation}, uses program analysis to
 avoid unnecessary coordination between servers in a \emph{sharded} database
@@
 
+\textbf{Branch and Merge.}
+Bayou~\cite{terry1995managing}, Dynamo~\cite{decandia2007dynamo}, and
+TARDiS~\cite{crooks2016tardis} all take a branch and merge approach to
+maintaining distributed invariants without coordination. With this approach,
+servers execute transactions without any coordination but keep track of the
+causal dependencies between transactions. Periodically, two servers merge
+states and invoke a user defined merge function to reconcile the divergent
+states. This approach does \emph{not} provide any formal guarantees that
+invariants are maintained. Its correctness depends on the correctness of the
+potentially complex user defined merge functions.
+
 \textbf{CRDTs.}

### 2018-04-30 doc/paper/sections/evaluation.tex
Shrinking paper.

@@
 Users specify objects, transactions, invariants, and segmentations in C++. Lucy
-then replicates the objects using segmented \invariantconfluence{} (or
-\invariantconfluence{} if the segmentation contains a single segment without
-any disallowed transactions). Clients send every transaction request to a
-randomly selected server. When a server receives a transaction request, it
-executes \algoref{TxnExecution} to attempt to execute the transaction locally.
-If the transaction requires global coordination, then the server forwards the
-transaction request to a predetermined leader. When the leader receives a
-transaction request, it sends a coordination request to all other servers. When
-a server receives a coordination request from the leader, it stops processing
-transactions and sends the leader its state in a coordination reply. When the
-leader receives coordination replies from all other servers, it executes the
-transaction, and then sends its state to the other servers. When a server
-receives a new state, it adopts the state, computes its new active segment, and
-resumes normal processing. After every 100 transactions processed, a server
-sends a merge request to a randomly selected server.
+then replicates the objects using segmented \invariantconfluence{}. Clients
+send every transaction request to a randomly selected server. When a server
+receives a transaction request, it executes \algoref{TxnExecution} to attempt
+to execute the transaction locally.  If the transaction requires global
+coordination, then the server forwards the transaction request to a
+predetermined leader. When the leader receives a transaction request, it
+broadcasts a coordination request to the other servers. When a server receives
+a coordination request from the leader, it stops processing transactions and
+sends the leader its state. When the leader receives the states of all other
+servers, it executes the transaction, and then sends its state to the other
+servers. When a server receives a new state, it adopts the state, computes its
+new active segment, and resumes normal processing. After every 100 transactions
+processed, a server sends a merge request to a randomly selected server.
 

### 2018-04-30 doc/paper/sections/related_work.tex
Shrinking paper.

@@
 
-\textbf{Program Analysis in Database Systems.}
-Using program analysis to improve the performance of database systems is not
-new. For example, it has been used to improve the performance of
-database-backed web applications~\cite{cheung2014using, wu2016transaction,
-ramachandra2012program} and used to improve the performance of optimistic
-concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
-on \invariantconfluence{} continues the theme of using program analysis to reap
-the performance benefits gained from understanding program semantics.
+% \textbf{Program Analysis in Database Systems.}
+% Using program analysis to improve the performance of database systems is not
+% new. For example, it has been used to improve the performance of
+% database-backed web applications~\cite{cheung2014using, wu2016transaction,
+% ramachandra2012program} and used to improve the performance of optimistic
+% concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
+% on \invariantconfluence{} continues the theme of using program analysis to reap
+% the performance benefits gained from understanding program semantics.

### 2018-04-30 doc/paper/sections/segmented_invariant_confluence.tex
Added a note about segment suggestion.

@@
 There are a couple of things worth noting about segmented
-\invariantconfluence{}, its system model, and its decision procedure. First,
-\invariantconfluence{} is a very coarse-grained property. If an object is
-\invariantconfluent{}, then we can replicate it with no coordination. If it is
-not \invariantconfluent{}, then we have no guarantees. There's no in-between.
-Segmented \invariantconfluence{}, on the other hand, is a much more
-fine-grained property that can be applied to applications with varying degrees
-of complexity. Segmented \invariantconfluence{} provides guarantees to complex
-applications that require a large number of segments and to simple applications
-with a smaller number of segments, whereas \invariantconfluence{} only provides
-guarantees to applications that can be segmented into a single segment.
+\invariantconfluence{}, its system model, and its decision procedure.
+%
+First, \invariantconfluence{} is a very coarse-grained property. If an object
+is \invariantconfluent{}, then we can replicate it with no coordination. If it
+is not \invariantconfluent{}, then we have no guarantees. There's no
+in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
+more fine-grained property that can be applied to applications with varying
+degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
+complex applications that require a large number of segments and to simple
+applications with a smaller number of segments, whereas \invariantconfluence{}
+only provides guarantees to applications that can be segmented into a single
+segment.
 
@@
 I$.
+
+Fifth, while a segmented \invariantconfluence{} decision procedure can help
+decide whether or not an object is segmented \invariantconfluent{}, it cannot
+currently help construct the segmentation. In the future, we plan to explore
+ways by which we can automatically suggest segmentations to the application
+designer.

### 2018-04-30 doc/paper/sections/invariant_confluence.tex
Fixed weird example 1.

@@
 \begin{example}\examplelabel{Ints}
-  $O = (\ints, +)$ is a distributed object consisting of integers merged by
-  addition; $t(x) = x + 1$ is a transaction which adds one to a state; and
+  $O = (\ints, \max)$ is a distributed object consisting of integers merged by
+  $\max$; $t(x) = x + 1$ is a transaction which adds one to a state; and
   $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are

### 2018-04-30 doc/paper/sections/introduction.tex
Slight tweaks to intro.

@@
 program that can automatically determine \invariantconfluence{} for us.
-\textbf{The first main thrust of this paper is to make progress towards a
-general-purpose \invariantconfluence{} decision procedure.}
+\textbf{The first main thrust of this paper is to design a general-purpose
+\invariantconfluence{} decision procedure.}
 
@@
   \item
-    We develop an understanding of why \invariantclosure{} is not necessary and
-    use this understanding to develop conditions under which it is both
-    necessary and sufficient (\secref{InvariantClosure}).
+    We develop an understanding of why \invariantclosure{} is not necessary for
+    \invariantconfluence{} and use this understanding to develop conditions
+    under which it is both necessary and sufficient
+    (\secref{InvariantClosure}).
 
@@
     We again exploit these conditions to design a novel non-trivial sufficient
-    condition for \invariantconfluence{}, \mergereducibility, that does not
-    depend on user interaction (\secref{MergeReduction}).
+    condition for \invariantconfluence{}, called \mergereducibility, that does
+    not depend on user interaction (\secref{MergeReduction}).
 

### 2018-04-30 doc/paper/sections/invariant_confluence.tex
Added more of a warmup to section 2.

@@
 modified or merged together~\cite{bailis2014coordination}. In this section, we
-make this informal notion of \invariantconfluence{} precise. We begin by
-introducing our system model of replicated objects and then present a formal
-definition of \invariantconfluence{}.
+make this informal notion of \invariantconfluence{} precise.
+
+We begin by introducing our system model of replicated objects in which a
+distributed object and accompanying invariant is replicated across a set of
+servers. Clients send transactions to servers, and a server executes a
+transaction so long as it maintains the invariant. Servers execute
+transactions without coordination, but to avoid state divergence, servers
+periodically gossip with one another and merge their replicas.
+%
+After we introduce the system model, we present a formal definition of
+\invariantconfluence{}.
 
 \subsection{System Model}
+
 A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states

### 2018-04-30 doc/paper/sections/related_work.tex
Shrunk paper.

@@
 
-\textbf{CRDTs.}
-CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
-semilattices with inflationary update methods. Due to their algebraic
-properties, CRDTs can be replicated with strong eventual consistency without
-the need for any coordination. Our definition of distributed objects and our
-\invariantconfluence{} system model are inspired directly by the corresponding
-definitions and system models in~\cite{shapiro2011conflict}. Also, an
-\invariantconfluent{} object does not necessarily have to be eventually
-consistent, so it is natural (though not necessary) to use CRDTs as
-\invariantconfluent{} distributed objects, achieving a combination of eventual
-consistency and invariant preservation. Our criteria in
-\thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
-algebraic properties of semilattices.
+% \textbf{CRDTs.}
+% CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
+% semilattices with inflationary update methods. Due to their algebraic
+% properties, CRDTs can be replicated with strong eventual consistency without
+% the need for any coordination. Our definition of distributed objects and our
+% \invariantconfluence{} system model are inspired directly by the corresponding
+% definitions and system models in~\cite{shapiro2011conflict}. Also, an
+% \invariantconfluent{} object does not necessarily have to be eventually
+% consistent, so it is natural (though not necessary) to use CRDTs as
+% \invariantconfluent{} distributed objects, achieving a combination of eventual
+% consistency and invariant preservation. Our criteria in
+% \thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
+% algebraic properties of semilattices.
 
@@
 execution. Moreover, Bloom and Dedalus are general purpose programming
-languages that can be used to implement a variety of distributed systems (e.g.
-a sharded database) which are outside of the scope of \invariantconfluence{}.
+languages that can be used to implement a variety of distributed systems which
+are outside of the scope of \invariantconfluence{}.
 

### 2018-05-01 doc/paper/sections/abstract.tex
Couple more tweaks.

@@
 \invariantconfluence{} called segmented \invariantconfluence{}, which allows us
-to replicate non-\invariantconfluent{} objects with a small and tunable amount
-of coordination. We implemented our theoretical findings in a prototype called
+to replicate non-\invariantconfluent{} objects with a small amount of
+coordination. We implemented our theoretical findings in a prototype called
 Lucy and found that our decision procedures efficiently handle workloads

### 2018-05-01 doc/paper/sections/evaluation.tex
Added eval table.
Couple more tweaks.

@@
 simple, yet practical examples. We then discuss how our tool can be used to
-analyze the TPC-C benchmark.
+analyze the TPC-C benchmark. A summary of these results is given in
+\tabref{EvalTable}.
+
+{\input{figures/eval_table.tex}}
 
@@
     We applied Lucy to this workload and in less than a second, it produced a
-    counterexample that---when labelled as reachable---confirming that the
+    counterexample that---when labelled as reachable---confirms that the
     workload is \emph{not} \invariantconfluent{}~\cite{bailis2014coordination}.
@@
   decreases by another factor of 2. With 90\% coordination-free transactions
-  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventual
-  consistency.
+  (i.e.\ 10\% decrements), we achieve only 10\% of the throughput of eventually
+  consistent replication.
 
@@
   low-coordination workloads, segmented \invariantconfluent{} replication can
-  offer almost an order of magnitude better scalability of linearizable
-  replication, but coordination decreases scalability superlinearly. Even
-  infrequent coordination can drastically reduce the scalability of segmented
-  \invariantconfluent{} replication with segmented \invariantconfluent{}
-  replication ultimately scaling worse than linearizable replication for
-  high-coordination workloads.
+  offer almost an order of magnitude better scalability compared to
+  linearizable replication, but coordination decreases scalability
+  superlinearly. Even infrequent coordination can drastically reduce the
+  scalability of segmented \invariantconfluent{} replication with segmented
+  \invariantconfluent{} replication ultimately scaling worse than linearizable
+  replication for high-coordination workloads.
 \end{benchmark}

### 2018-05-01 doc/paper/sections/interactive_decision_procedure.tex
Moved figure around.
Couple more tweaks.

@@
 \newcommand{\Helper}{\textsf{Helper}}
-\newcommand{\IsIclosed}{\textsf{IsIclosed}}
-\newcommand{\IsInvConfluent}{\textsf{IsInvConfluent}}
+\newcommand{\IsIclosed}{\textsc{IsIclosed}}
+\newcommand{\IsInvConfluent}{\textsc{IsInvConfluent}}
 
@@
 \begin{algorithm}[t]
-  \caption{Interactive invariantconfluence{} decision procedure}%
+  \caption{Interactive \invariantconfluence{} decision procedure}%
   \algolabel{InteractiveDecisionProcedure}
@@
 satisfying points are reachable, then \invariantclosure{} and
-invariantconfluence{} are equivalent. In this section, we present the
-interactive invariantconfluence{} decision procedure shown in
+\invariantconfluence{} are equivalent. In this section, we present the
+interactive \invariantconfluence{} decision procedure shown in
 \algoref{InteractiveDecisionProcedure} which takes advantage of this result.
@@
 from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
-the problem of invariantconfluence{} directly to the problem of
+the problem of \invariantconfluence{} directly to the problem of
 \invariantclosure{}, and the algorithm terminates.
@@
 in \figref{DecisionProcedureExample}.
+
+{\input{figures/decision_procedure_example}}
 
@@
 $NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
-iteratively increased, which removes unreachable states from $I$ until $I
-\subseteq \setst{s \in S}{\sTIreachablepredicate{s}}$.
+iteratively increased, which removes unreachable states from $I$ until $I$ is a
+subset of $\setst{s \in S}{\sTIreachablepredicate{s}}$.
 
@@
 
-{\input{figures/decision_procedure_example}}
-
 \subsection{Tolerating User Error}

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
Couple more tweaks.

@@
 \section{Segmented Invariant Confluence}\seclabel{SegmentedInvariantConfluence}
-\newcommand{\IsIclosed}{\textsf{IsIclosed}}
+\newcommand{\IsIclosed}{\textsc{IsIclosed}}
 
@@
   \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
-  \emph{not} \sTISconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
+  \emph{not} \sTIconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
   reachable, and merging these points yields $(42, 42)$ which violates the
@@
 
+\begin{itemize}
+  \item
     If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
     execute $t$. During global coordination, every server temporarily stops
-processing and transitions to state $s = s_1 \join \ldots \join s_n$, the join
-of every server's state. Then, every server speculatively executes $t$
-transitioning to state $t(s)$. If $t(s)$ violates the invariant (i.e.  $\lnot
-I(t(s))$), then every server aborts $t$ and reverts to state $s$.  Then, $p_i$
-replies to the client. If $t(s)$ satisfies the invariant (i.e. $I(t(s))$), then
-every server commits $t$ and remains in state $t(s)$. Every server then chooses
-the first segment $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new
-active segment. Note that such a segment is guaranteed to exist because the
-segment invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server
-will deterministically pick the same active segment. In fact, an invariant of
-the system model is that at any given point of normal processing, every server
-has the same active segment.
+    processing transactions and transitions to state $s = s_1 \join \ldots
+    \join s_n$, the join of every server's state. Then, every server
+    speculatively executes $t$ transitioning to state $t(s)$. If $t(s)$
+    violates the invariant (i.e.  $\lnot I(t(s))$), then every server aborts
+    $t$ and reverts to state $s$.  Then, $p_i$ replies to the client. If $t(s)$
+    satisfies the invariant (i.e.  $I(t(s))$), then every server commits $t$
+    and remains in state $t(s)$. Every server then chooses the first segment
+    $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new active segment.
+    Note that such a segment is guaranteed to exist because the segment
+    invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server will
+    deterministically pick the same active segment. In fact, an invariant of
+    the system model is that at any given point of normal processing, every
+    server has the same active segment.
 
-If $t \in T_i$, then $p_i$ executes $t$ immediately and without coordination.
-If $t(s_i)$ satisfies the \emph{active} invariant (i.e. $I_i(t(s_i))$), then
-$p_i$ commits $t$, stays in state $t(s_i)$, and replies to the client. If
-$t(s_i)$ violates the invariant (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts
-$t$, reverts to state $s_i$, and replies to the client. If $t(s_i)$ satisfies
-the invariant but violates the active invariant (i.e. $I(t(s_i))$ but $\lnot
-I_i(t(s_i))$), then $p_i$ reverts to state $s_i$ and initiates a round of
-global coordination to execute $t$, as described in the previous paragraph.
+  \item
+    If $t \in T_i$, then $p_i$ executes $t$ immediately and without
+    coordination.  If $t(s_i)$ satisfies the \emph{active} invariant (i.e.
+    $I_i(t(s_i))$), then $p_i$ commits $t$, stays in state $t(s_i)$, and
+    replies to the client. If $t(s_i)$ violates the \emph{global} invariant
+    (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts $t$, reverts to state $s_i$,
+    and replies to the client. If $t(s_i)$ satisfies the global invariant but
+    violates the active invariant (i.e.  $I(t(s_i))$ but $\lnot I_i(t(s_i))$),
+    then $p_i$ reverts to state $s_i$ and initiates a round of global
+    coordination to execute $t$, as described in the previous paragraph.
     Transaction execution is summarized in \algoref{TxnExecution}.
+\end{itemize}
 
@@
 invoke \algoref{InteractiveDecisionProcedure} $|I_i|$ times for every segment
-$(I_i, T_I)$, which is clearly infeasible given that $I_i$ can be large or even
+$(I_i, T_i)$, which is clearly infeasible given that $I_i$ can be large or even
 infinite.
@@
   \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
-  triple (closed, $s_1$, $s_2$) where if closed is false, then $s_1, s_2 \in I$
+  triple (closed, $s_1$, $s_2$). If closed is false, then $s_1, s_2 \in I$
   are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
@@
 Third, we can integrate a couple of optimizations into our system model to
-further reduce the amount of coordination required. First, if a server with
+further reduce the amount of coordination it requires. First, if a server with
 state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$

### 2018-05-01 doc/paper/sections/evaluation.tex
took a pass

@@
 \invariantconfluent{}. A concurrent insertion into $X$ and deletion from $Y$
-can lead to a state which violates the invariant. This object is not
+can lead to a state that violates the invariant. This object is not
 \invariantconfluent{} and therefore not \invariantclosed{}. Thus, previous
@@
 The second segment $(\setst{B, w}{w \neq \bot} \cap I, \emptyset)$ includes all
-auctions which have already been closed and forbids all transactions. This
+auctions that have already been closed and forbids all transactions. This
 segmentation captures the intuition that bids should be permitted only when the
@@
 First, segmented \invariantconfluent{} replication provides higher throughput
-and better scalability than linearizabile replication for workloads which
+and better scalability than linearizable replication for workloads that
 require little coordination (i.e.\ low-coordination workloads).
@@
 Second, the throughput and scalability of segmented \invariantconfluent{}
-replication decreases as we increase the fraction of transactions which require
+replication decreases as we increase the fraction of transactions that require
 coordination.
@@
 These hypotheses state that segmented \invariantconfluent{} is more performant
-than linearizabile replication for low-coordination workloads. But by how much?
+than linearizable replication for low-coordination workloads. But by how much?
 We also aim to measure the absolute performance and scalability benefits of
@@
 to confirm our hypotheses and discover the absolute performance benefits. The
-workloads themselves are trivial, but also not salient. Our objective is to
+workloads themselves are trivial, but also not salient. \jmh{Salient to what? I'm confused by the last sentence.} Our objective is to
 obtain a controlled measure of throughput and scalability as we vary workload
@@
   First, the relationship between segmented \invariantconfluent{} and
-  linearizable replication is analogous to the relationship between pessimistic
-  and optimistic concurrency control protocols with coordination playing the
-  role of contention. Linearizable replication pessimistically assumes that
+  linearizable replication is analogous to the relationship between optimistic
+  and pessimistic concurrency control protocols with coordination playing the
+  role of contention \jmh{I didn't understand that last clause about ``coordination playing the role of contention.} Linearizable replication pessimistically assumes that
   concurrently executing \emph{any} pair of transactions will lead to an
@@
   replication for low-coordination workloads, but is inferior for medium to
-  high contention workloads.
+  high contention workloads. \jmh{say why}
 
@@
 \begin{benchmark}
+\jmh{Can you get the figures to color ``eventual'' and ``linearizable'' the same in both?}
   In this benchmark, we measure the scale-out of segmented

### 2018-05-01 doc/paper/sections/interactive_decision_procedure.tex
Couple more small tweaks.
took a pass

@@
 \section{Interactive Decision Procedure}\seclabel{InteractiveDecisionProcedure}
-\newcommand{\Helper}{\textsf{Helper}}
+\newcommand{\Helper}{\textsc{Helper}}
 \newcommand{\IsIclosed}{\textsc{IsIclosed}}
@@
 witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
-but $\lnot I(s_1 \join s_2)$ (e.g. $s_1$ and $s_2$ from \exampleref{Z2}).  As
+but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As
 we mentioned earlier, we can (and do) implement the \invariantclosure{}
@@
 to grow as \Helper{} randomly explores the set of reachable states. As the user
-sees more and more examples of unreachable and reachable states, it becomes
+sees more and more examples of unreachable and reachable states, it often becomes
 easier and easier for them to recognize patterns that define which states are
@@
 %
-We can mitigate this in two ways. First, we can aggressively expand $R$. If a
+We can mitigate this in two ways. First, we can aggressively expand $R$ automatically. If a
 user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the

### 2018-05-01 doc/paper/sections/introduction.tex
took a pass

@@
 make a fundamental choice between weak and strong consistency. Replicating the
-data with strong consistency using a technique like state machine
+data with strong consistency using a technique like distributed transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state machine
 replication~\cite{schneider1990implementing, lamport1998part,
@@
 application designer to reason about the complex interleavings of operations
-that are allowed by these weak consistency models. In fact, weak consistency
+that are allowed by these weak consistency models. In particular, weak consistency
 models strip an application developer of one of the earliest and most effective
 tools that is used to reason about the execution of programs: application
-invariants~\cite{hoare1969axiomatic, balegas2015towards}. Even if every
+invariants~\cite{hoare1969axiomatic, balegas2015towards} and integrity constraints~\cite{IntegrityConstraintCitations}. Even if every
 transaction executing in a weakly consistent system individually maintains an
-application invariant, the system as a whole can produce invariant violating
+application invariant, the system as a whole can produce invariant-violating
 states.
@@
 still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
-introduced the notion of \invariantconfluence{} as a necessary and sufficient
+introduced the notion of \emph{\invariantconfluence{}} as a necessary and sufficient
 condition for when invariants can be maintained over replicated data without
@@
 the performance benefits of weak consistency and (some of) the correctness
-benefits of strong consistency. However, the task of identifying whether or not
-an object actually is \invariantconfluent{} remains a challenge. Bailis et al.\
+benefits of strong consistency.
+
+Unfortunately, to date, the task of identifying whether or not
+an object actually is \invariantconfluent{} has remained an exercise in human proof generation. Bailis et al.\
 manually categorized a set of common objects, transactions, and invariants
 (e.g.\ foreign key constraints on relations, linear constraints on integers) as
-\invariantconfluent{} or not, but ideally we would have a general purpose
+\invariantconfluent{} or not. Hand-written proofs of this sort are unreasonable to expect from programmers. Ideally we would have a general purpose
 program that can automatically determine \invariantconfluence{} for us.
 \textbf{The first main thrust of this paper is to design a general-purpose
-\invariantconfluence{} decision procedure.}
+\invariantconfluence{} decision procedure, and implement it in an interactive system.}
 
-Designing a general-purpose \invariantconfluence{} decision procedure is
+Designing such a general-purpose decision procedure is
 difficult\footnote{Actually, it's impossible. Determining whether an object is
@@
 decision procedure that works well in the common case.}. Many existing
-approaches instead develop a decision procedure for \invariantclosure{}, a
+approaches instead develop a decision procedure for \emph{\invariantclosure{}}, a
 sufficient (but not necessary) condition for
@@
 \invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
-property about reachability, about whether certain values of our replicated
+reachability property, checking whether certain values of our replicated
 object can actually be obtained in the execution of a system. We discover that
@@
 towards the conditions under which \invariantclosure{} and
-\invariantconfluence{} are equivalent. Meanwhile, the decision-procedure
-generates example reachable and unreachable states which help the user
+\invariantconfluence{} are equivalent. Meanwhile, the decision procedure
+generates examples of reachable and unreachable states which help the user
 recognize patterns that describe reachability. We also leverage our intuitions
 about reachability to develop a new sufficient condition for
-\invariantconfluence{}, dubbed \mergereducibility. \Mergereducibility{} covers
+\invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{} covers
 some cases that are not covered by \invariantclosure{}, and it can be
@@
 segmented \invariantconfluence{} allows us to replicate
-non-\invariantconfluent{} objects with \emph{minimal coordination}. The main
+non-\invariantconfluent{} objects with only \emph{occasional coordination}. The main
 idea is that invariant satisfying states are divided into a number of segments,

### 2018-05-01 doc/paper/sections/invariant_closure.tex
Couple more tweaks.
took a pass

@@
 \thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
-sufficient for \invariantconfluence{}. Intuitively, our system model ensures
+sufficient for \invariantconfluence{}. Intuitively, recall that our system model ensures
 that transaction execution preserves the invariant, so if merging states also
@@
   in which we draw each state $(x, y)$ as a point in space. The invariant
-  consists of the second and fourth quadrant, and the reachable states consist
-  of the fourth quadrant. From this, it's immediate that the reachable states
+  consists of the second and fourth quadrant, while the reachable states consist only
+  of the fourth quadrant. From this, it is immediate that the reachable states
   are a subset of the invariant, so $O$ is \invariantconfluent{}. However,
@@
 \invariantclosure{} is completely ignorant of reachability. As a result,
-invariant satisfying yet unreachable states like $s_1$ are the key hurdle
+invariant-satisfying yet unreachable states like $s_1$ are the key hurdle
 preventing \invariantclosure{} from being equivalent to \invariantconfluence{}.
-This is formalized by \thmref{IclosureEquivalentIconfluence}
+This is formalized by \thmref{IclosureEquivalentIconfluence}.
 
@@
   \[
-    \text{$I(s_0)$ and $O$ is \Iclosed{}} \iff \text{$O$ is \sTIconfluent{}}
+    (\text{$I(s_0)$ and $O$ is \Iclosed{}}) \iff \text{$O$ is \sTIconfluent{}}
   \]

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
took a pass

@@
 \begin{example}\examplelabel{CounreachableExample}
-  Let $O = (\ints^3 \times \ints^3, \join)$ be a
-  PN-Counter~\cite{shapiro2011comprehensive} replicated on three machines.
+  Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps positive and negative integer counts (dubbed a
+  PN-Counter~\cite{shapiro2011comprehensive}), replicated on three machines.
   Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
-  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment and decrement
-  the counter, the $i$th server increments $p_i$ and $n_i$ respectively, and
+  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment or decrement
+  the counter, the $i$th server increments $p_i$ or $n_i$ respectively, and
   $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
@@
   however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that two
+  \ticoreachable{T^+}{I} \jmh{Clarify here how the user recognizes this ...
+I believe that you're saying that the definition of $\Sigma$ implies that decrement requires coordination hence $n_i$ are kept consistent hence disagreeing $n_i$ isn't possible.} The user refines $NR$ with the observation that two
   states are coreachable if and only if they have the same values of $n_1, n_2,
@@
 \subsection{Discussion}
-There are a couple of things worth noting about segmented
+There are a few things worth noting about segmented
 \invariantconfluence{}, its system model, and its decision procedure.
@@
 Third, we can integrate a couple of optimizations into our system model to
-further reduce the amount of coordination it requires. First, if a server with
+further reduce the amount of coordination it requires. To begin, if a server with
 state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
@@
 decide whether or not an object is segmented \invariantconfluent{}, it cannot
-currently help construct the segmentation. In the future, we plan to explore
+currently help construct a segmentation. In the future, we plan to explore
 ways by which we can automatically suggest segmentations to the application

### 2018-05-01 doc/paper/sections/abstract.tex
took a pass

@@
 However, actually determining whether or not an object is \invariantconfluent{}
-is challenging (undecidable). In this paper, we establish conditions under
+is challenging: undecidable in general. In this paper, we establish conditions under
 which a commonly used sufficient condition for \invariantconfluence{} is both
@@
 interactive \invariantconfluence{} decision procedure and (b) a novel
-sufficient condition which can be checked automatically. We then take a step
+sufficient condition that can be checked automatically. We then take a step
 beyond \invariantconfluence{} and introduce a generalization of
@@
 coordination. We implemented our theoretical findings in a prototype called
-Lucy and found that our decision procedures efficiently handle workloads
-including foreign key constraints, rollups, escrow transactions, etc.  We also
+Lucy and found that our decision procedures efficiently handle common real-world workloads
+including foreign key constraints, rollups, escrow transactions, and others.  We also
 found that segmented \invariantconfluent{} replication can deliver up to an

### 2018-05-01 doc/paper/sections/invariant_confluence.tex
took a pass

@@
 \begin{example}\examplelabel{Ints}
-  $O = (\ints, \max)$ is a distributed object consisting of integers merged by
-  $\max$; $t(x) = x + 1$ is a transaction which adds one to a state; and
+  $O = (\ints, \max)$ is a distributed object consisting of integers merged by the
+  $\max$ function; $t(x) = x + 1$ is a transaction which adds one to a state; and
   $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are

### 2018-05-01 doc/paper/sections/related_work.tex
Couple more tweaks.
took a pass

@@
 %
-In~\cite{li2012making}, Li et al.\ also introduce invariant safety as a
+In~\cite{li2012making}, Li et al.\ introduce invariant safety as a
 sufficient (but not necessary) condition for RedBlue consistent objects to be
@@
 distributed systems. The accompanying CALM
-theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
+theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if and only if
 a program can be written in the monotone fragment of these languages, then
-there exists a coordination-free implementation of the program.  The CALM
-theorem provides guarantees about the outputs of a program, but it does not
-guarantee that invariants are maintained throughout the duration of an
-execution. Moreover, Bloom and Dedalus are general purpose programming
-languages that can be used to implement a variety of distributed systems which
+there exists a consistent, coordination-free implementation of the program.  The CALM
+theorem provides guarantees about the consistency of program outputs. It does not directly capture our notions of transactions or invariance maintenance during program execution.
+Moreover, Bloom and Dedalus are general purpose programming
+languages that can be used to implement a variety of distributed systems that
 are outside of the scope of \invariantconfluence{}.

### 2018-05-01 doc/paper/sections/introduction.tex
minor tweakage, mostly for space

@@
 make a fundamental choice between weak and strong consistency. Replicating the
-data with strong consistency using a technique like distributed transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state machine
+data with strong consistency using a technique like distributed transactions (e.g., ~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
 replication~\cite{schneider1990implementing, lamport1998part,
 liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
-oki1988viewstamped, lamport2005generalized, lamport2006fast} makes the
-application designer's life very easy. To the application developer, a strongly
+oki1988viewstamped, lamport2005generalized, lamport2006fast} \jmh{why so many citations? pick a few and say "e.g."} makes the
+application designer's life very easy. To the developer, a strongly
 consistent system behaves exactly like a single-threaded system running on a
@@
 tools that is used to reason about the execution of programs: application
-invariants~\cite{hoare1969axiomatic, balegas2015towards} and integrity constraints~\cite{IntegrityConstraintCitations}. Even if every
+invariants~\cite{hoare1969axiomatic, balegas2015towards} such as database integrity constraints~\cite{IntegrityConstraintCitations}. Even if every
 transaction executing in a weakly consistent system individually maintains an
@@
 availability too? Can we replicate a piece of data with weak consistency but
-still ensure that its invariants are maintained? Yes, sometimes. Bailis et al.\
+still ensure that its invariants are maintained? Yes ... sometimes. Bailis et al.\
 introduced the notion of \emph{\invariantconfluence{}} as a necessary and sufficient
@@
 program that can automatically determine \invariantconfluence{} for us.
-\textbf{The first main thrust of this paper is to design a general-purpose
-\invariantconfluence{} decision procedure, and implement it in an interactive system.}
+\textbf{The first main thrust of this paper is to make \invariantconfluence{} checkable:} to design a general-purpose
+\invariantconfluence{} decision procedure, and implement it in an interactive system.
 
 Designing such a general-purpose decision procedure is
-difficult\footnote{Actually, it's impossible. Determining whether an object is
-\invariantconfluent{} is, in general, undecidable. Still, we can develop a
-decision procedure that works well in the common case.}. Many existing
-approaches instead develop a decision procedure for \emph{\invariantclosure{}}, a
+difficult\footnote{In fact it's impossible. Determining 
+\invariantconfluence{} of an object is undecidable in general. Still, we can develop a
+decision procedure that works well in the common case.}. Many prior
+efforts instead develop a decision procedure for \emph{\invariantclosure{}}, a
 sufficient (but not necessary) condition for
 \invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
-approaches check to see if an object is \invariantclosed{}. If it is, then they
+approaches check whether an object is \invariantclosed{}. If it is, then they
 conclude that it is also \invariantconfluent{}. If it's not, then the current
@@
 
-\textbf{The second main thrust of this paper is to take a step beyond
-\invariantconfluence{} and develop a generalization of \invariantconfluence{}
-called segmented \invariantconfluence{}.} While \invariantconfluence{}
+\textbf{The second main thrust of this paper is to partially avoid coordination even in programs that require it}, by generalizing \invariantconfluence{} to a local property called \emph{segmented \invariantconfluence{}} over subsets of program states.
+While \invariantconfluence{}
 characterizes objects that can be replicated \emph{without any} coordination,
 segmented \invariantconfluence{} allows us to replicate
-non-\invariantconfluent{} objects with only \emph{occasional coordination}. The main
-idea is that invariant satisfying states are divided into a number of segments,
-and the set of allowable transactions within each segment is restricted.
-Within a segment, servers act without any coordination, synchronizing only to
+non-\invariantconfluent{} objects with only \emph{occasional} coordination. The main
+idea is to divide the set of invariant-satisfying states into \emph{segments}, with a restricted set of transactions allowed in each segment.
+Within a segment servers act without any coordination; they synchronize only to
 transition across segment boundaries. This design highlights the trade-off

### 2018-05-01 doc/paper/sections/abstract.tex
minor tweakage, mostly for space

@@
 \begin{abstract}
-Strongly consistent systems are easy to reason about but face fundamental
+Strongly consistent distributed systems are easy to reason about but face fundamental
 limitations in availability and performance. Weakly consistent systems can be
@@
 to replicate non-\invariantconfluent{} objects with a small amount of
-coordination. We implemented our theoretical findings in a prototype called
+coordination.
+
+We implemented these formalisms in a prototype called
 Lucy and found that our decision procedures efficiently handle common real-world workloads
-including foreign key constraints, rollups, escrow transactions, and others.  We also
+including foreign keys, rollups, escrow transactions, and more.  We also
 found that segmented \invariantconfluent{} replication can deliver up to an

### 2018-05-01 benchmark/plot.py
Updated plots with same colors.

@@ -79,8 +79,8 @@ def plot_bank_account(data):
 
         plt.figure(figsize=(8, 3))
         plt.semilogy(gossip_xs, gossip_ys, **EVENTUAL_FMT)
-        plt.semilogy(master_xs, master_ys, **IC_FMT)
         plt.semilogy(paxos_xs, paxos_ys, **LIN_FMT)
+        plt.semilogy(master_xs, master_ys, **IC_FMT)
         plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.2), prop={'size': 12})
         plt.xlabel('Decrement Frequency (fraction of workload)')
         plt.ylabel('Throughput (txns/s)')
@@ -104,8 +104,8 @@ def plot_bank_account(data):
 
         plt.figure(figsize=(8, 4))
         plt.semilogy(gossip_xs, gossip_ys, **EVENTUAL_FMT)
-        plt.semilogy(master_xs, master_ys, **IC_FMT)
         plt.semilogy(paxos_xs, paxos_ys, **LIN_FMT)
+        plt.semilogy(master_xs, master_ys, **IC_FMT)
         plt.legend(loc='upper center', ncol=3, bbox_to_anchor=(0.5, 1.15), prop={'size': 12})
         plt.xlabel('Number of Nodes')
         plt.ylabel('Throughput (txns/s)')

### 2018-05-01 doc/paper/sections/introduction.tex
Integrating Joe's feedback.

@@
 data with strong consistency using a technique like distributed transactions (e.g., ~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
-replication~\cite{schneider1990implementing, lamport1998part,
-liskov2012viewstamped, ongaro2014search, moraru2013there, vincent2015designing,
-oki1988viewstamped, lamport2005generalized, lamport2006fast} \jmh{why so many citations? pick a few and say "e.g."} makes the
+replication (e.g., ~\cite{schneider1990implementing, lamport1998part,
+liskov2012viewstamped, ongaro2014search}) makes the
 application designer's life very easy. To the developer, a strongly
@@
 
-\textbf{The second main thrust of this paper is to partially avoid coordination even in programs that require it}, by generalizing \invariantconfluence{} to a local property called \emph{segmented \invariantconfluence{}} over subsets of program states.
+\textbf{The second main thrust of this paper is to partially avoid coordination even in programs that require it}, by generalizing \invariantconfluence{} to a local property called \emph{segmented \invariantconfluence{}}.
 While \invariantconfluence{}

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
Integrating Joe's feedback.

@@
   however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I} \jmh{Clarify here how the user recognizes this ...
-I believe that you're saying that the definition of $\Sigma$ implies that decrement requires coordination hence $n_i$ are kept consistent hence disagreeing $n_i$ isn't possible.} The user refines $NR$ with the observation that two
+  \ticoreachable{T^+}{I} (all modifications of $(n_1, n_2, n_3)$ require global
+  coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
+  values).
+  The user refines $NR$ with the observation that two
   states are coreachable if and only if they have the same values of $n_1, n_2,

### 2018-05-01 doc/paper/sections/evaluation.tex
Integrating Joe's feedback.

@@
 to confirm our hypotheses and discover the absolute performance benefits. The
-workloads themselves are trivial, but also not salient. \jmh{Salient to what? I'm confused by the last sentence.} Our objective is to
+workloads themselves are trivial but are not the focus of our experiments. Our objective is to
 obtain a controlled measure of throughput and scalability as we vary workload
@@
   linearizable replication is analogous to the relationship between optimistic
-  and pessimistic concurrency control protocols with coordination playing the
-  role of contention \jmh{I didn't understand that last clause about ``coordination playing the role of contention.} Linearizable replication pessimistically assumes that
+  and pessimistic concurrency control protocols. Linearizable replication pessimistically assumes that
   concurrently executing \emph{any} pair of transactions will lead to an
@@
   replication can offer substantial performance benefits over linearizable
-  replication for low-coordination workloads, but is inferior for medium to
-  high contention workloads. \jmh{say why}
+  replication for low-coordination workloads. However, it is inferior for
+  medium to high contention workloads because the majority of transactions that
+  are sent to a server are eventually aborted and relayed to the leader. This
+  additional latency is avoided by linearizable replication which sends
+  transactions directly to the leader.
 
@@
 \begin{benchmark}
-\jmh{Can you get the figures to color ``eventual'' and ``linearizable'' the same in both?}
   In this benchmark, we measure the scale-out of segmented

### 2018-05-01 doc/paper/sections/related_work.tex
Integrating Joe's feedback.
Added section on Calvin, Janus, TAPIR.

@@
 
+\textbf{Serializable Distributed Databases.}
+In \secref{Evaluation}, we saw that segmented \invariantconfluent{} replication
+vastly outperforms linearizable replication for low coordination workloads, and
+it performs comparably or worse for medium and high coordination workloads.
+Distributed databases like Calvin~\cite{thomson2012calvin},
+Janus~\cite{mu2016consolidating}, and TAPIR~\cite{zhang2015building} employ
+algorithmic optimizations to implement serializable transactions with high
+throughput and low latency.
+%
+While segmented \invariantconfluent{} replication will likely always outperform
+serializable replication for low coordination workloads, these databases make
+serializable replication the most performant option for executing workloads
+that require a modest amount of coordination.
+
 \textbf{Branch and Merge.}
@@
 distributed systems. The accompanying CALM
-theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if and only if
-a program can be written in the monotone fragment of these languages, then
-there exists a consistent, coordination-free implementation of the program.  The CALM
-theorem provides guarantees about the consistency of program outputs. It does not directly capture our notions of transactions or invariance maintenance during program execution.
-Moreover, Bloom and Dedalus are general purpose programming
-languages that can be used to implement a variety of distributed systems that
-are outside of the scope of \invariantconfluence{}.
+theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
+and only if a program can be written in the monotone fragment of these
+languages, then there exists a consistent, coordination-free implementation of
+the program.  The CALM theorem provides guarantees about the consistency of
+program outputs. It does not directly capture our notions of transactions or
+invariant maintenance during program execution.  Moreover, Bloom and Dedalus
+are general purpose programming languages that can be used to implement a
+variety of distributed systems that are outside of the scope of
+\invariantconfluence{}.
 
-% \textbf{Program Analysis in Database Systems.}
-% Using program analysis to improve the performance of database systems is not
-% new. For example, it has been used to improve the performance of
-% database-backed web applications~\cite{cheung2014using, wu2016transaction,
-% ramachandra2012program} and used to improve the performance of optimistic
-% concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
-% on \invariantconfluence{} continues the theme of using program analysis to reap
-% the performance benefits gained from understanding program semantics.
+\textbf{Program Analysis in Database Systems.}
+Using program analysis to improve the performance of database systems is not
+new. For example, it has been used to improve the performance of
+database-backed web applications~\cite{cheung2014using, wu2016transaction,
+ramachandra2012program} and used to improve the performance of optimistic
+concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
+on \invariantconfluence{} continues the theme of using program analysis to reap
+the performance benefits gained from understanding program semantics.

### 2018-05-01 doc/paper/sections/introduction.tex
Honey, I shrunk the paper.

@@
 make a fundamental choice between weak and strong consistency. Replicating the
-data with strong consistency using a technique like distributed transactions (e.g., ~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
-replication (e.g., ~\cite{schneider1990implementing, lamport1998part,
-liskov2012viewstamped, ongaro2014search}) makes the
-application designer's life very easy. To the developer, a strongly
-consistent system behaves exactly like a single-threaded system running on a
-single node, so reasoning about the behavior of the system is
-simple~\cite{herlihy1990linearizability}.  Unfortunately, strong consistency is
-at odds with performance. The CAP theorem and PACELC theorem tell us that
-strongly consistent systems suffer from higher latency at best and
-unavailability at worst~\cite{gilbert2002brewer, brewer2012cap,
-abadi2012consistency}. On the other hand, weak consistency models like eventual
-consistency~\cite{vogels2009eventually}, PRAM
+data with strong consistency using a technique like distributed transactions
+(e.g., \cite{bernstein1981concurrency,mohan1986transaction}) or state machine
+replication (e.g., \cite{schneider1990implementing, lamport1998part,
+liskov2012viewstamped, ongaro2014search}) makes the application designer's life
+very easy. To the developer, a strongly consistent system behaves exactly like
+a single-threaded system running on a single node, so reasoning about the
+behavior of the system is simple~\cite{herlihy1990linearizability}.
+Unfortunately, strong consistency is at odds with performance. The CAP theorem
+and PACELC theorem tell us that strongly consistent systems suffer from higher
+latency at best and unavailability at worst~\cite{gilbert2002brewer,
+brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
+models like eventual consistency~\cite{vogels2009eventually}, PRAM
 consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
@@
 application designer to reason about the complex interleavings of operations
-that are allowed by these weak consistency models. In particular, weak consistency
-models strip an application developer of one of the earliest and most effective
-tools that is used to reason about the execution of programs: application
-invariants~\cite{hoare1969axiomatic, balegas2015towards} such as database integrity constraints~\cite{IntegrityConstraintCitations}. Even if every
-transaction executing in a weakly consistent system individually maintains an
-application invariant, the system as a whole can produce invariant-violating
-states.
+that are allowed by these weak consistency models. In particular, weak
+consistency models strip an application developer of one of the earliest and
+most effective tools that is used to reason about the execution of programs:
+application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
+database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
+if every transaction executing in a weakly consistent system individually
+maintains an application invariant, the system as a whole can produce
+invariant-violating states.
 
@@
 availability too? Can we replicate a piece of data with weak consistency but
-still ensure that its invariants are maintained? Yes ... sometimes. Bailis et al.\
-introduced the notion of \emph{\invariantconfluence{}} as a necessary and sufficient
-condition for when invariants can be maintained over replicated data without
-the need for any coordination~\cite{bailis2014coordination}. If an object is
-\invariantconfluent{} with respect to an invariant, we can replicate it with
-the performance benefits of weak consistency and (some of) the correctness
-benefits of strong consistency.
+still ensure that its invariants are maintained? Yes... sometimes. Bailis et
+al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
+sufficient condition for when invariants can be maintained over replicated data
+without the need for any coordination~\cite{bailis2014coordination}. If an
+object is \invariantconfluent{} with respect to an invariant, we can replicate
+it with the performance benefits of weak consistency and (some of) the
+correctness benefits of strong consistency.
 
-Unfortunately, to date, the task of identifying whether or not
-an object actually is \invariantconfluent{} has remained an exercise in human proof generation. Bailis et al.\
-manually categorized a set of common objects, transactions, and invariants
-(e.g.\ foreign key constraints on relations, linear constraints on integers) as
-\invariantconfluent{} or not. Hand-written proofs of this sort are unreasonable to expect from programmers. Ideally we would have a general purpose
-program that can automatically determine \invariantconfluence{} for us.
-\textbf{The first main thrust of this paper is to make \invariantconfluence{} checkable:} to design a general-purpose
-\invariantconfluence{} decision procedure, and implement it in an interactive system.
+Unfortunately, to date, the task of identifying whether or not an object
+actually is \invariantconfluent{} has remained an exercise in human proof
+generation. Bailis et al.\ manually categorized a set of common objects,
+transactions, and invariants (e.g.\ foreign key constraints on relations,
+linear constraints on integers) as \invariantconfluent{} or not. Hand-written
+proofs of this sort are unreasonable to expect from programmers. Ideally we
+would have a general purpose program that can automatically determine
+\invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
+to make \invariantconfluence{} checkable:} to design a general-purpose
+\invariantconfluence{} decision procedure, and implement it in an interactive
+system.
 
-Designing such a general-purpose decision procedure is
-difficult\footnote{In fact it's impossible. Determining
-\invariantconfluence{} of an object is undecidable in general. Still, we can develop a
-decision procedure that works well in the common case.}. Many prior
-efforts instead develop a decision procedure for \emph{\invariantclosure{}}, a
-sufficient (but not necessary) condition for
-\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
-approaches check whether an object is \invariantclosed{}. If it is, then they
-conclude that it is also \invariantconfluent{}. If it's not, then the current
-approaches are unable to conclude anything about whether or not the object is
-\invariantconfluent{}. In this paper, we take a step back and study the
-underlying reason \emph{why} \invariantclosure{} is not necessary for
-\invariantconfluence{}. We find that \invariantconfluence{} is fundamentally a
-reachability property, checking whether certain values of our replicated
-object can actually be obtained in the execution of a system. We discover that
-unreachable, yet invariant satisfying, states cause \invariantclosure{} to be
-an unnecessary condition. Using this understanding, we construct a set of
-modest conditions under which \invariantclosure{} and \invariantconfluence{}
-are in fact \emph{equivalent}, allowing us to reduce the problem of determining
-\invariantconfluence{} to that of determining \invariantclosure{}.
+Designing such a general-purpose decision procedure is difficult\footnote{In
+fact it's impossible. Determining \invariantconfluence{} of an object is
+undecidable in general. Still, we can develop a decision procedure that works
+well in the common case.}. Many prior efforts instead develop a decision
+procedure for \emph{\invariantclosure{}}, a sufficient (but not necessary)
+condition for \invariantconfluence{}~\cite{li2012making, li2014automating}. The
+existing approaches check whether an object is \invariantclosed{}. If it is,
+then they conclude that it is also \invariantconfluent{}. If it's not, then the
+current approaches are unable to conclude anything about whether or not the
+object is \invariantconfluent{}.
 
+In this paper, we take a step back and study the underlying reason \emph{why}
+\invariantclosure{} is not necessary for \invariantconfluence{}.
+%
+% We find that \invariantconfluence{} is fundamentally a reachability property,
+% checking whether certain values of our replicated object can actually be
+% obtained in the execution of a system. We discover that unreachable, yet
+% invariant satisfying, states cause \invariantclosure{} to be an unnecessary
+% condition.
+%
+Using this understanding, we construct a set of modest conditions under which
+\invariantclosure{} and \invariantconfluence{} are in fact \emph{equivalent},
+allowing us to reduce the problem of determining \invariantconfluence{} to that
+of determining \invariantclosure{}.
+%
 Then, we use these conditions to design a general-purpose interactive
-\invariantconfluence{} decision procedure. Users iteratively interact with the
-decision procedure, eliminating unreachable states from the invariant, moving
-towards the conditions under which \invariantclosure{} and
-\invariantconfluence{} are equivalent. Meanwhile, the decision procedure
-generates examples of reachable and unreachable states which help the user
-recognize patterns that describe reachability. We also leverage our intuitions
-about reachability to develop a new sufficient condition for
-\invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{} covers
-some cases that are not covered by \invariantclosure{}, and it can be
-automatically checked without any user interaction.
+\invariantconfluence{} decision procedure and a new sufficient condition for
+\invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
+covers some cases that are not covered by \invariantclosure{}, and it can be
+checked automatically.
 
-\textbf{The second main thrust of this paper is to partially avoid coordination even in programs that require it}, by generalizing \invariantconfluence{} to a local property called \emph{segmented \invariantconfluence{}}.
-While \invariantconfluence{}
-characterizes objects that can be replicated \emph{without any} coordination,
-segmented \invariantconfluence{} allows us to replicate
-non-\invariantconfluent{} objects with only \emph{occasional} coordination. The main
-idea is to divide the set of invariant-satisfying states into \emph{segments}, with a restricted set of transactions allowed in each segment.
-Within a segment servers act without any coordination; they synchronize only to
-transition across segment boundaries. This design highlights the trade-off
-between application complexity and coordination-freedom; more complex
-applications require more segments which require more coordination, and
-vice-versa.
+% Then, we use these conditions to design a general-purpose interactive
+% \invariantconfluence{} decision procedure. Users iteratively interact with the
+% decision procedure, eliminating unreachable states from the invariant, moving
+% towards the conditions under which \invariantclosure{} and
+% \invariantconfluence{} are equivalent. Meanwhile, the decision procedure
+% generates examples of reachable and unreachable states which help the user
+% recognize patterns that describe reachability. We also leverage our intuitions
+% about reachability to develop a new sufficient condition for
+% \invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
+% covers some cases that are not covered by \invariantclosure{}, and it can be
+% automatically checked without any user interaction.
+
+\textbf{The second main thrust of this paper is to partially avoid coordination
+even in programs that require it}, by generalizing \invariantconfluence{} to a
+local property called \emph{segmented \invariantconfluence{}}. While
+\invariantconfluence{} characterizes objects that can be replicated
+\emph{without any} coordination, segmented \invariantconfluence{} allows us to
+replicate non-\invariantconfluent{} objects with only \emph{occasional}
+coordination. The main idea is to divide the set of invariant-satisfying states
+into \emph{segments}, with a restricted set of transactions allowed in each
+segment.  Within a segment servers act without any coordination; they
+synchronize only to transition across segment boundaries. This design
+highlights the trade-off between application complexity and
+coordination-freedom; more complex applications require more segments which
+require more coordination, and vice-versa.
+
+% TODO: add bit on performance.
 
 In closing, here is an outline of the paper and of our contributions:
-\begin{itemize}
-  \item
+%
 We propose a novel expression-oriented definition of \invariantconfluence{}
 that is both formal and simple (\secref{InvariantConfluence}).
-
-  \item
+%
 We develop an understanding of why \invariantclosure{} is not necessary for
-    \invariantconfluence{} and use this understanding to develop conditions
-    under which it is both necessary and sufficient
-    (\secref{InvariantClosure}).
-
-  \item
-    We exploit these conditions to design a general-purpose interactive
-    decision procedure for \invariantconfluence{}
-    (\secref{InteractiveDecisionProcedure}).
-
-  \item
+\invariantconfluence{} and use this understanding to develop conditions under
+which it is both necessary and sufficient (\secref{InvariantClosure}).
+%
+We exploit these conditions to design a general-purpose interactive decision
+procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
+%
 We again exploit these conditions to design a novel non-trivial sufficient
-    condition for \invariantconfluence{}, called \mergereducibility, that does
-    not depend on user interaction (\secref{MergeReduction}).
-
-  \item
+condition for \invariantconfluence{}, called \mergereducibility.
+%
 We present segmented \invariantconfluence{}: a generalization of
 \invariantconfluence{} that uses a small amount of coordination to maintain
-    invariants for replicated objects that are otherwise not
-    \invariantconfluent{} (\secref{SegmentedInvariantConfluence}).
-
-  \item
+invariants for replicated objects that are otherwise not \invariantconfluent{}
+(\secref{SegmentedInvariantConfluence}).
+%
 We implement our decision procedures and sufficient conditions for both
-    \invariantconfluence{} and segmented \invariantconfluence{} in a system
-    called Lucy, then use Lucy to evaluate our methods (\secref{Evaluation}).
-\end{itemize}
+\invariantconfluence{} and segmented \invariantconfluence{} in a system called
+Lucy, then use Lucy to evaluate our methods (\secref{Evaluation}).

### 2018-05-01 doc/paper/sections/related_work.tex
Honey, I shrunk the paper.

@@
 workloads that the homeostasis protocol would execute with more coordination
-than a segmented \invariantconfluent{} execution.  The protocol generates
-global invariants, called global treaties, and servers operate without
-coordination so long as they preserve the treaty. When the treaty is violated,
-the nodes perform a round of coordination and establish a new treaty.  This
-mechanism of establishing invariants and operating without coordination so long
-as the invariants are maintained is very similar to our design of segmented
-\invariantconfluence{}.
+than a segmented \invariantconfluent{} execution. Moreover, the homeostasis and
+demarcation protocols' mechanism of establishing global invariants and
+operating without coordination so long as the invariants are maintained is very
+similar to our design of segmented \invariantconfluence{}.
 

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
Honey, I shrunk the paper.

@@
 
-\begin{itemize}
-  \item
 If $t \notin T_i$, then $p_i$ initiates a round of global coordination to
 execute $t$. During global coordination, every server temporarily stops
-    processing transactions and transitions to state $s = s_1 \join \ldots
-    \join s_n$, the join of every server's state. Then, every server
-    speculatively executes $t$ transitioning to state $t(s)$. If $t(s)$
-    violates the invariant (i.e.  $\lnot I(t(s))$), then every server aborts
-    $t$ and reverts to state $s$.  Then, $p_i$ replies to the client. If $t(s)$
-    satisfies the invariant (i.e.  $I(t(s))$), then every server commits $t$
-    and remains in state $t(s)$. Every server then chooses the first segment
-    $(I_i, T_i) \in \Sigma$ such that $I_i(t(s))$ to be the new active segment.
-    Note that such a segment is guaranteed to exist because the segment
-    invariants cover $I$. Moreover, $\Sigma$ is ordered, so every server will
-    deterministically pick the same active segment. In fact, an invariant of
-    the system model is that at any given point of normal processing, every
-    server has the same active segment.
+processing transactions and transitions to state $s = s_1 \join \ldots \join
+s_n$, the join of every server's state. Then, every server speculatively
+executes $t$ transitioning to state $t(s)$. If $t(s)$ violates the invariant
+(i.e.  $\lnot I(t(s))$), then every server aborts $t$ and reverts to state $s$.
+Then, $p_i$ replies to the client. If $t(s)$ satisfies the invariant (i.e.
+$I(t(s))$), then every server commits $t$ and remains in state $t(s)$. Every
+server then chooses the first segment $(I_i, T_i) \in \Sigma$ such that
+$I_i(t(s))$ to be the new active segment.  Note that such a segment is
+guaranteed to exist because the segment invariants cover $I$. Moreover,
+$\Sigma$ is ordered, so every server will deterministically pick the same
+active segment. In fact, an invariant of the system model is that at any given
+point of normal processing, every server has the same active segment.
 
-  \item
-    If $t \in T_i$, then $p_i$ executes $t$ immediately and without
+Otherwise, if $t \in T_i$, then $p_i$ executes $t$ immediately and without
 coordination.  If $t(s_i)$ satisfies the \emph{active} invariant (i.e.
-    $I_i(t(s_i))$), then $p_i$ commits $t$, stays in state $t(s_i)$, and
-    replies to the client. If $t(s_i)$ violates the \emph{global} invariant
-    (i.e. $\lnot I(t(s_i))$), then $p_i$ aborts $t$, reverts to state $s_i$,
-    and replies to the client. If $t(s_i)$ satisfies the global invariant but
-    violates the active invariant (i.e.  $I(t(s_i))$ but $\lnot I_i(t(s_i))$),
-    then $p_i$ reverts to state $s_i$ and initiates a round of global
-    coordination to execute $t$, as described in the previous paragraph.
-    Transaction execution is summarized in \algoref{TxnExecution}.
-\end{itemize}
+$I_i(t(s_i))$), then $p_i$ commits $t$, stays in state $t(s_i)$, and replies to
+the client. If $t(s_i)$ violates the \emph{global} invariant (i.e. $\lnot
+I(t(s_i))$), then $p_i$ aborts $t$, reverts to state $s_i$, and replies to the
+client. If $t(s_i)$ satisfies the global invariant but violates the active
+invariant (i.e.  $I(t(s_i))$ but $\lnot I_i(t(s_i))$), then $p_i$ reverts to
+state $s_i$ and initiates a round of global coordination to execute $t$, as
+described in the previous paragraph.  Transaction execution is summarized in
+\algoref{TxnExecution}.
 
@@
 
-Second, segmented \invariantconfluence{} naturally subsumes a distributed
-locking approach to replicating non-\invariantconfluent{} objects.  This
-approach first recognizes which transactions cannot be safely executed
-concurrently and then requires them to acquire a distributed lock before
-executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
-banking application with the invariant that all balances remain non-negative,
-concurrent deposits are permitted, but concurrent withdrawals can lead to
-invariant violations. Thus, we require that withdrawals acquire a distributed
-lock before executing. This example is exactly the same as
-\exampleref{CounreachableExample} which we handled by simply removing
-withdrawal transactions from our segmentation's set of transactions.
+% Second, segmented \invariantconfluence{} naturally subsumes a distributed
+% locking approach to replicating non-\invariantconfluent{} objects.  This
+% approach first recognizes which transactions cannot be safely executed
+% concurrently and then requires them to acquire a distributed lock before
+% executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
+% banking application with the invariant that all balances remain non-negative,
+% concurrent deposits are permitted, but concurrent withdrawals can lead to
+% invariant violations. Thus, we require that withdrawals acquire a distributed
+% lock before executing. This example is exactly the same as
+% \exampleref{CounreachableExample} which we handled by simply removing
+% withdrawal transactions from our segmentation's set of transactions.
 
-Third, we can integrate a couple of optimizations into our system model to
-further reduce the amount of coordination it requires. To begin, if a server with
-state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
-to execute, and $t(s_i)$ violates the active invariant but not the global
-invariant, instead of initiating a round of global coordination, $p_i$ can
-simply buffer $t$ for re-execution at a later time. While this increases the
-latency required to execute $t$, it's possible that after other transactions
-are executed, re-executing $t$ may lead to a state that either satisfies the
-active invariant or violates the global invariant. In either case, a round of
-global coordination is avoided. Similarly, servers can buffer transactions that
-require global coordination, executing an entire batch of these transactions
-during a single round of global coordination.
+% Second, we can integrate a couple of optimizations into our system model to
+% further reduce the amount of coordination it requires. To begin, if a server with
+% state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
+% to execute, and $t(s_i)$ violates the active invariant but not the global
+% invariant, instead of initiating a round of global coordination, $p_i$ can
+% simply buffer $t$ for re-execution at a later time. While this increases the
+% latency required to execute $t$, it's possible that after other transactions
+% are executed, re-executing $t$ may lead to a state that either satisfies the
+% active invariant or violates the global invariant. In either case, a round of
+% global coordination is avoided. Similarly, servers can buffer transactions that
+% require global coordination, executing an entire batch of these transactions
+% during a single round of global coordination.
 
-Fourth, a segmented \invariantconfluence{} decision procedure can also leverage
+Second, we can integrate a couple of optimizations into our system model to
+further reduce the amount of coordination it requires. For example, servers can
+buffer transactions that require global coordination, executing an entire batch
+of these transactions during a single round of global coordination.
+
+Third, a segmented \invariantconfluence{} decision procedure can also leverage
 \thmref{LatticeProperty} in addition to
@@
 
-Fifth, while a segmented \invariantconfluence{} decision procedure can help
+Fourth, while a segmented \invariantconfluence{} decision procedure can help
 decide whether or not an object is segmented \invariantconfluent{}, it cannot

### 2018-05-01 doc/paper/sections/introduction.tex
Added more on performance.

@@
 
-% TODO: add bit on performance.
+Finally, we present Lucy: an implementation of our decision procedures and a
+system for replicating \invariantconfluent{} and segmented
+\invariantconfluent{} objects. Using Lucy, we find that our decision
+procedures can efficiently handle a wide range of common workloads. For
+example, in \secref{Evaluation}, we apply Lucy to foreign key constraints,
+escrow transactions, an auction application, and the TPC-C benchmark. Lucy
+processes these workloads in less than half a second, and no workload requires
+more than 66 lines of code to specify. Moreover, we find that segmented
+\invariantconfluent{} replication can achieve 100x to 10x more throughput than
+linearizable replication for low-coordination workloads.
 
@@
 %
-We implement our decision procedures and sufficient conditions for both
-\invariantconfluence{} and segmented \invariantconfluence{} in a system called
-Lucy, then use Lucy to evaluate our methods (\secref{Evaluation}).
+We evaluate our methods using a prototype implementation called Lucy
+(\secref{Evaluation}).

### 2018-05-01 doc/paper/sections/conclusion.tex
Last tweaks!

@@
 \textbf{Acknowledgments.}
-The authors would like to thank Peter Alvaro, Alvin Cheung, Anthony Tan, and
-Cristina Teodoropol for fruitful discussion and feedback.
+The authors would like to thank Alan Fekete, Peter Alvaro, Alvin Cheung,
+Alexandra Meliou, Anthony Tan, and Cristina Teodoropol for fruitful discussion
+and feedback.
 %

### 2018-05-01 doc/paper/sections/evaluation.tex
Last tweaks!

@@
 Lucy also includes an implementation of the \invariantconfluence{} and
-segmented-invariant confluence system models in roughly 3,500 lines of C++.
+segmented \invariantconfluence{} system models in roughly 3,500 lines of C++.
 Users specify objects, transactions, invariants, and segmentations in C++. Lucy
@@
 
-These hypotheses state that segmented \invariantconfluent{} is more performant
-than linearizable replication for low-coordination workloads. But by how much?
-We also aim to measure the absolute performance and scalability benefits of
-segmented \invariantconfluent{} replication and how they vary as we vary the
-coordination required by a workload. We perform two controlled microbenchmarks
-to confirm our hypotheses and discover the absolute performance benefits. The
-workloads themselves are trivial but are not the focus of our experiments. Our objective is to
-obtain a controlled measure of throughput and scalability as we vary workload
-contention.
+These hypotheses state that segmented \invariantconfluent{} replication is more
+performant than linearizable replication for low-coordination workloads. But by
+how much?  We also aim to measure the absolute performance and scalability
+benefits of segmented \invariantconfluent{} replication and how they vary as we
+vary the coordination required by a workload. We perform two controlled
+microbenchmarks to confirm our hypotheses and discover the absolute performance
+benefits. The workloads themselves are trivial but are not the focus of our
+experiments. Our objective is to obtain a controlled measure of throughput and
+scalability as we vary workload contention.
 

### 2018-05-01 doc/paper/sections/segmented_invariant_confluence.tex
Added more on performance.
Last tweaks!

@@
 invariants of non-\invariantconfluent{} objects, requiring only a small amount
-of coordination.
+of coordination. In \secref{Evaluation}, we see that replicating a
+non-\invariantconfluent{} object with segmented \invariantconfluence{} can
+achieve between 100x and 10x more throughput than linearizable replication for
+certain workloads.
 
@@
 
-{\input{figures/segmented_z2}}
-
 \begin{example}\examplelabel{SegmentedZ2}
@@
 \end{example}
+
+{\input{figures/segmented_z2}}
 
@@
 to another segment, the coordination ensures that all servers begin in the same
-start state and with the same active replica, reestablishing the assumptions of
+start state and with the same active segment, reestablishing the assumptions of
 the \invariantconfluence{} system model.

### 2018-09-19 doc/paper/sections/introduction.tex
Removed confusing footnote in first section.
From reviewer 12:

> The first footnote is confusing. Should the footnote be simply
> embedded into the main text?

This commit removes the footnote and moves the footnote's text into the
main text.

@@
 
-Designing such a general-purpose decision procedure is difficult\footnote{In
-fact it's impossible. Determining \invariantconfluence{} of an object is
+Unfortunately, designing such a general-purpose decision procedure is
+impossible because determining the \invariantconfluence{} of an object is
 undecidable in general. Still, we can develop a decision procedure that works
-well in the common case.}. Many prior efforts instead develop a decision
-procedure for \emph{\invariantclosure{}}, a sufficient (but not necessary)
-condition for \invariantconfluence{}~\cite{li2012making, li2014automating}. The
-existing approaches check whether an object is \invariantclosed{}. If it is,
-then they conclude that it is also \invariantconfluent{}. If it's not, then the
-current approaches are unable to conclude anything about whether or not the
-object is \invariantconfluent{}.
+well in the common case.
+%
+For example, many prior efforts have developed decision procedures for
+\emph{\invariantclosure{}}, a sufficient (but not necessary) condition for
+\invariantconfluence{}~\cite{li2012making, li2014automating}. The existing
+approaches check whether an object is \invariantclosed{}. If it is, then they
+conclude that it is also \invariantconfluent{}. If it's not, then the current
+approaches are unable to conclude anything about whether or not the object is
+\invariantconfluent{}.
 

### 2018-09-19 doc/paper/sections/merge_reduction.tex
Shrinking paper.
Fixed confluence/closure typo.
From reviewer 12:

> The paper states that in the end of Section 3, "invariant closure is
> completely ignorant of reachability", whereas in Section 5, "invariant
> closure is fundamentally a property of reachability". Which one is
> correct?

Very keen eye by the reviewer! That was a typo. This commit fixes that.

@@
 \section{Merge Reduction}\seclabel{MergeReduction}
-In \secref{InvariantClosure}, we discussed how \invariantclosure{} is
+In \secref{InvariantClosure}, we discussed how \invariantconfluence{} is
 fundamentally a property of reachability and that \invariantclosure{} is
@@
 
-{\input{figures/merge_reducible}}
+% {\input{figures/merge_reducible}}
 

### 2018-09-19 doc/paper/sections/related_work.tex
Expanded on differences in related work.
From reviewer 8:

> The related work section should more strictly emphasize the difference
> of the work from previous work in the field.

I read through the related works section again and expanded on some of
the differences between our work and RedBlue consistency and explicit
consistency. I thought the differences in the other sections of the
related work were pretty clear.

@@
 sufficient (but not necessary) condition for RedBlue consistent objects to be
-\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{} for
-an operation-based system model (as opposed to the state based system model
-that we use). Li et al.\ develop sophisticated techniques for deciding
-invariant safety that involve calculating weakest preconditions. These
+\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{}.
+In~\cite{li2014automating}, Li et al.\ develop sophisticated techniques for
+deciding invariant safety that involve calculating weakest preconditions. These
 techniques are complementary to our work and can be used to improve the
 \invariantclosure{} subroutine used by our decision procedures.
+%
+In contrast with these techniques, our \invariantconfluence{} decision
+procedures can determine the \invariantconfluence{} of objects that are
+\emph{not} invariant safe.
 
@@
 this is a sufficient condition for explicit consistency. It is similar to
-criterion (3) in \thmref{LatticeProperty}. Balegas et al.\ also describe a
-variety of techniques---like conflict resolution, locking, and escrow
-transactions~\cite{o1986escrow}---that can be used to replicate workloads that
-do not meet their sufficient conditions. As discussed in \secref{Evaluation},
-simple forms of these techniques can be modelled with segmented
+criterion (3) in \thmref{LatticeProperty}. In our work, we take a step further
+and explore sufficient \emph{and necessary} conditions for
 \invariantconfluence{}.
+%
+Balegas et al.\ also describe a variety of techniques---like conflict
+resolution, locking, and escrow transactions~\cite{o1986escrow}---that can be
+used to replicate workloads that do not meet their sufficient conditions.
+Segmented \invariantconfluence{} is a general-purpose formalism that can be
+used to model simple forms of these techniques.
 

### 2018-09-19 doc/paper/sections/evaluation.tex
Added hw configuration for decision procedures.
From reviewer 12:

> The hardware configuration should be specified (in Section 7.1)
> before discussing the runtimes (in Section 7.2)

In section 7.1, we discuss the implementation of the Lucy decision
procedures and runtime. In section 7.2, we evaluate the decision
procedures. In section 7.3, we evaluate the runtime.

At the beginning of section 7.3, we describe the hardware on which we
run the 7.3 experiments. We never described the hardware configuration
of the experiments in 7.2. I think the reviewer is telling us to add it.
This commit does

@@
 \algoref{ArbitraryStartInteractiveDecisionProcedure}. The decision procedures
-are implemented in roughly 2,500 lines of Python. Users specify objects,
+are implemented in roughly 2,500 lines of Python. Programmers specify objects,
 transactions, and invariants in a small Python DSL and interact with the
-interactive decision procedures using an interactive Python console. We use
-Z3~\cite{de2008z3} to implement our \invariantclosure{} decision procedure,
+interactive decision procedures using an interactive Python console. Note that
+a programmer only has to run the decision procedures offline a single time to
+check the \invariantconfluence{} of their distributed object. The decision
+procedures do not have to be run online when transactions are being processed.
+
+We use Z3~\cite{de2008z3} to implement our \invariantclosure{} decision procedure,
 compiling an object and invariant into a formula that is satisfiable if and
@@
 simple, yet practical examples. We then discuss how our tool can be used to
-analyze the TPC-C benchmark. A summary of these results is given in
-\tabref{EvalTable}.
+analyze the TPC-C benchmark. All decision procedures were run on a MacBook Pro
+laptop with a 3.5 GHz Intel Core i7 processor and 16 GB of RAM. A summary of
+these results is given in \tabref{EvalTable}.
 

### 2018-09-19 doc/paper/sections/related_work.tex
Added CRDT section to related work.
From reviewer 11:

> A clear connection to what CRDTs are within scope would improve the
> contribution of this paper.

and

> You make a connection to Shapiro et al.'s work on CRDT and use a few
> examples that are popular (e.g. PN-Counter), but can you go further
> and connect to more complex structures like Sequence. This will allow
> you to reach a broader audience and let them clearly understand the
> broader implication of this work. (BTW, nice discussion of PK-FK).

Previously, we had a paragraph on CRDTs in the related works section. I
cut it out bec

@@
 
-% \textbf{CRDTs.}
-% CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
-% semilattices with inflationary update methods. Due to their algebraic
-% properties, CRDTs can be replicated with strong eventual consistency without
-% the need for any coordination. Our definition of distributed objects and our
-% \invariantconfluence{} system model are inspired directly by the corresponding
-% definitions and system models in~\cite{shapiro2011conflict}. Also, an
-% \invariantconfluent{} object does not necessarily have to be eventually
-% consistent, so it is natural (though not necessary) to use CRDTs as
-% \invariantconfluent{} distributed objects, achieving a combination of eventual
-% consistency and invariant preservation. Our criteria in
-% \thmref{LatticeProperty} also borrow ideas from CRDTs, exploiting the nice
-% algebraic properties of semilattices.
+\textbf{CRDTs.}
+CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
+semilattices with inflationary update methods. Due to their algebraic
+properties, CRDTs can be replicated with strong eventual consistency without
+the need for any coordination. Our definition of distributed objects and our
+\invariantconfluence{} system model are inspired directly by the corresponding
+definitions and system models in~\cite{shapiro2011conflict}.
+%
+CRDTs are eventually consistent but may not preserve invariants. Conversely,
+\invariantconfluent{} objects preserve invariants but may not be eventually
+consistent. Thus, it is natural (though not necessary) to use CRDTs as
+distributed objects. If a CRDT is determined to be \invariantconfluent{} with
+respect to a particular invariant and set of transactions, then it achieves a
+combination of strong eventual consistency and invariant preservation. Any CRDT
+(e.g., counters, sets, graphs, sequences) can be used for this purpose.
+%
+Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
+exploiting the nice algebraic properties of semilattices.
 

### 2018-09-19 doc/paper/sections/invariant_confluence.tex
Last tweaks!
Clarified read-only transactions.
From reviewer 11:

> It appears that there is an issue with readers who want to look at a
> dataset (e.g. compute an aggregate, or min). This triggers a merge?
> This issue should be clearly described.

Our formalism doesn't really discuss reads at all because reads don't
affect invariant confluence in any way. I'm guessing this is why the
reviewer got a little confused about what happens when you want to read
the object. I added a paragraph trying to clarify how invariant
confluence models reads.

--- 2018-04-30 doc/paper/sections/invariant_confluence.tex
@@
 A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of states
-and a binary merge operator $\join: S \times S \to S$ which merges two states
-into one. A \defword{transaction} $t: S \to S$ is a function which maps one
+and a binary merge operator $\join: S \times S \to S$ that merges two states
+into one. A \defword{transaction} $t: S \to S$ is a function that maps one
 state to another. An \defword{invariant} $I$ is a subset of $S$. Notationally,
@@
 \begin{example}\examplelabel{Ints}
-  $O = (\ints, \max)$ is a distributed object consisting of integers merged by
-  $\max$; $t(x) = x + 1$ is a transaction which adds one to a state; and
+  $O = (\ints, \max)$ is a distributed object consisting of integers merged by the
+  $\max$ function; $t(x) = x + 1$ is a transaction that adds one to a state; and
   $\setst{x \in \ints}{x \geq 0}$ is the invariant that states $x$ are
@@
 \end{example}
+
+Note that by modelling a transaction $t$ as a function $S \to S$, we focus
+exclusively on the effects that a transaction has on the object (i.e.\
+``writes'' to the object). Transactions are also free to read the value of the
+object, but these reads are not captured by our model because, as we'll see,
+they do not affect \invariantconfluence{}. For example, we could model any
+read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.
 
@@
 \end{itemize}
-Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists a
+Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an
 \sTIreachable{} expression $e$ that evaluates to $s$. Returning to

### 2018-09-19 doc/paper/sections/evaluation.tex
Consequence of TPC-C not being iconfluent.
From reviewer 11:

> During the TPC-C discussion, you should complete the description and
> let the reader know what it the implication of the sequence and unique
> identifiers being not invariant means for running that workload.

This commit tries to clarify the consequence of TPC-C not being
invariant confluent.

@@
     unique identifiers}. Consider, for example, invariant 2 which dictates that
-    a district's next order id \ttt{D\_NEXT\_O\_ID} is equal to the maximum
+    a district's next order ID \ttt{D\_NEXT\_O\_ID} is equal to the maximum
     order id \ttt{O\_ID} of orders within the district. The New Order
@@
     We model this workload with an integer for \ttt{D\_NEXT\_O\_ID} and a map
-    for \ttt{O\_ID} that maps order ids to order.
+    for \ttt{O\_ID} that maps order IDs to order.
     %
@@
     We applied Lucy to this workload and in less than a second, it produced a
-    counterexample that---when labelled as reachable---confirms that the
-    workload is \emph{not} \invariantconfluent{}~\cite{bailis2014coordination}.
+    counterexample that---when labelled as reachable---confirms Bailis et al.'s
+    finding that the workload is \emph{not}
+    \invariantconfluent{}~\cite{bailis2014coordination}. Thus, the TPC-C
+    benchmark requires some form of coordination to ensure unique and
+    sequential IDs. Alternatively, as Bailis et al. describe
+    in~\cite{bailis2014coordination}, the workload can be run coordination free
+    if we drop the requirement that IDs are assigned sequentially.
 % \end{enumerate}

### 2018-09-19 doc/paper/sections/segmented_invariant_confluence.tex
Remove optimizations + limitations of segmentation
This commit addresses two pieces of reviewer feedback.

From reviewer 12:

> The optimizations are simply vaguely discussed in Section 6.4.

I completeley removed the vague paragraph on optimizations. It is vague,
not super important, and the paper is a bit long.

Also from reviewer 12:

> Regarding segmented invariant confluence, it does not seem easy to
> find good "quality" segments, which relates to how usable Section 6
> could be. The paper may include some discussions on how segments are
> constructed.

We already had a paragraph saying

@@
 
-\subsection{Discussion}
-There are a few things worth noting about segmented
+\subsection{Discussion and Limitations}
+There are a couple of things worth noting about segmented
 \invariantconfluence{}, its system model, and its decision procedure.
@@
 
-Second, we can integrate a couple of optimizations into our system model to
-further reduce the amount of coordination it requires. For example, servers can
-buffer transactions that require global coordination, executing an entire batch
-of these transactions during a single round of global coordination.
+% Second, we can integrate a couple of optimizations into our system model to
+% further reduce the amount of coordination it requires. For example, servers can
+% buffer transactions that require global coordination, executing an entire batch
+% of these transactions during a single round of global coordination.
+%
+% Third, a segmented \invariantconfluence{} decision procedure can also leverage
+% \thmref{LatticeProperty} in addition to
+% \algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
+% criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
+% I$.
 
-Third, a segmented \invariantconfluence{} decision procedure can also leverage
-\thmref{LatticeProperty} in addition to
-\algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
-criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
-I$.
-
-Fourth, while a segmented \invariantconfluence{} decision procedure can help
+Second, while a segmented \invariantconfluence{} decision procedure can help
 decide whether or not an object is segmented \invariantconfluent{}, it cannot
-currently help construct a segmentation. In the future, we plan to explore
-ways by which we can automatically suggest segmentations to the application
-designer.
+currently help construct a segmentation. It is the responsibility of the
+programmer to think of a segmentation that is amenable to segmented
+\invariantconfluence{}, and this can be an onerous process. In the future, we
+plan to explore ways by which we can automatically suggest segmentations to the
+application designer to ease this process.

### 2018-09-19 doc/paper/sections/segmented_invariant_confluence.tex
Moved some stuff to tech report.
The paper is a bit long. I'm starting to move some content out of the
paper and into an unoffical technical report. I introduced a techreport
environment to selectively remove or include these parts.

@@
 \subsection{Discussion and Limitations}
-There are a couple of things worth noting about segmented
-\invariantconfluence{}, its system model, and its decision procedure.
+There are a few things worth noting about segmented \invariantconfluence{}, its
+system model, and its decision procedure.
 %
@@
 
-% Second, segmented \invariantconfluence{} naturally subsumes a distributed
-% locking approach to replicating non-\invariantconfluent{} objects.  This
-% approach first recognizes which transactions cannot be safely executed
-% concurrently and then requires them to acquire a distributed lock before
-% executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
-% banking application with the invariant that all balances remain non-negative,
-% concurrent deposits are permitted, but concurrent withdrawals can lead to
-% invariant violations. Thus, we require that withdrawals acquire a distributed
-% lock before executing. This example is exactly the same as
-% \exampleref{CounreachableExample} which we handled by simply removing
-% withdrawal transactions from our segmentation's set of transactions.
-
-% Second, we can integrate a couple of optimizations into our system model to
-% further reduce the amount of coordination it requires. To begin, if a server with
-% state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t \in I_i$
-% to execute, and $t(s_i)$ violates the active invariant but not the global
-% invariant, instead of initiating a round of global coordination, $p_i$ can
-% simply buffer $t$ for re-execution at a later time. While this increases the
-% latency required to execute $t$, it's possible that after other transactions
-% are executed, re-executing $t$ may lead to a state that either satisfies the
-% active invariant or violates the global invariant. In either case, a round of
-% global coordination is avoided. Similarly, servers can buffer transactions that
-% require global coordination, executing an entire batch of these transactions
-% during a single round of global coordination.
-
-% Second, we can integrate a couple of optimizations into our system model to
-% further reduce the amount of coordination it requires. For example, servers can
-% buffer transactions that require global coordination, executing an entire batch
-% of these transactions during a single round of global coordination.
-%
-% Third, a segmented \invariantconfluence{} decision procedure can also leverage
-% \thmref{LatticeProperty} in addition to
-% \algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
-% criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
-% I$.
-
 Second, while a segmented \invariantconfluence{} decision procedure can help
@@
 application designer to ease this process.
+
+\begin{techreport}
+
+  Third, segmented \invariantconfluence{} naturally subsumes a distributed
+  locking approach to replicating non-\invariantconfluent{} objects. This
+  approach first recognizes which transactions cannot be safely executed
+  concurrently and then requires them to acquire a distributed lock before
+  executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
+  banking application with the invariant that all balances remain non-negative,
+  concurrent deposits are permitted, but concurrent withdrawals can lead to
+  invariant violations. Thus, we require that withdrawals acquire a distributed
+  lock before executing. This example is exactly the same as
+  \exampleref{CounreachableExample} which we handled by simply removing
+  withdrawal transactions from our segmentation's set of transactions.
+
+  Fourth, we can integrate a couple of optimizations into our system model to
+  further reduce the amount of coordination it requires. To begin, if a server
+  with state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t
+  \in I_i$ to execute, and $t(s_i)$ violates the active invariant but not the
+  global invariant, instead of initiating a round of global coordination, $p_i$
+  can simply buffer $t$ for re-execution at a later time. While this increases
+  the latency required to execute $t$, it's possible that after other
+  transactions are executed, re-executing $t$ may lead to a state that either
+  satisfies the active invariant or violates the global invariant. In either
+  case, a round of global coordination is avoided. Similarly, servers can
+  buffer transactions that require global coordination, executing an entire
+  batch of these transactions during a single round of global coordination.
+
+  Fifth, a segmented \invariantconfluence{} decision procedure can also
+  leverage \thmref{LatticeProperty} in addition to
+  \algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
+  criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
+  I$.
+\end{techreport}

### 2018-09-20 doc/paper/sections/abstract.tex
Integrating Joe's feedback.
Discussed limitations + shrunk paper.
From reviewer 11:

> Discussion of limitations and practical use can be made clearer.

From reviewer 12:

> While the ideas of the paper are very nice, the paper may more
> concretely describe how practical the algorithms are. For example, the
> first decision procedure (Algorithm 1) involves human to augment NR.
> In general, what are the tricky classes of transactions that may
> require lot of human interactions? Similar comments apply to segmented
> invariant confluence (Algorithm 5). What classes of
> transactions/applicatio

@@
 \begin{abstract}
-Strongly consistent distributed systems are easy to reason about but face fundamental
-limitations in availability and performance. Weakly consistent systems can be
-implemented with very high performance but place a burden on the application
-developer to reason about complex interleavings of execution.
+Strongly consistent distributed systems are easy to reason about but face
+fundamental limitations in availability and performance. Weakly consistent
+systems can be implemented with very high performance but place a burden on the
+application developer to reason about complex interleavings of execution.
 \Invariantconfluence{} provides a formal framework for understanding when we
@@
 However, actually determining whether or not an object is \invariantconfluent{}
-is challenging: undecidable in general. In this paper, we establish conditions under
-which a commonly used sufficient condition for \invariantconfluence{} is both
-necessary and sufficient and use this condition to design (a) a general-purpose
-interactive \invariantconfluence{} decision procedure and (b) a novel
-sufficient condition that can be checked automatically. We then take a step
-beyond \invariantconfluence{} and introduce a generalization of
-\invariantconfluence{} called segmented \invariantconfluence{}, which allows us
-to replicate non-\invariantconfluent{} objects with a small amount of
-coordination.
+is challenging: undecidable in general.
 
-We implemented these formalisms in a prototype called
-Lucy and found that our decision procedures efficiently handle common real-world workloads
-including foreign keys, rollups, escrow transactions, and more.  We also
-found that segmented \invariantconfluent{} replication can deliver up to an
-order of magnitude more throughput than linearizable replication for low
-contention workloads and comparable throughput for medium to high contention
-workloads.
+In this paper, we establish conditions under which a commonly used sufficient
+condition for \invariantconfluence{} is both necessary and sufficient and use
+this condition to design (a) a general-purpose interactive
+\invariantconfluence{} decision procedure and (b) a novel sufficient condition
+that can be checked automatically. We then take a step beyond
+\invariantconfluence{} and introduce a generalization of \invariantconfluence{}
+called segmented \invariantconfluence{}, which allows us to replicate
+non-\invariantconfluent{} objects with a small amount of coordination.
+
+We implemented these formalisms in a prototype called Lucy and found that our
+decision procedures efficiently handle common real-world workloads including
+foreign keys, rollups, escrow transactions, and more.  We also found that
+segmented \invariantconfluent{} replication can deliver up to an order of
+magnitude more throughput than linearizable replication for low contention
+workloads and comparable throughput for medium to high contention workloads.
 \end{abstract}

### 2018-09-20 doc/paper/sections/interactive_decision_procedure.tex
Last tweaks!
Discussed limitations + shrunk paper.
From reviewer 11:

> Discussion of limitations and practical use can be made clearer.

From reviewer 12:

> While the ideas of the paper are very nice, the paper may more
> concretely describe how practical the algorithms are. For example, the
> first decision procedure (Algorithm 1) involves human to augment NR.
> In general, what are the tricky classes of transactions that may
> require lot of human interactions? Similar comments apply to segmented
> invariant confluence (Algorithm 5). What classes of
> transactions/applications that may req

@@
   \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
-  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g. $xy \leq
+  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g., $xy \leq
   0$).
@@
 
+\subsection{Limitations}
+Our interactive \invariantconfluence{} decision procedure has two main
+limitations. First, \algoref{InteractiveDecisionProcedure} requires an
+\invariantclosure{} decision procedure, but determining \invariantclosure{} is
+undecidable in general.
+%
+\begin{techreport}
+
+  For example, let $O_p = (S, \join)$ where $S$ is the set of all programs and
+  $s_1 \join s_2 = p$ for some fixed program $p$. Letting $I$ be the set of all
+  programs that terminate, determining if $O_p$ is \Iclosed{} is tantamount to
+  determining if $p$ terminates.
+\end{techreport}
+%
+In practice, we can implement an \invariantclosure{} decision procedure using
+an SMT solver like Z3 that works well on simple objects, invariants, and merge
+functions (e.g., integers, tuples, infinite sets, bit vectors, linear
+constraints, basic arithmetic, tuple projection, basic set operations, bit
+arithmetic). But, SMT solvers are mostly unable to analyze more complex
+constructs (e.g., finite lists~\cite{kroning2009proposal}, graphs, recursive
+algebraic data types, nonlinear constraints, merge functions that contain loops
+or recursion).
+
+Second, \algoref{InteractiveDecisionProcedure} relies on a user to identify
+unreachable states. As we saw in \figref{DecisionProcedureExample}, the set of
+unreachable states can sometimes be clear, especially if there's a noticeable
+pattern in the set of reachable states. However, if the set of transactions is
+large or complex or if the merge function is complex, then reasoning about
+unreachable states can be difficult.  Unlike with reachable states---where
+verifying that a state is reachable only requires thinking of a single way to
+reach the state---verifying that a state is unreachable requires a programmer
+to reason about a large number of system executions and conclude that
+\emph{none} of them can lead to the state. In the future, we plan on exploring
+ways to assist a user reason about unreachable states and discover some
+unreachable states automatically.
+
+\begin{techreport}
   \subsection{Tolerating User Error}
   \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
-procedure. It requires that a user classify states as reachable or unreachable.
-To err is human, so what happens when a user incorrectly classifies a state?
-There are two possible errors that can be made, and
+  procedure. It requires that a user classify states as reachable or
+  unreachable.  To err is human, so what happens when a user incorrectly
+  classifies a state?  There are two possible errors that can be made, and
   \algoref{InteractiveDecisionProcedure} can be made robust to both.
@@
   \textbf{A user can label an unreachable state as reachable.}
-In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and return
-false, concluding that $O$ is not \sTIconfluent{} even when it is.  This is
-inconvenient, but not catastrophic. We can modify
-\algoref{InteractiveDecisionProcedure} so that \Helper{} requires that whenever
-a user labels a state $s$ as \sTIreachable{}, they must also provide an
-\sTIreachable{} expression $e$ that evaluates to $s$. Here, $e$ acts a proof
-that certifies that $s$ is indeed reachable. This increases the burden on the
-user but completely eliminates this type of user error.
+  In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and
+  return false, concluding that $O$ is not \sTIconfluent{} even when it is.
+  This is inconvenient, but not catastrophic. We can modify
+  \algoref{InteractiveDecisionProcedure} so that \Helper{} requires that
+  whenever a user labels a state $s$ as \sTIreachable{}, they must also provide
+  an \sTIreachable{} expression $e$ that evaluates to $s$. Here, $e$ acts a
+  proof that certifies that $s$ is indeed reachable. This increases the burden
+  on the user but completely eliminates this type of user error.
 
   \textbf{A user can label a reachable state as unreachable.}
-In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is not
-\sTIconfluent. Thus, a user might falsely believe their distributed object to
-be \sTIconfluent{} even though it isn't, and eventually one replica of their
-distributed object might enter a state that violates the invariant.
+  In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is
+  not \sTIconfluent. Thus, a user might falsely believe their distributed
+  object to be \sTIconfluent{} even though it isn't, and eventually one replica
+  of their distributed object might enter a state that violates the invariant.
   %
-We can mitigate this in two ways. First, we can aggressively expand $R$ automatically. If a
-user ever labels a state $s$ as unreachable, but $s \in R$, we can notify the
-user of their mistake. Second, \Helper{} returns true when $O$ is \iclosed{(I -
-NR)} for some $NR$, so $O$ \emph{is} \sticonfluent{s_0}{T}{I-NR} (even though
-it might not be \sTIconfluent{}). Thus, when a user replicates their
-distributed object across a set of servers, they can deploy with the invariant
-$I - NR$ instead of $I$. If $NR$ is correct and only contains unreachable
-states, then deploying with $I - NR$ is equivalent to deploying with $I$. If
-$NR$ is incorrect and contains some \sTIreachable{} states, then some of these
-states are artificially made unreachable, but the system is still guaranteed to
-never produce a state which violates $I$.
+  We can mitigate this in two ways. First, we can aggressively expand $R$
+  automatically. If a user ever labels a state $s$ as unreachable, but $s \in
+  R$, we can notify the user of their mistake. Second, \Helper{} returns true
+  when $O$ is \iclosed{(I - NR)} for some $NR$, so $O$ \emph{is}
+  \sticonfluent{s_0}{T}{I-NR} (even though it might not be \sTIconfluent{}).
+  Thus, when a user replicates their distributed object across a set of
+  servers, they can deploy with the invariant $I - NR$ instead of $I$. If $NR$
+  is correct and only contains unreachable states, then deploying with $I - NR$
+  is equivalent to deploying with $I$. If $NR$ is incorrect and contains some
+  \sTIreachable{} states, then some of these states are artificially made
+  unreachable, but the system is still guaranteed to never produce a state
+  which violates $I$.
+\end{techreport}

### 2018-09-20 doc/paper/sections/merge_reduction.tex
Discussed limitations + shrunk paper.
From reviewer 11:

> Discussion of limitations and practical use can be made clearer.

From reviewer 12:

> While the ideas of the paper are very nice, the paper may more
> concretely describe how practical the algorithms are. For example, the
> first decision procedure (Algorithm 1) involves human to augment NR.
> In general, what are the tricky classes of transactions that may
> require lot of human interactions? Similar comments apply to segmented
> invariant confluence (Algorithm 5). What classes of
> transactions/applications that may require complex

@@
 
+\begin{techreport}
   The proof of \thmref{ReducibilityImpliesIconfluence} is a straightforward
@@
   equivalent merge-free reachable subexpression (which we can do because $O$ is
-merge-reducible). We repeat this process until $e$ has been completely replaced
-with an equivalent merge-free reachable expression $e'$. Because $I(s_0)$ and
-because our system model only executes transactions that preserve the
-invariant, $e'$ (and hence $e$) is guaranteed to satisfy the invariant. Thus,
-all reachable states satisfy the invariant, so $O$ is \invariantconfluent{}.
+  merge-reducible). We repeat this process until $e$ has been completely
+  replaced with an equivalent merge-free reachable expression $e'$. Because
+  $I(s_0)$ and because our system model only executes transactions that
+  preserve the invariant, $e'$ (and hence $e$) is guaranteed to satisfy the
+  invariant. Thus, all reachable states satisfy the invariant, so $O$ is
+  \invariantconfluent{}.
+\end{techreport}
 
@@
 
+\begin{techreport}
   The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
@@
   is reachable and evaluates to $e_1 \join e_2$.
+\end{techreport}
 
@@
 
-  Clearly, criteria (1), (2), and (4) are satisfied. Criterion (3) is also
-  satisfied. If $s$ is an arbitrary state that satisfies $I$, then it is
-  non-empty and contains, without loss of generality, only even integers. If
-  $t_1$ and $t_2$ are arbitrary transactions such that $I(t_1(s))$ and
-  $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty and contain only
-  even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly non-empty and contains
-  only even integers, so $I(t_1(s) \join t_2(s))$.
-  %
-  However, $O$ is not \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$.
-  Then, $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$,
-  $\lnot I(s_3)$.
+  Criteria (1), (2), (3) and (4) are all satisfied. However, $O$ is not
+  \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$. Then, $I(s_1)$ and
+  $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$, $\lnot I(s_3)$.
+
+  \begin{techreport}
+
+    Here's why criterion (3) is satisfied. If $s$ is an arbitrary state that
+    satisfies $I$, then it is non-empty and contains, without loss of generality,
+    only even integers. If $t_1$ and $t_2$ are arbitrary transactions such that
+    $I(t_1(s))$ and $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty
+    and contain only even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly
+    non-empty and contains only even integers, so $I(t_1(s) \join t_2(s))$.
+  \end{techreport}
 \end{example}

### 2018-09-20 doc/paper/sections/abstract.tex
Minor typo fixes, grammar fixes, and rewording.

@@
 However, actually determining whether or not an object is \invariantconfluent{}
-is challenging: undecidable in general.
+is challenging.
 
 In this paper, we establish conditions under which a commonly used sufficient
-condition for \invariantconfluence{} is both necessary and sufficient and use
+condition for \invariantconfluence{} is both necessary and sufficient, and we use
 this condition to design (a) a general-purpose interactive
@@
 that can be checked automatically. We then take a step beyond
-\invariantconfluence{} and introduce a generalization of \invariantconfluence{}
-called segmented \invariantconfluence{}, which allows us to replicate
+\invariantconfluence{} and introduce a generalization of \invariantconfluence{},
+called segmented \invariantconfluence{}, that allows us to replicate
 non-\invariantconfluent{} objects with a small amount of coordination.

### 2018-09-20 doc/paper/sections/evaluation.tex
Minor typo fixes, grammar fixes, and rewording.

@@
 in \algoref{InteractiveDecisionProcedure}, an implementation of a decision
-procedure which checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
+procedure that checks criteria (1) - (4) from \thmref{LatticeProperty}, and an
 implementation of the decision procedure described in

### 2018-09-20 doc/paper/sections/interactive_decision_procedure.tex
Minor typo fixes, grammar fixes, and rewording.

@@
 interactive \invariantconfluence{} decision procedure shown in
-\algoref{InteractiveDecisionProcedure} which takes advantage of this result.
+\algoref{InteractiveDecisionProcedure}, that takes advantage of this result.
 
@@
 \join)$, a start state $s_0$, a set of transactions $T$, and an invariant
-$I$\footnote{%
-  Formally, $I$ is a (possibly infinite) set of states (e.g. $\setst{(x, y) \in
-  \ints \times \ints}{xy \leq 0}$), but in practice, our implementation of
-  \IsInvConfluent{} takes in $I$ as a finite logical formula (e.g., $xy \leq
-  0$).
-}. The user then interacts with
-\algoref{InteractiveDecisionProcedure} to iteratively eliminate unreachable
-states from the invariant. Meanwhile, the algorithm leverages an
-\invariantclosure{} decision procedure to either (a) conclude that $O$ is or is
-not \sTIconfluent{} or (b) provide counterexamples to the user to help them
-eliminate unreachable states. After all unreachable states have been eliminated
-from the invariant, \thmref{IclosureEquivalentIconfluence} allows us to reduce
-the problem of \invariantconfluence{} directly to the problem of
-\invariantclosure{}, and the algorithm terminates.
+$I$. The user then interacts with \algoref{InteractiveDecisionProcedure} to
+iteratively eliminate unreachable states from the invariant. Meanwhile, the
+algorithm leverages an \invariantclosure{} decision procedure to either (a)
+conclude that $O$ is or is not \sTIconfluent{} or (b) provide counterexamples
+to the user to help them eliminate unreachable states. After all unreachable
+states have been eliminated from the invariant,
+\thmref{IclosureEquivalentIconfluence} allows us to reduce the problem of
+\invariantconfluence{} directly to the problem of \invariantclosure{}, and the
+algorithm terminates.
 %
@@
 \IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
-\IsIclosed$(O, I)$ which returns a triple (closed, $s_1$, $s_2$). closed is a
+\IsIclosed$(O, I)$ that returns a triple (closed, $s_1$, $s_2$). closed is a
 boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
@@
 \sTIconfluent{} and \IsInvConfluent{} returns false. Otherwise,
-\IsInvConfluent{} calls a helper function \Helper{} which---in addition to $O$,
+\IsInvConfluent{} calls a helper function \Helper{} that---in addition to $O$,
 $s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
@@
 which, as we just showed, is a subset of $I$.  Thus, $O$ is \sTIconfluent{}, so
-\Helper{} return true.
+\Helper{} returns true.
 
@@
 \subsection{Limitations}
-Our interactive \invariantconfluence{} decision procedure has two main
+Our interactive \invariantconfluence{} decision procedure has two
 limitations. First, \algoref{InteractiveDecisionProcedure} requires an
@@
 an SMT solver like Z3 that works well on simple objects, invariants, and merge
-functions (e.g., integers, tuples, infinite sets, bit vectors, linear
+operators (e.g., integers, tuples, infinite sets, bit vectors, linear
 constraints, basic arithmetic, tuple projection, basic set operations, bit
@@
 constructs (e.g., finite lists~\cite{kroning2009proposal}, graphs, recursive
-algebraic data types, nonlinear constraints, merge functions that contain loops
+algebraic data types, nonlinear constraints, merge operators that contain loops
 or recursion).
@@
 pattern in the set of reachable states. However, if the set of transactions is
-large or complex or if the merge function is complex, then reasoning about
+large or complex or if the merge operator is complex, then reasoning about
 unreachable states can be difficult.  Unlike with reachable states---where
@@
 \emph{none} of them can lead to the state. In the future, we plan on exploring
-ways to assist a user reason about unreachable states and discover some
+ways to help a user reason about unreachable states and ways to discover sets of
 unreachable states automatically.
@@
   unreachable, but the system is still guaranteed to never produce a state
-  which violates $I$.
+  that violates $I$.
 \end{techreport}

### 2018-09-20 doc/paper/sections/segmented_invariant_confluence.tex
Minor typo fixes, grammar fixes, and rewording.

@@
 non-\invariantconfluent{} object with segmented \invariantconfluence{} can
-achieve between 100x and 10x more throughput than linearizable replication for
+achieve between 10x and 100x more throughput than linearizable replication for
 certain workloads.
@@
 T_1), \ldots, (I_n, T_n)$ is a sequence (not a set) of $n$ segments $(I_i,
-T_i)$ where every $T_i$ is a subset of $T$. $O$ is \defword{segmented
-\invariantconfluent{}} with respect to $s_0$, $T$, $I$, and $\Sigma$,
-abbreviated \defword{\sTISconfluent}, if the following conditions hold:
+T_i)$ where every $T_i$ is a subset of $T$ and every $I_i \subseteq S$ is an
+invariant. $O$ is \defword{segmented \invariantconfluent{}} with respect to
+$s_0$, $T$, $I$, and $\Sigma$, abbreviated \defword{\sTISconfluent}, if the
+following conditions hold:
 \begin{itemize}
@@
 \end{theorem}
-\begin{proof}
+
 The proof of the forward direction is exactly the same as the proof of
@@
 \sTIreachable{}, so it satisfies the invariant.
-\end{proof}
 
@@
   transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
-  \IsIclosed{} again returns false with same states $s_1$ and $s_2$. This time,
+  \IsIclosed{} again returns false with the same states $s_1$ and $s_2$. This time,
   however, the user recognizes that the two states are not
@@
 
-Second, while a segmented \invariantconfluence{} decision procedure can help
+Second, while our segmented \invariantconfluence{} decision procedure can help
 decide whether or not an object is segmented \invariantconfluent{}, it cannot
@@
 programmer to think of a segmentation that is amenable to segmented
-\invariantconfluence{}, and this can be an onerous process. In the future, we
+\invariantconfluence{}. This can be an onerous process. In the future, we
 plan to explore ways by which we can automatically suggest segmentations to the

### 2018-09-20 doc/paper/sections/interactive_decision_procedure.tex
Small tweaks to avoid column overflow.

@@
 \IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
-\IsIclosed$(O, I)$ that returns a triple (closed, $s_1$, $s_2$). closed is a
-boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
-$s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a counterexample
-witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
-but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As
-we mentioned earlier, we can (and do) implement the \invariantclosure{}
-decision procedure using an SMT solver like Z3~\cite{de2008z3}.
+\IsIclosed$(O, I)$. \IsIclosed$(O, I)$ returns a triple (closed, $s_1$, $s_2$).
+closed is a boolean indicating whether $O$ is \Iclosed{}. If closed is true,
+then $s_1$ and $s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a
+counterexample witnessing the fact that $O$ is not \Iclosed{}. That is,
+$I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from
+\exampleref{Z2}).  As we mentioned earlier, we can (and do) implement the
+\invariantclosure{} decision procedure using an SMT solver like
+Z3~\cite{de2008z3}.
 

### 2019-08-15 doc/vldb_slides/iconfluence.tex
Progress towards VLDB slides!

\documentclass[xcolor={dvipsnames,svgnames,table}]{beamer}

\usepackage{amsmath}
\usepackage{centernot}
\usepackage{etoolbox}
\usepackage{mathtools}
\usepackage{python}
\usepackage{slides}
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}
\usetikzlibrary{calc}
\usetikzlibrary{positioning}

\title{Interactive Checks for Coordination Avoidance}
\author{Michael Whittaker, Joseph M. Hellerstein}
\date{}

\begin{document}
\begin{frame}
  \maketitle
\end{frame}

\tikzstyle{pic}=[inner sep=0pt]
\tikzstyle{quote}=[fill=gray!30, inner sep=12pt]

\newcommand{\vogels}{
  \node[pic] (vogels) at (0, 0) {
    \includegraphics[
      width=0.22\textwidth,
      trim=0 75 0 25,
      clip
    ]{assets/vogels.jpg}
  };
}

\newcommand{\hoare}{
  \node[pic, right=0.2cm of vogels] (hoare) {
    \includegraphics[width=0.22\textwidth]{assets/hoare.jpg}
  };
}

\newcommand{\lamport}{
  \node[pic, right=0.2cm of hoare] (lamport) {
    \includegraphics[
      width=0.22\textwidth,
      trim=250 700 300 200,
      clip
    ]{assets/lamport.jpg}
  };
}

\newcommand{\brewer}{
  \node[pic, right=0.2cm of lamport] (brewer) {
    \includegraphics[
      width=0.22\textwidth,
      trim=100 300 100 100,
      clip
    ]{assets/brewer.jpg}
  };
}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{You should use weak consistency. It's wicked fast!}
      };
      \path[fill=gray!30] (vogels.north) -- ++(0, 1) -- ++(0.5, 0) -- (vogels.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
        \Large\textit{What about the application invariants?!}
      };
      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
      };
      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \brewer
      \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
        \Large\textit{Not so fast!}
      };
      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \node[pic] (bailis) at (0, 0) {\includegraphics[width=0.25\textwidth]{assets/bailis.jpg}};
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{?`Porque no los dos?}
      };
      \path[fill=gray!30] (bailis.north) -- ++(0, 1.5) -- ++(1, 0) -- (bailis.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\end{document}

### 2019-08-15 doc/vldb_slides/iconfluence.tex
Added motivating bank example.

@@
 
-\tikzstyle{pic}=[inner sep=0pt]
-\tikzstyle{quote}=[fill=gray!30, inner sep=12pt]
-
-\newcommand{\vogels}{
-  \node[pic] (vogels) at (0, 0) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=0 75 0 25,
-      clip
-    ]{assets/vogels.jpg}
-  };
-}
-
-\newcommand{\hoare}{
-  \node[pic, right=0.2cm of vogels] (hoare) {
-    \includegraphics[width=0.22\textwidth]{assets/hoare.jpg}
-  };
-}
-
-\newcommand{\lamport}{
-  \node[pic, right=0.2cm of hoare] (lamport) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=250 700 300 200,
-      clip
-    ]{assets/lamport.jpg}
-  };
-}
-
-\newcommand{\brewer}{
-  \node[pic, right=0.2cm of lamport] (brewer) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=100 300 100 100,
-      clip
-    ]{assets/brewer.jpg}
-  };
-}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{You should use weak consistency. It's wicked fast!}
-      };
-      \path[fill=gray!30] (vogels.north) -- ++(0, 1) -- ++(0.5, 0) -- (vogels.north);
-    \end{tikzpicture}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
-        \Large\textit{What about the application invariants?!}
-      };
-      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
-    \end{tikzpicture}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \lamport
-      \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
-        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
-      };
-      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
-    \end{tikzpicture}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \lamport
-      \brewer
-      \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
-        \Large\textit{Not so fast!}
-      };
-      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
-    \end{tikzpicture}
-  \end{center}
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \node[pic] (bailis) at (0, 0) {\includegraphics[width=0.25\textwidth]{assets/bailis.jpg}};
-      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{?`Porque no los dos?}
-      };
-      \path[fill=gray!30] (bailis.north) -- ++(0, 1.5) -- ++(1, 0) -- (bailis.north);
-    \end{tikzpicture}
-  \end{center}
-\end{frame}
-
+{\input{sections/four_wise_men.tex}}
+{\input{sections/bank_example.tex}}
 \end{document}

### 2019-08-15 doc/vldb_slides/sections/bank_example.tex
Added motivating bank example.

\newcommand{\bankdistance}{4}
\tikzstyle{bank}=[
  draw,
  minimum width=1.5cm,
  minimum height=1cm,
  font=\huge
]
\tikzstyle{op}=[
  -latex,
  ultra thick
]

\newcommand{\drawbanks}[6]{{
  \newcommand{\banka}{#1}
  \newcommand{\bankb}{#2}
  \newcommand{\bankaa}{#3}
  \newcommand{\bankbb}{#4}
  \newcommand{\bankaaa}{#5}
  \newcommand{\bankbbb}{#6}

  \node[bank, label=above:{Replica $A$}] (a1) at (0,              0) {\banka};
  \node[bank                           ] (a2) at (0,             -3) {\bankaa};
  \node[bank                           ] (a3) at (0,             -6) {\bankaaa};
  \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\bankb};
  \node[bank                           ] (b2) at (\bankdistance, -3) {\bankbb};
  \node[bank                           ] (b3) at (\bankdistance, -6) {\bankbbb};

  \draw[op] (a1) -- (a2);
  \draw[op] (a2) -- (a3);
  \draw[op] (b1) -- (b2);
  \draw[op] (b2) -- (b3);
}}

\newcommand{\zigzag}[2]{{
  \newcommand{\banka}{#1}
  \newcommand{\bankb}{#2}

  \draw[red, op] ($(\banka) + (0, -0.7)$) -- ($(\bankb) + (0, -1.0)$);
  \draw[red, op] ($(\bankb) + (0, -1.0)$) -- ($(\banka) + (0, -1.3)$);
  \draw[red, op] ($(\banka) + (0, -1.3)$) -- ($(\bankb) + (0, -1.6)$);
  \draw[red, op] ($(\bankb) + (0, -1.6)$) -- ($(\banka) + (0, -1.9)$);
}}

\begin{frame}{Replicated Bank Account}
  \begin{center}
    \begin{tikzpicture}
      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance, 0) {\$50};
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \drawbanks{\$50}{\$50}
                {\$60}{\$60}
                {\$80}{\$80}

      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
      \draw[op] (b3) -- ($(b3) + (+2, 0)$) node[midway, above] {OK};

      \zigzag{a1}{b1}
      \zigzag{a2}{b2}
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Coordinating}
  \begin{center}
    \begin{tikzpicture}
      \drawbanks{\$50}{\$50}
                {\$20}{\$20}
                {\$20}{\$20}

      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
      \draw[op] (b3) -- ($(b3) + (+2, 0)$) node[midway, above] {NO};

      \zigzag{a1}{b1}
      \zigzag{a2}{b2}
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \drawbanks{\$50}{\$50}
                {\$60}{\$70}
                {\$80}{\$80}

      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
      \draw[op] (b2) -- ($(b2) + (+2, 0)$) node[midway, above] {OK};

      \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$+\$10$};
      \draw[dashed, op] (b2) -- (a3) node[near start, sloped, above] {$+\$20$};
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}{Avoiding Coordination}
  \begin{center}
    \begin{tikzpicture}
      \drawbanks{\$50}{\$50}
                {\$20}{\$10}
                {-\$20}{-\$20}

      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
      \draw[op] (b2) -- ($(b2) + (+2, 0)$) node[midway, above] {OK};

      \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$-\$30$};
      \draw[dashed, op] (b2) -- (a3) node[near start, sloped, above] {$-\$40$};
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \Huge
    Deposits don't require coordination to maintain invariants, but withdrawals
    do.
  \end{center}
\end{frame}

### 2019-08-15 doc/vldb_slides/sections/four_wise_men.tex
Added motivating bank example.

\tikzstyle{pic}=[inner sep=0pt]
\tikzstyle{quote}=[fill=gray!30, inner sep=12pt]

\newcommand{\vogels}{
  \node[pic] (vogels) at (0, 0) {
    \includegraphics[
      width=0.22\textwidth,
      trim=0 75 0 25,
      clip
    ]{assets/vogels.jpg}
  };
}

\newcommand{\hoare}{
  \node[pic, right=0.2cm of vogels] (hoare) {
    \includegraphics[width=0.22\textwidth]{assets/hoare.jpg}
  };
}

\newcommand{\lamport}{
  \node[pic, right=0.2cm of hoare] (lamport) {
    \includegraphics[
      width=0.22\textwidth,
      trim=250 700 300 200,
      clip
    ]{assets/lamport.jpg}
  };
}

\newcommand{\brewer}{
  \node[pic, right=0.2cm of lamport] (brewer) {
    \includegraphics[
      width=0.22\textwidth,
      trim=100 300 100 100,
      clip
    ]{assets/brewer.jpg}
  };
}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{You should use weak consistency. It's wicked fast!}
      };
      \path[fill=gray!30] (vogels.north) -- ++(0, 1) -- ++(0.5, 0) -- (vogels.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
        \Large\textit{What about the application invariants?!}
      };
      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
      };
      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \vogels
      \hoare
      \lamport
      \brewer
      \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
        \Large\textit{Not so fast!}
      };
      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \begin{tikzpicture}[xscale=1.3]
      \node[pic] (bailis) at (0, 0) {\includegraphics[width=0.25\textwidth]{assets/bailis.jpg}};
      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
        \Large\textit{?`Porque no los dos?}
      };
      \path[fill=gray!30] (bailis.north) -- ++(0, 1.5) -- ++(1, 0) -- (bailis.north);
    \end{tikzpicture}
  \end{center}
\end{frame}

### 2019-08-15 doc/vldb_slides/iconfluence.tex
Added invariant confluence definitions.

@@
 \usepackage{etoolbox}
+\usepackage{invariantconfluence}
+\usepackage{mathpartir}
 \usepackage{mathtools}
+\usepackage{pervasives}
 \usepackage{python}
@@
 {\input{sections/bank_example.tex}}
+{\input{sections/iconfluence.tex}}
+{\input{sections/iclosure.tex}}
+{\input{sections/decision_procedure.tex}}
+{\input{sections/segmented_iconfluence.tex}}
+{\input{sections/evaluation.tex}}
 \end{document}

### 2019-08-15 doc/vldb_slides/sections/iclosure.tex
Finished iclosure section.

\begin{frame}
  \Huge
  \begin{center}
    Goal: develop an invariant-confluence decision procedure.
  \end{center}
\end{frame}

\begin{frame}
  \Huge
  \begin{center}
    Reasoning about reachable states is hard.
  \end{center}
\end{frame}

\begin{frame}
  \Large
  We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
  y) \in S$. \pause For example,
  \begin{itemize}
    \item Even numbers are closed under addition.
    \item Odd numbers are \emph{not} closed under addition.
  \end{itemize}
\end{frame}

\begin{frame}
  \Large
  $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to an invariant
  $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying states are
  closed under merge. \pause That is, for every state $s_1, s_2 \in S$, if
  $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \implies \text{invariant confluence}
  \]
  \pause
  Transactions perserve invariants. If merging does too, then all reachable
  states satisfy the invariant.
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \implies \text{invariant confluence}
  \]
  \pause
  Transactions perserve invariants. If merging does too, then all reachable
  states satisfy the invariant.
\end{frame}

\begin{frame}
  \Large
  Checking invariant closure is more straightforward.
  \pause
  \begin{align*}
    \forall x_1, & y_1, x_2, y_2.\, \\
    \quad & x_1y_1 \leq 0 \land x_2y_2 \leq 0 \implies \\
    \quad & \max(x_1, x_2)\max(y_1, y_2) \leq 0
  \end{align*}
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \implies \text{invariant confluence}
  \]
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure}
    \xLeftarrow{\phantom{a}?\phantom{a}}
    \text{invariant confluence}
  \]
\end{frame}

\newcommand{\xmin}{-2}
\newcommand{\xmax}{2}
\newcommand{\ymin}{-2}
\newcommand{\ymax}{2}

% Axes.
\newcommand{\xyaxes}{
  \draw[] (\xmin.5, 0) to (\xmax.5, 0);
  \draw[] (0, \ymin.5) to (0, \ymax.5);
  \node at (\xmax + 1, 0) {$x$};
  \node at (0, \ymax + 1) {$y$};
}

% Quadrant 1.
\newcommand{\quadi}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {0, ..., \argxmax} {
    \foreach \y in {0, ..., \argymax} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 2.
\newcommand{\quadii}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {\argxmin, ..., 0} {
    \foreach \y in {0, ..., \argymax} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 3.
\newcommand{\quadiii}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {\argxmin, ..., 0} {
    \foreach \y in {\argymin, ..., 0} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% Quadrant 4.
\newcommand{\quadiv}[5]{{
  \newcommand{\argstyle}{#1}
  \newcommand{\argxmin}{#2}
  \newcommand{\argxmax}{#3}
  \newcommand{\argymin}{#4}
  \newcommand{\argymax}{#5}
  \foreach \x in {0, ..., \argxmax} {
    \foreach \y in {\argymin, ..., 0} {
      \node[\argstyle] (\x-\y) at (\x, \y) {};
    }
  }
}}

% State labels.
\newcommand{\statelabels}{
  \node[statelabel] at (0, 0) {$s_0$};
  \node[statelabel] at (-1, 1) {$s_1$};
  \node[statelabel] at (1, -1) {$s_2$};
  \node[statelabel] at (1, 1) {$s_3$};
}

\tikzstyle{point}=[shape=circle, fill=flatgray, inner sep=3pt]
\tikzstyle{inv}=[line width=0.75pt, draw=black]
\tikzstyle{pointinv}=[point, inv]
\tikzstyle{invregion}=[rounded corners, fill=flatgreen!50, draw=none]
\tikzstyle{reachableregion}=[rounded corners, fill=flatblue!50, draw=none]
\tikzstyle{statelabel}=[anchor=south west, inner sep=1pt]

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}

        \draw[-latex, ultra thick] (-1, 2) to (2, 2);
        \draw[-latex, ultra thick] (2, -1) to (2, 2);
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \Large
  \[
    \text{invariant closure} \centernot\impliedby \text{invariant confluence}
  \]
\end{frame}

\begin{frame}
  \begin{columns}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (\xmin.5, \ymax.5) rectangle (\xmax.5, \ymin.5);
          \draw[invregion] (\xmin.9, \ymax.9) rectangle (0.5, -0.5);
          \draw[invregion] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
          \draw (0.5, 0.5) to (0.5, \ymax.5);
          \draw (0.5, 0.5) to (\xmax.5, 0.5);
          \draw (-0.5, -0.5) to (-0.5, \ymin.5);
          \draw (-0.5, -0.5) to (\xmin.5, -0.5);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}

        \draw[-latex, ultra thick] (-1, 2) to (2, 2);
        \draw[-latex, ultra thick] (2, -1) to (2, 2);
      \end{tikzpicture}

      {\Huge Invariant}
    \end{column}
    \begin{column}{0.5\textwidth}
      \centering
      \begin{tikzpicture}[scale=1]
        \begin{scope}
          \clip (-1, 1) rectangle (\xmax.5, \ymin.5);
          \draw[reachableregion, draw=black] (-0.5, 0.5) rectangle (\xmax.9, \ymin.9);
        \end{scope}

        \xyaxes{}
        \quadi{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadii{point}{\xmin}{\xmax}{\ymin}{\ymax}
        \quadiv{pointinv}{\xmin}{\xmax}{\ymin}{\ymax}
      \end{tikzpicture}

      {\Huge Reachable}
    \end{column}
  \end{columns}
\end{frame}

\begin{frame}
  \Large
  If $I$ is a subset of reachable states, then
  \[
    \text{invariant closure} \iff \text{invariant confluence}
  \]
  \pause
  \begin{itemize}
    \item
      Forward direction: same as before.
    \item
      Backward direction: $I$ is equal to the set of reachable states.
      Reachable states are closed under merge, so so is $I$.
  \end{itemize}
\end{frame}

### 2019-08-15 doc/vldb_slides/sections/evaluation.tex
First crack at all slides done.

\begin{frame}[fragile]
  \begin{verbatim}
checker = InteractiveInvariantConfluenceChecker()
x = checker.int_max('x', 0) # An int, x, merged by max.
y = checker.int_max('y', 0) # An int, y, merged by max.
checker.add_transaction('inc_x', [x.assign(x + 1)])
checker.add_transaction('dec_y', [y.assign(y - 1)])
checker.add_invariant(x * y <= 0)
checker.check()\end{verbatim}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Foreign keys
    \item Auction application
    \item Escrow transactions
    \item TPC-C
  \end{itemize}

  \pause
  All run in less than a second and are implemented in less than 75 lines of
  specification.
\end{frame}

\begin{frame}
  \includegraphics[width=\textwidth]{assets/throughput_vs_fraction_16.pdf}
\end{frame}

\begin{frame}
  \begin{center}
    \Huge
    Come to our poster tomorrow night!
  \end{center}
\end{frame}

\begin{frame}
  \begin{center}
    \Huge
    Thank you
  \end{center}
\end{frame}

### 2019-08-16 doc/vldb_slides/sections/bank_example.tex
Started writing speaker notes.

@@
   \end{center}
+
+  \note{%
+    So, what is invariant confluence? Well, let's begin with an example.
+
+    Let's say we want to replicate the database shown here. This database is
+    the world's simplest database. It contains only a single number, which
+    represents my bank account balance. We decide to replicate this database
+    across two machines, replica A and replica B.
+
+    Users can issue transactions to the database, and every transacation is
+    either a deposit or a withdrawal. And, I have one critical invariant that I
+    want to maintain over my database. And that invariant is that my bank
+    account balance should be non-negative at all times. I'm not allowed to go
+    in the hole.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    One way to replicate our database is by using coordination to implement
+    strong consistency. For example, imagine a user issues a deposit of 10
+    dollars to replica A, and simultaneously, a user issues a deposit of 20
+    dollars to replica B. The two replicas can coordinate, for example by
+    running Paxos, and agree on a single total order in which to execute the
+    two transactions. Here, the replicas agree to deposit 10 dollars and then
+    deposit 20 dollars. Both transactions succeed, and the replicas send back OK
+    messages to the clients letting them know everything went smoothly.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Now, assume two clients issue concurrent withdrawal transactions of 30
+    dollars and 40 dollars. Again, the two replicas coordinate and agree to
+    execute the withdrawal of 30 dollars and then the withdrawal of 40 dollars.
+
+    The withdrawal of 30 dollars executes fine, brining my bank account balance
+    to 20 dollars, but the withdrawal of 40 dollars fails because I only have
+    20 dollars left. So replica A sends back an OK message to the client, but
+    replica B sends back a NO message letting the client know the transaction
+    failed. And while the transaction failed, you'll note that we never
+    violated our application invariant.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Now, instead of implementing our bank with strong consistency, we could
+    instead implement it with weak consistency and avoid coordination
+    altogether.
+
+    Now, when two clients issue concurrent deposit transactions, the two
+    replicas execute the transactions immediately without consulting one
+    another. And every once in a while, the two replicas communicate and fill
+    each other in on the transactions they missed.
+
+    Here, for example, replica A lets replica B know to deposit 10 dollars and
+    replica B lets replica A know to deposit 20 dollars. Notice that even
+    though we didn't perform any coordination, the two replicas never violate
+    my critical invariant of a non-negative balance.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    However, let's look what happens when two clients issue concurrent
+    withdrawals. Both withdrawals are executed locally and both succeed.
+    However, when the two replicas exchange transactions with one another, they
+    both enter an invariant violating state of negative twenty dollars!
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Now, these are just two examples, but if you think about it long enough,
+    they are representative of a trend. We can implement deposit transactions
+    without any coordination and we're guaranteed to never violate the
+    invariant. Withdrawal transactions, on the other hand, do require
+    coordination to avoid violating the invariant.
+
+    At a very high level, this is exactly the notion that invariant confluence
+    captures. We'll see that deposit transactions are invariant confluent and
+    thus do not require coordination while withdrawal transactions are not
+    invariant confluent and thus do require coordination.
+  }
 \end{frame}

### 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex
Started writing speaker notes.

@@
 \begin{frame}
+  \note{%
+    Hi everyone! Thanks so much for attending. I'm Michael, and today I'll be
+    talking about ``Interactive Checks for Coordination Avoidance''. This is
+    joint work with my advisor at UC Berkeley, Joe Hellerstein.
+
+    Imagine you're an application developer and you're trying to decide what
+    kind of consistency you'd like to use to replicate some of your data.
+  }
+\end{frame}
+
+\begin{frame}
   \begin{center}
@@
   \end{center}
+
+  \note{%
+    Someone like Werner Vogels would probably tell you to use weak consistency.
+    You can implement weak consistency super duper fast, and for many
+    applications, weak consistency is good enough.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
+    settle for weak consistency because a weakly consistent system is hard to
+    reason about. It robs us of a fundamental way we use to reason about
+    programs: application invariants. Even if every transaction in a weakly
+    consistent system preserves application invariants, the whole system may
+    not.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Leslie Lamport chimes in agreeingly. He encourages us to choose strong
+    consistency. Strongly consistent systems are much easier to reason about.
+    After all a strongly consistent system to end users looks exactly like a
+    system that's not even distributed.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
+    partition tolerance, oh my. Surely, strong consistency comes at the price
+    of availability and performance.
+
+    So we're in a bit of a pickle. Weak consistency is too confusing and strong
+    consistency is too heavy handed.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Thankfully, Peter Bailis comes to save the day and reminds us that perhaps
+    we can have the best of both worlds. In 2014, Peter and some other smart
+    folks defined invariant confluence as a condition under which we can
+    replicate a piece of data without the need for strongly consistent
+    coordination while maintaining the data's application invariants.
+
+    Invariant confluence will be the topic of this talk. In particular, I'll
+    define invariant confluence and then discuss how we can check invariant
+    confluence automatically.
+  }
 \end{frame}

### 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex
More speaker notes.

@@
     we can have the best of both worlds. In 2014, Peter and some other smart
-    folks defined invariant confluence as a condition under which we can
-    replicate a piece of data without the need for strongly consistent
-    coordination while maintaining the data's application invariants.
+    folks including Alan defined invariant confluence as a condition under
+    which we can replicate a piece of data without the need for strongly
+    consistent coordination while maintaining the data's application
+    invariants.
 

### 2019-08-16 doc/vldb_slides/sections/iclosure.tex
Added invariant closure notes.

@@
   \end{center}
+
+  \note{%
+    Now that we've defined invariant confluence, we can turn our attention to
+    the main goal of our paper. And that is to develop an invariant-confluence
+    decision procedure. Determining whether or not a distributed object is
+    invariant confluent is not easy to do by hand, so we'd like to develop a
+    decision procedure that can automatically check whether something is
+    invariant confluent for us.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Unfortunately, developing an invariant confluence decision procedure
+    straight up is not easy. Invariant confluence is fundamentally a property
+    about reachable states but reasoning automatically about reachable states
+    is hard. I'm not saying it's impossible, but certainly I don't know how to
+    do it.
+
+    Instead of reasoning about invariant confluence directly then, let's look
+    at a sufficient condition for invariant confluence called invariant
+    closure.
+  }
 \end{frame}
@@
   \end{itemize}
+
+  \note{%
+    First, a quick refresher on terminology. We say a set $S$ is closed under a
+    binary operator $f$ if for every $x$ and $y$ in $S$, $f(x, y)$ is also in
+    $S$. For example, even numbers are closed under addition because the sum of
+    any two even numbers is even. But odd numbers are not closed under addition
+    because the sum of two odd numbers is even.
+  }
 \end{frame}
@@
   $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
+
+  \note{%
+    We say that an object $O$ is invariant closed if invariant satisfying
+    states are closed under merge. That is, if an object is invariant closed,
+    then for every pair of states $s_1$ and $s_2$, if $s_1$ and $s_2$ satisfy
+    the invariant then so does $s_1 \join s_2$.
+  }
 \end{frame}
@@
   states satisfy the invariant.
-\end{frame}
 
-\begin{frame}
-  \Large
-  \[
-    \text{invariant closure} \implies \text{invariant confluence}
-  \]
-  \pause
-  Transactions perserve invariants. If merging does too, then all reachable
-  states satisfy the invariant.
+  \note{%
+    Invariant closure is a sufficient condition for invariant confluence. Why
+    is that? Well, transactions already preserve invariants. So if merging does
+    as well, then every reachable state always maintains the invariant.
+  }
 \end{frame}
@@
   \end{align*}
+
+  \note{%
+    The good thing about invariant closure is that it's much easier to check
+    automatically. for example, remember that example we had with pairs of
+    integers? We can pose whether or not that object is invariant closed as
+    this formula, and we can pass this formula directly to an SMT solver to
+    figure out whether it's invariant closed.
+  }
 \end{frame}
@@
   \]
+
+  \note{%
+    So to decide whether an object is invariant confluent, we can take the
+    object and ask if it's invariant closed. If it is, then it's also invariant
+    confluent and we're done. If it's not invariant closed, then what do we
+    know?
+  }
 \end{frame}
@@
   \]
+
+  \note{%
+    Well, that depends on whether invariant confluence implies invariant
+    closure.
+  }
 \end{frame}
@@
   \end{columns}
+
+  \note{%
+    To see if it does, let's revisit the example from before. Recall that in
+    this example, our object is invariant confluent.
+  }
 \end{frame}
@@
   \end{columns}
+
+  \note{%
+    But, note that our object is not invariant closed. The points $(-1, 2)$ and
+    $(2, -1)$ both satisfy the invariant, but their merge does not.
+  }
 \end{frame}
@@
   \]
+
+  \note{%
+    Thus, our object is invariant confluent but not invariant closed, so
+    invariant confluence does not imply invariant closure. This is unfortunate.
+    We can check for invariant closure but not invariant confluence, so we'd
+    like the two to be equivalent.
+
+    Are there any situations in which the two do happen to be equivalent?
+  }
 \end{frame}
@@
   \end{columns}
+
+  \note{%
+    Well, let's take a look at our example again. We noticed that our object is
+    not invariant closed because we can merge $(-1, 2)$ and $(2, -1)$ to
+    violate the invariant. But, notice that the point $(-1, 2)$ is not
+    reachable.
+  }
 \end{frame}
@@
   \end{itemize}
+
+  \note{%
+    This example is actually representative of the overall trend. An invariant
+    confluent object may not be invariant closed but only because we're merging
+    points that are unreachable. If the invariant is a subset of the reachable
+    states, then invariant closure and invariant confluence are identical.
+
+    Why? Well, the forward direction here is what we proved before. For the
+    backwards direction, we have that the invariant and the set of reachable
+    points are identical. The set of reachable states is closed under merge, so
+    so is the invariant, so the object is invariant closed.
+
+    I know I'm walking through a lot of theory pretty quickly, but if you
+    remember one slide, remember this one. If the invariant is a subset of the
+    reachable states, then invariant closure and invariant confluence are
+    identical.
+  }
 \end{frame}

### 2019-08-16 doc/vldb_slides/sections/evaluation.tex
Finished first cut at speaker notes.

@@
 checker.check()\end{verbatim}
+
+  \note{%
+    Finally, I'll give a teaser of our evaluation. We implemented our decision
+    procedure in Python. You can specify objects, transactions, and invariants
+    with code that looks like this.
+  }
 \end{frame}
@@
   specification.
+
+  \note{%
+    We specified a bunch of examples including things involving foreign keys,
+    an auction application, something involving escrow transactions, and TPC-C.
+    All the examples verify with a runtime of less then a second and every one
+    is implemented in fewer than 75 lines of specification.
+  }
 \end{frame}
@@
   \includegraphics[width=\textwidth]{assets/throughput_vs_fraction_16.pdf}
+
+  \note{%
+    We have some graphs showing that no coordination vastly outperforms
+    replicating with strong consistency, and how segmented invariant confluence
+    allows a smooth tradeoff between the two.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    I've jumped over a ton of details in the paper, and some things I didn't
+    even mention at all. If you're curious about anything, I encourage you to
+    come to our poster tomorrow night where I'm happy to chat.
+  }
 \end{frame}
@@
   \end{center}
+
+  \note{%
+    Thank you all for listening, and I'm happy to take questions.
+  }
 \end{frame}

### 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex
Added four wise men poem
Finished first cut at speaker notes.

@@
 
-    Imagine you're an application developer and you're trying to decide what
-    kind of consistency you'd like to use to replicate some of your data.
+    I'd like to begin my talk, if you don't mind, with a short fairy tale. The
+    fairy tale involves an application developer who has some data they need to
+    replicate. Unfortunately, the application developer cannot decide whether
+    to replicate with weak consistency or strong consistency. The application
+    developer stays up all night trying to decide but eventually falls asleep
+    at their desk. In their sleep, they are visited by four wise men.
+
+    % Imagine you're an application developer and you're trying to decide what
+    % kind of consistency you'd like to use to replicate some of your data.
+
   }
@@
   \note{%
-    Someone like Werner Vogels would probably tell you to use weak consistency.
-    You can implement weak consistency super duper fast, and for many
-    applications, weak consistency is good enough.
+    % Someone like Werner Vogels would probably tell you to use weak consistency.
+    % You can implement weak consistency super duper fast, and for many
+    % applications, weak consistency is good enough.
+
+    ``Weak. Weak. It's weak that you're seeking'', squeaked Werner Vogels.
+    His voice trembled, creaking, as he kept speaking.
+    ``Please learn that if you yearn your performance to peak,
+    then turn no further than weak.'' critiqued Werner.
+
+    Werner Vogels has a good point. You can implement weak consistency super
+    duper fast, and for many applications, weak consistency is good enough. But
+    anyway, let's continue.
   }
@@
   \note{%
-    But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
-    settle for weak consistency because a weakly consistent system is hard to
-    reason about. It robs us of a fundamental way we use to reason about
-    programs: application invariants. Even if every transaction in a weakly
-    consistent system preserves application invariants, the whole system may
-    not.
+    % But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
+    % settle for weak consistency because a weakly consistent system is hard to
+    % reason about. It robs us of a fundamental way we use to reason about
+    % programs: application invariants. Even if every transaction in a weakly
+    % consistent system preserves application invariants, the whole system may
+    % not.
+
+    ``No, no, no more'' groaned Tony Hoare.
+    He moaned, he lamented, his soul was tormented
+    ``Weak is not easy. Weak is just treason.
+    Please, these anomalies, make it so hard to reason.''
+
+    You see, Hoare encourages us not to settle for weak consistency because a
+    weakly consistent system is very hard to reason about.
   }
@@
   \note{%
-    Leslie Lamport chimes in agreeingly. He encourages us to choose strong
-    consistency. Strongly consistent systems are much easier to reason about.
-    After all a strongly consistent system to end users looks exactly like a
-    system that's not even distributed.
+    % Leslie Lamport chimes in agreeingly. He encourages us to choose strong
+    % consistency. Strongly consistent systems are much easier to reason about.
+    % After all a strongly consistent system to end users looks exactly like a
+    % system that's not even distributed.
+
+    ``You see, I agree'', says he, the ever mesmerizing Leslie.
+    ``Weak is just wrong, not right for long.
+    To defend your end users, you must go with strong.''
+
+    Leslie Lamport encourages us to choose strong consistency. Strongly
+    consistent systems are much easier to reason about. After all a strongly
+    consistent system to end users looks exactly like a system that's not even
+    distributed.
   }
@@
   \note{%
-    ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
-    partition tolerance, oh my. Surely, strong consistency comes at the price
-    of availability and performance.
+    % ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
+    % partition tolerance, oh my. Surely, strong consistency comes at the price
+    % of availability and performance.
+    %
+    % So we're in a bit of a pickle. Weak consistency is too confusing and strong
+    % consistency is too heavy handed.
 
-    So we're in a bit of a pickle. Weak consistency is too confusing and strong
-    consistency is too heavy handed.
+    ``Alas, not so fast'', Brewer the dapper chap clapped back.
+    ``Consistency weak but not weary, strong but not nearly
+    as fast as you think when you know my CAP theory.''
+
+    Eric Brewer is right. Surely, strong consistency comes at the price of
+    availability and performance. So our application developer is in a bit of a
+    pickle. Weak consistency is too confusing and strong consistency is too
+    heavy handed. Fortunately, a hero emerges to save the day.
   }
@@
   \note{%
-    Thankfully, Peter Bailis comes to save the day and reminds us that perhaps
-    we can have the best of both worlds. In 2014, Peter and some other smart
-    folks including Alan defined invariant confluence as a condition under
-    which we can replicate a piece of data without the need for strongly
-    consistent coordination while maintaining the data's application
-    invariants.
+    But then out teetered Peter, with swagger demeanor.
+    ``Why not have both?'' quoth Peter.
+    The end.
+
+    Peter Bailis comes to save the day and reminds us that perhaps we can have
+    the best of both worlds. In 2014, Peter and some other smart folks
+    including Alan defined invariant confluence as a condition under which we
+    can replicate a piece of data without the need for strongly consistent
+    coordination while maintaining the data's application invariants.
 

### 2019-08-16 doc/vldb_poster/sections/iconfluence.tex
Finished poster.

\begin{itemize}
  \item
    A \defword{distributed object} $O = (S, \join)$ consists of a set $S$ of
    states and a binary merge operator $\join: S \times S \to S$.
  \item
    A \defword{transaction} $t: S \to S$ is a function.
  \item
    An \defword{invariant} $I$ is a subset of $S$.
\end{itemize}

\tikzstyle{s0color}=[fill=flatred]
\tikzstyle{s1color}=[fill=flatgreen]
\tikzstyle{s2color}=[fill=flatdenim]
\tikzstyle{s3color}=[fill=flatorange]
\tikzstyle{s4color}=[fill=flatyellow]
\tikzstyle{s5color}=[fill=flatcyan]
\tikzstyle{s6color}=[fill=flatpurple]
\tikzstyle{s7color}=[fill=flatblue]
\tikzstyle{txnline}=[black, line width=6pt, -latex]

\newcommand{\internaltext}[1]{$\boldsymbol #1$}

\tikzstyle{phantomstate}=[%
  shape=circle, inner sep=12pt, draw=white, line width=4pt, fill=white]
\tikzstyle{state}=[%
  shape=circle, inner sep=12pt, draw=black, line width=4pt, text opacity=1,
  fill opacity=0.6]

\begin{center}
  \large
  \begin{tikzpicture}[scale=5]
    \node (p3) at (0.5, 3) {$p_1$};
    \node (p2) at (0.5, 2) {$p_2$};
    \node (p1) at (0.5, 1) {$p_3$};

    \tikzstyle{pline}=[gray, opacity=0.75, line width=6pt, -latex]
    \draw[pline] (p3) to (4.75, 3);
    \draw[pline] (p2) to (4.75, 2);
    \draw[pline] (p1) to (4.75, 1);

    % Top line.
    \node[phantomstate] (s03) at (1, 3) {$\phantom{s_0}$};
    \node[phantomstate] (s13) at (2, 3) {$\phantom{s_1}$};
    \node[phantomstate] (s23) at (3, 3) {$\phantom{s_3}$};
    \node[phantomstate] (s33) at (3.75, 3) {$\phantom{s_6}$};
    \node[state, s0color] (s03) at (1, 3) {\internaltext{s_0}};
    \node[state, s1color] (s13) at (2, 3) {\internaltext{s_1}};
    \node[state, s3color] (s23) at (3, 3) {\internaltext{s_3}};
    \node[state, s6color] (s33) at (3.75, 3) {\internaltext{s_6}};

    % Middle line.
    \node[phantomstate] (s02) at (1, 2) {$\phantom{s_0}$};
    \node[phantomstate] (s12) at (2, 2) {$\phantom{s_2}$};
    \node[phantomstate] (s22) at (3, 2) {$\phantom{s_4}$};
    \node[state, s0color] (s02) at (1, 2) {\internaltext{s_0}};
    \node[state, s2color] (s12) at (2, 2) {\internaltext{s_2}};
    \node[state, s4color] (s22) at (3, 2) {\internaltext{s_4}};

    % Bottom line.
    \node[phantomstate] (s01) at (1, 1) {$\phantom{s_0}$};
    \node[phantomstate] (s11) at (2, 1) {$\phantom{s_2}$};
    \node[phantomstate] (s21) at (3, 1) {$\phantom{s_5}$};
    \node[phantomstate] (s31) at (4.25, 1) {$\phantom{s_7}$};
    \node[state, s0color] (s01) at (1, 1) {\internaltext{s_0}};
    \node[state, s2color] (s11) at (2, 1) {\internaltext{s_2}};
    \node[state, s5color] (s21) at (3, 1) {\internaltext{s_5}};
    \node[state, s7color] (s31) at (4.25, 1) {\internaltext{s_7}};

    \tikzstyle{txntext}=[sloped, above]
    \draw[txnline] (s03) to node[txntext]{$t$} (s13);
    \draw[txnline] (s13) to node[txntext]{$v$} (s23);
    \draw[txnline] (s02) to node[txntext]{$u$} (s12);
    \draw[txnline] (s01) to node[txntext]{$u$} (s11);
    \draw[txnline] (s11) to node[txntext]{$w$} (s21);

    \draw[txnline] (s13) to node[txntext]{} (s22);
    \draw[txnline] (s12) to node[txntext]{} (s22);
    \draw[txnline] (s22) to node[txntext]{} (s33);
    \draw[txnline] (s23) to node[txntext]{} (s33);
    \draw[txnline] (s21) to node[txntext]{} (s31);
    \draw[txnline] (s33) to node[txntext]{} (s31);
  \end{tikzpicture}
\end{center}

\[
  e ::= s \mid t(e) \mid e_1 \join e_2
\]

\begin{center}
  \begin{tikzpicture}[scale=4]
    \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (j1) at (0, 0) {\internaltext{\join}};
    \draw (j1)++(-30:1) node[state, s5color, label={[label distance=-0.1cm] 90:$s_5$}] (w)            {\internaltext{w}};
    \draw (j1)++(210:1) node[state, s6color, label={[label distance=-0.1cm] 90:$s_6$}] (j2)           {\internaltext{\join}};
    \draw (w)++(-90:1)  node[state, s2color, label={[label distance=-0.2cm] 60:$s_2$}] (u1)           {\internaltext{u}};
    \draw (u1)++(-90:1) node[state, s0color]                                           (s4)           {\internaltext{s_0}};
    \draw (j2)++(225:1) node[state, s3color, label={[label distance=-0.1cm] 90:$s_3$}] (v)            {\internaltext{v}};
    \draw (v)++(-90:1)  node[state, s1color, label={[label distance=-0.2cm]120:$s_1$}] (t2)           {\internaltext{t}};
    \draw (t2)++(-90:1) node[state, s0color]                                           (s1)           {\internaltext{s_0}};
    \draw (j2)++(-45:1) node[state, s4color, label={[label distance=-0.1cm] 90:$s_4$}] (j3)           {\internaltext{\join}};
    \draw (j3)++(240:1) node[state, s1color, label={[label distance=-0.2cm]120:$s_1$}] (t3)           {\internaltext{t}};
    \draw (t3)++(-90:1) node[state, s0color]                                           (s2)           {\internaltext{s_0}};
    \draw (j3)++(-60:1) node[state, s2color, label={[label distance=-0.1cm] 90:$s_2$}] (u3)           {\internaltext{u}};
    \draw (u3)++(-90:1) node[state, s0color]                                           (s3)           {\internaltext{s_0}};

    \tikzstyle{astedge}=[line width=6pt]
    \draw[astedge] (j1) to (w) to (u1) to (s4);
    \draw[astedge] (j1) to (j2) to (v) to (t2) to (s1);
    \draw[astedge] (j1) to (j2) to (j3) to (t3) to (s2);
    \draw[astedge] (j1) to (j2) to (j3) to (u3) to (s3);
  \end{tikzpicture}
\end{center}

\begin{mathpar}
  \inferrule{ }{\sTIreachablepredicate{s_0}}

  \inferrule{\sTIreachablepredicate{e} \\ I(t(e))}
            {\sTIreachablepredicate{t(e)}}

  \inferrule{\sTIreachablepredicate{e_1} \\ \sTIreachablepredicate{e_2}}
            {\sTIreachablepredicate{e_1 \join e_2}}
\end{mathpar}

$O$ is \defword{\invariantconfluent} with respect to $s_0$, $T$, and $I$,
abbreviated \defword{\sTIconfluent{}}, if all reachable states satisfy the
invariant:
\[
  \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
\]

### 2019-08-18 doc/vldb_slides/sections/bank_example.tex
Tweaks, plus gutted poem.

@@
   \note{%
-    So, what is invariant confluence? Well, let's begin with an example.
+    We'll begin with an example.
 
@@
     want to maintain over my database. And that invariant is that my bank
-    account balance should be non-negative at all times. I'm not allowed to go
-    in the hole.
+    account balance should never be negative. My balance must be non-negative at all times.
   }
@@
   \note{%
-    Now, instead of implementing our bank with strong consistency, we could
-    instead implement it with weak consistency and avoid coordination
-    altogether.
+    Alternatively, instead of implementing our bank with strong consistency, we
+    could implement it with weak consistency and avoid coordination altogether.
 
@@
   \note{%
-    Now, these are just two examples, but if you think about it long enough,
-    they are representative of a trend. We can implement deposit transactions
-    without any coordination and we're guaranteed to never violate the
-    invariant. Withdrawal transactions, on the other hand, do require
+    We can learn a lot from these examples.
+
+    First, if we implement our bank with strong consistency, we never violate
+    our invariant.
+
+    Second, we can implement deposit transactions without any coordination and
+    we're still guaranteed to never violate the invariant.
+
+    Third, withdrawal transactions, unlike deposit transactions, do require
     coordination to avoid violating the invariant.
 
-    At a very high level, this is exactly the notion that invariant confluence
-    captures. We'll see that deposit transactions are invariant confluent and
-    thus do not require coordination while withdrawal transactions are not
-    invariant confluent and thus do require coordination.
+    At a very high level, invariant confluence tries to capture exactly which
+    transactions require coordination to maintain invariants and which do not.
+    In this example, deposit transactions are invariant confluent and thus do
+    not require coordination while withdrawal transactions are not invariant
+    confluent and thus do require coordination.
   }

### 2019-08-18 doc/vldb_slides/sections/four_wise_men.tex
Tweaks, plus gutted poem.

--- 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex
@@
     talking about ``Interactive Checks for Coordination Avoidance''. This is
-    joint work with my advisor at UC Berkeley, Joe Hellerstein.
+    joint work with my advisor at UC Berkeley, Joe Hellerstein. \\[12pt]
+
+    % I'd like to begin my talk with a short fairy tale. The fairy tale involves
+    % an application developer who has some data they need to replicate.
+    % Unfortunately, the application developer cannot decide whether to replicate
+    % with weak consistency or strong consistency. The application developer
+    % stays up all night trying to decide but eventually falls asleep at their
+    % desk. In their sleep, they are visited by four wise men. And so, the fairy
+    % tale begins:
 
@@
       \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{You should use weak consistency. It's wicked fast!}
+        \Large\textit{
+        Go with weak consistency. It's super duper fast and good enough for
+        many applications!
+        % Weak. Weak. It's weak that you're seeking.
+        % \pause
+        % Please learn that if you yearn your performance to peak, then turn no
+        % further than weak.
+      }
       };
@@
     You can implement weak consistency super duper fast, and for many
-    applications, weak consistency is good enough.
+    applications, weak consistency is good enough. \\[12pt]
+
+    % ``Weak. Weak. It's weak that you're seeking'', squeaked Werner Vogels.
+    % His voice trembled, creaking, as he kept speaking.
+    % ``Please learn that if you yearn your performance to peak,
+    % then turn no further than weak.'' critiqued Werner.
+    %
+    % Werner Vogels has a good point. You can implement weak consistency super
+    % duper fast, and for many applications, weak consistency is good enough.
+    % But, let's continue.
   }
@@
       \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
-        \Large\textit{What about the application invariants?!}
+        \Large\textit{
+          Weakly consistent systems are so hard to reason about!
+          % No. No. No more.
+          % \pause
+          % Weak is not easy. Weak is just treason. Please, these anomalies make
+          % it so hard to reason.
+        }
       };
-      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
+      \path[fill=gray!30] (hoare.north) -- ++(0, 1) -- ++(0.5, 0) -- (hoare.north);
     \end{tikzpicture}
@@
   \note{%
-    But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
-    settle for weak consistency because a weakly consistent system is hard to
-    reason about. It robs us of a fundamental way we use to reason about
-    programs: application invariants. Even if every transaction in a weakly
-    consistent system preserves application invariants, the whole system may
-    not.
+    Tony Hoare would encourage us not to settle for weak consistency because a
+    weakly consistent system is hard to reason about.
+
+    % ``No, no, no more'' groaned Tony Hoare.
+    % He moaned, he lamented, his soul was tormented
+    % ``Weak is not easy. Weak is just treason.
+    % Please, these anomalies, make it so hard to reason.''
+    %
+    % You see, Tony Hoare encourages us not to settle for weak consistency
+    % because a weakly consistent system is very hard to reason about.
   }
@@
       \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
-        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
+        \Large\textit{
+          Go with a strongly consistent system! I think there are some good
+          algorithms for that kind of stuff.
+          % You see, I agree.
+          % \pause
+          % Weak is just wrong, not right for long. To defend your end users, you
+          % must go with strong.
+        }
       };
-      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
+      \path[fill=gray!30] (lamport.north) -- ++(0, 0.5) -- ++(0.25, 0) -- (lamport.north);
     \end{tikzpicture}
@@
   \note{%
-    Leslie Lamport chimes in agreeingly. He encourages us to choose strong
+    Leslie Lamport would agree. He would encourage us to choose strong
     consistency. Strongly consistent systems are much easier to reason about.
@@
     system that's not even distributed.
+
+    % ``You see, I agree'', says he, the ever mesmerizing Leslie.
+    % ``Weak is just wrong, not right for long.
+    % To defend your end users, you must go with strong.''
+    %
+    % Leslie Lamport encourages us to choose strong consistency becausesStrongly
+    % consistent systems are much easier to reason about. After all a strongly
+    % consistent system to end users looks exactly like a system that's not even
+    % distributed.
   }
@@
       \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
-        \Large\textit{Not so fast!}
+        \Large\textit{
+          Not so fast!
+          % Alas, not so fast!
+          % \pause
+          % Consistency weak but not weary, strong but not nearly as fast as you
+          % think when you hear my CAP theory.
+        }
       };
-      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
+      \path[fill=gray!30] (brewer.north) -- ++(0, 1) -- ++(-1, 0) -- (brewer.north);
     \end{tikzpicture}
@@
   \note{%
-    ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
-    partition tolerance, oh my. Surely, strong consistency comes at the price
-    of availability and performance.
+    Eric Brewer would counter that strong consistency isn't perfect. Strong
+    consistency comes at the price of availability and performance. So we're in
+    a bit of a pickle. Weak consistency is too confusing and strong consistency
+    is too heavy handed.
 
-    So we're in a bit of a pickle. Weak consistency is too confusing and strong
-    consistency is too heavy handed.
+    %
+    % ``Alas, not so fast'', Brewer the dapper chap clapped back.
+    % ``Consistency weak but not weary, strong but not nearly
+    % as fast as you think when you hear my CAP theory.''
+    %
+    % Eric Brewer is right. Surely, strong consistency comes at the price of
+    % availability and performance. So our application developer is in a bit of a
+    % pickle. Weak consistency is too confusing and strong consistency is too
+    % heavy handed. Fortunately, a hero emerges to save the day.
   }
@@
       \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{?`Porque no los dos?}
+        \Large\textit{Why not have both?}
       };
@@
   \note{%
-    Thankfully, Peter Bailis comes to save the day and reminds us that perhaps
-    we can have the best of both worlds. In 2014, Peter and some other smart
-    folks defined invariant confluence as a condition under which we can
-    replicate a piece of data without the need for strongly consistent
-    coordination while maintaining the data's application invariants.
+    % Out teetered Peter, with swagger demeanor.
+    % ``Why not have both?'' quoth the Peter.
+    % The end.
 
-    Invariant confluence will be the topic of this talk. In particular, I'll
-    define invariant confluence and then discuss how we can check invariant
-    confluence automatically.
+    Fortunately, Peter Bailis can save the and reminding us of when in 2014, he
+    and some other smart folks including Alan defined invariant confluence as a
+    way to get the benefits of both weak and strong consistency. An invariant
+    confluent object can be replicated without coordination while maintaining
+    its application invariants.
+
+    Invariant confluence will be the topic of this talk, so let's build some
+    intuition on what invariant confluence is.
   }

### 2019-08-18 doc/vldb_slides/sections/iclosure.tex
Finished first cut at speaker notes.
Tweaks, plus gutted poem.

@@
     about reachable states but reasoning automatically about reachable states
-    is hard. I'm not saying it's impossible, but certainly I don't know how to
-    do it.
+    is hard.
 
@@
   \note{%
-    First, a quick refresher on terminology. We say a set $S$ is closed under a
-    binary operator $f$ if for every $x$ and $y$ in $S$, $f(x, y)$ is also in
-    $S$. For example, even numbers are closed under addition because the sum of
-    any two even numbers is even. But odd numbers are not closed under addition
-    because the sum of two odd numbers is even.
+    First, a quick refresher on what it meas for a set to be closed. We say a
+    set $S$ is closed under a binary operator $f$ if for every $x$ and $y$ in
+    $S$, $f(x, y)$ is also in $S$. For example, even numbers are closed under
+    addition because the sum of any two even numbers is even. But odd numbers
+    are not closed under addition because the sum of two odd numbers may not
+    be odd.
   }
@@
     Invariant closure is a sufficient condition for invariant confluence. Why
-    is that? Well, transactions already preserve invariants. So if merging does
-    as well, then every reachable state always maintains the invariant.
+    is that? Well for an object to be invariant confluent, every reachable
+    state must satisfy the invariant. A state is reachable if it can be reached
+    by executing transactions merging states. Transactions are already
+    guarnateed to preserve invariants, so if merging is also guaranteed to
+    preserve the invariant, then every reachable state always satisfies the
+    invariant.
   }
@@
     The good thing about invariant closure is that it's much easier to check
-    automatically. for example, remember that example we had with pairs of
+    automatically. Remember that example we had with pairs of
     integers? We can pose whether or not that object is invariant closed as
     this formula, and we can pass this formula directly to an SMT solver to
-    figure out whether it's invariant closed.
+    figure out whether it's invariant closed. This is not something we can do
+    with invariant confluence because it's hard to express reachability as a
+    simple formula like this one.
   }
@@
   \note{%
-    So to decide whether an object is invariant confluent, we can take the
+    To decide whether an object is invariant confluent, we can take the
     object and ask if it's invariant closed. If it is, then it's also invariant
@@
     To see if it does, let's revisit the example from before. Recall that in
-    this example, our object is invariant confluent.
+    this example, our object is invariant confluent. The set of reachable
+    states is a subset of the invariant.
   }
@@
     But, note that our object is not invariant closed. The points $(-1, 2)$ and
-    $(2, -1)$ both satisfy the invariant, but their merge does not.
+    $(2, -1)$ both satisfy the invariant, but if we merge them we get the point
+    $(2, 2)$, and the point $(2, 2)$ does not satisfy the invariant.
   }
@@
   \note{%
-    This example is actually representative of the overall trend. An invariant
-    confluent object may not be invariant closed but only because we're merging
-    points that are unreachable. If the invariant is a subset of the reachable
-    states, then invariant closure and invariant confluence are identical.
+    Turns out, this is not a coincidence. An invariant confluent object may not
+    be invariant closed but only because we're merging points that are
+    unreachable. If the invariant is a subset of the reachable states, then
+    invariant closure and invariant confluence are equivalent. \\[12pt]
 
     Why? Well, the forward direction here is what we proved before. For the
-    backwards direction, we have that the invariant and the set of reachable
-    points are identical. The set of reachable states is closed under merge, so
-    so is the invariant, so the object is invariant closed.
+    backwards direction, we have that the invariant is a subset of the set of
+    reachable points, and because our object is invariant confluent, the set of
+    reachable points is a subset of the invariant. So the invariant and the set
+    of reachable points are the same. The set of reachable states is closed
+    under merge, so the invariant is also, so the object is invariant closed.
+    \\[12pt]
 
-    I know I'm walking through a lot of theory pretty quickly, but if you
+    I know I've walked through a lot of theory pretty quickly, but if you
     remember one slide, remember this one. If the invariant is a subset of the
     reachable states, then invariant closure and invariant confluence are
-    identical.
+    equivalent. \\[12pt]
   }

### 2019-08-18 doc/vldb_slides/sections/evaluation.tex
Slide tweaks.

@@
     Finally, I'll give a teaser of our evaluation. We implemented our decision
-    procedure in Python. You can specify objects, transactions, and invariants
-    with code that looks like this.
+    procedure in Python and leveraged the Z3 decision procedure. You can
+    specify objects, transactions, and invariants with code that looks like
+    this. \\[12pt]
+
+    This code specifies the running example that we've been looking at
+    throughout the talk. $x$ is an integer merged by max that is initially 0.
+    $y$ is an integer merged by max that is initially 0. We have a transaction
+    incx that increments x and a function decy that decrements y. And, our
+    invariant is that the product of $x$ and $y$ is less than or equal to zero.
   }
@@
   \note{%
-    We specified a bunch of examples including things involving foreign keys,
-    an auction application, something involving escrow transactions, and TPC-C.
-    All the examples verify with a runtime of less then a second and every one
-    is implemented in fewer than 75 lines of specification.
+    In our paper, we specified a bunch of examples including things involving
+    foreign keys, an auction application, something involving escrow
+    transactions, and TPC-C.  All the examples verify with a runtime of less
+    then a second and every one is implemented in fewer than 75 lines of
+    specification.
   }

### 2019-08-18 doc/vldb_slides/sections/iclosure.tex
More slide tweaks.

@@
     the main goal of our paper. And that is to develop an invariant-confluence
-    decision procedure. Determining whether or not a distributed object is
-    invariant confluent is not easy to do by hand, so we'd like to develop a
-    decision procedure that can automatically check whether something is
-    invariant confluent for us.
+    decision procedure. As you saw in the previos example, determining whether
+    or not a distributed object is invariant confluent is not easy to do by
+    hand, so we'd like to develop a decision procedure that can automatically
+    check whether something is invariant confluent for us.
   }
@@
     be invariant closed but only because we're merging points that are
-    unreachable. If the invariant is a subset of the reachable states, then
-    invariant closure and invariant confluence are equivalent. \\[12pt]
+    unreachable. If the invariant is a subset of the reachable states---that
+    is, if every invariant satisfying point is reachable---then invariant
+    closure and invariant confluence are equivalent. \\[12pt]
 

### 2019-08-18 doc/vldb_slides/sections/bank_example.tex
Slimmed down talk.

@@
     want to maintain over my database. And that invariant is that my bank
-    account balance should never be negative. My balance must be non-negative at all times.
+    account balance should never be negative. My balance must be non-negative
+    at all times.
   }
@@
     two transactions. Here, the replicas agree to deposit 10 dollars and then
-    deposit 20 dollars. Both transactions succeed, and the replicas send back OK
-    messages to the clients letting them know everything went smoothly.
+    deposit 20 dollars. Both transactions succeed.
   }
@@
     Now, assume two clients issue concurrent withdrawal transactions of 30
-    dollars and 40 dollars. Again, the two replicas coordinate and agree to
-    execute the withdrawal of 30 dollars and then the withdrawal of 40 dollars.
+    dollars and 40 dollars. Again, the two replicas coordinate and agree an a
+    total order in which to execute transactions.
 
-    The withdrawal of 30 dollars executes fine, brining my bank account balance
-    to 20 dollars, but the withdrawal of 40 dollars fails because I only have
-    20 dollars left. So replica A sends back an OK message to the client, but
-    replica B sends back a NO message letting the client know the transaction
-    failed. And while the transaction failed, you'll note that we never
-    violated our application invariant.
+    They first withdraw 30 dollars, bringing my bank account balance to 20
+    dollars. They then try to withdraw of 40 dollars but abort the transaction
+    because I only have 40 dollars left. And while the transaction failed,
+    you'll note that we never violated our application invariant.
   }
@@
     another. And every once in a while, the two replicas communicate and fill
-    each other in on the transactions they missed.
-
-    Here, for example, replica A lets replica B know to deposit 10 dollars and
-    replica B lets replica A know to deposit 20 dollars. Notice that even
-    though we didn't perform any coordination, the two replicas never violate
-    my critical invariant of a non-negative balance.
+    each other in on the transactions they missed. Notice that even though we
+    didn't perform any coordination, the two replicas never violate my critical
+    invariant of a non-negative balance.
   }

### 2019-08-18 doc/vldb_slides/sections/evaluation.tex
More slide tweaks.
Slimmed down talk.

@@
 
-    This code specifies the running example that we've been looking at
-    throughout the talk. $x$ is an integer merged by max that is initially 0.
-    $y$ is an integer merged by max that is initially 0. We have a transaction
-    incx that increments x and a function decy that decrements y. And, our
-    invariant is that the product of $x$ and $y$ is less than or equal to zero.
+    % This code specifies the running example that we've been looking at
+    % throughout the talk. $x$ is an integer merged by max that is initially 0.
+    % $y$ is an integer merged by max that is initially 0. We have a transaction
+    % incx that increments x and a function decy that decrements y. And, our
+    % invariant is that the product of $x$ and $y$ is less than or equal to zero.
   }
@@
     allows a smooth tradeoff between the two.
+
+    % Fully invariant confluent objects don't require any coordination, partly
+    % invariant confluent objects require some coordination, and not at all
+    % invariant confluent objects require a lot of coordination.
   }

### 2019-08-18 doc/vldb_slides/sections/iclosure.tex
Slimmed down talk.

@@
 
-\begin{frame}
-  \Large
-  We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
-  y) \in S$. \pause For example,
-  \begin{itemize}
-    \item Even numbers are closed under addition.
-    \item Odd numbers are \emph{not} closed under addition.
-  \end{itemize}
-
-  \note{%
-    First, a quick refresher on what it meas for a set to be closed. We say a
-    set $S$ is closed under a binary operator $f$ if for every $x$ and $y$ in
-    $S$, $f(x, y)$ is also in $S$. For example, even numbers are closed under
-    addition because the sum of any two even numbers is even. But odd numbers
-    are not closed under addition because the sum of two odd numbers may not
-    be odd.
-  }
-\end{frame}
+% \begin{frame}
+%   \Large
+%   We say a set $S$ is \defword{closed under $f$} if for every $x, y \in S, f(x,
+%   y) \in S$. \pause For example,
+%   \begin{itemize}
+%     \item Even numbers are closed under addition.
+%     \item Odd numbers are \emph{not} closed under addition.
+%   \end{itemize}
+%
+%   \note{%
+%     First, a quick refresher on what it meas for a set to be closed. We say a
+%     set $S$ is closed under a binary operator $f$ if for every $x$ and $y$ in
+%     $S$, $f(x, y)$ is also in $S$. For example, even numbers are closed under
+%     addition because the sum of any two even numbers is even. But odd numbers
+%     are not closed under addition because the sum of two odd numbers may not
+%     be odd.
+%   }
+% \end{frame}
 
@@
   \note{%
-    Invariant closure is a sufficient condition for invariant confluence. Why
-    is that? Well for an object to be invariant confluent, every reachable
-    state must satisfy the invariant. A state is reachable if it can be reached
-    by executing transactions merging states. Transactions are already
-    guarnateed to preserve invariants, so if merging is also guaranteed to
-    preserve the invariant, then every reachable state always satisfies the
-    invariant.
+    Invariant closure is a sufficient condition for invariant confluence. I'll
+    refer you to the paper for the proof, but it's a very simple proof.
+
+    % Why is that? Well for an object to be invariant confluent, every reachable
+    % state must satisfy the invariant. A state is reachable if it can be reached
+    % by executing transactions merging states. Transactions are already
+    % guarnateed to preserve invariants, so if merging is also guaranteed to
+    % preserve the invariant, then every reachable state always satisfies the
+    % invariant.
   }
@@
     this formula, and we can pass this formula directly to an SMT solver to
-    figure out whether it's invariant closed. This is not something we can do
-    with invariant confluence because it's hard to express reachability as a
-    simple formula like this one.
+    figure out whether it's invariant closed.
+
+    % This is not something we can do with invariant confluence because it's hard
+    % to express reachability as a simple formula like this one.
   }
@@
   \note{%
-    To decide whether an object is invariant confluent, we can take the
+    To decide whether an object is invariant confluent, then, we can take the
     object and ask if it's invariant closed. If it is, then it's also invariant
@@
 
-    Why? Well, the forward direction here is what we proved before. For the
-    backwards direction, we have that the invariant is a subset of the set of
-    reachable points, and because our object is invariant confluent, the set of
-    reachable points is a subset of the invariant. So the invariant and the set
-    of reachable points are the same. The set of reachable states is closed
-    under merge, so the invariant is also, so the object is invariant closed.
-    \\[12pt]
-
-    I know I've walked through a lot of theory pretty quickly, but if you
-    remember one slide, remember this one. If the invariant is a subset of the
-    reachable states, then invariant closure and invariant confluence are
-    equivalent. \\[12pt]
+    Again, I'll refer you to the paper for a proof.
+    % Why? Well, the forward direction here is what we proved before. For the
+    % backwards direction, we have that the invariant is a subset of the set of
+    % reachable points, and because our object is invariant confluent, the set of
+    % reachable points is a subset of the invariant. So the invariant and the set
+    % of reachable points are the same. The set of reachable states is closed
+    % under merge, so the invariant is also, so the object is invariant closed.
+    % \\[12pt]
+    %
+    % I know I've walked through a lot of theory pretty quickly, but if you
+    % remember one slide, remember this one. If the invariant is a subset of the
+    % reachable states, then invariant closure and invariant confluence are
+    % equivalent. \\[12pt]
   }

### 2019-08-20 doc/vldb_slides/sections/bank_example.tex
Progress towards slide revisions.

@@
 \newcommand{\bankdistance}{4}
-\tikzstyle{bank}=[
-  draw,
-  minimum width=1.5cm,
-  minimum height=1cm,
-  font=\huge
-]
-\tikzstyle{op}=[
-  -latex,
-  ultra thick
-]
-
-\newcommand{\drawbanks}[6]{{
-  \newcommand{\banka}{#1}
-  \newcommand{\bankb}{#2}
-  \newcommand{\bankaa}{#3}
-  \newcommand{\bankbb}{#4}
-  \newcommand{\bankaaa}{#5}
-  \newcommand{\bankbbb}{#6}
-
-  \node[bank, label=above:{Replica $A$}] (a1) at (0,              0) {\banka};
-  \node[bank                           ] (a2) at (0,             -3) {\bankaa};
-  \node[bank                           ] (a3) at (0,             -6) {\bankaaa};
-  \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\bankb};
-  \node[bank                           ] (b2) at (\bankdistance, -3) {\bankbb};
-  \node[bank                           ] (b3) at (\bankdistance, -6) {\bankbbb};
-
-  \draw[op] (a1) -- (a2);
-  \draw[op] (a2) -- (a3);
-  \draw[op] (b1) -- (b2);
-  \draw[op] (b2) -- (b3);
-}}
+\tikzstyle{bank}=[draw, minimum width=1.5cm, minimum height=1cm, font=\huge]
+\tikzstyle{op}=[-latex, ultra thick]
 
@@
   \newcommand{\bankb}{#2}
-
-  \draw[red, op] ($(\banka) + (0, -0.7)$) -- ($(\bankb) + (0, -1.0)$);
-  \draw[red, op] ($(\bankb) + (0, -1.0)$) -- ($(\banka) + (0, -1.3)$);
-  \draw[red, op] ($(\banka) + (0, -1.3)$) -- ($(\bankb) + (0, -1.6)$);
-  \draw[red, op] ($(\bankb) + (0, -1.6)$) -- ($(\banka) + (0, -1.9)$);
+  \draw[gray, op] ($(\banka) + (0, -0.7)$) -- ($(\bankb) + (0, -1.0)$);
+  \draw[gray, op] ($(\bankb) + (0, -1.0)$) -- ($(\banka) + (0, -1.3)$);
+  \draw[gray, op] ($(\banka) + (0, -1.3)$) -- ($(\bankb) + (0, -1.6)$);
+  \draw[gray, op] ($(\bankb) + (0, -1.6)$) -- ($(\banka) + (0, -1.9)$);
 }}
@@
   \note{%
-    We'll begin with an example.
+    We'll begin with an example. \\[12pt]
 
-    Let's say we want to replicate the database shown here. This database is
-    the world's simplest database. It contains only a single number, which
-    represents my bank account balance. We decide to replicate this database
-    across two machines, replica A and replica B.
+    Let's say we have an object that we want to replicate. The object can be
+    something complex like a database or a key-value store, but to keep things
+    simple, let's say we want to replicate a single integer that represents my
+    bank account balance. We decide to replicate the integer across two
+    machines, replica A and replica B. \\[12pt]
 
-    Users can issue transactions to the database, and every transacation is
-    either a deposit or a withdrawal. And, I have one critical invariant that I
-    want to maintain over my database. And that invariant is that my bank
-    account balance should never be negative. My balance must be non-negative
-    at all times.
+    Users can issue transactions to our replicate object, and every
+    transacation is either a deposit or a withdrawal. And, I have one critical
+    invariant that I want to maintain at all times. And that invariant is that
+    my bank account balance should never be negative. \\[12pt]
   }
@@
     \begin{tikzpicture}
-      \drawbanks{\$50}{\$50}
-                {\$60}{\$60}
-                {\$80}{\$80}
-
+      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
+      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
       \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
       \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
-      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
-      \draw[op] (b3) -- ($(b3) + (+2, 0)$) node[midway, above] {OK};
+      \pause
 
       \zigzag{a1}{b1}
+      \node[bank] (a2) at (0, -3) {\$60};
+      \node[bank] (b2) at (\bankdistance, -3) {\$60};
+      \draw[op] (a1) to node[left] {$+\$10$} (a2);
+      \draw[op] (b1) to node[right] {$+\$10$} (b2);
+      \pause
+
       \zigzag{a2}{b2}
+      \node[bank] (a3) at (0, -6) {\$80};
+      \node[bank] (b3) at (\bankdistance, -6) {\$80};
+      \draw[op] (a2) to node[left] {$+\$20$} (a3);
+      \draw[op] (b2) to node[right] {$+\$20$} (b3);
     \end{tikzpicture}
@@
   \note{%
-    One way to replicate our database is by using coordination to implement
+    One way to replicate our object is by using coordination to implement
     strong consistency. For example, imagine a user issues a deposit of 10
-    dollars to replica A, and simultaneously, a user issues a deposit of 20
-    dollars to replica B. The two replicas can coordinate, for example by
-    running Paxos, and agree on a single total order in which to execute the
-    two transactions. Here, the replicas agree to deposit 10 dollars and then
-    deposit 20 dollars. Both transactions succeed.
+    dollars to replica A, and simultaneously, another user issues a deposit of
+    20 dollars to replica B. The two replicas can coordinate and agree on a
+    single total order in which to execute the two transactions. Here, the
+    replicas deposit 10 dollars and then deposit 20 dollars. Both transactions
+    succeed and our invariant is maintained.
   }
@@
     \begin{tikzpicture}
-      \drawbanks{\$50}{\$50}
-                {\$20}{\$20}
-                {\$20}{\$20}
-
+      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
+      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
       \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
       \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
-      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
-      \draw[op] (b3) -- ($(b3) + (+2, 0)$) node[midway, above] {NO};
+      \pause
 
       \zigzag{a1}{b1}
+      \node[bank] (a2) at (0, -3) {\$20};
+      \node[bank] (b2) at (\bankdistance, -3) {\$20};
+      \draw[op] (a1) to node[left] {$-\$30$} (a2);
+      \draw[op] (b1) to node[right] {$-\$30$} (b2);
+      \pause
+
       \zigzag{a2}{b2}
+      \node[bank] (a3) at (0, -6) {\$20};
+      \node[bank] (b3) at (\bankdistance, -6) {\$20};
+      \draw[op] (a2) to node[left] {$-\$40$} (a3);
+      \draw[op] (b2) to node[right] {$-\$40$} (b3);
     \end{tikzpicture}
@@
   \note{%
-    Now, assume two clients issue concurrent withdrawal transactions of 30
-    dollars and 40 dollars. Again, the two replicas coordinate and agree an a
-    total order in which to execute transactions.
-
-    They first withdraw 30 dollars, bringing my bank account balance to 20
-    dollars. They then try to withdraw of 40 dollars but abort the transaction
-    because I only have 40 dollars left. And while the transaction failed,
-    you'll note that we never violated our application invariant.
+    Now, consider what happens when two clients issue concurrent withdrawal
+    transactions of 30 dollars and 40 dollars. Again, the two replicas
+    coordinate. They first withdraw 30 dollars, bringing my bank account
+    balance to 20 dollars. They then try to withdraw of 40 dollars but abort
+    the transaction because I only have 20 dollars left. Note that even though
+    the withdrawal of 40 dollars failed, our invariant is always maintained.
   }
@@
     \begin{tikzpicture}
-      \drawbanks{\$50}{\$50}
-                {\$60}{\$70}
-                {\$80}{\$80}
-
+      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
+      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
       \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
       \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
-      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
-      \draw[op] (b2) -- ($(b2) + (+2, 0)$) node[midway, above] {OK};
+      \pause
 
+      \node[bank] (a2) at (0, -3) {\$60};
+      \node[bank] (b2) at (\bankdistance, -3) {\$70};
+      \draw[op] (a1) to node[left] {$+\$10$} (a2);
+      \draw[op] (b1) to node[right] {$+\$20$} (b2);
+      \pause
+
+      \node[bank] (a3) at (0, -6) {\$80};
+      \node[bank] (b3) at (\bankdistance, -6) {\$80};
+      \draw[op] (a2) to (a3);
+      \draw[op] (b2) to (b3);
       \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$+\$10$};
@@
     Alternatively, instead of implementing our bank with strong consistency, we
-    could implement it with weak consistency and avoid coordination altogether.
+    can implement it with weak consistency and avoid coordination altogether.
+    \\[12pt]
 
@@
     replicas execute the transactions immediately without consulting one
-    another. And every once in a while, the two replicas communicate and fill
-    each other in on the transactions they missed. Notice that even though we
-    didn't perform any coordination, the two replicas never violate my critical
-    invariant of a non-negative balance.
+    another. \\[12pt]
+
+    And every once in a while, the two replicas communicate and fill each other
+    in on the transactions they missed. Notice that even though we didn't
+    perform any coordination, the two replicas never violate the invariant.
+    \\[12pt]
   }
@@
     \begin{tikzpicture}
-      \drawbanks{\$50}{\$50}
-                {\$20}{\$10}
-                {-\$20}{-\$20}
-
+      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
+      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
       \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
       \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
-      \draw[op] (a2) -- ($(a2) + (-2, 0)$) node[midway, above] {OK};
-      \draw[op] (b2) -- ($(b2) + (+2, 0)$) node[midway, above] {OK};
+      \pause
 
+      \node[bank] (a2) at (0, -3) {\$20};
+      \node[bank] (b2) at (\bankdistance, -3) {\$10};
+      \draw[op] (a1) to node[left] {$-\$30$} (a2);
+      \draw[op] (b1) to node[right] {$-\$40$} (b2);
+      \pause
+
+      \node[bank] (a3) at (0, -6) {-\$20};
+      \node[bank] (b3) at (\bankdistance, -6) {-\$20};
+      \draw[op] (a2) to (a3);
+      \draw[op] (b2) to (b3);
       \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$-\$30$};
@@
   \note{%
-    However, let's look what happens when two clients issue concurrent
+    However, let's look at what happens when two clients issue concurrent
     withdrawals. Both withdrawals are executed locally and both succeed.
@@
 \begin{frame}
+  \Large
+
   \begin{center}
-    \Huge
-    Deposits don't require coordination to maintain invariants, but withdrawals
-    do.
+    \begin{enumerate}
+      \item
+        Strong consistency preserves the invariant.
+      \pause \item
+        Deposits don't require coordination.
+      \pause \item
+        Withdrawals do require coordination.
+    \end{enumerate}
   \end{center}
@@
   \note{%
-    We can learn a lot from these examples.
+    We can learn a lot from these examples. \\[12pt]
 
     First, if we implement our bank with strong consistency, we never violate
-    our invariant.
+    our invariant. \\[12pt]
 
     Second, we can implement deposit transactions without any coordination and
-    we're still guaranteed to never violate the invariant.
+    we're still guaranteed to never violate the invariant. \\[12pt]
 
     Third, withdrawal transactions, unlike deposit transactions, do require
-    coordination to avoid violating the invariant.
+    coordination to avoid violating the invariant. \\[12pt]
 
@@
     In this example, deposit transactions are invariant confluent and thus do
-    not require coordination while withdrawal transactions are not invariant
-    confluent and thus do require coordination.
+    not require coordination, while withdrawal transactions not invariant
+    confluent and thus do require coordination. \\[12pt]
   }

### 2019-08-20 doc/vldb_slides/sections/four_wise_men.tex
Slimmed down talk.
Progress towards slide revisions.

--- 2019-08-16 doc/vldb_slides/sections/four_wise_men.tex
@@
+\begin{frame}
   \tikzstyle{pic}=[inner sep=0pt]
-\tikzstyle{quote}=[fill=gray!30, inner sep=12pt]
+  \tikzstyle{thought}=[cloud, draw, aspect=4]
 
-\newcommand{\vogels}{
-  \node[pic] (vogels) at (0, 0) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=0 75 0 25,
-      clip
-    ]{assets/vogels.jpg}
+  \begin{center}
+    \begin{tikzpicture}[xscale=3, yscale=3]
+      \node[pic] at (0, 0) {%
+        \includegraphics[width=0.2\textwidth]{assets/person.jpg}
       };
-}
+      \pause
 
-\newcommand{\hoare}{
-  \node[pic, right=0.2cm of vogels] (hoare) {
-    \includegraphics[width=0.22\textwidth]{assets/hoare.jpg}
+      \node[thought, aspect=1] at (-0.5, 0.5) {};
+      \node[thought] at (-1, 1) {Weak Consistency?};
+      \pause
+
+      \node[thought, aspect=1] at (0.5, 0.5) {};
+      \node[thought] at (1, 1) {Strong Consistency?};
+      \pause
+
+      \node[anchor=west] at (-1.75, 0.25) {%
+        \textcolor{flatgreen}{\texttt{+}} Super fast
       };
-}
+      \node[anchor=west] at (-1.75, 0) {%
+        \textcolor{flatred}{\texttt{-}} Hard to reason about
+      };
+      \pause
 
-\newcommand{\lamport}{
-  \node[pic, right=0.2cm of hoare] (lamport) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=250 700 300 200,
-      clip
-    ]{assets/lamport.jpg}
+      \node[anchor=west] at (0.3, 0.25) {%
+        \textcolor{flatgreen}{\texttt{+}} Easy to reason about
       };
-}
+      \node[anchor=west] at (0.3, 0) {%
+        \textcolor{flatred}{\texttt{-}} Sacrifices availability
+      };
+    \end{tikzpicture}
+  \end{center}
 
-\newcommand{\brewer}{
-  \node[pic, right=0.2cm of lamport] (brewer) {
-    \includegraphics[
-      width=0.22\textwidth,
-      trim=100 300 100 100,
-      clip
-    ]{assets/brewer.jpg}
-  };
-}
+  \note{%
+    Imagine you're an application developer and you have some data that you
+    want to replicate. But, you're not sure whether to replicate the data with
+    weak consistency or with strong consistency. \\[12pt]
 
-\begin{frame}
-  \note{%
-    Hi everyone! Thanks so much for attending. I'm Michael, and today I'll be
-    talking about ``Interactive Checks for Coordination Avoidance''. This is
-    joint work with my advisor at UC Berkeley, Joe Hellerstein.
-
-    Imagine you're an application developer and you're trying to decide what
-    kind of consistency you'd like to use to replicate some of your data.
+    You can implement weak consistency super fast, but weakly consistent
+    systems are hard to reason about. Strongly consistent systems are much
+    easier to reason about, but they come at the cost of availability and
+    performance. \\[12pt]
   }
@@
 \begin{frame}
+  \tikzstyle{pic}=[inner sep=0pt]
+  \tikzstyle{quote}=[fill=gray!30, inner sep=12pt]
+
   \begin{center}
     \begin{tikzpicture}[xscale=1.3]
-      \vogels
+      \node[pic] (bailis) at (0, 0) {%
+        \includegraphics[width=0.25\textwidth]{assets/bailis.jpg}
+      };
       \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{You should use weak consistency. It's wicked fast!}
-      };
-      \path[fill=gray!30] (vogels.north) -- ++(0, 1) -- ++(0.5, 0) -- (vogels.north);
-    \end{tikzpicture}
-  \end{center}
-
-  \note{%
-    Someone like Werner Vogels would probably tell you to use weak consistency.
-    You can implement weak consistency super duper fast, and for many
-    applications, weak consistency is good enough.
-  }
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \node[quote, text width=0.8\textwidth, align=left] at (1, 3) {
-        \Large\textit{What about the application invariants?!}
-      };
-      \path[fill=gray!30] (hoare.north) -- ++(0, 1.5) -- ++(1, 0) -- (hoare.north);
-    \end{tikzpicture}
-  \end{center}
-
-  \note{%
-    But then Tony Hoare laments, ``no no no''. Hoare encourages us not to
-    settle for weak consistency because a weakly consistent system is hard to
-    reason about. It robs us of a fundamental way we use to reason about
-    programs: application invariants. Even if every transaction in a weakly
-    consistent system preserves application invariants, the whole system may
-    not.
-  }
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \lamport
-      \node[quote, text width=0.8\textwidth, align=left] at (2, 3) {
-        \Large\textit{Yeah, go with strong consistency! I have an algorithm for that :)}
-      };
-      \path[fill=gray!30] (lamport.north) -- ++(0, 1.5) -- ++(1, 0) -- (lamport.north);
-    \end{tikzpicture}
-  \end{center}
-
-  \note{%
-    Leslie Lamport chimes in agreeingly. He encourages us to choose strong
-    consistency. Strongly consistent systems are much easier to reason about.
-    After all a strongly consistent system to end users looks exactly like a
-    system that's not even distributed.
-  }
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \vogels
-      \hoare
-      \lamport
-      \brewer
-      \node[quote, text width=0.8\textwidth, align=left] at (3, 3) {
-        \Large\textit{Not so fast!}
-      };
-      \path[fill=gray!30] (brewer.north) -- ++(0, 1.5) -- ++(-1, 0) -- (brewer.north);
-    \end{tikzpicture}
-  \end{center}
-
-  \note{%
-    ``Aha, not so fast'' counters Eric Brewer! Consistency and availability and
-    partition tolerance, oh my. Surely, strong consistency comes at the price
-    of availability and performance.
-
-    So we're in a bit of a pickle. Weak consistency is too confusing and strong
-    consistency is too heavy handed.
-  }
-\end{frame}
-
-\begin{frame}
-  \begin{center}
-    \begin{tikzpicture}[xscale=1.3]
-      \node[pic] (bailis) at (0, 0) {\includegraphics[width=0.25\textwidth]{assets/bailis.jpg}};
-      \node[quote, text width=0.8\textwidth, align=left] at (0, 3) {
-        \Large\textit{?`Porque no los dos?}
+        \Large\textit{Why not have both?} Invariant confluence!
       };
@@
   \note{%
-    Thankfully, Peter Bailis comes to save the day and reminds us that perhaps
-    we can have the best of both worlds. In 2014, Peter and some other smart
-    folks including Alan defined invariant confluence as a condition under
-    which we can replicate a piece of data without the need for strongly
-    consistent coordination while maintaining the data's application
-    invariants.
+    Fortunately, Peter Bailis saves the day and reminds us of when in 2014, he
+    and some other smart folks including Alan and my advisor Joe defined
+    invariant confluence as a way to get the benefits of both weak and strong
+    consistency.
 
-    Invariant confluence will be the topic of this talk. In particular, I'll
-    define invariant confluence and then discuss how we can check invariant
-    confluence automatically.
+    Invariant confluence will be the topic of this talk, so let's build some
+    intuition on what invariant confluence is.
   }

### 2019-08-20 doc/vldb_slides/sections/iclosure.tex
More cuts.
Talk tweaks.

--- 2019-08-18 doc/vldb_slides/sections/iclosure.tex
@@
     about reachable states but reasoning automatically about reachable states
-    is hard.
+    is hard. \\[12pt]
 
@@
     at a sufficient condition for invariant confluence called invariant
-    closure.
+    closure. \\[12pt]
   }
@@
   \Large
-  $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to an invariant
-  $I$, abbreviated \defword{\Iclosed{}}, if invariant satisfying states are
-  closed under merge. \pause That is, for every state $s_1, s_2 \in S$, if
-  $I(s_1)$ and $I(s_2)$, then $I(s_1 \join s_2)$.
+  \begin{itemize}
+    \item
+      $O = (S, \join)$ is \defword{\invariantclosed{}} with respect to an
+      invariant $I$ if invariant satisfying states are closed under merge.
+    \pause\item
+      For every state $s_1, s_2 \in S$, if $I(s_1)$ and $I(s_2)$, then $I(s_1
+      \join s_2)$.
+  \end{itemize}
 
@@
   \]
-  \pause
-  Transactions perserve invariants. If merging does too, then all reachable
-  states satisfy the invariant.
 
   \note{%
-    Invariant closure is a sufficient condition for invariant confluence. Why
-    is that? Well for an object to be invariant confluent, every reachable
-    state must satisfy the invariant. A state is reachable if it can be reached
-    by executing transactions merging states. Transactions are already
-    guarnateed to preserve invariants, so if merging is also guaranteed to
-    preserve the invariant, then every reachable state always satisfies the
-    invariant.
+    Invariant closure is a sufficient condition for invariant confluence. I'll
+    refer you to the paper for the proof, but it's a very simple proof.
   }
@@
   Checking invariant closure is more straightforward.
+
+  \vspace{0.5in}
   \pause
-  \begin{align*}
+
+  \begin{tikzpicture}[xscale=6]
+    \node[draw] (formula) at (0, 0) {
+      $
+      \begin{aligned}
         \forall x_1, & y_1, x_2, y_2.\, \\
@@
         \quad & \max(x_1, x_2)\max(y_1, y_2) \leq 0
-  \end{align*}
+      \end{aligned}
+      $
+    };
+    \pause
+    \node (smt) at (1, 0) {SMT Solver};
+    \draw[-latex, ultra thick] (formula) to (smt);
+  \end{tikzpicture}
 
@@
     this formula, and we can pass this formula directly to an SMT solver to
-    figure out whether it's invariant closed. This is not something we can do
-    with invariant confluence because it's hard to express reachability as a
-    simple formula like this one.
+    figure out whether it's invariant closed.
   }
@@
 
-\newcommand{\xmin}{-2}
-\newcommand{\xmax}{2}
-\newcommand{\ymin}{-2}
-\newcommand{\ymax}{2}
-
-% Axes.
-\newcommand{\xyaxes}{
-  \draw[] (\xmin.5, 0) to (\xmax.5, 0);
-  \draw[] (0, \ymin.5) to (0, \ymax.5);
-  \node at (\xmax + 1, 0) {$x$};
-  \node at (0, \ymax + 1) {$y$};
-}
-
-% Quadrant 1.
-\newcommand{\quadi}[5]{{
-  \newcommand{\argstyle}{#1}
-  \newcommand{\argxmin}{#2}
-  \newcommand{\argxmax}{#3}
-  \newcommand{\argymin}{#4}
-  \newcommand{\argymax}{#5}
-  \foreach \x in {0, ..., \argxmax} {
-    \foreach \y in {0, ..., \argymax} {
-      \node[\argstyle] (\x-\y) at (\x, \y) {};
-    }
-  }
-}}
-
-% Quadrant 2.
-\newcommand{\quadii}[5]{{
-  \newcommand{\argstyle}{#1}
-  \newcommand{\argxmin}{#2}
-  \newcommand{\argxmax}{#3}
-  \newcommand{\argymin}{#4}
-  \newcommand{\argymax}{#5}
-  \foreach \x in {\argxmin, ..., 0} {
-    \foreach \y in {0, ..., \argymax} {
-      \node[\argstyle] (\x-\y) at (\x, \y) {};
-    }
-  }
-}}
-
-% Quadrant 3.
-\newcommand{\quadiii}[5]{{
-  \newcommand{\argstyle}{#1}
-  \newcommand{\argxmin}{#2}
-  \newcommand{\argxmax}{#3}
-  \newcommand{\argymin}{#4}
-  \newcommand{\argymax}{#5}
-  \foreach \x in {\argxmin, ..., 0} {
-    \foreach \y in {\argymin, ..., 0} {
-      \node[\argstyle] (\x-\y) at (\x, \y) {};
-    }
-  }
-}}
-
-% Quadrant 4.
-\newcommand{\quadiv}[5]{{
-  \newcommand{\argstyle}{#1}
-  \newcommand{\argxmin}{#2}
-  \newcommand{\argxmax}{#3}
-  \newcommand{\argymin}{#4}
-  \newcommand{\argymax}{#5}
-  \foreach \x in {0, ..., \argxmax} {
-    \foreach \y in {\argymin, ..., 0} {
-      \node[\argstyle] (\x-\y) at (\x, \y) {};
-    }
-  }
-}}
-
-% State labels.
-\newcommand{\statelabels}{
-  \node[statelabel] at (0, 0) {$s_0$};
-  \node[statelabel] at (-1, 1) {$s_1$};
-  \node[statelabel] at (1, -1) {$s_2$};
-  \node[statelabel] at (1, 1) {$s_3$};
-}
-
-\tikzstyle{point}=[shape=circle, fill=flatgray, inner sep=3pt]
-\tikzstyle{inv}=[line width=0.75pt, draw=black]
-\tikzstyle{pointinv}=[point, inv]
-\tikzstyle{invregion}=[rounded corners, fill=flatgreen!50, draw=none]
-\tikzstyle{reachableregion}=[rounded corners, fill=flatblue!50, draw=none]
-\tikzstyle{statelabel}=[anchor=south west, inner sep=1pt]
+\input{runningexample.tex}
 
@@
     We can check for invariant closure but not invariant confluence, so we'd
-    like the two to be equivalent.
+    like the two to be equivalent. \\[12pt]
 
     Are there any situations in which the two do happen to be equivalent?
+    \\[12pt]
   }
@@
   \]
-  \pause
-  \begin{itemize}
-    \item
-      Forward direction: same as before.
-    \item
-      Backward direction: $I$ is equal to the set of reachable states.
-      Reachable states are closed under merge, so so is $I$.
-  \end{itemize}
 
@@
 
-    Why? Well, the forward direction here is what we proved before. For the
-    backwards direction, we have that the invariant is a subset of the set of
-    reachable points, and because our object is invariant confluent, the set of
-    reachable points is a subset of the invariant. So the invariant and the set
-    of reachable points are the same. The set of reachable states is closed
-    under merge, so the invariant is also, so the object is invariant closed.
-    \\[12pt]
-
-    I know I've walked through a lot of theory pretty quickly, but if you
-    remember one slide, remember this one. If the invariant is a subset of the
-    reachable states, then invariant closure and invariant confluence are
-    equivalent. \\[12pt]
+    Again, I'll refer you to the paper for a proof. \\[12pt]
   }

### 2019-08-22 doc/vldb_slides/sections/bank_example.tex
Progress towards iconfluence slide revisions.

@@
-\newcommand{\bankdistance}{4}
-\tikzstyle{bank}=[draw, minimum width=1.5cm, minimum height=1cm, font=\huge]
+\newcommand{\bankdistance}{3}
+\tikzstyle{bank}=[draw, minimum width=1.5cm, minimum height=1cm, font=\huge,
+                  fill=white, rounded corners]
 \tikzstyle{op}=[-latex, ultra thick]
+\tikzstyle{pline}=[gray, opacity=0.75, thick, -latex]
+\newcommand{\plusten}{\textcolor{flatgreen}{+\$10}}
+\newcommand{\plustwenty}{\textcolor{flatblue}{+\$20}}
+\newcommand{\minusthirty}{\textcolor{flatgreen}{-\$30}}
+\newcommand{\minusforty}{\textcolor{flatblue}{-\$40}}
 
@@
   \newcommand{\bankb}{#2}
-  \draw[gray, op] ($(\banka) + (0, -0.7)$) -- ($(\bankb) + (0, -1.0)$);
-  \draw[gray, op] ($(\bankb) + (0, -1.0)$) -- ($(\banka) + (0, -1.3)$);
-  \draw[gray, op] ($(\banka) + (0, -1.3)$) -- ($(\bankb) + (0, -1.6)$);
-  \draw[gray, op] ($(\bankb) + (0, -1.6)$) -- ($(\banka) + (0, -1.9)$);
+  \draw[gray, op] ($(\banka) + (1.0, 0)$) -- ($(\bankb) + (1.5, 0)$);
+  \draw[gray, op] ($(\bankb) + (1.6, 0)$) -- ($(\banka) + (2.0, 0)$);
+  \draw[gray, op] ($(\banka) + (2.1, 0)$) -- ($(\bankb) + (2.5, 0)$);
+  \draw[gray, op] ($(\bankb) + (2.6, 0)$) -- ($(\banka) + (3.0, 0)$);
 }}
@@
     \begin{tikzpicture}
-      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
-      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance, 0) {\$50};
+      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
+      \draw[pline] (a1) to (9, \bankdistance);
+      \draw[pline] (b1) to (9, 0);
     \end{tikzpicture}
@@
     \begin{tikzpicture}
-      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
-      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
-      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
-      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
+      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_1$}] (b1) at (0,  0) {\$50};
+      \draw[pline] (a1) to (9, \bankdistance);
+      \draw[pline] (b1) to (9, 0);
+      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
+      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
       \pause
@@
       \zigzag{a1}{b1}
-      \node[bank] (a2) at (0, -3) {\$60};
-      \node[bank] (b2) at (\bankdistance, -3) {\$60};
-      \draw[op] (a1) to node[left] {$+\$10$} (a2);
-      \draw[op] (b1) to node[right] {$+\$10$} (b2);
+      \node[bank] (a2) at (4, \bankdistance) {\$60};
+      \node[bank] (b2) at (4, 0) {\$60};
+      \draw[op] (a1) to node[above] {$\plusten$} (a2);
+      \draw[op] (b1) to node[below] {$\plusten$} (b2);
       \pause
@@
       \zigzag{a2}{b2}
-      \node[bank] (a3) at (0, -6) {\$80};
-      \node[bank] (b3) at (\bankdistance, -6) {\$80};
-      \draw[op] (a2) to node[left] {$+\$20$} (a3);
-      \draw[op] (b2) to node[right] {$+\$20$} (b3);
+      \node[bank] (a3) at (8, \bankdistance) {\$80};
+      \node[bank] (b3) at (8, 0) {\$80};
+      \draw[op] (a2) to node[above] {$\plustwenty$} (a3);
+      \draw[op] (b2) to node[below] {$\plustwenty$} (b3);
     \end{tikzpicture}
@@
     \begin{tikzpicture}
-      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
-      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
-      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
-      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
+      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
+      \draw[pline] (a1) to (9, \bankdistance);
+      \draw[pline] (b1) to (9, 0);
+      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
+      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
       \pause
@@
       \zigzag{a1}{b1}
-      \node[bank] (a2) at (0, -3) {\$20};
-      \node[bank] (b2) at (\bankdistance, -3) {\$20};
-      \draw[op] (a1) to node[left] {$-\$30$} (a2);
-      \draw[op] (b1) to node[right] {$-\$30$} (b2);
+      \node[bank] (a2) at (4, \bankdistance) {\$20};
+      \node[bank] (b2) at (4, 0) {\$20};
+      \draw[op] (a1) to node[above] {$\minusthirty$} (a2);
+      \draw[op] (b1) to node[below] {$\minusthirty$} (b2);
       \pause
@@
       \zigzag{a2}{b2}
-      \node[bank] (a3) at (0, -6) {\$20};
-      \node[bank] (b3) at (\bankdistance, -6) {\$20};
-      \draw[op] (a2) to node[left] {$-\$40$} (a3);
-      \draw[op] (b2) to node[right] {$-\$40$} (b3);
+      \node[bank] (a3) at (8, \bankdistance) {\$20};
+      \node[bank] (b3) at (8, 0) {\$20};
+      \draw[op] (a2) to node[above] {$\minusforty$} (a3);
+      \draw[op] (b2) to node[below] {$\minusforty$} (b3);
     \end{tikzpicture}
@@
     \begin{tikzpicture}
-      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
-      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
-      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$+\$10$};
-      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$+\$20$};
+      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
+      \draw[pline] (a1) to (9, \bankdistance);
+      \draw[pline] (b1) to (9, 0);
+      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\plusten$};
+      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\plustwenty$};
       \pause
 
-      \node[bank] (a2) at (0, -3) {\$60};
-      \node[bank] (b2) at (\bankdistance, -3) {\$70};
-      \draw[op] (a1) to node[left] {$+\$10$} (a2);
-      \draw[op] (b1) to node[right] {$+\$20$} (b2);
+      \node[bank] (a2) at (4, \bankdistance) {\$60};
+      \node[bank] (b2) at (4, 0) {\$70};
+      \draw[op] (a1) to node[above] {$\plusten$} (a2);
+      \draw[op] (b1) to node[below] {$\plustwenty$} (b2);
       \pause
 
-      \node[bank] (a3) at (0, -6) {\$80};
-      \node[bank] (b3) at (\bankdistance, -6) {\$80};
-      \draw[op] (a2) to (a3);
-      \draw[op] (b2) to (b3);
-      \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$+\$10$};
-      \draw[dashed, op] (b2) -- (a3) node[near start, sloped, above] {$+\$20$};
+      \node[bank] (a3) at (8, \bankdistance) {\$80};
+      \node[bank] (b3) at (8, 0) {\$80};
+      \draw[op] (a2) to node[midway, above] {$\plustwenty$} (a3);
+      \draw[op] (b2) to node[midway, below] {$\plusten$} (b3);
+      \draw[dashed, op] (a2) -- ($(b3) + (-2, 0)$);
+      \draw[dashed, op] (b2) -- ($(a3) + (-2, 0)$);
     \end{tikzpicture}
@@
     \begin{tikzpicture}
-      \node[bank, label=above:{Replica $A$}] (a1) at (0, 0) {\$50};
-      \node[bank, label=above:{Replica $B$}] (b1) at (\bankdistance,  0) {\$50};
-      \draw[op] ($(a1) + (-2, 0)$) -- (a1) node[midway, above] {$-\$30$};
-      \draw[op] ($(b1) + (+2, 0)$) -- (b1) node[midway, above] {$-\$40$};
+      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
+      \draw[pline] (a1) to (9, \bankdistance);
+      \draw[pline] (b1) to (9, 0);
+      \draw[op] ($(a1) + (0, +2)$) -- (a1) node[midway, left] {$\minusthirty$};
+      \draw[op] ($(b1) + (0, -2)$) -- (b1) node[midway, left] {$\minusforty$};
       \pause
 
-      \node[bank] (a2) at (0, -3) {\$20};
-      \node[bank] (b2) at (\bankdistance, -3) {\$10};
-      \draw[op] (a1) to node[left] {$-\$30$} (a2);
-      \draw[op] (b1) to node[right] {$-\$40$} (b2);
+      \node[bank] (a2) at (4, \bankdistance) {\$20};
+      \node[bank] (b2) at (4, 0) {\$10};
+      \draw[op] (a1) to node[above] {$\minusthirty$} (a2);
+      \draw[op] (b1) to node[below] {$\minusforty$} (b2);
       \pause
 
-      \node[bank] (a3) at (0, -6) {-\$20};
-      \node[bank] (b3) at (\bankdistance, -6) {-\$20};
-      \draw[op] (a2) to (a3);
-      \draw[op] (b2) to (b3);
-      \draw[dashed, op] (a2) -- (b3) node[near start, sloped, above] {$-\$30$};
-      \draw[dashed, op] (b2) -- (a3) node[near start, sloped, above] {$-\$40$};
+      \node[bank] (a3) at (8, \bankdistance) {-\$20};
+      \node[bank] (b3) at (8, 0) {-\$20};
+      \draw[op] (a2) to node[above] {$\minusforty$} (a3);
+      \draw[op] (b2) to node[below] {$\minusthirty$} (b3);
+      \draw[dashed, op] (a2) -- ($(b3) + (-2, 0)$);
+      \draw[dashed, op] (b2) -- ($(a3) + (-2, 0)$);
     \end{tikzpicture}

### 2019-08-22 doc/vldb_slides/sections/iclosure.tex
Finished slide revisions.

@@
 \begin{frame}
-  \Huge
   \begin{center}
-    Reasoning about reachable states is hard.
+    \begin{tikzpicture}[xscale=3, yscale=2]
+      \node (confluence) at (2, 0) {\Large invariant confluence};
+      \pause
+
+      \node (hard) at (2, 1) {hard to check :(};
+      \draw[-latex, thick] (hard) to (confluence);
+      \pause
+
+      \node (closure) at (0, 0) {\Large invariant closure};
+      \node at ($(confluence)!0.5!(closure)$) {\Large $\implies$};
+      \pause
+
+      \node (easy) at (0, 1) {easy to check :)};
+      \draw[-latex, thick] (easy) to (closure);
+    \end{tikzpicture}
   \end{center}
@@
 
-    Instead of reasoning about invariant confluence directly then, let's look
+    Instead of reasoning about invariant confluence directly then, we'll look
     at a sufficient condition for invariant confluence called invariant
     closure. \\[12pt]
+
+    Unlike invariant confluence, we'll see that invariant closure is easy to
+    check. \\[12pt]
   }
@@
   \note{%
-    Invariant closure is a sufficient condition for invariant confluence. I'll
-    refer you to the paper for the proof, but it's a very simple proof.
+    As I mentioned earlier, invariant closure is a sufficient condition for
+    invariant confluence. I'll refer you to the paper for the proof, but it's a
+    very simple proof.
   }
@@
 \begin{frame}
-  \Large
-  \[
-    \text{invariant closure} \implies \text{invariant confluence}
-  \]
+  \begin{center}
+    \begin{tikzpicture}[xscale=3, yscale=2]
+      \node (confluence) at (2, 0) {\Large invariant confluence};
+      \node (closure) at (0, 0) {\Large invariant closure};
+      \node at ($(confluence)!0.5!(closure)$) {\Large $\implies$};
+      \node (hard) at (2, 1) {hard to check :(};
+      \draw[-latex, thick] (hard) to (confluence);
+      \node (easy) at (0, 1) {easy to check :)};
+      \draw[-latex, thick] (easy) to (closure);
+    \end{tikzpicture}
+  \end{center}
 
@@
     \text{invariant closure}
-    \xLeftarrow{\phantom{a}?\phantom{a}}
+    \xLeftarrow{\phantom{aa}?\phantom{aa}}
     \text{invariant confluence}
@@
 \begin{frame}
-  \Large
-  \[
-    \text{invariant closure} \centernot\impliedby \text{invariant confluence}
-  \]
+  \begin{center}
+    \begin{tikzpicture}[xscale=3, yscale=2]
+      \node (confluence) at (2, 0) {\Large invariant confluence};
+      \node (closure) at (0, 0) {\Large invariant closure};
+      \node at ($(confluence)!0.5!(closure)$) {\Large $\centernot\impliedby$};
+      \node (hard) at (2, 1) {hard to check :(};
+      \draw[-latex, thick] (hard) to (confluence);
+      \node (easy) at (0, 1) {easy to check :)};
+      \draw[-latex, thick] (easy) to (closure);
+    \end{tikzpicture}
+  \end{center}
 
@@
 \end{frame}
+
+\begin{frame}
+  \Large
+  If $I$ is a subset of reachable states, then
+  \begin{center}
+    \begin{tikzpicture}[xscale=3, yscale=2]
+      \node (confluence) at (2, 0) {\Large invariant confluence};
+      \node (closure) at (0, 0) {\Large invariant closure};
+      \node at ($(confluence)!0.5!(closure)$) {\Large $\iff$};
+      \node (hard) at (2, 1) {hard to check :(};
+      \draw[-latex, thick] (hard) to (confluence);
+      \node (easy) at (0, 1) {easy to check :)};
+      \draw[-latex, thick] (easy) to (closure);
+    \end{tikzpicture}
+  \end{center}
+
+  \note{%
+    This is good news. Invariant confluence is hard to check, but invariant
+    closure is easy to check. And when the invariant is a subset of the
+    reachable states, the two are equivalent. In this case, invariant
+    confluence also becomes easy to check.
+  }
+\end{frame}

### 2019-09-25 doc/vldb_slides/sections/bank_example.tex
Final edits.

@@
     \begin{tikzpicture}
-      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
-      \node[bank, label=left:{$p_1$}] (b1) at (0, 0) {\$50};
+      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_2$}] (b1) at (0, 0) {\$50};
       \draw[pline] (a1) to (9, \bankdistance);
@@
     bank account balance. We decide to replicate the integer across two
-    machines, replica A and replica B. \\[12pt]
+    machines, $p_1$ and $p_2$. \\[12pt]
 
-    Users can issue transactions to our replicate object, and every
+    Users can issue transactions to our replicated bank, and every
     transacation is either a deposit or a withdrawal. And, I have one critical
@@
     \begin{tikzpicture}
-      \node[bank, label=left:{$p_0$}] (a1) at (0, \bankdistance) {\$50};
-      \node[bank, label=left:{$p_1$}] (b1) at (0,  0) {\$50};
+      \node[bank, label=left:{$p_1$}] (a1) at (0, \bankdistance) {\$50};
+      \node[bank, label=left:{$p_2$}] (b1) at (0,  0) {\$50};
       \draw[pline] (a1) to (9, \bankdistance);
@@
     strong consistency. For example, imagine a user issues a deposit of 10
-    dollars to replica A, and simultaneously, another user issues a deposit of
-    20 dollars to replica B. The two replicas can coordinate and agree on a
+    dollars to $p_1$, and simultaneously, another user issues a deposit of
+    20 dollars to $p_2$. The two replicas can coordinate and agree on a
     single total order in which to execute the two transactions. Here, the

### 2020-01-13 doc/vldb_paper/figures/merge_reducible.tex
Renamed paper/ to vldb_paper/.
I'm working on the VLDB Journal submission now, so there's actually two
papers to deal with.

--- 2018-03-31 doc/paper/figures/merge_reducible.tex
@@
     \draw[rounded corners] (7, 1.5) rectangle ++(7, 2);
-    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\small Invariant-Confluent};
-    \node[inner sep=0pt, anchor=north] at (6, 6) {\small Invariant-Closed};
-    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\small Criteria (1) -- (4)};
-    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\small Merge-Reducible};
+    \node[inner sep=0pt, anchor=north] at (8, 7.5) {\scriptsize Invariant Confluent};
+    \node[inner sep=0pt, anchor=north] at (6, 6) {\scriptsize Invariant Closed};
+    \node[inner sep=0pt, anchor=south] at (10.5, 1.5) {\scriptsize Criteria (1) -- (4)};
+    \node[inner sep=0pt, anchor=south] at (10.5, 0.5) {\scriptsize Merge Reducible};
   \end{tikzpicture}
   \caption{%
-    The relationship between invariant-closure, merge-reducibility, criteria
-    (1) -- (4) from \thmref{LatticeProperty}, and invariant-confluence.
+    The relationship between \invariantclosure{}, \mergereducibility{},
+    criteria (1) -- (4) from \thmref{LatticeProperty}, and
+    \invariantconfluence{}.
   }\figlabel{IclosureVsReducible}

### 2020-01-13 doc/vldb_paper/sections/conclusion.tex
Renamed paper/ to vldb_paper/.
I'm working on the VLDB Journal submission now, so there's actually two
papers to deal with.

\section{Conclusion}
This paper revolved around two major contributions.
%
First, we developed a deeper understanding of \invariantclosure{} and
\invariantconfluence{} by looking at the two criteria with reachability in
mind. We found that \invariantclosure{} fails to incorporate a notion of
reachability, and using this intuition, we developed conditions under which
\invariantclosure{} and \invariantconfluence{} are equivalent. We implemented
this insight in an interactive \invariantconfluence{} decision procedure
that automatically checks whether an object is \invariantconfluent{}, with the
assistance of a programmer.

Second, we proposed a new consistency model and generalization of
\invariantconfluence{}, segmented \invariantconfluence{}, that can be used to
replicate non-\invariantconfluent{} objects with a small amount of coordination
while still preserving their invariants. We found that segmented
\invariantconfluence{} naturally subsumes existing techniques for maintaining
invariants of replicated objects (e.g.\ locking and escrow transactions), and
we developed an interactive decision procedure for segmented
\invariantconfluence{}.

Through our evaluation, we found that our decision procedures could analyze a
number of realistic workloads, each in less than a second. We also showed that
segmented \invariantconfluence{} can significantly outperform linearizable
replication for low-coordination workloads.

\textbf{Acknowledgments.}
The authors would like to thank Alan Fekete, Peter Alvaro, Alvin Cheung,
Alexandra Meliou, Anthony Tan, and Cristina Teodoropol for fruitful discussion
and feedback.
%
This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
Scotiabank, Splunk and VMware.

### 2020-01-13 doc/vldb_paper/sections/related_work.tex
Renamed paper/ to vldb_paper/.
I'm working on the VLDB Journal submission now, so there's actually two
papers to deal with.

--- 2018-09-19 doc/paper/sections/related_work.tex
@@
 determine whether a given token scheme is sufficient to guarantee that a global
-invariant is never broken. This model of consistency is very similar to
-segmented \invariantconfluence{}. The token based approach allows users to
+invariant is never broken. The token based approach allows users to
 specify certain conflicts that are not possible with segmented
-\invariantconfluence{}. A segmentation only allows transactions within a
+\invariantconfluence{} because a segmentation only allows transactions within a
 segment to acquire a single self-conflicting lock. However, segmented
@@
 Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
-exploiting the nice algebraic properties of semilattices.
+exploiting the algebraic properties of semilattices.
 
@@
 invariant maintenance during program execution.  Moreover, Bloom and Dedalus
-are general purpose programming languages that can be used to implement a
+are general-purpose programming languages that can be used to implement a
 variety of distributed systems that are outside of the scope of
@@
 
-\textbf{Program Analysis in Database Systems.}
-Using program analysis to improve the performance of database systems is not
-new. For example, it has been used to improve the performance of
-database-backed web applications~\cite{cheung2014using, wu2016transaction,
+\begin{techreport}
+
+  \textbf{Program Analysis in Database Systems.} Using program analysis to
+  improve the performance of database systems is not new. For example, it has
+  been used to improve the performance of database-backed web
+  applications~\cite{cheung2014using, wu2016transaction,
   ramachandra2012program} and used to improve the performance of optimistic
   concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
-on \invariantconfluence{} continues the theme of using program analysis to reap
-the performance benefits gained from understanding program semantics.
+  on \invariantconfluence{} continues the theme of using program analysis to
+  reap the performance benefits gained from understanding program semantics.
+\end{techreport}

### 2020-01-13 doc/the_vldb_journal_paper/sections/introduction.tex
First formatting of paper for The VLDB Journal.
Sitting around 15 pages of content. A 30% increase (as recommended by
the editors) would put us at about 20 pages of content, maybe 22 pages
overall. 25 page limit, so we should be good.

--- 2018-09-19 doc/paper/sections/introduction.tex
@@
 data with strong consistency using a technique like distributed transactions
-(e.g., \cite{bernstein1981concurrency,mohan1986transaction}) or state machine
-replication (e.g., \cite{schneider1990implementing, lamport1998part,
+(e.g.,~\cite{bernstein1981concurrency,mohan1986transaction}) or state machine
+replication (e.g.,~\cite{schneider1990implementing, lamport1998part,
 liskov2012viewstamped, ongaro2014search}) makes the application designer's life
@@
 proofs of this sort are unreasonable to expect from programmers. Ideally we
-would have a general purpose program that can automatically determine
+would have a general-purpose program that can automatically determine
 \invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
@@
 even in programs that require it}, by generalizing \invariantconfluence{} to a
-local property called \emph{segmented \invariantconfluence{}}. While
+property called \emph{segmented \invariantconfluence{}}. While
 \invariantconfluence{} characterizes objects that can be replicated
@@
 more than 66 lines of code to specify. Moreover, we find that segmented
-\invariantconfluent{} replication can achieve 100x to 10x more throughput than
+\invariantconfluent{} replication can achieve 10x to 100x more throughput than
 linearizable replication for low-coordination workloads.

### 2020-01-13 doc/the_vldb_journal_paper/sections/merge_reduction.tex
Added tech report sections of paper.

--- 2018-09-19 doc/paper/sections/merge_reduction.tex
@@
   transactions $T$, and an invariant $I$, if the following criteria are met,
-  then $O$ is \sTIconfluent{}.
+  then $O$ is \sTImergereducible{} (and therefore \sTIconfluent{}).
   \begin{enumerate}
     \item
-      $O$ is a semilattice.
+      $O$ is a join-semilattice.
 
@@
 
-The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
-through (4) are met, then $O$ is \sTImergereducible{}. From this, it's
-immediate that $O$ is \sTIconfluent{} by
-\thmref{ReducibilityImpliesIconfluence}. To show that $O$ is
-\sTImergereducible, we consider two merge-free reachable expressions
+To show that $O$ is \sTImergereducible, we consider two merge-free reachable
+expressions
 \[
@@
 \invariantclosure{} does not imply criteria (1) -- (4). Conversely, there are
-some objects which satisfy criteria (1) -- (4) which are not
+some objects that satisfy criteria (1) -- (4) that are not
 \invariantclosed{}. Here's one example.
@@
   set of the natural numbers. Our start state $s_0 = \set{0}$ is the set of
-  $0$. Let $t_Y(X) = X \cup Y$ be the transaction which unions $Y$ with its
+  $0$. Let $t_Y(X) = X \cup Y$ be the transaction that unions $Y$ with its
   argument $X$. Our set $T = \setst{t_Y}{Y \subseteq \nats}$ of transactions
@@
 
-  Clearly, criteria (1), (2), and (4) are satisfied. Criterion (3) is also
-  satisfied. If $s$ is an arbitrary state that satisfies $I$, then it is
-  non-empty and contains, without loss of generality, only even integers. If
-  $t_1$ and $t_2$ are arbitrary transactions such that $I(t_1(s))$ and
-  $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty and contain only
-  even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly non-empty and contains
-  only even integers, so $I(t_1(s) \join t_2(s))$.
-  %
-  However, $O$ is not \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$.
-  Then, $I(s_1)$ and $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$,
-  $\lnot I(s_3)$.
+  Criteria (1), (2), (3) and (4) are all satisfied. However, $O$ is not
+  \Iclosed{}. Let $s_1 = \set{0}$ and $s_2 = \set{1}$. Then, $I(s_1)$ and
+  $I(s_2)$, but letting $s_3 = s_1 \cup s_2 = \set{0, 1}$, $\lnot I(s_3)$.
+
+  Here's why criterion (3) is satisfied. If $s$ is an arbitrary state that
+  satisfies $I$, then it is non-empty and contains, without loss of generality,
+  only even integers. If $t_1$ and $t_2$ are arbitrary transactions such that
+  $I(t_1(s))$ and $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty
+  and contain only even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly
+  non-empty and contains only even integers, so $I(t_1(s) \join t_2(s))$.
 \end{example}
@@
 closed under merge. The fundamental difference is that criterion (3) restricts
-its attention to the merge of two states which are \emph{reachable} from a
+its attention to the merge of two states that are \emph{reachable} from a
 common ancestor state.

### 2020-01-13 doc/the_vldb_journal_paper/sections/introduction.tex
Minor introduction formatting changes.
No content changes, just some formatting changes.

@@
 brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
-models like eventual consistency~\cite{vogels2009eventually}, PRAM
+models like eventual consistency \cite{vogels2009eventually}, PRAM
 consistency~\cite{lipton1988pram}, causal consistency~\cite{ahamad1995causal},
@@
 availability too? Can we replicate a piece of data with weak consistency but
-still ensure that its invariants are maintained? Yes... sometimes. Bailis et
+still ensure that its invariants are maintained? Yes...\ sometimes. Bailis et
 al.\ introduced the notion of \emph{\invariantconfluence{}} as a necessary and
@@
 into \emph{segments}, with a restricted set of transactions allowed in each
-segment.  Within a segment servers act without any coordination; they
+segment. Within a segment, servers act without any coordination; they
 synchronize only to transition across segment boundaries. This design
@@
 more than 66 lines of code to specify. Moreover, we find that segmented
-\invariantconfluent{} replication can achieve 10x to 100x more throughput than
-linearizable replication for low-coordination workloads.
+\invariantconfluent{} replication can achieve 10$\times$ to 100$\times$ more
+throughput than linearizable replication for low-coordination workloads.
 
-In closing, here is an outline of the paper and of our contributions:
-%
+In closing, here is a summary of our contributions:
+\begin{itemize}
+  \item
     We propose a novel expression-oriented definition of \invariantconfluence{}
     that is both formal and simple (\secref{InvariantConfluence}).
-%
+  \item
     We develop an understanding of why \invariantclosure{} is not necessary for
-\invariantconfluence{} and use this understanding to develop conditions under
-which it is both necessary and sufficient (\secref{InvariantClosure}).
-%
-We exploit these conditions to design a general-purpose interactive decision
-procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
-%
+    \invariantconfluence{} and use this understanding to develop conditions
+    under which it is both necessary and sufficient
+    (\secref{InvariantClosure}).
+  \item
+    We exploit these conditions to design a general-purpose interactive
+    decision procedure for \invariantconfluence{}
+    (\secref{InteractiveDecisionProcedure}).
+  \item
     We again exploit these conditions to design a novel non-trivial sufficient
-condition for \invariantconfluence{}, called \mergereducibility.
-%
+    condition for \invariantconfluence{}, called \mergereducibility{}
+    (\secref{MergeReduction}).
+  \item
     We present segmented \invariantconfluence{}: a generalization of
     \invariantconfluence{} that uses a small amount of coordination to maintain
-invariants for replicated objects that are otherwise not \invariantconfluent{}
-(\secref{SegmentedInvariantConfluence}).
-%
-We evaluate our methods using a prototype implementation called Lucy
-(\secref{Evaluation}).
+    invariants for replicated objects that are otherwise not
+    \invariantconfluent{} (\secref{SegmentedInvariantConfluence}).
+
+  % \item
+  %   We evaluate our methods using a prototype implementation called Lucy
+  %   (\secref{Evaluation}).
+\end{itemize}

### 2020-01-13 doc/the_vldb_journal_paper/sections/invariant_confluence.tex
Minor invariant confluence section changes.
- Clarified that the merge operator doesn't need any special properties.
- Fixed the formatting. For some weird reason, equations in \[\] are not
  being centered. I have no idea why. I added \hfills to fix it though.
  A little jank, but oh well.

--- 2018-09-19 doc/paper/sections/invariant_confluence.tex
@@
 and a binary merge operator $\join: S \times S \to S$ that merges two states
-into one. A \defword{transaction} $t: S \to S$ is a function that maps one
-state to another. An \defword{invariant} $I$ is a subset of $S$. Notationally,
-we write $I(s)$ to denote that $s$ satisfies the invariant (i.e. $s \in I$)
-and $\lnot I(s)$ to denote that $s$ does not satisfy the invariant (i.e. $s
-\notin I$).
+into one. A \defword{transaction} $t: S \to S$ is a
+function that maps one state to another. An \defword{invariant} $I$ is a subset
+of $S$. Notationally, we write $I(s)$ to denote that $s$ satisfies the
+invariant (i.e. $s \in I$) and $\lnot I(s)$ to denote that $s$ does not satisfy
+the invariant (i.e. $s \notin I$).
 
@@
 
-Note that by modelling a transaction $t$ as a function $S \to S$, we focus
-exclusively on the effects that a transaction has on the object (i.e.\
-``writes'' to the object). Transactions are also free to read the value of the
-object, but these reads are not captured by our model because, as we'll see,
-they do not affect \invariantconfluence{}. For example, we could model any
-read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.
+Note that we do not assume any properties of $\join$, like associativity or
+commutativity. Also note that by modelling a transaction $t$ as a function $S
+\to S$, we focus exclusively on the effects that a transaction has on the
+object (i.e.\ ``writes'' to the object). Transactions are also free to read the
+value of the object, but these reads are not captured by our model because, as
+we'll see, they do not affect \invariantconfluence{}. For example, we could
+model any read-only transaction as a function $t$ where $t(s) = s$ for every $s
+\in S$.
 
@@
 system execution as a simple expression generated by the grammar
+%
 \[
+  \hfill
   e ::= s \mid t(e) \mid e_1 \join e_2
+  \hfill
 \]
+%
 where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
@@
 \[
+  \hfill
   \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
+  \hfill
 \]

### 2020-01-13 doc/the_vldb_journal_paper/sections/invariant_confluence.tex
Added definition equivalence proof.
Added proof that our definition of invariant confluence is the same as
Bailis' initial definition. Also started a list of changes that we'll
ultimately have to report to the reviewers.

@@
 \]
+
+\subsection{Equivalence to Existing Definition}
+Our definition of invariant confluence is different than the original
+definition given in~\cite{bailis2014coordination}, but the difference is merely
+cosmetic. We now prove that the two definitions are equivalent.
+
+We say an expression $e$ \defword{recursively satisfies $I$}, denoted
+$\Irec{e}$, if $e$ and all of $e$'s children satisfy $I$. That is,
+\begin{itemize}
+  \item
+    $\Irec{s}$ if $I(s)$,
+
+  \item
+    $\Irec{t(e)}$ if $\Irec{e}$ and $I(t(e))$, and
+
+  \item
+    $\Irec{e_1 \join e_2}$ if $\Irec{e_1}$, $\Irec{e_2}$, and $I(e_1 \join
+    e_2)$.
+\end{itemize}
+
+In~\cite{bailis2014coordination}, Bailis et al.\ define \sTIconfluence{} to
+mean that (a) the start state $s_0$ satisfies the invariant and (b) all
+\sTIreachable{} expressions recursively satisfying $I$ are closed under join.
+That is, $O$ is \sTIconfluent{} if $I(s_0)$ and
+\begin{gather*}
+  \hspace{0.75in} \forall e_1, e_2 \in \setst{e}{\sTIreachablepredicate{e}}.\; \\
+  \hspace{0.75in} \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)
+\end{gather*}
+
+\begin{theorem}\thmlabel{TwoIconfluenceDefs}
+  Consider a state based object $O = (S, \join)$, a start state $s_0$, a set of
+  transactions $T$, and an invariant $I$. The following two are equivalent:
+  \begin{enumerate}
+    \item
+      $\setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I$
+
+    \item
+      $I(s_0)$ and
+      $\forall e_1, e_2 \in \setst{e}{\sTIreachablepredicate{e}}.\;
+         \Irec{e_1} \land \Irec{e_2} \implies I(e_1 \join e_2)$
+  \end{enumerate}
+\end{theorem}
+\begin{proof}
+  First, we show that (1) implies (2). Trivially,
+  $\sTIreachablepredicate{s_0}$, so by (1), $I(s_0)$. Let $e_1$ and $e_2$ be
+  arbitrary \sTIreachable{} expressions. Then $e_1 \join e_2$ is also
+  reachable, so by (1), $I(e_1 \join e_2)$.
+
+  Next, we show that (2) implies (1). We prove by structural induction that for
+  all $e$, $\sTIreachablepredicate{e} \implies \Irec{e}$. From this, (1) is
+  immediate.
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      $I(s_0)$ by (2), so $\Irec{s_0}$
+
+    \item \textbf{Case $t(e)$.}
+      Let $t(e)$ be \sTIreachable{}. Then, $\sTIreachablepredicate{e}$ and
+      $I(t(e))$. By the inductive hypothesis, $\Irec{e}$, so by the definition
+      of $\Irec{\cdot}$, $\Irec{t(e)}$.
+
+    \item \textbf{Case $e_1 \join e_2$.}
+      Let $e_1 \join e_2$ be \sTIreachable{}. Then,
+      $\sTIreachablepredicate{e_1}$ and $\sTIreachablepredicate{e_2}$. By the
+      inductive hypothesis, $\Irec{e_1}$ and $\Irec{e_2}$. By $(2)$, $I(e_1
+      \join e_2)$. Thus, by the definition of $\Irec{\cdot}$, $\Irec{e_1 \join
+      e_2}$.
+  \end{itemize}
+\end{proof}

### 2020-01-14 doc/the_vldb_journal_paper/sections/invariant_closure.tex
Added proofs to invariant closure section.

--- 2018-05-01 doc/paper/sections/invariant_closure.tex
@@
 
+\begin{proof}
 \thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
-sufficient for \invariantconfluence{}. Intuitively, recall that our system model ensures
-that transaction execution preserves the invariant, so if merging states also
-preserves the invariant and if our start state satisfies the invariant, then
-inductively it is impossible for us to reach a state that doesn't satisfy the
-invariant.
+sufficient for \invariantconfluence{}. To prove this informally, recall that
+our system model ensures that transaction execution preserves the invariant, so
+if merging states also preserves the invariant and if our start state satisfies
+the invariant, then inductively it is impossible for us to reach a state that
+doesn't satisfy the invariant.
 
-This is good news because checking if an object is \invariantclosed{} is more
-straightforward than checking if it is \invariantconfluent{}. Existing systems
-typically use an SMT solver like Z3 to check if an object is
-\invariantclosed{}~\cite{de2008z3, balegas2015putting, gotsman2016cause}. If it
-is, then by \thmref{IclosureImpliesIconfluence}, it is \invariantconfluent{}.
-Unfortunately, \invariantclosure{} is \emph{not} necessary for
-\invariantconfluence{}, so if an object is \emph{not} \invariantclosed{}, these
-systems cannot conclude that the object is \emph{not} \invariantconfluent{}.
-The reason why \invariantclosure{} is not necessary for \invariantconfluence{}
-is best explained through an example.
+More formally, consider an arbitrary state $s$ that is reachable via expression
+$e$. We prove by structural induction that for every expression $e$, if $e$ is
+\sTIreachable{} then $I(e)$.
+\begin{itemize}
+  \item \textbf{Case $s_0$.}
+    $s_0$ is reachable and satisfies the invariant by assumption.
+
+  \item \textbf{Case $t(e)$.}
+    If $t(e)$ is reachable, then it satisfies the invariant by definition.
+
+  \item \textbf{Case $e_1 \join e_2$.}
+    Let $e_1 \join e_2$ be \sTIreachable{}. Then, $\sTIreachablepredicate{e_1}$
+    and $\sTIreachablepredicate{e_2}$. By the inductive hypothesis, $I(e_1)$
+    and $I(e_2)$. By \invariantclosure{}, $I(e_1 \join e_2)$.
+\end{itemize}
+\end{proof}
+
+\thmref{IclosureImpliesIconfluence} is good news because checking if an object
+is \invariantclosed{} is more straightforward than checking if it is
+\invariantconfluent{}. Existing systems typically use an SMT solver like Z3 to
+check if an object is \invariantclosed{}~\cite{de2008z3, balegas2015putting,
+gotsman2016cause}. If it is, then by \thmref{IclosureImpliesIconfluence}, it is
+\invariantconfluent{}.  Unfortunately, \invariantclosure{} is \emph{not}
+necessary for \invariantconfluence{}, so if an object is \emph{not}
+\invariantclosed{}, these systems cannot conclude that the object is \emph{not}
+\invariantconfluent{}. The reason why \invariantclosure{} is not necessary for
+\invariantconfluence{} is best explained through an example.
 
@@
   Let $O = (\ints \times \ints, \join)$ consist of pairs $(x, y)$ of integers
-  where $(x_1, y_1) \join (x_2, y_2) = (\max(x_1, x_2), \max(y_1, y_2))$. Our
-  start state $s_0 \in \ints \times \ints$ is the point $(0, 0)$. Our set $T$
-  of transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1, y)$
-  which increments $x$ and $t_{y-1}((x, y)) = (x, y - 1)$ which decrements $y$.
-  Our invariant $I = \setst{(x, y) \in \ints \times \ints}{xy \leq 0}$ consists
-  of all points $(x, y)$ where the product of $x$ and $y$ is non-positive.
+  where
+  \[
+    \hfill (x_1, y_1) \join (x_2, y_2) = (\max(x_1, x_2), \max(y_1, y_2)) \hfill
+  \]
+  Our start state $s_0 \in \ints \times \ints$ is the point $(0, 0)$. Our set
+  $T$ of transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1,
+  y)$ which increments $x$ and $t_{y-1}((x, y)) = (x, y - 1)$ which decrements
+  $y$.  Our invariant $I = \setst{(x, y) \in \ints \times \ints}{xy \leq 0}$
+  consists of all points $(x, y)$ where the product of $x$ and $y$ is
+  non-positive.
 
   The invariant and the set of reachable states are illustrated in \figref{Z2}
-  in which we draw each state $(x, y)$ as a point in space. The invariant
+  in which we draw each state $(x, y)$ as a point in $\ints^2$. The invariant
   consists of the second and fourth quadrant, while the reachable states consist only
@@
 \end{theorem}
+
 \begin{proof}
@@
   And because $O$ is \sTIconfluent{}, all reachable points, including $s_1
-  \join s_2$, satisfy the invariant.
+  \join s_2$, satisfy the invariant. Moreover $s_0$ is reachable, so $I(s_0)$
+  because $O$ is \sTIconfluent{}.
 \end{proof}

### 2020-01-14 doc/the_vldb_journal_paper/sections/invariant_confluence.tex
Added proofs to invariant closure section.

@@
 \subsection{Equivalence to Existing Definition}
-Our definition of invariant confluence is different than the original
+Our definition of \invariantconfluence{} is different than the original
 definition given in~\cite{bailis2014coordination}, but the difference is merely

### 2020-01-14 doc/the_vldb_journal_paper/sections/interactive_decision_procedure.tex
Added tech report sections of paper.
Expanded correctness proof of decision proc.

--- 2018-09-20 doc/paper/sections/interactive_decision_procedure.tex
@@
       \If {closed}
-        \Return true
+        \State \Return true
       \EndIf
@@
       \If{$s_1, s_2 \in R$}
-        \Return false
+        \State \Return false
       \EndIf
@@
 \IsInvConfluent{} assumes access to an \invariantclosure{} decision procedure
-\IsIclosed$(O, I)$ that returns a triple (closed, $s_1$, $s_2$). closed is a
-boolean indicating whether $O$ is \Iclosed{}. If closed is true, then $s_1$ and
-$s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a counterexample
-witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
-but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As
-we mentioned earlier, we can (and do) implement the \invariantclosure{}
-decision procedure using an SMT solver like Z3~\cite{de2008z3}.
+\IsIclosed$(O, I)$. The decision procedure \IsIclosed$(O, I)$ returns a triple
+(closed, $s_1$, $s_2$).  closed is a boolean indicating whether $O$ is
+\Iclosed{}. If closed is true, then $s_1$ and $s_2$ are null. If closed is
+false, then $s_1$ and $s_2$ are a counterexample witnessing the fact that $O$
+is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$
+(e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As we mentioned earlier, we can
+(and do) implement the \invariantclosure{} decision procedure using an SMT
+solver like Z3~\cite{de2008z3}.
 
@@
 $s_0$, $T$, and $I$---takes as arguments a set $R$ of \sTIreachable{} states
-and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, \Helper$(O,
-s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming $R$ and
-$NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes, $NR$ is
-iteratively increased, which removes unreachable states from $I$ until $I$ is a
-subset of $\setst{s \in S}{\sTIreachablepredicate{s}}$.
+and a set $NR$ of \sTIunreachable{} states. Like \IsInvConfluent, the function
+\Helper$(O, s_0, T, I, R, NR)$ returns whether $O$ is \sTIconfluent{} (assuming
+$R$ and $NR$ are correct).  As \algoref{InteractiveDecisionProcedure} executes,
+$NR$ is iteratively increased, which removes unreachable states from $I$ until
+$I$ is a subset of $\setst{s \in S}{\sTIreachablepredicate{s}}$.
 
@@
 \[
+  \hfill
   \setst{s \in S}{\stireachablepredicate{s_0}{T}{I-NR}{s}}
@@
     \subseteq I
+  \hfill
 \]
-Because $NR$ only contains \sTIunreachable{} states, then the set of
-\sTIreachable{} states is equal to set of \stireachable{s_0}{T}{I-NR}{} states
-which, as we just showed, is a subset of $I$.  Thus, $O$ is \sTIconfluent{}, so
-\Helper{} returns true.
+
+In this case, $O$ is \sTIconfluent{} because the set of \sTIreachable{} states
+is a subset of \stireachable{s_0}{T}{I-NR}{} states which is in turn a subset
+of $I$. We prove that now.
+\begin{proof}
+  We prove, by structural induction, that every \sTIreachable{} expression $e$
+  is also \stireachable{s_0}{T}{I-NR}{}.
+
+  \begin{itemize}
+    \item \textbf{Case $s_0$.}
+      $s_0$ is \stireachable{s_0}{T}{I-NR}{} by definition.
+
+    \item \textbf{Case $t(e)$.}
+      $e$ is \sTIreachable{}, so by the inductive hypothesis, it is also
+      \stireachable{s_0}{T}{I-NR}{}. Moreover, $t(e) \in I$ and $t(e) \notin
+      NR$ (by assumption), so $t(e) \in I - NR$. Thus, $t(e)$ is
+      \stireachable{s_0}{T}{I-NR}{}.
+
+    \item \textbf{Case $e_1 \join e_2$.}
+      By the inductive hypothesis, $e_1$ and $e_2$ are both
+      \stireachable{s_0}{T}{I-NR}{}, so $e_1 \join e_2$ is also
+      \stireachable{s_0}{T}{I-NR}{}.
+  \end{itemize}
+\end{proof}
 
 If \IsIclosed{} determines that $O$ is \emph{not} \iclosed{(I - NR)}, then we
-have a counterexample $s_1, s_2$. We want to determine whether $s_1$ and $s_2$
-are reachable or unreachable. We can do so in two ways.
+have a counterexample $s_1, s_2$. That is, $s_1, s_2 \in I-NR$, but $s_1 \join
+s_2 \notin I-NR$. We want to determine whether $s_1$ and $s_2$ are
+\sTIreachable{} or unreachable. We can do so in two ways.
 %
@@
 continue. If both $s_1$ and $s_2$ are \sTIreachable{}, then so is $s_1 \join
-s_2$, but $\lnot I(s_1 \join s_2)$. Thus, $O$ is not \sTIconfluent{}, so
-\Helper{} returns false. Otherwise, one of $s_1$ and $s_2$ is
-\sTIunreachable{}, so we recurse.
+s_2$, but $\lnot I(s_1 \join s_2)$ (because $s_1 \join s_2 \notin I - NR$).
+Thus, $O$ is not \sTIconfluent{}, so \Helper{} returns false. Otherwise, one of
+$s_1$ and $s_2$ is \sTIunreachable{}, so we recurse.
 
@@
 %
-\begin{techreport}
-
 For example, let $O_p = (S, \join)$ where $S$ is the set of all programs and
@@
 determining if $p$ terminates.
-\end{techreport}
 %
@@
 
-\begin{techreport}
 \subsection{Tolerating User Error}
@@
 that violates $I$.
-\end{techreport}

### 2020-01-14 doc/the_vldb_journal_paper/figures/merging_diagram.tex
Fleshed out merge reducibility section.

\begin{figure*}[ht]
  \centering

  \tikzstyle{s0color}=[fill=flatred]
  \tikzstyle{s1color}=[fill=flatgreen]
  \tikzstyle{s2color}=[fill=flatdenim]
  \tikzstyle{s3color}=[fill=flatorange]
  \tikzstyle{s4color}=[fill=flatyellow]
  \tikzstyle{s5color}=[fill=flatcyan]
  \tikzstyle{s6color}=[fill=flatpurple]
  \tikzstyle{s7color}=[fill=flatblue]
  \tikzstyle{astedge}=[thick]
  \tikzstyle{phantomstate}=[%
    shape=circle, inner sep=2pt, draw=white, line width=1pt, fill=white]
  \tikzstyle{state}=[%
    shape=circle, inner sep=2pt, draw=black, line width=1pt, text opacity=1,
    fill opacity=0.6]
  \newcommand{\internaltext}[1]{$\boldsymbol #1$}

  \begin{tikzpicture}[xscale=1.1]
    \begin{scope}[]
                         \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (s7) at (0, 0) {\internaltext{\join}};
      \draw (s7)++(210:1) node[state, s3color, label={[label distance=-0.1cm] 90:$s_3$}] (s3)           {\internaltext{\join}};
      \draw (s7)++(-30:1) node[state, s6color, label={[label distance=-0.1cm] 90:$s_6$}] (s6)           {\internaltext{\join}};
      \draw (s3)++(240:1) node[state, s1color, label={[label distance=-0.2cm] 120:$s_1$}](s1)           {\internaltext{t}};
      \draw (s3)++(-60:1) node[state, s2color, label={[label distance=-0.2cm] 60:$s_2$}] (s2)           {\internaltext{u}};
      \draw (s6)++(240:1) node[state, s4color, label={[label distance=-0.2cm] 240:$s_4$}](s4)           {\internaltext{v}};
      \draw (s6)++(-60:1) node[state, s5color, label={[label distance=-0.2cm] 60:$s_5$}] (s5)           {\internaltext{w}};
      \draw (s1)++(-90:1) node[state, s0color]                                           (n1)           {\internaltext{s_0}};
      \draw (s2)++(-90:1) node[state, s0color]                                           (n2)           {\internaltext{s_0}};
      \draw (s4)++(-90:1) node[state, s0color]                                           (n4)           {\internaltext{s_0}};
      \draw (s5)++(-90:1) node[state, s0color]                                           (n5)           {\internaltext{s_0}};
      \draw[astedge] (s7) to (s3) to (s1) to (n1);
      \draw[astedge] (s7) to (s3) to (s2) to (n2);
      \draw[astedge] (s7) to (s6) to (s4) to (n4);
      \draw[astedge] (s7) to (s6) to (s5) to (n5);
      \coordinate (A) at (s5);
    \end{scope}
    \begin{scope}[xshift=125]
      \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (s7) at (0, 0) {\internaltext{\join}};
      \draw (s7)++(210:1) node[state, s3color, label={[label distance=-0.1cm] 90:$s_3$}] (s3)           {\internaltext{p}};
      \draw (s7)++(-30:1) node[state, s6color, label={[label distance=-0.1cm] 90:$s_6$}] (s6)           {\internaltext{\join}};
      \draw (s6)++(240:1) node[state, s4color, label={[label distance=-0.2cm] 240:$s_4$}](s4)           {\internaltext{v}};
      \draw (s6)++(-60:1) node[state, s5color, label={[label distance=-0.2cm] 60:$s_5$}] (s5)           {\internaltext{w}};
      \draw (s3)++(-90:1) node[state, s0color]                                           (n3)           {\internaltext{s_0}};
      \draw (s4)++(-90:1) node[state, s0color]                                           (n4)           {\internaltext{s_0}};
      \draw (s5)++(-90:1) node[state, s0color]                                           (n5)           {\internaltext{s_0}};
      \draw[astedge] (s7) to (s3) to (n3);
      \draw[astedge] (s7) to (s6) to (s4) to (n4);
      \draw[astedge] (s7) to (s6) to (s5) to (n5);
    \end{scope}
    \begin{scope}[xshift=250]
      \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (s7) at (0, 0) {\internaltext{\join}};
      \draw (s7)++(210:1) node[state, s3color, label={[label distance=-0.1cm] 90:$s_3$}] (s3)           {\internaltext{p}};
      \draw (s7)++(-30:1) node[state, s6color, label={[label distance=-0.1cm] 90:$s_6$}] (s6)           {\internaltext{q}};
      \draw (s3)++(-90:1) node[state, s0color]                                           (n3)           {\internaltext{s_0}};
      \draw (s6)++(-90:1) node[state, s0color]                                           (n6)           {\internaltext{s_0}};
      \draw[astedge] (s7) to (s3) to (n3);
      \draw[astedge] (s7) to (s6) to (n6);
    \end{scope}
    \begin{scope}[xshift=350]
                         \node[state, s7color, label={[label distance=-0.1cm] 90:$s_7$}] (s7) at (0, 0) {\internaltext{r}};
      \draw (s7)++(-90:1) node[state, s0color]                                           (n7)           {\internaltext{s_0}};
      \draw[astedge] (s7) to (n7);
    \end{scope}

    \draw[-latex, line width=3pt] (A) ++ (0.55, 0) to ++(1, 0);
    \draw[-latex, line width=3pt] (A) ++ (5, 0) to ++(1, 0);
    \draw[-latex, line width=3pt] (A) ++ (9, 0) to ++(1, 0);
  \end{tikzpicture}

  \caption{%
    An illustration of the proof of \thmref{ReducibilityImpliesIconfluence}. We
    begin with a reachable expression and convert it into a merge-free
    reachable expression by repeatedly replacing the merge of two merge-free
    reachable subexpressions with an equivalent merge-free reachable
    expression. In this example, we first replace $t(s_0) \join u(s_0)$ with
    $p(s_0)$. We then replace $v(s_0) \join w(s_0)$ with $q(s_0)$. Finally, we
    replace $p(s_0) \join q(s_0)$ with $r(s_0)$.
  }\figlabel{MergingDiagram}
\end{figure*}

### 2020-01-14 doc/the_vldb_journal_paper/sections/merge_reduction.tex
Fleshed out merge reducibility section.

@@
 
+\begin{example}\examplelabel{MergeReducible}
+  Consider the distributed object $O = (\ints, \max)$ consisting of integers
+  merged by the max function. Our start state state $s_0 = 0$ and our invariant
+  $I = \setst{x \in \ints}{x \geq 0}$. Our set $T$ of transactions is the
+  infinite set $T = \setst{t_y}{y \in \ints}$ where $t_y(x) = x + y$ is a
+  transaction that adds $y$ to the state. For example, $t_2$ is a transaction
+  that adds $2$ to the state, and $t_{-3}$ is a transaction which subtracts $3$
+  from the state.
+  %
+  $O$ is \sTImergereducible{}. Consider two merge-free \sTIreachable{}
+  expressions $e_1$ and $e_2$ that evaluate to states $x_1$ and $x_2$. Without
+  loss of generality, assume $x_1 \geq x_2$. Then, we can replace $e_1 \join
+  e_2$ (which evaluates to $x_1$) with $e_1$. We can also replace it with
+  $t_{x_1}(0)$.
+\end{example}
+
+\begin{example}\examplelabel{NotMergeReducible}
+  Consider the distributed object $O = (\set{X \subseteq \ints}, \join)$ in
+  which each state is a set of integers and where $X_1 \join X_2 = \set{y}$
+  where $y = \sum_{x \in X_1} x + \sum_{x \in X_2} x$. Our start state state
+  $s_0 = \emptyset{}$ and our invariant $I = \setst{X}{\forall x \in X.\ x \
+  \text{is even}}$. Our set $T$ of transactions is the set $T = \set{t_0, t_2,
+  t_4}$ where $t_i(X) = X \cup \set{i}$ is a transaction that adds $i$ to the
+  state. For example, $t_2(\set{0}) = \set{0, 2}$.
+  %
+  $O$ is \emph{not} \sTImergereducible{}. Consider the two merge-free
+  \sTIreachable{} expressions $e_1 = t_2(\emptyset)$ and $e_2 =
+  t_4(\emptyset)$. $e_1 \join e_2$ evaluates to $\set{6}$, but there does not
+  exist a merge-free expression that evaluates to $\set{6}$.
+\end{example}
+
+{\input{figures/merging_diagram.tex}}
+
 \begin{theorem}\thmlabel{ReducibilityImpliesIconfluence}
@@
 
-The proof of \thmref{ReducibilityImpliesIconfluence} is a straightforward
-induction. We begin with an \sTIreachable{} expression $e$ and repeatedly
-replace any subexpression that merges two merge-free subexpressions with an
-equivalent merge-free reachable subexpression (which we can do because $O$ is
-merge-reducible). We repeat this process until $e$ has been completely replaced
-with an equivalent merge-free reachable expression $e'$. Because $I(s_0)$ and
-because our system model only executes transactions that preserve the
-invariant, $e'$ (and hence $e$) is guaranteed to satisfy the invariant. Thus,
-all reachable states satisfy the invariant, so $O$ is \invariantconfluent{}.
+\begin{proof}
+  Intuitively, the proof of \thmref{ReducibilityImpliesIconfluence} is a
+  straightforward induction. We begin with an \sTIreachable{} expression $e$
+  and repeatedly replace any subexpression that merges two merge-free
+  subexpressions with an equivalent merge-free reachable subexpression (which
+  we can do because $O$ is merge-reducible). We repeat this process until $e$
+  has been completely replaced with an equivalent merge-free reachable
+  expression $e'$. Because $I(s_0)$ and because our system model only executes
+  transactions that preserve the invariant, $e'$ (and hence $e$) is guaranteed
+  to satisfy the invariant. Thus, all reachable states satisfy the invariant,
+  so $O$ is \invariantconfluent{}. An illustration of this idea is given in
+  \figref{MergingDiagram}.
+
+  More formally, we prove by structural induction on $e$, that for all
+  \sTIreachable{} expressions $e$, there exists a merge-free \sTIreachable{}
+  expression $e'$ such that $eval(e) = eval(e')$.
+  \begin{itemize}
+    \item \textbf{Case 1: $e = s_0$.}
+      Trivially, $e' = s_0$.
+
+    \item \textbf{Case 2: $e = t(e_1)$.}
+      $e_1$ is \sTIreachable{}, so by the inductive hypothesis, there exists a
+      merge-free \sTIreachable{} expression $e_1'$ such that $eval(e_1) =
+      eval(e_1')$. $t(e_1)$ is \sTIreachable{}, so $I(t(e_1))$. Because
+      $eval(e_1) = eval(e_1')$, we know also that $I(t(e_1'))$. Thus, $t(e_1')$
+      is \sTIreachable{} (and join free), so we can let $e' = t(e_1')$.
+
+    \item \textbf{Case 3: $e = e_1 \join e_2$.}
+      $e_1$ and $e_2$ are \sTIreachable{}, so by the inductive hypothesis,
+      there exists equivalent merge-free \sTIreachable{} expressions $e_1'$ and
+      $e_2'$. $O$ is \sTImergereducible{}, so there exists an equivalent
+      merge-free \sTIreachable{} expression $e'$.
+  \end{itemize}
+
+  Consider an arbitrary \sTIreachable{} expression $e$ and it's equivalent
+  merge-free \sTIreachable{} counterpart $e'$. $e'$ is either $s_0$ or
+  $t(e'')$.  In either case, it satisfies the invariant, so $O$ is
+  \sTIconfluent{}.
+\end{proof}
+
+Note that while merge-reducibility is a sufficient condition for
+\invariantconfluence{}, it is not necessary. The object in
+\exampleref{NotMergeReducible} is \invariantconfluent{} but not
+merge-reducible.
 
@@
     \item
-      $O$ is a join-semilattice.
+      $O$ is a join-semilattice. That is, $\join$ is associative ($(x \join y)
+      \join z = x \join (y \join z)$), commutative ($x \join y = y \join x$),
+      and idempotent ($x \join x = x$).
 
@@
 
-To show that $O$ is \sTImergereducible, we consider two merge-free reachable
-expressions
+\newcommand{\bart}[1]{\overline{t_{#1}}}
+\newcommand{\baru}[1]{\overline{u_{#1}}}
+\newcommand{\bartu}[2]{\bart{#1}(\baru{#2}(s_0))}
+\begin{proof}
+  Let
   \[
-  e_1 = t_1(\ldots(t_n(s_0))\ldots)
+    \hfill
+    e_1 = t_n(t_{n-1}(\ldots (t_1(s_0))\ldots))
+    \hfill
+  \]
+  and
+  \[
+    \hfill
+    e_2 = u_m(u_{m-1}(\ldots (u_1(s_0))\ldots))
+    \hfill
+  \]
+  be two arbitrary merge-free \sTIreachable{} expressions.
+  For ease of notation, let
+  \[
+    \hfill
+    \bart{i} = t_i(\ldots(t_1(s_0))\ldots)
     \quad\text{and}\quad
-  e_2 = u_1(\ldots(u_m(s_0))\ldots)
+    \baru{j} = u_j(\ldots(u_1(s_0))\ldots)
+    \hfill
   \]
-Then, we inductively use criterion (3) with criteria (1) and (2) to show that
-\[
-  e_3 = t_1(\ldots(
-  t_n(u_1(\ldots(u_m(s_0))\ldots))
-  )\ldots)
-\]
-is reachable and evaluates to $e_1 \join e_2$.
+  We want to show that there exists some merge-free \sTIreachable{} expression
+  that is equivalent to $e_1 \join e_2$.
+  %
+  To do so, we prove by strong induction on $k \in \nats$ that if $k = i + j$
+  where $0 \leq i \leq n$ and $0 \leq j \leq m$, $\bartu{i}{j}$ is
+  \sTIreachable{} and $eval(\bartu{i}{j}) = eval(\bart{i}(s_0) \join
+  \baru{j}(s_0))$.
+  \begin{itemize}
+    \item \textbf{Case $k = 0$.}
+      $i = j = 0$, so $\bartu{0}{0} = s_0$ which is trivially \sTIreachable{}
+      and equivalent to $\bart{0}(s_0) \join \baru{0}(s_0) = s_0 \join s_0$
+      which evaluates to $s_0$ (because $\join$ is idempotent).
+
+    \item \textbf{Case $k = 1$.}
+      Without loss of generality, assume $i = 1$ and $j = 0$. Then,
+      $
+        \bartu{1}{0} = t_1(s_0)
+      $ which is \sTIreachable{} because it is a subexpression of $\bart{n}$
+      which is \sTIreachable{}. Moreover, it is equivalent to $\bart{1}(s_0)
+      \join \baru{0}(s_0) = t_1(s_0) \join s_0$ which evaluates to $s_{t_1}
+      \join s_0 \join s_0 = s_{t_1} \join s_0 = t_1(s_0)$ for some $s_{t_1} \in
+      S$.
+
+    \item \textbf{Case $k \geq 2$.}
+      If $i = 0$, then $j = k$ and $\baru{k}(s_0)$ is \sTIreachable{} because
+      it is a subexpression of $\baru{m}(s_0)$. Also, it is equivalent to
+      $\bart{0}(s_0) \join \baru{k}(s_0)$ which evaluates to $\baru{k}(s_0)$.
+      %
+      The symmetric result holds if $j = 0$.
+
+      Otherwise, $i, j > 1$. Let
+      \begin{align*}
+        e_{i-1,j-1} &= \bartu{i-1}{j-1} \\
+        e_{i,j-1} &= \bartu{i}{j-1} \\
+        e_{i-1,j} &= \bartu{i-1}{j}
+      \end{align*}
+      By the inductive hypothesis, $e_{i-1,j-1}$, $e_{i,j-1}$, and $e_{i-1,j}$
+      are all \sTIreachable{}. By condition 3 (with $s = eval(e_{i-1,j-1})$,
+      $t_1$ = $t_i$, and $t_2 = u_j$), $I(e_{i,j-1} \join e_{i-1,j})$.
+      $e_{i,j-1} \join e_{i-1,j} = t_i(e_{i-1,j}) = u_j(e_{i,j-1}) =
+      \bartu{i}{j}$, so $I(\bartu{i}{j})$. Therefore, $\bartu{i}{j}$ is
+      \sTIreachable{}.
+  \end{itemize}
+\end{proof}
+
+{\input{figures/diamond_proof_diagram.tex}}
+
+An illustration of this proof is given in \figref{DiamondProofDiagram}. We
+arrange the expressions $e_1$ and $e_2$ as the left and top edges of a square
+grid. Each point in the grid represents a state (with $s_0$ in the top left
+corner), and each edge represents the application of a transaction. A state is
+circled if we know it satisfies the invariant.
+%
+Condition (1) and (2) tell us that the order in which we apply transactions is
+immaterial. Thus, if we begin at the top left of the square and walk to any
+other point in the square, applying transactions along the way, it does not
+matter which path we take. They are all equivalent. Condition (4) tells us that
+the top-left corner satisfies the invariant. We induct to repeatedly apply
+condition (3) to ``fill in'' the square, one block at a time. In iteration $k$,
+we discover that all points with a Manhattan distance of $k$ from the top left
+corner satisfy the invariant. Ultimately, we conclude that the bottom left
+corner (i.e., $e_1 \join e_2$) satisfies the invariant and is equivalent to
+$\bartu{n}{m}$.
 
@@
 
-% {\input{figures/merge_reducible}}
+{\input{figures/merge_reducible}}
 
@@
 \invariantconfluent{} despite it not being \invariantclosed{}.
+
+The relationship between \invariantconfluence{}, \invariantclosure{},
+merge-reducibility, and criteria (1)-(4) is illustrated in
+\figref{IclosureVsReducible}.

### 2020-01-14 doc/the_vldb_journal_paper/sections/segmented_invariant_confluence.tex
Added tech report sections of paper.
Revised Segmented Invariant Confluence section.

--- 2018-09-20 doc/paper/sections/segmented_invariant_confluence.tex
@@
 non-\invariantconfluent{} object with segmented \invariantconfluence{} can
-achieve between 10x and 100x more throughput than linearizable replication for
-certain workloads.
+achieve between 10$\times$ and 100$\times$ more throughput than linearizable
+replication for certain workloads.
 
@@
 \begin{example}\examplelabel{SegmentedZ2}
-  Consider again the object $O = (\ints \times \ints, \join)$, transactions $T
-  = \set{t_{x+1}, t_{y-1}}$, and invariant $I = \setst{(x, y)}{xy \leq 0}$ from
-  \exampleref{Z2}, but now let the start state $s_0$ be $(-42, 42)$. $O$ is
-  \emph{not} \sTIconfluent{} because the points $(0, 42)$ and $(42, 0)$ are
-  reachable, and merging these points yields $(42, 42)$ which violates the
-  invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1, T_1)$, $(I_2,
-  T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
+  Consider again the object $O = (\ints \times \ints, \join)$, the set of
+  transactions $T = \set{t_{x+1}, t_{y-1}}$, and the invariant $I = \setst{(x,
+  y)}{xy \leq 0}$ from \exampleref{Z2}, but now let the start state $s_0$ be
+  $(-42, 42)$. $O$ is \emph{not} \sTIconfluent{} because the points $(0, 42)$
+  and $(42, 0)$ are reachable, and merging these points yields $(42, 42)$ which
+  violates the invariant. However, $O$ is \sTISconfluent{} for $\Sigma = (I_1,
+  T_1)$, $(I_2, T_2)$, $(I_3, T_3)$, $(I_4, T_4)$ where
   \begin{align*}
-    I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
-    I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
-    I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
-    I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
+    \qquad I_1 &= \setst{(x, y)}{x < 0, y > 0}       &T_1 &= \set{t_{x+1}, t_{y-1}} \\
+    \qquad I_2 &= \setst{(x, y)}{x \geq 0, y \leq 0} &T_2 &= \set{t_{x+1}, t_{y-1}} \\
+    \qquad I_3 &= \setst{(x, y)}{x = 0}              &T_3 &= \set{t_{y-1}} \\
+    \qquad I_4 &= \setst{(x, y)}{y = 0}              &T_4 &= \set{t_{x+1}}
   \end{align*}
@@
 time, a server designates one of the segments in $\Sigma$ as its
-\defword{active segment}.  Initially, every server chooses the first segment
-$(I_i, T_i) \in \Sigma$ such that $I_i(s_0)$ to be its active segment. We'll
-see momentarily the significance of the active segment.
+\defword{active segment}.  Initially, every server selects its active segment
+to be the first segment $(I_i, T_i) \in \Sigma$ such that $I_i(s_0)$. We'll see
+momentarily the significance of the active segment.
 
@@
 Transaction execution in the new system model, on the other hand, is a bit more
-involved. Consider a server $s_i$ with active segment $(I_i, T_i)$. A client
+involved. Consider a server $p_i$ with active segment $(I_i, T_i)$. A client
 can request that $p_i$ execute a transaction $t \in T$. We consider what
@@
     \Else{}
-      \If{$I_i(t(s_i))$} Commit $t$
-      \ElsIf{$\lnot I(t(s_i))$} Abort $t$
-      \Else{} Execute $t$ with global coordination
+      \If{$I_i(t(s_i))$}
+        \State Commit $t$
+      \ElsIf{$\lnot I(t(s_i))$}
+        \State Abort $t$
+      \Else{}
+        \State Execute $t$ with global coordination
       \EndIf{}
@@
 
+\begin{example}
+  Consider the object $O = (\nats, \max)$ consisting of natural numbers merged
+  by the max function. Let $I = \nats$. Let $T = \set{t_{2\times}, t_{+10}}$
+  where $t_{2\times}(x) = 2x$ and $t_{+10} = x + 10$. The states $3$ and $16$
+  are \TIcoreachable{} because both are \stireachable{s}{T}{I} from $s = 3$
+  ($16 = 2 \times 3 + 10$). The states $3$ and $5$, however, are not
+  \TIcoreachable{}. The only state in which $3$ is reachable is $3$, but from
+  this state, the smallest reachable number larger than $3$ is $6$.
+\end{example}
+
 \begin{theorem}\thmlabel{ExtendedIclosureEquivalentIconfluence}
@@
   \[
+    \hfill
     \text{$O$ is \Iclosed{}}
     \iff
-    \text{$O$ is \sticonfluent{s}{T}{I} for every $s \in I$}
+    \forall s \in I.\ \text{$O$ is \sticonfluent{s}{T}{I}}
+    \hfill
   \]
@@
 
+\begin{proof}
   The proof of the forward direction is exactly the same as the proof of
@@
   For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
-The two points are \TIcoreachable{}, so there exists some state $s_0$ from
-which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
-\sTIreachable{}, so it satisfies the invariant.
+  By assumption, the two points are \TIcoreachable{}, so there exists some
+  state $s_0$ from which they can be reached. $O$ is \sTIconfluent{} and $s_1
+  \join s_2$ is \sTIreachable{}, so it satisfies the invariant.
+\end{proof}
 
@@
 
-\begin{techreport}
-
 Third, segmented \invariantconfluence{} naturally subsumes a distributed
@@
 with state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t
-  \in I_i$ to execute, and $t(s_i)$ violates the active invariant but not the
+\in T_i$ to execute, and $t(s_i)$ violates the active invariant but not the
 global invariant, instead of initiating a round of global coordination, $p_i$
@@
 I$.
-\end{techreport}
+
+Sixth, a naive implementation of our segmented invariant confluence system
+model is \emph{not} fault tolerant. A round of global coordination requires
+that \emph{every} server be non-faulty. If even a single server fails, then
+every round of global coordination will fail to complete. Fortunately, we can
+leverage well-established means of ensuring fault tolerance without difficulty.
+For example, we can replicate every server using a state machine replication
+protocol like MultiPaxos~\cite{lamport1998part, lamport2001paxos} or
+Raft~\cite{ongaro2014search}. Doing so will introduce a constant slowdown for
+each (now replicated) server, but servers remain independent and scalable.

### 2020-01-14 doc/the_vldb_journal_paper/sections/conclusion.tex
Added Peter Bailis to acks.

--- 2020-01-13 doc/vldb_paper/sections/conclusion.tex
@@
 \textbf{Acknowledgments.}
-The authors would like to thank Alan Fekete, Peter Alvaro, Alvin Cheung,
-Alexandra Meliou, Anthony Tan, and Cristina Teodoropol for fruitful discussion
-and feedback.
+The authors would like to thank Alan Fekete, Alexandra Meliou, Alvin Cheung,
+Anthony Tan, Cristina Teodoropol Peter Alvaro, and Peter Bailis, for fruitful
+discussion and feedback.
 %

### 2020-01-14 doc/the_vldb_journal_paper/sections/evaluation.tex
Re-formatted evaluation.

--- 2018-09-20 doc/paper/sections/evaluation.tex
@@
 
-Consider again the PN-Counter $s = (p_1, p_2, p_3), (n_1, n_2, n_3)$ from
-\exampleref{CounreachableExample} replicated on three servers with transactions
-to increment and decrement the PN-Counter. In
+Consider again, from \exampleref{CounreachableExample}, the PN-Counter $s =
+(p_1, p_2, p_3)$, $(n_1, n_2, n_3)$ replicated on three servers with
+transactions to increment and decrement the PN-Counter. In
 \exampleref{CounreachableExample}, we found that concurrent decrements violate
@@
 that will allow us to perform concurrent decrements that respect the escrowed
-value. The first segment $(\setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{p_1, p_2,
-p_3 \geq k \land n_1, n_2, n_3 \leq k}, T)$ allows for concurrent increments
-and decrements so long as every $p_i \geq k$ and every $n_i \leq k$.
-Intuitively, this segment represents the situation in which every server has
-escrowed a value of $k$. They can decrement freely, so long as they don't
-exceed their escrow budget of $k$. The second segment is the one presented in
-\exampleref{CounreachableExample} which prohibits concurrent decrements. We ran
-our decision procedure on this example and it concluded that it was segmented
-\invariantconfluent{} in less than a tenth of a second and without any human
-interaction.
+value. The first segment is $(I_k, T)$ where
+\[
+  I_k = \setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{n_1, n_2, n_3 \leq k \leq p_1, p_2, p_3}
+\]
+This segment allows for concurrent increments and decrements so long as every
+$p_i \geq k$ and every $n_i \leq k$.  Intuitively, this segment represents the
+situation in which every server has escrowed a value of $k$. They can decrement
+freely, so long as they don't exceed their escrow budget of $k$. The second
+segment is the one presented in \exampleref{CounreachableExample} which
+prohibits concurrent decrements. We ran our decision procedure on this example
+and it concluded that it was segmented \invariantconfluent{} in less than a
+tenth of a second and without any human interaction.
 
 \example[TPC-C]\examplelabel{TpccEval}
-TPC-C is a ubiquitous OLTP database benchmark with a workload that models a
-simple warehousing application~\cite{difallah2013oltp}. The TPC-C specification
+TPC-C is a ubiquitous OLTP benchmark with a workload that models a simple
+warehousing application~\cite{difallah2013oltp}. The TPC-C specification
 outlines twelve ``consistency requirements'' (read invariants) that govern the
@@
 %   \item
-    Three of the twelve invariants are \textbf{foreign key constraints}.  As
+    Three of the twelve invariants involve \textbf{foreign key constraints}.  As
     discussed in \exampleref{ForeignKeysEval}, our decision procedures can

### 2020-01-15 doc/the_vldb_journal_paper/sections/operation_based_invariant_confluence.tex
Added operation-based invariant confluence.

\section{Operation Based Invariant Confluence}
In the system model we described, a server $p_i$ periodically sends its state
$s_i$ to some other server $p_j$ for merging. In this ``state-based'' model,
states are sent between replicas but transactions are not. Borrowing a trick
from CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive}, we can define
an alternate, but equivalent, ``operation-based'' system model in which
transactions are sent between replicas but states are not. Though the two
models are equivalent, the operation-based approach is sometimes more natural.
For example, with the operation-based approach, we can replace the PN-counter
from \exampleref{CounreachableExample} with a simple integer.

\subsection{System Model}
A \defword{distributed operation-based object} is a set $O = S$ of states. Note
that we do not have a merge function like we did with state-based objects.
%
An \defword{operation-based transaction} $t: S \to (S \to S)$ is a function
that maps a state $s$ to a \defword{shadow transaction} $t(s): S \to
S$~\cite{li2014automating}.
%
The definition of an invariant is the same in the state-based and
operation-based models.

\begin{example}
  $\nats$ is a distributed operation-based object. $t: \nats \to (\nats \to
  \nats)$ is an operation-based transaction where $t(x)(y) = x + y$. That is,
  given a state $x$, $t(x)$ is the function $f_x$ where $f_x(y) = x + y$.
\end{example}

In our operation-based system model, a distributed object $O$ is replicated
across a set $p_1, \ldots, p_n$ of $n$ servers. Each server $p_i$ manages a
replica $s_i \in O$ of the replicated object. Every server begins with a start
state $s_0 \in S$, a fixed set $T$ of transactions, and an invariant $I$.
Servers repeatedly perform one of two actions.

First, a client can contact a server $p_i$ and request that it execute a
transaction $t \in T$. $p_i$ speculatively executes $t(s_i)$, transitioning
from state $s_i$ to state $t(s_i)(s_i)$. If $t(s_i)(s_i)$ does not satisfy the
invariant, then $p_i$ aborts the transaction and reverts to state $s_i$.
Otherwise, $p_i$ commits the transaction and remains in state $t(s_i)(s_i)$. It
also broadcasts the shadow transaction $t(s_i)$ in an exactly-once manner to
the rest of the servers.

Second, $p_j$ can receive a shadow transaction $t(s_i)$ from some other server
$p_i$. When $p_j$ receives $t(s_i)$, it transitions from its state $s_j$ to
state $t(s_i)(s_j)$. When $p_j$ receives a shadow transaction, it must execute
it, even if $\lnot I(t(s_i)(s_j))$.

Informally, $O$ is \invariantconfluent{} with respect to $s_0$, $T$, and $I$ if
every replica $s_1, \ldots, s_n$ is guaranteed to always satisfy the invariant
$I$ in every possible execution of the system.

% As with state-based objects, there are three ways to think about
% operation-based invariant-confluence: a process-based approach, a graph-based
% approach, and an expression-based approach. All three approaches are
% illustrated in \figref{OpBasedModels} and are described in more detail below.
%
% \input{figs/opbased_iconfluence.fig}
%
% \subsubsection{Process-Based}
% As with the state-based approach, the process-based model is most similar to
% the process model described by Shapiro et al.\ in~\cite{shapiro2011conflict}.
% We are given an operation-based object $O$, a start state $s_0$, a set of
% transactions $T$, an invariant $I$, and a set $\seq{p}{1}{n}$ of $n$
% processors. Each processors begins with state $s_0$. A processor $p_i$ can do
% one of two things.
%
% \begin{itemize}
%   \item
%     $p_i$ can execute a transaction $t$ on its current state $s$ and then
%     transitions from state $s$ to state $t(s)(s)$, given that $I(t(s)(s))$. If
%     $\lnot I(t(s)(s))$, then $p_i$ will not perform $t$. If $p_i$ does execute
%     $t$, then it broadcasts $t(s)$ to all other processors exactly once.
%
%   \item
%     $p_i$ can receive a broadcast shadow transaction $t(s_j)$ from some other
%     processor $p_j$. When $p_i$ receives $t(s_j)$, it transitions from its
%     state $s_i$ to state $t(s_j)(s_i)$. When $p_i$ receives a shadow
%     transaction, it must execute it, even if $\lnot I(t(s_j)(s_i))$.
% \end{itemize}
%
% $O$ is \sTIconfluent{} if every reachable state (including $s_0$) satisfies the
% invariant.
%

\subsection{Expression-Based Formalism}
To define operation-based \invariantconfluence{} formally, we represent a state
produced by an operation-based system execution as a simple expression
generated by the grammar
%
\[
  \hfill
  e ::= s \mid t(e_1)(e_2)
  \hfill
\]
%
where $s$ represents a state in $S$ and $t$ represents a transaction in $T$. As
an example, consider the system execution in \figref{OpSystemExecution} in
which a distributed object is replicated across servers $p_1$, $p_2$, and
$p_3$. Server $p_3$ begins with state $s_0$, receives transaction $t$,
transitions to state $s_1$ by executing shadow transaction $t(s_0)$,
transitions to state $s_3$ by executing shadow transaction $u(s_0)$, and then
transitions to state $s_7$ by executing shadow transaction $v(s_2)$.  In
\figref{OpExpression}, we see the abstract syntax tree of the corresponding
expression for state $s_7$.

{\input{figures/operation_based_invariant_confluent_definitions.tex}}

We say an expression $e$ is \defword{\sTIreachable{}} if it corresponds to a
valid execution of our system model. Formally, we define
$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
following equations:
\begin{itemize}
  \item
    $\sTIreachablepredicate{s_0}$.
  \item
    For all expressions $e_1, e_2$ and for all transactions $t$ in $T$, if
    $\sTIreachablepredicate{e_1}$, $\sTIreachablepredicate{e_2}$, and
    $I(t(e_1)(e_2))$, then $\sTIreachablepredicate{t(e_1)(e_2)}$.
\end{itemize}

Finally, we say $O$ is \invariantconfluent{} with respect to $s_0$, $T$, and
$I$, abbreviated \sTIconfluent{}, if all reachable states satisfy the
invariant:
\[
  \hfill
  \setst{s \in S}{\sTIreachablepredicate{s}} \subseteq I
  \hfill
\]

### 2020-01-19 doc/the_vldb_journal_paper/changelog.tex
Added changelog describing paper changes.
Since our journal submission is an extended version of our conference
submission, we are required to increase the content by 30% and to write
a note describing the changes. This commit includes that note.

\documentclass[12pt]{article}

\usepackage[letterpaper, margin=1in]{geometry}
\usepackage{natbib}

\begin{document}
\begin{center}
  {\huge Change Log for ``Interactive Checks for Coordination Avoidance''}

  Michael Whittaker, Joseph M. Hellerstein
\end{center}

Here, we describe the additions that we made to our VLDB 2019 conference
version of ``Interactive Checks for Coordination Avoidance''.

\begin{itemize}
  \item
    In our conference paper, we present a definition of invariant confluence
    that is slightly different than the original definition from Bailis et
    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
    We added a  subsection (Section 2.3) clarifying that the difference in
    definitions is superficial. We describe Bailis' original definition using
    our formalism and prove that the two definitions are equivalent.

  \item
    We expanded the correctness proof of our interactive decision procedure. A
    key step in understanding the correctness of our interactive invariant
    confluence decision procedure---namely that $I$-reachable and
    $(I-NR)$-reachable states were equal---was omitted from our conference
    paper. We added a formal proof of this fact as well as a few other minor
    elaborations.

  \item
    One of the core contributions of our paper is an interactive invariant
    confluence decision procedure. Here, ``interactive'' means that the
    decision procedure relies on human input to guide the decision procedure.
    To err is human. It's possible for a person to provide the decision
    procedure with incorrect information. We added a section (Section 4.3)
    describing how to mitigate these errors.

  \item
    We added a formal proof that criteria 1 through 4 are sufficient for
    invariant confluence. The idea behind the proof is straightforward but the
    details are tedious. To help readers see the forest through the trees, we
    also added an illustration of the key idea of the proof.

  \item
    Our conference paper presents formal definitions of distributed objects and
    invariant confluence. The definitions revolve around servers gossiping
    state to one another and merging the states together using a binary merge
    operator. This approach is called a ``state-based'' approach. We added a
    section (Section 7) outlining a dual ``operation-based'' approach in which
    transactions, rather than states, are gossiped. This duality between
    state-based and operation-based definitions is well established in the CRDT
    literature.

  \item
    We clarified that the merge operator of a distributed object does not have
    to satisfy any special mathematical properties like associativity,
    commutativity, or idempotence. Our results are more general and do not rely
    on any of these assumptions.

  \item
    We added a formal proof that invariant closure is sufficient for invariant
    confluence.

  \item
    We added a formal proof of our key result equating invariant closure and
    invariant confluence.

  \item
    In our conference paper, we asserted that checking for invariant closure is
    undecidable in general. We added a brief example showing \emph{why}
    invariant closure is undecidable, by reducing it to the Halting Problem.

  \item
    We added examples of merge-reducible and non-merge-reducible objects to
    make it easier for the reader to understand the definition of
    merge-reducibility.

  \item
    We added a formal proof that merge-reducibility is a sufficient condition
    for invariant-confluence. Similarly, we added a counterexample showing that
    merge-reducibility is not a necessary condition for invariant-confluence.

  \item
    In our conference paper, we provide an example of a distributed object that
    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
    closed. In the example, we asserted that criteria 1 through 4 are satisfied
    but did not elaborate on why. We added a brief explanation making it clear
    why the criteria are met.

  \item
    We added a Euler diagram illustrating the relationship between invariant
    confluence, invariant closure, merge reducibility, and criteria 1 through
    4.

  \item
    We added an example of co-reachable and co-unreachable states to help the
    reader understand the definition of co-reachability.

  \item
    We added a paragraph explaining the relationship between segmented
    invariant confluence and a distributed locking approach to maintaining
    invariants (that can be found in related works).

  \item
    We added a paragraph explaining a couple of optimizations that can be made
    to our segmented invariant confluence system model to reduce the frequency
    of global coordination.

  \item
    We added a paragraph explaining how a segmented invariant confluence
    decision procedure can leverage our merge reducibility results in addition
    to our results on invariant closure.

  \item
    We clarified the fault tolerance limitations of segmented invariant
    confluence. A naive implementation of our segmented invariant confluence
    decision procedure does not guarantee liveness, but there are
    straightforward techniques we can leverage to introduce liveness. We added
    an explanation of this to the paper.

  \item
    We added a ``Program Analysis in Database Systems'' paragraph to our
    related works section. In it, we describe other efforts to use program
    analysis to improve the performance of database systems.
\end{itemize}

\bibliographystyle{abbrv}
\bibliography{references}
\end{document}

### 2020-01-19 doc/the_vldb_journal_paper/figures/eval_table.tex
Formatted captions for Journal requirements.
They don't want punctuation at the end of a caption, and they want table
captions to describe the contents.

\begin{table}[t]
  \centering
  \caption{%
    \exampleref{TwoIntsEval} to \exampleref{TpccEval} summary. For every
    example, we list the time required (in seconds) to run our decision
    procedures to completion (excluding human input time) as well as the number
    of lines of code to specify the example
  }%
  \tablabel{EvalTable}
  \begin{tabular}{lrr}
    \toprule
    Example                                          & Run time (s) & Lines of code \\\midrule
    \ref{exa:TwoIntsEval}                            & 0.09         & 7 \\
    \ref{exa:ForeignKeysEval} (all transactions)     & 0.06         & 8 \\
    \ref{exa:ForeignKeysEval} (limited transactions) & 0.09         & 10 \\
    \ref{exa:AuctionEval}                            & 0.04         & 21 \\
    \ref{exa:EscrowTransactionsEval}                 & 0.09         & 49 \\
    \ref{exa:TpccEval} (Invariant 1)                 & 0.46         & 66 \\
    \ref{exa:TpccEval} (Invariant 2)                 & 0.44         & 33 \\
    \bottomrule
  \end{tabular}
\end{table}

### 2020-01-19 doc/the_vldb_journal_paper/sections/evaluation.tex
Formatted captions for Journal requirements.
They don't want punctuation at the end of a caption, and they want table
captions to describe the contents.

@@
     Segmented \invariantconfluent{} replication throughput versus coordination
-    induced by executing disallowed decrement transactions.
+    induced by executing disallowed decrement transactions
   }\figlabel{ThroughputVsGlobalSyncs}
@@
     Eventually consistent replication and linearizable replication are not
-    affected by workload.
+    affected by workload
   }\figlabel{VaryNodes}

### 2020-01-19 doc/the_vldb_journal_paper/sections/conclusion.tex
Typo fixes.

--- 2020-01-13 doc/vldb_paper/sections/conclusion.tex
@@
 \textbf{Acknowledgments.}
-The authors would like to thank Alan Fekete, Peter Alvaro, Alvin Cheung,
-Alexandra Meliou, Anthony Tan, and Cristina Teodoropol for fruitful discussion
-and feedback.
+The authors would like to thank Alan Fekete, Alexandra Meliou, Alvin Cheung,
+Anthony Tan, Cristina Teodoropol, Peter Alvaro, and Peter Bailis, for fruitful
+discussion and feedback.
 %

### 2020-01-19 doc/the_vldb_journal_paper/sections/merge_reduction.tex
Typo fixes.

@@
 we discover that all points with a Manhattan distance of $k$ from the top left
-corner satisfy the invariant. Ultimately, we conclude that the bottom left
+corner satisfy the invariant. Ultimately, we conclude that the bottom right
 corner (i.e., $e_1 \join e_2$) satisfies the invariant and is equivalent to
@@
   only even integers. If $t_1$ and $t_2$ are arbitrary transactions such that
-  $I(t_1(s))$ and $I(t_2(s))$, then $t_1(s)$ and $t_1(s)$ are also non-empty
+  $I(t_1(s))$ and $I(t_2(s))$, then $t_1(s)$ and $t_2(s)$ are also non-empty
   and contain only even integers. Thus, $t_1(s) \cup t_2(s)$ is clearly

### 2020-01-19 doc/the_vldb_journal_paper/sections/operation_based_invariant_confluence.tex
Typo fixes.

@@
     $\sTIreachablepredicate{e_1}$, $\sTIreachablepredicate{e_2}$, and
-    $I(t(e_1)(e_2))$, then $\sTIreachablepredicate{t(e_1)(e_2)}$.
+    $I(t(e_1)(e_1))$, then $\sTIreachablepredicate{t(e_1)(e_2)}$.
 \end{itemize}

### 2020-01-28 doc/the_vldb_journal_paper/changelog.tex
Added forgotten point to changelog.
Added references to changelog.
Thanks Joe for the feedback.

@@
 \begin{itemize}
-  \item
+  \item (Section 2.1)
+    We clarified that the merge operator of a distributed object does not have
+    to satisfy any special mathematical properties like associativity,
+    commutativity, or idempotence. Our results are more general and do not rely
+    on any of these assumptions.
+
+  \item (Section 2.3)
     In our conference paper, we present a definition of invariant confluence
@@
 
-  \item
+  \item (Section 3, Theorem 2)
+    We added a formal proof that invariant closure is sufficient for invariant
+    confluence.
+
+  \item (Section 3, Theorem 3)
+    We added a formal proof of our key result equating invariant closure and
+    invariant confluence.
+
+  \item (Section 4.1)
     We expanded the correctness proof of our interactive decision procedure. A
@@
 
-  \item
+  \item (Section 4.2)
+    In our conference paper, we asserted that checking for invariant closure is
+    undecidable in general. We added a brief example showing \emph{why}
+    invariant closure is undecidable, by reducing it to the Halting Problem.
+
+  \item (Section 4.3)
     One of the core contributions of our paper is an interactive invariant
@@
     decision procedure relies on human input to guide the decision procedure.
-    To err is human. It's possible for a person to provide the decision
-    procedure with incorrect information. We added a section (Section 4.3)
-    describing how to mitigate these errors.
+    It's possible for a person to provide the decision procedure with incorrect
+    information. We added a subsection (Section 4.3) describing how to mitigate
+    these errors.
 
-  \item
+  \item (Section 5, Examples 3 and 4)
+    We added examples of merge-reducible and non-merge-reducible objects to
+    make it easier for the reader to understand the definition of
+    merge-reducibility.
+
+  \item (Section 5, Theorem 4 and Figure 4)
+    We added a formal proof that merge-reducibility is a sufficient condition
+    for invariant-confluence. We also added a diagram illustrating the key idea
+    behind the proof. Similarly, we added a counterexample showing that
+    merge-reducibility is not a necessary condition for invariant-confluence.
+
+  \item (Section 5, Theorem 5 and Figure 5)
     We added a formal proof that criteria 1 through 4 are sufficient for
@@
 
-  \item
+  \item (Section 5, Example 5)
+    In our conference paper, we provide an example of a distributed object that
+    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
+    closed. In the example, we asserted that criteria 1 through 4 are satisfied
+    but did not elaborate on why. We added a brief explanation making it clear
+    why the criteria are met.
+
+  \item (Section 5, Figure 6)
+    We added an Euler diagram illustrating the relationship between invariant
+    confluence, invariant closure, merge reducibility, and criteria 1 through
+    4.
+
+  \item (Section 6.3, Example 7)
+    We added an example of co-reachable and co-unreachable states to help the
+    reader understand the definition of co-reachability.
+
+  \item (Section 6.4)
+    We added a paragraph explaining the relationship between segmented
+    invariant confluence and a distributed locking approach to maintaining
+    invariants (that can be found in related works).
+
+  \item (Section 6.4)
+    We added a paragraph explaining a couple of optimizations that can be made
+    to our segmented invariant confluence system model to reduce the frequency
+    of global coordination.
+
+  \item (Section 6.4)
+    We added a paragraph explaining how a segmented invariant confluence
+    decision procedure can leverage our merge reducibility results in addition
+    to our results on invariant closure.
+
+  \item (Section 6.4)
+    We clarified the fault tolerance limitations of segmented invariant
+    confluence. A naive implementation of our segmented invariant confluence
+    decision procedure does not guarantee liveness, but there are
+    straightforward techniques we can leverage to introduce liveness. We added
+    an explanation of this to the paper.
+
+  \item (Section 7)
     Our conference paper presents formal definitions of distributed objects and
@@
 
-  \item
-    We clarified that the merge operator of a distributed object does not have
-    to satisfy any special mathematical properties like associativity,
-    commutativity, or idempotence. Our results are more general and do not rely
-    on any of these assumptions.
-
-  \item
-    We added a formal proof that invariant closure is sufficient for invariant
-    confluence.
-
-  \item
-    We added a formal proof of our key result equating invariant closure and
-    invariant confluence.
-
-  \item
-    In our conference paper, we asserted that checking for invariant closure is
-    undecidable in general. We added a brief example showing \emph{why}
-    invariant closure is undecidable, by reducing it to the Halting Problem.
-
-  \item
-    We added examples of merge-reducible and non-merge-reducible objects to
-    make it easier for the reader to understand the definition of
-    merge-reducibility.
-
-  \item
-    We added a formal proof that merge-reducibility is a sufficient condition
-    for invariant-confluence. Similarly, we added a counterexample showing that
-    merge-reducibility is not a necessary condition for invariant-confluence.
-
-  \item
-    In our conference paper, we provide an example of a distributed object that
-    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
-    closed. In the example, we asserted that criteria 1 through 4 are satisfied
-    but did not elaborate on why. We added a brief explanation making it clear
-    why the criteria are met.
-
-  \item
-    We added a Euler diagram illustrating the relationship between invariant
-    confluence, invariant closure, merge reducibility, and criteria 1 through
-    4.
-
-  \item
-    We added an example of co-reachable and co-unreachable states to help the
-    reader understand the definition of co-reachability.
-
-  \item
-    We added a paragraph explaining the relationship between segmented
-    invariant confluence and a distributed locking approach to maintaining
-    invariants (that can be found in related works).
-
-  \item
-    We added a paragraph explaining a couple of optimizations that can be made
-    to our segmented invariant confluence system model to reduce the frequency
-    of global coordination.
-
-  \item
-    We added a paragraph explaining how a segmented invariant confluence
-    decision procedure can leverage our merge reducibility results in addition
-    to our results on invariant closure.
-
-  \item
-    We clarified the fault tolerance limitations of segmented invariant
-    confluence. A naive implementation of our segmented invariant confluence
-    decision procedure does not guarantee liveness, but there are
-    straightforward techniques we can leverage to introduce liveness. We added
-    an explanation of this to the paper.
-
-  \item
+  \item (Section 9)
     We added a ``Program Analysis in Database Systems'' paragraph to our

### 2020-01-28 doc/the_vldb_journal_paper/changelog.tex
Minor changelog reword.

@@
     confluence decision procedure---namely that $I$-reachable and
-    $(I-NR)$-reachable states were equal---was omitted from our conference
+    $(I-NR)$-reachable states are equal---was omitted from our conference
     paper. We added a formal proof of this fact as well as a few other minor

### 2020-05-16 doc/sigmod_paper/invariant_confluence.tex
Copied VLDB to SIGMOD.
I copied over the VLDB paper but ported it to the SIGMOD style files.
Now, we just have to trim!

\documentclass{research4cacm}

\usepackage[noend]{algpseudocode}
\usepackage{algorithm}
\usepackage{balance}
\usepackage{booktabs}
\usepackage{invariantconfluence}
\usepackage{listings}
\usepackage{pervasives}
\usepackage{printlen}
\usepackage{python}
\usepackage{subcaption}
\usepackage{tikz}
\usetikzlibrary{arrows}
\usetikzlibrary{calc}
\usetikzlibrary{patterns}

\begin{document}
\title{%
  TODO%
  \thanks{%
    The original version of this paper is entitled ``Interactive Checks for
    Coordination Avoidance'' and was published in PVLDB Vol.\ 12.
  }%
}

\numberofauthors{2}
\author{
\alignauthor
Michael Whittaker\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{mjwhittaker@berkeley.edu}
\alignauthor
Joseph M. Hellerstein\\
       \affaddr{UC Berkeley}\\
       \affaddr{Berkeley, CA}\\
       \email{hellerstein@berkeley.edu}
}

\maketitle

{\input{sections/abstract}}
{\input{sections/introduction}}
{\input{sections/invariant_confluence}}
{\input{sections/invariant_closure}}
{\input{sections/interactive_decision_procedure}}
{\input{sections/merge_reduction}}
{\input{sections/segmented_invariant_confluence.tex}}
{\input{sections/evaluation}}
{\input{sections/related_work}}
{\input{sections/conclusion}}

%ACKNOWLEDGMENTS are optional
\section{Acknowledgments}
This section is optional; it is a location for you
to acknowledge grants, funding, editing assistance, collaboration, access to
uncommon, but necessary, hardware or software etc.
In the present case, for example, the
authors would like to thank Gerald Murray of ACM for
his help in codifying this \textit{Author's Guide}
and the \textbf{.cls} and \textbf{.tex} files that it describes.

\bibliographystyle{abbrv} % standard abbrv style
\bibliography{references}
\balancecolumns
\end{document}

### 2020-05-16 doc/sigmod_paper/sections/merge_reduction.tex
Copied VLDB to SIGMOD.
I copied over the VLDB paper but ported it to the SIGMOD style files.
Now, we just have to trim!

--- 2018-09-20 doc/paper/sections/merge_reduction.tex
@@
 \begin{techreport}
-  The proof of \thmref{LatticeProperty} involves proving that if criteria (1)
-  through (4) are met, then $O$ is \sTImergereducible{}. From this, it's
-  immediate that $O$ is \sTIconfluent{} by
-  \thmref{ReducibilityImpliesIconfluence}. To show that $O$ is
-  \sTImergereducible, we consider two merge-free reachable expressions
+  To show that $O$ is \sTImergereducible, we consider two merge-free reachable
+  expressions
   \[

### 2020-05-16 doc/sigmod_paper/invariant_confluence.tex
Cut merge reduction section; split acks out.

@@
 {\input{sections/interactive_decision_procedure}}
-{\input{sections/merge_reduction}}
 {\input{sections/segmented_invariant_confluence.tex}}
@@
 {\input{sections/conclusion}}
-
-%ACKNOWLEDGMENTS are optional
-\section{Acknowledgments}
-This section is optional; it is a location for you
-to acknowledge grants, funding, editing assistance, collaboration, access to
-uncommon, but necessary, hardware or software etc.
-In the present case, for example, the
-authors would like to thank Gerald Murray of ACM for
-his help in codifying this \textit{Author's Guide}
-and the \textbf{.cls} and \textbf{.tex} files that it describes.
+{\input{sections/acknowledgements}}
 

### 2020-05-16 doc/sigmod_paper/sections/conclusion.tex
Cut merge reduction section; split acks out.

--- 2020-01-13 doc/vldb_paper/sections/conclusion.tex
@@
 replication for low-coordination workloads.
-
-\textbf{Acknowledgments.}
-The authors would like to thank Alan Fekete, Peter Alvaro, Alvin Cheung,
-Alexandra Meliou, Anthony Tan, and Cristina Teodoropol for fruitful discussion
-and feedback.
-%
-This research is supported in part by DHS Award HSHQDC-16-3-00083, NSF CISE
-Expeditions Award CCF-1139158, and gifts from Alibaba, Amazon Web Services, Ant
-Financial, CapitalOne, Ericsson, GE, Google, Huawei, Intel, IBM, Microsoft,
-Scotiabank, Splunk and VMware.

### 2020-05-16 doc/sigmod_paper/sections/abstract.tex
Copied VLDB to SIGMOD.
I copied over the VLDB paper but ported it to the SIGMOD style files.
Now, we just have to trim!
Fixed copyright.
The publisher gave us some specific instructions on what the copyright
should say. Hopefully, I've done this right.

--- 2018-09-20 doc/paper/sections/abstract.tex
@@
 In this paper, we establish conditions under which a commonly used sufficient
-condition for \invariantconfluence{} is both necessary and sufficient, and we use
-this condition to design (a) a general-purpose interactive
-\invariantconfluence{} decision procedure and (b) a novel sufficient condition
-that can be checked automatically. We then take a step beyond
-\invariantconfluence{} and introduce a generalization of \invariantconfluence{},
-called segmented \invariantconfluence{}, that allows us to replicate
-non-\invariantconfluent{} objects with a small amount of coordination.
-
+condition for \invariantconfluence{} is both necessary and sufficient, and we
+use this condition to design a general-purpose interactive
+\invariantconfluence{} decision procedure. We then take a step beyond
+\invariantconfluence{} and introduce a generalization of
+\invariantconfluence{}, called segmented \invariantconfluence{}, that allows us
+to replicate non-\invariantconfluent{} objects with a small amount of
+coordination.
+%
 We implemented these formalisms in a prototype called Lucy and found that our
 decision procedures efficiently handle common real-world workloads including
-foreign keys, rollups, escrow transactions, and more.  We also found that
-segmented \invariantconfluent{} replication can deliver up to an order of
-magnitude more throughput than linearizable replication for low contention
-workloads and comparable throughput for medium to high contention workloads.
+%
+% foreign keys, rollups, escrow transactions, and more.
+foreign keys, escrow transactions, and more.
+%
+We also found that segmented \invariantconfluent{} replication can deliver up
+to an order of magnitude more throughput than linearizable replication.
+% for low contention workloads and comparable throughput for medium to high
+% contention workloads.
 \end{abstract}

### 2020-05-16 doc/sigmod_paper/sections/conclusion.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

@@
 %
-First, we developed a deeper understanding of \invariantclosure{} and
-\invariantconfluence{} by looking at the two criteria with reachability in
-mind. We found that \invariantclosure{} fails to incorporate a notion of
+First,
+% we developed a deeper understanding of \invariantclosure{} and
+% \invariantconfluence{} by looking at the two criteria with reachability in
+% mind.
+we found that \invariantclosure{} fails to incorporate a notion of
 reachability, and using this intuition, we developed conditions under which
@@
 assistance of a programmer.
-
+%
 Second, we proposed a new consistency model and generalization of
@@
 replicate non-\invariantconfluent{} objects with a small amount of coordination
-while still preserving their invariants. We found that segmented
-\invariantconfluence{} naturally subsumes existing techniques for maintaining
-invariants of replicated objects (e.g.\ locking and escrow transactions), and
-we developed an interactive decision procedure for segmented
-\invariantconfluence{}.
+while still preserving their invariants.
+% We found that segmented
+% \invariantconfluence{} naturally subsumes existing techniques for maintaining
+% invariants of replicated objects (e.g.\ locking and escrow transactions), and
+% we developed an interactive decision procedure for segmented
+% \invariantconfluence{}.
 
-Through our evaluation, we found that our decision procedures could analyze a
-number of realistic workloads, each in less than a second. We also showed that
-segmented \invariantconfluence{} can significantly outperform linearizable
-replication for low-coordination workloads.
+% Through our evaluation, we found that our decision procedures could analyze a
+% number of realistic workloads, each in less than a second. We also showed that
+% segmented \invariantconfluence{} can significantly outperform linearizable
+% replication for low-coordination workloads.

### 2020-05-16 doc/sigmod_paper/sections/interactive_decision_procedure.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

--- 2018-09-20 doc/paper/sections/interactive_decision_procedure.tex
@@
 
-\subsection{The Decision Procedure}
+% \subsection{The Decision Procedure}
 A user provides \algoref{InteractiveDecisionProcedure} with an object $O = (S,
@@
 
-\subsection{Limitations}
-Our interactive \invariantconfluence{} decision procedure has two
-limitations. First, \algoref{InteractiveDecisionProcedure} requires an
-\invariantclosure{} decision procedure, but determining \invariantclosure{} is
-undecidable in general.
+% \subsection{Limitations}
+% Our interactive \invariantconfluence{} decision procedure has two
+% limitations. First, \algoref{InteractiveDecisionProcedure} requires an
+% \invariantclosure{} decision procedure, but determining \invariantclosure{} is
+% undecidable in general.
+% %
+% In practice, we can implement an \invariantclosure{} decision procedure using
+% an SMT solver like Z3 that works well on simple objects, invariants, and merge
+% operators (e.g., integers, tuples, infinite sets, bit vectors, linear
+% constraints, basic arithmetic, tuple projection, basic set operations, bit
+% arithmetic). But, SMT solvers are mostly unable to analyze more complex
+% constructs (e.g., finite lists~\cite{kroning2009proposal}, graphs, recursive
+% algebraic data types, nonlinear constraints, merge operators that contain loops
+% or recursion).
 %
-\begin{techreport}
-
-  For example, let $O_p = (S, \join)$ where $S$ is the set of all programs and
-  $s_1 \join s_2 = p$ for some fixed program $p$. Letting $I$ be the set of all
-  programs that terminate, determining if $O_p$ is \Iclosed{} is tantamount to
-  determining if $p$ terminates.
-\end{techreport}
-%
-In practice, we can implement an \invariantclosure{} decision procedure using
-an SMT solver like Z3 that works well on simple objects, invariants, and merge
-operators (e.g., integers, tuples, infinite sets, bit vectors, linear
-constraints, basic arithmetic, tuple projection, basic set operations, bit
-arithmetic). But, SMT solvers are mostly unable to analyze more complex
-constructs (e.g., finite lists~\cite{kroning2009proposal}, graphs, recursive
-algebraic data types, nonlinear constraints, merge operators that contain loops
-or recursion).
-
-Second, \algoref{InteractiveDecisionProcedure} relies on a user to identify
-unreachable states. As we saw in \figref{DecisionProcedureExample}, the set of
-unreachable states can sometimes be clear, especially if there's a noticeable
-pattern in the set of reachable states. However, if the set of transactions is
-large or complex or if the merge operator is complex, then reasoning about
-unreachable states can be difficult.  Unlike with reachable states---where
-verifying that a state is reachable only requires thinking of a single way to
-reach the state---verifying that a state is unreachable requires a programmer
-to reason about a large number of system executions and conclude that
-\emph{none} of them can lead to the state. In the future, we plan on exploring
-ways to help a user reason about unreachable states and ways to discover sets of
-unreachable states automatically.
-
-\begin{techreport}
-  \subsection{Tolerating User Error}
-  \algoref{InteractiveDecisionProcedure} is an \emph{interactive} decision
-  procedure. It requires that a user classify states as reachable or
-  unreachable.  To err is human, so what happens when a user incorrectly
-  classifies a state?  There are two possible errors that can be made, and
-  \algoref{InteractiveDecisionProcedure} can be made robust to both.
-
-  \textbf{A user can label an unreachable state as reachable.}
-  In this case, \Helper{} might erroneously find $s_1$ and $s_2 \in R$ and
-  return false, concluding that $O$ is not \sTIconfluent{} even when it is.
-  This is inconvenient, but not catastrophic. We can modify
-  \algoref{InteractiveDecisionProcedure} so that \Helper{} requires that
-  whenever a user labels a state $s$ as \sTIreachable{}, they must also provide
-  an \sTIreachable{} expression $e$ that evaluates to $s$. Here, $e$ acts a
-  proof that certifies that $s$ is indeed reachable. This increases the burden
-  on the user but completely eliminates this type of user error.
-
-  \textbf{A user can label a reachable state as unreachable.}
-  In this case, \IsIclosed$(O, I - NR)$ might return true, even though $O$ is
-  not \sTIconfluent. Thus, a user might falsely believe their distributed
-  object to be \sTIconfluent{} even though it isn't, and eventually one replica
-  of their distributed object might enter a state that violates the invariant.
-  %
-  We can mitigate this in two ways. First, we can aggressively expand $R$
-  automatically. If a user ever labels a state $s$ as unreachable, but $s \in
-  R$, we can notify the user of their mistake. Second, \Helper{} returns true
-  when $O$ is \iclosed{(I - NR)} for some $NR$, so $O$ \emph{is}
-  \sticonfluent{s_0}{T}{I-NR} (even though it might not be \sTIconfluent{}).
-  Thus, when a user replicates their distributed object across a set of
-  servers, they can deploy with the invariant $I - NR$ instead of $I$. If $NR$
-  is correct and only contains unreachable states, then deploying with $I - NR$
-  is equivalent to deploying with $I$. If $NR$ is incorrect and contains some
-  \sTIreachable{} states, then some of these states are artificially made
-  unreachable, but the system is still guaranteed to never produce a state
-  that violates $I$.
-\end{techreport}
+% Second, \algoref{InteractiveDecisionProcedure} relies on a user to identify
+% unreachable states. As we saw in \figref{DecisionProcedureExample}, the set of
+% unreachable states can sometimes be clear, especially if there's a noticeable
+% pattern in the set of reachable states. However, if the set of transactions is
+% large or complex or if the merge operator is complex, then reasoning about
+% unreachable states can be difficult.  Unlike with reachable states---where
+% verifying that a state is reachable only requires thinking of a single way to
+% reach the state---verifying that a state is unreachable requires a programmer
+% to reason about a large number of system executions and conclude that
+% \emph{none} of them can lead to the state. In the future, we plan on exploring
+% ways to help a user reason about unreachable states and ways to discover sets of
+% unreachable states automatically.

### 2020-05-16 doc/sigmod_paper/sections/introduction.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

@@
 Then, we use these conditions to design a general-purpose interactive
-\invariantconfluence{} decision procedure and a new sufficient condition for
-\invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
-covers some cases that are not covered by \invariantclosure{}, and it can be
-checked automatically.
+\invariantconfluence{} decision procedure.
 
@@
 system for replicating \invariantconfluent{} and segmented
-\invariantconfluent{} objects. Using Lucy, we find that our decision
-procedures can efficiently handle a wide range of common workloads. For
-example, in \secref{Evaluation}, we apply Lucy to foreign key constraints,
-escrow transactions, an auction application, and the TPC-C benchmark. Lucy
-processes these workloads in less than half a second, and no workload requires
-more than 66 lines of code to specify. Moreover, we find that segmented
-\invariantconfluent{} replication can achieve 10x to 100x more throughput than
-linearizable replication for low-coordination workloads.
+\invariantconfluent{} objects. Using Lucy, we find that our decision procedures
+can efficiently handle a wide range of common workloads.  For example, in
+\secref{Evaluation}, we apply Lucy to foreign key constraints and escrow
+transactions. Lucy processes these workloads in less than half a second, and
+neither workload requires more than 66 lines of code to specify. Moreover, we
+find that segmented \invariantconfluent{} replication can achieve 10x to 100x
+more throughput than linearizable replication for low-coordination workloads.
 
-In closing, here is an outline of the paper and of our contributions:
-%
-We propose a novel expression-oriented definition of \invariantconfluence{}
-that is both formal and simple (\secref{InvariantConfluence}).
-%
-We develop an understanding of why \invariantclosure{} is not necessary for
-\invariantconfluence{} and use this understanding to develop conditions under
-which it is both necessary and sufficient (\secref{InvariantClosure}).
-%
-We exploit these conditions to design a general-purpose interactive decision
-procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
-%
-We again exploit these conditions to design a novel non-trivial sufficient
-condition for \invariantconfluence{}, called \mergereducibility.
-%
-We present segmented \invariantconfluence{}: a generalization of
-\invariantconfluence{} that uses a small amount of coordination to maintain
-invariants for replicated objects that are otherwise not \invariantconfluent{}
-(\secref{SegmentedInvariantConfluence}).
-%
-We evaluate our methods using a prototype implementation called Lucy
-(\secref{Evaluation}).
+% In closing, here is an outline of the paper and of our contributions:
+% %
+% We propose a novel expression-oriented definition of \invariantconfluence{}
+% that is both formal and simple (\secref{InvariantConfluence}).
+% %
+% We develop an understanding of why \invariantclosure{} is not necessary for
+% \invariantconfluence{} and use this understanding to develop conditions under
+% which it is both necessary and sufficient (\secref{InvariantClosure}).
+% %
+% We exploit these conditions to design a general-purpose interactive decision
+% procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
+% %
+% We again exploit these conditions to design a novel non-trivial sufficient
+% condition for \invariantconfluence{}, called \mergereducibility.
+% %
+% We present segmented \invariantconfluence{}: a generalization of
+% \invariantconfluence{} that uses a small amount of coordination to maintain
+% invariants for replicated objects that are otherwise not \invariantconfluent{}
+% (\secref{SegmentedInvariantConfluence}).
+% %
+% We evaluate our methods using a prototype implementation called Lucy
+% (\secref{Evaluation}).

### 2020-05-16 doc/sigmod_paper/sections/invariant_confluence.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

--- 2018-09-19 doc/paper/sections/invariant_confluence.tex
@@
 
-Note that by modelling a transaction $t$ as a function $S \to S$, we focus
-exclusively on the effects that a transaction has on the object (i.e.\
-``writes'' to the object). Transactions are also free to read the value of the
-object, but these reads are not captured by our model because, as we'll see,
-they do not affect \invariantconfluence{}. For example, we could model any
-read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.
+% Note that by modelling a transaction $t$ as a function $S \to S$, we focus
+% exclusively on the effects that a transaction has on the object (i.e.\
+% ``writes'' to the object). Transactions are also free to read the value of the
+% object, but these reads are not captured by our model because, as we'll see,
+% they do not affect \invariantconfluence{}. For example, we could model any
+% read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.
 
@@
 s_j$. Servers periodically merge states with one another in order to keep their
-states loosely synchronized\footnote{%
-  Notably, if $O$ is a CRDT---i.e. $O$ is a semilattice and every transaction
-  $t \in T$ is inflationary---then this periodic merging ensures that $O$ is
-  strongly eventually consistent~\cite{shapiro2011conflict}.
-}.
+states loosely synchronized.
+% \footnote{%
+%   Notably, if $O$ is a CRDT---i.e. $O$ is a semilattice and every transaction
+%   $t \in T$ is inflationary---then this periodic merging ensures that $O$ is
+%   strongly eventually consistent~\cite{shapiro2011conflict}.
+% }.
 Note that unlike with transactions, servers \emph{cannot} abort a merge; server

### 2020-05-16 doc/sigmod_paper/sections/related_work.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

\section{Related Work}
% Other possibilities:
%   - Towards Fast Invariant Preservation in Geo-replicated Systems
%   - Feral Concurrency Control
%   - The CISE Tool: Proving Weakly-Consistent Applications Correct
%   - Declarative Programming over Eventually Consistent Data Stores
%   - Extending Eventually Consistent Cloud Databases for Enforcing Numeric Invariants
% In this section, we detail four alternate approaches to maintaining invariants
% without coordination: RedBlue consistency and Sieve, the homeostasis and
% demarcation protocols, explicit consistency, and token based invariant
% confluence. We then survey other bodies of related work.

\newcommand{\sieve}{\textsc{SIEVE}}
% \textbf{RedBlue Consistency and \sieve.}
RedBlue consistency is a consistency model that sits between causal consistency
and linearizability~\cite{li2012making}.
%
% With RedBlue consistency, every operation is manually labelled as either red or
% blue. All operations are executed with causal consistency, but with the added
% restrictions that red operations are executed in a single total order embedded
% within the causal ordering.
%
In~\cite{li2012making}, Li et al.\ introduce invariant safety as a
sufficient (but not necessary) condition for RedBlue consistent objects to be
\invariantconfluent{}. Invariant safety is an analog of \invariantclosure{}.
In~\cite{li2014automating}, Li et al.\ develop sophisticated techniques for
deciding invariant safety that involve calculating weakest preconditions.
% These
% techniques are complementary to our work and can be used to improve the
% \invariantclosure{} subroutine used by our decision procedures.
%
% In contrast with these techniques, our \invariantconfluence{} decision
% procedures can determine the \invariantconfluence{} of objects that are
% \emph{not} invariant safe.

% \textbf{The Demarcation and Homeostasis Protocols.}
The homeostasis protocol~\cite{roy2015homeostasis}, a generalization of the
demarcation protocol~\cite{barbara1994demarcation}, uses program analysis to
avoid unnecessary coordination between servers in a \emph{sharded} database
(whereas \invariantconfluence{} targets \emph{replicated} databases).
% The
% protocol guarantees that transactions are executed with observational
% equivalence with respect to some serial execution of the transactions. This
% means that intermediate states may be inconsistent, but externally observable
% side effects and the final database state are consistent.  The observational
% equivalence guaranteed by the homeostasis protocol is stronger than the
% guarantees of \invariantconfluence{}. As a result, there are invariants and
% workloads that the homeostasis protocol would execute with more coordination
% than a segmented \invariantconfluent{} execution. Moreover, the homeostasis and
% demarcation protocols' mechanism of establishing global invariants and
% operating without coordination so long as the invariants are maintained is very
% similar to our design of segmented \invariantconfluence{}.

% \textbf{Explicit Consistency.}
Explicit consistency~\cite{balegas2015towards} is a consistency model that
combines \invariantconfluence{} and causal consistency, similar to RedBlue
consistency with invariant safety.
% To determine if a workload is amenable to
% explicitly consistent replication, Balegas et al.\ determine if all pairs of
% transactions can be concurrently executed on the same start state without
% violating the invariant~\cite{balegas2015towards}. Balegas et al.\ argue that
% this is a sufficient condition for explicit consistency. It is similar to
% criterion (3) in \thmref{LatticeProperty}. In our work, we take a step further
% and explore sufficient \emph{and necessary} conditions for
% \invariantconfluence{}.
%
Balegas et al.\ also describe a variety of techniques---like conflict
resolution, locking, and escrow transactions~\cite{o1986escrow}---that can be
used to replicate workloads that do not meet their sufficient conditions.
Segmented \invariantconfluence{} is a general-purpose formalism that can be
used to model simple forms of these techniques.

% \textbf{Token Based Invariant Confluence.}
In~\cite{gotsman2016cause}, Gotsman et al.\ discuss a hybrid token based
consistency model that generalizes a family of consistency models including
causal consistency, sequential consistency, and RedBlue consistency.
% An
% application designer defines a set of tokens and specifies which pairs of
% tokens conflict, and transactions acquire some subset of the tokens when they
% execute. This allows the application designer to specify which transactions
% conflict with one another.  Gotsman et al.\ develop sufficient conditions to
% determine whether a given token scheme is sufficient to guarantee that a global
% invariant is never broken.
The token based approach allows users to
specify certain conflicts that are not possible with segmented
\invariantconfluence{} because a segmentation only allows transactions within a
segment to acquire a single self-conflicting lock. However, segmented
\invariantconfluence{} also introduces the notion of invariant segmentation,
which cannot be emulated with the token based approach.
% For example, it is
% difficult to emulate escrow transactions with the token based approach.

% \textbf{Serializable Distributed Databases.}
% In \secref{Evaluation}, we saw that segmented \invariantconfluent{} replication
% vastly outperforms linearizable replication for low coordination workloads, and
% it performs comparably or worse for medium and high coordination workloads.
% Distributed databases like Calvin~\cite{thomson2012calvin},
% Janus~\cite{mu2016consolidating}, and TAPIR~\cite{zhang2015building} employ
% algorithmic optimizations to implement serializable transactions with high
% throughput and low latency.
% %
% While segmented \invariantconfluent{} replication will likely always outperform
% serializable replication for low coordination workloads, these databases make
% serializable replication the most performant option for executing workloads
% that require a modest amount of coordination.

% \textbf{Branch and Merge.}
Bayou~\cite{terry1995managing}, Dynamo~\cite{decandia2007dynamo}, and
TARDiS~\cite{crooks2016tardis} all take a branch and merge approach to
maintaining distributed invariants without coordination. With this approach,
servers execute transactions without any coordination but keep track of the
causal dependencies between transactions. Periodically, two servers merge
states and invoke a user defined merge function to reconcile the divergent
states. This approach does \emph{not} provide any formal guarantees that
invariants are maintained.
% Its correctness depends on the correctness of the
% potentially complex user defined merge functions.

% \textbf{CRDTs.}
CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
semilattices with inflationary update methods. Due to their algebraic
properties, CRDTs can be replicated with strong eventual consistency without
the need for any coordination. Our definition of distributed objects and our
\invariantconfluence{} system model are inspired directly by the corresponding
definitions and system models in~\cite{shapiro2011conflict}.

%
% CRDTs are eventually consistent but may not preserve invariants. Conversely,
% \invariantconfluent{} objects preserve invariants but may not be eventually
% consistent. Thus, it is natural (though not necessary) to use CRDTs as
% distributed objects. If a CRDT is determined to be \invariantconfluent{} with
% respect to a particular invariant and set of transactions, then it achieves a
% combination of strong eventual consistency and invariant preservation. Any CRDT
% (e.g., counters, sets, graphs, sequences) can be used for this purpose.
% %
% Finally, our criteria in \thmref{LatticeProperty} also borrow ideas from CRDTs,
% exploiting the algebraic properties of semilattices.

% \textbf{CALM Theorem.}
% Bloom~\cite{alvaro2010boom, alvaro2011consistency, conway2012logic} and its
% formalism, Dedalus~\cite{alvaro2011dedalus, alvaro2013declarative}, are
% declarative Datalog-based programming languages that are designed to program
% distributed systems. The accompanying CALM
% theorem~\cite{hellerstein2010declarative, ameloot2013relational} states that if
% and only if a program can be written in the monotone fragment of these
% languages, then there exists a consistent, coordination-free implementation of
% the program.  The CALM theorem provides guarantees about the consistency of
% program outputs. It does not directly capture our notions of transactions or
% invariant maintenance during program execution.  Moreover, Bloom and Dedalus
% are general-purpose programming languages that can be used to implement a
% variety of distributed systems that are outside of the scope of
% \invariantconfluence{}.

\begin{techreport}

  \textbf{Program Analysis in Database Systems.} Using program analysis to
  improve the performance of database systems is not new. For example, it has
  been used to improve the performance of database-backed web
  applications~\cite{cheung2014using, wu2016transaction,
  ramachandra2012program} and used to improve the performance of optimistic
  concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
  on \invariantconfluence{} continues the theme of using program analysis to
  reap the performance benefits gained from understanding program semantics.
\end{techreport}

### 2020-05-16 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.

--- 2018-09-20 doc/paper/sections/segmented_invariant_confluence.tex
@@
 
-The proof of the forward direction is exactly the same as the proof of
-\thmref{IclosureImpliesIconfluence}. Transactions always maintain the
-invariant, so if merge does as well, then every reachable state must satisfy
-the invariant.
-%
-For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
-The two points are \TIcoreachable{}, so there exists some state $s_0$ from
-which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
-\sTIreachable{}, so it satisfies the invariant.
+% The proof of the forward direction is exactly the same as the proof of
+% \thmref{IclosureImpliesIconfluence}. Transactions always maintain the
+% invariant, so if merge does as well, then every reachable state must satisfy
+% the invariant.
+% %
+% For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
+% The two points are \TIcoreachable{}, so there exists some state $s_0$ from
+% which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
+% \sTIreachable{}, so it satisfies the invariant.
 
@@
 set of \emph{coreachable} states, but otherwise, the core of the two algorithms
-is the same.\footnote{%
-  Another small difference is that \IsIclosed{} behaves differently in
-  \algoref{InteractiveDecisionProcedure} and
-  \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
-  \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
-  triple (closed, $s_1$, $s_2$). If closed is false, then $s_1, s_2 \in I$
-  are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
-  \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
-  $s_2$ are null.
-}
+is the same.
+% \footnote{%
+%   Another small difference is that \IsIclosed{} behaves differently in
+%   \algoref{InteractiveDecisionProcedure} and
+%   \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
+%   \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
+%   triple (closed, $s_1$, $s_2$). If closed is false, then $s_1, s_2 \in I$
+%   are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
+%   \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
+%   $s_2$ are null.
+% }
 Now, a segmented \invariantconfluence{} decision procedure, can simply invoke
@@
 
-\subsection{Discussion and Limitations}
-There are a few things worth noting about segmented \invariantconfluence{}, its
-system model, and its decision procedure.
+% \subsection{Discussion and Limitations}
+% There are a few things worth noting about segmented \invariantconfluence{}, its
+% system model, and its decision procedure.
+% %
+% First, \invariantconfluence{} is a very coarse-grained property. If an object
+% is \invariantconfluent{}, then we can replicate it with no coordination. If it
+% is not \invariantconfluent{}, then we have no guarantees. There's no
+% in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
+% more fine-grained property that can be applied to applications with varying
+% degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
+% complex applications that require a large number of segments and to simple
+% applications with a smaller number of segments, whereas \invariantconfluence{}
+% only provides guarantees to applications that can be segmented into a single
+% segment.
 %
-First, \invariantconfluence{} is a very coarse-grained property. If an object
-is \invariantconfluent{}, then we can replicate it with no coordination. If it
-is not \invariantconfluent{}, then we have no guarantees. There's no
-in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
-more fine-grained property that can be applied to applications with varying
-degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
-complex applications that require a large number of segments and to simple
-applications with a smaller number of segments, whereas \invariantconfluence{}
-only provides guarantees to applications that can be segmented into a single
-segment.
-
-Second, while our segmented \invariantconfluence{} decision procedure can help
-decide whether or not an object is segmented \invariantconfluent{}, it cannot
-currently help construct a segmentation. It is the responsibility of the
-programmer to think of a segmentation that is amenable to segmented
-\invariantconfluence{}. This can be an onerous process. In the future, we
-plan to explore ways by which we can automatically suggest segmentations to the
-application designer to ease this process.
-
-\begin{techreport}
-
-  Third, segmented \invariantconfluence{} naturally subsumes a distributed
-  locking approach to replicating non-\invariantconfluent{} objects. This
-  approach first recognizes which transactions cannot be safely executed
-  concurrently and then requires them to acquire a distributed lock before
-  executing~\cite{balegas2015putting, gotsman2016cause}. For example, in a
-  banking application with the invariant that all balances remain non-negative,
-  concurrent deposits are permitted, but concurrent withdrawals can lead to
-  invariant violations. Thus, we require that withdrawals acquire a distributed
-  lock before executing. This example is exactly the same as
-  \exampleref{CounreachableExample} which we handled by simply removing
-  withdrawal transactions from our segmentation's set of transactions.
-
-  Fourth, we can integrate a couple of optimizations into our system model to
-  further reduce the amount of coordination it requires. To begin, if a server
-  with state $s_i$ and active segment $(I_i, T_i)$ receives a transaction $t
-  \in I_i$ to execute, and $t(s_i)$ violates the active invariant but not the
-  global invariant, instead of initiating a round of global coordination, $p_i$
-  can simply buffer $t$ for re-execution at a later time. While this increases
-  the latency required to execute $t$, it's possible that after other
-  transactions are executed, re-executing $t$ may lead to a state that either
-  satisfies the active invariant or violates the global invariant. In either
-  case, a round of global coordination is avoided. Similarly, servers can
-  buffer transactions that require global coordination, executing an entire
-  batch of these transactions during a single round of global coordination.
-
-  Fifth, a segmented \invariantconfluence{} decision procedure can also
-  leverage \thmref{LatticeProperty} in addition to
-  \algoref{ArbitraryStartInteractiveDecisionProcedure}. If an object $O$ meets
-  criteria (1) - (3), then it is \sticonfluent{s}{T}{I} for every state $s \in
-  I$.
-\end{techreport}
+% Second, while our segmented \invariantconfluence{} decision procedure can help
+% decide whether or not an object is segmented \invariantconfluent{}, it cannot
+% currently help construct a segmentation. It is the responsibility of the
+% programmer to think of a segmentation that is amenable to segmented
+% \invariantconfluence{}. This can be an onerous process. In the future, we
+% plan to explore ways by which we can automatically suggest segmentations to the
+% application designer to ease this process.

### 2020-05-17 doc/sigmod_paper/sections/related_work.tex
More aggressive cutting.
Need to polish now since I hastily chopped a bunch of stuff.

@@
 % \textbf{Branch and Merge.}
-Bayou~\cite{terry1995managing}, Dynamo~\cite{decandia2007dynamo}, and
-TARDiS~\cite{crooks2016tardis} all take a branch and merge approach to
-maintaining distributed invariants without coordination. With this approach,
-servers execute transactions without any coordination but keep track of the
-causal dependencies between transactions. Periodically, two servers merge
-states and invoke a user defined merge function to reconcile the divergent
-states. This approach does \emph{not} provide any formal guarantees that
-invariants are maintained.
+% Bayou~\cite{terry1995managing}, Dynamo~\cite{decandia2007dynamo}, and
+% TARDiS~\cite{crooks2016tardis} all take a branch and merge approach to
+% maintaining distributed invariants without coordination.
+% With this approach,
+% servers execute transactions without any coordination but keep track of the
+% causal dependencies between transactions. Periodically, two servers merge
+% states and invoke a user defined merge function to reconcile the divergent
+% states.
+% This approach does \emph{not} provide any formal guarantees that
+% invariants are maintained.
 % Its correctness depends on the correctness of the
@@
 % \textbf{CRDTs.}
-CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
-semilattices with inflationary update methods. Due to their algebraic
-properties, CRDTs can be replicated with strong eventual consistency without
-the need for any coordination. Our definition of distributed objects and our
-\invariantconfluence{} system model are inspired directly by the corresponding
-definitions and system models in~\cite{shapiro2011conflict}.
+% CRDTs~\cite{shapiro2011conflict, shapiro2011comprehensive} are distributed
+% semilattices with inflationary update methods. Due to their algebraic
+% properties, CRDTs can be replicated with strong eventual consistency without
+% the need for any coordination. Our definition of distributed objects and our
+% \invariantconfluence{} system model are inspired directly by the corresponding
+% definitions and system models in~\cite{shapiro2011conflict}.
 
@@
 % \invariantconfluence{}.
-
-\begin{techreport}
-
-  \textbf{Program Analysis in Database Systems.} Using program analysis to
-  improve the performance of database systems is not new. For example, it has
-  been used to improve the performance of database-backed web
-  applications~\cite{cheung2014using, wu2016transaction,
-  ramachandra2012program} and used to improve the performance of optimistic
-  concurrency control on multi-core machines~\cite{wu2016transaction}. Our work
-  on \invariantconfluence{} continues the theme of using program analysis to
-  reap the performance benefits gained from understanding program semantics.
-\end{techreport}

### 2020-05-17 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
More aggressive cutting.
Need to polish now since I hastily chopped a bunch of stuff.

@@
 
-\begin{example}\examplelabel{CounreachableExample}
-  Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps positive and negative integer counts (dubbed a
-  PN-Counter~\cite{shapiro2011comprehensive}), replicated on three machines.
-  Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
-  integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment or decrement
-  the counter, the $i$th server increments $p_i$ or $n_i$ respectively, and
-  $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
-  (0, 0, 0)$; our set $T$ of transactions consists of increment and decrement;
-  and our invariant $I$ is that the value of $s$ is non-negative.
-
-  Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
-  with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
-  1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
-  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
-  concurrent decrements, which we can forbid in a simple one-segment
-  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
-  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
-  \IsIclosed{} again returns false with the same states $s_1$ and $s_2$. This time,
-  however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I} (all modifications of $(n_1, n_2, n_3)$ require global
-  coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
-  values).
-  The user refines $NR$ with the observation that two
-  states are coreachable if and only if they have the same values of $n_1, n_2,
-  n_3$. After this, \IsIclosed{} and
-  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
-\end{example}
+% \begin{example}\examplelabel{CounreachableExample}
+%   Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps positive and negative integer counts (dubbed a
+%   PN-Counter~\cite{shapiro2011comprehensive}), replicated on three machines.
+%   Every state $s = $ $(p_1, p_2, p_3),$ $(n_1, n_2, n_3)$ represents the
+%   integer $(p_1 + p_2 + p_3) - (n_1 + n_2 + n_3)$. To increment or decrement
+%   the counter, the $i$th server increments $p_i$ or $n_i$ respectively, and
+%   $\join$ computes an element-wise maximum. Our start state $s_0 = (0, 0, 0),
+%   (0, 0, 0)$; our set $T$ of transactions consists of increment and decrement;
+%   and our invariant $I$ is that the value of $s$ is non-negative.
+%
+%   Applying \algoref{InteractiveDecisionProcedure}, \IsIclosed{} returns false
+%   with the states $s_1 = (1, 0, 0), (0, 1, 0)$ and $s_2 = (1, 0, 0), (0, 0,
+%   1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
+%   \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
+%   concurrent decrements, which we can forbid in a simple one-segment
+%   segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
+%   transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
+%   \IsIclosed{} again returns false with the same states $s_1$ and $s_2$. This time,
+%   however, the user recognizes that the two states are not
+%   \ticoreachable{T^+}{I} (all modifications of $(n_1, n_2, n_3)$ require global
+%   coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
+%   values).
+%   The user refines $NR$ with the observation that two
+%   states are coreachable if and only if they have the same values of $n_1, n_2,
+%   n_3$. After this, \IsIclosed{} and
+%   \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
+% \end{example}
 

### 2020-05-17 doc/sigmod_paper/sections/introduction.tex
Polished up paper, now 8 pages.

@@
 transactions. Lucy processes these workloads in less than half a second, and
-neither workload requires more than 66 lines of code to specify. Moreover, we
-find that segmented \invariantconfluent{} replication can achieve 10x to 100x
-more throughput than linearizable replication for low-coordination workloads.
+no workload requires more than 66 lines of code to specify.
+% Moreover, we find that segmented \invariantconfluent{} replication can achieve
+% 10x to 100x more throughput than linearizable replication for low-coordination
+% workloads.
 

### 2020-05-17 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
Polished up paper, now 8 pages.

--- 2020-05-16 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
@@
 invariants of non-\invariantconfluent{} objects, requiring only a small amount
-of coordination. In \secref{Evaluation}, we see that replicating a
-non-\invariantconfluent{} object with segmented \invariantconfluence{} can
-achieve between 10x and 100x more throughput than linearizable replication for
-certain workloads.
+of coordination.
+% In \secref{Evaluation}, we see that replicating a non-\invariantconfluent{}
+% object with segmented \invariantconfluence{} can achieve between 10x and 100x
+% more throughput than linearizable replication for certain workloads.
 
@@
 
-% The proof of the forward direction is exactly the same as the proof of
-% \thmref{IclosureImpliesIconfluence}. Transactions always maintain the
-% invariant, so if merge does as well, then every reachable state must satisfy
-% the invariant.
-% %
-% For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
-% The two points are \TIcoreachable{}, so there exists some state $s_0$ from
-% which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
-% \sTIreachable{}, so it satisfies the invariant.
+The proof of the forward direction is exactly the same as the proof of
+\thmref{IclosureImpliesIconfluence}. Transactions always maintain the
+invariant, so if merge does as well, then every reachable state must satisfy
+the invariant.
+%
+For the reverse direction, consider two arbitrary states $s_1, s_2 \in I$.
+The two points are \TIcoreachable{}, so there exists some state $s_0$ from
+which they can be reached. $O$ is \sTIconfluent{} and $s_1 \join s_2$ is
+\sTIreachable{}, so it satisfies the invariant.
 
@@
 \begin{example}\examplelabel{CounreachableExample}
-  Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps positive and negative integer counts (dubbed a
+  Let $O = (\ints^3 \times \ints^3, \join)$ be an object that separately keeps
+  positive and negative integer counts (dubbed a
   PN-Counter~\cite{shapiro2011comprehensive}), replicated on three machines.
@@
   however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I} (all modifications of $(n_1, n_2, n_3)$ require global
-  coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
-  values).
+  \ticoreachable{T^+}{I}.
+  % (all modifications of $(n_1, n_2, n_3)$ require global
+  % coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
+  % values).
   The user refines $NR$ with the observation that two

### 2020-05-22 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
Joe feedback.

@@
 set of transitions $T$, and an invariant $I$. A segmentation $\Sigma = (I_1,
-T_1), \ldots, (I_n, T_n)$ is a sequence (not a set) of $n$ segments $(I_i,
-T_i)$ where every $T_i$ is a subset of $T$ and every $I_i \subseteq S$ is an
-invariant. $O$ is \defword{segmented \invariantconfluent{}} with respect to
-$s_0$, $T$, $I$, and $\Sigma$, abbreviated \defword{\sTISconfluent}, if the
-following conditions hold:
+T_1), \ldots, (I_n, T_n)$ is a sequence of $n$ segments $(I_i, T_i)$ where
+every $T_i$ is a subset of $T$ and every $I_i \subseteq S$ is an invariant.
+Note that $\Sigma$ is a sequence, not a set. The reason for this will become
+clear in the next subsection. $O$ is \defword{segmented \invariantconfluent{}}
+with respect to $s_0$, $T$, $I$, and $\Sigma$, abbreviated
+\defword{\sTISconfluent}, if the following conditions hold:
 \begin{itemize}
@@
 guaranteed to exist because the segment invariants cover $I$. Moreover,
-$\Sigma$ is ordered, so every server will deterministically pick the same
-active segment. In fact, an invariant of the system model is that at any given
-point of normal processing, every server has the same active segment.
+$\Sigma$ is ordered, as described in the previous subsection, so every server
+will deterministically pick the same active segment. In fact, an invariant of
+the system model is that at any given point of normal processing, every server
+has the same active segment.
 
@@
   segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
-  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
+  transactions. Applying \algoref{ArbitraryStartInteractiveDecisionProcedure},
   \IsIclosed{} again returns false with the same states $s_1$ and $s_2$. This time,

### 2020-05-22 doc/sigmod_paper/sections/abstract.tex
Polished up paper, now 8 pages.
Cleaned up commented out text.

@@
 to replicate non-\invariantconfluent{} objects with a small amount of
-coordination.
-%
-We implemented these formalisms in a prototype called Lucy and found that our
-decision procedures efficiently handle common real-world workloads including
-%
-% foreign keys, rollups, escrow transactions, and more.
-foreign keys, escrow transactions, and more.
-%
-We also found that segmented \invariantconfluent{} replication can deliver up
-to an order of magnitude more throughput than linearizable replication.
-% for low contention workloads and comparable throughput for medium to high
-% contention workloads.
+coordination. We implemented these formalisms in a prototype called Lucy and
+found that our decision procedures efficiently handle common real-world
+workloads including foreign keys, escrow transactions, and more.
 \end{abstract}

### 2020-05-22 doc/sigmod_paper/sections/conclusion.tex
Cleaned up commented out text.

--- 2020-05-16 doc/sigmod_paper/sections/conclusion.tex
@@
 \section{Conclusion}
-This paper revolved around two major contributions.
+This paper revolved around two major contributions. First, we found that
+\invariantclosure{} fails to incorporate a notion of reachability, and using
+this intuition, we developed conditions under which \invariantclosure{} and
+\invariantconfluence{} are equivalent. We implemented this insight in an
+interactive \invariantconfluence{} decision procedure that automatically checks
+whether an object is \invariantconfluent{}, with the assistance of a
+programmer.
 %
-First, we developed a deeper understanding of \invariantclosure{} and
-\invariantconfluence{} by looking at the two criteria with reachability in
-mind. We found that \invariantclosure{} fails to incorporate a notion of
-reachability, and using this intuition, we developed conditions under which
-\invariantclosure{} and \invariantconfluence{} are equivalent. We implemented
-this insight in an interactive \invariantconfluence{} decision procedure
-that automatically checks whether an object is \invariantconfluent{}, with the
-assistance of a programmer.
-
 Second, we proposed a new consistency model and generalization of
@@
 replicate non-\invariantconfluent{} objects with a small amount of coordination
-while still preserving their invariants. We found that segmented
-\invariantconfluence{} naturally subsumes existing techniques for maintaining
-invariants of replicated objects (e.g.\ locking and escrow transactions), and
-we developed an interactive decision procedure for segmented
-\invariantconfluence{}.
-
-Through our evaluation, we found that our decision procedures could analyze a
-number of realistic workloads, each in less than a second. We also showed that
-segmented \invariantconfluence{} can significantly outperform linearizable
-replication for low-coordination workloads.
+while still preserving their invariants.

### 2020-05-22 doc/sigmod_paper/sections/invariant_confluence.tex
Joe feedback.
Cleaned up commented out text.

@@
 
-% Note that by modelling a transaction $t$ as a function $S \to S$, we focus
-% exclusively on the effects that a transaction has on the object (i.e.\
-% ``writes'' to the object). Transactions are also free to read the value of the
-% object, but these reads are not captured by our model because, as we'll see,
-% they do not affect \invariantconfluence{}. For example, we could model any
-% read-only transaction as a function $t$ where $t(s) = s$ for every $s \in S$.
-
 In our system model, a distributed object $O$ is replicated across a set $p_1,
@@
 s_j$. Servers periodically merge states with one another in order to keep their
-states loosely synchronized.
-% \footnote{%
-%   Notably, if $O$ is a CRDT---i.e. $O$ is a semilattice and every transaction
-%   $t \in T$ is inflationary---then this periodic merging ensures that $O$ is
-%   strongly eventually consistent~\cite{shapiro2011conflict}.
-% }.
-Note that unlike with transactions, servers \emph{cannot} abort a merge; server
-$p_j$ must transition from $s_j$ to $s_i \join s_j$ whether or not $s_i \join
-s_j$ satisfies the invariant.
+states loosely synchronized. Note that unlike with transactions, servers
+\emph{cannot} abort a merge; server $p_j$ must transition from $s_j$ to $s_i
+\join s_j$ whether or not $s_i \join s_j$ satisfies the invariant.
 
@@
 valid execution of our system model. Formally, we define
-$\sTIreachablepredicate{e}$ to be the smallest predicate that satisfies the
-following equations:
+$\sTIreachablepredicate{e}$ to be the predicate that satisfies the
+following conditions:
 \begin{itemize}

### 2020-05-22 doc/sigmod_paper/sections/abstract.tex
Small fixes.

@@
 to replicate non-\invariantconfluent{} objects with a small amount of
-coordination. We implemented these formalisms in a prototype called Lucy and
-found that our decision procedures efficiently handle common real-world
+coordination. We implement these formalisms in a prototype called Lucy and
+find that our decision procedures efficiently handle common real-world
 workloads including foreign keys, escrow transactions, and more.

### 2020-05-22 doc/sigmod_paper/sections/introduction.tex
Cleaned up commented out text.
Small fixes.

@@
 make a fundamental choice between weak and strong consistency. Replicating the
-data with strong consistency using a technique like distributed transactions
-(e.g., \cite{bernstein1981concurrency,mohan1986transaction}) or state machine
-replication (e.g., \cite{schneider1990implementing, lamport1998part,
-liskov2012viewstamped, ongaro2014search}) makes the application designer's life
+data with strong consistency using a technique like distributed
+transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state
+machine replication~\cite{schneider1990implementing, lamport1998part,
+liskov2012viewstamped, ongaro2014search} makes the application designer's life
 very easy. To the developer, a strongly consistent system behaves exactly like
@@
 proofs of this sort are unreasonable to expect from programmers. Ideally we
-would have a general-purpose program that can automatically determine
+would have a general-purpose program that could automatically determine
 \invariantconfluence{} for us.  \textbf{The first main thrust of this paper is
@@
 In this paper, we take a step back and study the underlying reason \emph{why}
-\invariantclosure{} is not necessary for \invariantconfluence{}.
-%
-% We find that \invariantconfluence{} is fundamentally a reachability property,
-% checking whether certain values of our replicated object can actually be
-% obtained in the execution of a system. We discover that unreachable, yet
-% invariant satisfying, states cause \invariantclosure{} to be an unnecessary
-% condition.
-%
-Using this understanding, we construct a set of modest conditions under which
+\invariantclosure{} is not necessary for \invariantconfluence{}. Using this
+understanding, we construct a set of modest conditions under which
 \invariantclosure{} and \invariantconfluence{} are in fact \emph{equivalent},
 allowing us to reduce the problem of determining \invariantconfluence{} to that
-of determining \invariantclosure{}.
-%
-Then, we use these conditions to design a general-purpose interactive
-\invariantconfluence{} decision procedure.
-
-% Then, we use these conditions to design a general-purpose interactive
-% \invariantconfluence{} decision procedure. Users iteratively interact with the
-% decision procedure, eliminating unreachable states from the invariant, moving
-% towards the conditions under which \invariantclosure{} and
-% \invariantconfluence{} are equivalent. Meanwhile, the decision procedure
-% generates examples of reachable and unreachable states which help the user
-% recognize patterns that describe reachability. We also leverage our intuitions
-% about reachability to develop a new sufficient condition for
-% \invariantconfluence{}, dubbed \emph{\mergereducibility}. \Mergereducibility{}
-% covers some cases that are not covered by \invariantclosure{}, and it can be
-% automatically checked without any user interaction.
+of determining \invariantclosure{}. Then, we use these conditions to design a
+general-purpose interactive \invariantconfluence{} decision procedure.
 
@@
 \secref{Evaluation}, we apply Lucy to foreign key constraints and escrow
-transactions. Lucy processes these workloads in less than half a second, and
-no workload requires more than 66 lines of code to specify.
-% Moreover, we find that segmented \invariantconfluent{} replication can achieve
-% 10x to 100x more throughput than linearizable replication for low-coordination
-% workloads.
-
-% In closing, here is an outline of the paper and of our contributions:
-% %
-% We propose a novel expression-oriented definition of \invariantconfluence{}
-% that is both formal and simple (\secref{InvariantConfluence}).
-% %
-% We develop an understanding of why \invariantclosure{} is not necessary for
-% \invariantconfluence{} and use this understanding to develop conditions under
-% which it is both necessary and sufficient (\secref{InvariantClosure}).
-% %
-% We exploit these conditions to design a general-purpose interactive decision
-% procedure for \invariantconfluence{} (\secref{InteractiveDecisionProcedure}).
-% %
-% We again exploit these conditions to design a novel non-trivial sufficient
-% condition for \invariantconfluence{}, called \mergereducibility.
-% %
-% We present segmented \invariantconfluence{}: a generalization of
-% \invariantconfluence{} that uses a small amount of coordination to maintain
-% invariants for replicated objects that are otherwise not \invariantconfluent{}
-% (\secref{SegmentedInvariantConfluence}).
-% %
-% We evaluate our methods using a prototype implementation called Lucy
-% (\secref{Evaluation}).
+transactions. Lucy processes these workloads in less than half a second, and no
+workload requires more than 66 lines of code to specify.

### 2020-06-07 doc/sigmod_paper/sections/introduction.tex
Crunched paper.
Got rid of some superfluous citations and got rid of some semi-obvious
pseudocode.

@@
 data with strong consistency using a technique like distributed
-transactions~\cite{bernstein1981concurrency,mohan1986transaction} or state
-machine replication~\cite{schneider1990implementing, lamport1998part,
-liskov2012viewstamped, ongaro2014search} makes the application designer's life
+transactions~\cite{bernstein1981concurrency} or state
+machine replication~\cite{lamport1998part} makes the application designer's life
 very easy. To the developer, a strongly consistent system behaves exactly like
@@
 most effective tools that is used to reason about the execution of programs:
-application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
+application invariants~\cite{balegas2015towards} such as
 database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even

### 2020-06-07 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
Cleaned up commented out text.
Crunched paper.
Got rid of some superfluous citations and got rid of some semi-obvious
pseudocode.

@@
 of coordination.
-% In \secref{Evaluation}, we see that replicating a non-\invariantconfluent{}
-% object with segmented \invariantconfluence{} can achieve between 10x and 100x
-% more throughput than linearizable replication for certain workloads.
 
@@
 state $s_i$ and initiates a round of global coordination to execute $t$, as
-described in the previous paragraph.  Transaction execution is summarized in
-\algoref{TxnExecution}.
+described in the previous paragraph.
+%
+% Transaction execution is summarized in \algoref{TxnExecution}.
 
-\begin{algorithm}[t]
-  \caption{%
-    Transaction execution in the segmented \invariantconfluence{} system model}%
-  \algolabel{TxnExecution}
-  \begin{algorithmic}
-    \If{$t \notin T_i$}
-      \State Execute $t$ with global coordination
-    \Else{}
-      \If{$I_i(t(s_i))$} Commit $t$
-      \ElsIf{$\lnot I(t(s_i))$} Abort $t$
-      \Else{} Execute $t$ with global coordination
-      \EndIf{}
-    \EndIf{}
-  \end{algorithmic}
-\end{algorithm}
+% \begin{algorithm}[t]
+%   \caption{%
+%     Transaction execution in the segmented \invariantconfluence{} system model}%
+%   \algolabel{TxnExecution}
+%   \begin{algorithmic}
+%     \If{$t \notin T_i$}
+%       \State Execute $t$ with global coordination
+%     \Else{}
+%       \If{$I_i(t(s_i))$} Commit $t$
+%       \ElsIf{$\lnot I(t(s_i))$} Abort $t$
+%       \Else{} Execute $t$ with global coordination
+%       \EndIf{}
+%     \EndIf{}
+%   \end{algorithmic}
+% \end{algorithm}
 
@@
 set of \emph{coreachable} states, but otherwise, the core of the two algorithms
-is the same.
-% \footnote{%
-%   Another small difference is that \IsIclosed{} behaves differently in
-%   \algoref{InteractiveDecisionProcedure} and
-%   \algoref{ArbitraryStartInteractiveDecisionProcedure}. In
-%   \algoref{ArbitraryStartInteractiveDecisionProcedure}, \IsIclosed{} returns a
-%   triple (closed, $s_1$, $s_2$). If closed is false, then $s_1, s_2 \in I$
-%   are two states not in $NR$ such that $I(s_1)$ and $I(s_2)$ but $\lnot I(s_1
-%   \join s_2)$. If no such states exist, then closed is true, and $s_1$ and
-%   $s_2$ are null.
-% }
-Now, a segmented \invariantconfluence{} decision procedure, can simply invoke
-\algoref{ArbitraryStartInteractiveDecisionProcedure} once on each segment.
+is the same. Now, a segmented \invariantconfluence{} decision procedure, can
+simply invoke \algoref{ArbitraryStartInteractiveDecisionProcedure} once on each
+segment.
 
@@
   however, the user recognizes that the two states are not
-  \ticoreachable{T^+}{I}.
-  % (all modifications of $(n_1, n_2, n_3)$ require global
-  % coordination, so it is impossible for $s_1$ and $s_2$ to differ on these
-  % values).
-  The user refines $NR$ with the observation that two
+  \ticoreachable{T^+}{I}. The user refines $NR$ with the observation that two
   states are coreachable if and only if they have the same values of $n_1, n_2,
@@
 \end{example}
-
-% \subsection{Discussion and Limitations}
-% There are a few things worth noting about segmented \invariantconfluence{}, its
-% system model, and its decision procedure.
-% %
-% First, \invariantconfluence{} is a very coarse-grained property. If an object
-% is \invariantconfluent{}, then we can replicate it with no coordination. If it
-% is not \invariantconfluent{}, then we have no guarantees. There's no
-% in-between.  Segmented \invariantconfluence{}, on the other hand, is a much
-% more fine-grained property that can be applied to applications with varying
-% degrees of complexity. Segmented \invariantconfluence{} provides guarantees to
-% complex applications that require a large number of segments and to simple
-% applications with a smaller number of segments, whereas \invariantconfluence{}
-% only provides guarantees to applications that can be segmented into a single
-% segment.
-%
-% Second, while our segmented \invariantconfluence{} decision procedure can help
-% decide whether or not an object is segmented \invariantconfluent{}, it cannot
-% currently help construct a segmentation. It is the responsibility of the
-% programmer to think of a segmentation that is amenable to segmented
-% \invariantconfluence{}. This can be an onerous process. In the future, we
-% plan to explore ways by which we can automatically suggest segmentations to the
-% application designer to ease this process.

### 2020-06-10 doc/sigmod_paper/sections/introduction.tex
Joe feedback edits + crunch for space.

@@
 latency at best and unavailability at worst~\cite{gilbert2002brewer,
-brewer2012cap, abadi2012consistency}. On the other hand, weak consistency
+abadi2012consistency}. On the other hand, weak consistency
 models like eventual consistency~\cite{vogels2009eventually}, PRAM
@@
 application invariants~\cite{balegas2015towards} such as
-database integrity constraints~\cite{grefen1993integrity, gupta1993local}. Even
+database integrity constraints~\cite{grefen1993integrity}. Even
 if every transaction executing in a weakly consistent system individually

### 2020-06-14 doc/sigmod_paper/sections/interactive_decision_procedure.tex
Joe feedback.
Cleaned up commented out text.
Got rid of orphaned text.

@@
 closed is a boolean indicating whether $O$ is \Iclosed{}. If closed is true,
-then $s_1$ and $s_2$ are null. If closed is false, then $s_1$ and $s_2$ are a
-counterexample witnessing the fact that $O$ is not \Iclosed{}. That is,
-$I(s_1)$ and $I(s_2)$, but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from
-\exampleref{Z2}).  As we mentioned earlier, we can (and do) implement the
-\invariantclosure{} decision procedure using an SMT solver like
-Z3~\cite{de2008z3}.
+then $s_1$ and $s_2$ are null. Otherwise, $s_1$ and $s_2$ are a counterexample
+witnessing the fact that $O$ is not \Iclosed{}. That is, $I(s_1)$ and $I(s_2)$,
+but $\lnot I(s_1 \join s_2)$ (e.g., $s_1$ and $s_2$ from \exampleref{Z2}).  As
+we mentioned earlier, we can (and do) implement the \invariantclosure{}
+decision procedure using an SMT solver like Z3~\cite{de2008z3}.
 
@@
 First, we can randomly generate a set of reachable states and add them to $R$.
-If $s_1$ or $s_2$ is in $R$, then we know they are reachable.
+If $s_1$ or $s_2$ is in $R$, then they are reachable.
 %
-Second, we can prompt the user to tell us directly whether or not the states
+Second, we can prompt the user to tell us directly whether the states
 are reachable or unreachable.
@@
 can also refine $I$ by augmenting $R$ and $NR$ arbitrarily (see
-\figref{DecisionProcedureExample} for an example). In this step, we also make
+\figref{DecisionProcedureExample} for example). In this step, we also make
 sure that $s_0 \notin NR$ since we know that $s_0$ is reachable.
@@
 augment $NR$ and eliminate a large number of unreachable states from the
-invariant. Once $NR$ has been sufficiently augmented to the point that $I - NR$
-is a subset of the reachable states, \thmref{IclosureEquivalentIconfluence}
-guarantees that the algorithm will terminate after one more call to \IsIclosed.
-
-% \subsection{Limitations}
-% Our interactive \invariantconfluence{} decision procedure has two
-% limitations. First, \algoref{InteractiveDecisionProcedure} requires an
-% \invariantclosure{} decision procedure, but determining \invariantclosure{} is
-% undecidable in general.
-% %
-% In practice, we can implement an \invariantclosure{} decision procedure using
-% an SMT solver like Z3 that works well on simple objects, invariants, and merge
-% operators (e.g., integers, tuples, infinite sets, bit vectors, linear
-% constraints, basic arithmetic, tuple projection, basic set operations, bit
-% arithmetic). But, SMT solvers are mostly unable to analyze more complex
-% constructs (e.g., finite lists~\cite{kroning2009proposal}, graphs, recursive
-% algebraic data types, nonlinear constraints, merge operators that contain loops
-% or recursion).
-%
-% Second, \algoref{InteractiveDecisionProcedure} relies on a user to identify
-% unreachable states. As we saw in \figref{DecisionProcedureExample}, the set of
-% unreachable states can sometimes be clear, especially if there's a noticeable
-% pattern in the set of reachable states. However, if the set of transactions is
-% large or complex or if the merge operator is complex, then reasoning about
-% unreachable states can be difficult.  Unlike with reachable states---where
-% verifying that a state is reachable only requires thinking of a single way to
-% reach the state---verifying that a state is unreachable requires a programmer
-% to reason about a large number of system executions and conclude that
-% \emph{none} of them can lead to the state. In the future, we plan on exploring
-% ways to help a user reason about unreachable states and ways to discover sets of
-% unreachable states automatically.
+invariant. See \figref{DecisionProcedureExample}, for example. Once $NR$ has
+been sufficiently augmented to the point that $I - NR$ is a subset of the
+reachable states, \thmref{IclosureEquivalentIconfluence} guarantees that the
+algorithm will terminate after one more call to \IsIclosed.

### 2020-06-14 doc/sigmod_paper/sections/introduction.tex
Got rid of orphaned text.

@@
 most effective tools that is used to reason about the execution of programs:
-application invariants~\cite{balegas2015towards} such as
+application invariants~\cite{hoare1969axiomatic, balegas2015towards} such as
 database integrity constraints~\cite{grefen1993integrity}. Even
@@
 coordination-freedom; more complex applications require more segments which
-require more coordination, and vice-versa.
+require more coordination.
 

### 2020-06-14 doc/sigmod_paper/sections/invariant_closure.tex
First round of cuts.
The paper is now 8 pages without citations. We need 8 pages with
citations, so we need to trim another page or so. I'll probably shrink
the evaluation a bit more, shrink the related works a bit more, shrink
segmented invariant confluence a bit more and see what we end up with.
Got rid of orphaned text.

--- 2018-05-01 doc/paper/sections/invariant_closure.tex
@@
 \thmref{IclosureImpliesIconfluence} states that \invariantclosure{} is
-sufficient for \invariantconfluence{}. Intuitively, recall that our system model ensures
+sufficient for \invariantconfluence{}. Intuitively, our system model ensures
 that transaction execution preserves the invariant, so if merging states also
@@
   where $(x_1, y_1) \join (x_2, y_2) = (\max(x_1, x_2), \max(y_1, y_2))$. Our
-  start state $s_0 \in \ints \times \ints$ is the point $(0, 0)$. Our set $T$
-  of transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1, y)$
+  start state $s_0 \in \ints \times \ints$ is $(0, 0)$. Our set $T$ of
+  transactions consists of two transactions: $t_{x+1}((x, y)) = (x + 1, y)$
   which increments $x$ and $t_{y-1}((x, y)) = (x, y - 1)$ which decrements $y$.
@@
 In \exampleref{Z2}, $s_1$ and $s_2$ witness the fact that $O$ is not
-\invariantclosed{}, but $s_1$ is not reachable. This is not particular to
+\invariantclosed{}, but $s_1$ is \emph{not reachable}. This is not particular to
 \exampleref{Z2}. In fact, it is fundamentally the reason why
@@
 \end{theorem}
-\begin{proof}
+
 The forward direction of \thmref{IclosureEquivalentIconfluence} follows
@@
 \join s_2$, satisfy the invariant.
-\end{proof}

### 2020-06-14 doc/sigmod_paper/sections/invariant_confluence.tex
Got rid of orphaned text.

@@
 We begin by introducing our system model of replicated objects in which a
-distributed object and accompanying invariant is replicated across a set of
-servers. Clients send transactions to servers, and a server executes a
-transaction so long as it maintains the invariant. Servers execute
+distributed object and an invariant are replicated across a set of
+servers. Clients send transactions to servers, and servers execute
+transactions so long as they maintain the invariant. Servers execute
 transactions without coordination, but to avoid state divergence, servers
@@
 %
-After we introduce the system model, we present a formal definition of
-\invariantconfluence{}.
+% After we introduce the system model, we present a formal definition of
+% \invariantconfluence{}.
 
@@
 
+\vfill\null
+
 \begin{example}\examplelabel{Ints}
@@
 \ldots, p_n$ of $n$ servers. Each server $p_i$ manages a replica $s_i \in S$ of
-the replicated object. Every server begins with a start state $s_0 \in S$, a
+the object. Every server begins with a start state $s_0 \in S$, a
 fixed set $T$ of transactions, and an invariant $I$. Servers repeatedly perform
@@
 following conditions:
+\pagebreak
 \begin{itemize}
@@
   \item
-    For all expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
+    For expressions $e_1$ and $e_2$, if $\sTIreachablepredicate{e_1}$ and
     $\sTIreachablepredicate{e_2}$, then $\sTIreachablepredicate{e_1 \join
@@
 \end{itemize}
+
 Similarly, we say a state $s \in S$ is \sTIreachable{} if there exists an

### 2020-06-14 doc/sigmod_paper/sections/related_work.tex
Polished up paper, now 8 pages.
Joe feedback.
Got rid of orphaned text.

@@
 \newcommand{\sieve}{\textsc{SIEVE}}
-% \textbf{RedBlue Consistency and \sieve.}
-RedBlue consistency is a consistency model that sits between causal consistency
-and linearizability~\cite{li2012making}.
+RedBlue consistency~\cite{li2012making}, is a consistency model that sits
+between causal consistency and linearizability.
 %
@@
 deciding invariant safety that involve calculating weakest preconditions.
-% These
-% techniques are complementary to our work and can be used to improve the
-% \invariantclosure{} subroutine used by our decision procedures.
+These techniques are complementary to our work and can be used to improve the
+\invariantclosure{} subroutine used by our decision procedures.
 %
@@
 used to replicate workloads that do not meet their sufficient conditions.
-Segmented \invariantconfluence{} is a general-purpose formalism that can be
+Segmented \invariantconfluence{} is a formalism that can be
 used to model simple forms of these techniques.
@@
 specify certain conflicts that are not possible with segmented
-\invariantconfluence{} because a segmentation only allows transactions within a
-segment to acquire a single self-conflicting lock. However, segmented
+\invariantconfluence{}.
+% because a segmentation only allows transactions within a
+% segment to acquire a single self-conflicting lock.
+However, segmented
 \invariantconfluence{} also introduces the notion of invariant segmentation,
 which cannot be emulated with the token based approach.
-% For example, it is
-% difficult to emulate escrow transactions with the token based approach.
+For example, it is
+difficult to emulate escrow transactions with the token based approach.
 

### 2020-06-14 doc/sigmod_paper/sections/segmented_invariant_confluence.tex
Got rid of orphaned text.

@@
 %
-Transaction execution in the new system model, on the other hand, is a bit more
+Transaction execution in the new system model, on the other hand, is more
 involved. Consider a server $s_i$ with active segment $(I_i, T_i)$. A client
-can request that $p_i$ execute a transaction $t \in T$. We consider what
-happens when $t \in T_i$ and $t \notin T_i$ separately.
+can request that $p_i$ execute a transaction $t$. We consider what happens when
+$t \in T_i$ and when $t \notin T_i$.
 
@@
 then generalize \thmref{IclosureEquivalentIconfluence}. Two states $s_1, s_2
-\in I$ are \defword{coreachable} with respect to $T$ and $I$, abbreviated
-\defword{\TIcoreachable}, if there exists some state $s_0 \in I$ such that
-$s_1$ and $s_2$ are both \sTIreachable{}.
+\in I$ are \defword{coreachable} with respect to a set of transactions $T$ and
+an invariant $I$, abbreviated \defword{\TIcoreachable}, if there exists some
+state $s_0 \in I$ such that $s_1$ and $s_2$ are both \sTIreachable{}.
 
@@
   states are coreachable if and only if they have the same values of $n_1, n_2,
-  n_3$. After this, \IsIclosed{} and
-  \algoref{ArbitraryStartInteractiveDecisionProcedure} return true.
+  n_3$. After this, \IsIclosed{} (and thus \textsc{Helper}) returns true, and
+  \algoref{ArbitraryStartInteractiveDecisionProcedure} terminates.
 \end{example}

### 2020-08-06 doc/the_vldb_journal_paper/rebuttal.tex
Put in reviewer feedback.

\documentclass[9pt]{article}

\usepackage[hidelinks]{hyperref}
\usepackage{enumitem}
\usepackage{float}
\usepackage{multicol}
\usepackage{pervasives}
\usepackage{vldblite}

% Reviewer quotes.
\definecolor{ReviewerDarkGray}{HTML}{37474F}
\newenvironment{reviewerquote}
{\list{}{\leftmargin=\parindent\rightmargin=0in}\item[] \itshape \color{ReviewerDarkGray}}%
{\endlist}

\newtoggle{showfeedback}
\toggletrue{showfeedback}
% \togglefalse{showfeedback}
\NewEnviron{feedback}{
  \iftoggle{showfeedback}{
    \begin{itemize}[leftmargin=*]
      \item[] \itshape \textcolor{ReviewerDarkGray}{\BODY}
    \end{itemize}
  }{}
}

% Turn on the full reviewer comments if the turnonshowfeedback flag was passed
% in from the command line. This was taken from [1].
%
% [1]: https://stackoverflow.com/a/1466610/3187068
% \ifdefined\turnonshowfeedback
%   \toggletrue{showfeedback}
% \else
%   \togglefalse{showfeedback}
% \fi

\begin{document}
\begin{center}
  \ttlfnt Response to Reviewer Comments
\end{center}

\begin{multicols*}{2}
Below, you'll find our response to reviewer comments and our revisions. All
changes made to the paper are shown in \markrevisions{pink}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 1}
\paragraph{Reviewer 1: W1}
\begin{feedback}
  More details about the DSL to specify the invariants with more complex
  examples would add a lot of value to the paper
\end{feedback}
\paragraph{Reviewer 1: W2}
\begin{feedback}
  More experiments with some realistic workloads (like TPCC) to show the
  benefit of the proposed concepts in action are needed
\end{feedback}
\paragraph{Reviewer 1: W3}
\begin{feedback}
  Even though the automatic segmentation is left for future work, an example
  with a manual mapping of segments in a realistic workload (like TPCC) would
  show-case its potential
\end{feedback}
\paragraph{Reviewer 1: W4}
\begin{feedback}
  The main weak point of this specific submission is whether it has enough new
  material to to qualify for being an extended version
\end{feedback}
\paragraph{Reviewer 1: D1}
\begin{feedback}
  W4 touches upon the "new material" for this extended version of the paper,
  and the above weak points W1, W2 and W3 can all serve as additional material.
\end{feedback}
\paragraph{Reviewer 1: D2}
\begin{feedback}
  W1 will essentially help the reader to quickly see the value of the proposed
  definitions and theorems with some specification examples that will be able
  to capture more interesting use-cases.
\end{feedback}
\paragraph{Reviewer 1: D3}
\begin{feedback}
  W2 would help to quantify the benefits that come from being able to have some
  invariant confluent objects in an otherwise strongly consistent setting.
\end{feedback}
\paragraph{Reviewer 1: D4}
\begin{feedback}
  W3 would really help to showcase the value of segmentation in practice.
  Ideally, we want to see that in a specific application there is a logical
  mapping between segments and data, and then use this work as an example for
  potential future work that will be fully automated as the authors suggest.
\end{feedback}
\paragraph{Reviewer 1: D5}
\begin{feedback}
  In fact, this automation would definitely strength the paper dramatically if
  it were added (but it is not required for this extended version to be enough)
\end{feedback}
\paragraph{Reviewer 1: D6}
\begin{feedback}
  It is clear that the authors added more material to the journal paper,
  however, it does not substantially augment the messages of the paper --
  addressing W1-W3 will help in that direction (for example section 7 ends
  without any enough discussion of the practical connections of the proposed
  idea)
\end{feedback}
\paragraph{Reviewer 1: D7}
\begin{feedback}
  In the intro we read about "the trade-off between application complexity and
  coordination-freedom", it would be great to see some practical examples later
  in the paper
\end{feedback}
\paragraph{Reviewer 1: D8}
\begin{feedback}
  In section 4.2, more details about what can be done with respect to help a
  user reason about unreachable states would be great
\end{feedback}
\paragraph{Reviewer 1: D9}
\begin{feedback}
  Figure 5 should be moved before figure 6
\end{feedback}
\paragraph{Reviewer 1: D10}
\begin{feedback}
  in page 3 we read "The culprit is concurrent decrements, which we can forbid
  in a simple one-segment segmentation ...", this sentence does not parse well
\end{feedback}
\paragraph{Reviewer 1: D11}
\begin{feedback}
  In 6.4 we read "This example is exactly the same as Example 8 which we
  handled by simply removing withdrawal transactions from our segmentation's
  set of transactions." How can one identify transactions to remove, and is the
  removal practically done?
\end{feedback}
\paragraph{Reviewer 1: D12}
\begin{feedback}
  In 8.1 we read "When a server receives a coordination request from the
  leader, it stops processing transactions and sends the leader its state."
  what happens to the in-flight transactions? Are the cancelled or allowed to
  finish? Is the system simply not admitting new transactions?
\end{feedback}

\paragraph{Reviewer 1: D13}
\begin{feedback}
  In 8.1 again, at the end, more discussion about the tradeoff of
  fault-tolerance would be interesting.
\end{feedback}

\paragraph{Reviewer 1: Typos}
\begin{feedback}
  end of first column of page 3 "that it execute a transaction" --> "that it
  executes a transaction"

  section 5, in example 3, "Our start state state" --> two times state

  section 8.2 in example 12: "and a optional" --> "and an optional"
\end{feedback}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 3}
\paragraph{Reviewer 3: Overview}
\begin{feedback}
  This manuscript is an extension of the VLDB 2019 conference version of
  "Interactive Checks for Coordination Avoidance". The original PVLDB paper is
  a solid piece of work that is well written and easy to follow.  It was
  published in September 2018. In this manuscript, the authors elaborated more
  by adding more examples, proofs, and figures to make it easier for the reader
  to understand specific concepts and definitions. Some readers may find that
  the original paper is clear and enough without these additions. Moreover, the
  authors added two new sections, namely Section 4.3 and 7, which do not show a
  significant improvement to the original contribution. In almost two years, it
  was expected that the authors will address some of the future work mentioned
  in the original paper, such as "In the future, we plan on exploring ways to
  help a user reason about unreachable states and ways to discover sets of
  unreachable states automatically." Overall, the authors did not manage to
  show that the additions were significant enough to warrant publication.
\end{feedback}

\paragraph{Reviewer 3: Section 2.1}
\begin{feedback}
  The authors revised a small part of this section to clarify that the merge
  operator of a distributed object does not have to satisfy any special
  mathematical properties like associativity and commutativity.
\end{feedback}

\paragraph{Reviewer 3: Section 2.3}
\begin{feedback}
  The authors added this section to prove that their definition of invariant
  confluence and the original definition from Bailis et al.  [1] are
  equivalent. The authors did not clarify why it is important to prove that
  both definitions are equivalent.
\end{feedback}

\paragraph{Reviewer 3: Section 3}
\begin{feedback}
  The authors revised a small part of the original section to be in the form of
  a proof. For example in the PVLDB paper, the authors said: "The forward
  direction of Theorem 2 follows immediately from Theorem 1. The backward
  direction holds because any two invariant satisfying states s1 and s2 must be
  reachable (by assumption), so their join s1 t s2 is also reachable. And
  because O is (s0, T, I)-confluent, all reachable points, including s1 t s2,
  satisfy the invariant." And in the VLDBJ submission, the authors said:
  "Proof. The forward direction of Theorem 3 follows immediately from Theorem
  2. The backward direction holds because any two invariant satisfying states
  s1 and s2 must be reachable (by assumption), so their join s1 ⊔ s2 is also
  reachable. And because O is (s0, T, I)- confluent, all reachable points,
  including s1 ⊔s2, satisfy the invariant. Moreover s0 is reachable, so I(s0)
  because O is (s0,T,I)-confluent. "
\end{feedback}

\paragraph{Reviewer 3: Section 4.1}
\begin{feedback}
  the authors expanded their correctness proof of the interactive decision
  procedure and added a few other minor elaborations. The original proof in the
  PVLDB paper was clear enough to the reader.
\end{feedback}

\paragraph{Reviewer 3: Section 4.2}
\begin{feedback}
  In Section 4.2, the authors added a very brief example showing why invariant
  closure is undecidable. The reader can find original section 4.2 is clear
  enough. Instead in both versions of section 4.2, i.e., PVLDB and VLDBJ, the
  authors said: "In the future, we plan on exploring ways to help a user reason
  about unreachable states and ways to discover sets of unreachable states
  automatically." After almost two years of publishing the PVLDB version of
  this submission, this manuscript would be much stronger, if the authors
  explored these ways.
\end{feedback}

\paragraph{Reviewer 3: Section 4.3}
\begin{feedback}
  Section 4.3 is newly added to clarify that the decision procedure relies on
  human input to guide the decision procedure and explain how to mitigate these
  errors.
\end{feedback}

\paragraph{Reviewer 3: Section 5}
\begin{feedback}
  In section 5, the authors added more examples, proof, and Figure 4 to provide
  a better understanding of merge-reducible, non-merge- reducible objects.
  Still, the reader can easily understand these concepts from the original
  paper.
\end{feedback}

\paragraph{Reviewer 3: Section 6}
\begin{feedback}
  In Section 6, the authors elaborated more to clarify the concepts and
  definitions in this section. Some readers may find the original paper was
  enough to understand these concepts and definitions.
\end{feedback}

\paragraph{Reviewer 3: Section 7}
\begin{feedback}
  The authors added Section 7 to highlight a dual "operation-based" approach.
  This is not a significant improvement to the original contribution.
\end{feedback}

\paragraph{Reviewer 3: Reviewer 3: Other Comments}
\begin{feedback}
  Figure 5 should appear before Figure 6.
\end{feedback}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Reviewer 4}
\paragraph{Reviewer 4: Overview}
\begin{feedback}
  The paper is very well written, and it makes substantial improvement over the
  VLDB 2019 conference version.
\end{feedback}

\paragraph{Reviewer 4: 1}
\begin{feedback}
  In the introduction the concept of "invariant" is missing some context. Where
  does this invariant come from? It looks like this invariant is a given
  invariant for the program executing on the replicated object, and it is not
  necessarily a tight invariant.  Clarifying this in advance would be useful,
  as I realized this only in the beginning of Section 2, I read this:
  "Informally, a replicated object is invariant confluent *with respect to an
  invariant* if every replica of the object is guaranteed to satisfy the
  invariant despite the possibility of different replicas being concurrently
  modified or merged together." Mentioning the system model explicitly in the
  introduction also would help: "a distributed object and accompanying
  invariant is replicated across a set of servers."
\end{feedback}

\paragraph{Reviewer 4: 2}
\begin{feedback}
  Invariant confluence is about safety, however, could there be a liveness part
  of the problem worth investigating/discussing for replicated object
  management? Would it be possible that forgoing coordination across replicas
  could jeopardize/violate some liveness properties?
\end{feedback}

\paragraph{Reviewer 4: 3}
\begin{feedback}
  Mergeable Replicated Data Types paper (OOPSLA'19) is relevant work. The paper
  does not cite MRDT understandably because it is very recent work, however
  including a discussion of this in the related work would be helpful.
\end{feedback}

\paragraph{Reviewer 4: 4}
\begin{feedback}
  Section 4 is about invariant tightening, which has a broader applicability in
  distributed systems/programs. Unfortunately the techniques used here are not
  new. Involving the user for identifying unreachable states in not practical,
  and the discussion on tolerating user errors is weak as it does not offer a
  remedy. Would using some heuristics and abstracting state provide any help
  here?
\end{feedback}

\paragraph{Reviewer 4: 5}
\begin{feedback}
  Section 6 on segmented invariant confluence is nice. It would be nice if
  Section 6.4 could be improved. I liked the fourth point in Section 6.4, would
  it be possible to elaborate on this. The sixth point however is bad, as
  replicating every server with state-machine-replication is overkill. Would it
  be possible to use use a membership service approach (as in Vertical Paxos
  and Chain Replication configuration) to prevent a single crash to flunk
  global coordination?
\end{feedback}

\paragraph{Reviewer 4: 6}
\begin{feedback}
  Is there plans to make Lucy implementation open source?
\end{feedback}

\paragraph{Reviewer 4: 7}
\begin{feedback}
  In the throughput evaluation of segmented invariant confluence systems, what
  would be the effects/overhead of faulttolerance? (Related to item 5 above.)
\end{feedback}

\paragraph{Reviewer 4: 8}
\begin{feedback}
  Would it be possible to elaborate more on escrow transaction implementation.
  It was not clear to me from Example 13.
\end{feedback}

\bibliographystyle{abbrv}
\bibliography{references}
\end{multicols*}
\end{document}

### 2020-08-06 doc/the_vldb_journal_paper/make.sh

@@ -1,0 +1,11 @@
+#! /usr/bin/env bash
+
+set -euo pipefail
+
+main() {
+    latexmk -pdf rebuttal.tex
+    latexmk -pdf iconfluence_whittaker.tex
+    pdfunite rebuttal.pdf iconfluence_whittaker.pdf iconfluence_whittaker_revisions.pdf
+}
+
+main "$@"

### 2020-08-07 doc/the_vldb_journal_paper/rebuttal.tex
Start on revisions.

@@
 \usepackage{float}
+\usepackage{invariantconfluence}
 \usepackage{multicol}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: W2}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: W3}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: W4}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D1}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D2}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D3}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D4}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D5}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D6}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D7}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D8}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D9}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D10}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D11}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
+
 \paragraph{Reviewer 1: D12}
@@
 \end{feedback}
+\textcolor{red}{TODO.}
 
@@
   This manuscript is an extension of the VLDB 2019 conference version of
-  "Interactive Checks for Coordination Avoidance". The original PVLDB paper is
-  a solid piece of work that is well written and easy to follow.  It was
+  ``Interactive Checks for Coordination Avoidance''. The original PVLDB paper
+  is a solid piece of work that is well written and easy to follow. It was
   published in September 2018. In this manuscript, the authors elaborated more
@@
   was expected that the authors will address some of the future work mentioned
-  in the original paper, such as "In the future, we plan on exploring ways to
+  in the original paper, such as ``In the future, we plan on exploring ways to
   help a user reason about unreachable states and ways to discover sets of
-  unreachable states automatically." Overall, the authors did not manage to
+  unreachable states automatically.'' Overall, the authors did not manage to
   show that the additions were significant enough to warrant publication.
 \end{feedback}
+We address this feedback in detail below. We also disagree that the time
+between publication and now should have any bearing on the revisions.
 
@@
 \end{feedback}
+Yup. This was requested by many readers who assumed early on that the merge
+operator did require some special properties. In Bailis et al.'s original
+paper, for example, they assume the merge operator is associative, commutative,
+and idempotent~\cite{bailis2014coordination}.
 
@@
   The authors added this section to prove that their definition of invariant
-  confluence and the original definition from Bailis et al.  [1] are
-  equivalent. The authors did not clarify why it is important to prove that
-  both definitions are equivalent.
+  confluence and the original definition from Bailis et al.\ are equivalent.
+  The authors did not clarify why it is important to prove that both
+  definitions are equivalent.
 \end{feedback}
+Without a proof of the two definitions being equivalent, it's possible that our
+notion of \invariantconfluence{} and Bailis et al.'s original notion differ.
+If so, none of our theorems or decision procedures could be applied to Bailis et
+al.'s findings. We'd be talking about two completely different concepts.
+Proving they are equivalent unifies the work. It's similar to proving that
+relational algebra, SQL, and relational calculus are all equally expressive.
+Without this knowledge, proofs using the different definitions would be
+mutually incompatible.
 
@@
   The authors revised a small part of the original section to be in the form of
-  a proof. For example in the PVLDB paper, the authors said: "The forward
+  a proof. For example in the PVLDB paper, the authors said: ``The forward
   direction of Theorem 2 follows immediately from Theorem 1. The backward
@@
   because O is (s0, T, I)-confluent, all reachable points, including s1 t s2,
-  satisfy the invariant." And in the VLDBJ submission, the authors said:
-  "Proof. The forward direction of Theorem 3 follows immediately from Theorem
+  satisfy the invariant.'' And in the VLDBJ submission, the authors said:
+  ``Proof. The forward direction of Theorem 3 follows immediately from Theorem
   2. The backward direction holds because any two invariant satisfying states
@@
   including s1 ⊔s2, satisfy the invariant. Moreover s0 is reachable, so I(s0)
-  because O is (s0,T,I)-confluent. "
+  because O is (s0,T,I)-confluent.''
 \end{feedback}
+Yup, a minor but helpful improvement.
 
@@
 \end{feedback}
+We're happy that the original proof was clear to the reviewer, but we felt it
+was not clear to most people, including us.
 
@@
   enough. Instead in both versions of section 4.2, i.e., PVLDB and VLDBJ, the
-  authors said: "In the future, we plan on exploring ways to help a user reason
-  about unreachable states and ways to discover sets of unreachable states
-  automatically." After almost two years of publishing the PVLDB version of
-  this submission, this manuscript would be much stronger, if the authors
-  explored these ways.
+  authors said: ``In the future, we plan on exploring ways to help a user
+  reason about unreachable states and ways to discover sets of unreachable
+  states automatically.'' After almost two years of publishing the PVLDB
+  version of this submission, this manuscript would be much stronger, if the
+  authors explored these ways.
 \end{feedback}
+Automatically discovering unreachable states is hard, really hard. As we
+describe in the paper, it is easy to verify that a state is reachable, simply
+give an expression that evaluates to it. Verifying that a state is unreachable
+is significantly harder, as we have to create a proof showing that there does
+not exist any execution in which the state is reached. Things like the Collatz
+conjecture show that reasoning about the reachability or unreachability of even
+simple programs can be tremendously difficult. If we developed any substantial
+techniques to identify unreachable states in a general way, we feel that would
+constitute a strong publication on its own.
 
@@
 \end{feedback}
+Yes, and we feel this is one of the core additions to the paper. The two
+biggest drawbacks of interactivity are (1) the time and labor required of a
+user and (2) the possibility of introducing errors. We break down the two types
+of user errors (i.e.\ labelling an unreachable state as reachable and labelling
+a reachable state as unreachable) and provide solutions for both, eliminating
+the possibility of an erroneously verified \invariantconfluent{} object from
+violating an invariant in deployment. This addition is very important for
+anyone deploying \invariantconfluent{} objects in the wild. It's also
+surprising and non-obvious. We originally did not think that our decision
+procedures could be made tolerant to user errors.
 
@@
 \end{feedback}
+We strongly disagree. We believe that Section 5 was the most difficult section
+to understand in our original publication, mostly because we did not have
+enough space to fully explain all of our definitions, theorems, and proofs. The
+statement and proof of Theorem 4 (in our original publication) is particularly
+nuanced. We think that our added proofs greatly clarify this section, and our
+figures illustrating the intuition behind the proofs are especially helpful.
 
@@
 \end{feedback}
+We added an example to the paper to better explain co-reachable and
+co-unreachable states.  We agree that this was a minor bit of clarification,
+and that the section was already relatively clear. However, we also made a
+number of other additions to Section 6. We explained the difference between
+segmented \invariantconfluence{} and a distributed locking approach to
+maintaining invariants~\cite{balegas2015putting, gotsman2016cause}. We
+discussed two performance optimizations to our segmented \invariantconfluence{}
+system model.  We discussed how segmented \invariantconfluence{} can leverage
+our results on \mergereducibility{}. We elaborated on the fault-tolerance of
+our model. We believe that, taken together, these additions are significant.
 
@@
 \begin{feedback}
-  The authors added Section 7 to highlight a dual "operation-based" approach.
+  The authors added Section 7 to highlight a dual ``operation-based'' approach.
   This is not a significant improvement to the original contribution.
 \end{feedback}
+We disagree. The dual operation-based definitions of \invariantconfluence{} are
+both theoretically interesting and practically important. Theoretically, as
+with work on CRDTs, having dual models of the same concept can be helpful in
+writing proofs and building intuition. Some properties are more naturally
+expressed in the state-based model, while others are more natural in the
+operation-based model. The operation-based model also has practical importance.
+Many CRDTs are easier to understand as state-based objects, but are more
+efficiently implemented as operation-based objects. By introducing
+operation-based definitions of \invariantconfluence{}, we allow system
+designers to choose whichever form of CRDT is more efficient. As we describe in
+the paper, for example, the operation-based definitions let us replace a
+PN-counter with a more efficient integer.
 
@@
 \end{feedback}
+Thanks! We appreciate it!
 
@@
 \end{feedback}
+Forgoing coordination across replicas actually \emph{increases} liveness,
+rather than jeopardizing or violating it. Any alive replica of an
+\invariantconfluent{} object can service client requests and execute
+transactions. This is the strongest form of liveness we can hope to achieve. In
+fact, if we aim for strongly consistent replication, then the CAP theorem tells
+us that total availability (the kind that \invariantconfluence{} achieves) is
+impossible~\cite{brewer2012cap, gilbert2002brewer}.
 
@@
 \end{feedback}
+{
+  Yes. Lucy is available online at \urlstyle{rm}
+  \url{https://github.com/mwhittaker/enforced_invariant_confluence}. We also
+  added this link to Section 8 of the paper.
+}
 

### 2020-08-07 doc/the_vldb_journal_paper/rebuttal.tex
More progress on rebuttal.

@@
 \begin{feedback}
-  in page 3 we read "The culprit is concurrent decrements, which we can forbid
-  in a simple one-segment segmentation ...", this sentence does not parse well
+  in page 3 we read ``The culprit is concurrent decrements, which we can forbid
+  in a simple one-segment segmentation ...'', this sentence does not parse well
 \end{feedback}
-\textcolor{red}{TODO.}
+We reworded the sentence to be easier to read.
 
@@
 \begin{feedback}
-  In 6.4 we read "This example is exactly the same as Example 8 which we
+  In 6.4 we read ``This example is exactly the same as Example 8 which we
   handled by simply removing withdrawal transactions from our segmentation's
-  set of transactions." How can one identify transactions to remove, and is the
-  removal practically done?
+  set of transactions.'' How can one identify transactions to remove, and is
+  the removal practically done?
 \end{feedback}
-\textcolor{red}{TODO.}
+Users must manually identify transactions to remove. More generally,
+identifying a transaction to remove from a segmentation is a special case of
+forming a segmentation, and we do not provide any automated assistance on how
+to form a segmentation. Is the removal practically done? That depends on the
+workload. For some workloads, identifying invariant-violating transactions can
+be easy. The bank withdrawal example (Example 8). Other, more complex
+workloads, can be difficult to reason about. Forming segmentations for these
+workloads is not easy. We think that automatically forming segmentations is an
+interesting avenue for future research.
 
@@
 \begin{feedback}
-  In 8.1 we read "When a server receives a coordination request from the
-  leader, it stops processing transactions and sends the leader its state."
+  In 8.1 we read ``When a server receives a coordination request from the
+  leader, it stops processing transactions and sends the leader its state.''
   what happens to the in-flight transactions? Are the cancelled or allowed to
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+In-flight transactions are deferred until after the round of coordination is
+done. They aren't cancelled, and they are allowed to finish, but only after the
+coordination is done. The entire system stops processing transactions
+temporarily.  Technically, the system can still \emph{admit} new transactions
+during coordination, but it cannot execute them.
+
+We describe this in Section 6.2, for example, when we say that if ``$t \notin
+T_i$, then $p_i$ initiates a round of global coordination to execute $t$.
+During global coordination, every server temporarily stops processing
+transactions.''
+
+Your question also reinforces the conclusion of Figure 10. If coordination is
+frequent, then segmented \invariantconfluence{} becomes slow, even slower than
+linearizable replication. Segmented \invariantconfluent{} works well only when
+coordination is infrequent.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+Fixed.
 

### 2020-08-07 doc/the_vldb_journal_paper/sections/segmented_invariant_confluence.tex
Typo fixes.
More progress on rebuttal.

@@
   by the max function. Let $I = \nats$. Let $T = \set{t_{2\times}, t_{+10}}$
-  where $t_{2\times}(x) = 2x$ and $t_{+10} = x + 10$. The states $3$ and $16$
-  are \TIcoreachable{} because both are \stireachable{s}{T}{I} from $s = 3$
-  ($16 = 2 \times 3 + 10$). The states $3$ and $5$, however, are not
+  where $t_{2\times}(x) = 2x$ and $t_{+10}(x) = x + 10$. The states $3$ and
+  $16$ are \TIcoreachable{} because both are \stireachable{s}{T}{I} from $s =
+  3$ ($16 = 2 \times 3 + 10$). The states $3$ and $5$, however, are not
   \TIcoreachable{}. The only state in which $3$ is reachable is $3$, but from
@@
   1)$. Both are reachable, so $O$ is not \sTIconfluent{} and
-  \algoref{InteractiveDecisionProcedure} returns false.  The culprit is
-  concurrent decrements, which we can forbid in a simple one-segment
-  segmentation $\Sigma = (I, T^+)$ where $T^+$ consists only of increment
-  transactions. Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
+  \algoref{InteractiveDecisionProcedure} returns false.
+  \begin{revisions}
+    $O$ is not \sTIconfluent{} because of concurrent decrements. We can forbid
+    concurrent decrements using a simple one-segment segmentation $\Sigma = (I,
+    T^+)$ where $T^+$ consists only of increment transactions.
+  \end{revisions}
+  Applying, \algoref{ArbitraryStartInteractiveDecisionProcedure},
   \IsIclosed{} again returns false with the same states $s_1$ and $s_2$. This time,
@@
 every round of global coordination will fail to complete. Fortunately, we can
-leverage well-established means of ensuring fault tolerance without difficulty.
-For example, we can replicate every server using a state machine replication
-protocol like MultiPaxos~\cite{lamport1998part, lamport2001paxos} or
+leverage well-established means of ensuring fault tolerance. For example, we
+can replicate every server using a state machine replication protocol like
+MultiPaxos~\cite{lamport1998part, lamport2001paxos} or
 Raft~\cite{ongaro2014search}. Doing so will introduce a constant slowdown for

### 2020-08-07 doc/the_vldb_journal_paper/rebuttal.tex
More work on revisions.

@@
 \end{feedback}
-Forgoing coordination across replicas actually \emph{increases} liveness,
-rather than jeopardizing or violating it. Any alive replica of an
-\invariantconfluent{} object can service client requests and execute
-transactions. This is the strongest form of liveness we can hope to achieve. In
-fact, if we aim for strongly consistent replication, then the CAP theorem tells
-us that total availability (the kind that \invariantconfluence{} achieves) is
-impossible~\cite{brewer2012cap, gilbert2002brewer}.
+Forgoing coordination across replicas in favor of \invariantconfluence{}
+actually \emph{increases} liveness, rather than jeopardizing or violating it.
+Any alive replica of an \invariantconfluent{} object can service client
+requests and execute transactions. This is the strongest form of liveness we
+can hope to achieve. In fact, if we aim for strongly consistent replication,
+then the CAP theorem tells us that total availability (the kind that
+\invariantconfluence{} achieves) is impossible~\cite{brewer2012cap,
+gilbert2002brewer}.
+
+Note, however, that a segmented \invariantconfluent{} object does suffer from
+liveness problems, as we discuss in Section 6.4. Specifically, transactions
+that require global coordination are live only as long as every replica is
+live. We discuss this more in Reviewer 4: 5 below.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We added a couple paragraphs to Example 13 with a concrete example of the
+optimization described in our fourth point. We also added a forward reference
+in Section 6.4 to Example 13.
 
-\paragraph{Reviewer 4: 6}
+Your comment on our sixth point is interesting. We hadn't thought of that
+before. Coincidentally, since our initial publication, we spent some time
+researching membership changes in state machine replication protocols---which
+you can find here:
+\url{https://mwhittaker.github.io/publications/matchmaker_paxos.html}---so we
+can answer your question with a bit of confidence. Assume a client sends a
+transactions to a server $p$. The transaction happens to not need global
+coordination, so $p$ executes the transaction locally and sends back the result
+to the client. This scenario makes membership changes difficult for the
+following reason. Consider a different execution in which $p$ is partitioned
+from the rest of the servers. The system erroneously thinks that $p$ has failed
+and reconfigures to some other set of servers. However, $p$ and the client are
+oblivious to this. The client can still contact $p$ and have transactions
+executed, even though the system has moved to a completely new set of servers.
+This can lead to situations that are impossible without membership changes. For
+example, $p$ may be in a different active segment than the rest of the servers.
+
+This problem can be mitigated by making stronger assumptions (e.g., a failure
+detector, clock synchrony assumptions with leases). Alternatively, we could
+require that a client contact multiple servers to execute a transaction and
+then use quorum intersection to perform membership changes safely. We think
+that this is a very interesting direction of future work, but unfortunately (as
+with all membership change protocols) is very nuanced and very easy to get
+wrong. It is out of the scope of our paper.
+
 \begin{feedback}
@@
 \end{feedback}
-{
-  Yes. Lucy is available online at \urlstyle{rm}
+Yes. Lucy is available online at
 \url{https://github.com/mwhittaker/enforced_invariant_confluence}. We also
 added this link to Section 8 of the paper.
-}
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+If we replicate every server with a state machine replication protocol, the
+throughput of the segmented \invariantconfluent{} object certainly
+decreases. However, since ever server can still independently execute
+transactions, the throughput of the system would still be scalable for
+workloads that don't require frequent coordination.
+%
+The performance overheads of more complex forms of fault tolerance (as
+described in item 5 above) are out of scope for our paper.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We added two more paragraphs of explanation to Example 13, including a concrete
+example. Hopefully, that makes things clearer.
 

### 2020-08-07 doc/the_vldb_journal_paper/sections/evaluation.tex
Start on revisions.
More progress on rebuttal.
More work on revisions.

@@
 our decision procedures and system models.
+\begin{revisions}
+  All code is available online:
+  \urlstyle{rm}
+  \url{https://github.com/mwhittaker/enforced_invariant_confluence}.
+\end{revisions}
 
@@
 We now consider a simple auction system introduced in~\cite{gotsman2016cause}.
-Our object consists of a set $B$ of integer-valued bids and a optional winning
+Our object consists of a set $B$ of integer-valued bids and an optional winning
 bid $w$. Initially, $B = \emptyset$ and $w = \bot$ (indicating that there is no
@@
 \[
-  I_k = \setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}{n_1, n_2, n_3 \leq k \leq p_1, p_2, p_3}
+  I_k = \setst{(p_1, p_2, p_3), (n_1, n_2, n_3)}%
+              {n_1, n_2, n_3 \leq k \leq p_1, p_2, p_3}
 \]
@@
 segment is the one presented in \exampleref{CounreachableExample} which
-prohibits concurrent decrements. We ran our decision procedure on this example
-and it concluded that it was segmented \invariantconfluent{} in less than a
-tenth of a second and without any human interaction.
+prohibits concurrent decrements.
+
+\begin{revisions}
+  For example, assume that $k = 3$ and that the three states are $s_1 = s_2 =
+  s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 4, 5), (3, 3, 3)$. Here, the
+  value of the PN-counter is $(3 + 4 + 5) - (3 + 3 + 3) = 3$, which is
+  non-negative as expected. If server 1 receives a decrement request, it
+  \emph{cannot} execute the decrement without global coordination because doing
+  so would cause $n_1$ to exceed the escrow threshold $k = 3$. Naively, it
+  seems safe. Server 1 would transition to state $(3, 3, 5), (3, 3, 3)$ which
+  has non-negative value $2$. However, if servers 2 and 3 concurrently perform
+  decrements of their own, the three servers' states would merge to a negative
+  value. Thus, the decrement requires global coordination to rule out the
+  possibility of concurrent decrements.
+
+  On the other hand, if server 2 receives a decrement request, it can safely
+  execute it locally and transition to state $(3, 3, 5), (3, 3, 3)$. This is
+  safe because server 2 knows that every server $i$ maintains $p_i \geq n_i$.
+  Thus, after merging, the value of the PN-counter is $\sum_i p_i - \sum_i n_i
+  = \sum_i (p_i - n_i)$ which is the sum of non-negative terms. Intuitively, if
+  every server agrees to never go in debt, then the PN-counter in aggregate
+  never goes in debt.
+\end{revisions}
+
+We ran our decision procedure on this example and it concluded that it was
+segmented \invariantconfluent{} in less than a tenth of a second and without
+any human interaction.
+
+\begin{revisions}
+  In \secref{SegmentedDiscussion}, we discussed an optimization in which a
+  server receives a transaction $t$ and concludes that the transaction requires
+  global coordination. The server can defer the execution of $t$,
+  executing other transactions in the mean time. When the server tries to
+  execute $t$ a second time, the transaction may no longer require global
+  coordination.
+
+  Escrow transactions are a good example where this optimization works
+  particularly well. Assume that $k = 3$ and that the three states are $s_1 =
+  s_2 = s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 3, 3), (3, 3, 3)$. Here,
+  no server can process a decrement transaction locally. If a server receives a
+  decrement transaction, it must execute it with global coordination. However,
+  if the server instead defers the execution of the decrement transaction and
+  happens to receive a number of increment transactions, then it can execute
+  the decrement locally, without the need for global coordination.
+\end{revisions}
 

### 2020-08-07 doc/the_vldb_journal_paper/sections/segmented_invariant_confluence.tex
More work on revisions.

@@
 
-\subsection{Discussion and Limitations}
+\subsection{Discussion and Limitations}~\seclabel{SegmentedDiscussion}
 There are a few things worth noting about segmented \invariantconfluence{}, its
@@
 satisfies the active invariant or violates the global invariant. In either
-case, a round of global coordination is avoided. Similarly, servers can
-buffer transactions that require global coordination, executing an entire
-batch of these transactions during a single round of global coordination.
+case, a round of global coordination is avoided. \markrevisions{Later, in
+\exampleref{EscrowTransactionsEval}, we'll see a concrete example of this
+optimization.}
+%
+Similarly, servers can buffer transactions that require global coordination,
+executing an entire batch of these transactions during a single round of global
+coordination.
 

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex
More work on revisions.
Mostly reviewer 1 and a couple of reviewer 4 left.

@@
 \end{feedback}
-\textcolor{red}{TODO.}
+Fixed!
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We agree that involving the user for identifying unreachable states is not
+ideal, and in many cases not practical. We think our discussion of tolerating
+user errors is also not ideal, but not weak. By requiring users to provide
+proofs of reachability and by deploying the invariant $I - NR$ instead of $I$,
+invariant violations are impossible.
+%
+Heuristics and abstracting state and other techniques from invariant tightening
+would probably help, but they are out of scope of the paper. As you mention,
+the techniques are not new. We view it as an orthogonal but complementary line
+of research.
 

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex
More revisions.
Only four left!

@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We added a link to our DSL implementation on GitHub to the beginning of Section
+8. The GitHub repository also contains a couple of examples that didn't make it
+into the paper. We also added a bit more explanation of the DSL to Section 8.1.
+We agree the paper would benefit from more examples, but given that the paper
+already has over 10 examples of invariant confluence with varying degrees of
+complexity from simple (e.g., Example 10) to complex (e.g, Example 14), we feel
+the added benefit of yet another example does not justify the effort.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See above. We agree that having a number of realistic workloads in the paper
+would be beneficial, but specifying, verifying, and explaining every workload
+requires a significant amount of work for a marginal increase in the clarity of
+the paper.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See above.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See D6.
 
@@
 \begin{feedback}
-  W4 touches upon the "new material" for this extended version of the paper,
+  W4 touches upon the ``new material'' for this extended version of the paper,
   and the above weak points W1, W2 and W3 can all serve as additional material.
 \end{feedback}
-\textcolor{red}{TODO.}
+See W1, W2, W3, and D6.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See W1.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See W2.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See W3.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We think that automated segmentation would be such a significant contribution
+that it would warrant a publication on its own.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+We disagree. Our paper is largely theoretical. We provide new, and arguably
+improved, definitions of \invariantconfluence{}. We characterize the
+relationship between \invariantclosure{} and \invariantconfluence{}. We exploit
+the relationship to develop an \invariantclosure{} decision procedure. We
+invent the notion of \mergereducibility{} and characterize its relationship
+with \invariantclosure{}. We invent the notion of segmented
+\invariantconfluence{}, and develop another decision procedure.
+
+Addressing W1-W3 would certainly strengthen the paper. We don't argue that.
+But, these three points focus exclusively on the empirical evaluation of our
+techniques, rather than the theoretical underpinnings, which we feel is the
+core message of the paper. This is why we focused our revisions on the
+theoretical aspects of the paper. We believe that these additions are
+substantial. Repeating from our change log:
+\begin{itemize}
+  \item (Section 2.3)
+    In our conference paper, we present a definition of invariant confluence
+    that is slightly different than the original definition from Bailis et
+    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
+    We added a subsection (Section 2.3) clarifying that the difference in
+    definitions is superficial. We describe Bailis' original definition using
+    our formalism and prove that the two definitions are equivalent.
+
+  \item (Section 3, Theorem 2)
+    We added a formal proof that invariant closure is sufficient for invariant
+    confluence.
+
+  \item (Section 3, Theorem 3)
+    We added a formal proof of our key result equating invariant closure and
+    invariant confluence.
+
+  \item (Section 4.1)
+    We expanded the correctness proof of our interactive decision procedure. A
+    key step in understanding the correctness of our interactive invariant
+    confluence decision procedure---namely that $I$-reachable and
+    $(I-NR)$-reachable states are equal---was omitted from our conference
+    paper. We added a formal proof of this fact as well as a few other minor
+    elaborations.
+
+  \item (Section 4.2)
+    In our conference paper, we asserted that checking for invariant closure is
+    undecidable in general. We added a brief example showing \emph{why}
+    invariant closure is undecidable, by reducing it to the Halting Problem.
+
+  \item (Section 4.3)
+    One of the core contributions of our paper is an interactive invariant
+    confluence decision procedure. Here, ``interactive'' means that the
+    decision procedure relies on human input to guide the decision procedure.
+    It's possible for a person to provide the decision procedure with incorrect
+    information. We added a subsection (Section 4.3) describing how to mitigate
+    these errors.
+
+  \item (Section 5, Examples 3 and 4)
+    We added examples of merge-reducible and non-merge-reducible objects to
+    make it easier for the reader to understand the definition of
+    merge-reducibility.
+
+  \item (Section 5, Theorem 4 and Figure 4)
+    We added a formal proof that merge-reducibility is a sufficient condition
+    for invariant-confluence. We also added a diagram illustrating the key idea
+    behind the proof. Similarly, we added a counterexample showing that
+    merge-reducibility is not a necessary condition for invariant-confluence.
+
+  \item (Section 5, Theorem 5 and Figure 5)
+    We added a formal proof that criteria 1 through 4 are sufficient for
+    invariant confluence. The idea behind the proof is straightforward but the
+    details are tedious. To help readers see the forest through the trees, we
+    also added an illustration of the key idea of the proof.
+
+  \item (Section 5, Example 5)
+    In our conference paper, we provide an example of a distributed object that
+    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
+    closed. In the example, we asserted that criteria 1 through 4 are satisfied
+    but did not elaborate on why. We added a brief explanation making it clear
+    why the criteria are met.
+
+  \item (Section 6.3, Example 7)
+    We added an example of co-reachable and co-unreachable states to help the
+    reader understand the definition of co-reachability.
+
+  \item (Section 6.4)
+    We added a paragraph explaining the relationship between segmented
+    invariant confluence and a distributed locking approach to maintaining
+    invariants (that can be found in related works).
+
+  \item (Section 6.4)
+    We added a paragraph explaining a couple of optimizations that can be made
+    to our segmented invariant confluence system model to reduce the frequency
+    of global coordination.
+
+  \item (Section 6.4)
+    We added a paragraph explaining how a segmented invariant confluence
+    decision procedure can leverage our merge reducibility results in addition
+    to our results on invariant closure.
+
+  \item (Section 6.4)
+    We clarified the fault tolerance limitations of segmented invariant
+    confluence. A naive implementation of our segmented invariant confluence
+    decision procedure does not guarantee liveness, but there are
+    straightforward techniques we can leverage to introduce liveness. We added
+    an explanation of this to the paper.
+
+  \item (Section 7)
+    Our conference paper presents formal definitions of distributed objects and
+    invariant confluence. The definitions revolve around servers gossiping
+    state to one another and merging the states together using a binary merge
+    operator. This approach is called a ``state-based'' approach. We added a
+    section (Section 7) outlining a dual ``operation-based'' approach in which
+    transactions, rather than states, are gossiped. This duality between
+    state-based and operation-based definitions is well established in the CRDT
+    literature.
+\end{itemize}
 
@@
 \begin{feedback}
-  In the intro we read about "the trade-off between application complexity and
-  coordination-freedom", it would be great to see some practical examples later
+  In the intro we read about ``the trade-off between application complexity and
+  coordination-freedom'', it would be great to see some practical examples later
   in the paper
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See Reviewer 4: 5, 7.
 

### 2020-08-08 doc/the_vldb_journal_paper/sections/evaluation.tex
More revisions.
Only four left!

@@
 \begin{revisions}
-  All code is available online:
+  All code and all the specifications described below are available online:
   \urlstyle{rm}
-  \url{https://github.com/mwhittaker/enforced_invariant_confluence}.
+  \url{https://github.com/mwhittaker/enforced_invariant_confluence}. The
+  repository also contains a couple more example specifications.
 \end{revisions}
@@
 transactions, and invariants in a small Python DSL and interact with the
-interactive decision procedures using an interactive Python console. Note that
-a programmer only has to run the decision procedures offline a single time to
-check the \invariantconfluence{} of their distributed object. The decision
-procedures do not have to be run online when transactions are being processed.
+interactive decision procedures using an interactive Python console.
+\begin{revisions}
+  The Python DSL includes a library of CRDTs including integers, booleans,
+  tuples, sets, option types, and maps. It supports invariants involving
+  equalities, inequalities, arithmetic, boolean logic, and basic set operations
+  (e.g, union, intersection).
+\end{revisions}\
+Note that a programmer only has to run the decision procedures offline a single
+time to check the \invariantconfluence{} of their distributed object. The
+decision procedures do not have to be run online when transactions are being
+processed.
 
-We use Z3~\cite{de2008z3} to implement our \invariantclosure{} decision procedure,
-compiling an object and invariant into a formula that is satisfiable if and
-only if the object is \emph{not} \invariantclosed{}. If the object is
+We use Z3~\cite{de2008z3} to implement our \invariantclosure{} decision
+procedure, compiling an object and invariant into a formula that is satisfiable
+if and only if the object is \emph{not} \invariantclosed{}. If the object is
 \invariantclosed{}, then Z3 concludes that the formula is unsatisfiable.
@@
   Escrow transactions are a good example where this optimization works
-  particularly well. Assume that $k = 3$ and that the three states are $s_1 =
-  s_2 = s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 3, 3), (3, 3, 3)$. Here,
-  no server can process a decrement transaction locally. If a server receives a
-  decrement transaction, it must execute it with global coordination. However,
-  if the server instead defers the execution of the decrement transaction and
-  happens to receive a number of increment transactions, then it can execute
-  the decrement locally, without the need for global coordination.
+  particularly well. Assume that $k = 3$ and that the three states are $(3, 3,
+  3), (3, 3, 3)$. Here, no server can process a decrement transaction locally.
+  If a server receives a decrement transaction, it must execute it with global
+  coordination. However, if the server instead defers the execution of the
+  decrement transaction and happens to receive a number of increment
+  transactions, then it can execute the decrement locally, without the need for
+  global coordination.
 \end{revisions}

### 2020-08-08 doc/the_vldb_journal_paper/rebuttal.tex
More work on revisions.
Only two left!

@@
 \begin{feedback}
-  end of first column of page 3 "that it execute a transaction" --> "that it
-  executes a transaction"
+  end of first column of page 3 ``that it execute a transaction'' --> ``that it
+  executes a transaction''
 
-  section 5, in example 3, "Our start state state" --> two times state
+  section 5, in example 3, ``Our start state state'' --> two times state
 
-  section 8.2 in example 12: "and a optional" --> "and an optional"
+  section 8.2 in example 12: ``and a optional'' --> ``and an optional''
 \end{feedback}
@@
 \begin{feedback}
-  In the introduction the concept of "invariant" is missing some context. Where
+  In the introduction the concept of ``invariant'' is missing some context. Where
   does this invariant come from? It looks like this invariant is a given
@@
   as I realized this only in the beginning of Section 2, I read this:
-  "Informally, a replicated object is invariant confluent *with respect to an
+  ``Informally, a replicated object is invariant confluent *with respect to an
   invariant* if every replica of the object is guaranteed to satisfy the
   invariant despite the possibility of different replicas being concurrently
-  modified or merged together." Mentioning the system model explicitly in the
-  introduction also would help: "a distributed object and accompanying
-  invariant is replicated across a set of servers."
+  modified or merged together.'' Mentioning the system model explicitly in the
+  introduction also would help: ``a distributed object and accompanying
+  invariant is replicated across a set of servers.''
 \end{feedback}
-\textcolor{red}{TODO.}
+We re-worded the introduction with your suggestion and also added a short
+example of \invariantconfluence{} to the introduction.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+Thanks for pointing out this paper. We hadn't heard of it before, and it is
+definitely relevant to our work on \invariantconfluence{}. We added a
+discussion of MRDTs to our related works section.
 

### 2020-08-08 doc/the_vldb_journal_paper/sections/introduction.tex
More work on revisions.
Only two left!

@@
 without the need for any coordination~\cite{bailis2014coordination}. If an
-object is \invariantconfluent{} with respect to an invariant, we can replicate
-it with the performance benefits of weak consistency and (some of) the
-correctness benefits of strong consistency.
+object is \invariantconfluent{} with respect to an \markrevisions{accompanying}
+invariant, we can replicate the \markrevisions{object and invariant} with the
+performance benefits of weak consistency and (some of) the correctness benefits
+of strong consistency.
+\begin{revisions}
+  For example, a replicated bank account balance (represented as an integer) is
+  \invariantconfluent{} with respect the invariant of a non-negative balance,
+  so long as we only allow deposits.
+\end{revisions}
 

### 2020-08-08 doc/the_vldb_journal_paper/sections/related_work.tex
Added tech report sections of paper.
More work on revisions.
Only two left!

--- 2020-01-13 doc/vldb_paper/sections/related_work.tex
@@
 
+\begin{revisions}
+  \textbf{Mergeable Replicated Data Types.}
+  Mergeable Replicated Data Types (MRDTs)~\cite{kaki2019mergeable}, like CRDTs,
+  are eventually consistent replicated objects. CRDTs get their eventual
+  consistency by merging two CRDT replicas using a commutative, associative,
+  and idempotent merge function. MRDTs instead merge two divergent replicas
+  along with their least common ancestor. MRDTs rely on an abstraction function
+  $\alpha$ and concretization function $\gamma$ to map abstract data types to
+  the domain of relations. Given these two functions, one can derive an
+  appropriate merge function.
+
+  Like CRDTs, MRDTs are eventually consistent but may not preserve invariants,
+  while \invariantconfluent{} objects preserve invariants but may not be
+  eventually consistent. An interesting direction for future work would be to
+  generalize \invariantconfluence{}'s use of binary merge operators to use the
+  ternary merge operators used by MRDTs.
+\end{revisions}
+
 \textbf{CALM Theorem.}
@@
 
-\begin{techreport}
-
 \textbf{Program Analysis in Database Systems.} Using program analysis to
@@
 reap the performance benefits gained from understanding program semantics.
-\end{techreport}

### 2020-08-09 doc/the_vldb_journal_paper/rebuttal.tex
Another rebuttal point.
Only one left!
Finished first crack at rebuttal.

@@
 \end{feedback}
-\textcolor{red}{TODO.}
+The trade-off between application complexity and coordination-freedom is shown
+in Examples 2, 6, 12, and 13. Very simple examples (Example 2) don't require
+any segmentation. More complex initial states (Example 6), more complex
+transactions (Example 12), and more complex optimizations (Example 13) require
+segmentation. We agree though that these examples don't demonstrate the
+trade-off in an airtight way. The trade-off we describe in the introduction is
+written to give some intuition to the reader, but ultimately the notion of
+complexity is subjective, and the number of segments of an application is up to
+the user.
 
@@
 \end{feedback}
-\textcolor{red}{TODO.}
+See Reviewer 3: Section 4.2.
 

### 2020-08-10 doc/the_vldb_journal_paper/rebuttal.tex
Hopefully final revisions...
...before catching typos of course :)

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section*{Overview}
+Our original publication on \invariantconfluence{} is largely theoretical. We
+provide new, and arguably improved, definitions of \invariantconfluence{}. We
+characterize the relationship between \invariantclosure{} and
+\invariantconfluence{}. We exploit the relationship to develop an
+\invariantclosure{} decision procedure. We invent the notion of
+\mergereducibility{} and characterize its relationship with
+\invariantclosure{}. We invent the notion of segmented \invariantconfluence{},
+and develop another decision procedure.
+
+When our paper was invited to this journal, we took the opportunity to improve
+the rigor and clarity of the core contributions of the paper. We added numerous
+proofs to verify the theorems in the paper, and we added many examples and
+illustrations providing valuable intuition behind the key results. We also
+added a significant amount of new material:
+\begin{itemize}
+  \item (Section 2.3)
+    In our conference paper, we present a definition of invariant confluence
+    that is slightly different than the original definition from Bailis et
+    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
+    We added a subsection (Section 2.3) clarifying that the difference in
+    definitions is superficial. We describe Bailis' original definition using
+    our formalism and prove that the two definitions are equivalent.
+
+  \item (Section 4.3)
+    One of the core contributions of our paper is an interactive invariant
+    confluence decision procedure. Here, ``interactive'' means that the
+    decision procedure relies on human input to guide the decision procedure.
+    It's possible for a person to provide the decision procedure with incorrect
+    information. We added a subsection (Section 4.3) describing how to mitigate
+    these errors.
+
+  \item (Section 6.4)
+    We added a paragraph explaining the relationship between segmented
+    invariant confluence and a distributed locking approach to maintaining
+    invariants (that can be found in related works).
+
+  \item (Section 6.4)
+    We added a paragraph explaining a couple of optimizations that can be made
+    to our segmented invariant confluence system model to reduce the frequency
+    of global coordination.
+
+  \item (Section 6.4)
+    We added a paragraph explaining how a segmented invariant confluence
+    decision procedure can leverage our merge reducibility results in addition
+    to our results on invariant closure.
+
+  \item (Section 6.4)
+    We clarified the fault tolerance limitations of segmented invariant
+    confluence. A naive implementation of our segmented invariant confluence
+    decision procedure does not guarantee liveness, but there are
+    straightforward techniques we can leverage to introduce liveness. We added
+    an explanation of this to the paper.
+
+  \item (Section 7)
+    Our conference paper presents formal definitions of distributed objects and
+    invariant confluence. The definitions revolve around servers gossiping
+    state to one another and merging the states together using a binary merge
+    operator. This approach is called a ``state-based'' approach. We added a
+    section (Section 7) outlining a dual ``operation-based'' approach in which
+    transactions, rather than states, are gossiped. This duality between
+    state-based and operation-based definitions is well established in the CRDT
+    literature.
+\end{itemize}
+
+The requested revisions, however, fall mainly into two categories. First, we
+have requests to expand our empirical evaluation. Unfortunately, an expanded
+evaluation requires a significant amount of work for very little additional
+technical contribution. The applications of our theoretical results are
+important (which is why we included them in the paper originally), but it is
+not the focus of the paper. We believe that future work focused more on the
+application of \invariantconfluence{} would leverage the key insights of our
+paper and would benefit from the increased clarity and rigor of our journal
+submission.
+
+Second, we have requests to complete some of our suggested future work, with
+arguments that the two years since our original publication should be enough
+time to do so. We disagree that the time since our original publication should
+have any bearing on our journal submission. Since our publication, we have been
+working in new research areas. We have not spent the last two years working on
+\invariantconfluence{}. Moreover, the suggested future work involves a
+significant research undertaking that would involve solving many fundamental
+research problems with implications well beyond \invariantconfluence{}. We
+believe that any such results would more than warrant a publication on their
+own.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section*{Reviewer 1}
@@
 We added a link to our DSL implementation on GitHub to the beginning of Section
-8. The GitHub repository also contains a couple of examples that didn't make it
-into the paper. We also added a bit more explanation of the DSL to Section 8.1.
-We agree the paper would benefit from more examples, but given that the paper
+8. We also added a bit more explanation of the DSL to Section 8.1.  The paper
 already has over 10 examples of invariant confluence with varying degrees of
-complexity from simple (e.g., Example 10) to complex (e.g, Example 14), we feel
-the added benefit of yet another example does not justify the effort.
+complexity from simple (e.g., Example 10) to complex (e.g., Example 14), and we
+feel that additional examples will offer little additional technical insight.
 
@@
 \end{feedback}
-See above. We agree that having a number of realistic workloads in the paper
-would be beneficial, but specifying, verifying, and explaining every workload
-requires a significant amount of work for a marginal increase in the clarity of
-the paper.
+We agree that having a number of realistic workloads in the paper would be
+beneficial, but specifying, verifying, and explaining every workload requires a
+significant amount of work for a marginal increase in the clarity of the paper.
+The focus of the paper is the foundations, not the applications of these ideas.
 
@@
 \end{feedback}
-See D6.
+See our Overview section above.
 
@@
 \end{feedback}
-We disagree. Our paper is largely theoretical. We provide new, and arguably
-improved, definitions of \invariantconfluence{}. We characterize the
-relationship between \invariantclosure{} and \invariantconfluence{}. We exploit
-the relationship to develop an \invariantclosure{} decision procedure. We
-invent the notion of \mergereducibility{} and characterize its relationship
-with \invariantclosure{}. We invent the notion of segmented
-\invariantconfluence{}, and develop another decision procedure.
-
-Addressing W1-W3 would certainly strengthen the paper. We don't argue that.
-But, these three points focus exclusively on the empirical evaluation of our
-techniques, rather than the theoretical underpinnings, which we feel is the
-core message of the paper. This is why we focused our revisions on the
-theoretical aspects of the paper. We believe that these additions are
-substantial. Repeating from our change log:
-\begin{itemize}
-  \item (Section 2.3)
-    In our conference paper, we present a definition of invariant confluence
-    that is slightly different than the original definition from Bailis et
-    al.~\cite{bailis2014coordination}, but we do not comment on the difference.
-    We added a subsection (Section 2.3) clarifying that the difference in
-    definitions is superficial. We describe Bailis' original definition using
-    our formalism and prove that the two definitions are equivalent.
-
-  \item (Section 3, Theorem 2)
-    We added a formal proof that invariant closure is sufficient for invariant
-    confluence.
-
-  \item (Section 3, Theorem 3)
-    We added a formal proof of our key result equating invariant closure and
-    invariant confluence.
-
-  \item (Section 4.1)
-    We expanded the correctness proof of our interactive decision procedure. A
-    key step in understanding the correctness of our interactive invariant
-    confluence decision procedure---namely that $I$-reachable and
-    $(I-NR)$-reachable states are equal---was omitted from our conference
-    paper. We added a formal proof of this fact as well as a few other minor
-    elaborations.
-
-  \item (Section 4.2)
-    In our conference paper, we asserted that checking for invariant closure is
-    undecidable in general. We added a brief example showing \emph{why}
-    invariant closure is undecidable, by reducing it to the Halting Problem.
-
-  \item (Section 4.3)
-    One of the core contributions of our paper is an interactive invariant
-    confluence decision procedure. Here, ``interactive'' means that the
-    decision procedure relies on human input to guide the decision procedure.
-    It's possible for a person to provide the decision procedure with incorrect
-    information. We added a subsection (Section 4.3) describing how to mitigate
-    these errors.
-
-  \item (Section 5, Examples 3 and 4)
-    We added examples of merge-reducible and non-merge-reducible objects to
-    make it easier for the reader to understand the definition of
-    merge-reducibility.
-
-  \item (Section 5, Theorem 4 and Figure 4)
-    We added a formal proof that merge-reducibility is a sufficient condition
-    for invariant-confluence. We also added a diagram illustrating the key idea
-    behind the proof. Similarly, we added a counterexample showing that
-    merge-reducibility is not a necessary condition for invariant-confluence.
-
-  \item (Section 5, Theorem 5 and Figure 5)
-    We added a formal proof that criteria 1 through 4 are sufficient for
-    invariant confluence. The idea behind the proof is straightforward but the
-    details are tedious. To help readers see the forest through the trees, we
-    also added an illustration of the key idea of the proof.
-
-  \item (Section 5, Example 5)
-    In our conference paper, we provide an example of a distributed object that
-    satisfies criteria 1 through 4 outlined in Theorem 4 but is not invariant
-    closed. In the example, we asserted that criteria 1 through 4 are satisfied
-    but did not elaborate on why. We added a brief explanation making it clear
-    why the criteria are met.
-
-  \item (Section 6.3, Example 7)
-    We added an example of co-reachable and co-unreachable states to help the
-    reader understand the definition of co-reachability.
-
-  \item (Section 6.4)
-    We added a paragraph explaining the relationship between segmented
-    invariant confluence and a distributed locking approach to maintaining
-    invariants (that can be found in related works).
-
-  \item (Section 6.4)
-    We added a paragraph explaining a couple of optimizations that can be made
-    to our segmented invariant confluence system model to reduce the frequency
-    of global coordination.
-
-  \item (Section 6.4)
-    We added a paragraph explaining how a segmented invariant confluence
-    decision procedure can leverage our merge reducibility results in addition
-    to our results on invariant closure.
-
-  \item (Section 6.4)
-    We clarified the fault tolerance limitations of segmented invariant
-    confluence. A naive implementation of our segmented invariant confluence
-    decision procedure does not guarantee liveness, but there are
-    straightforward techniques we can leverage to introduce liveness. We added
-    an explanation of this to the paper.
-
-  \item (Section 7)
-    Our conference paper presents formal definitions of distributed objects and
-    invariant confluence. The definitions revolve around servers gossiping
-    state to one another and merging the states together using a binary merge
-    operator. This approach is called a ``state-based'' approach. We added a
-    section (Section 7) outlining a dual ``operation-based'' approach in which
-    transactions, rather than states, are gossiped. This duality between
-    state-based and operation-based definitions is well established in the CRDT
-    literature.
-\end{itemize}
+See our Overview section above.
 
@@
 transactions (Example 12), and more complex optimizations (Example 13) require
-segmentation. We agree though that these examples don't demonstrate the
-trade-off in an airtight way. The trade-off we describe in the introduction is
-written to give some intuition to the reader, but ultimately the notion of
-complexity is subjective, and the number of segments of an application is up to
-the user.
+segmentation. The trade-off we describe in the introduction is written to give
+some intuition to the reader, but ultimately the notion of complexity is
+subjective, and the number of segments of an application is up to the user.
 
@@
 \end{feedback}
-Users must manually identify transactions to remove. More generally,
-identifying a transaction to remove from a segmentation is a special case of
-forming a segmentation, and we do not provide any automated assistance on how
-to form a segmentation. Is the removal practically done? That depends on the
-workload. For some workloads, identifying invariant-violating transactions can
-be easy. The bank withdrawal example (Example 8). Other, more complex
+In the current work, users must manually identify transactions to remove. More
+generally, identifying a transaction to remove from a segmentation is a special
+case of forming a segmentation, and we do not provide any automated assistance
+on how to form a segmentation. Is the removal practically done? That depends on
+the workload. For some workloads, identifying invariant-violating transactions
+can be easy, like the bank withdrawal example (Example 8). Other, more complex
 workloads, can be difficult to reason about. Forming segmentations for these
@@
 \end{feedback}
-In-flight transactions are deferred until after the round of coordination is
-done. They aren't cancelled, and they are allowed to finish, but only after the
-coordination is done. The entire system stops processing transactions
-temporarily.  Technically, the system can still \emph{admit} new transactions
-during coordination, but it cannot execute them.
-
-We describe this in Section 6.2, for example, when we say that if ``$t \notin
-T_i$, then $p_i$ initiates a round of global coordination to execute $t$.
-During global coordination, every server temporarily stops processing
-transactions.''
+Thanks for catching this. We added some clarification to Section 8.1.  In
+short, in-flight transactions are deferred until after the round of
+coordination is done. They aren't cancelled, and they are allowed to finish,
+but only after the coordination is done. The entire system stops processing
+transactions temporarily. Technically, the system can still \emph{admit} new
+transactions during coordination, but it cannot execute them.
 
@@
 \begin{feedback}
-  end of first column of page 3 ``that it execute a transaction'' --> ``that it
-  executes a transaction''
+  end of first column of page 3 ``that it execute a transaction'' $\to$ ``that
+  it executes a transaction''
 
-  section 5, in example 3, ``Our start state state'' --> two times state
+  section 5, in example 3, ``Our start state state'' $\to$ two times state
 
-  section 8.2 in example 12: ``and a optional'' --> ``and an optional''
+  section 8.2 in example 12: ``and a optional'' $\to$ ``and an optional''
 \end{feedback}
@@
 \end{feedback}
-Without a proof of the two definitions being equivalent, it's possible that our
-notion of \invariantconfluence{} and Bailis et al.'s original notion differ.
-If so, none of our theorems or decision procedures could be applied to Bailis et
-al.'s findings. We'd be talking about two completely different concepts.
-Proving they are equivalent unifies the work. It's similar to proving that
-relational algebra, SQL, and relational calculus are all equally expressive.
-Without this knowledge, proofs using the different definitions would be
-mutually incompatible.
+We believe that proving equivalent definitions are, in fact, equivalent is a
+fundamental tenant of theoretical research. Without a proof of the two
+definitions being equivalent, it's possible that our notion of
+\invariantconfluence{} and Bailis et al.'s original notion differ.  If so, none
+of our theorems or decision procedures could be applied to Bailis et al.'s
+findings. We'd be talking about two completely different concepts.  Proving
+they are equivalent unifies the work. Without this knowledge, proofs using the
+different definitions would be mutually incompatible.
 
@@
 \end{feedback}
-Yup, a minor but helpful improvement.
+Yup, a minor but (we hope) useful improvement.
 
@@
 is significantly harder, as we have to create a proof showing that there does
-not exist any execution in which the state is reached. Things like the Collatz
-conjecture show that reasoning about the reachability or unreachability of even
-simple programs can be tremendously difficult. If we developed any substantial
-techniques to identify unreachable states in a general way, we feel that would
-constitute a strong publication on its own.
+not exist any execution in which the state is reached. Open mathematical
+problems like the Collatz conjecture~\cite{mottaanalysis} show that reasoning
+about the reachability or unreachability of even simple programs can be
+tremendously difficult. If we developed any substantial techniques to identify
+unreachable states in a general way, we feel that would constitute a strong
+publication on its own.
 
@@
 \url{https://mwhittaker.github.io/publications/matchmaker_paxos.html}---so we
-can answer your question with a bit of confidence. Assume a client sends a
-transactions to a server $p$. The transaction happens to not need global
+can answer your question with quite a bit of confidence. Assume a client sends
+a transaction to a server $p$. The transaction happens to not need global
 coordination, so $p$ executes the transaction locally and sends back the result

### 2020-08-10 doc/the_vldb_journal_paper/sections/evaluation.tex
Hopefully final revisions...
...before catching typos of course :)

@@
 a coordination request from the leader, it stops processing transactions and
-sends the leader its state. When the leader receives the states of all other
-servers, it executes the transaction, and then sends its state to the other
-servers. When a server receives a new state, it adopts the state, computes its
-new active segment, and resumes normal processing. After every 100 transactions
-processed, a server sends a merge request to a randomly selected server.
+sends the leader its state. \markrevisions{All in-flight transactions are
+deferred until the global communication is complete.} When the leader receives
+the states of all other servers, it executes the transaction, and then sends
+its state to the other servers. When a server receives a new state, it adopts
+the state, computes its new active segment, and resumes normal processing.
+After every 100 transactions processed, a server sends a merge request to a
+randomly selected server.
 

### 2020-08-10 doc/the_vldb_journal_paper/rebuttal.tex
Final tweaks.

@@
 \invariantconfluence{}. We exploit the relationship to develop an
-\invariantclosure{} decision procedure. We invent the notion of
+\invariantconfluence{} decision procedure. We invent the notion of
 \mergereducibility{} and characterize its relationship with
-\invariantclosure{}. We invent the notion of segmented \invariantconfluence{},
-and develop another decision procedure.
+\invariantclosure{} and \invariantconfluence{}. We invent the notion of
+segmented \invariantconfluence{} and develop a segmented \invariantconfluence{}
+decision procedure.
 
 When our paper was invited to this journal, we took the opportunity to improve
-the rigor and clarity of the core contributions of the paper. We added numerous
-proofs to verify the theorems in the paper, and we added many examples and
-illustrations providing valuable intuition behind the key results. We also
-added a significant amount of new material:
+the rigor and clarity of the core contributions of the paper: its theoretical
+foundations. We added numerous proofs to verify the theorems in the paper, and
+we added many examples and illustrations providing valuable intuition behind
+the key results. We also added a significant amount of new material:
 \begin{itemize}
@@
 \end{feedback}
-See W1, W2, W3, and D6.
+See W1, W2, W3, and our Overview section above.
 
@@
 \end{feedback}
-We address this feedback in detail below. We also disagree that the time
-between publication and now should have any bearing on the revisions.
+We address this feedback in detail below.
 
@@
 We believe that proving equivalent definitions are, in fact, equivalent is a
-fundamental tenant of theoretical research. Without a proof of the two
+fundamental tenet of theoretical research. Without a proof of the two
 definitions being equivalent, it's possible that our notion of
@@
 We're happy that the original proof was clear to the reviewer, but we felt it
-was not clear to most people, including us.
+was not clear to most people.
 
@@
 user and (2) the possibility of introducing errors. We break down the two types
-of user errors (i.e.\ labelling an unreachable state as reachable and labelling
+of user errors (i.e.\ labeling an unreachable state as reachable and labeling
 a reachable state as unreachable) and provide solutions for both, eliminating
@@
 
-Note, however, that a segmented \invariantconfluent{} object does suffer from
-liveness problems, as we discuss in Section 6.4. Specifically, transactions
-that require global coordination are live only as long as every replica is
-live. We discuss this more in Reviewer 4: 5 below.
+Note, however, that a \emph{segmented} \invariantconfluent{} object does suffer
+from liveness problems, as we discuss in Section 6.4. Specifically,
+transactions that require global coordination are live only as long as every
+replica is live. We discuss this more in Reviewer 4: 5 below.
 
@@
 can answer your question with quite a bit of confidence. Assume a client sends
-a transaction to a server $p$. The transaction happens to not need global
-coordination, so $p$ executes the transaction locally and sends back the result
-to the client. This scenario makes membership changes difficult for the
+a transaction to a server $p$. The transaction happens to \emph{not} need
+global coordination, so $p$ executes the transaction locally and sends back the
+result to the client. This scenario makes membership changes difficult for the
 following reason. Consider a different execution in which $p$ is partitioned
@@
 
+\paragraph{Reviewer 4: 6}
 \begin{feedback}

### 2020-08-10 doc/the_vldb_journal_paper/sections/evaluation.tex
Final tweaks.

@@
   For example, assume that $k = 3$ and that the three states are $s_1 = s_2 =
-  s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 4, 5), (3, 3, 3)$. Here, the
-  value of the PN-counter is $(3 + 4 + 5) - (3 + 3 + 3) = 3$, which is
+  s_3 = (p_1, p_2, p_3), (n_1, n_2, n_3) = (3, 3, 3), (3, 2, 1)$. Here, the
+  value of the PN-counter is $(3 + 3 + 3) - (3 + 2 + 1) = 3$, which is
   non-negative as expected. If server 1 receives a decrement request, it
@@
   so would cause $n_1$ to exceed the escrow threshold $k = 3$. Naively, it
-  seems safe. Server 1 would transition to state $(3, 3, 5), (3, 3, 3)$ which
+  seems safe. Server 1 would transition to state $(3, 3, 3), (4, 2, 1)$ which
   has non-negative value $2$. However, if servers 2 and 3 concurrently perform
@@
   On the other hand, if server 2 receives a decrement request, it can safely
-  execute it locally and transition to state $(3, 3, 5), (3, 3, 3)$. This is
+  execute it locally and transition to state $(3, 3, 3), (3, 3, 1)$. This is
   safe because server 2 knows that every server $i$ maintains $p_i \geq n_i$.
@@
   Escrow transactions are a good example where this optimization works
-  particularly well. Assume that $k = 3$ and that the three states are $(3, 3,
-  3), (3, 3, 3)$. Here, no server can process a decrement transaction locally.
-  If a server receives a decrement transaction, it must execute it with global
-  coordination. However, if the server instead defers the execution of the
-  decrement transaction and happens to receive a number of increment
-  transactions, then it can execute the decrement locally, without the need for
-  global coordination.
+  particularly well. Assume that instead of using PN-counters, we implement a
+  replicated bank account balance as a tuple of timestamped integers $(x_1,
+  x_2, \ldots, x_n)$, with the value of the bank account being the sum of the
+  integers, with merging done element-wise, and with higher timestamps
+  overriding lower timestamps~\cite{wu2019anna}. Every time a sever $s_i$
+  increments or decrements its integer $x_i$, it increases the integer's
+  timestamp. We consider the segmentation in which decrements do not require
+  global coordination so long as every individual integer remains non-negative.
+
+  If a server's integer is zero, it cannot process a decrement transaction
+  locally, it must execute it with global coordination. However, if the server
+  instead defers the execution of the decrement transaction and happens to
+  receive a number of increment transactions, then it can execute the decrement
+  locally, without the need for global coordination.
 \end{revisions}
